### 2015-11-09 imaging-paper.Rnw
setting things up

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

% declare the path(s) where your graphic files are
\graphicspath{{Figure/}{Images/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

%---------------------------------------------------
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
\usepackage{color}
% \usepackage[dvipsnames,svgnames]{xcolor}
% \usepackage{ulem}
% \usepackage[section]{placeins}
\usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
% \usepackage{sidecap}
\usepackage{multirow}
\usepackage{bbm}
\usepackage{url}
\usepackage{xr}
\usepackage{xr-hyper}
\usepackage{afterpage}

%---------------------------------------------------

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=section,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

\externaldocument{appendix}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Theory/Model}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}

%% Paper title.

\title{Matching bullets}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Eric Hare and Heike Hofmann}
\authorfooter{
%% insert punctuation at end of each item
\item Eric Hare is a graduate student in the Department of Statistics at Iowa State University. Email: erichare@iastate.edu.
\item Heike Hofmann is Professor of Statistics at Iowa State University. Email: hofmann@iastate.edu.
}

%other entries to be set up for journal
\shortauthortitle{Hare \MakeLowercase{\textit{et al.}}: Matching Bullets}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{
abstract goes here - this is actually not part of the paper, but goes on a separate page.
} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Data visualization, Statistical graphics, Statistical computing.}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

\CCScatlist{ % not used in journal version
 \CCScat{H.2.8.c}{Data and knowledge visualization}%
{Database applications}{Database management};
 \CCScat{G.3.n}{Statistical Computing}{Probability and Statistics}{Mathematics of Computing}
}

%% Uncomment below to include a teaser figure.
   \teaser{
      \centering
Teaser image wouldn't be bad, let's see whether there is  space
%          \includegraphics[height=\textwidth]{lineup.png}
    \caption{\protect\label{fig:teaser} Teaser image shows the final results. }
  }

<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
#rm(list=ls())
#wd <- getwd()
library(extrafont)
library(knitr)
opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|", dev="CairoPDF")
datadir <- "data"
imgdir <- "Figure/"
codedir <- "code"

options(replace.assign=TRUE,scipen=3)
@

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

Rough todo list:
\begin{enumerate}
\item Description of the problem: Standard operating practice, 2009 NAS report
\item Description of the Results
\item Walkthrough of the process
\end{enumerate}

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{references}

\end{document}

### 2015-11-10 imaging-paper.Rnw
Beginning some work on the introduction

@@
 
+Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel.
+This process has broad applicability in terms of the conviction of criminals in the United States criminal justice system.
+Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone some more significant scrutiny.
+In 2005, in \emph{United States vs.
+Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar.
+
 \bibliographystyle{abbrv}

### 2015-11-10 imaging-paper.Rnw
More intro

@@
 Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar.
+Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence.
+
+The new scrutiny of the procedure culminated with the 2009 National Academy of Sciences report.
+The report concluded that minimal research had been done exploring the reliability of these ballistic identification methods.
+In fact, the report says the forensic evidence "including, for example, bite marks and rearm and toolmark identifications is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
 

### 2015-11-10 imaging-paper.Rnw
added reference to NAS report

--- 2015-11-10 imaging-paper.Rnw
@@
 
+% Quick Intro
 Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel.
@@
 Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar.
+Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence.
+
+The new scrutiny of the procedure culminated with the 2009 National Academy of Sciences report \cite{NAS:2009}.
+The report concluded that minimal research had been done exploring the reliability of these ballistic identification methods.
+In fact, the report says the forensic evidence ``including, for example, bite marks and rearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
+
+% Statement of the problem
 

### 2015-11-15 imaging-paper.Rnw
Sketch the later sections
Start some discussion of the land images

@@
 
+% Methodology
+Throughout this section, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study.
+Ten Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
+3D topographical images are provided in format called x3p.
+The x3p format was developed as a means for forensic examiners to easily exchange images in a standard format.
+It was adopted by the OpenFMC (Open Forensic Metrology Consortium), a group of academic, industry, and government Firearm Forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+
+Each fired bullet is provided as a set of six x3p files, where each file is a surface scan between two bullet grooves, called a ``land".
+An example of plotting one of these lands is given in Figures \ref{fig:sidex3p} and \ref{fig:topx3p}, which show side and top profiles of the land respectively.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\linewidth]{images/sidex3p.png}
+  \caption{An example of a scan of a bullet land, viewed from the side.}
+  \label{fig:sidex3p}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\linewidth]{images/topx3p.png}
+  \caption{An example of a scan of a bullet land, viewed from the top.}
+  \label{fig:topx3p}
+\end{figure}
+
+% Application
+
+% Conclusion
+
 \bibliographystyle{abbrv}

### 2015-11-15 imaging-paper.Rnw
Residual talk

@@
 \usepackage{times}
+\usepackage{float}
 
@@
 
+The primary goal in this type of forensic analysis is to identify markings.
+Gun barrels should produce similar markings on bullets fired consecutively, and the careful analysis of these surface images can help us determine whether two bullets are a match.
+However, this process has traditionally required forensic experts to perform the analysis, also required physical examination of the bullets.
+We've produced a framework which, in software, allows for the analysis of the surface topologies, and also transcribes the markings into a 2D plotting framework which we believe makes identification of matches easier for the lay person.
+
+An initial naive approach leads us to fixing a particular value of a coordinate, and then producing a plot of the heights across values of the other coordinate.
+Figure \ref{fig:fixedX} gives a plot resulting from such a procedure.
+It can be seem that the global structure of the land dominates the appearance of the plot.
+The grooves can be clearly identified on the left and right side, and the curvature of the surface is the most visible feature in the middle.
+
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, out.width='\\textwidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across fixed x values.
+Note that the global features dominate any deviations, which would correspond to markings.'>>=
+library(RColorBrewer)
+library(ggplot2)
+library(dplyr)
+library(x3pr)
+library(x3prplus)
+
+datadir <- "~/GitHub/x3prproto/images/Hamby252_3DX3P1of2/"
+
+br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
+dbr111 <- fortify_x3p(br111)
+
+pars <- data.frame(getCircle(dbr111$y, dbr111$value))
+dbr111$theta <- acos((dbr111$y-pars$x0)/pars$radius)/pi*180
+dbr111 <- dbr111 %>% mutate(
+  xpred = cos(theta/180*pi)*pars$radius + pars$x0,
+  ypred = sin(theta/180*pi)*pars$radius + pars$y0
+)
+
+qplot(data=subset(dbr111, x <= 100 & x >= 95), y, value, colour=factor(x)) +
+  geom_line(aes(x=xpred, y=ypred, group=x), 
+            colour="black", size=0.25) +
+  scale_colour_brewer("x", palette="Paired") + 
+  theme_bw() + 
+  theme(legend.position="bottom") + coord_equal()
+@
+
+It's clear that our next approach must be to model the signal in order to focus on the noise, given that the noise should correspond to bullet markings.
+
 % Application

### 2015-11-15 imaging-paper.Rnw
bit more on x3p file format

@@
 % Methodology
-Throughout this section, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study.
+Throughout this section, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study \cite{hamby:2009}.
 Ten Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
-3D topographical images are provided in format called x3p.
-The x3p format was developed as a means for forensic examiners to easily exchange images in a standard format.
-It was adopted by the OpenFMC (Open Forensic Metrology Consortium), a group of academic, industry, and government Firearm Forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+3D topographical images are provided on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
+x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
+%The x3p format was developed as a means for forensic examiners to easily exchange images in a standard format. 
+x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium), a group of academic, industry, and government Firearm Forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+
+In the Hamby Study, x3p files were created using a NanoFocus lens at 20x resolution. \hh{not sure where this sentences best goes.}
 
 Each fired bullet is provided as a set of six x3p files, where each file is a surface scan between two bullet grooves, called a ``land".
-An example of plotting one of these lands is given in Figures \ref{fig:sidex3p} and \ref{fig:topx3p}, which show side and top profiles of the land respectively.
+An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.
 

### 2015-11-16 imaging-paper.Rnw
personal highlighter
typo
couple of comments

@@
 \usepackage{color}
+
+\newcommand{\hh}[1]{{\color{magenta} #1}}
 % \usepackage[dvipsnames,svgnames]{xcolor}
@@
 
-In the Hamby Study, x3p files were created using a NanoFocus lens at 20x resolution. \hh{not sure where this sentences best goes.}
+In the Hamby Study, x3p files were created using a NanoFocus lens at 20x magnification. \hh{not sure where this sentences best goes.}
 
@@
 It's clear that our next approach must be to model the signal in order to focus on the noise, given that the noise should correspond to bullet markings.
+\hh{XXX let's not call the markings noise - that has the conotation of them being random.
+The markings are also a signal, it is just dominated by the stronger signal of the land structure.
+We often go through a process of multiple steps to get close to the more interesting, smaller signal.
+}
+\hh{XXX on a side note - I've got a different relative path to the x3p bullet files ...
+should we assume a data directory inside the folder?
+Then we don't have to push the files on a github repo again, but are using the same file structure?}
 

### 2015-11-16 imaging-paper.Rnw
Address comments, begin discussion of cylindrical fit

@@
 
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, out.width='\\textwidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across fixed x values.
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, out.width='\\linewidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across fixed x values.
 Note that the global features dominate any deviations, which would correspond to markings.'>>=
@@
 
-It's clear that our next approach must be to model the signal in order to focus on the noise, given that the noise should correspond to bullet markings.
-\hh{XXX let's not call the markings noise - that has the conotation of them being random.
-The markings are also a signal, it is just dominated by the stronger signal of the land structure.
-We often go through a process of multiple steps to get close to the more interesting, smaller signal.
-}
+It's clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
+
+\subsection{Cylindrical Fit}
+
+One reasonable approach is to fit a circle to the curve and compute the residuals from this fit.
+The scatterplot below shows the residuals of such a fit.
+The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
+For cross sections of $x$ values between 75 and 80, there is a residual circular structure that does not show up for all cross sections.
+
+<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+qplot(data=subset(dbr111, x <= 80 & x >=75), y, value-ypred, colour=factor(x)) +
+  scale_colour_brewer("x", palette="Paired") + 
+  theme_bw() + 
+  theme(legend.position="bottom")
+@
+
+<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+qplot(data=subset(dbr111, x <= 5), y, value-ypred, colour=factor(x)) +
+  scale_colour_brewer("x", palette="Paired") + 
+  theme_bw() + 
+  theme(legend.position="bottom")
+@
+
 \hh{XXX on a side note - I've got a different relative path to the x3p bullet files ...

### 2015-11-16 imaging-paper.Rnw
Residual plots

@@
 
+A single cylinder as a fit is unlikely to be a particularly good fit; because there seem to be quite massive deformations in vertical direction.
+We therefore fit a circle for each cross section of the bullet, and compare the figure again:
+
+<<bullet1, echo=FALSE, cache=TRUE>>=
+db1 <- NULL
+for (i in 1:6) {
+  bname <- sprintf(file.path(datadir, "Br1 Bullet 1-%d.x3p"), i)
+  dbi <- fortify_x3p(read.x3p(bname))
+  dbi$part <- i
+  db1 <- rbind(db1, dbi)
+}
+
+db1 <- db1 %>% group_by(part, x) %>% do (
+    data.frame(., predCircle(.$y, .$value))
+  )
+@
+
+<<bullet2, echo=FALSE, cache=TRUE>>=
+db2 <- NULL
+for (i in 1:6) {
+  bname <- sprintf(file.path(datadir, "Br1 Bullet 2-%d.x3p"), i)
+  dbi <- fortify_x3p(read.x3p(bname))
+  dbi$part <- i
+  db2 <- rbind(db2, dbi)
+}
+
+db2 <- db2 %>% group_by(part, x) %>% do (
+    data.frame(., predCircle(.$y, .$value))
+  )
+
+@
+
+First attempt to match - something is wrong with  piece 3 of bullet 2.
+
+<<echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x)) + 
+  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
+  theme_bw() + theme(legend.position="bottom") + 
+  geom_point(aes(y, resid), colour="black", size=1, alpha=0.2, 
+             data = filter(db1, part==1, x <= 80, x >=75)[,c("y", "resid")]) + ggtitle("Part 1 of Bullet 1 in black")
+@
+
 \hh{XXX on a side note - I've got a different relative path to the x3p bullet files ...

### 2015-11-17 imaging-paper.Rnw
couple of comments

@@
 
-Rough todo list:
+Rough outline:
 \begin{enumerate}
@@
 \end{enumerate}
+
+\hh{I'm putting in footnotes for URLs at the moment, they should probably be included properly in the references.}
 
@@
 Throughout this section, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study \cite{hamby:2009}.
-Ten Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
-3D topographical images are provided on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
+Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
+
+\hh{XXX here's a paragraph missing: rifling in the barrel is creating striation marks on the bullet during the firing process.
+These marks are claimed to be unique to the barrel - we need to find a citation for that from the SOP manual.
+}
+
+3D topographical images were taken of these bullets using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
 x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
 %The x3p format was developed as a means for forensic examiners to easily exchange images in a standard format. 
-x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium), a group of academic, industry, and government Firearm Forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government Firearm Forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
 
-In the Hamby Study, x3p files were created using a NanoFocus lens at 20x magnification. \hh{not sure where this sentences best goes.}
-
-Each fired bullet is provided as a set of six x3p files, where each file is a surface scan between two bullet grooves, called a ``land".
+Each fired bullet is provided as a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.
@@
 
-datadir <- "~/GitHub/x3prproto/images/Hamby252_3DX3P1of2/"
+#datadir <- "~/GitHub/x3prproto/images/Hamby252_3DX3P1of2/"
 
@@
 
-\hh{XXX on a side note - I've got a different relative path to the x3p bullet files ...
-should we assume a data directory inside the folder?
-Then we don't have to push the files on a github repo again, but are using the same file structure?}
-
 % Application

### 2015-11-17 code/process_bullets.R
color scheme

@@ -133,20 +133,34 @@ qplot(x, y, fill = resid, data=LOF, geom="tile", facets=~bullet) +
   scale_fill_gradient(limits=c(-5,5), low = "grey5", high="grey95") + theme_bw()
 
 qplot(x, y, fill = resid, data=LOF, geom="tile", facets=~bullet) + 
-  scale_fill_gradient(limits=c(-5,5), low = "grey5", high="darkgoldenrod1") + theme_bw()
+  scale_fill_gradient2(limits=c(-5,5), low = "grey5", mid = "gold4", #mid="darkgoldenrod1", 
+                       high = "lightgoldenrod1") + theme_bw()
+
+
+qplot(x, y, fill = resid, data=LOF, geom="tile", facets=~bullet) + 
+  scale_fill_gradientn(limits=c(-5,5), 
+                       colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
+                       values = c(0,0.5, 0.75, 0.9, 1)) + theme_bw()
+
 
 subLOF1 <- LOF %>% filter(bullet == 1, x <= 75)
 subLOF2 <- LOF %>% filter(bullet == 2, x > 75, x < 100)
 
-subLOF$y <- subLOF$y %/% .64
+#subLOF$y <- subLOF$y %/% .64
 #subLOF1$y <- subLOF1$y + 11 # why doesn't this work??
 subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
 
-qplot(x, y, fill = resid, colour=I(NA), data=filter(subLOF, y >450, y < 550), geom="tile") + 
-  scale_fill_gradient(limits=c(-5,5), low = "grey5", high="darkgoldenrod1") +
-#  scale_colour_gradient(limits=c(-5,5), low = "grey5", high="darkgoldenrod1") +
+qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
+  scale_fill_gradientn(limits=c(-5,5), 
+                       colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
+                       values = c(0,0.5, 0.75, 0.9, 1)) +
   theme_bw()
 
+qplot(x, y, fill = 5*sign(resid)*sqrt(abs(resid/5)), colour=I(NA), data=subLOF, geom="tile") + 
+  scale_fill_gradientn(limits=c(-5,5), 
[…]

### 2015-11-17 imaging-paper.Rnw
references to Thompson's work

@@
 
-\hh{XXX here's a paragraph missing: rifling in the barrel is creating striation marks on the bullet during the firing process.
+\hh{XXX here's a paragraph missing: rifling in the barrel is creating striation marks on the bullet during the firing process (and grooves).
 These marks are claimed to be unique to the barrel - we need to find a citation for that from the SOP manual.
-}
+Roger Thompson's writeup on firearm identification cites AFTE 1992 procedures.\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}, \footnote{AFTE 1992 theory of identification \url{http://projects.nfstc.org/firearms/module09/fir_m09_t05_05.htm}} }
 
-3D topographical images were taken of these bullets using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
+3D topographical images of each bullet were taken using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
 x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
@@
 
-Each fired bullet is provided as a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
+Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.

### 2015-11-17 imaging-paper.Rnw
AFTE statements

@@
 These marks are claimed to be unique to the barrel - we need to find a citation for that from the SOP manual.
-Roger Thompson's writeup on firearm identification cites AFTE 1992 procedures.\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}, \footnote{AFTE 1992 theory of identification \url{http://projects.nfstc.org/firearms/module09/fir_m09_t05_05.htm}} }
+Roger Thompson's writeup on firearm identification cites AFTE 1992 procedures.\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}, \footnote{AFTE 1992 theory of identification \url{http://projects.nfstc.org/firearms/module09/fir_m09_t05_05.htm}} includes the following statements, both on page 34: `The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in ``sufficient agreement".' (my highlighting) and `Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows.'}
 

### 2015-11-17 imaging-paper.Rnw
Matching part 2 of bullet 1

@@
 
-#datadir <- "~/GitHub/x3prproto/images/Hamby252_3DX3P1of2/"
+datadir <- "~/GitHub/x3prproto/images/Hamby252_3DX3P1of2/"
 
@@
 
-A single cylinder as a fit is unlikely to be a particularly good fit; because there seem to be quite massive deformations in vertical direction.
-We therefore fit a circle for each cross section of the bullet, and compare the figure again:
+A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in vertical direction.
+We therefore fit a circle for each cross section of the bullet.
+Figure \ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line.
+There is some global structure evident that makes focusing on the deviations a bit challenging.
 
@@
 
-First attempt to match - something is wrong with  piece 3 of bullet 2.
-
-<<echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+<<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.', fig.pos='H'>>=
 qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x)) + 
@@
 @
+
+If we perform a similar routine for part 2 of bullet 1, we obtain the figure give in Figure \ref{fig:circlefits2}.
+
+<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.', fig.pos='H'>>=
+qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x)) + 
+  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
+  theme_bw() + theme(legend.position="bottom") + 
+  geom_point(aes(y, resid), colour="black", size=1, alpha=0.2, 
+             data = filter(db1, part==2, x <= 80, x >=75)[,c("y", "resid")]) + ggtitle("Part 2 of Bullet 1 in black")
+@
+
+It can be seen pretty clearly at this point that the deviations resulting from the grooves dominates the residual structure that would correspond to markings.
+This is unfortunate, as the grooves themselves will not typically contain markings produced by a gun barrel.
+We can also see the residual circular structure which suggests our circle fit is not entirely appropriate.
+
+\subsection{Loess Fit}
 

### 2015-11-17 imaging-paper.Rnw
circle fit

@@
 
+\section{Circular Fit}
+
+Assume, $n$ data points $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ are given that are (approximately) located on a circle.
+We want to estimate the location of the center and radius of the best fitting circle.
+
+We minimize the following expression:
+\begin{equation}\label{eq:circle}
+D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2
+\end{equation}
+We do this by differentiating $D$ with respect to $r, a,$ and $b$:
+let us assume that $x_i$ and $y_i$ are centralized (i.e.
+$\sum x_i = \sum_i y_i = 0$).
+Note, if they are not, make a note of the current means, subtract them now and add them to $(\hat{a}, \hat{b})$ at the end.
+
+\noindent
+The  derivate of $D$ with respect to $r$ is:
+\begin{eqnarray*}
+\frac{d}{dr} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 r = \\
+&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right) 
+\end{eqnarray*}
+For a minimum then holds:
+\begin{equation}\label{eq:rmin}
+\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2 
+\end{equation}
+%
+%
+The  derivative of $D$ with respect to $a$ is:
+\begin{eqnarray*}
+\frac{d}{da} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 (x_i - a) = \\
+&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]
+\end{eqnarray*}
+Using (\ref{eq:rmin}) for $nr^2$  in the equation above we get:
+\begin{eqnarray*}
+\frac{d}{da} D &=& -4 \left[  \sum_i a(x_i-a)^2 +  \sum_i a(y_i-b)^2  + \right. \\
+&& \phantom{-4 \ \ } \left . \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]  = \\
+&=& -4 \left[ \sum_i (x_i-a)^2 (a + x_i - a)  + \sum_i (x_i - a + a) (y_i - b)^2 \right] = \\
+&=& -4 \left[ \sum_i (x_i-a)^2 x_i   + \sum_i x_i  (y_i - b)^2 \right] \stackrel{\sum_i x_i = \sum_i y_i = 0}{=} \\
+&=& -4 \left[ \sum_i x_i^3   + \sum_i x_i y_i^2  - 2a s_{xx} - 2b s_{xy} \right], 
+\end{eqnarray*}
+where $s_{xx} = \sum_i x_i^2, s_{xy} = \sum_i x_i y_i$ and $s_{yy} = \sum_i y_i^2$.
+
+\noindent
+Likewise, we get for the derivative of $D$ with respect to $b$:
+\begin{eqnarray*}
+\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right] 
+\end{eqnarray*}
+For the minimum we therefore get a system of two linear equations in $a$ and $b$:
+\begin{eqnarray*}
+2 s_{xx} a + 2 s_{xy} b = c_1 && \text{ with } c_1 = \sum_i x_i^3 + x_i y_i^2 \\
+2 s_{xy} a + 2 s_{yy} b = c_2 &&\text{ with } c_2 = \sum_i x_i^2 y_i + y_i^3.
+\end{eqnarray*}
+This resolves to 
+\begin{eqnarray*}
+\hat{a} &=& \frac{c_1 s_{yy} - c_2 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2},\\
+\hat{b} &=& \frac{c_2 s_{xx} - c_1 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2}, \text{ and}\\
+\hat{r^2} &=& \frac{1}{n}s_{xx} + \frac{1}{n}s_{yy} + \hat{a}^2 + \hat{b}^2.
+\end{eqnarray*}
 \end{document}

### 2015-11-17 imaging-paper.Rnw
i thought i had pushed all ...
Sections + removing the grooves

@@
 
-% Quick Intro
+\section{Introduction}
 Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel.
@@
 
-% Statement of the problem
+\section{Problem Statement}
 
-% Methodology
+\section{Methodology}
 Throughout this section, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study \cite{hamby:2009}.
@@
 
-% Application
+To improve upon the previous algorithm, the first step must be to eliminate the grooves from the images under the assumption that they do not contain relevant information for determining a match.
+Fortunately, the location and appearance of the grooves in these 2d plots is quite consistent.
+Because a bullet is scanned groove to groove, the groove should appear within approximately scan points from each end.
+That is, on y coordinate system ranging from 0 to 1000 as these images, are, the grooves will appear between $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
+Hence, the algorith for removing the grooves can be described as follows:
 
-% Conclusion
+1. Fix an x coordinate value.
+2. Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
+3. For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minimums.
+4. Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove)
+5. Remove all y values below the minimum and above the maximum
+
+\section{Application}
+
+\section{Conclusion}
 

### 2015-11-17 imaging-paper.Rnw
lines as geom

--- 2015-11-17 imaging-paper.Rnw
@@
 
-qplot(data=subset(dbr111, x <= 100 & x >= 95), y, value, colour=factor(x)) +
+qplot(data=subset(dbr111, x <= 100 & x >= 95), y, value, colour=factor(x), geom="line", size=I(1)) +
   geom_line(aes(x=xpred, y=ypred, group=x), 
-            colour="black", size=0.25) +
-  scale_colour_brewer("x", palette="Paired") + 
+            colour="grey30", size=0.25) +
+  scale_colour_brewer("Cross-sections at x", palette="Paired") + 
+  ylab(expression(paste("Surface Measurements (in ",mu,m,")"))) + 
   theme_bw() + 
@@
 <<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
-qplot(data=subset(dbr111, x <= 80 & x >=75), y, value-ypred, colour=factor(x)) +
+qplot(data=subset(dbr111, x <= 80 & x >=75), y, value-ypred, colour=factor(x), geom="line", size=I(1)) +
   scale_colour_brewer("x", palette="Paired") + 
@@
 <<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
-qplot(data=subset(dbr111, x <= 5), y, value-ypred, colour=factor(x)) +
+qplot(data=subset(dbr111, x <= 5), y, value-ypred, colour=factor(x),
+      geom="line", size=I(1)) +
   scale_colour_brewer("x", palette="Paired") + 
@@
 <<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.', fig.pos='H'>>=
-qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x)) + 
+qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x), geom="line", size=I(1)) + 
   facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_point(aes(y, resid), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==1, x <= 80, x >=75)[,c("y", "resid")]) + ggtitle("Part 1 of Bullet 1 in black")
+  geom_line(aes(y, resid, group = x), colour="black", size=1, alpha=0.2, 
+             data = filter(db1, part==1, x <= 80, x >=75)[,c("y", "resid", "x")]) + ggtitle("Part 1 of Bullet 1 in black")
 @
 
-If we perform a similar routine for part 2 of bullet 1, we obtain the figure give in Figure \ref{fig:circlefits2}.
+If we perform a similar routine for part 2 of bullet 1, we obtain the figure given in Figure \ref{fig:circlefits2}.
 
 <<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.', fig.pos='H'>>=
-qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x)) + 
+qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x), geom="line", size=I(1)) + 
   facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_point(aes(y, resid), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==2, x <= 80, x >=75)[,c("y", "resid")]) + ggtitle("Part 2 of Bullet 1 in black")
+  geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
+             data = filter(db1, part==2, x <= 80, x >=75)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
 @

### 2015-11-17 imaging-paper.Rnw
Work on the loess section

--- 2015-11-17 imaging-paper.Rnw
@@
 
-1. Fix an x coordinate value.
-2. Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
-3. For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minimums.
-4. Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove)
-5. Remove all y values below the minimum and above the maximum
+\begin{enumerate}
+    \item Fix an x coordinate value.
+    \item Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
+    \item For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minimums.
+    \item Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove)
+    \item Remove all y values below the minimum and above the maximum
+\end{enumerate}
+
+Next, we fit a Loess regression to the data
+
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit for part 1 of bullet 1', fig.pos='H', message=FALSE>>=
+br111 <- get_bullet(file.path("GitHub", "x3prproto", "images", "Hamby252_3DX3P1of2", "Br1 Bullet 1-1.x3p"))
+br111.groove <- get_grooves(br111)
+
+my.loess <- fit_loess(br111, br111.groove)
+
+my.loess$fitted
+@
+
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit residuals for part 1 of bullet 1', fig.pos='H'>>=
+my.loess <- fit_loess(br111, br111.groove)
+
+my.loess$resid
+@
 

### 2015-11-17 code/process_bullets.R

@@ -12,62 +12,6 @@ library(pastecs)
 Sys.setenv("plotly_username" = "erichare")
 Sys.setenv("plotly_api_key" = "xd0oxpeept")
 
-get_bullet <- function(path, x = 99.84) {
-    br111 <- read.x3p(path)
-    dbr111 <- fortify_x3p(br111)
-    
-    dbr111.fixx <- dbr111[dbr111$x == x,]
-    
-    return(dbr111.fixx)
-}
-
-get_grooves <- function(bullet, trim = 125) {
-    smoothed <- c(rep(NA, 10), rollapply(bullet$value, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
-    subsmooth <- tail(head(smoothed, n = -trim), n = -trim)
-    
-    min.left <- which.min(subsmooth[1:(length(subsmooth) %/% 2)])
-    min.right <- which.min(subsmooth[(length(subsmooth) %/% 2):length(subsmooth)])
-        
-    p <- qplot(data=bullet, y, value) +
-        theme_bw() + coord_equal() +
-        geom_vline(xintercept = bullet$y[min.left + trim], colour = "red") +
-        geom_vline(xintercept = bullet$y[min.right + trim + (length(subsmooth) %/% 2)], colour = "blue")
-    
-    return(list(groove = c(bullet$y[min.left + trim], 
-                           bullet$y[min.right + trim + (length(subsmooth) %/% 2)]), plot = p))
-}
-
-fit_loess <- function(bullet, groove) {
-    bullet_filter <- subset(bullet, !is.na(value) & y > groove$groove[1] & y < groove$groove[2])
-    my.loess <- loess(value ~ y, data = bullet_filter)
-    bullet_filter$fitted <- fitted(my.loess)
-    bullet_filter$resid <- resid(my.loess)
-    
-    # filter out most extreme residuals
-    bullet_filter$abs_resid <-  abs(bullet_filter$resid)
-    cutoff <- quantile(bullet_filter$abs_resid, probs = c(0.995))
-    bullet_filter$chop <- bullet_filter$abs_resid > cutoff
-
-    qplot(data = bullet_filter, y, resid, colour = chop) +
-      theme_bw() 
-    bullet_filter <- subset(bullet_filter, chop != TRUE)
-    
-    qplot(data = bullet_filter, y, value) +
[…]

### 2015-11-17 imaging-paper.Rnw

--- 2015-11-17 imaging-paper.Rnw
@@
 
-% Application
+To improve upon the previous algorithm, the first step must be to eliminate the grooves from the images under the assumption that they do not contain relevant information for determining a match.
+Fortunately, the location and appearance of the grooves in these 2d plots is quite consistent.
+Because a bullet is scanned groove to groove, the groove should appear within approximately scan points from each end.
+That is, on y coordinate system ranging from 0 to 1000 as these images, are, the grooves will appear between $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
+Hence, the algorith for removing the grooves can be described as follows:
 
-% Conclusion
+\begin{enumerate}
+    \item Fix an x coordinate value.
+    \item Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
+    \item For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minimums.
+    \item Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove)
+    \item Remove all y values below the minimum and above the maximum
+\end{enumerate}
+
+Next, we fit a Loess regression to the data
+
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit for part 1 of bullet 1', fig.pos='H', message=FALSE>>=
+br111 <- get_bullet(file.path("GitHub", "x3prproto", "images", "Hamby252_3DX3P1of2", "Br1 Bullet 1-1.x3p"))
+br111.groove <- get_grooves(br111)
+
+my.loess <- fit_loess(br111, br111.groove)
+
+my.loess$fitted
+@
+
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit residuals for part 1 of bullet 1', fig.pos='H'>>=
+my.loess <- fit_loess(br111, br111.groove)
+
+my.loess$resid
+@
+
+\section{Application}
+
+\section{Conclusion}
 

### 2015-11-17 imaging-paper.Rnw
constants changed

--- 2015-11-17 imaging-paper.Rnw
@@
 
-qplot(data=subset(dbr111, x <= 100 & x >= 95), y, value, colour=factor(x)) +
+qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 95*1.5625^2), y, value, colour=factor(x), geom="line", size=I(1)) +
   geom_line(aes(x=xpred, y=ypred, group=x), 
-            colour="black", size=0.25) +
-  scale_colour_brewer("x", palette="Paired") + 
+            colour="grey30", size=0.25) +
+  scale_colour_brewer("Cross-sections at x", palette="Paired") + 
+  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
   theme_bw() + 
@@
 <<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
-qplot(data=subset(dbr111, x <= 80 & x >=75), y, value-ypred, colour=factor(x)) +
+qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred, colour=factor(x), geom="line", size=I(1)) +
   scale_colour_brewer("x", palette="Paired") + 
@@
 <<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
-qplot(data=subset(dbr111, x <= 5), y, value-ypred, colour=factor(x)) +
+qplot(data=subset(dbr111, x <= 5*1.5625^2), y, value-ypred, colour=factor(x),
+      geom="line", size=I(1)) +
   scale_colour_brewer("x", palette="Paired") + 
@@
 <<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.', fig.pos='H'>>=
-qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x)) + 
+qplot(y, resid, data=subset(db2, x <= 80*1.5625^2 & x >=75*1.5625^2), colour=factor(x), geom="line", size=I(1)) + 
   facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_point(aes(y, resid), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==1, x <= 80, x >=75)[,c("y", "resid")]) + ggtitle("Part 1 of Bullet 1 in black")
+  geom_line(aes(y, resid, group = x), colour="black", size=1, alpha=0.2, 
+             data = filter(db1, part==1, x <= 80*1.5625^2, x >=75*1.5625^2)[,c("y", "resid", "x")]) + ggtitle("Part 1 of Bullet 1 in black")
 @
 
-If we perform a similar routine for part 2 of bullet 1, we obtain the figure give in Figure \ref{fig:circlefits2}.
+If we perform a similar routine for part 2 of bullet 1, we obtain the figure given in Figure \ref{fig:circlefits2}.
 
 <<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.', fig.pos='H'>>=
-qplot(y, resid, data=subset(db2, x <= 80 & x >=75), colour=factor(x)) + 
+qplot(y, resid, data=subset(db2, x <= 80*1.5625^2 & x >=75*1.5625^2), colour=factor(x), geom="line", size=I(1)) + 
   facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_point(aes(y, resid), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==2, x <= 80, x >=75)[,c("y", "resid")]) + ggtitle("Part 2 of Bullet 1 in black")
+  geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
+             data = filter(db1, part==2, x <= 80*1.5625^2, x >=75*1.5625^2)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
 @

### 2015-11-17 imaging-paper.Rnw
manual match

@@
 That is, on y coordinate system ranging from 0 to 1000 as these images, are, the grooves will appear between $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
-Hence, the algorith for removing the grooves can be described as follows:
+Hence, the algorithm for removing the grooves can be described as follows:
 
@@
 
+\begin{figure}
+\includegraphics[width=\textwidth]{images/side-by-side-manual-match.png}
+\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of two lands from different bullets.}}
+\end{figure}
 \section{Application}

### 2015-11-17 code/process_bullets.R

@@ -43,13 +43,13 @@ crosscuts <- crosscuts[crosscuts > 50]
 crosscuts <- crosscuts[crosscuts < 150]
 
 list_of_fits <- lapply(crosscuts, function(x) {
-  br111 <- get_bullet(path, x)
+  br111 <- get_bullet(path, x = x)
   br111.groove <- get_grooves(br111)
   br111.groove$plot
   fit_loess(br111, br111.groove)
 })
 
-lof <- lapply(list_of_fits, function(x) x$data) %>% bind_rows
+lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
 qplot(x, y, fill = resid, data=lof, geom="tile") + 
   scale_fill_gradient2() + theme_bw()
 
@@ -66,7 +66,7 @@ list_of_fits2 <- lapply(crosscuts, function(x) {
   fit_loess(br111, br111.groove)
 })
 
-lof2 <- lapply(list_of_fits2, function(x) x$data) %>% bind_rows
+lof2 <- lapply(list_of_fits2, function(x) x$resid$data) %>% bind_rows
 lof$bullet <- 1
 lof2$bullet <- 2
 
@@ -88,11 +88,9 @@ qplot(x, y, fill = resid, data=LOF, geom="tile", facets=~bullet) +
                        values = c(0,0.5, 0.75, 0.9, 1)) + theme_bw()
 
 
-subLOF1 <- LOF %>% filter(bullet == 1, x <= 75)
-subLOF2 <- LOF %>% filter(bullet == 2, x > 75, x < 100)
+subLOF1 <- LOF %>% filter(bullet == 1, x <= 100)
+subLOF2 <- LOF %>% filter(bullet == 2, x > 100)
 
-#subLOF$y <- subLOF$y %/% .64
-#subLOF1$y <- subLOF1$y + 11 # why doesn't this work??
 subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
 
 qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
@@ -100,6 +98,18 @@ qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") +
                        colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
                        values = c(0,0.5, 0.75, 0.9, 1)) +
   theme_bw()
+
[…]

### 2015-11-17 imaging-paper.Rnw
figure width
added code and different view

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{images/side-by-side-manual-match.png}
+<<side-by-side, echo=FALSE, eval=FALSE>>=
+paths <- file.path(datadir, dir(datadir))
+
+br111 <- read.x3p(paths[5])
+crosscuts <- unique(fortify_x3p(br111)$x)
+crosscuts <- crosscuts[crosscuts > 50]
+crosscuts <- crosscuts[crosscuts < 150]
+
+list_of_fits <- lapply(crosscuts, function(x) {
+  br111 <- get_bullet(paths[5], x = x)
+  br111.groove <- get_grooves(br111)
+  br111.groove$plot
+  fit_loess(br111, br111.groove)
+})
+
+lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
+
+br111 <- read.x3p(paths[7])
+crosscuts <- unique(fortify_x3p(br111)$x)
+crosscuts <- crosscuts[crosscuts > 50]
+crosscuts <- crosscuts[crosscuts < 150]
+
+list_of_fits2 <- lapply(crosscuts, function(x) {
+  br111 <- get_bullet(paths[7], x)
+  br111.groove <- get_grooves(br111)
+  br111.groove$plot
+  fit_loess(br111, br111.groove)
+})
+
+lof2 <- lapply(list_of_fits2, function(x) x$resid$data) %>% bind_rows
+lof$bullet <- 1
+lof2$bullet <- 2
+
+LOF <- rbind(lof, lof2)
+
+subLOF1 <- LOF %>% filter(bullet == 1, x <= 100)
+subLOF2 <- LOF %>% filter(bullet == 2, x > 100)
+
+subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
+subLOF1$y <- subLOF1$y + 19*1.5625 # working now!!!
+subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
+
+qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
+  scale_fill_gradientn(limits=c(-5,5), 
+                       colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
+                       values = c(0,0.5, 0.75, 0.9, 1)) +
+  theme_bw()
+@
+
+\includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
 \caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of two lands from different bullets.}}
 \end{figure}
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{images/match-rgl.png}
+\caption{\label{fig:manualmatch}\hh{Different view of the manually adjusted side-by-side comparison of two lands from different bullets.}}
+\end{figure}
+
 \section{Application}

### 2015-11-17 code/process_bullets.R

@@ -109,13 +109,22 @@ qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") +
                        values = c(0,0.5, 0.75, 0.9, 1)) +
   theme_bw()
 
+ggsave("../code/images/side-by-side-manual-match.png", width=5, height=5)
 
+library(reshape2)
+z <- acast(data=subLOF, x~y, value.var="resid")
+nbcol = 256
+color = rev(rainbow(256, start = 0/6, end = 2/6))
+zcol = cut(z, nbcol)
+persp3d(z, type="surface", col=color[zcol], aspect=c(1,1,0.2))
 
 qplot(x, y, fill = 5*sign(resid)*sqrt(abs(resid/5)), colour=I(NA), data=subLOF, geom="tile") + 
   scale_fill_gradientn(limits=c(-5,5), 
                        colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
                        values = c(0,0.5, 0.75, 0.9, 1)) +
   theme_bw()
+
+
 
 qplot(y, resid, colour=resid, data=subset(LOF, x == 99.84), facets=~bullet) + 
   scale_colour_gradient2(limits=c(-5,5)) + theme_bw()

### 2015-11-18 imaging-paper.Rnw
image name
moving images

@@
 
+\begin{figure}[hbtp]
+  \centering
+    \subfloat[View from the side.\label{fig:sidex3p}]{%
+      \includegraphics[width=\linewidth]{images/sidex3p.png}
+    }
+    \hfill
+    \subfloat[View from the top.\label{fig:topx3p}]{%
+    \includegraphics[width=\linewidth]{images/topx3p.png}
+    }
+\caption{An example of a scan of a bullet land.}
+\end{figure}
 Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\linewidth]{images/sidex3p.png}
-  \caption{An example of a scan of a bullet land, viewed from the side.}
-  \label{fig:sidex3p}
-\end{figure}
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\linewidth]{images/topx3p.png}
-  \caption{An example of a scan of a bullet land, viewed from the top.}
-  \label{fig:topx3p}
-\end{figure}
 
@@
 
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.width=6, out.width='\\linewidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across fixed x values.
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2.5, fig.width=6, out.width='\\linewidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across fixed x values.
 Note that the global features dominate any deviations, which would correspond to markings.'>>=
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{images/match-rgl.png}
+\includegraphics[width=\columnwidth]{images/matchup-rgl.png}
 \caption{\label{fig:manualmatch}\hh{Different view of the manually adjusted side-by-side comparison of two lands from different bullets.}}

### 2015-11-18 turk20/turk20_analysis.Rmd
Add the app and some other related files, x3prproto not needed

---
title: "Turk20 Analysis"
author: "Eric Hare"
date: "November 2, 2015"
output: html_document
---

```{r}
results <- read.csv("turk20_results.csv")
users <- read.csv("turk20_users.csv")
details <- read.csv("picture-details.csv")
```

```{r}
library(ggplot2)
library(dplyr)

merged <- merge(results, details[,c("pic_id", "obs_plot_location")])
merged$response_no <- factor(merged$response_no)

merged_filter <- merged %>% filter(!(nick_name %in% c("AG7F3QKOQJGYR", "558ab68afdf99b6567d0081d")))

lapply(unique(merged$pic_id), function(id) {
    subbed <- merged[merged$pic_id == id,]
    subbed$dataplot <- subbed$response_no == subbed$obs_plot_location

    qplot(response_no, geom = "bar", data = subbed, fill = dataplot) +   
        scale_x_discrete(drop = FALSE) +
        ylim(c(0, 20)) +
        theme_bw() +
        ggtitle(paste("Distribution of Responses for Plot", id))
})

pvalues <- sapply(unique(merged$pic_id), function(id) {
    subbed <- merged[merged$pic_id == id,]
    subbed$dataplot <- subbed$response_no == subbed$obs_plot_location

    m <- 12
    K <- nrow(subbed)
    x <- sum(subbed$dataplot)

    sum(choose(K, x:K) * (1 / m)^(x:K) * ((m - 1) / m)^(K - x:K))
})

sapply(unique(merged_filter$pic_id), function(id) {
    subbed <- merged_filter[merged_filter$pic_id == id,]
    subbed$dataplot <- subbed$response_no == subbed$obs_plot_location

    sum(subbed$dataplot)
})

pvalues < .05
```

### 2015-11-18 app/server.R

@@ -1,0 +1,121 @@
+library(shiny)
+library(x3pr)
+library(x3prplus)
+library(plotly)
+library(gridExtra)
+library(zoo)
+
+source(system.file("gui/view", "helpers.R", package = "x3pr"))
+
+Sys.setenv("plotly_username" = "erichare")
+Sys.setenv("plotly_api_key" = "xd0oxpeept")
+
+get_grooves <- function(bullet) {
+    left <- subset(bullet, y <= 500)
+    right <- subset(bullet, y > 500)
+    
+    leftsmooth <- c(NA, NA, rollapply(left$value, 5, function(x) mean(x, na.rm = TRUE)), NA, NA)
+    rightsmooth <- c(NA, NA, rollapply(right$value, 5, function(x) mean(x, na.rm = TRUE)), NA, NA)
+    
+    final.left <- left[which.min(leftsmooth[-(1:100)]) + 100,]
+    final.right <- right[which.min(rightsmooth[-((length(rightsmooth) - 99):length(rightsmooth))]),]
+    
+    p <- qplot(data = bullet, y, value) +
+        theme_bw() + coord_equal() +
+        geom_vline(xintercept = final.left$y, colour = "red") +
+        geom_vline(xintercept = final.right$y, colour = "blue")
+    
+    return(list(groove = c(final.left$y, final.right$y), plot = p))
+}
+
+fit_loess <- function(bullet, groove) {
+    
+    bullet_filter <- subset(bullet, !is.na(value) & y > groove$groove[1] & y < groove$groove[2])
+    my.loess <- loess(value ~ y, data = bullet_filter)
+    bullet_filter$fitted <- fitted(my.loess)
+    bullet_filter$resid <- resid(my.loess)
+    
+    #qplot(data = bullet_filter, y, value) +
+    #    theme_bw() + coord_equal() +
+    #    geom_smooth()
+    
+    qplot(data = bullet_filter, y, resid) +
+        theme_bw()
+}
+
+shinyServer(function(input, output, session) {
+    
+    bullet1 <- reactive({
+        if (is.null(input$file1)) return(read.x3p("images/Hamby252_3DX3P1of2/Br1 Bullet 2-5.x3p"))
+        else return(read.x3p(input$file1$datapath))
+    })
+    
+    bullet2 <- reactive({
+        if (is.null(input$file2)) return(read.x3p("images/Hamby252_3DX3P1of2/Br1 Bullet 1-3.x3p"))
[…]

### 2015-11-18 app/ui.R

@@ -1,0 +1,40 @@
+library(shiny)
+library(shinythemes)
+library(plotly)
+
+source(system.file("gui/view", "helpers.R", package="x3pr"))
+
+shinyUI(fluidPage(theme = shinytheme("cerulean"),
+    headerPanel("Bullet Rotation Prototype"),
+    
+    sidebarLayout(
+        sidebarPanel(width = 3,
+            h4("Bullet Options"),
+            fileInput("file1", "First Bullet"),
+            fileInput("file2", "Second Bullet"),
+            
+            hr(),
+            
+            h4("Plot Options"),
+            sliderInput("subsample", "Subsample Factor", min = 1, max = 20, value = 2),
+            sliderInput("ambient_lighting", "Ambient Lighting", min = 0, max = 1, step = 0.1, value = 0.8),
+            sliderInput("diffuse_lighting", "Diffuse Lighting", min = 0, max = 1, step = 0.1, value = 0.8),
+            sliderInput("specular_lighting", "Specular Lighting", min = 0, max = 2, step = 0.05, value = 0.05),
+            sliderInput("roughness_lighting", "Roughness Lighting", min = 0, max = 1, step = 0.1, value = 0.5),
+            sliderInput("fresnel_lighting", "Fresnel Lighting", min = 0, max = 5, step = 0.1, value = 0.2),
+            
+            
+            hr(),
+            
+            h4("Residuals"),
+            sliderInput("xcoord", "X Coordinate", min = 1, max = 1000, value = 99, step = 1),
+            
+            actionButton("compute", "Compute Residuals")
+        ),
+        
+        mainPanel(width = 9,
+            plotlyOutput("trendPlot", height = "700px"),
+            plotOutput("residuals")
+        )
+    )
+))

### 2015-11-18 app/server.R
Take away x3pr dependency so we dont load rgl

@@ -1,5 +1,5 @@
 library(shiny)
-library(x3pr)
+library(XML)
 library(x3prplus)
 library(plotly)
 library(gridExtra)
@@ -46,13 +46,13 @@ fit_loess <- function(bullet, groove) {
 shinyServer(function(input, output, session) {
     
     bullet1 <- reactive({
-        if (is.null(input$file1)) return(read.x3p("images/Hamby252_3DX3P1of2/Br1 Bullet 2-5.x3p"))
-        else return(read.x3p(input$file1$datapath))
+        if (is.null(input$file1)) return(x3pr::read.x3p("images/Hamby252_3DX3P1of2/Br1 Bullet 2-5.x3p"))
+        else return(x3pr::read.x3p(input$file1$datapath))
     })
     
     bullet2 <- reactive({
-        if (is.null(input$file2)) return(read.x3p("images/Hamby252_3DX3P1of2/Br1 Bullet 1-3.x3p"))
-        else return(read.x3p(input$file2$datapath))
+        if (is.null(input$file2)) return(x3pr::read.x3p("images/Hamby252_3DX3P1of2/Br1 Bullet 1-3.x3p"))
+        else return(x3pr::read.x3p(input$file2$datapath))
     })
     
     theSurface <- reactive({

### 2015-11-18 imaging-paper.Rnw
Ignore the pdf files
single datadir path

@@
 opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|", dev="CairoPDF")
-datadir <- "data"
 imgdir <- "Figure/"
 codedir <- "code"
+datadir <- "app/images/Hamby252_3DX3P1of2/"
 
@@
 library(x3prplus)
-
-datadir <- "~/GitHub/x3prproto/images/Hamby252_3DX3P1of2/"
 

### 2015-11-18 imaging-paper.Rnw
aspect ratio figure 2

@@
     }
-\caption{An example of a scan of a bullet land.}
+\caption{An example of a scan of a bullet land.
+\hh{XXX is the code for these pictures somewhere in the code folder?
+Let's put a link/note to it here.
+The aspect ratio of these 3D files should be adjusted such that it faithfully shows the actual geography.
+That should happen in the function plot3D.x3p.file }}
 \end{figure}

### 2015-11-18 app/server.R
Update app to overlay residual plots on same plot

@@ -10,39 +10,6 @@ source(system.file("gui/view", "helpers.R", package = "x3pr"))
 Sys.setenv("plotly_username" = "erichare")
 Sys.setenv("plotly_api_key" = "xd0oxpeept")
 
-get_grooves <- function(bullet) {
-    left <- subset(bullet, y <= 500)
-    right <- subset(bullet, y > 500)
-    
-    leftsmooth <- c(NA, NA, rollapply(left$value, 5, function(x) mean(x, na.rm = TRUE)), NA, NA)
-    rightsmooth <- c(NA, NA, rollapply(right$value, 5, function(x) mean(x, na.rm = TRUE)), NA, NA)
-    
-    final.left <- left[which.min(leftsmooth[-(1:100)]) + 100,]
-    final.right <- right[which.min(rightsmooth[-((length(rightsmooth) - 99):length(rightsmooth))]),]
-    
-    p <- qplot(data = bullet, y, value) +
-        theme_bw() + coord_equal() +
-        geom_vline(xintercept = final.left$y, colour = "red") +
-        geom_vline(xintercept = final.right$y, colour = "blue")
-    
-    return(list(groove = c(final.left$y, final.right$y), plot = p))
-}
-
-fit_loess <- function(bullet, groove) {
-    
-    bullet_filter <- subset(bullet, !is.na(value) & y > groove$groove[1] & y < groove$groove[2])
-    my.loess <- loess(value ~ y, data = bullet_filter)
-    bullet_filter$fitted <- fitted(my.loess)
-    bullet_filter$resid <- resid(my.loess)
-    
-    #qplot(data = bullet_filter, y, value) +
-    #    theme_bw() + coord_equal() +
-    #    geom_smooth()
-    
-    qplot(data = bullet_filter, y, resid) +
-        theme_bw()
-}
-
 shinyServer(function(input, output, session) {
     
     bullet1 <- reactive({
@@ -112,10 +79,16 @@ shinyServer(function(input, output, session) {
         b1.groove <- get_grooves(values$fort1_fixed)
         b2.groove <- get_grooves(values$fort2_fixed)
         
-        p1 <- fit_loess(values$fort1_fixed, b1.groove)
-        p2 <- fit_loess(values$fort2_fixed, b2.groove)
+        l1 <- fit_loess(values$fort1_fixed, b1.groove)
+        l2 <- fit_loess(values$fort2_fixed, b2.groove)
         
-        grid.arrange(p1, p2)
[…]

### 2015-11-18 app/ui.R

@@ -27,7 +27,7 @@ shinyUI(fluidPage(theme = shinytheme("cerulean"),
             hr(),
             
             h4("Residuals"),
-            sliderInput("xcoord", "X Coordinate", min = 1, max = 1000, value = 99, step = 1),
+            sliderInput("xcoord", "X Coordinate", min = 1, max = 1000, value = 136, step = 1),
             
             actionButton("compute", "Compute Residuals")
         ),

### 2015-11-18 imaging-paper.Rnw
more details on spin

@@
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.
+\hh{The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artefact, but a direct and expected consequence of the spin induced by the rifling during the firing process.
+Depending on whether a barrel is rifled clock-wise or anti-clockwise, the striations have a left or right tilt.
+The direction of the rifling is a class characteristic.
+XXX rotation in Ruger?
+twist rate?}
 

### 2015-11-18 imaging-paper.Rnw
discussion of alternatives

@@
 \item Walkthrough of the process
+\item \hh{XXX Alternative approaches: IBIS database and the NRC report against it\footnote{Fox report on the NRC report \url{http://www.foxnews.com/story/2008/03/05/nationwide-firearms-database-shot-down-by-government-sponsored-study.html}.
+NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
 \end{enumerate}

### 2015-11-20 imaging-paper.Rnw
Remove duplicated introduction
start on automatic alignment

@@
 %% the only exception to this rule is the \firstsection command
+
 \firstsection{Introduction}
@@
 
-\section{Introduction}
 Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel.
@@
 
-<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.', fig.pos='H'>>=
+<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.', fig.pos='H', cache=TRUE>>=
 qplot(y, resid, data=subset(db2, x <= 80*1.5625^2 & x >=75*1.5625^2), colour=factor(x), geom="line", size=I(1)) + 
@@
 @
+
+\section{Approach to automatic matching}
+\hh{XXX all tentative}
+Applying the  loess fit to a range of different surface cross cuts (see Figures~\ref{fig:manualmatch} and Figure~\ref{fig:manualmatch}, with cross cuts from 50 to 150) shows the striation marks across two bullets.
+Cross cuts of bullet 1 are shown on the left (cross cuts with values below 100) and cross cuts of bullets 2 are shown on the right (cross cuts with values above 100).
+A lot of the striation marks align, allowing for an easy visual assessment of the shared similarities between the two bullets.
 
@@
 \includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
-\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of two lands from different bullets.}}
+\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch}\hh{Different view of the manually adjusted side-by-side comparison of two lands from different bullets.}}
+\caption{\label{fig:manualmatch}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
 \end{figure}
@@
 \section{Application}
+We've implemented a web interface to automate many of the steps outlined in the previous section.
+The web interface, developed using Shiny[ref], allows a user to upload two x3p files representing scans of a bullet between two grooves.
+A 3D viewer, using Plotly[ref], allows the user to rotate and zoom to examine surface features of the images.
+A screenshot of the application is available in Figure \ref{fig:app}.
+
+\begin{figure*}
+\includegraphics[width=\textwidth]{images/app.png}
+\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
+\end{figure*}
 

### 2015-11-20 code/process_bullets.R

@@ -174,10 +174,28 @@ ggplot(aes(x=y, y=resid, group=bullet), data=LOFx) +
 # a span of 0.03 looks decent
 
 ggplot(aes(x=y, y=resid, colour=factor(bullet), group=bullet), data=LOFx) +
-  geom_point() +
+#  geom_point() +
   geom_line(aes(y=l50), size=0.5) + 
   scale_color_brewer(palette="Set1") +
 #  facet_grid(bullet~.) + 
   theme_bw()
 
+subLOFx1 <- subset(LOFx, bullet==1)
+subLOFx2 <- subset(LOFx, bullet==2)
+subLOFx1$y <- subLOFx1$y + 19*1.5625 # working now!!!
+subLOFx <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+ggplot(aes(x=y, y=resid, colour=factor(bullet), group=bullet), data=subLOFx) +
+  #  geom_point() +
+  geom_line(aes(y=l50), size=0.5) + 
+  scale_color_brewer(palette="Set1") +
+  #  facet_grid(bullet~.) + 
+  theme_bw()
+
+# create a signature by using a threshold value - 
+subLOFx$r05 <- sign(subLOFx$l30) * as.numeric(abs(subLOFx$l30) > .75)
+qplot(data=subLOFx, x=y, y=l30, colour=factor(r05), geom="line", group=bullet) + facet_grid(bullet~.) +
+  geom_point(aes(y=r05))
+
+qplot(data=subLOFx, x=y, y=bullet, fill=factor(r05), geom="tile") + ylim(c(0,3))
 

### 2015-11-20 imaging-paper.Rnw
Introduce the application

@@
 
-\section{Approach to automatic matching}
-\hh{XXX all tentative}
-Applying the  loess fit to a range of different surface cross cuts (see Figures~\ref{fig:manualmatch} and Figure~\ref{fig:manualmatch}, with cross cuts from 50 to 150) shows the striation marks across two bullets.
-Cross cuts of bullet 1 are shown on the left (cross cuts with values below 100) and cross cuts of bullets 2 are shown on the right (cross cuts with values above 100).
-A lot of the striation marks align, allowing for an easy visual assessment of the shared similarities between the two bullets.
-
 \begin{figure}
@@
 \includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
-\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of two lands from different bullets.}}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch}\hh{Different view of the manually adjusted side-by-side comparison of two lands from different bullets.}}
 \end{figure}

### 2015-11-20 code/matching.R
added code for matching

@@ -1,0 +1,71 @@
+library(x3pr)
+library(x3prplus)
+
+
+processBullets <- function(paths, x = 100) {
+  br111 <- read.x3p(paths[1])
+  crosscuts <- unique(fortify_x3p(br111)$x)
+  crosscuts <- crosscuts[crosscuts >= min(x)]
+  crosscuts <- crosscuts[crosscuts <= max(x)]
+  
+  LOF <- lapply(paths, function(path) {
+    list_of_fits <- lapply(crosscuts, function(x) {
+      br111 <- get_bullet(path, x = x)
+      br111.groove <- get_grooves(br111)
+      br111.groove$plot
+      fit_loess(br111, br111.groove)
+    })
+    lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
+    lof$path <- path
+    path <- gsub("app.*//", "", as.character(path))
+    lof$bullet <- gsub(".x3p", "", path)
+    
+    lof
+  })
+  LOF %>% bind_rows()
+}
+
+smoothloess <- function(x, y, span, sub = 2) {
+  dat <- data.frame(x, y)
+  indx <- sub *(1: (nrow(dat) %/% sub))
+  subdat <- dat[indx, ]
+  lwp <- with(subdat, loess(y~x,span=span))
+  predict(lwp, newdata = dat)
+}
+
+datadir <- "app/images/Hamby252_3DX3P1of2/"
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y + 19*1.5625 # working now!!!
+lof <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+
+
+# couple of side by side comparisons
+qplot(bullet, y, fill = resid, data=lof, geom="tile") + 
+  scale_fill_gradient2() + theme_bw()
+
+qplot(y, resid, group=bullet, colour=bullet, data=lof, geom="line") + 
+  scale_colour_brewer(palette="Set1") + theme_bw() +
+  theme(legend.position="bottom")
+
+
+lof <- lof %>% group_by(bullet) %>% mutate(
[…]

### 2015-11-20 imaging-paper.Rnw
two cross cuts

--- 2015-11-20 imaging-paper.Rnw
@@
 
+<<twolands100, echo=FALSE>>=
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y + 19*1.5625 # working now!!!
+lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+@
+\begin{figure}
+\centering
+    \subfloat[Crosscuts at $x = 100$.\label{fig:crosscut}]{%
+<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
+qplot(y, resid, group=bullet, colour=bullet, data=lof, geom="line") + 
+  scale_colour_brewer("", palette="Set1") + theme_bw() +
+  theme(legend.position = c(.9, .2))
+@
+    }
+    \hfill
+    \subfloat[Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}]{%
+<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
+qplot(y, resid, group=bullet, colour=bullet, data=lofX, geom="line") + 
+  scale_colour_brewer(palette="Set1") + theme_bw() +
+  theme(legend.position="none")
+@
+    }
+\end{figure}
+
 \section{Application}

### 2015-11-20 code/matching.R

@@ -66,6 +66,9 @@ qplot(data=lof, x=y, y=l30, colour=type, geom="line", group=bullet) +
   geom_point(aes(y=r05))
 
 qplot(data=lof, x=y, y=bullet, fill=type, geom="tile") +
-  scale_fill_brewer(palette="Set1")
+  scale_fill_brewer(palette="Set1") + 
+  theme_bw() +
+  ylab("") +
+  theme(legend.position="bottom")
 
 

### 2015-11-20 imaging-paper.Rnw
move sections around
figure outline of matching ...

--- 2015-11-20 imaging-paper.Rnw
@@
 
+\section{Application}
+We've implemented a web interface to automate many of the steps outlined in the previous section.
+The web interface, developed using Shiny[ref], allows a user to upload two x3p files representing scans of a bullet between two grooves.
+A 3D viewer, using Plotly[ref], allows the user to rotate and zoom to examine surface features of the images.
+A screenshot of the application is available in Figure \ref{fig:app}.
+
+\begin{figure*}
+\includegraphics[width=\textwidth]{images/app.png}
+\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
+\end{figure*}
+
 \section{Approach to automatic matching}
@@
 \end{figure}
+\begin{figure}
+\includegraphics[width=\columnwidth]{images/matchup-rgl.png}
+\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\end{figure}
+
+<<twolands100, cache=TRUE,  echo=FALSE>>=
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y + 19*1.5625 # working now!!!
+lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+@
+\begin{figure}
+\centering
+    \subfloat[Crosscuts at $x = 100$.\label{fig:crosscut}]{%
+<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
+qplot(y, resid, group=bullet, colour=bullet, data=lof, 
+      geom="line", alpha=I(0.6)) + 
+  scale_colour_brewer("", palette="Set1") + theme_bw() +
+  theme(legend.position = c(.9, .2))
+@
+    }
+    \hfill
+    \subfloat[Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}]{%
+<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
+qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
+      geom="line", alpha=I(0.6)) + 
+  scale_colour_brewer(palette="Set1") + theme_bw() +
+  theme(legend.position="none")
+@
+    }
+\caption{\label{fig:cross100} Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
+\end{figure}
+
+\hh{Using (for now) a single crosscut across the surface, we extract the values and remove the class characteristics using a loess fit as before.
+This enables us to focus on the individual characterists.
+Striation marks show up in this representations as peaks and valleys.
+Figure~\ref{fig:cross100} shows lines of the surface crosscut at $x = 100$.
+A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.}
 
 \begin{figure}
-\includegraphics[width=\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\centering
+    \subfloat[Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}]{%
+<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth', warning=FALSE>>=
+lofX <- lofX %>% group_by(bullet) %>% mutate(
+  l30 = smoothloess(y, resid, span = 0.03)
+)
+lofX$r05 <- 0.75* sign(lofX$l30) * as.numeric(abs(lofX$l30) > .75)
+lofX$type <- factor(lofX$r05)
+levels(lofX$type) <- c("groove", NA, "peak")
+
+qplot(data=lofX, y, resid, colour=bullet, size=I(1)) +
+  geom_line(aes(y=l30), colour="grey30") +
+  facet_grid(bullet~.) +
+  scale_colour_brewer("", palette="Set1") +
+  theme_bw() +
+  theme(legend.position="none")
+@
+    }
+    \hfill
+    \subfloat[Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}]{%
+<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='.49\\textwidth', warning=FALSE>>=
+qplot(data=lofX, x=y, y=l30, colour=type, geom="line", group=bullet) + 
+  facet_grid(bullet~.) +
+  ylab("Individual characteristics") + 
+  geom_point(aes(y=r05)) +
+  scale_colour_brewer(palette="Set2", na.value = "grey90") +
+  theme_bw() +
+  theme(legend.position="bottom")
+@
+    }
+\hfill
+    \subfloat[ \label{fig:smoothcutc}]{%
+<<smoothtile, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=2.5, out.width='.49\\textwidth'>>=
+qplot(data=lofX, x=y, y=bullet, fill=type, geom="tile") +
+  scale_fill_brewer(palette="Set2", na.value="grey90") + 
+  theme_bw() +
+  ylab("") +
+  theme(legend.position="bottom")
+@
+    }    
+\caption{\label{fig:smoothcut} Processing surface crosscuts.}
 \end{figure}
-
-\section{Application}
-We've implemented a web interface to automate many of the steps outlined in the previous section.
-The web interface, developed using Shiny[ref], allows a user to upload two x3p files representing scans of a bullet between two grooves.
-A 3D viewer, using Plotly[ref], allows the user to rotate and zoom to examine surface features of the images.
-A screenshot of the application is available in Figure \ref{fig:app}.
-
-\begin{figure*}
-\includegraphics[width=\textwidth]{images/app.png}
-\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
-\end{figure*}
 

### 2015-11-20 imaging-paper.Rnw
Discuss lighting options

@@
 
+Several lighting options are available to tweak in order to highlight markings in on the bullets more strongly:
+
+\begin{enumerate}
+    \item \textbf{Ambient Lighting} - Lighting which is used to simulate the light naturally present in the environment.
+    \item \textbf{Diffuse Lighting} - Setting this value higher will cause the surface to re-emit a higher proportion of light.
+    \item \textbf{Specular Lighting} - The lighting which appears on the surface when illuminated.
+    \item \textbf{Roughness Lighting} - The amount of light reflected by the roughness of the surface.
+    \item \textbf{Fresnel Lighting} - Higher values create a wider, softer lighting across the surface.
+\end{enumerate}
+
 \section{Approach to automatic matching}

### 2015-11-20 imaging-paper.Rnw
start and end of lines defined

@@
     \subfloat[Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}]{%
-<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth', warning=FALSE>>=
+<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.49\\textwidth', warning=FALSE>>=
 lofX <- lofX %>% group_by(bullet) %>% mutate(
@@
 
-qplot(data=lofX, y, resid, colour=bullet, size=I(1)) +
-  geom_line(aes(y=l30), colour="grey30") +
+qplot(data=lofX, y, resid, colour=bullet, size=I(0.5), geom="line", group=bullet) +
+  geom_line(aes(y=l30), colour="grey30", size=0.75) +
   facet_grid(bullet~.) +
@@
 <<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='.49\\textwidth', warning=FALSE>>=
-qplot(data=lofX, x=y, y=l30, colour=type, geom="line", group=bullet) + 
+qplot(data=lofX, x=y, y=l30, colour=type, geom="line", 
+      group=bullet, size = I(0.75)) + 
   facet_grid(bullet~.) +
   ylab("Individual characteristics") + 
-  geom_point(aes(y=r05)) +
+#  geom_point(aes(y=r05)) +
+  geom_rect(xmin = 200, xmax = 2260, ymin = -.75, ymax = 0.75, colour=NA, fill=alpha("grey90", alpha =0.5)) +
   scale_colour_brewer(palette="Set2", na.value = "grey90") +
@@
 \hfill
-    \subfloat[ \label{fig:smoothcutc}]{%
+    \subfloat[Overlaps of the same color indicate matching striation marks.  \label{fig:smoothcutc}]{%
 <<smoothtile, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=2.5, out.width='.49\\textwidth'>>=
@@
     }    
-\caption{\label{fig:smoothcut} Processing surface crosscuts.}
+\caption{\label{fig:match} Matching striation marks: smooth (a), cut (b), and match (c).}
 \end{figure}
+
+\hh{Figure~\ref{fig:match} gives an overview of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; XXX need to vary over this threshold level next).
+Two lines are defined as a match, if their peak/groove base at the threshold overlaps.
+If there is no overlap, a peak/groove leads to a mismatch.
+We then count the number of maximum consecutive matches.
+}
 

### 2015-11-20 code/matching.R

@@ -40,7 +40,7 @@ lof <- processBullets(paths = images[c(5,7)], x = 100)
 
 subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
 subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y + 19*1.5625 # working now!!!
+subLOFx1$y <- subLOFx1$y + 20*1.5625 # working now!!!
 lof <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
 
 
@@ -57,6 +57,8 @@ qplot(y, resid, group=bullet, colour=bullet, data=lof, geom="line") +
 lof <- lof %>% group_by(bullet) %>% mutate(
   l30 = smoothloess(y, resid, span = 0.03)
 )
+lof <- subset(lof, !is.na(l30))
+
 lof$r05 <- 0.75* sign(lof$l30) * as.numeric(abs(lof$l30) > .75)
 lof$type <- factor(lof$r05)
 levels(lof$type) <- c("groove", NA, "peak")
@@ -71,4 +73,33 @@ qplot(data=lof, x=y, y=bullet, fill=type, geom="tile") +
   ylab("") +
   theme(legend.position="bottom")
 
+lof <- lof %>% group_by(bullet) %>% mutate(
+  diff = c(0, diff(r05)),
+  mins = list(c(1, which(diff != 0))),
+  maxs = list(c(which(diff != 0)-1, n())),
+  n = n(),
+  minMarks = list(mins[[1]][!is.na(type[mins[[1]]])]),
+  maxMarks = list(maxs[[1]][!is.na(type[maxs[[1]]])])
+)
+
+marks <- lof %>% group_by(bullet) %>% summarise(
+  mins = mins[1],
+  maxs = maxs[1],
+  n = n[1],
+  minMarks = minMarks[1],
+  maxMarks = maxMarks[1]
+)
+
+mins <- marks$minMarks[1][[1]]
+maxs <- marks$maxMarks[1][[1]]
+for (i in 2:nrow(marks)) {
+  mins <- c(mins, marks$n[i-1] + marks$minMarks[i][[1]])
+  maxs <- c(maxs, marks$n[i-1] + marks$maxMarks[i][[1]])
+}
+
+# start and end of the marks
+qplot(data=lof, x=y, y=l30, colour=type, geom="line", group=bullet) + 
[…]

### 2015-11-21 code/matching.R
matching

@@ -1,6 +1,7 @@
 library(x3pr)
 library(x3prplus)
 
+library(dplyr)
 
 processBullets <- function(paths, x = 100) {
   br111 <- read.x3p(paths[1])
@@ -40,7 +41,7 @@ lof <- processBullets(paths = images[c(5,7)], x = 100)
 
 subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
 subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y + 20*1.5625 # working now!!!
+subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
 lof <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
 
 
@@ -103,3 +104,81 @@ qplot(data=lof, x=y, y=l30, colour=type, geom="line", group=bullet) +
   geom_point(aes(y=r05), data=lof[mins,]) +
   geom_point(aes(y=r05), data=lof[maxs,])
 
+matches <- lof[mins, c("y", "bullet", "type")]
+matches$yend <- unlist(lof[maxs, c("y")])
+
+qplot(y = bullet, yend=bullet, x=y, xend = yend, data=matches, 
+      geom="segment", colour=type, size=I(10))
+
+##################
+# different attempt
+
+
+matches <- lof %>% group_by(y) %>% summarise(
+  potential = (length(unique(type)) == 1),
+  allnas = sum(is.na(type))/n(),
+  type1 = na.omit(type)[1],
+  type = paste(type, sep="|", collapse="|"),
+  n = n()
+)
+
+
+
+
+qplot(x = y, y = potential, colour=type1, data=matches)
+
+qplot(data=lof, x=y, y=bullet, fill=type, geom="tile") +
+  geom_point(aes(x = y, fill=type1), y=1.5, data=subset(matches, potential & allnas < 1)) +
+#  geom_point(aes(x = y), y=1.5, pch = "x", data=subset(matches, !potential)) +
+  scale_fill_brewer(palette="Set1") + 
+  theme_bw() +
+  ylab("") +
+  theme(legend.position="bottom")
+
+
+# lines are defined by white space in between 
+matches$id <- cumsum(matches$allnas == 1) + 1
+matches$lineid <- as.numeric(matches$allnas != 1) * matches$id
+
+isMatch <- function(id, type) {
[…]

### 2015-11-21 code/matching.R
annotated matches

@@ -160,7 +160,7 @@ isMatch <- function(id, type) {
 lines <- matches %>% group_by(lineid) %>% summarise(
   meany = mean(y, na.rm=T),
   miny = min(y, na.rm=T),
-  maxy = max(y, na.rm=T),
+  maxy = max(y, na.rm=T) + 1.5625,
   match = isMatch(lineid, type),
   type = type1[1]
 )
@@ -180,5 +180,6 @@ ggplot() +
   scale_fill_brewer(palette="Set1") + 
   theme_bw() +
   ylab("") +
-  theme(legend.position="bottom") 
-  
+  theme(legend.position="bottom") +
+  geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
+  geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))

### 2015-11-21 imaging-paper.Rnw
biasotti added to references

@@
 subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y + 19*1.5625 # working now!!!
+subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
 lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+# smooth
+lofX <- lofX %>% group_by(bullet) %>% mutate(
+  l30 = smoothloess(y, resid, span = 0.03)
+)
+# cut at .75
+lofX$r05 <- 0.75* sign(lofX$l30) * as.numeric(abs(lofX$l30) > .75)
+lofX$type <- factor(lofX$r05)
+levels(lofX$type) <- c("groove", NA, "peak")
+@
+
+<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
+matches <- lofX %>% group_by(y) %>% summarise(
+  potential = (length(unique(type)) == 1),
+  allnas = sum(is.na(type))/n(),
+  type1 = na.omit(type)[1],
+  type = paste(type, sep="|", collapse="|"),
+  n = n()
+)
+
+matches$id <- cumsum(matches$allnas == 1) + 1
+matches$lineid <- as.numeric(matches$allnas != 1) * matches$id
+
+isMatch <- function(id, type) {
+  if (id[1] == 0) return(FALSE)
+#  browser()
+  types <- strsplit(type, split = "|", fixed=TRUE) 
+  t1 <- sapply(types, function(x) x[1])
+  t2 <- sapply(types, function(x) x[2])
+  if (all(t1 == "NA")) return(FALSE)
+  if (all(is.na(t2))) return(FALSE)
+  if (all(t2 == "NA")) return(FALSE)
+
+  peak <- length(grep("peak", c(t1, t2))) > 0
+  groove <- length(grep("groove", c(t1, t2))) > 0
+  if (peak & groove) return(FALSE)
+
+  return(TRUE)
+}
+
+lines <- matches %>% group_by(lineid) %>% summarise(
+  meany = mean(y, na.rm=T),
+  miny = min(y, na.rm=T),
+  maxy = max(y, na.rm=T) + 1.5625,
+  match = isMatch(lineid, type),
+  type = type1[1]
+)
+
 @
@@
 <<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.49\\textwidth', warning=FALSE>>=
-lofX <- lofX %>% group_by(bullet) %>% mutate(
-  l30 = smoothloess(y, resid, span = 0.03)
-)
-lofX$r05 <- 0.75* sign(lofX$l30) * as.numeric(abs(lofX$l30) > .75)
-lofX$type <- factor(lofX$r05)
-levels(lofX$type) <- c("groove", NA, "peak")
-
-qplot(data=lofX, y, resid, colour=bullet, size=I(0.5), geom="line", group=bullet) +
-  geom_line(aes(y=l30), colour="grey30", size=0.75) +
+qplot(data=lofX, y, resid, colour=bullet, size=I(0.75), geom="line", group=bullet) +
+  geom_line(aes(y=l30), colour="grey30", size=0.5) +
   facet_grid(bullet~.) +
@@
   geom_rect(xmin = 200, xmax = 2260, ymin = -.75, ymax = 0.75, colour=NA, fill=alpha("grey90", alpha =0.5)) +
-  scale_colour_brewer(palette="Set2", na.value = "grey90") +
+  scale_colour_brewer(palette="Set2", na.value = NA) +
   theme_bw() +
@@
 \hfill
-    \subfloat[Overlaps of the same color indicate matching striation marks.  \label{fig:smoothcutc}]{%
-<<smoothtile, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=2.5, out.width='.49\\textwidth'>>=
-qplot(data=lofX, x=y, y=bullet, fill=type, geom="tile") +
-  scale_fill_brewer(palette="Set2", na.value="grey90") + 
+    \subfloat[Grey backgrounds identify a line on one of the bullets. Overlaps of the same color indicate matching striation marks.
+    Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}]{%
+<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.49\\textwidth'>>=
+ggplot() +
+  geom_rect(aes(xmin = miny, xmax = maxy), ymin = 0.25, ymax=2.75, fill="grey80", data = subset(lines, lineid != 0)) +
+  geom_tile(data=lofX, aes(x=y, y=bullet, fill=type)) + 
+  scale_y_discrete(expand = c(0.2,0)) +
+  scale_fill_brewer(palette="Set2") + 
   theme_bw() +
   ylab("") +
-  theme(legend.position="bottom")
+  theme(legend.position="bottom") +
+  geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
+  geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
 @
@@
 We then count the number of maximum consecutive matches.
+In this example, the number of consecutive matching striations (CMS) is fifteen, a very high number indicative of a match between the bullets.
+For lead bullets, such as this, Biasotti \cite{biasotti:1959} XXX need more references XXX considered four or more consecutive lines to be sufficient evidence of a match.
 }
+
+\hh{XXX next step: how does this matching work in known non-matches? distribution of CMS under a null of no match? }
 

### 2015-11-21 imaging-paper.Rnw
number of consecutive matches

@@
 )
+
 # cut at .75
-lofX$r05 <- 0.75* sign(lofX$l30) * as.numeric(abs(lofX$l30) > .75)
+threshold <- .75
+lofX$r05 <- threshold* sign(lofX$l30) * as.numeric(abs(lofX$l30) > threshold)
 lofX$type <- factor(lofX$r05)
@@
 
+\hh{XXX what would be good is to get the line identifications and project them back onto the surface. }
+
 \section{Conclusion}

### 2015-11-21 imaging-paper.Rnw
varying the threshold

@@
 
-\hh{Figure~\ref{fig:match} gives an overview of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; XXX need to vary over this threshold level next).
+\hh{Figure~\ref{fig:match} gives an overview of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; obviously this is a crucial parameter.
+For more discussion see Figure~\ref{fig:thresholds}).
 Two lines are defined as a match, if their peak/groove base at the threshold overlaps.
@@
 \hh{XXX what would be good is to get the line identifications and project them back onto the surface. }
+
+\paragraph{Varying threshold}
+
+In order to address the subjectivity involved in choosing a threshold level, we investigate this parameter more closely.
+Figure~\ref{fig:thresholds} shows the relationship between threshold and the maximal number of consecutive matching striae.
+For thresholds between 0.75 and 0.85 the number of CMS peaks at fifteen.
+On average, there are 6.6 CMS across the range of thresholds.
+\begin{figure}
+<<thresholds, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '0.49\\textwidth'>>=
+thresholds <- seq(0.3, 1.5, by = 0.025)
+CMS <- thresholds %>% lapply(function(threshold) {
+ lines <- striation_identify(lofX, threshold = threshold)
+ cms <- CMS(lines$match)
+ data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
+}) %>% bind_rows()
+
+qplot(threshold, maxCMS, data=CMS) + theme_bw() +
+  ylab("Number of maximal CMS") +
+  ylim(c(0,15))
+@
+\caption{\label{fig:thresholds}Varying the threshold levels results in different number of consecutive matching striae between the two bullets.
+\hh{XXX Our suggestion is, to aim for the threshold with the largest number of CMS.}}
+\end{figure}
 

### 2015-11-21 code/matching.R

@@ -183,3 +183,15 @@ ggplot() +
   theme(legend.position="bottom") +
   geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
   geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
+
+
+thresholds <- seq(0.3, 1.5, by = 0.05)
+CMS <- thresholds %>% lapply(function(threshold) {
+ lines <- striation_identify(lofX, threshold = threshold)
+ cms <- CMS(lines$match)
+ data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
+}) %>% bind_rows()
+
+qplot(threshold, maxCMS, data=CMS) + theme_bw() +
+  ylab("Number of maximal CMS") +
+  ylim(c(0,15))

### 2015-11-21 code/matching.R
second barrel pictures

@@ -1,7 +1,13 @@
 library(x3pr)
 library(x3prplus)
-
 library(dplyr)
+library(ggplot2)
+library(gridExtra)
+
+datadir <- "app/images/Hamby252_3DX3P1of2/"
+images <- file.path(datadir, dir(datadir))
+
+
 
 processBullets <- function(paths, x = 100) {
   br111 <- read.x3p(paths[1])
@@ -34,8 +40,6 @@ smoothloess <- function(x, y, span, sub = 2) {
   predict(lwp, newdata = dat)
 }
 
-datadir <- "app/images/Hamby252_3DX3P1of2/"
-images <- file.path(datadir, dir(datadir))
 
 lof <- processBullets(paths = images[c(5,7)], x = 100)
 
@@ -195,3 +199,27 @@ CMS <- thresholds %>% lapply(function(threshold) {
 qplot(threshold, maxCMS, data=CMS) + theme_bw() +
   ylab("Number of maximal CMS") +
   ylim(c(0,15))
+
+###############
+# how do things look like between non-matches?
+
+
+list_of_matches <- lapply(8:12, function(i) {
+  lof <- processBullets(paths = images[c(5,i)], x = 100)
+  lines <- striation_identify(lof, threshold = 0.75)
+  title <- gsub("app.*//","", images[i])
+  title <- gsub(".x3p","", title)
+  ggplot() +
+    theme_bw() + 
+    geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=subset(lines, lineid!=0)) +
+    geom_line(aes(x = y, y = resid, colour = bullet),  data = lof) +
+    scale_colour_brewer(palette="Set1") +
+    theme(legend.position = "none") + 
+    ylim(c(-6,6)) +
+    geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match & lineid !=0)) +
+    geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match & lineid !=0)) +
+    ggtitle(title)
+})
+
+grid.arrange(list_of_matches[[1]], list_of_matches[[2]], list_of_matches[[3]], 
+             list_of_matches[[4]], list_of_matches[[5]], ncol = 2)

### 2015-11-21 imaging-paper.Rnw
Add scales as a dependency

@@
 library(ggplot2)
+library(scales)
 library(dplyr)

### 2015-11-21 code/matching.R
other barrel

@@ -204,15 +204,23 @@ qplot(threshold, maxCMS, data=CMS) + theme_bw() +
 # how do things look like between non-matches?
 
 
-list_of_matches <- lapply(8:12, function(i) {
-  lof <- processBullets(paths = images[c(5,i)], x = 100)
-  lines <- striation_identify(lof, threshold = 0.75)
+list_of_matches <- lapply(19:24, function(i) {
+  lof <- processBullets(paths = images[c(18,i)], x = 100)
+  mins <- lof %>% group_by(bullet) %>% summarise(miny = min(y))
+  subLOFx1 <- subset(lof, bullet==unique(lof$bullet)[1])
+  subLOFx2 <- subset(lof, bullet==unique(lof$bullet)[2]) 
+  browser()
+  
+  subLOFx1$y <- subLOFx1$y + diff(mins$miny) # working now!!!
+  lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+  
+  lines <- striation_identify(lofX, threshold = 0.75)
   title <- gsub("app.*//","", images[i])
   title <- gsub(".x3p","", title)
   ggplot() +
     theme_bw() + 
     geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=subset(lines, lineid!=0)) +
-    geom_line(aes(x = y, y = resid, colour = bullet),  data = lof) +
+    geom_line(aes(x = y, y = resid, colour = bullet),  data = lofX) +
     scale_colour_brewer(palette="Set1") +
     theme(legend.position = "none") + 
     ylim(c(-6,6)) +
@@ -222,4 +230,5 @@ list_of_matches <- lapply(8:12, function(i) {
 })
 
 grid.arrange(list_of_matches[[1]], list_of_matches[[2]], list_of_matches[[3]], 
-             list_of_matches[[4]], list_of_matches[[5]], ncol = 2)
+             list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
+             ncol = 2)

### 2015-11-23 code/matching.R
fiddling with horizontal alignment. no solution yet

@@ -206,21 +206,27 @@ qplot(threshold, maxCMS, data=CMS) + theme_bw() +
 # biggest problem right now: horizontal alignment. 
 list_of_matches <- lapply(19:24, function(i) {
   lof <- processBullets(paths = images[c(18,i)], x = 100)
+  lof <- lof %>% group_by(bullet) %>% mutate(
+    l30 = smoothloess(y, resid, span = 0.03)
+  )
+  lof$l30 <- pmin(5, lof$l30)
+  lof$l30 <- pmax(-5, lof$l30)
+  
   subLOFx1 <- subset(lof, bullet==unique(lof$bullet)[1])
   subLOFx2 <- subset(lof, bullet==unique(lof$bullet)[2]) 
-  ccf <- ccf(subLOFx1$resid, subLOFx2$resid, plot = FALSE)
+  ccf <- ccf(subLOFx1$l30, subLOFx2$l30, plot = TRUE, lag.max=100, na.action = na.omit)
   lag <- ccf$lag[which.max(ccf$acf)]
-  subLOFx1$y <- subLOFx1$y + lag * 1.5625 # working now!!!
+  subLOFx1$y <- subLOFx1$y - 3*lag * 1.5625 # amount of shifting should just be lag * y.inc
   lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
 
-  browser()  
-  lines <- striation_identify(lofX, threshold = 0.75)
+#  browser()  
+  lines <- striation_identify(lofX, threshold = .75)
   title <- gsub("app.*//","", images[i])
   title <- gsub(".x3p","", title)
   ggplot() +
     theme_bw() + 
     geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=subset(lines, lineid!=0)) +
-    geom_line(aes(x = y, y = resid, colour = bullet),  data = lofX) +
+    geom_line(aes(x = y, y = l30, colour = bullet),  data = lofX) +
     scale_colour_brewer(palette="Set1") +
     theme(legend.position = "none") + 
     ylim(c(-6,6)) +

### 2015-11-23 imaging-paper.Rnw
got to match the striation_identify with the rest again

@@
 )
-
+lines <- subset(lines, lineid != 0)
 @

### 2015-11-23 code/matching.R

@@ -205,22 +205,18 @@ qplot(threshold, maxCMS, data=CMS) + theme_bw() +
 
 # biggest problem right now: horizontal alignment. 
 list_of_matches <- lapply(19:24, function(i) {
-  lof <- processBullets(paths = images[c(18,i)], x = 100)
-  lof <- lof %>% group_by(bullet) %>% mutate(
-    l30 = smoothloess(y, resid, span = 0.03)
-  )
-  lof$l30 <- pmin(5, lof$l30)
-  lof$l30 <- pmax(-5, lof$l30)
+  lof <- processBullets(paths = images[c(5,7)], x = 100)
+  lof <- bulletSmooth(lof)
   
   subLOFx1 <- subset(lof, bullet==unique(lof$bullet)[1])
   subLOFx2 <- subset(lof, bullet==unique(lof$bullet)[2]) 
   ccf <- ccf(subLOFx1$l30, subLOFx2$l30, plot = TRUE, lag.max=100, na.action = na.omit)
   lag <- ccf$lag[which.max(ccf$acf)]
-  subLOFx1$y <- subLOFx1$y - 3*lag * 1.5625 # amount of shifting should just be lag * y.inc
+  subLOFx1$y <- subLOFx1$y - 2 * lag * 1.5625 # amount of shifting should just be lag * y.inc
   lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
 
 #  browser()  
-  lines <- striation_identify(lofX, threshold = .75)
+  lines <- striation_identify(lofX, threshold = 0.75)
   title <- gsub("app.*//","", images[i])
   title <- gsub(".x3p","", title)
   ggplot() +
@@ -228,7 +224,7 @@ list_of_matches <- lapply(19:24, function(i) {
     geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=subset(lines, lineid!=0)) +
     geom_line(aes(x = y, y = l30, colour = bullet),  data = lofX) +
     scale_colour_brewer(palette="Set1") +
-    theme(legend.position = "none") + 
+    theme(legend.position = "bottom") + 
     ylim(c(-6,6)) +
     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match & lineid !=0)) +
     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match & lineid !=0)) +

### 2015-11-23 code/matching.R
found the error in the horizontal adjustment

@@ -205,26 +205,41 @@ qplot(threshold, maxCMS, data=CMS) + theme_bw() +
 
 # biggest problem right now: horizontal alignment. 
 list_of_matches <- lapply(19:24, function(i) {
-  lof <- processBullets(paths = images[c(5,7)], x = 100)
+  cat(i)
+  lof <- processBullets(paths = images[c(15,i)], x = 100)
   lof <- bulletSmooth(lof)
   
   subLOFx1 <- subset(lof, bullet==unique(lof$bullet)[1])
   subLOFx2 <- subset(lof, bullet==unique(lof$bullet)[2]) 
-  ccf <- ccf(subLOFx1$l30, subLOFx2$l30, plot = TRUE, lag.max=100, na.action = na.omit)
+  
+  # ccf assumes that both time series start at the same y
+  subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
+  subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
+  ccf <- ccf(subLOFx1$l30, subLOFx2$l30, plot = FALSE, lag.max=150, na.action = na.omit)
   lag <- ccf$lag[which.max(ccf$acf)]
-  subLOFx1$y <- subLOFx1$y - 2 * lag * 1.5625 # amount of shifting should just be lag * y.inc
+  
+  subLOFx1$y <- subLOFx1$y -  lag * 1.5625 # amount of shifting should just be lag * y.inc
   lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
 
 #  browser()  
-  lines <- striation_identify(lofX, threshold = 0.75)
+  CMS <- thresholds %>% lapply(function(threshold) {
+    lines <- striation_identify(lofX, threshold = threshold)
+    cms <- CMS(lines$match)
+    data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
+  }) %>% bind_rows()
+  
+  threshold = CMS$threshold[which.max(CMS$maxCMS)]
+  lines <- striation_identify(lofX, threshold = threshold)
   title <- gsub("app.*//","", images[i])
   title <- gsub(".x3p","", title)
   ggplot() +
     theme_bw() + 
     geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=subset(lines, lineid!=0)) +
     geom_line(aes(x = y, y = l30, colour = bullet),  data = lofX) +
[…]

### 2015-11-23 imaging-paper.Rnw
have a look at th ebarrel 2 comparisons!

@@
 
+\begin{figure}
+<<barrel2, echo=FALSE, fig.width = 7.5, fig.height=7.5, out.width='\\linewidth', warning=FALSE>>=
+library(grid)
+library(gridExtra)
+thresholds <- seq(0.3, 1.5, by = 0.05)
+list_of_matches <- lapply(19:24, function(i) {
+  lof <- processBullets(paths = images[c(15,i)], x = 100)
+  lof <- bulletSmooth(lof)
+
+  subLOFx1 <- subset(lof, bullet==unique(lof$bullet)[1])
+  subLOFx2 <- subset(lof, bullet==unique(lof$bullet)[2])
+
+  # ccf assumes that both time series start at the same y
+  subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
+  subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
+  ccf <- ccf(subLOFx1$l30, subLOFx2$l30, plot = FALSE, lag.max=150, na.action = na.omit)
+  lag <- ccf$lag[which.max(ccf$acf)]
+
+  subLOFx1$y <- subLOFx1$y -  lag * 1.5625 # amount of shifting should just be lag * y.inc
+  lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+#  browser()  
+  CMS <- thresholds %>% lapply(function(threshold) {
+    lines <- striation_identify(lofX, threshold = threshold)
+    cms <- CMS(lines$match)
+    data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
+  }) %>% bind_rows()
+
+  threshold = CMS$threshold[which.max(CMS$maxCMS)]
+  lines <- striation_identify(lofX, threshold = threshold)
+  title <- gsub("app.*//","", images[i])
+  title <- gsub(".x3p","", title)
+  ggplot() +
+    theme_bw() + 
+    geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=subset(lines, lineid!=0)) +
+    geom_line(aes(x = y, y = l30, colour = bullet),  data = lofX) +
+    geom_hline(yintercept = threshold) +
+    geom_hline(yintercept = - threshold) +
+    scale_colour_brewer(palette="Set1") +
+    theme(legend.position = "none") + 
+    ylim(c(-6,6)) + ylab("") + xlab("") +
+    geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match & lineid !=0)) +
+    geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match & lineid !=0)) +
+    ggtitle(title) +
+    theme(plot.margin=unit(c(1,0,0,0), unit="line"))
+})
+
+grid.arrange(list_of_matches[[1]], list_of_matches[[2]], list_of_matches[[3]], 
+             list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
+             ncol = 2)
+@
+\caption{\label{fig:barrel2}All lands from the second bullet shot from barrel 2 are compared to the third land of the first bullet shot from barrel 2 (in red).
+The largest number of CMS is seven, observed with land four of the second bullet.
+}
+\end{figure}
+
 \hh{XXX what would be good is to get the line identifications and project them back onto the surface. }

### 2015-11-23 imaging-paper.Rnw
Discussion of residual plot creation in application

@@
 
+The application includes a button which allows the automatic creation of a residual plot.
+An x coordinate value can be selected, and a cross-section of the two bullet at the x value will be taken.
+The previously described Loess regression procedure will be applied to the two cross sections.
+A residual plot will be produced beneath the surface plots allowing for an assessment of the similarity of the markings.
+It is intended that users will perform this procedure at different x values in order to see whether the markings indicate a match at different points along the bullet.
+
 \section{Approach to automatic matching}

### 2015-11-23 imaging-paper.Rnw
Fix image to use only a single x value

@@
 
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2.5, fig.width=6, out.width='\\linewidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across fixed x values.
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2.5, fig.width=6, out.width='\\linewidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across a fixed x value.
 Note that the global features dominate any deviations, which would correspond to markings.'>>=
@@
 
-qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 95*1.5625^2), y, value, colour=factor(x), geom="line", size=I(1)) +
+qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 99*1.5625^2), y, value, geom="line", size=I(1)) +
   geom_line(aes(x=xpred, y=ypred, group=x), 
             colour="grey30", size=0.25) +
-  scale_colour_brewer("Cross-sections at x", palette="Paired") + 
   ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 

### 2015-11-23 imaging-paper.Rnw
Paragraph about rotating images

@@
 A screenshot of the application is available in Figure \ref{fig:app}.
+The application is available at the url \url{http://erichare.shinyapps.io/x3prproto}.
 
@@
 \end{figure*}
+
+The mouse is the primary control input for interacting with the surface plots.
+One can mouse over the plots to view the coordinate values for each point.
+Clicking and dragging anywhere on the plot will allow rotation.
+Scrolling up and down allows the user to zoom in and out respectively.
+There are also a set of controls available at the top right of the page, which will allow the selection of some other rotation methods, taking of screenshots, and reseting the camera to the default view.
 

### 2015-11-23 imaging-paper.Rnw
cross-correlation added

@@
 
-subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
+#subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
 subLOF1$y <- subLOF1$y + 19*1.5625 # working now!!!
@@
 levels(lofX$type) <- c("groove", NA, "peak")
+@
+
+<<twolands100adjust, echo=FALSE>>=
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
+subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
+ccf <- ccf(subLOFx1$resid, subLOFx2$resid, lag.max=100, plot=FALSE)
+lag <- ccf$lag[which.max(ccf$acf)]
+subLOFx1$y <- subLOFx1$y - lag*1.5625 
+lofY <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
 @
@@
 <<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
-qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
+# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
+#       geom="line", alpha=I(0.6)) + 
+#   scale_colour_brewer(palette="Set1") + theme_bw() +
+#   theme(legend.position="none")
+qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
       geom="line", alpha=I(0.6)) + 
@@
 
-\hh{Using (for now) a single crosscut across the surface, we extract the values and remove the class characteristics using a loess fit as before.
+Using (for now) a single crosscut across the surface, we extract the values and remove the class characteristics using a loess fit as before.
 This enables us to focus on the individual characterists.
@@
 Figure~\ref{fig:cross100} shows lines of the surface crosscut at $x = 100$.
-A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.}
+A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.
+\hh{This horizontal shift is based on the cross-correlation between the two surface crosscuts.
+Let $f(t)$ and $g(t)$ define the values of the crosscut at $t$, then the cross-correlation between $f$ and $g$ at lag $k$ is defined as
+\[
+(f * g) (k) = \sum_t f(t+k) g(t),
+\]
+with suitable defined limits of the summation. 
+}
+\begin{figure}
+<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '\\linewidth'>>=
+lag <- ccf$lag[which.max(ccf$acf)]
+qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
+  theme_bw() + ylab("Correlation") + xlab("Lag k") +
+  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
+  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100))
+@
+\caption{\label{fig:ccf}Cross-correlation between the two crosscuts shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
+At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the crosscuts.
+Figure~\ref{fig:crosscutX} shows the lag-shifted crosscuts.}
+\end{figure}
 
@@
 
-\hh{Figure~\ref{fig:match} gives an overview of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; obviously this is a crucial parameter.
+\hh{Figure~\ref{fig:match} gives an overview of the rest of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; obviously this is a crucial parameter.
 For more discussion see Figure~\ref{fig:thresholds}).

### 2015-11-23 imaging-paper.Rnw
fiddling with ccf figure

@@
 \]
-with suitable defined limits of the summation. 
+with suitable defined limits of the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation). 
 }
@@
   geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
-  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100))
+  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
+                     minor_breaks = c(-75,-25,25, 75))
 @

### 2015-11-23 imaging-paper.Rnw
change in style file

@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-%\documentclass[electronic,journal]{vgtc}     % electronic version, journal
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Large Colored Title Article
+% LaTeX Template
+% Version 1.1 (25/11/12)
+%
+% This template has been downloaded from:
+% http://www.LaTeXTemplates.com
+%
+% Original author:
+% Frits Wenneker (http://www.howtotex.com)
+%
+% License:
+% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier. Further, ``electronic'' includes
-%% hyperreferences for more convenient online viewing.
+%----------------------------------------------------------------------------------------
+%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
+%----------------------------------------------------------------------------------------
 
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
+\documentclass[DIV=calc, paper=letter, fontsize=10pt, twocolumn]{scrartcl}	 % A4 paper and 11pt font size
 
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused.
-
-%% These three lines bring in essential packages: ``mathptmx'' for Type 1
-%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
-%% for proper handling of the times font family.
-
-\usepackage{mathptmx}
-\usepackage{graphicx}
-\usepackage{times}
-\usepackage{float}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
+\usepackage[USenglish]{babel}
+\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
+\usepackage{amsmath,amsfonts,amsthm} % Math packages
+\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
+\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
+\usepackage{booktabs} % Horizontal rules in tables
+\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
 
@@
 
-%---------------------------------------------------
-\usepackage[cmex10]{amsmath}
-\usepackage{amssymb}
+\usepackage{sectsty} % Enables custom section titles
+\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands
+
+\usepackage{fancyhdr} % Needed to define custom headers/footers
+\pagestyle{fancy} % Enables the custom headers/footers
+\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")
+
+\usepackage{url}
+
+% Headers - all currently empty
+\lhead{}
+\chead{}
+\rhead{}
+
+% Footers
+\lfoot{}
+\cfoot{}
+\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % "Page 1 of 2"
+
+\renewcommand{\headrulewidth}{0.0pt} % No header rule
+\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule
+
 \usepackage{color}
+\newcommand{\hh}[1]{{\color{magenta} #1}}
 
-\newcommand{\hh}[1]{{\color{magenta} #1}}
-% \usepackage[dvipsnames,svgnames]{xcolor}
-% \usepackage{ulem}
-% \usepackage[section]{placeins}
-\usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
-% \usepackage{sidecap}
-\usepackage{multirow}
-\usepackage{bbm}
-\usepackage{url}
-\usepackage{xr}
-\usepackage{xr-hyper}
-\usepackage{afterpage}
+\usepackage{lettrine} % Package to accentuate the first letter of the text
+\newcommand{\initial}[1]{ % Defines the command and style for the first letter
+\lettrine[lines=3,lhang=0.3,nindent=0em]{
+\color{DarkGoldenrod}
+{\textsf{#1}}}{}}
 
-%---------------------------------------------------
+\usepackage{subfig}
 
-%% This turns references into clickable hyperlinks.
-\usepackage[bookmarks,backref=section,linkcolor=black]{hyperref} %,colorlinks
-\hypersetup{
-  pdfauthor = {},
-  pdftitle = {},
-  pdfsubject = {},
-  pdfkeywords = {},
-  colorlinks=true,
-  linkcolor= black,
-  citecolor= black,
-  pageanchor=true,
-  urlcolor = black,
-  plainpages = false,
-  linktocpage
-}
+%----------------------------------------------------------------------------------------
+%	TITLE SECTION
+%----------------------------------------------------------------------------------------
 
-\externaldocument{appendix}
+\usepackage{titling} % Allows custom title configuration
 
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
+\newcommand{\HorRule}{\color{DarkGoldenrod} \rule{\linewidth}{1pt}} % Defines the gold horizontal rule around the title
 
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Theory/Model}
+\pretitle{\vspace{-30pt} \begin{flushleft} \HorRule \fontsize{50}{50} \usefont{OT1}{phv}{b}{n} \color{DarkRed} \selectfont} % Horizontal rule before the title
 
-%% allow for this line if you want the electronic option to work properly
-\vgtcinsertpkg
+\title{Matching Bullets} % Your article title
 
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
+\posttitle{\par\end{flushleft}\vskip 0.5em} % Whitespace under the title
 
-%% Paper title.
+\preauthor{\begin{flushleft}\large \lineskip 0.5em \usefont{OT1}{phv}{b}{sl} \color{DarkRed}} % Author font configuration
 
-\title{Matching bullets}
+\author{Eric Hare, Heike Hofmann} % Your name
 
-%% This is how authors are specified in the journal style
+\postauthor{\footnotesize \usefont{OT1}{phv}{m}{sl} \color{Black} % Configuration for the institution name
 
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Eric Hare and Heike Hofmann}
-\authorfooter{
-%% insert punctuation at end of each item
-\item Eric Hare is a graduate student in the Department of Statistics at Iowa State University. Email: erichare@iastate.edu.
-\item Heike Hofmann is Professor of Statistics at Iowa State University. Email: hofmann@iastate.edu.
-}
+Department of Statistics and Statistical Laboratory,
+Iowa State University% Your institution
 
-%other entries to be set up for journal
-\shortauthortitle{Hare \MakeLowercase{\textit{et al.}}: Matching Bullets}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
+\par\end{flushleft}\HorRule} % Horizontal rule after the title
 
-%% Abstract section.
-\abstract{
-abstract goes here - this is actually not part of the paper, but goes on a separate page.
-} % end of abstract
+\date{} % Add a date here if you would like one to appear underneath the title block
+
+%----------------------------------------------------------------------------------------
+
+\begin{document}
+
+\maketitle % Print the title
+
+\thispagestyle{fancy} % Enabling the custom headers/footers for the first page
+
+%----------------------------------------------------------------------------------------
+%	ABSTRACT
+%----------------------------------------------------------------------------------------
+
+% The first character should be within \initial{}
+\initial{H}\textbf{ere is some sample text to show the initial in the introductory paragraph of this template article.
+The color and lineheight of the initial can be modified in the preamble of this document.}
+
+%----------------------------------------------------------------------------------------
+%	ARTICLE CONTENTS
+%----------------------------------------------------------------------------------------
 
@@
 %% please capitalize first letter and insert punctuation after last keyword
-\keywords{Data visualization, Statistical graphics, Statistical computing.}
-
-%% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/class/1998/> for details.
-%% The ``\CCScat'' command takes four arguments.
-
-\CCScatlist{ % not used in journal version
- \CCScat{H.2.8.c}{Data and knowledge visualization}%
-{Database applications}{Database management};
- \CCScat{G.3.n}{Statistical Computing}{Probability and Statistics}{Mathematics of Computing}
-}
-
-%% Uncomment below to include a teaser figure.
-   \teaser{
-      \centering
-Teaser image wouldn't be bad, let's see whether there is  space
-%          \includegraphics[height=\textwidth]{lineup.png}
-    \caption{\protect\label{fig:teaser} Teaser image shows the final results. }
-  }
+\paragraph{Keywords:} Data visualization, Statistical graphics, Statistical computing.
 
@@
 
-%% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
-
-%% Copyright space is enabled by default as required by guidelines.
-%% It is disabled by the 'review' option or via the following command:
-% \nocopyrightspace
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\firstsection{Introduction}
-
-\maketitle
+\section{Introduction}
 

### 2015-11-23 imaging-paper.Rnw
working through pictures; exchanged subfig with subcaption

@@
 
-\usepackage{subfig}
+\usepackage{subcaption}
 
@@
   \centering
-    \subfloat[View from the side.\label{fig:sidex3p}]{%
+\begin{subfigure}[b]{\linewidth}
+\caption{View from the side.\label{fig:sidex3p}}{%
       \includegraphics[width=\linewidth]{images/sidex3p.png}
     }
-    \hfill
-    \subfloat[View from the top.\label{fig:topx3p}]{%
+\end{subfigure}    
+\begin{subfigure}[b]{\linewidth}
+    \caption{View from the top.\label{fig:topx3p}}{%
     \includegraphics[width=\linewidth]{images/topx3p.png}
     }
-\caption{An example of a scan of a bullet land.
-\hh{XXX is the code for these pictures somewhere in the code folder?
-Let's put a link/note to it here.
-The aspect ratio of these 3D files should be adjusted such that it faithfully shows the actual geography.
-That should happen in the function plot3D.x3p.file }}
+\end{subfigure}
+\caption{An example of a scan of a bullet land from one groove to the next. }
 \end{figure}
@@
 
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2.5, fig.width=6, out.width='\\linewidth', fig.pos='t!', fig.cap='Side profile of the height values of a bullet land across a fixed x value.
-Note that the global features dominate any deviations, which would correspond to markings.'>>=
+\begin{figure}
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
 library(RColorBrewer)
@@
             colour="grey30", size=0.25) +
-  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
+#  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
+  ylab("") +
   theme_bw() + 
-  theme(legend.position="bottom") + coord_equal()
+  theme(legend.position="bottom") #+ coord_equal()
 @
+\caption{\label{fig:fixedX}Side profile of the surface measurements (in $\mu m$) of a bullet land across a fixed $x$ value.
+Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
+\end{figure}
 
@@
 One reasonable approach is to fit a circle to the curve and compute the residuals from this fit.
-The scatterplot below shows the residuals of such a fit.
+
+For this, assume that $n$ data points $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ are given that are (approximately) located on a circle.
+We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.
+
+We minimize the following expression:
+\begin{equation}\label{eq:circle}
+D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2
+\end{equation}
+We do this by differentiating $D$ with respect to $r, a,$ and $b$:
+let us assume that $x_i$ and $y_i$ are centralized (i.e.
+$\sum x_i = \sum_i y_i = 0$).
+Note, if they are not, make a note of the current means, subtract them now and add them to $(\hat{a}, \hat{b})$ at the end.
+
+\noindent
+The  derivate of $D$ with respect to $r$ is:
+\begin{eqnarray*}
+\frac{d}{dr} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 r = \\
+&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right) 
+\end{eqnarray*}
+For a minimum then holds:
+\begin{equation}\label{eq:rmin}
+\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2 
+\end{equation}
+%
+%
+The  derivative of $D$ with respect to $a$ is:
+\begin{eqnarray*}
+\frac{d}{da} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 (x_i - a) = \\
+&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]
+\end{eqnarray*}
+Using (\ref{eq:rmin}) for $nr^2$  in the equation above we get:
+\begin{eqnarray*}
+\frac{d}{da} D &=& -4 \left[  \sum_i a(x_i-a)^2 +  \sum_i a(y_i-b)^2  + \right. \\
+&& \phantom{-4 \ \ } \left . \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]  = \\
+&=& -4 \left[ \sum_i (x_i-a)^2 (a + x_i - a)  + \right.\\
+&& \phantom{-4 \ \ } \left .\sum_i (x_i - a + a) (y_i - b)^2 \right] = \\
+&=& -4 \left[ \sum_i (x_i-a)^2 x_i   + \sum_i x_i  (y_i - b)^2 \right] 
+\stackrel{\begin{array}{c}\sum_i x_i = 0\\
+\sum_i y_i = 0\end{array}}{=} \\
+&=& -4 \left[ \sum_i x_i^3   + \sum_i x_i y_i^2  - 2a s_{xx} - 2b s_{xy} \right], 
+\end{eqnarray*}
+where $s_{xx} = \sum_i x_i^2, s_{xy} = \sum_i x_i y_i$ and $s_{yy} = \sum_i y_i^2$.
+
+\noindent
+Likewise, we get for the derivative of $D$ with respect to $b$:
+\begin{eqnarray*}
+\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right] 
+\end{eqnarray*}
+For the minimum we therefore get a system of two linear equations in $a$ and $b$:
+\begin{eqnarray*}
+2 s_{xx} a + 2 s_{xy} b = c_1 && \text{ with } c_1 = \sum_i x_i^3 + x_i y_i^2 \\
+2 s_{xy} a + 2 s_{yy} b = c_2 &&\text{ with } c_2 = \sum_i x_i^2 y_i + y_i^3.
+\end{eqnarray*}
+This resolves to 
+\begin{eqnarray*}
+\hat{a} &=& \frac{c_1 s_{yy} - c_2 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2},\\
+\hat{b} &=& \frac{c_2 s_{xx} - c_1 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2}, \text{ and}\\
+\hat{r^2} &=& \frac{1}{n}s_{xx} + \frac{1}{n}s_{yy} + \hat{a}^2 + \hat{b}^2.
+\end{eqnarray*}
+
+The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
-For cross sections of $x$ values between 75 and 80, there is a residual circular structure that does not show up for all cross sections.
+For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.
 
-<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
-qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred, colour=factor(x), geom="line", size=I(1)) +
-  scale_colour_brewer("x", palette="Paired") + 
+\begin{figure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:residuala}Residual structure of a surface crosscut at $x = 1.5625$ (bottom of the bullet).}
+<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
+      geom="line", size=I(1)) +
+#  scale_colour_brewer("x", palette="Paired") + 
   theme_bw() + 
@@
 @
-
-<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
-qplot(data=subset(dbr111, x <= 5*1.5625^2), y, value-ypred, colour=factor(x),
+\end{subfigure}    
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:residualb} Residual structure of a surface crosscut at $x = 100.00$}
+<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
+qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
       geom="line", size=I(1)) +
@@
 @
+\end{subfigure}
+\caption{\label{fig:residual} Residual structure of  circular fits at two different cross-section.
+The residual structure in (b) shows systematic structure, indicating that the fit is not entirely appropriate.
+}
+\end{figure}
 
@@
 \centering
-    \subfloat[Crosscuts at $x = 100$.\label{fig:crosscut}]{%
+\begin{subfigure}[t]{\linewidth}
+\caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
 <<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
@@
     }
-    \hfill
-    \subfloat[Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}]{%
+\end{subfigure}    
+\begin{subfigure}[t]{\linewidth}
+    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
 <<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
@@
     }
+\end{subfigure}
 \caption{\label{fig:cross100} Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
@@
 \centering
-    \subfloat[Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}]{%
+    \begin{subfigure}[t]{\linewidth}
+\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
 <<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.49\\textwidth', warning=FALSE>>=
@@
     }
-    \hfill
-    \subfloat[Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}]{%
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}}{%
 <<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='.49\\textwidth', warning=FALSE>>=
@@
     }
-\hfill
-    \subfloat[Grey backgrounds identify a line on one of the bullets. Overlaps of the same color indicate matching striation marks.
-    Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}]{%
+\end{subfigure}    
+\begin{subfigure}[t]{\linewidth}
+    \caption{Grey backgrounds identify a line on one of the bullets. Overlaps of the same color indicate matching striation marks.
+    Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}}{%
 <<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.49\\textwidth'>>=
@@
     }
+\end{subfigure}
 \caption{\label{fig:match} Matching striation marks: smooth (a), cut (b), and match (c).}
@@
 
-\section{Circular Fit}
-
-Assume, $n$ data points $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ are given that are (approximately) located on a circle.
-We want to estimate the location of the center and radius of the best fitting circle.
-
-We minimize the following expression:
-\begin{equation}\label{eq:circle}
-D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2
-\end{equation}
-We do this by differentiating $D$ with respect to $r, a,$ and $b$:
-let us assume that $x_i$ and $y_i$ are centralized (i.e.
-$\sum x_i = \sum_i y_i = 0$).
-Note, if they are not, make a note of the current means, subtract them now and add them to $(\hat{a}, \hat{b})$ at the end.
-
-\noindent
-The  derivate of $D$ with respect to $r$ is:
-\begin{eqnarray*}
-\frac{d}{dr} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 r = \\
-&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right) 
-\end{eqnarray*}
-For a minimum then holds:
-\begin{equation}\label{eq:rmin}
-\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2 
-\end{equation}
-%
-%
-The  derivative of $D$ with respect to $a$ is:
-\begin{eqnarray*}
-\frac{d}{da} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 (x_i - a) = \\
-&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]
-\end{eqnarray*}
-Using (\ref{eq:rmin}) for $nr^2$  in the equation above we get:
-\begin{eqnarray*}
-\frac{d}{da} D &=& -4 \left[  \sum_i a(x_i-a)^2 +  \sum_i a(y_i-b)^2  + \right. \\
-&& \phantom{-4 \ \ } \left . \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]  = \\
-&=& -4 \left[ \sum_i (x_i-a)^2 (a + x_i - a)  + \sum_i (x_i - a + a) (y_i - b)^2 \right] = \\
-&=& -4 \left[ \sum_i (x_i-a)^2 x_i   + \sum_i x_i  (y_i - b)^2 \right] \stackrel{\sum_i x_i = \sum_i y_i = 0}{=} \\
-&=& -4 \left[ \sum_i x_i^3   + \sum_i x_i y_i^2  - 2a s_{xx} - 2b s_{xy} \right], 
-\end{eqnarray*}
-where $s_{xx} = \sum_i x_i^2, s_{xy} = \sum_i x_i y_i$ and $s_{yy} = \sum_i y_i^2$.
-
-\noindent
-Likewise, we get for the derivative of $D$ with respect to $b$:
-\begin{eqnarray*}
-\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right] 
-\end{eqnarray*}
-For the minimum we therefore get a system of two linear equations in $a$ and $b$:
-\begin{eqnarray*}
-2 s_{xx} a + 2 s_{xy} b = c_1 && \text{ with } c_1 = \sum_i x_i^3 + x_i y_i^2 \\
-2 s_{xy} a + 2 s_{yy} b = c_2 &&\text{ with } c_2 = \sum_i x_i^2 y_i + y_i^3.
-\end{eqnarray*}
-This resolves to 
-\begin{eqnarray*}
-\hat{a} &=& \frac{c_1 s_{yy} - c_2 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2},\\
-\hat{b} &=& \frac{c_2 s_{xx} - c_1 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2}, \text{ and}\\
-\hat{r^2} &=& \frac{1}{n}s_{xx} + \frac{1}{n}s_{yy} + \hat{a}^2 + \hat{b}^2.
-\end{eqnarray*}
 \end{document}

### 2015-11-23 imaging-paper.Rnw
bit of editing

@@
 
-It's clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
+It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
 
@@
   theme_bw() + 
+  geom_hline(yintercept = 0, colour="grey50") +
+  ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
   theme(legend.position="bottom")
@@
       geom="line", size=I(1)) +
-  scale_colour_brewer("x", palette="Paired") + 
+#  scale_colour_brewer("x", palette="Paired") + 
+  geom_hline(yintercept = 0, colour="grey50") +
   theme_bw() + 
+  ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
   theme(legend.position="bottom")
@@
 \caption{\label{fig:residual} Residual structure of  circular fits at two different cross-section.
-The residual structure in (b) shows systematic structure, indicating that the fit is not entirely appropriate.
+Both residual plots show systematic structures, indicating that the fit is not entirely appropriate.
 }
@@
 We therefore fit a circle for each cross section of the bullet.
-Figure \ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line.
+Figure~\ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line.
 There is some global structure evident that makes focusing on the deviations a bit challenging.
@@
 
-<<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.', fig.pos='H'>>=
-qplot(y, resid, data=subset(db2, x <= 80*1.5625^2 & x >=75*1.5625^2), colour=factor(x), geom="line", size=I(1)) + 
+\begin{figure}[hbtp]
+  \centering
+\begin{subfigure}[b]{\linewidth}
+\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}{%
+<<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
+      geom="line", size=I(1), colour=I("grey30")) + 
   facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
@@
   geom_line(aes(y, resid, group = x), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==1, x <= 80*1.5625^2, x >=75*1.5625^2)[,c("y", "resid", "x")]) + ggtitle("Part 1 of Bullet 1 in black")
+             data = filter(db1, part==1, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 1 of Bullet 1 in black")
 @
-
-If we perform a similar routine for part 2 of bullet 1, we obtain the figure given in Figure \ref{fig:circlefits2}.
-
-<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.', fig.pos='H', cache=TRUE>>=
-qplot(y, resid, data=subset(db2, x <= 80*1.5625^2 & x >=75*1.5625^2), colour=factor(x), geom="line", size=I(1)) + 
+}
+\end{subfigure}    
+\begin{subfigure}[b]{\linewidth}
+    \caption{Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.\label{fig:circlefits2}}{%
+<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE,  cache=TRUE>>=
+qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
+      geom="line", size=I(1), colour=I("grey30")) + 
   facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
@@
   geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==2, x <= 80*1.5625^2, x >=75*1.5625^2)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
+             data = filter(db1, part==2, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
 @    
+}
+\end{subfigure}
+\caption{An example of a scan of a bullet land from one groove to the next. }
+\end{figure}
+
+If we perform a similar routine for part 2 of bullet 1, we obtain the figure given in Figure \ref{fig:circlefits2}.
 
@@
 
-Next, we fit a Loess regression to the data
+Next, we fit a Loess regression to the data.
+
+\hh{XXX could you include a series of plots here, walking through the individual steps of this list? I like the ones that you've already got in the loess fit function ... }
 
@@
 \caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
-<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
+<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
 qplot(y, resid, group=bullet, colour=bullet, data=lof, 
@@
     \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
-<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.49\\textwidth'>>=
+<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
 # qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
@@
 \caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
-<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.49\\textwidth', warning=FALSE>>=
+<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
 qplot(data=lofX, y, resid, colour=bullet, size=I(0.75), geom="line", group=bullet) +
@@
 \caption{Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}}{%
-<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='.49\\textwidth', warning=FALSE>>=
+<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='\\textwidth', warning=FALSE>>=
 qplot(data=lofX, x=y, y=l30, colour=type, geom="line", 
@@
     Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}}{%
-<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.49\\textwidth'>>=
+<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth'>>=
 ggplot() +
@@
 }
+
+\paragraph{Varying threshold}
+
+In order to address the subjectivity involved in choosing a threshold level, we investigate this parameter more closely.
+Figure~\ref{fig:thresholds} shows the relationship between threshold and the maximal number of consecutive matching striae.
+For thresholds between 0.75 and 0.85 the number of CMS peaks at fifteen.
+On average, there are 6.6 CMS across the range of thresholds.
+\begin{figure}
+<<thresholds, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '0.49\\textwidth'>>=
+thresholds <- seq(0.3, 1.5, by = 0.025)
+CMS <- thresholds %>% lapply(function(threshold) {
+ lines <- striation_identify(lofX, threshold = threshold)
+ cms <- CMS(lines$match)
+ data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
+}) %>% bind_rows()
+
+qplot(threshold, maxCMS, data=CMS) + theme_bw() +
+  ylab("Number of maximal CMS") +
+  ylim(c(0,15))
+@
+\caption{\label{fig:thresholds}Varying the threshold levels results in different number of consecutive matching striae between the two bullets.
+\hh{XXX Our suggestion is, to aim for the threshold with the largest number of CMS.}}
+\end{figure}
 
@@
 
-\paragraph{Varying threshold}
-
-In order to address the subjectivity involved in choosing a threshold level, we investigate this parameter more closely.
-Figure~\ref{fig:thresholds} shows the relationship between threshold and the maximal number of consecutive matching striae.
-For thresholds between 0.75 and 0.85 the number of CMS peaks at fifteen.
-On average, there are 6.6 CMS across the range of thresholds.
-\begin{figure}
-<<thresholds, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '0.49\\textwidth'>>=
-thresholds <- seq(0.3, 1.5, by = 0.025)
-CMS <- thresholds %>% lapply(function(threshold) {
- lines <- striation_identify(lofX, threshold = threshold)
- cms <- CMS(lines$match)
- data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
-}) %>% bind_rows()
-
-qplot(threshold, maxCMS, data=CMS) + theme_bw() +
-  ylab("Number of maximal CMS") +
-  ylim(c(0,15))
-@
-\caption{\label{fig:thresholds}Varying the threshold levels results in different number of consecutive matching striae between the two bullets.
-\hh{XXX Our suggestion is, to aim for the threshold with the largest number of CMS.}}
-\end{figure}
-
 \section{Conclusion}

### 2015-11-23 imaging-paper.Rnw
teensy comments

@@
 
-For this, assume that $n$ data points $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ are given that are (approximately) located on a circle.
+For this, assume that $n$ data points are given in form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle.
 We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.
@@
 
+\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.

### 2015-11-24 imaging-paper.Rnw
couple ore editorial comments

@@
   geom_hline(yintercept = 0, colour="grey50") +
-  ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
+  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
   theme(legend.position="bottom")
@@
   theme_bw() + 
-  ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
+  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
   theme(legend.position="bottom")
@@
 \caption{\label{fig:residual} Residual structure of  circular fits at two different cross-section.
-Both residual plots show systematic structures, indicating that the fit is not entirely appropriate.
+Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.
 }
@@
 A single cylinder as a fit is unlikely to be a particularly good fit, because  there seem to be quite massive deformations in vertical direction.
+\hh{Even when we fit a circle to each cross section of the bullet, we can not address all of these issues.
+While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.} \hh{XXX move the rest of the section to the Loess Fit (except for the last paragraph).}
 We therefore fit a circle for each cross section of the bullet.
-Figure~\ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line.
+\hh{XXX I don't think that we need the next figure.
+Instead of a circular fit, let's do the same thing after a loess fit.
+} Figure~\ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line.
 There is some global structure evident that makes focusing on the deviations a bit challenging.

### 2015-11-24 imaging-paper.Rnw
Include float to fix some errors in building
Describe rifling procedure, clean up footnotes

@@
 \usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
+\usepackage{float}
 
@@
 \newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\eh}[1]{{\color{blue} #1}}
 
@@
 \item Walkthrough of the process
-\item \hh{XXX Alternative approaches: IBIS database and the NRC report against it\footnote{Fox report on the NRC report \url{http://www.foxnews.com/story/2008/03/05/nationwide-firearms-database-shot-down-by-government-sponsored-study.html}.
+\item \hh{XXX Alternative approaches: IBIS database and the NRC report against it\footnote{Fox report on the NRC report \url{http://www.foxnews.com/story/2008/03/05/nationwide-firearms-
+database-shot-down-by-government-sponsored-study.html}.
 NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
@@
 
-\hh{XXX here's a paragraph missing: rifling in the barrel is creating striation marks on the bullet during the firing process (and grooves).
-These marks are claimed to be unique to the barrel - we need to find a citation for that from the SOP manual.
-Roger Thompson's writeup on firearm identification cites AFTE 1992 procedures.\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}, \footnote{AFTE 1992 theory of identification \url{http://projects.nfstc.org/firearms/module09/fir_m09_t05_05.htm}} includes the following statements, both on page 34: `The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in ``sufficient agreement".' (my highlighting) and `Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows.'}
+Rifling in the barrel creates striation marks on the bullet during the firing process (and grooves).
+These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article.
+``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows." \footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}, \footnote{AFTE 1992 theory of identification \url{http://projects.nfstc.org/firearms/module09/fir_m09_t05_05.htm}}
 

### 2015-11-24 imaging-paper.Rnw
Shorten fox news URL so it doesn't overlap

@@
 \item Walkthrough of the process
-\item \hh{XXX Alternative approaches: IBIS database and the NRC report against it\footnote{Fox report on the NRC report \url{http://www.foxnews.com/story/2008/03/05/nationwide-firearms-
-database-shot-down-by-government-sponsored-study.html}.
+\item \hh{XXX Alternative approaches: IBIS database and the NRC report against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
 NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }

### 2015-11-24 imaging-paper.Rnw
Work in the discussion on the tilt of the lines

@@
 
-\hh{I'm putting in footnotes for URLs at the moment, they should probably be included properly in the references.}
-
 Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel.
@@
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.
-\hh{The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artefact, but a direct and expected consequence of the spin induced by the rifling during the firing process.
-Depending on whether a barrel is rifled clock-wise or anti-clockwise, the striations have a left or right tilt.
+The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process.
+Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt.
 The direction of the rifling is a class characteristic.
-XXX rotation in Ruger?
-twist rate?}
 
-The primary goal in this type of forensic analysis is to identify markings.
+The primary goal in this type of forensic analysis is to identify markings in an attempt to match two bullets.
 Gun barrels should produce similar markings on bullets fired consecutively, and the careful analysis of these surface images can help us determine whether two bullets are a match.

### 2015-11-24 imaging-paper.Rnw
added references for plotly and shiny

@@
 We've implemented a web interface to automate many of the steps outlined in the previous section.
-The web interface, developed using Shiny[ref], allows a user to upload two x3p files representing scans of a bullet between two grooves.
-A 3D viewer, using Plotly[ref], allows the user to rotate and zoom to examine surface features of the images.
+The web interface, developed using Shiny \cite{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
+A 3D viewer, using Plotly \cite{plotly}, allows the user to rotate and zoom to examine surface features of the images.
 A screenshot of the application is available in Figure \ref{fig:app}.

### 2015-11-24 imaging-paper.Rnw
Fitted and residual images for Loess fit

@@
 Next, we fit a Loess regression to the data.
-
-\hh{XXX could you include a series of plots here, walking through the individual steps of this list? I like the ones that you've already got in the loess fit function ... }
+Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlayed against the processed image of part 1 of bullet 1.
+The fit seems to do a reaosnable job of capturing the structure of the image.
+Figure \ref{fig:loess_resid} shows the residuals from this fit.
+It is clear that there is a rather significant amount of autocorrelation, which make sense given the presence of striations and markings.
+Certainly, tuning the parameters for this regression, such as the Loess span, could prove to do an even better job at making the markings more clear.
 

### 2015-11-24 imaging-paper.Rnw
Images for steps of Loess fitting

@@
 library(x3prplus)
+library(grid)
+library(gridExtra)
+library(zoo)
 
@@
 \begin{enumerate}
-    \item Fix an x coordinate value.
-    \item Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
-    \item For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minimums.
-    \item Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove)
-    \item Remove all y values below the minimum and above the maximum
+    \item Fix an x coordinate value. (Top panel of Figure \ref{fig:loess_step12})
+    \item Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$. (Bottom panel of Figure \ref{fig:loess_step12})
+    \item For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minimums. (Top panel of Figure \ref{fig:loess_step34})
+    \item Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove) (Bottom panel of Figure \ref{fig:loess_step34})
+    \item Remove all y values below the minimum and above the maximum (Figure \ref{fig:loess_step5})
 \end{enumerate}
@@
 
+<<loess_step12, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Steps 1 and 2 of the Loess fit algorithm.
+The top panel is a plot of height values for part 1 of bullet 1 across different values of y, with x held constant at 243.75.
+The bottom panel is the same plot, but with the first and last 100 y values trimmed.', fig.pos='H', message=FALSE>>=
+br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
+
+p1 <- qplot(y, value, data = br111_bullet) + theme_bw()
+p2 <- qplot(y, value, data = tail(head(br111_bullet, n = -100), n = -100)) + theme_bw()
+
+grid.arrange(p1, p2, nrow = 2)
+@
+
+<<loess_step34, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Steps 3 and 4 of the Loess fitting algorithm.
+The top panel is a plot of the smoothed height values, and the bottom panel is a plot of the detected minimums.', fig.pos='H', message=FALSE>>=
+subdata <- tail(head(br111_bullet, n = -100), n = -100)
+subdata$smoothed <- c(rep(NA, 10), rollapply(subdata$value, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
+p3 <- qplot(y, smoothed, data = tail(head(subdata, n = -10), n = -10)) + theme_bw()
+
+curdata <- tail(head(subdata, n = -10), n = -10)
+min.left <- which.min(curdata$value[1:(length(curdata$value) %/% 2)])
+min.right <- which.min(curdata$value[(length(curdata$value) %/% 2):length(curdata$value)])
+
+p4 <- qplot(data=curdata, y, value) +
+        theme_bw() +
+        geom_vline(xintercept = curdata$y[min.left], colour = "red") +
+        geom_vline(xintercept = curdata$y[min.right + (length(curdata$y) %/% 2)], colour = "blue")
+
+grid.arrange(p3, p4, nrow = 2)
+@
+
+<<loess_step5, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Step 5 of the Loess fitting algorithm.
+The plot shows the result after trimming values below the first minimum, and above the second minimum.', fig.pos='H', message=FALSE>>=
+final.dat <- subset(curdata, y > curdata$y[min.left + 5] & y < curdata$y[min.right + (length(curdata$y) %/% 2) - 5])
+
+qplot(data=final.dat, y, value) +
+    theme_bw()
+@
+
 <<loess_fit, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit for part 1 of bullet 1', fig.pos='H', message=FALSE>>=
-br111 <- get_bullet(file.path("GitHub", "x3prproto", "images", "Hamby252_3DX3P1of2", "Br1 Bullet 1-1.x3p"))
-br111.groove <- get_grooves(br111)
 
-my.loess <- fit_loess(br111, br111.groove)
+br111.groove <- get_grooves(br111_bullet)
+
+my.loess <- fit_loess(br111_bullet, br111.groove)
 
@@
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit residuals for part 1 of bullet 1', fig.pos='H'>>=
-my.loess <- fit_loess(br111, br111.groove)
+my.loess <- fit_loess(br111_bullet, br111.groove)
 
@@
 <<barrel2, echo=FALSE, fig.width = 7.5, fig.height=7.5, out.width='\\linewidth', warning=FALSE>>=
-library(grid)
-library(gridExtra)
 thresholds <- seq(0.3, 1.5, by = 0.05)

### 2015-11-24 imaging-paper.Rnw
added a couple of references

@@
 \item Walkthrough of the process
-\item \hh{XXX Alternative approaches: IBIS database and the NRC report against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
+\item \hh{XXX Alternative approaches: IBIS database and the NRC report\cite{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
 NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
@@
 The report concluded that minimal research had been done exploring the reliability of these ballistic identification methods.
-In fact, the report says the forensic evidence ``including, for example, bite marks and rearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
+In fact, the report says the forensic evidence ``including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
 
@@
 Rifling in the barrel creates striation marks on the bullet during the firing process (and grooves).
-These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article.
+These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article \cite{afte:1992}.
 ``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows." \footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}, \footnote{AFTE 1992 theory of identification \url{http://projects.nfstc.org/firearms/module09/fir_m09_t05_05.htm}}
@@
 In this example, the number of consecutive matching striations (CMS) is fifteen, a very high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti \cite{biasotti:1959} XXX need more references XXX considered four or more consecutive lines to be sufficient evidence of a match.
-}
+For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
+XXX need more references XXX \cite{nichols:2003} XXX what does the paper say?
+AFTE is so frustrating!
+XXX }
 

### 2015-11-25 imaging-paper.Rnw
Lame attempt at one of the teaser sentences

@@
 % The first character should be within \initial{}
-\initial{H}\textbf{ere is some sample text to show the initial in the introductory paragraph of this template article.
-The color and lineheight of the initial can be modified in the preamble of this document.}
+\initial{F}\textbf{orensic examiners try to determine whether two guns were fired from the same barrel by analyzing them under a microscope in controlled conditions. Can we make the task simpler?}
 

### 2015-11-26 code/complete-hamby.R
trying for a complete analysis of the Hamby study

@@ -1,0 +1,43 @@
+library(x3pr)
+library(x3prplus)
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+
+knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
+knowns <- file.path(knowndatadir, dir(knowndatadir))
+
+unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
+
+
+###############
+# can we identify the barrels the unknown bullets came from?
+
+# match first unknown land
+for (j in 3:90) {
+
+reslist <- lapply(knowns, function(x) {
+  cat(x)
+  cat("\n")
+  bulletGetMaxCMS(x, unknowns[j])
+})
+save(reslist, file=sprintf("data/unkn%d.RData", j))
+}
+
+cmsdist <- sapply(reslist, function(x) x$maxCMS)
+qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
+res <- reslist[[which.max(cmsdist)]]  
+
+
+  ggplot() +
+    theme_bw() + 
+    geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+    geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
+    geom_hline(yintercept = res$threshold) +
+    geom_hline(yintercept = - res$threshold) +
+    scale_colour_brewer(palette="Set1") +
+    theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+    ylim(c(-6,6)) +
+    geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+    geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) 

### 2015-11-28 imaging-paper.Rnw
trying for a complete analysis of the Hamby study
Start refactoring that second circle figure into loess

@@
   \centering
-\begin{subfigure}[b]{\linewidth}
 \caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}{%
@@
 @
-}
-\end{subfigure}    
-\begin{subfigure}[b]{\linewidth}
-    \caption{Circle fit for each cross section of bullet 2, with part 2 of bullet 1 overlayed.\label{fig:circlefits2}}{%
-<<circlefits2, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE,  cache=TRUE>>=
-qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
-      geom="line", size=I(1), colour=I("grey30")) + 
-  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
-  theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==2, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
-@    
-}
-\end{subfigure}
-\caption{An example of a scan of a bullet land from one groove to the next. }
+
 \end{figure}
-
-If we perform a similar routine for part 2 of bullet 1, we obtain the figure given in Figure \ref{fig:circlefits2}.
 
@@
 br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
+br112_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-2.x3p", sep = "/"))
+br113_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-3.x3p", sep = "/"))
+br114_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-4.x3p", sep = "/"))
+br115_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
+br116_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-6.x3p", sep = "/"))
 
@@
 my.loess$resid
+@
+
+<<loessfits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+br111.groove <- get_grooves(br111_bullet)
+br112.groove <- get_grooves(br112_bullet)
+br113.groove <- get_grooves(br113_bullet)
+br114.groove <- get_grooves(br114_bullet)
+br115.groove <- get_grooves(br115_bullet)
+br116.groove <- get_grooves(br116_bullet)
+
+my.loess1 <- fit_loess(br111_bullet, br111.groove)
+my.loess2 <- fit_loess(br112_bullet, br112.groove)
+my.loess3 <- fit_loess(br113_bullet, br113.groove)
+my.loess4 <- fit_loess(br114_bullet, br114.groove)
+my.loess5 <- fit_loess(br115_bullet, br115.groove)
+my.loess6 <- fit_loess(br116_bullet, br116.groove)
+
+qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
+      geom="line", size=I(1), colour=I("grey30")) + 
+  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
+  theme_bw() + theme(legend.position="bottom") + 
+  geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
+             data = filter(db1, part==2, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
 @
@@
 Figure~\ref{fig:thresholds} shows the relationship between threshold and the maximal number of consecutive matching striae.
-For thresholds between 0.75 and 0.85 the number of CMS peaks at fifteen.
+For thresholds between 0.7 and 0.9 the number of CMS peaks at fifteen.
 On average, there are 6.6 CMS across the range of thresholds.

### 2015-11-28 imaging-paper.Rnw
Produce the figure

@@
 library(zoo)
+library(tidyr)
 
@@
 br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
-br112_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-2.x3p", sep = "/"))
-br113_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-3.x3p", sep = "/"))
-br114_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-4.x3p", sep = "/"))
-br115_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
-br116_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-6.x3p", sep = "/"))
+
+br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
+br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
+br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
+br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
+br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
+br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))
 
@@
 
+We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from part 1 of bullet 1.
+This is available in Figure \ref{fig:loessfits}.
+
 <<loessfits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
-br111.groove <- get_grooves(br111_bullet)
-br112.groove <- get_grooves(br112_bullet)
-br113.groove <- get_grooves(br113_bullet)
-br114.groove <- get_grooves(br114_bullet)
-br115.groove <- get_grooves(br115_bullet)
-br116.groove <- get_grooves(br116_bullet)
+br121.groove <- get_grooves(br121_bullet)
+br122.groove <- get_grooves(br122_bullet)
+br123.groove <- get_grooves(br123_bullet)
+br124.groove <- get_grooves(br124_bullet)
+br125.groove <- get_grooves(br125_bullet)
+br126.groove <- get_grooves(br126_bullet)
 
-my.loess1 <- fit_loess(br111_bullet, br111.groove)
-my.loess2 <- fit_loess(br112_bullet, br112.groove)
-my.loess3 <- fit_loess(br113_bullet, br113.groove)
-my.loess4 <- fit_loess(br114_bullet, br114.groove)
-my.loess5 <- fit_loess(br115_bullet, br115.groove)
-my.loess6 <- fit_loess(br116_bullet, br116.groove)
+my.loess1 <- fit_loess(br121_bullet, br121.groove)
+my.loess2 <- fit_loess(br122_bullet, br122.groove)
+my.loess3 <- fit_loess(br123_bullet, br123.groove)
+my.loess4 <- fit_loess(br124_bullet, br124.groove)
+my.loess5 <- fit_loess(br125_bullet, br125.groove)
+my.loess6 <- fit_loess(br126_bullet, br126.groove)
 
-qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
+my.loessactual <- fit_loess(br111_bullet, br111.groove)
+
+minrows <- min(length(my.loess1$data$resid),
+               length(my.loess2$data$resid),
+               length(my.loess3$data$resid),
+               length(my.loess4$data$resid),
+               length(my.loess5$data$resid),
+               length(my.loess6$data$resid),
+               length(my.loessactual$data$resid))
+
+my.df <- data.frame(y = my.loess1$data$y[1:minrows], resid1 = my.loess1$data$resid[1:minrows],
+                    resid2 = my.loess2$data$resid[1:minrows], resid3 = my.loess3$data$resid[1:minrows],
+                    resid4 = my.loess4$data$resid[1:minrows], resid5 = my.loess5$data$resid[1:minrows],
+                    resid6 = my.loess6$data$resid[1:minrows])
+
+melt.df <- my.df %>%
+    gather(key = part, value = resid, resid1:resid6)
+melt.df$part <- gsub("resid", "", melt.df$part)
+
+actual.df <- my.loessactual$data[1:minrows,]
+
+qplot(y, resid, data=melt.df, #colour=factor(x), 
       geom="line", size=I(1), colour=I("grey30")) + 
@@
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group =x), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==2, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 2 of Bullet 1 in black")
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=1, alpha=0.2) + ggtitle("Part 1 of Bullet 1 in black")
 @

### 2015-11-28 imaging-paper.Rnw
Redo end of circular fit section

@@
 A single cylinder as a fit is unlikely to be a particularly good fit, because  there seem to be quite massive deformations in vertical direction.
-\hh{Even when we fit a circle to each cross section of the bullet, we can not address all of these issues.
-While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.} \hh{XXX move the rest of the section to the Loess Fit (except for the last paragraph).}
-We therefore fit a circle for each cross section of the bullet.
-\hh{XXX I don't think that we need the next figure.
-Instead of a circular fit, let's do the same thing after a loess fit.
-} Figure~\ref{fig:circlefits} illustrates an attempt at matching one of the six lands to the first part of bullet 1, which is shown as a black line.
-There is some global structure evident that makes focusing on the deviations a bit challenging.
+Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, we can not address all of these issues.
+While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.
 
@@
   \centering
-\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}{%
+\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}%
 <<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
@@
 @
-
 \end{figure}
-
-It can be seen pretty clearly at this point that the deviations resulting from the grooves dominates the residual structure that would correspond to markings.
-This is unfortunate, as the grooves themselves will not typically contain markings produced by a gun barrel.
-We can also see the residual circular structure which suggests our circle fit is not entirely appropriate.
 

### 2015-11-29 code/evaluation-hamby.R
couple more matches/non-matches

@@ -42,6 +42,7 @@ CCFs <- CCFs[order(as.character(CCFs$b2)),]
 CCFs$b2 <- factor(as.character(CCFs$b2))
 
 idx <- which(CCFs$ccf > 0.6 & CCFs$distr.dist < 1.25)
+idx <- c(21, 84)
 for ( i in idx) {
   load(CCFs$data[i])
   res <- reslist[[CCFs$resID[i]]]  

### 2015-11-29 imaging-paper.Rnw
attempt at including perfect scenario in the paper ... need more discussion

@@
 
+\hh{We can use the Hamby study to run all unkowns against the known bullets. When things go perfectly, they look like in Figure~\ref{fig:hamby-perfect}.}
+
+\begin{figure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{Maximal number of CMS between Bullet land and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
+<<cms, echo=FALSE, fig.width=5, fig.height=3, out.width='\\textwidth'>>=
+load("data/unkn17.RData")
+cmsdist <- sapply(reslist, function(x) x$maxCMS)
+
+qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
+@
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{Overlaid lines of the land of the unknown bullet and the land with the top matching CMS.}
+<<top, echo=FALSE, fig.width=5, fig.height=3, out.width='\\textwidth'>>=
+res <- reslist[[which.max(cmsdist)]]  
+#res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
+
+print(ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
+  geom_hline(yintercept = res$threshold) +
+  geom_hline(yintercept = - res$threshold) +
+  scale_colour_brewer(palette="Set1") +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+
+@
+
+\caption{Overlaid lines of the land of the unknown bullet and the land with the second top matching CMS.}
+<<top2, echo=FALSE, fig.width=5, fig.height=3, out.width='\\textwidth'>>=
+#res <- reslist[[which.max(cmsdist)]]  
+res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
+
+print(ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
+  geom_hline(yintercept = res$threshold) +
+  geom_hline(yintercept = - res$threshold) +
+  scale_colour_brewer(palette="Set1") +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+
+@
+\end{subfigure}
+\end{figure}
+
 \hh{XXX what would be good is to get the line identifications and project them back onto the surface. }

### 2015-11-29 code/complete-hamby.R

@@ -35,6 +35,15 @@ for (j in 1:90) {
 }
 
 
+for (j in 1) {
+  reslist <- lapply(knowns, function(x) {
+    cat(x)
+    cat("\n")
+    bulletGetMaxCMS(x, unknowns[j], crosscut = 112.5)
+  })
+  save(reslist, file=sprintf("data/unkn%d.RData", j))
+}
+
 
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
 qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")

### 2015-11-29 code/evaluation-hamby.R

@@ -43,6 +43,7 @@ CCFs$b2 <- factor(as.character(CCFs$b2))
 
 idx <- which(CCFs$ccf > 0.6 & CCFs$distr.dist < 1.25)
 idx <- c(21, 84)
+idx <- which(CCFs$b2 == "Ukn Bullet B-1")
 for ( i in idx) {
   load(CCFs$data[i])
   res <- reslist[[CCFs$resID[i]]]  
@@ -71,8 +72,10 @@ idx <- which(maxCMS == 7) # one is a match
 idx <- which(maxCMS == 6) # two are a match
 
 
+i <- 1
 
 for (i in idx) {
+i <- i+1
 load(datas[i])
 
 cmsdist <- sapply(reslist, function(x) x$maxCMS)

### 2015-11-29 imaging-paper.Rnw
matching land D-5

@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Maximal number of CMS between Bullet land and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
-<<cms, echo=FALSE, fig.width=5, fig.height=3, out.width='\\textwidth'>>=
+\caption{Maximal number of CMS between unknown Bullet D land 5 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
+<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
 load("data/unkn17.RData")
@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Overlaid lines of the land of the unknown bullet and the land with the top matching CMS.}
-<<top, echo=FALSE, fig.width=5, fig.height=3, out.width='\\textwidth'>>=
+\caption{Overlaid lines of the land of the unknown bullet D-5 and the land with the top matching CMS.}
+<<top, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
 res <- reslist[[which.max(cmsdist)]]  
@@
 
-\caption{Overlaid lines of the land of the unknown bullet and the land with the second top matching CMS.}
-<<top2, echo=FALSE, fig.width=5, fig.height=3, out.width='\\textwidth'>>=
+\caption{Overlaid lines of the land of the unknown bullet D-5 and the land with the second top matching CMS.}
+<<top2, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
 #res <- reslist[[which.max(cmsdist)]]  
@@
 \end{subfigure}
+\caption{\label{fig:hamby-perfect}Showcase scenario of when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
 \end{figure}

### 2015-11-29 imaging-paper.Rnw
attempt at crosscut discussion

@@
 Next, we fit a Loess regression to the data.
-Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlayed against the processed image of part 1 of bullet 1.
+Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
 The fit seems to do a reaosnable job of capturing the structure of the image.
@@
 On average, there are 6.6 CMS across the range of thresholds.
+In the autmatic matching routine, several thresholds are considered and the threshold resulting in the higest number of consecutive atching striae is chosen (and reported).
 \begin{figure}
@@
 \caption{\label{fig:thresholds}Varying the threshold levels results in different number of consecutive matching striae between the two bullets.
-\hh{XXX Our suggestion is, to aim for the threshold with the largest number of CMS.}}
+By default, the threshold with the highest CMS is chosen automatically.}
+\end{figure}
+
+\paragraph{Varying crosscuts}
+
+\begin{figure}
+<<crosscuts-vary, echo=FALSE, fig.width = 14, fig.height = 21, out.width = '\\linewidth', cache=TRUE, warnings = FALSE>>=
+paths <- file.path(datadir, dir(datadir))
+crosscuts <- seq(100, 450, by = 25)
+lof <- processBullets(paths[5], x = crosscuts)
+lof$bullet <- paste(lof$bullet, lof$x)
+qplot(data=lof, x=y, y=resid, colour=factor(x), facets=~x)
+
+reslist <- lapply(crosscuts[-1], function(cc) {
+  b2 <- subset(lof, x %in% c(cc, 100))
+  lofX <- bulletSmooth(b2)
+  lofX <- bulletAlign(lofX)  
+  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+
+  lines <- striation_identify(lofX, threshold = threshold)
+  maxCMS <- maxCMS(lines$match==TRUE)
+  list(maxCMS = maxCMS, threshold=threshold, lines=lines, bullets=lofX)
+})
+
+lop <- lapply(reslist, function(res) {
+ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
+  geom_hline(yintercept = res$threshold) +
+  geom_hline(yintercept = - res$threshold) +
+  scale_colour_brewer(palette="Set1") +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
+})
+
+grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
+             lop[[7]], lop[[8]], lop[[9]], lop[[10]], lop[[11]], lop[[12]], 
+             lop[[13]], lop[[14]],
+             ncol = 2)
+@
+\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts. Crosscut at $x = 100$ is compared to crosscuts every 25$\mu m$. \hh{XXX result?}}
 \end{figure}
@@
 
+\hh{XXX needs more work: how does crosscut impact the results?  For the same bullet: how far apart can crosscuts be to still result in a match?}
+
+\hh{XXX another idea for the rest of the paper: use some datamining methods to match the Hamby bullets }
+
 \section{Conclusion}

### 2015-11-30 imaging-paper.Rnw
description of results from varying crosscut levels

@@
 \begin{figure}
-<<crosscuts-vary, echo=FALSE, fig.width = 14, fig.height = 21, out.width = '\\linewidth', cache=TRUE, warnings = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 14, fig.height = 21, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
 lof$bullet <- paste(lof$bullet, lof$x)
-qplot(data=lof, x=y, y=resid, colour=factor(x), facets=~x)
 

### 2015-11-30 imaging-paper.Rnw
discussion of crosscuts

@@
 \paragraph{Varying crosscuts}
-
+The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up.
+For bullet land 1-5 from barrel 1 crosscuts at intervals of 50$\mu m$ between 150$\mu m$  and 450$\mu m$  are compared to a crosscut at 100$\mu m$.
+The initially almost perfect match (consolidated by a CMS of nineteen) between the  crosscuts deteriorates with increasing distance to a very questionable match at $x = 350\mu m$ and non-matches further apart.
+Only if the crosscut is within 200$\mu m$ do we get a good visual match.
+Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections.
 \begin{figure}
@@
 paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100, 450, by = 25)
+crosscuts <- seq(100, 450, by = 50)
 lof <- processBullets(paths[5], x = crosscuts)
@@
 grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
-             lop[[7]], lop[[8]], lop[[9]], lop[[10]], lop[[11]], lop[[12]], 
-             lop[[13]], lop[[14]],
-             ncol = 2)
+             lop[[7]],
+             ncol = 1)
 @
-\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts. Crosscut at $x = 100$ is compared to crosscuts every 25$\mu m$. \hh{XXX result?}}
+\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts.
+Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$.
+With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 19 to only 3 at $x = 400$ and $x = 450$.
+}
 \end{figure}

### 2015-11-30 imaging-paper.Rnw
crosscut sections for bullet 1 barrel 3.

@@
 Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections.
+\hh{XXX We're not doing that yet, but it might help - particular for bullets that have larger break-offs at the bottom.}
 \begin{figure}
@@
 Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$.
-With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 19 to only 3 at $x = 400$ and $x = 450$.
-}
+With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 19 to only 3 at $x = 400\mu m$ and $x = 450\mu m$.
+}
+\end{figure}
+
+\hh{XXX Does this give us a way to automatically check for the bottom of the bullet?
+-- We could try to take two cross sections - maybe 50 or 100 $\mu m$ apart.
+If they don't match, we need to go a bit higher with the lower cross cut.
+Attempt at this is shown in Figure~\ref{fig:crosscuts2}.}
+
+\begin{figure}
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+paths <- file.path(datadir, dir(datadir))
+crosscuts <- seq(100,225, by=25)
+lof <- processBullets(paths[37], x = crosscuts)
+lof$bullet <- paste(lof$bullet, lof$x)
+
+reslist <- lapply(1:length(crosscuts[-1]), function(i) {
+  b2 <- subset(lof, x %in% crosscuts[i:(i+1)])
+  lofX <- bulletSmooth(b2)
+  lofX <- bulletAlign(lofX)  
+  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+
+  lines <- striation_identify(lofX, threshold = threshold)
+  maxCMS <- maxCMS(lines$match==TRUE)
+  list(maxCMS = maxCMS, threshold=threshold, lines=lines, bullets=lofX)
+})
+
+lop <- lapply(reslist, function(res) {
+ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
+  geom_hline(yintercept = res$threshold) +
+  geom_hline(yintercept = - res$threshold) +
+  scale_colour_brewer(palette="Set1") +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
+})
+
+grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
+             lop[[4]], lop[[5]], 
+             ncol=1)
+@
+\caption{\label{fig:crosscuts2}Varying crosscuts for barrel 3, bullet 1.
+Initially, the match between crosscuts 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet.
+At a crosscut level of $175\mu m$ the bullet pattern stabilizes.
+For this land, matches should not be attempted to lower crosscuts.}
 \end{figure}

### 2015-11-30 imaging-paper.Rnw
bullet 311 needed for building paper

@@
 \begin{figure}
-<<crosscuts-vary, echo=FALSE, fig.width = 14, fig.height = 21, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
   geom_hline(yintercept = - res$threshold) +
-  scale_colour_brewer(palette="Set1") +
+  scale_colour_brewer("", palette="Set1") +
   theme(legend.position = c(1,1), legend.justification=c(1,1)) + 

### 2015-11-30 imaging-paper.Rnw
just fiddling with the legends

@@
   scale_colour_brewer("", palette="Set1") +
-  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
   ylim(c(-6,6)) +

### 2015-11-30 imaging-paper.Rnw
extending discussion on dealing with break-offs

@@
 
+\paragraph{Horizontal Alignment}
 Using (for now) a single crosscut across the surface, we extract the values and remove the class characteristics using a loess fit as before.
@@
 A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.
-\hh{This horizontal shift is based on the cross-correlation between the two surface crosscuts.
+This horizontal shift is based on the cross-correlation between the two surface crosscuts.
 Let $f(t)$ and $g(t)$ define the values of the crosscut at $t$, then the cross-correlation between $f$ and $g$ at lag $k$ is defined as
@@
 with suitable defined limits of the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).
-}
+
 \begin{figure}
@@
 
-\hh{Figure~\ref{fig:match} gives an overview of the rest of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; obviously this is a crucial parameter.
+Figure~\ref{fig:match} gives an overview of the rest of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; obviously this is a crucial parameter.
 For more discussion see Figure~\ref{fig:thresholds}).
@@
 For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
-XXX need more references XXX \cite{nichols:2003} XXX what does the paper say?
+\hh{XXX need more references XXX \cite{nichols:2003} XXX what does the paper say?
 AFTE is so frustrating!
@@
 
-\hh{XXX Does this give us a way to automatically check for the bottom of the bullet?
--- We could try to take two cross sections - maybe 50 or 100 $\mu m$ apart.
-If they don't match, we need to go a bit higher with the lower cross cut.
-Attempt at this is shown in Figure~\ref{fig:crosscuts2}.}
+\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
+By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, such as in a break-off, or in a stable section, where we can expect to find matches reasonably well.
+In the approach here, we choose a slightly higher crosscut until we find a section with a stable pattern.
+This process is shown  in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
 

### 2015-11-30 code/complete-hamby.R
adding crosscut values into the automatic matching

@@ -5,7 +5,7 @@ library(ggplot2)
 library(gridExtra)
 
 knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-knowns <- file.path(knowndatadir, dir(knowndatadir))
+knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 
 unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
@@ -14,13 +14,30 @@ unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 ###############
 # can we identify the barrels the unknown bullets came from?
 
+
+# match unknown land using crosscuts
+ccs <- read.csv("crosscuts.csv")
+for (j in 1:90) {
+  
+  reslist <- lapply(knowns, function(x) {
+    cat(x)
+    cat("\n")
+    
+    bulletGetMaxCMS(x, unknowns[j], 
+                    crosscut = ccs$cc[which(ccs$path==x)], 
+                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
+  })
+  save(reslist, file=sprintf("data/unkn%d.RData", j))
+}
+
+
 # match first unknown land
-for (j in 3:90) {
+for (j in 1:90) {
 
 reslist <- lapply(knowns, function(x) {
   cat(x)
   cat("\n")
-  bulletGetMaxCMS(x, unknowns[j])
+  bulletGetMaxCMS(x, unknowns[j], check=TRUE)
 })
 save(reslist, file=sprintf("data/unkn%d.RData", j))
 }
@@ -35,7 +52,7 @@ for (j in 1:90) {
 }
 
 
-for (j in 1) {
+for (j in 2:90) {
   reslist <- lapply(knowns, function(x) {
     cat(x)
     cat("\n")

### 2015-11-30 imaging-paper.Rnw
some restructuring, re-running the whole hamby evaluation again ... that'll take another 24h

@@
 \paragraph{Keywords:} Data visualization, Statistical graphics, Statistical computing.
+
+\tableofcontents
 
@@
   \centering
-\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}%
-<<circlefits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+<<circlefits, echo=FALSE, fig.width=7, fig.height=10.5, out.width='\\linewidth', warning=FALSE>>=
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
       geom="line", size=I(1), colour=I("grey30")) + 
-  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
+  facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
@@
 @
+\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}%
 \end{figure}
@@
 Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
-The fit seems to do a reaosnable job of capturing the structure of the image.
+The fit seems to do a reasonable job of capturing the structure of the image.
 Figure \ref{fig:loess_resid} shows the residuals from this fit.
@@
 
-<<loess_step12, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Steps 1 and 2 of the Loess fit algorithm.
+<<loess_step12, echo=FALSE, fig.width=7, fig.height=10.5, out.width='\\linewidth', warning=FALSE, fig.cap='Steps 1 and 2 of the Loess fit algorithm.
 The top panel is a plot of height values for part 1 of bullet 1 across different values of y, with x held constant at 243.75.
@@
 
-grid.arrange(p1, p2, nrow = 2)
+grid.arrange(p1, p2, nrow = 3)
 @
@@
 \section{Application}
-We've implemented a web interface to automate many of the steps outlined in the previous section.
+We have implemented a web interface to automate many of the steps outlined in the previous section.
 The web interface, developed using Shiny \cite{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
@@
 
-\paragraph{Horizontal Alignment}
+\subsection{Horizontal Alignment}
 Using (for now) a single crosscut across the surface, we extract the values and remove the class characteristics using a loess fit as before.
@@
 
-\paragraph{Varying threshold}
+\subsection{Varying threshold}
 
@@
 
-\paragraph{Varying crosscuts}
+\subsection{Varying crosscuts}
 The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up.
@@
   lofX <- bulletSmooth(b2)
-  lofX <- bulletAlign(lofX)  
+  bAlign = bulletAlign(lofX)
+  lofX <- bAlign$bullet  
   threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
@@
   maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, threshold=threshold, lines=lines, bullets=lofX)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, threshold=threshold, lines=lines, bullets=lofX)
 })
+
+ccfs <- sapply(reslist, function(res) res$ccf)
 
@@
   lofX <- bulletSmooth(b2)
-  lofX <- bulletAlign(lofX)  
+  bAlign <- bulletAlign(lofX)
+  lofX <- bAlign$bullet  
   threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
@@
   maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, threshold=threshold, lines=lines, bullets=lofX)
+  list(maxCMS = maxCMS, threshold=threshold, ccf = bAlign$ccf, lines=lines, bullets=lofX)
 })
+
+ccfs <- sapply(reslist, function(res) res$ccf)
 
@@
 At a crosscut level of $175\mu m$ the bullet pattern stabilizes.
-For this land, matches should not be attempted to lower crosscuts.}
+For this land, matches should not be attempted at lower crosscuts.}
 \end{figure}

### 2015-11-30 code/evaluation-hamby.R

@@ -1,5 +1,5 @@
-setwd("data")
-datas <- dir(pattern="RData")
+
+datas <- file.path("data", dir("data", pattern="RData"))
 
 maxCMS <- sapply(datas, function(x) {
   load(x)
@@ -41,6 +41,7 @@ CCFs$resID <- rep(1:120, length=nrow(CCFs))
 CCFs <- CCFs[order(as.character(CCFs$b2)),]
 CCFs$b2 <- factor(as.character(CCFs$b2))
 
+idx <- which(CCFs$cms >= 6)
 idx <- which(CCFs$ccf > 0.6 & CCFs$distr.dist < 1.25)
 idx <- c(21, 84)
 idx <- which(CCFs$b2 == "Ukn Bullet B-1")

### 2015-12-01 imaging-paper.Rnw
loess fit - pictures

@@
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
-      geom="line", size=I(1), colour=I("grey30")) + 
+      geom="line", size=I(1), colour=I("grey70")) + 
   facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group = x), colour="black", size=1, alpha=0.2, 
-             data = filter(db1, part==1, x == 100)[,c("y", "resid", "x")]) + ggtitle("Part 1 of Bullet 1 in black")
-@
-\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlayed.\label{fig:circlefits}}%
+  geom_line(aes(y, resid, group = x), colour="black", size=1, 
+             data = filter(db1, part==1, x == 100)[,c("y", "resid", "x")]) +
+  ylab("Residuals from circular fit") +
+  ggtitle("Part 1 of Bullet 1 in black")
+@
+\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlaid.\label{fig:circlefits}}%
 \end{figure}
@@
 \begin{enumerate}
-    \item Fix an x coordinate value. (Top panel of Figure \ref{fig:loess_step12})
-    \item Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$. (Bottom panel of Figure \ref{fig:loess_step12})
-    \item For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minimums. (Top panel of Figure \ref{fig:loess_step34})
-    \item Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove) (Bottom panel of Figure \ref{fig:loess_step34})
-    \item Remove all y values below the minimum and above the maximum (Figure \ref{fig:loess_step5})
+    \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step12}).
+    \item Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$ (Figure~\ref{fig:loess_step12}).
+    \item For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step34}).
+    \item Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove) (Figure~\ref{fig:loess_step34}).
+    \item Remove all $y$ values left and right of the identified minima in vertical direction (Figure \ref{fig:loess_step34}).
 \end{enumerate}
@@
 Certainly, tuning the parameters for this regression, such as the Loess span, could prove to do an even better job at making the markings more clear.
-
-<<loess_step12, echo=FALSE, fig.width=7, fig.height=10.5, out.width='\\linewidth', warning=FALSE, fig.cap='Steps 1 and 2 of the Loess fit algorithm.
-The top panel is a plot of height values for part 1 of bullet 1 across different values of y, with x held constant at 243.75.
-The bottom panel is the same plot, but with the first and last 100 y values trimmed.', fig.pos='H', message=FALSE>>=
+\hh{XXX did you follow up on that?
+What is the effect of using a different span?}
+
+\begin{figure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_step12}Steps 1 and 2 of the Loess fit algorithm.
+For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.
+The first and last 100 values of y are trimmed (shown in grey).}
+<<loess_step12, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
 br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
@@
 
-p1 <- qplot(y, value, data = br111_bullet) + theme_bw()
-p2 <- qplot(y, value, data = tail(head(br111_bullet, n = -100), n = -100)) + theme_bw()
-
-grid.arrange(p1, p2, nrow = 3)
-@
-
-<<loess_step34, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Steps 3 and 4 of the Loess fitting algorithm.
-The top panel is a plot of the smoothed height values, and the bottom panel is a plot of the detected minimums.', fig.pos='H', message=FALSE>>=
+p1 <- qplot(y, value, data = br111_bullet, size=I(1), colour=I("grey60"), alpha=I(.25)) + theme_bw()
+p1 + geom_point(aes(y, value), data = tail(head(br111_bullet, n = -100), n = -100))
+#p2 <- qplot(y, value, data = tail(head(br111_bullet, n = -100), n = -100)) + theme_bw()
+
+#grid.arrange(p1, p2, nrow = 3)
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_step34}Steps 3 - 5 of the Loess fitting algorithm.
+After smoothing the $y$ values left and right minima are detected (marked by vertical lines).
+Values outside these boundaries are removed (shown in grey)}
+<<loess_step34, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
 subdata <- tail(head(br111_bullet, n = -100), n = -100)
 subdata$smoothed <- c(rep(NA, 10), rollapply(subdata$value, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
-p3 <- qplot(y, smoothed, data = tail(head(subdata, n = -10), n = -10)) + theme_bw()
+p3 <- qplot(y, smoothed, data = tail(head(subdata, n = -10), n = -10)) + theme_bw() + xlim(c(0,2500))
 
@@
 
-p4 <- qplot(data=curdata, y, value) +
+p4 <- qplot(data=curdata, y, value, size=I(1), colour=I("grey60"), alpha=I(.25)) +
         theme_bw() +
         geom_vline(xintercept = curdata$y[min.left], colour = "red") +
-        geom_vline(xintercept = curdata$y[min.right + (length(curdata$y) %/% 2)], colour = "blue")
-
-grid.arrange(p3, p4, nrow = 2)
-@
-
-<<loess_step5, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Step 5 of the Loess fitting algorithm.
-The plot shows the result after trimming values below the first minimum, and above the second minimum.', fig.pos='H', message=FALSE>>=
+        geom_vline(xintercept = curdata$y[min.right + (length(curdata$y) %/% 2)], colour = "blue") + xlim(c(0,2500))
+
+#grid.arrange(p3, p4, nrow = 2)
 final.dat <- subset(curdata, y > curdata$y[min.left + 5] & y < curdata$y[min.right + (length(curdata$y) %/% 2) - 5])
 
-qplot(data=final.dat, y, value) +
-    theme_bw()
-@
-
-<<loess_fit, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit for part 1 of bullet 1', fig.pos='H', message=FALSE>>=
+p4 + geom_point(data=final.dat, aes(y, value))
+@
+\end{subfigure}
+%' \begin{subfigure}[t]{\linewidth}
+%' \caption{\label{fig:loess_step5}Step 5 of the Loess fitting algorithm. The plot shows the result after trimming values below the first minimum, and above the second minimum.}
+%' <<loess_step5, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+%' final.dat <- subset(curdata, y > curdata$y[min.left + 5] & y < curdata$y[min.right + (length(curdata$y) %/% 2) - 5])
+%' 
+%' qplot(data=final.dat, y, value) +
+%'     theme_bw() + xlim(c(0,2500))
+%' @
+%' \end{subfigure}
+\caption{Overview of all five steps of the loess algorithm to identify and remove big bullet strucure.}
+\end{figure}
+
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit for part 1 of bullet 1', fig.pos='hbtp', message=FALSE>>=
 
@@
 
-<<loess_resid, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit residuals for part 1 of bullet 1', fig.pos='H'>>=
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit residuals for part 1 of bullet 1', fig.pos='hbtp'>>=
 my.loess <- fit_loess(br111_bullet, br111.groove)
@@
 qplot(y, resid, data=melt.df, #colour=factor(x), 
-      geom="line", size=I(1), colour=I("grey30")) + 
+      geom="line", size=I(1), colour=I("grey70")) + 
   facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=1, alpha=0.2) + ggtitle("Part 1 of Bullet 1 in black")
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=1) + ggtitle("Part 1 of Bullet 1 in black")
 @

### 2015-12-01 code/evaluation-hamby.R

@@ -19,7 +19,7 @@ CCFs <- plyr::ldply(datas, function(x) {
     
     subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
     subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
- #     browser()
+#      browser()
     ccf <- ccf(subLOFx1$val, subLOFx2$val, plot = FALSE, lag.max=200, na.action = na.omit)
     lag <- ccf$lag[which.max(ccf$acf)]
     incr <- min(diff(sort(unique(subLOFx1$y))))
@@ -30,7 +30,11 @@ CCFs <- plyr::ldply(datas, function(x) {
     idx2 <- which(subLOFx2$y %in% ys)
     distr.dist <- mean((subLOFx1$val[idx1] - subLOFx2$val[idx2])^2, na.rm=T)
     
-    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), distr.dist=distr.dist, b1=b12[1], b2=b12[2])
+    # feature extraction
+    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), distr.dist=distr.dist, 
+               b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
+               num.matches = sum(res$lines$match), 
+               num.mismatches = sum(!res$lines$match))
   })
   ccf$cms <- cmsdist
   ccf$data <- x
@@ -41,13 +45,23 @@ CCFs$resID <- rep(1:120, length=nrow(CCFs))
 CCFs <- CCFs[order(as.character(CCFs$b2)),]
 CCFs$b2 <- factor(as.character(CCFs$b2))
 
-qplot(data=CCFs, x=ccf, y=cms)
+CCFs$match <- NA
+idx <- which(CCFs$cms >= 9) # all are matches, visually confirmed
+CCFs$match[idx] <- TRUE
+write.csv(CCFs, file="bullet-stats.csv", row.names=FALSE)
 
 
-idx <- which(CCFs$cms >= 6)
+qplot(data=CCFs, x=ccf, y=cms, colour=match)
+qplot(data=CCFs, x=ccf, y=distr.dist)
+
+
+idx <- which(CCFs$cms >= 9)
+
+
+
 idx <- which(CCFs$ccf > 0.6 & CCFs$distr.dist < 1.25)
 idx <- c(21, 84)
[…]

### 2015-12-01 imaging-paper.Rnw
bullet-stats.csv updated

@@
   \centering
-<<circlefits, echo=FALSE, fig.width=7, fig.height=10.5, out.width='\\linewidth', warning=FALSE>>=
+<<circlefits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
@@
 
-<<loess_fit, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit for part 1 of bullet 1', fig.pos='hbtp', message=FALSE>>=
+\begin{figure}[hbtp]
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_fit} Loess fit for land 1 of bullet 1.}
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
 
@@
 @
-
-<<loess_resid, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE, fig.cap='Loess fit residuals for part 1 of bullet 1', fig.pos='hbtp'>>=
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_resid} Loess fit residuals for land 1 of bullet 1.}
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
 my.loess <- fit_loess(br111_bullet, br111.groove)
@@
 @
+\end{subfigure}
+\caption{Fit and residuals of a loess fit to land 1 of bullet 1 (Barrel 1).}
+\end{figure}
 
 We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from part 1 of bullet 1.
-This is available in Figure \ref{fig:loessfits}.
-
-<<loessfits, echo=FALSE, fig.width=10, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+This is shown in Figure~\ref{fig:loessfits}.
+
+\begin{figure}[hbtp]
+<<loessfits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
 br121.groove <- get_grooves(br121_bullet)
@@
       geom="line", size=I(1), colour=I("grey70")) + 
-  facet_wrap(~part) + scale_colour_brewer(palette="Paired") +
+  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=1) + ggtitle("Part 1 of Bullet 1 in black")
-@
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=1) + ggtitle("Land 1 of Bullet 1 in black")
+@    
+\caption{Residuals of a loess fit to bullets 1 and 2 from barrel 1. Land 1 of bullet 1 is shown in black. }
+\end{figure}
 

### 2015-12-01 code/complete-hamby.R
update on matches

@@ -30,6 +30,73 @@ for (j in 1:90) {
   save(reslist, file=sprintf("data/unkn%d.RData", j))
 }
 
+# match some known lands
+br4 <- knowns[grep("Br4", knowns)]
+reslist <- list()
+for (j in 1) {
+  res <- lapply(br4[7:12], function(x) {
+    cat(x)
+    cat("\n")
+    
+    bulletGetMaxCMS(x, br4[j], 
+                    crosscut = ccs$cc[which(ccs$path==x)], 
+                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
+  })
+  reslist <- c(reslist, res)
+}
+save(reslist, file=sprintf("data/br4-%d.RData", j))
+
+br <- knowns[grep("Br8", knowns)]
+reslist <- list()
+for (j in 2) {
+  res <- lapply(br[7:12], function(x) {
+    cat(x)
+    cat("\n")
+    
+    bulletGetMaxCMS(x, br[j], 
+                    crosscut = ccs$cc[which(ccs$path==x)], 
+                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
+  })
+  reslist <- c(reslist, res)
+}
+save(reslist, file=sprintf("data/br8-%d.RData", j))
+
+br <- knowns[grep("Br9", knowns)]
+reslist <- list()
+for (j in 1:6) {
+  res <- lapply(br[7:12], function(x) {
+    cat(x)
+    cat("\n")
+    
+    bulletGetMaxCMS(x, br[j], 
+                    crosscut = ccs$cc[which(ccs$path==x)], 
+                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
+  })
+  reslist <- c(reslist, res)
+}
+cmsdist <- sapply(reslist, function(x) x$maxCMS)
+qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
+
+save(reslist, file=sprintf("data/br9.RData", j))
+
+
+br <- knowns[grep("Br10", knowns)]
+reslist <- list()
+for (j in 2) {
+  res <- lapply(br[7:12], function(x) {
+    cat(x)
+    cat("\n")
+    
+    bulletGetMaxCMS(x, br[j], 
+                    crosscut = ccs$cc[which(ccs$path==x)], 
[…]

### 2015-12-01 code/evaluation-hamby.R

@@ -50,6 +50,7 @@ idx <- which(CCFs$cms >= 9) # all are matches, visually confirmed
 CCFs$match[idx] <- TRUE
 write.csv(CCFs, file="bullet-stats.csv", row.names=FALSE)
 
+idx <- which(CCFs$cms == 8) # 
 
 qplot(data=CCFs, x=ccf, y=cms, colour=match)
 qplot(data=CCFs, x=ccf, y=distr.dist)

### 2015-12-01 code/complete-hamby.R
match for barrel 9

@@ -63,7 +63,7 @@ save(reslist, file=sprintf("data/br8-%d.RData", j))
 
 br <- knowns[grep("Br9", knowns)]
 reslist <- list()
-for (j in 1:6) {
+for (j in 1) {
   res <- lapply(br[7:12], function(x) {
     cat(x)
     cat("\n")

### 2015-12-01 imaging-paper.Rnw
Redo the loess groove detection section

@@
 Fortunately, the location and appearance of the grooves in these 2d plots is quite consistent.
-Because a bullet is scanned groove to groove, the groove should appear within approximately scan points from each end.
-That is, on y coordinate system ranging from 0 to 1000 as these images, are, the grooves will appear between $y \in \{0, 100\}$ and $y \in \{900, 1000\}$.
-Hence, the algorithm for removing the grooves can be described as follows:
+By noting that height values beginning at $y = 0$ reach a maximum around the peak of the groove, we can then follow the descent of the height values to valley of the groove.
+This marks the point at which we should trim the image.
+We repeat a similar procedure for the right-hand groove by starting at the maximum y value and working backwards.
+The procedure can be described as follows:
 
@@
     \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step12}).
-    \item Remove values from $y \in \{0, 100\}$ and $y \in \{900, 1000\}$ (Figure~\ref{fig:loess_step12}).
-    \item For each y value from 100 to 900, compute a rolling average to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step34}).
-    \item Determine the minimum in the first half (corresponding to the left groove) and the minimum in the right half (corresponding to the right groove) (Figure~\ref{fig:loess_step34}).
+    \item For each y value, compute a rolling average to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step12}).
+    \item Determine the peak of the left groove by finding the first value with a negative discretized derivative.
+Determine the valley of the left groove by finding the first value with a positive discretized derivative after the value of the peak.
+(Figure~\ref{fig:loess_step34}).
+    \item Determine the peak of the right groove by finding the last value with a positive discretized derivative.
+Determine the valley of the right groove by finding the last value with a negative discretized derivative before the value of the peak.
+(Figure~\ref{fig:loess_step34}).
     \item Remove all $y$ values left and right of the identified minima in vertical direction (Figure \ref{fig:loess_step34}).
@@
 br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
-
 br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
@@
 p1 + geom_point(aes(y, value), data = tail(head(br111_bullet, n = -100), n = -100))
-#p2 <- qplot(y, value, data = tail(head(br111_bullet, n = -100), n = -100)) + theme_bw()
-
-#grid.arrange(p1, p2, nrow = 3)
+
+subdata <- br111_bullet
+subdata$smoothed <- c(rep(NA, 10), rollapply(subdata$value, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
+subdata$smoothed2 <- c(rep(NA, 10), rollapply(subdata$smoothed, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
+p2 <- qplot(y, smoothed2, data = subdata) + theme_bw() + xlim(c(0,2500))
+
+p2
 @
@@
 \caption{\label{fig:loess_step34}Steps 3 - 5 of the Loess fitting algorithm.
-After smoothing the $y$ values left and right minima are detected (marked by vertical lines).
+After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
 Values outside these boundaries are removed (shown in grey)}
 <<loess_step34, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-subdata <- tail(head(br111_bullet, n = -100), n = -100)
-subdata$smoothed <- c(rep(NA, 10), rollapply(subdata$value, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
-p3 <- qplot(y, smoothed, data = tail(head(subdata, n = -10), n = -10)) + theme_bw() + xlim(c(0,2500))
-
-curdata <- tail(head(subdata, n = -10), n = -10)
-min.left <- which.min(curdata$value[1:(length(curdata$value) %/% 2)])
-min.right <- which.min(curdata$value[(length(curdata$value) %/% 2):length(curdata$value)])
-
-p4 <- qplot(data=curdata, y, value, size=I(1), colour=I("grey60"), alpha=I(.25)) +
-        theme_bw() +
-        geom_vline(xintercept = curdata$y[min.left], colour = "red") +
-        geom_vline(xintercept = curdata$y[min.right + (length(curdata$y) %/% 2)], colour = "blue") + xlim(c(0,2500))
-
-#grid.arrange(p3, p4, nrow = 2)
-final.dat <- subset(curdata, y > curdata$y[min.left + 5] & y < curdata$y[min.right + (length(curdata$y) %/% 2) - 5])
-
-p4 + geom_point(data=final.dat, aes(y, value))
+peak_ind <- head(which(diff(subdata$smoothed2) < 0), n = 1)
+    groove_ind <- head(which(diff(tail(subdata$smoothed2, n = -(peak_ind + 10))) > 0), n = 1) + peak_ind + 10
+
+    peak_ind2 <- tail(which(diff(subdata$smoothed2) > 0), n = 1)
+    groove_ind2 <- tail(which(diff(head(subdata$smoothed2, n = -(length(subdata$smoothed2) - peak_ind2 + 10))) < 0), n = 1)
+
+    p4 <- qplot(subdata$y[subdata$y < subdata$y[groove_ind]], subdata$smoothed2[subdata$y < subdata$y[groove_ind]], colour=I("grey60"), alpha=I(.25)) +
+        ylim(c(min(subdata$value[!is.nan(subdata$value)]) - 25, max(subdata$value[!is.nan(subdata$value)]) + 25)) +
+        xlim(c(min(subdata$y) - 25, max(subdata$y) + 25)) +
+        geom_point(aes(x, y), data = data.frame(x = subdata$y[subdata$y > subdata$y[groove_ind2]], y = subdata$smoothed2[subdata$y > subdata$y[groove_ind2]]), colour=I("grey60"), alpha=I(.25)) +
+        geom_point(inherit.aes = FALSE, aes(x, y), data = data.frame(x = subdata$y[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]], y = subdata$smoothed2[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]])) +
+        theme_bw() +
+        geom_vline(xintercept = subdata$y[peak_ind], colour = "red") +
+        geom_vline(xintercept = subdata$y[groove_ind], colour = "blue") +
+        geom_vline(xintercept = subdata$y[peak_ind2], colour = "red") +
+        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue")
+
+    p4
 @

### 2015-12-01 imaging-paper.Rnw
Small correction in the loess section

@@
 For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.
-The first and last 100 values of y are trimmed (shown in grey).}
+The height values are smoothed to remove noise in the deviations.}
 <<loess_step12, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 
-p1 <- qplot(y, value, data = br111_bullet, size=I(1), colour=I("grey60"), alpha=I(.25)) + theme_bw()
-p1 + geom_point(aes(y, value), data = tail(head(br111_bullet, n = -100), n = -100))
+p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw()
+p1
 

### 2015-12-02 imaging-paper.Rnw
bullet-stats and crosscuts are dependent on previous version of xpr3plus

@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_resid} Loess fit residuals for land 1 of bullet 1.}
+\caption{\label{fig:loess_resid} Residuals of Loess fit for land 1 of bullet 1.}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
@@
       geom="line", size=I(1), colour=I("grey70")) + 
-  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
+  facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 

### 2015-12-02 code/checkBullets.R
update to bullet-stats.csv

@@ -6,6 +6,14 @@ library(ggplot2)
 knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
 knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 
+crosscuts <- sapply(unknowns, function(x) {
+  cat(x)
+  crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=12.5))
+  cat(crosscut)
+  cat("\n")
+  crosscut
+})
+
 
 dframe <- data.frame(
   path = knowns,

### 2015-12-02 code/evaluation-hamby.R

@@ -1,11 +1,13 @@
 
 datas <- file.path("data", dir("data", pattern="RData"))
+datas <- datas[grep("data/u.*", datas)]
 
 maxCMS <- sapply(datas, function(x) {
   load(x)
   cmsdist <- sapply(reslist, function(x) x$maxCMS)
   max(cmsdist)
 })
+
 
 CCFs <- plyr::ldply(datas, function(x) {
   load(x)
@@ -19,7 +21,7 @@ CCFs <- plyr::ldply(datas, function(x) {
     
     subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
     subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-#      browser()
+ #     browser()
     ccf <- ccf(subLOFx1$val, subLOFx2$val, plot = FALSE, lag.max=200, na.action = na.omit)
     lag <- ccf$lag[which.max(ccf$acf)]
     incr <- min(diff(sort(unique(subLOFx1$y))))
@@ -34,7 +36,8 @@ CCFs <- plyr::ldply(datas, function(x) {
     data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), distr.dist=distr.dist, 
                b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
                num.matches = sum(res$lines$match), 
-               num.mismatches = sum(!res$lines$match))
+               num.mismatches = sum(!res$lines$match), 
+               non_cms = x3prplus::maxCMS(res$lines$match))
   })
   ccf$cms <- cmsdist
   ccf$data <- x
@@ -50,21 +53,9 @@ idx <- which(CCFs$cms >= 9) # all are matches, visually confirmed
 CCFs$match[idx] <- TRUE
 write.csv(CCFs, file="bullet-stats.csv", row.names=FALSE)
 
-idx <- which(CCFs$cms == 8) # 
 
-qplot(data=CCFs, x=ccf, y=cms, colour=match)
-qplot(data=CCFs, x=ccf, y=distr.dist, colour=match)
-
-
-idx <- which(CCFs$cms < 8 & CCFs$ccf > .85) # nothing
-idx <- which(CCFs$cms < 8 & CCFs$ccf > 0.5 & CCFs$distr.dist < 1.25)
-
-
-
-idx <- which(CCFs$ccf > 0.6 & CCFs$distr.dist < 1.25)
-idx <- c(21, 84)
[…]

### 2015-12-02 imaging-paper.Rnw
update to bullet-stats.csv
start on evaluation of the Hamby study

@@
 \end{subfigure}
-\caption{\label{fig:hamby-perfect}Showcase scenario of when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
+\caption{\label{fig:hamby-perfect}Showcase scenario  when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
+\end{figure}
+
+\section{Evaluation}
+\hh{XXX a lot more about the Hamby study.
+Fifteen unknown bullets (i.e.
+90 lands) compared to two times ten known bullets (i.e.
+120 lands) makes $90 \times 120 = 10,800$ land to land comparisons, out of which 180 are actual matches.
+It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches.
+}
+Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
+All 28 pairs of lands with at least nine CMS in common are matches.
+For smaller number of CMS, the percentage of false positives quickly increases.
+\begin{figure}
+\begin{subfigure}[t]{\linewidth}
+\caption{Distribution of maximal CMS.}
+<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
+bstats <- read.csv("data/bullet-stats.csv")
+ggplot(data=bstats) + 
+  geom_bar(aes(x=factor(cms))) + theme_bw() +
+  theme(legend.position="bottom") + 
+  xlab("maximum CMS")
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{All bullet pairs with a maximum of at least nine CMS are matches.}
+<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
+bstats <- read.csv("data/bullet-stats.csv")
+ggplot(data=bstats) + 
+  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
+  theme(legend.position="bottom") + 
+  xlab("maximum CMS")
+@
+\end{subfigure}
+\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
 \end{figure}

### 2015-12-02 code/evaluation-hamby.R

@@ -89,3 +89,4 @@ CCFs <- merge(CCFs, mm[,c("id","value")], by.x="b1", by.y="value")
 CCFs <- merge(CCFs, mm[,c("id","value")], by.x="b2", by.y="value")
 CCFs$match <- CCFs$id.x == CCFs$id.y
 
+write.csv(CCFs, file="data/bullet-stats.csv", row.names=FALSE)

### 2015-12-02 imaging-paper.Rnw
fixed non CMS runs

@@
 
+\begin{figure}
+<<distr.dist, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+ggplot(data=bstats) + 
+    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match)) +
+    theme(legend.position="bottom") + 
+    xlab("maximum CMS")
+@
+\caption{If the distance between the crosscuts is small, a smaller number of CMS might already be an indication of a match.}
+\end{figure}
+
 \hh{XXX what would be good is to get the line identifications and project them back onto the surface. }

### 2015-12-02 code/evaluation-hamby.R

@@ -37,7 +37,7 @@ CCFs <- plyr::ldply(datas, function(x) {
                b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
                num.matches = sum(res$lines$match), 
                num.mismatches = sum(!res$lines$match), 
-               non_cms = x3prplus::maxCMS(res$lines$match))
+               non_cms = x3prplus::maxCMS(!res$lines$match))
   })
   ccf$cms <- cmsdist
   ccf$data <- x

### 2015-12-02 code/checkBullets.R
adding the old results for comparison

@@ -6,7 +6,7 @@ library(ggplot2)
 knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
 knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 
-crosscuts <- sapply(unknowns, function(x) {
+crosscuts1 <- sapply(knowns, function(x) {
   cat(x)
   crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=12.5))
   cat(crosscut)
@@ -15,23 +15,22 @@ crosscuts <- sapply(unknowns, function(x) {
 })
 
 
-dframe <- data.frame(
-  path = knowns,
-  cc = crosscuts
-)
-
 
 unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
-crosscuts <- sapply(unknowns, function(x) {
+crosscuts2 <- sapply(unknowns, function(x) {
+  cat(x)
   crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=12.5))
+  cat(crosscut)
+  cat("\n")
+  crosscut
 })
 
-dframe2 <- data.frame(
-  path = unknowns,
-  cc = crosscuts
+dframe <- data.frame(
+  path = c(knowns, unknowns),
+  cc = c(crosscuts1, crosscuts2)
 )
 
 
-write.csv(rbind(dframe, dframe2), file="crosscuts.csv", row.names=FALSE)
+write.csv(dframe, file="crosscuts.csv", row.names=FALSE)

### 2015-12-02 imaging-paper.Rnw
adding the old results for comparison
'stable' crosscuts

@@
 Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections.
-\hh{XXX We're not doing that yet, but it might help - particular for bullets that have larger break-offs at the bottom.}
 \begin{figure}
@@
 
-\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
-By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, such as in a break-off, or in a stable section, where we can expect to find matches reasonably well.
-In the approach here, we choose a slightly higher crosscut until we find a section with a stable pattern.
+%\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
+By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, indicative of a break-off, or in a stable section, where we can expect to find matches reasonably well.
+In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
 This process is shown  in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
-
+\hh{`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. }
 \begin{figure}
@@
 
-\hh{We can use the Hamby study to run all unkowns against the known bullets. When things go perfectly, they look like in Figure~\ref{fig:hamby-perfect}.}
+\hh{XXX what would be good is to get the line identifications and project them back onto the surface. }
+
+\hh{XXX needs more work: how does crosscut impact the results?  For the same bullet: how far apart can crosscuts be to still result in a match?}
+
+\section{Evaluation}
+
+\hh{We can use the Hamby study to run all unkowns against the known bullets. When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}.}
 
@@
 \end{figure}
-
-\section{Evaluation}
 \hh{XXX a lot more about the Hamby study.
@@
 
-\hh{XXX what would be good is to get the line identifications and project them back onto the surface. }
-
-\hh{XXX needs more work: how does crosscut impact the results?  For the same bullet: how far apart can crosscuts be to still result in a match?}
-
 \hh{XXX another idea for the rest of the paper: use some datamining methods to match the Hamby bullets }
@@
 
+\hh{Things we discussed}
+\begin{enumerate}
+\item adding more non-matches into the mix is going to make things more realistic. 
+\end{enumerate}
 \bibliographystyle{abbrv}

### 2015-12-02 code/test_getgrooves.R
Some small tweaks to the testing for get grooves

@@ -3,9 +3,15 @@ library(x3prplus)
 library(ggplot2)
 
 for (file in dir("app/images/Hamby252_3DX3P1of2")) {
-    x <- get_grooves(get_bullet(file.path("app/images/Hamby252_3DX3P1of2", file), x = 100), smoothfactor = 41)
+    mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", file), x = 150)
+    x <- get_grooves(mybullet)
     cat(file, "\n")
     print(x$plot)
+    my.loess <- fit_loess(mybullet, x)
+    print(my.loess$resid)
     value <- readline("Press Enter to Continue, or q to Quit")
     if (value == "q") break;
 }
+
+mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", "Br1 Bullet 1-5.x3p"), x = 100)
+x <- get_grooves(mybullet)

### 2015-12-03 imaging-paper.Rnw
Test the get peaks function

@@
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
-my.loess <- fit_loess(br111_bullet, br111.groove)
-
 my.loess$resid

### 2015-12-03 code/test_getpeaks.R

@@ -6,12 +6,9 @@ for (file in dir("app/images/Hamby252_3DX3P1of2")) {
     mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", file), x = 150)
     x <- get_grooves(mybullet)
     cat(file, "\n")
-    print(x$plot)
-    #my.loess <- fit_loess(mybullet, x)
-    #print(my.loess$resid)
+    my.loess <- fit_loess(mybullet, x)
+    result <- get_peaks(my.loess$data)
+    print(result$plot)
     value <- readline("Press Enter to Continue, or q to Quit")
     if (value == "q") break;
 }
-
-mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", "Br1 Bullet 1-5.x3p"), x = 100)
-x <- get_grooves(mybullet)

### 2015-12-03 code/test_getgrooves.R
Small change to testgrooves

@@ -3,7 +3,7 @@ library(x3prplus)
 library(ggplot2)
 
 for (file in dir("app/images/Hamby252_3DX3P1of2")) {
-    mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", file), x = 150)
+    mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", file), x = 100)
     x <- get_grooves(mybullet)
     cat(file, "\n")
     print(x$plot)
@@ -12,6 +12,3 @@ for (file in dir("app/images/Hamby252_3DX3P1of2")) {
     value <- readline("Press Enter to Continue, or q to Quit")
     if (value == "q") break;
 }
-
-mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", "Br1 Bullet 1-5.x3p"), x = 100)
-x <- get_grooves(mybullet)

### 2015-12-03 imaging-paper.Rnw
start of discussing the automated matching algorithm

@@
 
+\hh{The automated matching routine is then as follows:
+Identify a stable region for each bullet and pick the lowest crosscut level in this region (assumption here: lower crosscut levels show better expression of individual marks XXX need to still check)
+
+All of the other steps are done on pairs of bullets:
+\begin{enumerate}
+\item Smooth the two crosscuts by a loess with a very small span. 
+\item Use cross-correlation to find the best alignment along the crosscuts: Shift one of the crosscuts by the lag indicated by the cross-correlation function.
+\item Based on a starting threshold level of $\theta$, identify the maximal number of CMS (consecutive matching striae) between the two crosscuts.
+Vary $\theta$, such that the maximum number of CMS is reached.
+\end{enumerate}
+}
 \hh{XXX next step: how does this matching work in known non-matches? distribution of CMS under a null of no match? }

### 2015-12-03 code/evaluation-hamby.R

@@ -90,3 +90,12 @@ CCFs <- merge(CCFs, mm[,c("id","value")], by.x="b2", by.y="value")
 CCFs$match <- CCFs$id.x == CCFs$id.y
 
 write.csv(CCFs, file="data/bullet-stats.csv", row.names=FALSE)
+
+# diagnostics
+
+library(ggplot2)
+qplot(factor(cms), data=CCFs)
+ggplot(data=CCFs) + geom_bar(aes(x=factor(cms), fill=match), position="fill")
+ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match))
+ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=ccf, colour=match)) + facet_wrap(~match)
+ggplot(data=CCFs) + geom_bar(aes(x=factor(non_cms), fill=match), position="fill")

### 2015-12-03 imaging-paper.Rnw
change to line identification

@@
 \item Use cross-correlation to find the best alignment along the crosscuts: Shift one of the crosscuts by the lag indicated by the cross-correlation function.
-\item Based on a starting threshold level of $\theta$, identify the maximal number of CMS (consecutive matching striae) between the two crosscuts.
-Vary $\theta$, such that the maximum number of CMS is reached.
+\item Based on a starting threshold level of $\theta$, identify areas of the crosscuts whose absolute value are above the thresholds.
+These areas are regions of potential matches between the crosscuts.
+If all of the values of the crosscuts stay on the same side, the area is called a matching line.
+If only one crosscut exceeds the threshold, or if the two crosscuts move into opposite directions over the threshold, the area is a non-match.
+\item Identify the maximal number of CMS (consecutive matching striae) between the two crosscuts by counting consecutive matching lines.
+Vary $\theta$ in (3), such that a maximum number of CMS is reached.
 \end{enumerate}
@@
   theme_bw() + 
-  geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_rect(aes(xmin=xmin, xmax=xmax), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
   geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
-  geom_hline(yintercept = res$threshold) +
-  geom_hline(yintercept = - res$threshold) +
+#  geom_hline(yintercept = res$threshold) +
+#  geom_hline(yintercept = - res$threshold) +
   scale_colour_brewer(palette="Set1") +

### 2015-12-03 code/checkBullets.R

@@ -34,3 +34,22 @@ dframe <- data.frame(
 
 
 write.csv(dframe, file="crosscuts.csv", row.names=FALSE)
+ccs <- read.csv("data/crosscuts.csv")
+
+knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
+knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
+
+bullets <- lapply(c(knowns, unknowns), function(path) {
+  lof <- processBullets(paths = path, x = ccs$cc[which(ccs$path==path)], check=FALSE)
+  peaks <- get_peaks(lof)
+  peaks$extrema <- c(peaks$peaks, peaks$valleys)
+  peaks$type <- c(rep(1, length(peaks$peaks)), rep(-1, length(peaks$valleys)))
+  idx <- order(peaks$extrema)
+  peaks$extrema <- peaks$extrema[idx]
+  peaks$type <- peaks$type[idx]
+  diffs <- diff(peaks$extrema)
+  lines <- data.frame(xmin = peaks$extrema-c(diffs[1],diffs)/3,
+                      xmax=peaks$extrema+c(diffs,diffs[length(diffs)])/3, 
+                      type=peaks$type)
+  list(path=path, bullet=lof, peaks=peaks, lines=lines)
+})

### 2015-12-03 code/complete-hamby.R

@@ -17,12 +17,34 @@ unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
 # match unknown land using crosscuts
 ccs <- read.csv("data/crosscuts.csv")
+load("data/bullets.RData")
+bPaths <- sapply(bullets, function(x) x$path)
+
+
 for (j in 1:90) {
   
   reslist <- lapply(knowns, function(x) {
     cat(x)
     cat("\n")
+#    browser()
+    b1 <- bullets[[which(bPaths == x)]]
+    b2 <- bullets[[which(bPaths == unknowns[j])]]
     
+    bulletGetMaxCMSXXX(b1, b2)
+  })
+  save(reslist, file=sprintf("data/unkn%d.RData", j))
+}
+
+
+
+# match unknown land using crosscuts
+ccs <- read.csv("data/crosscuts.csv")
+for (j in 1:90) {
+  
+  reslist <- lapply(knowns, function(x) {
+    cat(x)
+    cat("\n")
+    browser()
     bulletGetMaxCMS(x, unknowns[j], 
                     crosscut = ccs$cc[which(ccs$path==x)], 
                     crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)

### 2015-12-03 code/evaluation-hamby.R

@@ -53,6 +53,7 @@ idx <- which(CCFs$cms >= 9) # all are matches, visually confirmed
 CCFs$match[idx] <- TRUE
 write.csv(CCFs, file="bullet-stats.csv", row.names=FALSE)
 
+idx <- which(CCFs$match & CCFs$cms <= 6)
 
 for ( i in idx) {
   

### 2015-12-03 imaging-paper.Rnw
updating results

@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Maximal number of CMS between unknown Bullet D land 5 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown Bullet F land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
-load("data/unkn17.RData")
+load("data/unkn27.RData")
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Overlaid lines of the land of the unknown bullet D-5 and the land with the top matching CMS.}
+\caption{Overlaid lines of the land of the unknown bullet F-3 and the land with the top matching CMS.}
 <<top, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
@@
 
-\caption{Overlaid lines of the land of the unknown bullet D-5 and the land with the second top matching CMS.}
+\caption{Overlaid lines of the land of the unknown bullet F-3 and the land with the second top matching CMS.}
 <<top2, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
@@
 Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
-All 28 pairs of lands with at least nine CMS in common are matches.
+All 23 pairs of lands with at least twelve CMS in common are matches.
 For smaller number of CMS, the percentage of false positives quickly increases.
+Taking cross-correlation into account (see Figure~\ref{ccfs-match}) increases the number of matches.
+A cross-correlation of above 0.75 in the Hamby study is indicative of a match.
+There are 67 land-land comparisons with a cross-correlation of above 0.75.
 \begin{figure}
@@
 <<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
-bstats <- read.csv("data/bullet-stats.csv")
+#bstats <- read.csv("data/bullet-stats.csv")
 ggplot(data=bstats) + 
@@
 \begin{figure}
-<<distr.dist, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
 ggplot(data=bstats) + 
-    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match)) +
-    theme(legend.position="bottom") + 
+    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match), alpha=0.5) +
+   facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
+  ylab("Distance between crosscuts") +
+  theme_bw() + 
+  theme(legend.position="none") +
     xlab("maximum CMS")
@@
 \caption{If the distance between the crosscuts is small, a smaller number of CMS might already be an indication of a match.}
+\end{figure}
+
+\begin{figure}
+<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
+ggplot(data=bstats) + theme_bw() + 
+  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + ylab("Cross-correlation") +
+  theme(legend.position="none")
+@
+\caption{If the cross-correlations between the crosscuts is high, a smaller number of CMS might already be an indication of a match.}
+\end{figure}
+
+\begin{figure}
+<<cms-outlier, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+idx <- which(bstats$cms==11 & !bstats$match)[1]
+bstats[idx,]
+load(as.character(bstats$data[idx]))
+res <- reslist[[bstats$resID[idx]]]
+print(ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
+#  geom_hline(yintercept = res$threshold) +
+#  geom_hline(yintercept = - res$threshold) +
+  scale_colour_brewer(palette="Set1") +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+@
+\caption{Wrongfully identified match between two bullet lands with a maximum CMS of 11.}
 \end{figure}

### 2015-12-03 imaging-paper.Rnw
broken link

@@
 For smaller number of CMS, the percentage of false positives quickly increases.
-Taking cross-correlation into account (see Figure~\ref{ccfs-match}) increases the number of matches.
+Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches.
 A cross-correlation of above 0.75 in the Hamby study is indicative of a match.
@@
 idx <- which(bstats$cms==11 & !bstats$match)[1]
-bstats[idx,]
 load(as.character(bstats$data[idx]))

### 2015-12-04 imaging-paper.Rnw
slowsly adjusting the figures to the new matching routine

@@
 \begin{subfigure}[t]{\linewidth}
-\caption{Values above/below a threshold level are defined as peaks and grooves. \label{fig:smoothcutb}}{%
+\caption{Using a rolling median peaks and grooves are identified for each crosscut. \label{fig:smoothcutb}}{%
 <<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='\\textwidth', warning=FALSE>>=
-qplot(data=lofX, x=y, y=l30, colour=type, geom="line", 
-      group=bullet, size = I(0.75)) + 
-  facet_grid(bullet~.) +
-  ylab("Individual characteristics") + 
-#  geom_point(aes(y=r05)) +
-  geom_rect(xmin = 200, xmax = 2260, ymin = -.75, ymax = 0.75, colour=NA, fill=alpha("grey90", alpha =0.5)) +
-  scale_colour_brewer(palette="Set2", na.value = NA) +
-  theme_bw() +
-  theme(legend.position="bottom")
+  lof <- processBullets(paths = images[c(5,7)], x = 100)
+  lof <- bulletSmooth(lof)
+  bAlign = bulletAlign(lof)
+  lofX <- bAlign$bullet
+
+  b12 <- unique(lof$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 30)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 30)
+  p1 <- peaks1$plot + xlab("y") + ylab("smoothed crosscut") + ggtitle(b12[1])
+  p2 <- peaks2$plot + xlab("y") + ylab("smoothed crosscut") + ggtitle(b12[2])
+grid.arrange(p1, p2, ncol=1)
 @
@@
 \begin{subfigure}[t]{\linewidth}
-    \caption{Grey backgrounds identify a line on one of the bullets. Overlaps of the same color indicate matching striation marks.
-    Matching striations are marked by an `o', mismatches are marked by an `x'.   \label{fig:smoothcutd}}{%
+    \caption{Rectangles in the back identify a line on one of the bullets.
+Matching striation marks are indicated by color filled rectangles and arked by an `o'.
+Mismatches are filled in grey and  marked by an `x'.
+\label{fig:smoothcutd}}{%
 <<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth'>>=
+peaks1$lines$bullet <- b12[1]
+peaks2$lines$bullet <- b12[2]
+
+lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
 ggplot() + 
-  geom_rect(aes(xmin = miny, xmax = maxy), ymin = 0.25, ymax=2.75, fill="grey80", data = subset(lines, lineid != 0)) +
-  geom_tile(data=lofX, aes(x=y, y=bullet, fill=type)) + 
-  scale_y_discrete(expand = c(0.2,0)) +
-  scale_fill_brewer(palette="Set2") + 
-  theme_bw() +
-  ylab("") +
+  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6,  data = lines, alpha=0.2) +
   theme(legend.position="bottom") +
-  geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
-  geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
-@
-    }
-\end{subfigure}
-\caption{\label{fig:match} Matching striation marks: smooth (a), cut (b), and match (c).}
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
+  ylim(c(-6,6)) + theme_bw() +
+  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, colour=bullet)) +
+  scale_colour_brewer("", palette="Set1") +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
+  theme(legend.position="none") +
+  ylab(expression("Crosscut values in",mu,m)) +
+  xlab("y")
+
+# ggplot() +
+#   geom_rect(aes(xmin = miny, xmax = maxy), ymin = 0.25, ymax=2.75, fill="grey80", data = subset(lines, lineid != 0)) +
+#   geom_tile(data=lofX, aes(x=y, y=bullet, fill=type)) + 
+#   scale_y_discrete(expand = c(0.2,0)) +
+#   scale_fill_brewer(palette="Set2") + 
+#   theme_bw() +
+#   ylab("") +
+#   theme(legend.position="bottom") +
+#   geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
+#   geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
+@
+    }
+\end{subfigure}
+\caption{\label{fig:match} Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between crosscuts (c).}
 \end{figure}
@@
 
-\subsection{Varying threshold}
-
-In order to address the subjectivity involved in choosing a threshold level, we investigate this parameter more closely.
+%' \subsection{Varying threshold}
+%' 
+%' In order to address the subjectivity involved in choosing a threshold level, we investigate this parameter more closely.
 Figure~\ref{fig:thresholds} shows the relationship between threshold and the maximal number of consecutive matching striae.
@@
 In the autmatic matching routine, several thresholds are considered and the threshold resulting in the higest number of consecutive atching striae is chosen (and reported).
-\begin{figure}
-<<thresholds, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '0.49\\textwidth'>>=
-thresholds <- seq(0.3, 1.5, by = 0.025)
-CMS <- thresholds %>% lapply(function(threshold) {
- lines <- striation_identify(lofX, threshold = threshold)
- cms <- CMS(lines$match)
- data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
-}) %>% bind_rows()
-
-qplot(threshold, maxCMS, data=CMS) + theme_bw() +
-  ylab("Number of maximal CMS") +
-  ylim(c(0,15))
-@
-\caption{\label{fig:thresholds}Varying the threshold levels results in different number of consecutive matching striae between the two bullets.
+%' \begin{figure}
+%' <<thresholds, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '0.49\\textwidth'>>=
+%' thresholds <- seq(0.3, 1.5, by = 0.025)
+%' CMS <- thresholds %>% lapply(function(threshold) {
+%'  lines <- striation_identify(lofX, threshold = threshold)
+%'  cms <- CMS(lines$match)
+%'  data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
+%' }) %>% bind_rows()
+%' 
+%' qplot(threshold, maxCMS, data=CMS) + theme_bw() +
+%'   ylab("Number of maximal CMS") +
+%'   ylim(c(0,15))
+%' @
+%' \caption{\label{fig:thresholds}Varying the threshold levels results in different number of consecutive matching striae between the two bullets.
 By default, the threshold with the highest CMS is chosen automatically.}
-\end{figure}
+%' \end{figure}
 
@@
 \begin{enumerate}
-\item adding more non-matches into the mix is going to make things more realistic. 
+\item the Hamby study only includes 35 bullets - this is not a particularly realistic scenario for automatic matching. Adding more non-matches into the mix is going to make things more realistic. 
 \end{enumerate}

### 2015-12-04 code/checkBullets.R

@@ -2,10 +2,15 @@ library(x3pr)
 library(x3prplus)
 library(dplyr)
 library(ggplot2)
+library(zoo)
 
 knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
 knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 
+unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
+
+if (!file.exists("data/crosscuts.csv")) {
 crosscuts1 <- sapply(knowns, function(x) {
   cat(x)
   crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=12.5))
@@ -16,8 +21,6 @@ crosscuts1 <- sapply(knowns, function(x) {
 
 
 
-unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
-unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
 crosscuts2 <- sapply(unknowns, function(x) {
   cat(x)
@@ -33,23 +36,14 @@ dframe <- data.frame(
 )
 
 
-write.csv(dframe, file="crosscuts.csv", row.names=FALSE)
+write.csv(dframe, file="data/crosscuts.csv", row.names=FALSE)
+} 
 ccs <- read.csv("data/crosscuts.csv")
 
-knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 
 bullets <- lapply(c(knowns, unknowns), function(path) {
   lof <- processBullets(paths = path, x = ccs$cc[which(ccs$path==path)], check=FALSE)
   peaks <- get_peaks(lof)
-  peaks$extrema <- c(peaks$peaks, peaks$valleys)
-  peaks$type <- c(rep(1, length(peaks$peaks)), rep(-1, length(peaks$valleys)))
-  idx <- order(peaks$extrema)
-  peaks$extrema <- peaks$extrema[idx]
-  peaks$type <- peaks$type[idx]
-  diffs <- diff(peaks$extrema)
-  lines <- data.frame(xmin = peaks$extrema-c(diffs[1],diffs)/3,
-                      xmax=peaks$extrema+c(diffs,diffs[length(diffs)])/3, 
-                      type=peaks$type)
[…]

### 2015-12-04 code/complete-hamby.R

@@ -3,6 +3,7 @@ library(x3prplus)
 library(dplyr)
 library(ggplot2)
 library(gridExtra)
+library(zoo)
 
 knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
 knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
@@ -17,8 +18,8 @@ unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
 # match unknown land using crosscuts
 ccs <- read.csv("data/crosscuts.csv")
-load("data/bullets.RData")
-bPaths <- sapply(bullets, function(x) x$path)
+#load("data/bullets.RData")
+#bPaths <- sapply(bullets, function(x) x$path)
 
 
 for (j in 1:90) {
@@ -27,10 +28,10 @@ for (j in 1:90) {
     cat(x)
     cat("\n")
 #    browser()
-    b1 <- bullets[[which(bPaths == x)]]
-    b2 <- bullets[[which(bPaths == unknowns[j])]]
-    
-    bulletGetMaxCMSXXX(b1, b2)
+    cc1 <- ccs$cc[which(ccs$path == x)]
+    cc2 <- ccs$cc[which(ccs$path == unknowns[j])]
+
+    bulletGetMaxCMSXXX(x, unknowns[j], cc1, cc2)
   })
   save(reslist, file=sprintf("data/unkn%d.RData", j))
 }

### 2015-12-04 imaging-paper.Rnw
fig 14 and 15

@@
   scale_colour_brewer("", palette="Set1") +
-  theme_bw() +
+  theme_bw() + ylab("Residuals of Loess Fit") + 
   theme(legend.position="none")
@@
   peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 30)
-  p1 <- peaks1$plot + xlab("y") + ylab("smoothed crosscut") + ggtitle(b12[1])
-  p2 <- peaks2$plot + xlab("y") + ylab("smoothed crosscut") + ggtitle(b12[2])
+
+  library(RColorBrewer)
+  col <- brewer.pal(3, "Set1")
+  p1 <- ggplot() + theme_bw() +
+    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, data=peaks1$lines,  alpha=0.2) +
+    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks1$lines) + scale_colour_brewer(palette="Set2") + 
+    scale_fill_brewer(palette="Set2") +
+    theme(legend.position="none") + 
+    geom_line(aes(x=y, y=l30, group=bullet), data=subset(lofX, bullet==b12[1])) +
+  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
+    ggtitle(b12[1])+ xlim(c(0, 2000))
+
+  p2 <- ggplot() + theme_bw() +
+    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, data=peaks2$lines,  alpha=0.2) +
+    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks2$lines) + scale_colour_brewer(palette="Set2") + 
+    scale_fill_brewer(palette="Set2") +
+    theme(legend.position="none") + 
+    geom_line(aes(x=y, y=l30, group=bullet), data=subset(lofX, bullet==b12[2])) +
+  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
+    ggtitle(b12[2]) + xlim(c(0, 2000))
 grid.arrange(p1, p2, ncol=1)
@@
 \label{fig:smoothcutd}}{%
-<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth'>>=
+<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth', warning=FALSE>>=
 peaks1$lines$bullet <- b12[1]
@@
 ggplot() + 
-  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6,  data = lines, alpha=0.2) +
+  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
   theme(legend.position="bottom") +
@@
   geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
-  ylim(c(-6,6)) + theme_bw() +
+  ylim(c(-6,6.5)) + theme_bw() +
   geom_line(data=lofX, aes(x=y, y=l30, group=bullet, colour=bullet)) +
@@
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
-  theme(legend.position="none") +
-  ylab(expression("Crosscut values in",mu,m)) +
+    theme(legend.position = c(1,1), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
   xlab("y")
-
-# ggplot() +
-#   geom_rect(aes(xmin = miny, xmax = maxy), ymin = 0.25, ymax=2.75, fill="grey80", data = subset(lines, lineid != 0)) +
-#   geom_tile(data=lofX, aes(x=y, y=bullet, fill=type)) + 
-#   scale_y_discrete(expand = c(0.2,0)) +
-#   scale_fill_brewer(palette="Set2") + 
-#   theme_bw() +
-#   ylab("") +
-#   theme(legend.position="bottom") +
-#   geom_text(aes(x = meany), y= .4, label= "x", data = subset(lines, !match & lineid !=0)) +
-#   geom_text(aes(x = meany), y= .4, label= "o", data = subset(lines, match & lineid !=0))
 @
@@
 reslist <- lapply(crosscuts[-1], function(cc) {
+#  browser()
   b2 <- subset(lof, x %in% c(cc, 100))
@@
   lofX <- bAlign$bullet
-  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-
-  lines <- striation_identify(lofX, threshold = threshold)
+    b12 <- unique(b2$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
+
+#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+#  lines <- striation_identify(lofX, threshold = threshold)
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
   maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, threshold=threshold, lines=lines, bullets=lofX)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
 })
@@
   theme_bw() + 
-  geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
   geom_line(aes(x = y, y = l30, colour = bullet),   data = res$bullets) +
-  geom_hline(yintercept = res$threshold) +
-  geom_hline(yintercept = - res$threshold) +
-  scale_colour_brewer("", palette="Set1") +
+  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
   theme(legend.position = c(1,1.2), legend.justification=c(1,1),
@@
   lofX <- bAlign$bullet
-  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-
-  lines <- striation_identify(lofX, threshold = threshold)
+
+  b12 <- unique(b2$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
+
+#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+#  lines <- striation_identify(lofX, threshold = threshold)
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
   maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, threshold=threshold, ccf = bAlign$ccf, lines=lines, bullets=lofX)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
 })

### 2015-12-04 imaging-paper.Rnw
figures completed

@@
   scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
   theme(legend.position = c(1,1.2), legend.justification=c(1,1),
@@
 \begin{figure}
-<<barrel2, echo=FALSE, fig.width = 7.5, fig.height=7.5, out.width='\\linewidth', warning=FALSE>>=
-thresholds <- seq(0.3, 1.5, by = 0.05)
-list_of_matches <- lapply(19:24, function(i) {
-  lof <- processBullets(paths = images[c(15,i)], x = 100)
+<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='\\linewidth', warning=FALSE>>=
+
+list_of_matches <- lapply(25:30, function(i) {
+  lof <- processBullets(paths = images[c(31,i)], x = 100)
   lof <- bulletSmooth(lof)
 
-  subLOFx1 <- subset(lof, bullet==unique(lof$bullet)[1])
-  subLOFx2 <- subset(lof, bullet==unique(lof$bullet)[2])
-
-  # ccf assumes that both time series start at the same y
-  subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
-  subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-  ccf <- ccf(subLOFx1$l30, subLOFx2$l30, plot = FALSE, lag.max=150, na.action = na.omit)
-  lag <- ccf$lag[which.max(ccf$acf)]
-
-  subLOFx1$y <- subLOFx1$y -  lag * 1.5625 # amount of shifting should just be lag * y.inc
-  lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
-
-#  browser()  
-  CMS <- thresholds %>% lapply(function(threshold) {
-    lines <- striation_identify(lofX, threshold = threshold)
-    cms <- CMS(lines$match)
-    data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
-  }) %>% bind_rows()
-
-  threshold = CMS$threshold[which.max(CMS$maxCMS)]
-  lines <- striation_identify(lofX, threshold = threshold)
-  title <- gsub("app.*//","", images[i])
-  title <- gsub(".x3p","", title)
+  bAlign = bulletAlign(lof)
+  lofX <- bAlign$bullet
+
+  b12 <- unique(lof$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
+  peaks1$lines$bullet = b12[1]
+  peaks2$lines$bullet = b12[2]
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
   ggplot() +
     theme_bw() + 
-    geom_rect(aes(xmin=miny, xmax=maxy), ymin=-6, ymax=5, fill="grey90", data=subset(lines, lineid!=0)) +
+    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=6, data=lines, alpha=0.2) +
     geom_line(aes(x = y, y = l30, colour = bullet),  data = lofX) +
-    geom_hline(yintercept = threshold) +
-    geom_hline(yintercept = - threshold) +
-    scale_colour_brewer(palette="Set1") +
-    theme(legend.position = "none") + 
+#    geom_hline(yintercept = threshold) +
+#    geom_hline(yintercept = - threshold) +
+    scale_colour_brewer("",palette="Set1") +
+    scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
     ylim(c(-6,6)) + ylab("") + xlab("") +
-    geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match & lineid !=0)) +
-    geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match & lineid !=0)) +
-    ggtitle(title) +
+    geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
+    geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
     theme(plot.margin=unit(c(1,0,0,0), unit="line"))
@@
              list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
-             ncol = 2)
-@
-\caption{\label{fig:barrel2}All lands from the second bullet shot from barrel 2 are compared to the third land of the first bullet shot from barrel 2 (in red).
-The largest number of CMS is seven, observed with land four of the second bullet.
+             ncol = 1)
+@
+\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to the first land of the second bullet shot from barrel 2 (in blue).
+The largest number of CMS is eleven, observed with land six of the second bullet.
 }
@@
   theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
   geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
@@
 #  geom_hline(yintercept = - res$threshold) +
-  scale_colour_brewer(palette="Set1") +
+  scale_colour_brewer("",palette="Set1") +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
   theme(legend.position = c(1,1), legend.justification=c(1,1)) + 

### 2015-12-04 code/complete-hamby.R

@@ -31,7 +31,7 @@ for (j in 1:90) {
     cc1 <- ccs$cc[which(ccs$path == x)]
     cc2 <- ccs$cc[which(ccs$path == unknowns[j])]
 
-    bulletGetMaxCMSXXX(x, unknowns[j], cc1, cc2)
+    bulletGetMaxCMSXXX(x, unknowns[j], cc1, cc2, span=30)
   })
   save(reslist, file=sprintf("data/unkn%d.RData", j))
 }

### 2015-12-05 imaging-paper.Rnw
decision tree included

@@
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
-load("data/unkn27.RData")
+load("data-35/unkn27.RData")
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
@@
 <<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
-bstats <- read.csv("data/bullet-stats.csv")
+bstats <- read.csv("data-35/bullet-stats.csv")
 ggplot(data=bstats) + 
@@
 <<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
-#bstats <- read.csv("data/bullet-stats.csv")
 ggplot(data=bstats) + 
@@
 
-\begin{figure}
-<<cms-outlier, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
-idx <- which(bstats$cms==11 & !bstats$match)[1]
-load(as.character(bstats$data[idx]))
-res <- reslist[[bstats$resID[idx]]]
-print(ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax), ymin=-6, ymax=5, fill="grey90", data=res$lines) +
-  geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
-#  geom_hline(yintercept = res$threshold) +
-#  geom_hline(yintercept = - res$threshold) +
-  scale_colour_brewer(palette="Set1") +
-  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-@
-\caption{Wrongfully identified match between two bullet lands with a maximum CMS of 11.}
-\end{figure}
-
 \hh{XXX another idea for the rest of the paper: use some datamining methods to match the Hamby bullets }
+Using recursive partitioning, we find a decision tree predicting matches based on features derived from the image files.
+The resulting tree is shown in Figure~\ref{fig:tree}.
+A total of 93 lands are being matched correctly.
+Interestingly, the number of consecutive matching striae is not featured in this evaluation.
+Instead, cross-correlation (ccf) between the crosscuts is featured heavily.
+Aside from cross correlation, the number of matches between crosscuts is involved in the decision.
+
+\begin{figure}
+<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+library(rpart)
+library(rpart.plot)
+CCFs <- read.csv("data-35/bullet-stats.csv")
+includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y"))
+rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
+prp(rp1)
+
+CCFs$pred35 <- predict(rp1)
+@
+\caption{\label{fig:tree}Decision tree of matching bullets based on recursive partitioning.}
+\end{figure}
+
+<<pred:table, echo=FALSE, results='asis'>>=
+library(xtable)
+xtable(xtabs(~I(pred35>0.5)+match, data=CCFs))
+@
+
+\begin{figure}
+<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+CCFs$bullet <- gsub("-[0-9]$", "", CCFs$b2)
+bullets <- CCFs %>% group_by(bullet) %>% summarize(
+  n = n(),
+  pred = sum(pred35 > 0.5)
+)
+
+qplot(data=bullets, pred, reorder(bullet, pred/n)) + 
+  theme_bw()  + ylab("") + 
+  scale_x_continuous("Number of correctly predicted land-to-land matches", 
+                     breaks = 3*0:4, limits=c(0,12))
+@
+\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree.
+For each bullet at least one land is correctly matched.
+Unknown bullet B is the hardest for the algorithm to match.}
+\end{figure}
 

### 2015-12-05 code/evaluation-hamby.R

@@ -1,6 +1,6 @@
 
-datas <- file.path("data", dir("data", pattern="RData"))
-datas <- datas[grep("data/u.*", datas)]
+datas <- file.path("data-30", dir("data-30", pattern="RData"))
+datas <- datas[grep("data-30/u.*", datas)]
 
 maxCMS <- sapply(datas, function(x) {
   load(x)
@@ -90,24 +90,26 @@ CCFs <- merge(CCFs, mm[,c("id","value")], by.x="b1", by.y="value")
 CCFs <- merge(CCFs, mm[,c("id","value")], by.x="b2", by.y="value")
 CCFs$match <- CCFs$id.x == CCFs$id.y
 
-write.csv(CCFs, file="data/bullet-stats.csv", row.names=FALSE)
+write.csv(CCFs, file="data-30/bullet-stats.csv", row.names=FALSE)
 
 # diagnostics
 
 library(ggplot2)
-CCFs <- read.csv("data-35/bullet-stats.csv")
+CCFs <- read.csv("data-30/bullet-stats.csv")
 
 qplot(factor(cms), data=CCFs)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(cms), fill=match), position="fill")
+sum(CCFs$cms >= 12)
 ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match))
 ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=ccf, colour=match)) + facet_wrap(~match)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(non_cms), fill=match), position="fill")
+qplot(ccf, num.matches, geom="jitter", data=CCFs, colour=match, alpha=0.1) + facet_wrap(~match)
 
 library(rpart)
+library(rpart.plot)
 includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y"))
 rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
-plot(rp1)
-text(rp1) 
+prp(rp1)
 
 qplot(ccf, num.matches, geom="jitter", data=CCFs, colour=match, alpha=0.1) + facet_wrap(~match)
 
@@ -129,16 +131,24 @@ qplot(ccf, num.matches, geom="jitter", data=subCCFs, colour=dist)
 
[…]

### 2015-12-05 imaging-paper.Rnw
fixing description

@@
 \item Use cross-correlation to find the best alignment along the crosscuts: Shift one of the crosscuts by the lag indicated by the cross-correlation function.
-\item Based on a starting threshold level of $\theta$, identify areas of the crosscuts whose absolute value are above the thresholds.
+\item Using a rolling average, identify peaks and grooves for each of the crosscuts. We define a region around the extrema on each side as one third of the distance to the next extrema. 
+%Based on a starting threshold level of $\theta$, identify areas of the crosscuts whose absolute value are above the thresholds.
 These areas are regions of potential matches between the crosscuts.
@@
 If only one crosscut exceeds the threshold, or if the two crosscuts move into opposite directions over the threshold, the area is a non-match.
+\item Match grooves and peaks between crosscuts. \hh{XXX  needs some more details.} 
 \item Identify the maximal number of CMS (consecutive matching striae) between the two crosscuts by counting consecutive matching lines. 
-Vary $\theta$ in (3), such that a maximum number of CMS is reached.
 \end{enumerate}
@@
 @
-\caption{If the cross-correlations between the crosscuts is high, a smaller number of CMS might already be an indication of a match.}
+\caption{\label{fig:ccfs-match}If the cross-correlations between the crosscuts is high, a smaller number of CMS might already be an indication of a match.}
 \end{figure}
@@
 Aside from cross correlation, the number of matches between crosscuts is involved in the decision.
+This is not in contradiction to earlier papers emphasizing the value of CMS on visual assessments of bullet matches.
 

### 2015-12-05 code/evaluation-hamby.R

@@ -1,6 +1,7 @@
-
-datas <- file.path("data-30", dir("data-30", pattern="RData"))
-datas <- datas[grep("data-30/u.*", datas)]
+span <- 35
+dataStr <- sprintf("data-%d", span)
+datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
+datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
 
 maxCMS <- sapply(datas, function(x) {
   load(x)
@@ -48,10 +49,25 @@ CCFs$resID <- rep(1:120, length=nrow(CCFs))
 CCFs <- CCFs[order(as.character(CCFs$b2)),]
 CCFs$b2 <- factor(as.character(CCFs$b2))
 
-CCFs$match <- NA
-idx <- which(CCFs$cms >= 9) # all are matches, visually confirmed
-CCFs$match[idx] <- TRUE
-write.csv(CCFs, file="bullet-stats.csv", row.names=FALSE)
+
+matches <- read.csv("csvs/matches.csv", header=FALSE, stringsAsFactors = FALSE)
+matches$V3 <- paste("Ukn Bullet",matches$V3)
+matches$V4 <- paste("Ukn Bullet",matches$V4)
+matches$V5 <- paste("Ukn Bullet",matches$V5)
+matches$id <- 1:nrow(matches)
+
+library(reshape2)
+mm <- melt(matches, id.var="id")
+mm <- subset(mm, value != "Ukn Bullet ")
+
+CCFs <- merge(CCFs, mm[,c("id","value")], by.x="b1", by.y="value")
+CCFs <- merge(CCFs, mm[,c("id","value")], by.x="b2", by.y="value")
+CCFs$match <- CCFs$id.x == CCFs$id.y
+
+write.csv(CCFs, file=file.path(dataStr, "bullet-stats.csv"), row.names=FALSE)
+
+# diagnostics
+
 
 idx <- which(CCFs$match & CCFs$cms <= 6)
 
@@ -59,7 +75,7 @@ for ( i in idx) {
   
   load(CCFs$data[i])
   res <- reslist[[CCFs$resID[i]]]  
-
+  
   ch <- scan()
   print(ggplot() +
           theme_bw() + 
@@ -76,26 +92,8 @@ for ( i in idx) {
 }
 
 
-matches <- read.csv("csvs/matches.csv", header=FALSE, stringsAsFactors = FALSE)
[…]

### 2015-12-05 imaging-paper.Rnw
adding node sizes to decision tree

@@
 
-Figure~\ref{fig:match} gives an overview of the rest of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if it exceeds a pre-set threshold (of 0.75$\mu m$; obviously this is a crucial parameter.
-For more discussion see Figure~\ref{fig:thresholds}).
+Figure~\ref{fig:match} gives an overview of the rest of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if \hh{XXX }
+
 Two lines are defined as a match, if their peak/groove base at the threshold overlaps.
@@
 In this example, the number of consecutive matching striations (CMS) is fifteen, a very high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
-\hh{XXX need more references XXX \cite{nichols:2003} XXX what does the paper say?
-AFTE is so frustrating!
-XXX }
-
-%' \subsection{Varying threshold}
-%' 
-%' In order to address the subjectivity involved in choosing a threshold level, we investigate this parameter more closely.
-Figure~\ref{fig:thresholds} shows the relationship between threshold and the maximal number of consecutive matching striae.
-For thresholds between 0.7 and 0.9 the number of CMS peaks at fifteen.
-On average, there are 6.6 CMS across the range of thresholds.
-In the autmatic matching routine, several thresholds are considered and the threshold resulting in the higest number of consecutive atching striae is chosen (and reported).
-%' \begin{figure}
-%' <<thresholds, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '0.49\\textwidth'>>=
-%' thresholds <- seq(0.3, 1.5, by = 0.025)
-%' CMS <- thresholds %>% lapply(function(threshold) {
-%'  lines <- striation_identify(lofX, threshold = threshold)
-%'  cms <- CMS(lines$match)
-%'  data.frame(threshold=threshold, maxCMS = as.numeric(rev(names(cms)))[1])
-%' }) %>% bind_rows()
-%' 
-%' qplot(threshold, maxCMS, data=CMS) + theme_bw() +
-%'   ylab("Number of maximal CMS") +
-%'   ylim(c(0,15))
-%' @
-%' \caption{\label{fig:thresholds}Varying the threshold levels results in different number of consecutive matching striae between the two bullets.
-By default, the threshold with the highest CMS is chosen automatically.}
-%' \end{figure}
+For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches  \cite{nichols:2003}.
 
@@
 The resulting tree is shown in Figure~\ref{fig:tree}.
-A total of 93 lands are being matched correctly.
+A total of 93 lands is being matched correctly.
 Interestingly, the number of consecutive matching striae is not featured in this evaluation.
@@
 Aside from cross correlation, the number of matches between crosscuts is involved in the decision.
-This is not in contradiction to earlier papers emphasizing the value of CMS on visual assessments of bullet matches.
+Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
+Out of the two, cross-correlation is the variable with more  predictive power.
+This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2d microscopic images of bullets.
+Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitalized version of the images that preserves the 3D surface structure.
 
@@
 rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
-prp(rp1)
+
+per <- rp1$frame$yval # predictive probability for each node in the tree
+prp(rp1, extra = 101, fallen.leaves=TRUE, box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
 

### 2015-12-05 code/evaluation-hamby.R

@@ -152,4 +152,5 @@ qplot(bullet, data=CCFs, fill=match) + coord_flip()
 qplot(bullet, data=CCFs, fill=pred35 > 0.5) + coord_flip()
 
 library(rattle)
-fancyRpartPlot(rp1)		
+fancyRpartPlot(rp1, sub="")		
+prp(rp1, extra = 101, fallen.leaves=TRUE)

### 2015-12-05 imaging-paper.Rnw
couple of citations

@@
 library(tidyr)
+library(rpart)
+library(rpart.plot)
 
@@
 \end{subfigure}
-\caption{\label{fig:match} Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between crosscuts (c).}
+\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between crosscuts (c).}
 \end{figure}
@@
 \hh{XXX another idea for the rest of the paper: use some datamining methods to match the Hamby bullets }
-Using recursive partitioning, we find a decision tree predicting matches based on features derived from the image files.
+Using recursive partitioning, we find a decision tree \cite{breiman:1984, rpart, rpart.plot} predicting matches based on features derived from the image files.
 The resulting tree is shown in Figure~\ref{fig:tree}.
@@
 
+\paragraph{Acknowledgment}: thanks to the men and women behind R\cite{R}, and the authors of the packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
 \section{Conclusion}
@@
 \end{enumerate}
+
 \bibliographystyle{abbrv}

### 2015-12-06 imaging-paper.Rnw
added heights of peaks to stats

@@
 
-\paragraph{Acknowledgment}: thanks to the men and women behind R\cite{R}, and the authors of the packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
+\paragraph{Acknowledgment} Thanks to the men and women behind the software R\cite{R}, and the authors of the R packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
 \section{Conclusion}

### 2015-12-06 code/complete-hamby.R

@@ -11,29 +11,28 @@ knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
+span <- 35
+dataStr <- sprintf("data-%d", span)
 
 ###############
 # can we identify the barrels the unknown bullets came from?
 
 
 # match unknown land using crosscuts
-ccs <- read.csv("data/crosscuts.csv")
-#load("data/bullets.RData")
-#bPaths <- sapply(bullets, function(x) x$path)
-
+ccs <- read.csv(file.path(dataStr, "crosscuts.csv"))
 
 for (j in 1:90) {
   
   reslist <- lapply(knowns, function(x) {
     cat(x)
     cat("\n")
-#    browser()
+
     cc1 <- ccs$cc[which(ccs$path == x)]
     cc2 <- ccs$cc[which(ccs$path == unknowns[j])]
 
-    bulletGetMaxCMSXXX(x, unknowns[j], cc1, cc2, span=30)
+    bulletGetMaxCMSXXX(x, unknowns[j], cc1, cc2, span=span)
   })
-  save(reslist, file=sprintf("data-30/unkn%d.RData", j))
+  save(reslist, file=file.path(dataStr, sprintf("unkn%d.RData", j)))
 }
 
 

### 2015-12-06 code/evaluation-hamby.R

@@ -22,7 +22,7 @@ CCFs <- plyr::ldply(datas, function(x) {
     
     subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
     subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
- #     browser()
+
     ccf <- ccf(subLOFx1$val, subLOFx2$val, plot = FALSE, lag.max=200, na.action = na.omit)
     lag <- ccf$lag[which.max(ccf$acf)]
     incr <- min(diff(sort(unique(subLOFx1$y))))
@@ -31,14 +31,19 @@ CCFs <- plyr::ldply(datas, function(x) {
     ys <- intersect(subLOFx1$y, subLOFx2$y)
     idx1 <- which(subLOFx1$y %in% ys)
     idx2 <- which(subLOFx2$y %in% ys)
-    distr.dist <- mean((subLOFx1$val[idx1] - subLOFx2$val[idx2])^2, na.rm=T)
+    distr.dist <- mean((subLOFx1$val[idx1] - subLOFx2$val[idx2])^2, na.rm=TRUE)
+    distr.sd <- sd(subLOFx1$val, na.rm=TRUE) + sd(subLOFx2$val, na.rm=TRUE)
     
     # feature extraction
-    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), distr.dist=distr.dist, 
+    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), 
+               distr.dist=distr.dist, 
+               distr.sd = distr.sd,
                b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
                num.matches = sum(res$lines$match), 
                num.mismatches = sum(!res$lines$match), 
-               non_cms = x3prplus::maxCMS(!res$lines$match))
+               non_cms = x3prplus::maxCMS(!res$lines$match),
+               sumpeaks = sum(abs(res$lines$heights[res$lines$match]))
+               )
   })
   ccf$cms <- cmsdist
   ccf$data <- x
@@ -70,20 +75,23 @@ write.csv(CCFs, file=file.path(dataStr, "bullet-stats.csv"), row.names=FALSE)
 
 
 idx <- which(CCFs$match & CCFs$cms <= 6)
+idx <- which(!CCFs$match & CCFs$ccf > .7)
[…]

### 2015-12-07 imaging-paper.Rnw
fixed figure 17

@@
 For bullet land 1-5 from barrel 1 crosscuts at intervals of 50$\mu m$ between 150$\mu m$  and 450$\mu m$  are compared to a crosscut at 100$\mu m$.
-The initially almost perfect match (consolidated by a CMS of nineteen) between the  crosscuts deteriorates with increasing distance to a very questionable match at $x = 350\mu m$ and non-matches further apart.
-Only if the crosscut is within 200$\mu m$ do we get a good visual match.
+The initially almost perfect match (consolidated by a CMS of fourteen) between the  crosscuts deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.
+Only if the crosscut is within 150$\mu m$ do we get a good visual match.
 Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections.
@@
 Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$.
-With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 19 to only 3 at $x = 400\mu m$ and $x = 450\mu m$.
+With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 15 to  3 or fewer at $x = 300\mu m$ and higher.
 }
@@
 #res <- reslist[[which.max(cmsdist)]]  
-res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
+res <- reslist[[11]]  # number 2
 
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{All bullet pairs with a maximum of at least nine CMS are matches.}
+\caption{All land-to-land comparisons a maximum number of CMS of twelve or more are matches.}
 <<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
@@
 The resulting tree is shown in Figure~\ref{fig:tree}.
-A total of 93 lands is being matched correctly.
+A total of 105 lands is being matched correctly.
 Interestingly, the number of consecutive matching striae is not featured in this evaluation.
@@
 Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
-Out of the two, cross-correlation is the variable with more  predictive power.
+Out of the two, cross-correlation has higher  predictive power.
 This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2d microscopic images of bullets.
@@
 \item the Hamby study only includes 35 bullets - this is not a particularly realistic scenario for automatic matching. Adding more non-matches into the mix is going to make things more realistic. 
+\item parameter optimization is going to drive some of the matches
+\item involve multiple crosscuts for each bullet in a match?
 \end{enumerate}
 
+Things to be dicussed in the future
+\begin{enumerate}
+\item 
+\end{enumerate}
 \bibliographystyle{abbrv}

### 2015-12-07 code/complete-hamby.R

@@ -11,7 +11,7 @@ knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
-span <- 35
+span <- 25
 dataStr <- sprintf("data-%d", span)
 
 ###############

### 2015-12-07 code/evaluation-hamby.R

@@ -1,4 +1,4 @@
-span <- 35
+span <- 30
 dataStr <- sprintf("data-%d", span)
 datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
 datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
@@ -113,11 +113,13 @@ qplot(ccf, num.matches, geom="jitter", data=CCFs, colour=match, alpha=I(0.5)) + 
 qplot(cms, sumpeaks, geom="jitter", data=CCFs, colour=match, alpha=I(0.5)) + facet_wrap(~match)
 qplot(cms, distr.sd, geom="jitter", data=CCFs, colour=match, alpha=I(0.5)) 
 
+qplot(factor(cms), fill=match, data=CCFs) + facet_wrap(~b2)
+
 library(rpart)
 library(rpart.plot)
 includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y"))
 rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
-prp(rp1)
+prp(rp1, extra = 101)
 
 qplot(ccf, num.matches, geom="jitter", data=CCFs, colour=match, alpha=0.1) + facet_wrap(~match)
 
@@ -161,6 +163,4 @@ CCFs$bullet <- gsub("-[0-9]$", "", CCFs$b2)
 qplot(bullet, data=CCFs, fill=match) + coord_flip()
 qplot(bullet, data=CCFs, fill=pred35 > 0.5) + coord_flip()
 
-library(rattle)
-fancyRpartPlot(rp1, sub="")		
 prp(rp1, extra = 101, fallen.leaves=TRUE)

### 2015-12-07 code/complete-hamby.R
cleaning up the code and adding a few comments

@@ -51,155 +51,4 @@ for (j in 1:90) {
 }
 
 
-
-# match unknown land using crosscuts
-ccs <- read.csv("data/crosscuts.csv")
-for (j in 1:90) {
-  
-  reslist <- lapply(knowns, function(x) {
-    cat(x)
-    cat("\n")
-    browser()
-    bulletGetMaxCMS(x, unknowns[j], 
-                    crosscut = ccs$cc[which(ccs$path==x)], 
-                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
-  })
-  save(reslist, file=sprintf("data/unkn%d.RData", j))
-}
-
-# match some known lands
-br4 <- knowns[grep("Br4", knowns)]
-reslist <- list()
-for (j in 1) {
-  res <- lapply(br4[7:12], function(x) {
-    cat(x)
-    cat("\n")
-    
-    bulletGetMaxCMS(x, br4[j], 
-                    crosscut = ccs$cc[which(ccs$path==x)], 
-                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
-  })
-  reslist <- c(reslist, res)
-}
-save(reslist, file=sprintf("data/br4-%d.RData", j))
-
-br <- knowns[grep("Br8", knowns)]
-reslist <- list()
-for (j in 2) {
-  res <- lapply(br[7:12], function(x) {
-    cat(x)
-    cat("\n")
-    
-    bulletGetMaxCMS(x, br[j], 
-                    crosscut = ccs$cc[which(ccs$path==x)], 
-                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
-  })
-  reslist <- c(reslist, res)
-}
-save(reslist, file=sprintf("data/br8-%d.RData", j))
-
-br <- knowns[grep("Br9", knowns)]
-reslist <- list()
-for (j in 1) {
-  res <- lapply(br[7:12], function(x) {
-    cat(x)
-    cat("\n")
-    
-    bulletGetMaxCMS(x, br[j], 
-                    crosscut = ccs$cc[which(ccs$path==x)], 
-                    crosscut2 = ccs$cc[which(ccs$path==unknowns[j])], check=TRUE)
-  })
-  reslist <- c(reslist, res)
-}
-cmsdist <- sapply(reslist, function(x) x$maxCMS)
-qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
-
[…]

### 2015-12-07 imaging-paper.Rnw
cleaning up the code and adding a few comments
Rewrite the explanation of loess groove smoothing

@@
 \begin{enumerate}
-    \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step12}).
-    \item For each y value, compute a rolling average to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step12}).
-    \item Determine the peak of the left groove by finding the first value with a negative discretized derivative.
-Determine the valley of the left groove by finding the first value with a positive discretized derivative after the value of the peak.
-(Figure~\ref{fig:loess_step34}).
-    \item Determine the peak of the right groove by finding the last value with a positive discretized derivative.
-Determine the valley of the right groove by finding the last value with a negative discretized derivative before the value of the peak.
-(Figure~\ref{fig:loess_step34}).
-    \item Remove all $y$ values left and right of the identified minima in vertical direction (Figure \ref{fig:loess_step34}).
+    \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step1}).
+    \item For each $y$ value, compute a rolling average with a particular \textbf{smoothing factor} to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step2}).
+    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor above. (Figure~\ref{fig:loess_step3}).
+    \item Determine the peak of the left groove by finding the first doubled-smoothed value $y_i$ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$.
+Call this peak $p_{left}$.
+(Figure~\ref{fig:loess_step47}).
+    \item Determine the valley of the left groove by finding the first doubled-smoothed $y_j > p_{left}$ such that $y_j < y_{j - 1}$ and $y_j < y_{j + 1}$. Call this valley $v_{left}$.
+    \item Reverse the order of the $y$ values and repeat the previous two steps to find the peak and valley of the right groove, $(p_{right}, v_{right})$.
+    \item Remove all $y_i < v_{left}$ and $y_i > v_{right}$. (Figure \ref{fig:loess_step47}).
 \end{enumerate}
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step12}Steps 1 and 2 of the Loess fit algorithm.
-For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.
-The height values are smoothed to remove noise in the deviations.}
-<<loess_step12, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.}
+<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
 br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
@@
 p1
-
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_step2}Step 2 of the Loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
+<<loess_step2, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
 subdata <- br111_bullet
-subdata$smoothed <- c(rep(NA, 10), rollapply(subdata$value, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
-subdata$smoothed2 <- c(rep(NA, 10), rollapply(subdata$smoothed, 21, function(x) mean(x, na.rm = TRUE)), rep(NA, 10))
-p2 <- qplot(y, smoothed2, data = subdata) + theme_bw() + xlim(c(0,2500))
-
+smoothfactor <- 35
+
+value_filled <- na.fill(subdata$value, "extend")
+smoothed <- rollapply(value_filled, smoothfactor, function(x) mean(x))
+smoothed_truefalse <- rollapply(smoothed, smoothfactor, function(x) mean(x))
+
+p2 <- qplot(1:length(smoothed), smoothed) + theme_bw()
 p2
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step34}Steps 3 - 5 of the Loess fitting algorithm.
+\caption{\label{fig:loess_step3}Step 3 of the Loess fit algorithm. The smoothed height values are smoothed one more time with the same smoothing factor.}
+<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+p3 <- qplot(1:length(smoothed_truefalse), smoothed_truefalse) + theme_bw()
+p3
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_step47}Steps 4 - 7 of the Loess fitting algorithm.
 After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
 Values outside these boundaries are removed (shown in grey)}
-<<loess_step34, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-peak_ind <- head(which(diff(subdata$smoothed2) < 0), n = 1)
-    groove_ind <- head(which(diff(tail(subdata$smoothed2, n = -(peak_ind + 10))) > 0), n = 1) + peak_ind + 10
-
-    peak_ind2 <- tail(which(diff(subdata$smoothed2) > 0), n = 1)
-    groove_ind2 <- tail(which(diff(head(subdata$smoothed2, n = -(length(subdata$smoothed2) - peak_ind2 + 10))) < 0), n = 1)
-
-    p4 <- qplot(subdata$y[subdata$y < subdata$y[groove_ind]], subdata$smoothed2[subdata$y < subdata$y[groove_ind]], colour=I("grey60"), alpha=I(.25)) +
+<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)
+
+    peak_ind_smoothed <- head(which(rollapply(smoothed_truefalse, 3, function(x) which.max(x) == 2)), n = 1)
+    peak_ind <- peak_ind_smoothed + floor(lengthdiff / 2)
+    groove_ind <- head(which(rollapply(tail(smoothed_truefalse, n = -peak_ind_smoothed), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind
+
+    peak_ind2_smoothed_temp <- head(which(rollapply(rev(smoothed_truefalse), 3, function(x) which.max(x) == 2)), n = 1)
+    peak_ind2_temp <- peak_ind2_smoothed_temp + floor(lengthdiff / 2)
+    groove_ind2_temp <- head(which(rollapply(tail(rev(smoothed_truefalse), n = -peak_ind2_smoothed_temp), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind2_temp
+
+    peak_ind2 <- length(subdata$value) - peak_ind2_temp + 1
+    groove_ind2 <- length(subdata$value) - groove_ind2_temp + 1
+
+    p4 <- qplot(subdata$y[subdata$y < subdata$y[groove_ind]], subdata$value[subdata$y < subdata$y[groove_ind]], colour=I("grey60"), alpha=I(.25)) +
         ylim(c(min(subdata$value[!is.nan(subdata$value)]) - 25, max(subdata$value[!is.nan(subdata$value)]) + 25)) +
         xlim(c(min(subdata$y) - 25, max(subdata$y) + 25)) +
-        geom_point(aes(x, y), data = data.frame(x = subdata$y[subdata$y > subdata$y[groove_ind2]], y = subdata$smoothed2[subdata$y > subdata$y[groove_ind2]]), colour=I("grey60"), alpha=I(.25)) +
-        geom_point(inherit.aes = FALSE, aes(x, y), data = data.frame(x = subdata$y[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]], y = subdata$smoothed2[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]])) +
+        geom_point(aes(x, y), data = data.frame(x = subdata$y[subdata$y > subdata$y[groove_ind2]], y = subdata$value[subdata$y > subdata$y[groove_ind2]]), colour=I("grey60"), alpha=I(.25)) +
+        geom_point(inherit.aes = FALSE, aes(x, y), data = data.frame(x = subdata$y[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]], y = subdata$value[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]])) +
         theme_bw() +
@@
 \begin{enumerate}
-\item 
+\item reduce the stretch of crosscuts and match smaller pieces. This allows for both a more realistic scenario, as well as a quantification of the individuality of the segment.
 \end{enumerate}

### 2015-12-07 code/evaluation-hamby.R
single file (bullet-stats.csv) for three spans: 25, 30, 35

@@ -124,7 +124,7 @@ prp(rp1, extra = 101)
 qplot(ccf, num.matches, geom="jitter", data=CCFs, colour=match, alpha=0.1) + facet_wrap(~match)
 
 CCFs$pred35 <- predict(rp1)
-xtabs(~pred35+match, data=CCFs)
+xtabs(~(pred35>.5)+match, data=CCFs)
 
 # the ccf/num.matches distribution of non-matches looks like a normal distribution - 
 # let's find a mean and compute (a standardized) distance from that
@@ -164,3 +164,37 @@ qplot(bullet, data=CCFs, fill=match) + coord_flip()
 qplot(bullet, data=CCFs, fill=pred35 > 0.5) + coord_flip()
 
 prp(rp1, extra = 101, fallen.leaves=TRUE)
+
+#####################
+# throw in all three evaluations into the mix:
+
+bstats <- NULL
+for (i in c(35, 30, 25)) {
+  dataStr <- sprintf("data-%d", i)
+  temp <- read.csv(file.path(dataStr, "bullet-stats.csv"))
+  includes <- setdiff(names(temp), c("b1", "b2", "data", "resID", "id.x", "id.y"))
+  temp$diffx <- with(temp, abs(x1-x2))
+  temp$perc_matches <- with(temp, num.matches/(num.matches+num.mismatches))
+  rp <- rpart(match~., data=temp[,includes])
+  
+  prp(rp, extra = 101)
+  ch <- scan()
+  temp$pred <- predict(rp)
+  temp$span <- i
+  bstats <- rbind(bstats, temp)
+}
+
+
+
+xtabs(~(pred>0.5)+match+span, data=bstats)
+
+bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
+bullets <- bstats %>% group_by(bullet, span) %>% summarize(
+  n = n(),
+  pred = sum(pred > 0.5)
+)
+
+qplot(data=bullets, pred, reorder(bullet, pred/n),  colour=factor(span), shape = factor(span)) + 
+  theme_bw()  + ylab("") + 
+  scale_x_continuous("Number of correctly predicted land-to-land matches", 
+                     breaks = 3*0:4, limits=c(0,12)) 

### 2015-12-07 imaging-paper.Rnw
Redescribe the caption
Smoothing factor paragraph

@@
 
-Next, we fit a Loess regression to the data.
-Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
-The fit seems to do a reasonable job of capturing the structure of the image.
-Figure \ref{fig:loess_resid} shows the residuals from this fit.
-It is clear that there is a rather significant amount of autocorrelation, which make sense given the presence of striations and markings.
-Certainly, tuning the parameters for this regression, such as the Loess span, could prove to do an even better job at making the markings more clear.
-\hh{XXX did you follow up on that?
-What is the effect of using a different span?}
-
 \begin{figure}
@@
 \end{subfigure}
-%' \begin{subfigure}[t]{\linewidth}
-%' \caption{\label{fig:loess_step5}Step 5 of the Loess fitting algorithm. The plot shows the result after trimming values below the first minimum, and above the second minimum.}
-%' <<loess_step5, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-%' final.dat <- subset(curdata, y > curdata$y[min.left + 5] & y < curdata$y[min.right + (length(curdata$y) %/% 2) - 5])
-%' 
-%' qplot(data=final.dat, y, value) +
-%'     theme_bw() + xlim(c(0,2500))
-%' @
-%' \end{subfigure}
-\caption{Overview of all five steps of the loess algorithm to identify and remove big bullet strucure.}
-\end{figure}
+\caption{Overview of all seven steps of the smoothing algorithm to identify and remove grooves from the bullet images.}
+\end{figure}
+
+The ``smoothing factor" described in the algorithm represents the window size to use for a rolling average.
+Higher values of the smoothing factor therefore lead to more smoothed results.
+In the groove detection portion of this algorithm, we wish to remove most of the deviation but still maintain enough signal in the groove to detect it.
+Empirically, a value of 35 for the smoothing factor seems to work quite well.
+It is important to note that the smoothing pass is done twice.
+That is, the smoothed data is once again smoothed by computing a new rolling average with the same smoothing factor.
+This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass [ref].
+
+Next, we fit a Loess regression to the data.
+Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
+The fit seems to do a reasonable job of capturing the structure of the image.
+Figure \ref{fig:loess_resid} shows the residuals from this fit.
+It is clear that there is a rather significant amount of autocorrelation, which make sense given the presence of striations and markings.
+Certainly, tuning the parameters for this regression, such as the Loess span, could prove to do an even better job at making the markings more clear.
+\hh{XXX did you follow up on that?
+What is the effect of using a different span?}
 

### 2015-12-07 code/evaluation-hamby.R
random forest doesn't seem to help

@@ -198,3 +198,7 @@ qplot(data=bullets, pred, reorder(bullet, pred/n),  colour=factor(span), shape =
   theme_bw()  + ylab("") + 
   scale_x_continuous("Number of correctly predicted land-to-land matches", 
                      breaks = 3*0:4, limits=c(0,12)) 
+
+write.csv(bstats, "csvs/bullet-stats.csv", row.names=TRUE)
+
+

### 2015-12-07 imaging-paper.Rnw
Fix broken figure reference
random forest doesn't seem to help
land 1 of Bullet 1-1 exchanged for land 5

@@
 
-br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
+br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
 dbr111 <- fortify_x3p(br111)
@@
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
-      geom="line", size=I(1), colour=I("grey70")) + 
+      geom="line", size=I(.75), colour=I("grey70")) + 
   facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group = x), colour="black", size=1, 
-             data = filter(db1, part==1, x == 100)[,c("y", "resid", "x")]) +
+  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
+             data = filter(db1, part==5, x == 100)[,c("y", "resid", "x")]) +
   ylab("Residuals from circular fit") +
-  ggtitle("Part 1 of Bullet 1 in black")
-@
-\caption{Circle fit for each cross section of bullet 2, with part 1 of bullet 1 overlaid.\label{fig:circlefits}}%
+  ggtitle("Land 5 of Bullet 1 in black")
+@
+\caption{Circle fit for each cross section of bullet 2, with land 5 of bullet 1 overlaid.\label{fig:circlefits}}%
 \end{figure}
@@
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-1.x3p", sep = "/"))
+br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
 br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
@@
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
-my.loess$resid
-@
-\end{subfigure}
-\caption{Fit and residuals of a loess fit to land 1 of bullet 1 (Barrel 1).}
+my.loess$resid + ylab("Residual of Loess Fit")
+@
+\end{subfigure}
+\caption{Fit and residuals of a loess fit to land 5 of bullet 1 (Barrel 1).}
 \end{figure}
@@
 qplot(y, resid, data=melt.df, #colour=factor(x), 
-      geom="line", size=I(1), colour=I("grey70")) + 
+      geom="line", size=I(.75), colour=I("grey70")) + 
   facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=1) + ggtitle("Land 1 of Bullet 1 in black")
-@    
-\caption{Residuals of a loess fit to bullets 1 and 2 from barrel 1. Land 1 of bullet 1 is shown in black. }
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Land 5 of Bullet 1 in black")
+@    
+\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Land 5 of bullet 1 is shown in black. }
 \end{figure}
@@
 
+%' 
+%' <<rforest, echo=FALSE>>=
+%' library(randomForest)
+%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=1000)
+%' CCFs$forest <- predict(rtrees)
+%' importance(rtrees)
+%' @
+%'
+
 \paragraph{Acknowledgment} Thanks to the men and women behind the software R\cite{R}, and the authors of the R packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.

### 2015-12-07 imaging-paper.Rnw
couple of small edits

@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.}
+\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for land 5 of bullet 1 are plotted across different values of y.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 Next, we fit a Loess regression to the data.
-Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
+Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of land 5 of bullet 1.
 The fit seems to do a reasonable job of capturing the structure of the image.
@@
 
-We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from part 1 of bullet 1.
+We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from land 5 of bullet 1.
 This is shown in Figure~\ref{fig:loessfits}.
@@
 A total of 105 lands is being matched correctly.
-Interestingly, the number of consecutive matching striae is not featured in this evaluation.
-Instead, cross-correlation (ccf) between the crosscuts is featured heavily.
-Aside from cross correlation, the number of matches between crosscuts is involved in the decision.
+Interestingly, the number of consecutive matching striae is not featuring as an important variable in this evaluation.
+Instead, cross-correlation (ccf) between the crosscuts is very predominant.
+Aside from cross correlation, the number of matches (num.matches) and the cumulative height of the joint peaks (sumpeaks) between crosscuts is involved in the decision.
 Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
@@
 per <- rp1$frame$yval # predictive probability for each node in the tree
-prp(rp1, extra = 101, fallen.leaves=TRUE, box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
+prp(rp1, extra = 101,  box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
 

### 2015-12-08 imaging-paper.Rnw
land 5 not land 1 of bullet 1

@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_fit} Loess fit for land 1 of bullet 1.}
+\caption{\label{fig:loess_fit} Loess fit for land 5 of bullet 1.}
 <<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_resid} Residuals of Loess fit for land 1 of bullet 1.}
+\caption{\label{fig:loess_resid} Residuals of Loess fit for land 5 of bullet 1.}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
@@
 \includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
-\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land 5 of bullet 1 and land 1 of bullet 2.}}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land 5 of bullet 1 and land 1 of bullet 2.}}
 \end{figure}

### 2015-12-08 imaging-paper.Rnw
slowly getting rid of all the little things

@@
 
-options(replace.assign=TRUE,scipen=3)
+options(replace.assign=TRUE,scipen=3, digits=2)
 @
@@
 
-\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
+%\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
+\hh{In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$, which is consistent with a 9 mm bullet fired in a Ruger P-85. }
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
@@
 A 3D viewer, using Plotly \cite{plotly}, allows the user to rotate and zoom to examine surface features of the images.
-A screenshot of the application is available in Figure \ref{fig:app}.
+A screenshot of the application is shown in Figure~\ref{fig:app}.
 The application is available at the url \url{http://erichare.shinyapps.io/x3prproto}.
@@
 \includegraphics[width=\textwidth]{images/app.png}
-\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
+\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.
+\hh{The shot shows two lands from the Hamby study.
+Although the lands are a match, this is visually very hard to confirm.}}}
 \end{figure*}
@@
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
-    theme(legend.position = c(1,1), legend.justification=c(1,1),
+    theme(legend.position = c(1,1.2), legend.justification=c(1,1),
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
@@
 
+\subsection{Automatic Matching Algorithm}
 \hh{The automated matching routine is then as follows:
@@
 
-\hh{XXX needs more work: how does crosscut impact the results?  For the same bullet: how far apart can crosscuts be to still result in a match?}
-
 \section{Evaluation}
@@
 \item involve multiple crosscuts for each bullet in a match?
+\item assessing different crosscuts at the same bullet allows us to assess a land's quality in terms of how stable it is -- generally, break-off patterns are less stable. }
+
 \end{enumerate}

### 2015-12-08 code/checkBullets.R

@@ -10,34 +10,27 @@ knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
-if (!file.exists("data/crosscuts.csv")) {
-crosscuts1 <- sapply(knowns, function(x) {
+if (!file.exists("csvs/crosscuts.csv")) {
+crosscuts <- sapply(c(knowns, unknowns), function(x) {
   cat(x)
-  crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=12.5))
+  crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=25))
   cat(crosscut)
   cat("\n")
   crosscut
 })
 
-
-
-
-crosscuts2 <- sapply(unknowns, function(x) {
-  cat(x)
-  crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=12.5))
-  cat(crosscut)
-  cat("\n")
-  crosscut
-})
 
 dframe <- data.frame(
   path = c(knowns, unknowns),
-  cc = c(crosscuts1, crosscuts2)
+  cc = crosscuts
 )
 
 
-write.csv(dframe, file="data/crosscuts.csv", row.names=FALSE)
+write.csv(dframe, file="csvs/crosscuts-25.csv", row.names=FALSE)
 } 
+
+
+
 ccs <- read.csv("data/crosscuts.csv")
 
 

### 2015-12-08 code/complete-hamby.R

@@ -11,15 +11,15 @@ knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
 unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
-span <- 25
-dataStr <- sprintf("data-%d", span)
+span <- 35
+dataStr <- sprintf("data-%d-25", span) # using crosscuts-25.csv
 
 ###############
 # can we identify the barrels the unknown bullets came from?
 
 
 # match unknown land using crosscuts
-ccs <- read.csv(file.path(dataStr, "crosscuts.csv"))
+ccs <- read.csv("csvs/crosscuts-25.csv")
 
 for (j in 1:90) {
   
@@ -36,19 +36,7 @@ for (j in 1:90) {
 }
 
 
-for (j in 1:90) {
-  
-  reslist <- lapply(knowns, function(x) {
-    cat(x)
-    cat("\n")
-    #    browser()
-    cc1 <- ccs$cc[which(ccs$path == x)]
-    cc2 <- ccs$cc[which(ccs$path == unknowns[j])]
-    
-    bulletGetMaxCMSXXX(x, unknowns[j], cc1, cc2, span=25)
-  })
-  save(reslist, file=sprintf("data-25/unkn%d.RData", j))
-}
+
 
 
   

### 2015-12-08 imaging-paper.Rnw
adding - yet another bullet-stats csv file.

@@
 library(rpart.plot)
-CCFs <- read.csv("data-35/bullet-stats.csv")
+CCFs <- read.csv("data-35-25/bullet-stats.csv")
 includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y"))
@@
 
-CCFs$pred35 <- predict(rp1)
+CCFs$pred <- predict(rp1)
 @
@@
 library(xtable)
-xtable(xtabs(~I(pred35>0.5)+match, data=CCFs))
+xtable(xtabs(~I(pred>0.75)+match, data=CCFs))
 @
@@
 %' library(randomForest)
-%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=1000)
-%' CCFs$forest <- predict(rtrees)
+%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
+%' qplot(1:nrow(rtrees$err.rate), OOB, geom="line", data=data.frame(rtrees$err.rate)) + theme_bw() + ylim(c(0,NA))
+%' CCFs$forest <- predict(rtrees, type="prob")[,2]
 %' importance(rtrees)
+%' qplot(pred, forest, geom="jitter", data=CCFs, colour=match)
 %' @

### 2015-12-08 code/evaluation-hamby.R

@@ -1,5 +1,5 @@
-span <- 30
-dataStr <- sprintf("data-%d", span)
+span <- 35
+dataStr <- sprintf("data-%d-25", span)
 datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
 datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
 
@@ -76,7 +76,7 @@ write.csv(CCFs, file=file.path(dataStr, "bullet-stats.csv"), row.names=FALSE)
 
 idx <- which(CCFs$match & CCFs$cms <= 6)
 idx <- which(!CCFs$match & CCFs$ccf > .7)
-idx <- which(!CCFs$match & CCFs$cms == 11)
+idx <- which(!CCFs$match & CCFs$cms == 15)
 idx <- which(CCFs$cms == 11)
 
 for ( i in idx) {
@@ -105,7 +105,7 @@ CCFs <- read.csv(file.path(dataStr, "bullet-stats.csv"))
 
 qplot(factor(cms), data=CCFs)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(cms), fill=match), position="fill")
-sum(CCFs$cms >= 12)
+sum(CCFs$cms >= 16)
 ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match))
 ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=ccf, colour=match)) + facet_wrap(~match)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(non_cms), fill=match), position="fill")

### 2015-12-08 imaging-paper.Rnw
added points in the conclusion

@@
 \item the Hamby study only includes 35 bullets - this is not a particularly realistic scenario for automatic matching. Adding more non-matches into the mix is going to make things more realistic. 
-\item parameter optimization is going to drive some of the matches
+\item The matching algorithm is sensitive to the parameter choices made.
+The distance between crosscuts (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
+Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 57.8375$\mu m$ is used, corresponding to a window of 17 values to the left and the right of an observation).
+While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.
+
 \item involve multiple crosscuts for each bullet in a match?
 \item assessing different crosscuts at the same bullet allows us to assess a land's quality in terms of how stable it is -- generally, break-off patterns are less stable. }
-
+\item we get probabilities for matches. 
 \end{enumerate}

### 2015-12-08 imaging-paper.Rnw
Tukey reference, couple of axis labels

@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step2}Step 2 of the Loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
+% \begin{subfigure}[t]{\linewidth}
+% \caption{\label{fig:loess_step2}Step 2 of the Loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
 <<loess_step2, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 
-p2 <- qplot(1:length(smoothed), smoothed) + theme_bw()
-p2
-@
-\end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step3}Step 3 of the Loess fit algorithm. The smoothed height values are smoothed one more time with the same smoothing factor.}
+p2 <- qplot(1.5625*(1:length(smoothed)), smoothed) + theme_bw() + xlab("y")
+#p2
+@
+%\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_step3}Steps 2 \& 3 of the Loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35.}
 <<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-p3 <- qplot(1:length(smoothed_truefalse), smoothed_truefalse) + theme_bw()
+xloc=250
+p3 <- ggplot() +
+  annotate("rect",xmin=xloc-17*1.5625, xmax=xloc+18*1.5625, ymin=45, ymax=180, fill=alpha("orange", alpha=0.5)) + 
+  geom_point(aes(x=(1:length(smoothed_truefalse))*1.5625, y=smoothed_truefalse)) + theme_bw() +
+    geom_vline(xintercept=xloc, colour="grey50") + xlab("y") + ylab("")
+
 p3
@@
         geom_vline(xintercept = subdata$y[peak_ind2], colour = "red") +
-        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue")
+        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue") +
+      xlab("y") + ylab("")
 
@@
 In the groove detection portion of this algorithm, we wish to remove most of the deviation but still maintain enough signal in the groove to detect it.
-Empirically, a value of 35 for the smoothing factor seems to work quite well.
+Empirically, a window of 35 observations for the smoothing factor seems to work quite well.
+%currently a window of 57.8375$\mu m$ is used, corresponding to a window of 17 values to the left and the right of an observation
 It is important to note that the smoothing pass is done twice.
 That is, the smoothed data is once again smoothed by computing a new rolling average with the same smoothing factor.
-This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass [ref].
+This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass \cite{tukey:1977}.
+\hh{Did Tukey say, what the goal of the twicing was?
+- also, do you have a page number for the twicing?}
 
@@
 The distance between crosscuts (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
-Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 57.8375$\mu m$ is used, corresponding to a window of 17 values to the left and the right of an observation).
+Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 35 observations is used).
 While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.
@@
 \item involve multiple crosscuts for each bullet in a match?
-\item assessing different crosscuts at the same bullet allows us to assess a land's quality in terms of how stable it is -- generally, break-off patterns are less stable. }
+\item assessing different crosscuts at the same bullet allows us to assess a land's quality in terms of how stable it is -- generally, break-off patterns are less stable. 
 \item we get probabilities for matches. 

### 2015-12-08 imaging-paper.Rnw
collapsed steps 2 and 3 of the loess fit into one

@@
     \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step1}).
-    \item For each $y$ value, compute a rolling average with a particular \textbf{smoothing factor} to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step2}).
-    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor above. (Figure~\ref{fig:loess_step3}).
+    \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \textbf{smoothing factor}. (Figure~\ref{fig:loess_step3}).
+%    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor as above. (Figure~\ref{fig:loess_step3}).
     \item Determine the peak of the left groove by finding the first doubled-smoothed value $y_i$ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$.
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step3}Steps 2 \& 3 of the Loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35.}
+\caption{\label{fig:loess_step3}Step 2 of the Loess fit algorithm. The  height values are smoothed twice with a smoothing factor of 35.}
 <<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 p3 <- ggplot() +
-  annotate("rect",xmin=xloc-17*1.5625, xmax=xloc+18*1.5625, ymin=45, ymax=180, fill=alpha("orange", alpha=0.5)) + 
+  annotate("rect",xmin=xloc-17*1.5625, xmax=xloc+18*1.5625, ymin=80, ymax=200, fill=alpha("orange", alpha=0.5)) + 
   geom_point(aes(x=(1:length(smoothed_truefalse))*1.5625, y=smoothed_truefalse)) + theme_bw() +
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step47}Steps 4 - 7 of the Loess fitting algorithm.
+\caption{\label{fig:loess_step47}Steps 3 -- 6 of the Loess fitting algorithm.
 After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
@@
 \end{subfigure}
-\caption{Overview of all seven steps of the smoothing algorithm to identify and remove grooves from the bullet images.}
+\caption{Overview of all six steps of the smoothing algorithm to identify and remove grooves from the bullet images.}
 \end{figure}
@@
 This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass \cite{tukey:1977}.
-\hh{Did Tukey say, what the goal of the twicing was?
+\hh{XXX Did Tukey say, what the goal of the twicing was?
 - also, do you have a page number for the twicing?}
@@
 Next, we fit a Loess regression to the data.
-Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of land 5 of bullet 1.
-The fit seems to do a reasonable job of capturing the structure of the image.
+Figure \ref{fig:loess_fit} provides a look at the Loess fit in blue, overlaid against the processed image of land 5 of bullet 1.
+The fit seems to do a reasonable job of capturing the structure.
 Figure \ref{fig:loess_resid} shows the residuals from this fit.
-It is clear that there is a rather significant amount of autocorrelation, which make sense given the presence of striations and markings.
+\hh{XXX what purpose does the next paragraph have?
+I think you just want end this section, but it brings up two new ideas: autocorrelation and dependency on the span.
+The span obviously doesn't get any resolution ...
+so it's just muddying the waters at the moment ...
+XXX}
+It is clear that there is a rather significant amount of autocorrelation, which makes sense given the presence of striations and markings.
 Certainly, tuning the parameters for this regression, such as the Loess span, could prove to do an even better job at making the markings more clear.

### 2015-12-08 imaging-paper.Rnw
Some small typos

--- 2015-12-08 imaging-paper.Rnw
@@
 
-options(replace.assign=TRUE,scipen=3, digits=2)
+options(replace.assign=TRUE,scipen=3)
 @
@@
 
-%\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
+\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
-\hh{In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$, which is consistent with a 9 mm bullet fired in a Ruger P-85. }
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for land 5 of bullet 1 are plotted across different values of y.}
+\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 Next, we fit a Loess regression to the data.
-Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of land 5 of bullet 1.
+Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
 The fit seems to do a reasonable job of capturing the structure of the image.
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_fit} Loess fit for land 5 of bullet 1.}
+\caption{\label{fig:loess_fit} Loess fit for land 1 of bullet 1.}
 <<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_resid} Residuals of Loess fit for land 5 of bullet 1.}
+\caption{\label{fig:loess_resid} Residuals of Loess fit for land 1 of bullet 1.}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
@@
 
-We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from land 5 of bullet 1.
+We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from part 1 of bullet 1.
 This is shown in Figure~\ref{fig:loessfits}.
@@
 A 3D viewer, using Plotly \cite{plotly}, allows the user to rotate and zoom to examine surface features of the images.
-A screenshot of the application is shown in Figure~\ref{fig:app}.
+A screenshot of the application is available in Figure \ref{fig:app}.
 The application is available at the url \url{http://erichare.shinyapps.io/x3prproto}.
@@
 \includegraphics[width=\textwidth]{images/app.png}
-\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.
-\hh{The shot shows two lands from the Hamby study.
-Although the lands are a match, this is visually very hard to confirm.}}}
+\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
 \end{figure*}
@@
 Clicking and dragging anywhere on the plot will allow rotation.
-Scrolling up and down allows the user to zoom in and out respectively.
+Scrolling up and down allows the user to zoom out and and respectively.
 There are also a set of controls available at the top right of the page, which will allow the selection of some other rotation methods, taking of screenshots, and reseting the camera to the default view.
@@
 
-The application includes a button which allows the automatic creation of a residual plot.
+The application includes a button which allows the automatic creation of a residual plot from the Loess fits of two bullets.
 An x coordinate value can be selected, and a cross-section of the two bullet at the x value will be taken.
@@
 \includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
-\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land 5 of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land 5 of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
 \end{figure}
@@
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
-    theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+    theme(legend.position = c(1,1), legend.justification=c(1,1),
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
@@
 
-\subsection{Automatic Matching Algorithm}
 \hh{The automated matching routine is then as follows:
@@
 
+\hh{XXX needs more work: how does crosscut impact the results?  For the same bullet: how far apart can crosscuts be to still result in a match?}
+
 \section{Evaluation}
@@
 A total of 105 lands is being matched correctly.
-Interestingly, the number of consecutive matching striae is not featuring as an important variable in this evaluation.
-Instead, cross-correlation (ccf) between the crosscuts is very predominant.
-Aside from cross correlation, the number of matches (num.matches) and the cumulative height of the joint peaks (sumpeaks) between crosscuts is involved in the decision.
+Interestingly, the number of consecutive matching striae is not featured in this evaluation.
+Instead, cross-correlation (ccf) between the crosscuts is featured heavily.
+Aside from cross correlation, the number of matches between crosscuts is involved in the decision.
 Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
@@
 library(rpart.plot)
-CCFs <- read.csv("data-35-25/bullet-stats.csv")
+CCFs <- read.csv("data-35/bullet-stats.csv")
 includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y"))
@@
 per <- rp1$frame$yval # predictive probability for each node in the tree
-prp(rp1, extra = 101,  box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
-
-CCFs$pred <- predict(rp1)
+prp(rp1, extra = 101, fallen.leaves=TRUE, box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
+
+CCFs$pred35 <- predict(rp1)
 @
@@
 library(xtable)
-xtable(xtabs(~I(pred>0.75)+match, data=CCFs))
+xtable(xtabs(~I(pred35>0.5)+match, data=CCFs))
 @
@@
 %' library(randomForest)
-%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
-%' qplot(1:nrow(rtrees$err.rate), OOB, geom="line", data=data.frame(rtrees$err.rate)) + theme_bw() + ylim(c(0,NA))
-%' CCFs$forest <- predict(rtrees, type="prob")[,2]
+%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=1000)
+%' CCFs$forest <- predict(rtrees)
 %' importance(rtrees)
-%' qplot(pred, forest, geom="jitter", data=CCFs, colour=match)
 %' @
@@
 \item the Hamby study only includes 35 bullets - this is not a particularly realistic scenario for automatic matching. Adding more non-matches into the mix is going to make things more realistic. 
-\item The matching algorithm is sensitive to the parameter choices made.
-The distance between crosscuts (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
-Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 57.8375$\mu m$ is used, corresponding to a window of 17 values to the left and the right of an observation).
-While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.
-
+\item parameter optimization is going to drive some of the matches
 \item involve multiple crosscuts for each bullet in a match?
-\item assessing different crosscuts at the same bullet allows us to assess a land's quality in terms of how stable it is -- generally, break-off patterns are less stable. }
-\item we get probabilities for matches. 
 \end{enumerate}

### 2015-12-08 imaging-paper.Rnw
Some questions and notes
more axis labels, some language

@@
 \begin{enumerate}
-\item Description of the problem: Standard operating practice, 2009 NAS report
-\item Description of the Results
-\item Walkthrough of the process
+\item DONE Description of the problem: Standard operating practice, 2009 NAS report
+\item DONE Description of the Results
+\item DONE Walkthrough of the process
 \item \hh{XXX Alternative approaches: IBIS database and the NRC report\cite{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
@@
 \section{Problem Statement}
+\eh{XXX do we still need this section?}
 
@@
 @
-\caption{Circle fit for each cross section of bullet 2, with land 5 of bullet 1 overlaid.\label{fig:circlefits}}%
+\caption{Circle fit for each cross section of bullet 2, with land 5 of bullet 1 overlaid.\label{fig:circlefits} \hh{Land 5 of bullet 1 matches best with the first land of the second bullet.}}%
 \end{figure}
@@
 \subsection{Loess Fit}
-
-To improve upon the previous algorithm, the first step must be to eliminate the grooves from the images under the assumption that they do not contain relevant information for determining a match.
+In an approach to improve upon the previous algorithm, we first have to eliminate 
+%To improve upon the previous algorithm, the first step must be to eliminate
+the grooves from the images under the assumption that they do not contain relevant information for determining a match.
 Fortunately, the location and appearance of the grooves in these 2d plots is quite consistent.
-By noting that height values beginning at $y = 0$ reach a maximum around the peak of the groove, we can then follow the descent of the height values to valley of the groove.
-This marks the point at which we should trim the image.
-We repeat a similar procedure for the right-hand groove by starting at the maximum y value and working backwards.
+Surface measurements %beginning at $y = 0$ 
+reach local maxima around the peak of the groove at either end of the range of $y$, we can then follow the descent of the surface measurements inwards to the valley of the groove. 
+This marks the point at which we should trim the image. %We repeat a similar procedure for the right-hand groove by starting at the maximum y value and working backwards. 
 The procedure can be described as follows:
@@
 \begin{enumerate}
-    \item Fix an $x$ coordinate value (Figure~\ref{fig:loess_step1}).
-    \item For each $y$ value, compute a rolling average with a particular \textbf{smoothing factor} to smooth out any deviations occurring near the minima. (Figure~\ref{fig:loess_step2}).
-    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor above. (Figure~\ref{fig:loess_step3}).
-    \item Determine the peak of the left groove by finding the first doubled-smoothed value $y_i$ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$.
-Call this peak $p_{left}$.
-(Figure~\ref{fig:loess_step47}).
-    \item Determine the valley of the left groove by finding the first doubled-smoothed $y_j > p_{left}$ such that $y_j < y_{j - 1}$ and $y_j < y_{j + 1}$. Call this valley $v_{left}$.
+    \item Fix an $x$ value (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
+    \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \textbf{smoothing factor}.
+(Figure~\ref{fig:loess_step3}, smoothing factor of 35, corresponding to 55$\mu m$).
+%    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor as above. (Figure~\ref{fig:loess_step3}).
+    \item Determine the location of the peak of the left groove by finding the first double-smoothed value $y_i$ that is the maximum within its smoothing window (e.g.\ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$, $i = 1, ..., 17$).
+We call the location of this peak $p_{left}$ (see Figure~\ref{fig:loess_step47}).
+    %\hh{XXX By peak, do you mean the location of the peak or its height or both?} XXX never mind, found the answer below
+    \item Similarly, determine the location of the valley of the left groove by finding the first double-smoothed $y_j$ that is the minimum within its smoothing window.
+Call the location of this valley $v_{left}$.
     \item Reverse the order of the $y$ values and repeat the previous two steps to find the peak and valley of the right groove, $(p_{right}, v_{right})$.
-    \item Remove all $y_i < v_{left}$ and $y_i > v_{right}$. (Figure \ref{fig:loess_step47}).
+    \item Trim the surface measurements to values within the two grooves (i.e.\ remove all records with $y_i < v_{left}$ and $y_i > v_{right}$). (Figure \ref{fig:loess_step47}).
 \end{enumerate}
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm. For a fixed crosscut (x held constant at 243.75) height values for part 1 of bullet 1 are plotted across different values of y.}
+\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm.
+For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for land 5 of bullet 1 are plotted across different values of $y$.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 
-p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw()
+p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
 p1
@@
 
-p2 <- qplot(1:length(smoothed), smoothed) + theme_bw()
-p2
-@
-\end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step3}Step 3 of the Loess fit algorithm. The smoothed height values are smoothed one more time with the same smoothing factor.}
+p2 <- qplot(1.5625*(1:length(smoothed)), smoothed) + theme_bw() + xlab("y") +
+  ylab("Surface measurements")
+#p2
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:loess_step3}Step 2 of the Loess fit algorithm.
+The  height values are smoothed twice with a smoothing factor of 35.
+The orange rectangle shows an example of the smoothing window.
+Groove and peak will be detected, if they are not within the same window.}
 <<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-p3 <- qplot(1:length(smoothed_truefalse), smoothed_truefalse) + theme_bw()
+xloc=250
+p3 <- ggplot() +
+  annotate("rect",xmin=xloc-17*1.5625, xmax=xloc+18*1.5625, ymin=80, ymax=200, fill=alpha("orange", alpha=0.5)) + 
+  geom_point(aes(x=(1:length(smoothed_truefalse))*1.5625, y=smoothed_truefalse)) + theme_bw() +
+    geom_vline(xintercept=xloc, colour="grey50") + xlab("y") + ylab("Surface measurements")
+
 p3
@@
         geom_vline(xintercept = subdata$y[peak_ind2], colour = "red") +
-        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue")
+        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue") +
+      xlab("y") + ylab("Surface measurements")
 
@@
 
-my.loess$fitted
+my.loess$fitted + ylab("Surface measurements")
 @
@@
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
-my.loess$resid + ylab("Residual of Loess Fit")
+my.loess$resid + ylab("Residuals of Loess Fit")
 @
@@
 
-We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from part 1 of bullet 1.
+We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from land 5 of bullet 1.
 This is shown in Figure~\ref{fig:loessfits}.
+\hh{The first land of bullet 2 is still the best  match for land 5, but now the details of the match are more pronounced.}
 
@@
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Land 5 of Bullet 1 in black")
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Land 5 of Bullet 1 in black") + ylab("Residuals of Loess Fit")
 @    
@@
 \end{subfigure}
-\caption{\label{fig:cross100} Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
+\caption{\label{fig:cross100}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
 \end{figure}
@@
 \begin{enumerate}
-\item Smooth the two crosscuts by a loess with a very small span. 
-\item Use cross-correlation to find the best alignment along the crosscuts: Shift one of the crosscuts by the lag indicated by the cross-correlation function.
-\item Using a rolling average, identify peaks and grooves for each of the crosscuts. We define a region around the extrema on each side as one third of the distance to the next extrema. 
+\item Smooth the two crosscuts by a loess with a very small span (see Figure~\ref{fig:smooth}). 
+\item Use cross-correlation to find the best alignment along the crosscuts: Shift one of the crosscuts by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
+\item Using a rolling average, identify peaks and grooves for each of the crosscuts.
+We define a region around the extrema on each side as one third of the distance to the next extrema (see Figure~\ref{fig:smoothcutb}).
 %Based on a starting threshold level of $\theta$, identify areas of the crosscuts whose absolute value are above the thresholds.
@@
 If only one crosscut exceeds the threshold, or if the two crosscuts move into opposite directions over the threshold, the area is a non-match.
-\item Match grooves and peaks between crosscuts. \hh{XXX  needs some more details.} 
+\item Match grooves and peaks between crosscuts: based on the regions around the extrema as defined above, we identify joint regions between two crosscuts as those areas, in which at least two of the individual regions touch.
+A joint region is defined as the smallest interval that covers all of the touching regions.
+A joint region is called a match between the crosscuts, if all of the crosscut-specific regions are of the same type of extrema, i.e.\ either all peaks or all valleys.
+In Figure~\ref{fig:smoothcutc} all matches are colored corresponding to their type of extrema.
+Non-matching regions are left grey.
 \item Identify the maximal number of CMS (consecutive matching striae) between the two crosscuts by counting consecutive matching lines. 
@@
 library(rpart.plot)
-CCFs <- read.csv("data-35/bullet-stats.csv")
-includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y"))
+CCFs <- read.csv("data-35-25/bullet-stats.csv")
+includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest"))
 rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
@@
 library(xtable)
-xtable(xtabs(~I(pred35>0.5)+match, data=CCFs))
+xtable(xtabs(~I(pred>0.5)+match, data=CCFs))
 @

### 2015-12-08 imaging-paper.Rnw
black/grey for some of the lines

@@
 library(rpart.plot)
+
+cols = c(alpha("grey60", alpha=0.6), alpha("black", 0.5))
 
@@
 qplot(y, resid, group=bullet, colour=bullet, data=lof, 
-      geom="line", alpha=I(0.6)) + 
-  scale_colour_brewer("", palette="Set1") + theme_bw() +
-  theme(legend.position = c(.9, .2))
+      geom="line") + 
+  scale_colour_manual("", values=cols) + theme_bw() +
+  theme(legend.position = c(.9, .2),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylab("Residuals of Loess fit")
 @
@@
 qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
-      geom="line", alpha=I(0.6)) + 
-  scale_colour_brewer(palette="Set1") + theme_bw() +
-  theme(legend.position="none")
+      geom="line") + 
+  scale_colour_manual("", values=cols) + theme_bw() +
+  theme(legend.position="none") + ylab("Residuals of Loess fit")
 @
@@
 <<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
-qplot(data=lofX, y, resid, colour=bullet, size=I(0.75), geom="line", group=bullet) +
-  geom_line(aes(y=l30), colour="grey30", size=0.5) +
+qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
+  geom_line(aes(y=l30), colour="grey70", size=0.5) +
   facet_grid(bullet~.) +
-  scale_colour_brewer("", palette="Set1") +
-  theme_bw() + ylab("Residuals of Loess Fit") + 
+#  scale_colour_manual("", values=cols) +
+  theme_bw() + ylab("") + 
   theme(legend.position="none")
@@
 \caption{Using a rolling median peaks and grooves are identified for each crosscut. \label{fig:smoothcutb}}{%
-<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=4, out.width='\\textwidth', warning=FALSE>>=
+<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
+#images <- file.path(datadir, dir(datadir))
   lof <- processBullets(paths = images[c(5,7)], x = 100)
@@
   peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 30)
-
-  library(RColorBrewer)
-  col <- brewer.pal(3, "Set1")
-  p1 <- ggplot() + theme_bw() +
-    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, data=peaks1$lines,  alpha=0.2) +
-    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks1$lines) + scale_colour_brewer(palette="Set2") + 
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+  peaks <- rbind(peaks1$lines, peaks2$lines)
+
+  ggplot() + theme_bw() +
+    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
+              data=peaks,  alpha=0.2) +
+    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks) +
+    scale_colour_brewer(palette="Set2") + 
     scale_fill_brewer(palette="Set2") +
     theme(legend.position="none") + 
-    geom_line(aes(x=y, y=l30, group=bullet), data=subset(lofX, bullet==b12[1])) +
-  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
-    ggtitle(b12[1])+ xlim(c(0, 2000))
-
-  p2 <- ggplot() + theme_bw() +
-    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, data=peaks2$lines,  alpha=0.2) +
-    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks2$lines) + scale_colour_brewer(palette="Set2") + 
-    scale_fill_brewer(palette="Set2") +
-    theme(legend.position="none") + 
-    geom_line(aes(x=y, y=l30, group=bullet), data=subset(lofX, bullet==b12[2])) +
-  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
-    ggtitle(b12[2]) + xlim(c(0, 2000))
-grid.arrange(p1, p2, ncol=1)
+    facet_grid(bullet~.) +
+    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
+  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
 @

### 2015-12-08 imaging-paper.Rnw
adding linetype to bullets

@@
   ylim(c(-6,6.5)) + theme_bw() +
-  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, colour=bullet)) +
-  scale_colour_brewer("", palette="Set1") +
+  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
+#  scale_colour_brewer("", palette="Set1") +
+  scale_linetype_discrete("") +
+  scale_colour_manual("", values=cols) +
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
-    theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
         legend.background = element_rect(fill=alpha('white', 0.4))) + 

### 2015-12-08 imaging-paper.Rnw
couple of comments

--- 2015-12-08 imaging-paper.Rnw
@@
 That is, the smoothed data is once again smoothed by computing a new rolling average with the same smoothing factor.
-This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass [ref].
+This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass \cite{tukey:1977}.
+\hh{XXX Did Tukey say, what the goal of the twicing was?
+- also, do you have a page number for the twicing?}
 
@@
 Figure \ref{fig:loess_resid} shows the residuals from this fit.
+\hh{XXX what purpose does the next paragraph have?
+I think you just want end this section, but it brings up two new ideas: autocorrelation and dependency on the span.
+The span obviously doesn't get any resolution ...
+so it's just muddying the waters at the moment ...
+XXX}
 It is clear that there is a rather significant amount of autocorrelation, which make sense given the presence of striations and markings.
@@
   ylim(c(-6,6.5)) + theme_bw() +
-  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, colour=bullet)) +
-  scale_colour_brewer("", palette="Set1") +
+  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
+  scale_linetype_discrete("") +
+  scale_colour_manual("", values=cols) +
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
-    theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
@@
   geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, colour = bullet),   data = res$bullets) +
+  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets) +
   scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
+  scale_linetype_discrete("") +
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
@@
   geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
+    ylab("")
 })
@@
   geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, colour = bullet),   data = res$bullets) +
-  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
+  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
+  scale_linetype_discrete("") +
+#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +

### 2015-12-09 imaging-paper.Rnw
couple more figures adjusted

@@
   geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets) +
+  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
   scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
@@
     geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=6, data=lines, alpha=0.2) +
-    geom_line(aes(x = y, y = l30, colour = bullet),  data = lofX) +
+    geom_line(aes(x = y, y = l30, linetype = bullet),  data = lofX, alpha=0.6) +
 #    geom_hline(yintercept = threshold) +
 #    geom_hline(yintercept = - threshold) +
-    scale_colour_brewer("",palette="Set1") +
+#    scale_colour_brewer("",palette="Set1") +
+    scale_linetype_discrete("") +
     scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
@@
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
-load("data-35/unkn27.RData")
+load("data-35-25/unkn27.RData")
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
@@
   geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
-  geom_line(aes(x = y, y = l30, colour = bullet),  data = res$bullets) +
+  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
 #  geom_hline(yintercept = res$threshold) +
 #  geom_hline(yintercept = - res$threshold) +
-  scale_colour_brewer("",palette="Set1") +
+#  scale_colour_brewer("",palette="Set1") +
+  scale_linetype_discrete("") +
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
@@
 Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
-All 23 pairs of lands with at least twelve CMS in common are matches.
+All 29 pairs of lands with at least sixteen CMS in common are matches.
 For smaller number of CMS, the percentage of false positives quickly increases.
@@
 <<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
-bstats <- read.csv("data-35/bullet-stats.csv")
+bstats <- read.csv("data-35-25/bullet-stats.csv")
 ggplot(data=bstats) + 
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{All land-to-land comparisons a maximum number of CMS of twelve or more are matches.}
+\caption{All land-to-land comparisons a maximum number of CMS of sixteen or more are matches.}
 <<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=

### 2015-12-09 imaging-paper.Rnw
Couple of little additions

@@
 This process has broad applicability in terms of the conviction of criminals in the United States criminal justice system.
-Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone some more significant scrutiny.
+Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny.
 In 2005, in \emph{United States vs.
 Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar.
-Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence.
+Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence \eh{XXX we probably need a reference here}.
 

### 2015-12-09 imaging-paper.Rnw
Rewrite the end of the intro + beginning of methodology

@@
 
-\section{Problem Statement}
-\eh{XXX do we still need this section?}
-
-\section{Methodology}
-Throughout this section, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study \cite{hamby:2009}.
-Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
-
 Rifling in the barrel creates striation marks on the bullet during the firing process (and grooves).
@@
 
+The primary goal in this type of forensic analysis is to identify markings in an attempt to match two bullets.
+Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help us determine whether two bullets are a match.
+However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
+We've produced a framework which, in software, allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier for the lay person.
+
+\section{Methodology}
+Throughout this paper, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study \cite{hamby:2009}.
+Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
+
 3D topographical images of each bullet were taken using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
 x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
-%The x3p format was developed as a means for forensic examiners to easily exchange images in a standard format. 
 x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government Firearm Forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
@@
 The direction of the rifling is a class characteristic.
-
-The primary goal in this type of forensic analysis is to identify markings in an attempt to match two bullets.
-Gun barrels should produce similar markings on bullets fired consecutively, and the careful analysis of these surface images can help us determine whether two bullets are a match.
-However, this process has traditionally required forensic experts to perform the analysis, also required physical examination of the bullets.
-We've produced a framework which, in software, allows for the analysis of the surface topologies, and also transcribes the markings into a 2D plotting framework which we believe makes identification of matches easier for the lay person.
 

### 2015-12-09 imaging-paper.Rnw
Linetype for circle fits

@@
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
-      geom="line", size=I(.75), colour=I("grey70")) + 
+      geom="line", size=I(.75), linetype=I(3)) + 
   facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
+  geom_line(aes(y, resid, group = x), linetype=1, size=.75,
              data = filter(db1, part==5, x == 100)[,c("y", "resid", "x")]) +

### 2015-12-09 imaging-paper.Rnw
Do the same for figure 7

@@
 qplot(y, resid, data=melt.df, #colour=factor(x), 
-      geom="line", size=I(.75), colour=I("grey70")) + 
+      geom="line", size=I(.75), linetype=I(3)) + 
   facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Land 5 of Bullet 1 in black") + ylab("Residuals of Loess Fit")
+  geom_line(data = actual.df, aes(y, resid), linetype=1, size=.75) + ggtitle("Land 5 of Bullet 1 in black") + ylab("Residuals of Loess Fit")
 @    

### 2015-12-09 imaging-paper.Rnw
conclusions

@@
 per <- rp1$frame$yval # predictive probability for each node in the tree
-prp(rp1, extra = 101, fallen.leaves=TRUE, box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
+prp(rp1, extra = 101, box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
 
@@
 \item the Hamby study only includes 35 bullets - this is not a particularly realistic scenario for automatic matching. Adding more non-matches into the mix is going to make things more realistic. 
-\item parameter optimization is going to drive some of the matches
+\item The matching algorithm is sensitive to the parameter choices made.
+The distance between crosscuts (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
+Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 57.8375$\mu m$ is used, corresponding to a window of 17 values to the left and the right of an observation).
+While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.
 \item involve multiple crosscuts for each bullet in a match?
+\item we get probabilities for matches.
+
 \end{enumerate}

### 2015-12-09 imaging-paper.Rnw
Revert the change to linetypes for figures 4 and 7

@@
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
-      geom="line", size=I(.75), linetype=I(3)) + 
+      geom="line", size=I(.75), colour=I("grey70")) + 
   facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group = x), linetype=1, size=.75,
+  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
              data = filter(db1, part==5, x == 100)[,c("y", "resid", "x")]) +
@@
 qplot(y, resid, data=melt.df, #colour=factor(x), 
-      geom="line", size=I(.75), linetype=I(3)) + 
+      geom="line", size=I(.75), colour=I("grey70")) + 
   facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), linetype=1, size=.75) + ggtitle("Land 5 of Bullet 1 in black") + ylab("Residuals of Loess Fit")
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Land 5 of Bullet 1 in black") + ylab("Residuals of Loess Fit")
 @    

### 2015-12-09 imaging-paper.Rnw
caught another land/part

@@
 
-\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
+\hh{In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$, which is consistent with a 9 mm bullet fired in a Ruger P-85. }
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
 For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.
+
+-%\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
 
@@
 %' library(randomForest)
-%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=1000)
+%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
 %' CCFs$forest <- predict(rtrees)

### 2015-12-09 imaging-paper.Rnw
update on numbers in results
expanded on random forest discussion

@@
 A cross-correlation of above 0.75 in the Hamby study is indicative of a match.
-There are 67 land-land comparisons with a cross-correlation of above 0.75.
+There are 97 land-land comparisons with a cross-correlation of above 0.75.
 \begin{figure}
@@
 CCFs <- read.csv("data-35-25/bullet-stats.csv")
-includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest"))
+includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
 rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
@@
   n = n(),
-  pred = sum(pred35 > 0.5)
+  pred = sum(pred > 0.5),
+  forest = sum(forest > 0.5)
 )
@@
 For each bullet at least one land is correctly matched.
-Unknown bullet B is the hardest for the algorithm to match.}
-\end{figure}
-
-%' 
-%' <<rforest, echo=FALSE>>=
-%' library(randomForest)
-%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
-%' CCFs$forest <- predict(rtrees)
-%' importance(rtrees)
-%' @
-%'
+Unknown bullets B and M are the hardest for the algorithm to match.}
+\end{figure}
+
+<<rforest, echo=FALSE, message=FALSE>>=
+library(randomForest)
+library(reshape2)
+rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
+
+errors <- data.frame(rtrees$err.rate)
+errors$id <- 1:nrow(errors)
+CCFs$forest <- predict(rtrees, type="prob")[,2]
+@
+
+Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate.
+
+\begin{figure}
+<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='\\linewidth'>>=
+qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
+@
+\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
+\end{figure}
+
+\begin{table}
+\centering
+<<importance, echo=FALSE, results='asis'>>=
+imp <- data.frame(importance(rtrees))
+imp$Variable <- row.names(imp)
+imp <- imp[order(-imp$MeanDecreaseGini),]
+names(imp)[1] <- "Mean Decrease Gini"
+print(xtable(imp[,2:1]), include.rownames=FALSE, floating = FALSE)
+@
+\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches. }
+\end{table}
+
+\begin{figure}
+<<tree:forest, echo=FALSE, fig.width=5, fig.height=6, out.width='\\linewidth'>>=
+qplot(pred, forest, geom="point", data=CCFs, colour=match) + theme_bw() +
+  theme(legend.position="bottom")
+@
+\caption{\label{fig:tree-forest}Scatterplot of prediction results from the tree (horizontal) and the forest (vertical).
+Each point corresponds to a land-to-land comparison.
+Matches are indicated in blue.}
+\end{figure}
 

### 2015-12-09 imaging-paper.Rnw
discussion of automated matching

@@
 Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence \eh{XXX we probably need a reference here}.
+\hh{XXX have you tried the Exoneration Registry at michigan? http://www.law.umich.edu/special/exoneration/Pages/about.aspx}
 
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step2}Step 2 of the Loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
+%\begin{subfigure}[t]{\linewidth}
+% \caption{\label{fig:loess_step2}Step 2 of the Loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
 <<loess_step2, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 @
-\end{subfigure}
+%\end{subfigure}
 \begin{subfigure}[t]{\linewidth}
@@
 
-\section{Application}
+\section{Web Application}
 We have implemented a web interface to automate many of the steps outlined in the previous section.
@@
 \section{Approach to automatic matching}
-\hh{XXX all tentative}
 Applying the  loess fit to a range of different surface cross cuts (see Figures~\ref{fig:manualmatch} and Figure~\ref{fig:manualmatch}, with cross cuts from 50 to 150) shows the striation marks across two bullets.
@@
 Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$.
-With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 15 to  3 or fewer at $x = 300\mu m$ and higher.
+With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
 }
@@
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6)) +
+  ylim(c(-6,6)) + ylab("") +
   geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
@@
 
+\subsection{Automated matching algorithm}
 \hh{The automated matching routine is then as follows:
@@
 We define a region around the extrema on each side as one third of the distance to the next extrema (see Figure~\ref{fig:smoothcutb}).
-%Based on a starting threshold level of $\theta$, identify areas of the crosscuts whose absolute value are above the thresholds.
-These areas are regions of potential matches between the crosscuts.
-If all of the values of the crosscuts stay on the same side, the area is called a matching line.
-If only one crosscut exceeds the threshold, or if the two crosscuts move into opposite directions over the threshold, the area is a non-match.
 \item Match grooves and peaks between crosscuts: based on the regions around the extrema as defined above, we identify joint regions between two crosscuts as those areas, in which at least two of the individual regions touch.
@@
 A joint region is called a match between the crosscuts, if all of the crosscut-specific regions are of the same type of extrema, i.e.\ either all peaks or all valleys.
-In Figure~\ref{fig:smoothcutc} all matches are colored corresponding to their type of extrema.
+In Figure~\ref{fig:match} all matches are colored corresponding to their type of extrema.
 Non-matching regions are left grey.
@@
 @
-\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to the first land of the second bullet shot from barrel 2 (in blue).
-The largest number of CMS is eleven, observed with land six of the second bullet.
+\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to the first land of the second bullet shot from barrel 2.
+The largest number of CMS is eleven, observed with land six of the first bullet.
 }
@@
 \section{Evaluation}
-
-\hh{We can use the Hamby study to run all unkowns against the known bullets. When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}.}
+Using the James Hamby study, we run all unkown bullets against all known bullets for matches.
+The algorithm is using land-to-land matches, i.e.
+we are running $15 \times 6 = 90$ lands from unkown bullets against $2 \times 10 \times 6 = 120$ lands from known bullets, yielding a total of $10,800$ comparisons with 180 known land-to-land matches.
+ When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land F-3 and all 120 known lands.
+Two lands show a high CMS.
+These correspond to the known matches, shown in Figures~\ref{fig:hamby-perfect}b and c.
+ Unfortunately, not all results are as clear cut. 
+It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches to get an overview of what we can reasonably expect from an automated match.
 
@@
 \end{figure}
-\hh{XXX a lot more about the Hamby study.
-Fifteen unknown bullets (i.e.
-90 lands) compared to two times ten known bullets (i.e.
-120 lands) makes $90 \times 120 = 10,800$ land to land comparisons, out of which 180 are actual matches.
-It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches.
-}
+
 Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
@@
 A cross-correlation of above 0.75 in the Hamby study is indicative of a match.
-There are 97 land-land comparisons with a cross-correlation of above 0.75.
+There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
 \begin{figure}
@@
   theme(legend.position="bottom") + 
+  scale_fill_brewer(palette="Paired") +
   xlab("maximum CMS")
@@
   ylab("Distance between crosscuts") +
+  scale_colour_brewer(palette="Paired") +
   theme_bw() + 
@@
 ggplot(data=bstats) + theme_bw() + 
-  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + ylab("Cross-correlation") +
+  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + 
+  facet_wrap(~match, labeller="label_both") + 
+    scale_colour_brewer(palette="Paired") +
+  xlab("maximum CMS") + ylab("Cross-correlation") +
   theme(legend.position="none")
@@
 \hh{XXX another idea for the rest of the paper: use some datamining methods to match the Hamby bullets }
-Using recursive partitioning, we find a decision tree \cite{breiman:1984, rpart, rpart.plot} predicting matches based on features derived from the image files.
+Using recursive partitioning, we fit a decision tree \cite{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
 The resulting tree is shown in Figure~\ref{fig:tree}.
-A total of 105 lands is being matched correctly.
-Interestingly, the number of consecutive matching striae is not featured in this evaluation.
-Instead, cross-correlation (ccf) between the crosscuts is featured heavily.
-Aside from cross correlation, the number of matches between crosscuts is involved in the decision.
+A total of 123 lands is being matched correctly.
+Interestingly, the number of consecutive matching striae only features as the last step in this evaluation.
+At this point in the tree, the first false positive matches are introduced into the results: out of the ten land-to-land comparisons three are mis-matched.
+Instead of CMS, cross-correlation (ccf) between the crosscuts is very important in the matching process  by the decision tree.
+Aside from  cross correlation, the number of matches (num.matches) between crosscuts and the distance between crosscuts (distr.dist) is involved in the decision.
 Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
@@
 This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2d microscopic images of bullets.
-Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitalized version of the images that preserves the 3D surface structure.
+Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitalized version of the images that preserves the 3d surface structure.
 
@@
 library(rpart.plot)
+library(RColorBrewer)
+vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
+
 CCFs <- read.csv("data-35-25/bullet-stats.csv")
@@
 per <- rp1$frame$yval # predictive probability for each node in the tree
-prp(rp1, extra = 101, box.col=c("grey90", "orange")[as.numeric(per > 0.5)+1])
+
+prp(rp1, extra = 101, box.col=vals[as.numeric(per > 0.5)+1])
 
@@
 
-<<pred:table, echo=FALSE, results='asis'>>=
-library(xtable)
-xtable(xtabs(~I(pred>0.5)+match, data=CCFs))
-@
+\hh{XXX still need discussion of the results from random forests \cite{breiman:2001, randomForest}}
 
@@
 \begin{figure}
-<<tree:forest, echo=FALSE, fig.width=5, fig.height=6, out.width='\\linewidth'>>=
-qplot(pred, forest, geom="point", data=CCFs, colour=match) + theme_bw() +
-  theme(legend.position="bottom")
+<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
+qplot(pred, forest, geom="jitter", data=CCFs, colour=match) + theme_bw() +
+  theme(legend.position="bottom") +
+  geom_vline(xintercept=0.5, colour="grey50") +
+  geom_hline(yintercept=0.5, colour="grey50") + 
+  xlab("Predictions from decision tree") +
+  ylab("Predictions from random forest")   +
+  scale_colour_brewer(palette="Paired")
 @
@@
 Each point corresponds to a land-to-land comparison.
-Matches are indicated in blue.}
+Known matches are indicated in blue.
+Matches predicted by the decision tree are on the right half of the plot, matches predicted by the random forest are in the upper half.
+}
 \end{figure}
@@
 \item we get probabilities for matches. 
-
+\item We don't have a solution for subclass characteristics
 \end{enumerate}

### 2015-12-09 imaging-paper.Rnw
Correct a typo
xtable as a dependency

@@
 library(rpart.plot)
+library(xtable)
 
@@
 \section{Evaluation}
-Using the James Hamby study, we run all unkown bullets against all known bullets for matches.
+Using the James Hamby study, we run all unknown bullets against all known bullets for matches.
 The algorithm is using land-to-land matches, i.e.

### 2015-12-09 imaging-paper.Rnw
trying to make figures show up closer to their text

@@
 \usepackage{float}
+\usepackage{afterpage}
 
@@
 \usepackage{subcaption}
+
+    \renewcommand{\topfraction}{0.99}	% max fraction of floats at top
+    \renewcommand{\bottomfraction}{0.99}	% max fraction of floats at bottom
 
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 \begin{subfigure}[t]{\linewidth}
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<side-by-side, echo=FALSE, eval=FALSE>>=
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[hbtp]
 \includegraphics[width=\columnwidth]{images/matchup-rgl.png}
@@
 @
-\begin{figure}
+\begin{figure}[hbtp]
 \centering
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '\\linewidth'>>=
@@
 
-\begin{figure}
+\subsection{Varying crosscuts}
+The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up.
+For bullet land 1-5 from barrel 1 crosscuts at intervals of 50$\mu m$ between 150$\mu m$  and 450$\mu m$  are compared to a crosscut at 100$\mu m$.
+The initially almost perfect match (consolidated by a CMS of fourteen) between the  crosscuts deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.
+Only if the crosscut is within 150$\mu m$ do we get a good visual match.
+Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections.
+\begin{figure}[hbtp]
+<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+paths <- file.path(datadir, dir(datadir))
+crosscuts <- seq(100, 450, by = 50)
+lof <- processBullets(paths[5], x = crosscuts)
+lof$bullet <- paste(lof$bullet, lof$x)
+
+reslist <- lapply(crosscuts[-1], function(cc) {
+#  browser()
+  b2 <- subset(lof, x %in% c(cc, 100))
+  lofX <- bulletSmooth(b2)
+  bAlign = bulletAlign(lofX)
+  lofX <- bAlign$bullet
+    b12 <- unique(b2$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
+
+#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+#  lines <- striation_identify(lofX, threshold = threshold)
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
+  maxCMS <- maxCMS(lines$match==TRUE)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
+})
+
+ccfs <- sapply(reslist, function(res) res$ccf)
+
+lop <- lapply(reslist, function(res) {
+ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
+  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
+  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
+  scale_linetype_discrete("") +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
+    ylab("") + xlab("")
+})
+
+grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
+             lop[[7]],
+             ncol = 1)
+@
+\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts.
+Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$.
+With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
+}
+\end{figure}
+\afterpage{\clearpage}
+
+%\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
+By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, indicative of a break-off, or in a stable section, where we can expect to find matches reasonably well.
+In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
+This process is shown  in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
+\hh{`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. }
+\begin{figure}[hbtp]
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+paths <- file.path(datadir, dir(datadir))
+crosscuts <- seq(100,225, by=25)
+lof <- processBullets(paths[37], x = crosscuts)
+lof$bullet <- paste(lof$bullet, lof$x)
+
+reslist <- lapply(1:length(crosscuts[-1]), function(i) {
+  b2 <- subset(lof, x %in% crosscuts[i:(i+1)])
+  lofX <- bulletSmooth(b2)
+  bAlign <- bulletAlign(lofX)
+  lofX <- bAlign$bullet
+
+  b12 <- unique(b2$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
+
+#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+#  lines <- striation_identify(lofX, threshold = threshold)
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
+  maxCMS <- maxCMS(lines$match==TRUE)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
+})
+
+ccfs <- sapply(reslist, function(res) res$ccf)
+
+lop <- lapply(reslist, function(res) {
+ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
+  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
+  scale_linetype_discrete("") +
+#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylim(c(-6,6)) + ylab("") + xlab("") + 
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
+})
+
+grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
+             lop[[4]], lop[[5]], 
+             ncol=1)
+@
+\caption{\label{fig:crosscuts2}Varying crosscuts for barrel 3, bullet 1.
+Initially, the match between crosscuts 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet.
+At a crosscut level of $175\mu m$ the bullet pattern stabilizes.
+For this land, matches should not be attempted at lower crosscuts.}
+\end{figure}
+
+\begin{figure}[hbtp]
 \centering
@@
 \end{figure}
+\afterpage{\clearpage}
+
+\subsection{Automated matching algorithm}
 
@@
 
-\subsection{Varying crosscuts}
-The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up.
-For bullet land 1-5 from barrel 1 crosscuts at intervals of 50$\mu m$ between 150$\mu m$  and 450$\mu m$  are compared to a crosscut at 100$\mu m$.
-The initially almost perfect match (consolidated by a CMS of fourteen) between the  crosscuts deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.
-Only if the crosscut is within 150$\mu m$ do we get a good visual match.
-Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections.
-\begin{figure}
-<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
-paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100, 450, by = 50)
-lof <- processBullets(paths[5], x = crosscuts)
-lof$bullet <- paste(lof$bullet, lof$x)
-
-reslist <- lapply(crosscuts[-1], function(cc) {
-#  browser()
-  b2 <- subset(lof, x %in% c(cc, 100))
-  lofX <- bulletSmooth(b2)
-  bAlign = bulletAlign(lofX)
-  lofX <- bAlign$bullet
-    b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
-
-#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-#  lines <- striation_identify(lofX, threshold = threshold)
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
-})
-
-ccfs <- sapply(reslist, function(res) res$ccf)
-
-lop <- lapply(reslist, function(res) {
-ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
-  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
-  scale_linetype_discrete("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
-    ylab("")
-})
-
-grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
-             lop[[7]],
-             ncol = 1)
-@
-\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts.
-Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$.
-With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
-}
-\end{figure}
-
-%\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
-By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, indicative of a break-off, or in a stable section, where we can expect to find matches reasonably well.
-In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
-This process is shown  in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
-\hh{`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. }
-\begin{figure}
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
-paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100,225, by=25)
-lof <- processBullets(paths[37], x = crosscuts)
-lof$bullet <- paste(lof$bullet, lof$x)
-
-reslist <- lapply(1:length(crosscuts[-1]), function(i) {
-  b2 <- subset(lof, x %in% crosscuts[i:(i+1)])
-  lofX <- bulletSmooth(b2)
-  bAlign <- bulletAlign(lofX)
-  lofX <- bAlign$bullet
-
-  b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
-
-#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-#  lines <- striation_identify(lofX, threshold = threshold)
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
-})
-
-ccfs <- sapply(reslist, function(res) res$ccf)
-
-lop <- lapply(reslist, function(res) {
-ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
-  scale_linetype_discrete("") +
-#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6)) + ylab("") +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
-})
-
-grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
-             lop[[4]], lop[[5]], 
-             ncol=1)
-@
-\caption{\label{fig:crosscuts2}Varying crosscuts for barrel 3, bullet 1.
-Initially, the match between crosscuts 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet.
-At a crosscut level of $175\mu m$ the bullet pattern stabilizes.
-For this land, matches should not be attempted at lower crosscuts.}
-\end{figure}
-
-\subsection{Automated matching algorithm}
 \hh{The automated matching routine is then as follows:
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='\\linewidth', warning=FALSE>>=
@@
 \end{figure}
+\afterpage{clearpage}
 
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 \begin{subfigure}[t]{.49\textwidth}
@@
 There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
-\begin{figure}
+%
+\begin{figure}[hbtp]
 \begin{subfigure}[t]{\linewidth}
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:distr.dist}Relationship between maximal CMS, distances between crosscuts and matches.}
 <<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
@@
 @
-\caption{If the distance between the crosscuts is small, a smaller number of CMS might already be an indication of a match.}
-\end{figure}
-
-\begin{figure}
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:ccfs-match}Relationship between maximal CMS, cross-correlation between crosscuts and matches.}
 <<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
@@
 @
-\caption{\label{fig:ccfs-match}If the cross-correlations between the crosscuts is high, a smaller number of CMS might already be an indication of a match.}
-\end{figure}
-
-\hh{XXX another idea for the rest of the paper: use some datamining methods to match the Hamby bullets }
+\end{subfigure}
+\caption{A small distance between crosscuts or a large cross-correlation between crosscuts are indicative of matches between lands.}
+\end{figure}
+
 Using recursive partitioning, we fit a decision tree \cite{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='\\linewidth'>>=
@@
 
-\begin{figure}
+\begin{figure}[hbtp]
 <<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=

### 2015-12-09 imaging-paper.Rnw
expanded on discussion of varying crosscuts

@@
 The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up.
-For bullet land 1-5 from barrel 1 crosscuts at intervals of 50$\mu m$ between 150$\mu m$  and 450$\mu m$  are compared to a crosscut at 100$\mu m$.
-The initially almost perfect match (consolidated by a CMS of fourteen) between the  crosscuts deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.
-Only if the crosscut is within 150$\mu m$ do we get a good visual match.
-Given that alignment in the microscope is not perfect, we should take this into account in the automatic matching routine by evaluating matches at several cross sections.
+This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which crosscut level was used in a matching, but on the other hand we can use multiple crosscuts of the same bullet for an initial assessment of its quality.Both of these situations are assessed and discussed in more detail below:
+Figure~\ref{fig:crosscuts} shows a sequence of crosscuts for bullet land 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$.
+These  are compared to a crosscut at 100$\mu m$.
+Initially this comparison constitutes an almost perfect match (consolidated by a CMS of fourteen) between the two crosscuts.
+However, the match quickly deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.
+Only if the crosscut is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
+Given that we have to expect some variation in nominally the same crosscut levels due to manual alignments in microscopes, we should take crosscut levels into account in the automatic matching routine by evaluating matches at several cross sections.
 \begin{figure}[hbtp]
@@
 %\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below} 
-By comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section, indicative of a break-off, or in a stable section, where we can expect to find matches reasonably well.
+An opportunity that arises from comparing multiple crosscuts of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
 In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
 This process is shown in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
-\hh{`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. }
+`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. 
 \begin{figure}[hbtp]

### 2015-12-09 imaging-paper.Rnw
fiddling with figure placement

@@
 The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up.
-This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which crosscut level was used in a matching, but on the other hand we can use multiple crosscuts of the same bullet for an initial assessment of its quality.Both of these situations are assessed and discussed in more detail below:
+This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which crosscut level was used in a matching, but on the other hand we can use multiple crosscuts of the same bullet for an initial assessment of its quality.
+Both of these situations are assessed and discussed in more detail below:
 Figure~\ref{fig:crosscuts} shows a sequence of crosscuts for bullet land 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$.
@@
 \end{figure}
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 
@@
 \end{figure}
+\afterpage{\clearpage}
 

### 2015-12-09 imaging-paper.Rnw
added reference and expanded conclusion

@@
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
-\hh{In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$, which is consistent with a 9 mm bullet fired in a Ruger P-85. }
+\hh{In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.
+Both of these estimates are consistent with a 9 mm bullet fired in a Ruger P-85.}
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
@@
 
-Figure~\ref{fig:match} gives an overview of the rest of the automated matching routine: after smoothing the scan with a loess of a very small span, shown in (a), we define a striation mark as a groove or a peak if \hh{XXX }
-
-Two lines are defined as a match, if their peak/groove base at the threshold overlaps.
-If there is no overlap, a peak/groove leads to a mismatch.
-We then count the number of maximum consecutive matches.
-In this example, the number of consecutive matching striations (CMS) is fifteen, a very high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches  \cite{nichols:2003}.
-
-\hh{The automated matching routine is then as follows:
-Identify a stable region for each bullet and pick the lowest crosscut level in this region (assumption here: lower crosscut levels show better expression of individual marks XXX need to still check)
-
-All of the other steps are done on pairs of bullets:
+Figure~\ref{fig:match} gives an overview of the automated matching routine: 
+We first identify a stable region for each bullet and pick the lowest crosscut level in this region \hh{(assumption here: lower crosscut levels show better expression of individual marks XXX need to still check)}
+
+All of the other steps are done on pairs of bullet lands:
 \begin{enumerate}
@@
 \item Identify the maximal number of CMS (consecutive matching striae) between the two crosscuts by counting consecutive matching lines. 
+In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
+For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches  \cite{nichols:2003}.
 \end{enumerate}
-}
+
 \hh{XXX next step: how does this matching work in known non-matches? distribution of CMS under a null of no match? }
@@
 \caption{Overlaid lines of the land of the unknown bullet F-3 and the land with the top matching CMS.}
-<<top, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
+<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
 res <- reslist[[which.max(cmsdist)]]  
@@
 \caption{Overlaid lines of the land of the unknown bullet F-3 and the land with the second top matching CMS.}
-<<top2, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
+<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
 #res <- reslist[[which.max(cmsdist)]]  
@@
 While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.
+\item reducing subjectivity: `the concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner' \cite{miller:1998}.
 \item involve multiple crosscuts for each bullet in a match?

### 2015-12-09 imaging-paper.Rnw
couple of edits

@@
 These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article \cite{afte:1992}.
-``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows." \footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}, \footnote{AFTE 1992 theory of identification \url{http://projects.nfstc.org/firearms/module09/fir_m09_t05_05.htm}}
+``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows." %\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}
 
@@
 \section{Approach to automatic matching}
-Applying the  loess fit to a range of different surface cross cuts (see Figures~\ref{fig:manualmatch} and Figure~\ref{fig:manualmatch}, with cross cuts from 50 to 150) shows the striation marks across two bullets.
+Applying the  loess fit to a range of different surface cross cuts (see Figures~\ref{fig:manualmatch} and~\ref{fig:manualmatch-rgl}, with cross cuts from 50 to 150) shows the striation marks across two bullets.
 Cross cuts of bullet 1 are shown on the left (cross cuts with values below 100) and cross cuts of bullets 2 are shown on the right (cross cuts with values above 100).
@@
 \afterpage{\clearpage}
+
+\subsection{Varying smoothing factor}
+\hh{XXX it would be good to have a discussion of the smoothing factor XXX}
 
@@
 \end{figure}
-\afterpage{clearpage}
 

### 2015-12-09 imaging-paper.Rnw
another picture: mismatch with CMS 15

@@
 
+\begin{figure}
+<<strange-res, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+load("data-35-25/unkn82.RData")
+res <- reslist[[97]]
+
+print(ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
+  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
+#  geom_hline(yintercept = res$threshold) +
+#  geom_hline(yintercept = - res$threshold) +
+#  scale_colour_brewer("",palette="Set1") +
+  scale_linetype_discrete("") +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+@
+\caption{\label{fig:mismatch}Known mismatch with a surprisingly large number of maximal consecutive matching striae (fifteen) and a small distance between the crosscuts. }
+\end{figure}
+
 \begin{figure}[hbtp]

### 2015-12-09 imaging-paper.Rnw
Bullet 1-5

@@
   ylab("Residuals from circular fit") +
-  ggtitle("Land 5 of Bullet 1 in black")
-@
-\caption{Circle fit for each cross section of bullet 2, with land 5 of bullet 1 overlaid.\label{fig:circlefits} \hh{Land 5 of bullet 1 matches best with the first land of the second bullet.}}%
+  ggtitle("Bullet 1-5 in black")
+@
+\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} \hh{Bullet 1-5 matches best with the first land of the second bullet.}}%
 \end{figure}
@@
 \caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm.
-For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for land 5 of bullet 1 are plotted across different values of $y$.}
+For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_fit} Loess fit for land 1 of bullet 1.}
+\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
 <<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_resid} Residuals of Loess fit for land 1 of bullet 1.}
+\caption{\label{fig:loess_resid} Residuals of Loess fit for bullet 1-5.}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
@@
 \end{subfigure}
-\caption{Fit and residuals of a loess fit to land 5 of bullet 1 (Barrel 1).}
-\end{figure}
-
-We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from land 5 of bullet 1.
+\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
+\end{figure}
+
+We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet 1-5.
 This is shown in Figure~\ref{fig:loessfits}.
@@
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Land 5 of Bullet 1 in black") + ylab("Residuals of Loess Fit")
-@    
-\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Land 5 of bullet 1 is shown in black. }
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Bullet 1-5 in black") + ylab("Residuals of Loess Fit")
+@    
+\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet 1-5 is shown in black. }
 \end{figure}

### 2015-12-09 imaging-paper.Rnw
axis labels and some rearrangement

@@
 #  scale_colour_brewer("",palette="Set1") +
-  scale_linetype_discrete("") +
+  scale_linetype_discrete("") + ylab("") +
   scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
@@
 All 29 pairs of lands with at least sixteen CMS in common are matches.
-For smaller number of CMS, the percentage of false positives quickly increases.
-Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches.
-A cross-correlation of above 0.75 in the Hamby study is indicative of a match.
-There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
+However, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller \cite{miller:1998}.
+Figure~\ref{fig:mismatch} shows a known mismatch, XXXX
 %
@@
 
-\begin{figure}
-<<strange-res, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+\begin{figure}[hbtp]
+<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\linewidth', warning=FALSE>>=
 load("data-35-25/unkn82.RData")
@@
 
+For smaller number of CMS, the percentage of false positives quickly increases.
+Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches.
+A cross-correlation of above 0.75 in the Hamby study is indicative of a match.
+There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
 \begin{figure}[hbtp]

### 2015-12-10 imaging-paper.Rnw
added to the table of importance results ... needs more work

@@
 All 29 pairs of lands with at least sixteen CMS in common are matches.
-However, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller \cite{miller:1998}.
-Figure~\ref{fig:mismatch} shows a known mismatch, XXXX
+There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
+On average, there are 2.2 maximal CMS between known non-matches (with a standard deviation of 1.4).
+Known matches share on average 8.2 maximal CMS, with a standard deviation of 5.8.
+Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller \cite{miller:1998}.
+Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share fifteen consecutively matched striae.
+The number of peaks and valleys, their heights and depths as well as the overall pattern are in relative good agreement.
+\hh{XXX something re-assuring that visually we can tell them apart?
+I don't know - this could just as well be a case where the crosscuts are very different.
+}
 %
@@
 
-For smaller number of CMS, the percentage of false positives quickly increases.
-Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches.
-A cross-correlation of above 0.75 in the Hamby study is indicative of a match.
+For smaller numbers of CMS, the percentage of false positives quickly increases.
+However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two crosscuts (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
 There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
@@
 \end{figure}
+
+Similarly, the Euclidean distance between two crosscuts across the range of loess residuals is strongly correlated to known matches.
+Out of the 48 pairs of land-to-land crosscuts with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of crosscuts that both are very flat with peaks and vallyes of less than 1.5$\mu m$).
+The average distance between known matches in the Hamby study is $1.9\mu m$ with a standard deviation of $2.3\mu m$; while known mismatches have an average  distance of  3.3$\mu m$ with a standard deviation of $1.9\mu m$.
 
@@
 imp <- imp[order(-imp$MeanDecreaseGini),]
-names(imp)[1] <- "Mean Decrease Gini"
-print(xtable(imp[,2:1]), include.rownames=FALSE, floating = FALSE)
+names(imp)[1] <- "Importance"
+imp <- imp[,2:1]
+imp$meanMatch <- sapply(imp$Variable, function(var) {
+  mean(subset(CCFs, match)[,var])
+})
+imp$sdMatch <- sapply(imp$Variable, function(var) {
+  sd(subset(CCFs, match)[,var])
+})
+imp$meanNonMatch <- sapply(imp$Variable, function(var) {
+  mean(subset(CCFs, !match)[,var])
+})
+imp$sdNonMatch <- sapply(imp$Variable, function(var) {
+  sd(subset(CCFs, !match)[,var])
+})
+imp$meanMatch <- round(imp$meanMatch,1)
+imp$sdMatch <- sprintf("(%.2f)", imp$sdMatch)
+imp$meanNonMatch <- round(imp$meanNonMatch,1)
+imp$sdNonMatch <- sprintf("(%.2f)", imp$sdNonMatch)
+
+print(xtable(imp, align="clrrrrr"), include.rownames=FALSE, floating = FALSE)
 @

### 2015-12-10 imaging-paper.Rnw
edited table a bit more

@@
 Similarly, the Euclidean distance between two crosscuts across the range of loess residuals is strongly correlated to known matches.
-Out of the 48 pairs of land-to-land crosscuts with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of crosscuts that both are very flat with peaks and vallyes of less than 1.5$\mu m$).
+Out of the 48 pairs of land-to-land crosscuts with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of crosscuts that both are very flat with peaks and valleys of less than 1.5$\mu m$).
 The average distance between known matches in the Hamby study is $1.9\mu m$ with a standard deviation of $2.3\mu m$; while known mismatches have an average  distance of  3.3$\mu m$ with a standard deviation of $1.9\mu m$.
+Both of these additional variables show large, if not significant, differences between matches and non-matches.
+Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of  a set of feature variables derived from the 3d topological surface measurements.
+There are quite large differences in these averages, indicating that we can successfully employ machine learning methods to distinguish matches from non-matches.
 
@@
 
-\begin{table}
+\begin{table*}[tbhp]
 \centering
@@
 imp$meanMatch <- round(imp$meanMatch,1)
-imp$sdMatch <- sprintf("(%.2f)", imp$sdMatch)
+imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
 imp$meanNonMatch <- round(imp$meanNonMatch,1)
-imp$sdNonMatch <- sprintf("(%.2f)", imp$sdNonMatch)
-
-print(xtable(imp, align="clrrrrr"), include.rownames=FALSE, floating = FALSE)
-@
-\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches. }
-\end{table}
+imp$sdNonMatch <- sprintf("(%5.2f)", imp$sdNonMatch)
+
+names(imp)[-(1:2)] <- c("Match", "(sd)", "Non-Match", "(sd)")
+print(xtable(imp, align="clrrlrl"), include.rownames=FALSE, floating = FALSE)
+@
+\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches.
+Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
+Match and Non-Match show averages of each of the variables for macthes and non-matches only.
+In parentheses behind the means, standard deviations are given}
+\end{table*}
 

### 2015-12-10 imaging-paper.Rnw
explanation of decision tree leaves

@@
 @
-\caption{\label{fig:tree}Decision tree of matching bullets based on recursive partitioning.}
-\end{figure}
-
-\hh{XXX still need discussion of the results from random forests \cite{breiman:2001, randomForest}}
+\caption{\label{fig:tree}Decision tree of matching bullets based on recursive partitioning.
+The rectangular nodes are the leaves, giving a short summary consisting of the number of observations in the leaf (bottom left), the corresponding percentage of the total (bottom right).
+The number at the top shows the fraction of these observations that are a match.
+A 1 or a 0 therefore indicate a homogeneous (or perfect) node.
+}
+\end{figure}
+
+Another benefit is that we can not only apply a single decision tree, but several hundred of them to combine in a random forest \cite{breiman:2001, randomForest}. This allows us to quantify XXXX
 
@@
 imp$meanNonMatch <- round(imp$meanNonMatch,1)
-imp$sdNonMatch <- sprintf("(%5.2f)", imp$sdNonMatch)
+imp$sdNonMatch <- sprintf("(%6.2f)", imp$sdNonMatch)
 

### 2015-12-10 imaging-paper.Rnw
Clean up section introducing loess.

@@
 Figure \ref{fig:loess_resid} shows the residuals from this fit.
-\hh{XXX what purpose does the next paragraph have?
-I think you just want end this section, but it brings up two new ideas: autocorrelation and dependency on the span.
-The span obviously doesn't get any resolution ...
-so it's just muddying the waters at the moment ...
-XXX}
-It is clear that there is a rather significant amount of autocorrelation, which make sense given the presence of striations and markings.
-Certainly, tuning the parameters for this regression, such as the Loess span, could prove to do an even better job at making the markings more clear.
-\hh{XXX did you follow up on that?
-What is the effect of using a different span?}
+We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet 1-5.
+This is shown in Figure~\ref{fig:loessfits}.
+Bullet 2-1 is still the best match for bullet 1-5, but now the details of the match are more pronounced.
 
@@
 \end{figure}
-
-We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet 1-5.
-This is shown in Figure~\ref{fig:loessfits}.
-\hh{The first land of bullet 2 is still the best  match for land 5, but now the details of the match are more pronounced.}
 

### 2015-12-10 imaging-paper.Rnw
Describe twicing a little bit better

@@
 That is, the smoothed data is once again smoothed by computing a new rolling average with the same smoothing factor.
-This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass \cite{tukey:1977}.
-\hh{XXX Did Tukey say, what the goal of the twicing was?
-- also, do you have a page number for the twicing?}
+This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass and then added back to the result \cite{tukey:1977}.
+This has the effect of introducing a little bit more variance into the smoothed data.
+We instead performed a second smoothing pass on the smoothed data, which has the effect of weighting observations near the center highest, while the weights linearly drop off as you reach the end of the smoothing window or smoothing factor.
 

### 2015-12-10 imaging-paper.Rnw
Disclaimer on website

@@
 In the groove detection portion of this algorithm, we wish to remove most of the deviation but still maintain enough signal in the groove to detect it.
-Empirically, a value of 35 for the smoothing factor seems to work quite well.
+Empirically, a value of 35 for the smoothing factor seems to work quite well (later in the manuscript we investigate varying the smoothing factor).
 It is important to note that the smoothing pass is done twice.
@@
 The application includes a button which allows the automatic creation of a residual plot from the Loess fits of two bullets.
-An x coordinate value can be selected, and a cross-section of the two bullet at the x value will be taken.
+An x coordinate value can be selected as a cross cut, and a cross-section of the two bullet at the x value will be taken.
 The previously described Loess regression procedure will be applied to the two cross sections.
@@
 It is intended that users will perform this procedure at different x values in order to see whether the markings indicate a match at different points along the bullet.
+
+Enhancements to the web application are still in progress. In particular, much of the automatic matching functionality described in the next section will be added in the future.
 

### 2015-12-10 imaging-paper.Rnw
Start on smoothing factor variation, much more coming

@@
 @
-\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} \hh{Bullet 1-5 matches best with the first land of the second bullet.}}%
+\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with the bullet 2-1}
 \end{figure}
@@
 \subsection{Varying smoothing factor}
-\hh{XXX it would be good to have a discussion of the smoothing factor XXX}
+As mentioned, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
+In particular, for detecting and removing the grooves prior to fitting a Loess regression, we selected a smoothing factor of 35
 

### 2015-12-10 imaging-paper.Rnw
Correct some small typos

@@
 
--%\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
+%\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
 
@@
 \caption{\label{fig:residual} Residual structure of circular fits at two different cross-section.
-Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.
-}
-\end{figure}
-
-A single cylinder as a fit is unlikely to be a particularly good fit, because  there seem to be quite massive deformations in vertical direction.
+Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
+\end{figure}
+
+A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction.
 Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, we can not address all of these issues.
@@
 @
-\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with the bullet 2-1}
+\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with the bullet 2-1.}
 \end{figure}

### 2015-12-10 imaging-paper.Rnw
More smoothing factor work...

@@
 As mentioned, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a Loess regression, we selected a smoothing factor of 35
+In particular, for detecting and removing the grooves prior to fitting a Loess regression, we selected a smoothing factor of 35.
+We can investigate the performance of groove detection across different smoothing factors.
+
+\begin{figure}[hbtp]
+\centering
+\begin{subfigure}[t]{\linewidth}
+\caption{Crosscuts at $x = 100$.\label{fig:smoothfac1}}{%
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=2, out.width='\\textwidth'>>=
+get_grooves(br111_bullet, smoothfactor = 35, smoothplot = TRUE)$plot
+@
+    }
+\end{subfigure}    
+\begin{subfigure}[t]{\linewidth}
+    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:smoothfac2}}{%
+<<smoothfac2, echo=FALSE, fig.width=7, fig.height=2, out.width='\\textwidth'>>=
+get_grooves(br111_bullet, smoothfactor = 5, smoothplot = TRUE)$plot
+@
+    }
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:smoothfac3}}{%
+<<smoothfac3, echo=FALSE, fig.width=7, fig.height=2, out.width='\\textwidth'>>=
+get_grooves(br111_bullet, smoothfactor = 55, smoothplot = TRUE)$plot
+@
+    }
+\end{subfigure}
+\caption{\label{fig:varysmooth}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
+\end{figure}
 

### 2015-12-10 imaging-paper.Rnw
Tweak the smoothing factor stuff

--- 2015-12-10 imaging-paper.Rnw
@@
 As mentioned, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a Loess regression, we selected a smoothing factor of 35
+In particular, for detecting and removing the grooves prior to fitting a Loess regression, we selected a smoothing factor of 35.
+We can investigate the performance of groove detection across different smoothing factors.
+Figure \ref{fig:varysmooth}...
+
+\begin{figure}[hbtp]
+\centering
+\begin{subfigure}[t]{\linewidth}
+\caption{Groove detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+get_grooves(br111_bullet, smoothfactor = 35, smoothplot = TRUE)$plot
+@
+    }
+\end{subfigure}    
+\begin{subfigure}[t]{\linewidth}
+    \caption{Groove detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
+<<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+get_grooves(br111_bullet, smoothfactor = 5, smoothplot = TRUE)$plot
+@
+    }
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+    \caption{Groove detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
+<<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+get_grooves(br111_bullet, smoothfactor = 65, smoothplot = TRUE)$plot
+@
+    }
+\end{subfigure}
+\caption{\label{fig:varysmooth} Groove detection at smoothing factors of 35, 5, and 65 respectively.
+Note that a smoothing factor of 5 yields enough variation prior to the groove that the right groove is not detected properly, while a smoothing factor of 65 over-smoothes and causes the peak on the right to disappear, so that the peak and groove are both not detected properly.}
+\end{figure}
 

### 2015-12-10 imaging-paper.Rnw
Correct a small warning issue

@@
 \caption{Groove detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
 get_grooves(br111_bullet, smoothfactor = 35, smoothplot = TRUE)$plot
@@
     \caption{Groove detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
-<<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+<<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
 get_grooves(br111_bullet, smoothfactor = 5, smoothplot = TRUE)$plot
@@
     \caption{Groove detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
-<<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+<<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
 get_grooves(br111_bullet, smoothfactor = 65, smoothplot = TRUE)$plot

### 2015-12-10 imaging-paper.Rnw
more on random forests

@@
 
-Another benefit is that we can not only apply a single decision tree, but several hundred of them to combine in a random forest \cite{breiman:2001, randomForest}. This allows us to quantify XXXX
-
+<<rforest, echo=FALSE, message=FALSE>>=
+library(randomForest)
+library(reshape2)
+rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
+
+errors <- data.frame(rtrees$err.rate)
+errors$id <- 1:nrow(errors)
+CCFs$forest <- predict(rtrees, type="prob")[,2]
+@
+
+Another benefit is that we can not only apply a single decision tree, but several hundred of them to combine in a random forest \cite{breiman:2001, randomForest}.
+In a random forest multiple trees are fitted.
+For each of these trees, only two thirds of the observations are used for fitting, while predictions for the remaining third  are used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate.
+Because errors are determined from the one third of held-back observations, this error rate is called the out-of bag (OOB) error.
+Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate for 300 trees. 
+%
+\begin{figure}[hbtp]
+<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='\\linewidth'>>=
+qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
+@
+\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
+\end{figure}
+%
+After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0056.
+This is a weighted average between false positives of 0.00028 and false negatives of 0.32.
+Note that these error rates are based on land-to-land comparisons and will be much lower for bullet-to-bullet matches.
+In the case of the Hamby study, the error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched (see Figure~\ref{fig:treeresult}).
 \begin{figure}[hbtp]
@@
 )
-
-qplot(data=bullets, pred, reorder(bullet, pred/n)) + 
+library(reshape2)
+bm <- melt(data=bullets, id.var=c("bullet", "n"))
+
+bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
+qplot(data=bm, value, reorder(bullet, value/n), size=I(2.5), 
+      shape=prediction, colour=prediction, alpha=I(0.6)) + 
   theme_bw()  + ylab("") + 
+#  scale_shape_manual(values=c(1,2)) +
   scale_x_continuous("Number of correctly predicted land-to-land matches", 
-                     breaks = 3*0:4, limits=c(0,12))
+                     breaks = 3*0:4, limits=c(0,12),
+                     minor_breaks = 1:11) +
+  theme(legend.position = "bottom")
 @
 \caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree.
-For each bullet at least one land is correctly matched.
+Both results for the tree as well as the random forest are shown.
+For each bullet at least two lands are correctly matched.
 Unknown bullets B and M are the hardest for the algorithm to match.}
 \end{figure}
-
-<<rforest, echo=FALSE, message=FALSE>>=
-library(randomForest)
-library(reshape2)
-rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
-
-errors <- data.frame(rtrees$err.rate)
-errors$id <- 1:nrow(errors)
-CCFs$forest <- predict(rtrees, type="prob")[,2]
-@
-
-Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate.
-
-\begin{figure}[hbtp]
-<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='\\linewidth'>>=
-qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
-@
-\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
-\end{figure}
-
+%
+However, this does not carry a lot of weight because of the few bullet comparisons in question: fifteen unknown bullets are successfully matched to two pairs of ten bullets.
+Matching bullets can only be tested realistically in a much bigger test case.
+Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.
+whenever the predicted probability of a match exceeds 0.5, a match is declared.
+Basing this decision on 0.5 is not necessary.
+In practice, examiners are allowed a third option of `inconclusive'.
+On a probability spectrum of outcomes we can therefore introduce an interval of `inconclusive' results in the middle of the spectrum.
+
+Besides a probabilistic quantification of matches, we get from a random forest an assessment of the importance of each of the features derived from the bullets' 3d topological surface measurements.
+Table~\ref{tab:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree.
 \begin{table*}[tbhp]
@@
 imp$meanNonMatch <- round(imp$meanNonMatch,1)
-imp$sdNonMatch <- sprintf("(%6.2f)", imp$sdNonMatch)
-
+imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
+row.names(imp) <- 1:nrow(imp)
 names(imp)[-(1:2)] <- c("Match", "(sd)", "Non-Match", "(sd)")
-print(xtable(imp, align="clrrlrl"), include.rownames=FALSE, floating = FALSE)
+print(xtable(imp, align="clrrcrc"), floating = FALSE)
 @
@@
 \end{table*}
+
+The rows in the table are ordered according to descending importance.
+The variable with the most predictive power by far is cross-correlation (ccf), followed by the number of matches (num.matches) and the Euclidean distance between the crosscuts (distr.dist).
+The maximal consecutively matching striae (cms) is found in seventh position.
 

### 2015-12-10 imaging-paper.Rnw
fixed typo
acknowledgement
random forest ... done?

@@
 On a probability spectrum of outcomes we can therefore introduce an interval of `inconclusive' results in the middle of the spectrum.
+Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis).
+Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only 13 cases result in a different decision.
+
+\begin{figure}[hbtp]
+<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
+qplot(pred, forest, geom="jitter", data=CCFs, colour=match) + theme_bw() +
+  theme(legend.position="bottom") +
+  geom_vline(xintercept=0.5, colour="grey50") +
+  geom_hline(yintercept=0.5, colour="grey50") + 
+  xlab("Predictions from decision tree") +
+  ylab("Predictions from random forest")   +
+  scale_colour_brewer(palette="Paired")
+@
+\caption{\label{fig:tree-forest}Scatterplot of prediction results from the tree (horizontal) and the forest (vertical).
+Each point corresponds to a land-to-land comparison.
+Known matches are indicated in blue.
+Matches predicted by the decision tree are on the right half of the plot, matches predicted by the random forest are in the upper half.
+}
+\end{figure}
 
@@
 Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
-Match and Non-Match show averages of each of the variables for macthes and non-matches only.
+Match and Non-Match show averages of each of the variables for matches and non-matches only.
 In parentheses behind the means, standard deviations are given}
@@
 The variable with the most predictive power by far is cross-correlation (ccf), followed by the number of matches (num.matches) and the Euclidean distance between the crosscuts (distr.dist).
-The maximal consecutively matching striae (cms) is found in seventh position.
-
-\begin{figure}[hbtp]
-<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
-qplot(pred, forest, geom="jitter", data=CCFs, colour=match) + theme_bw() +
-  theme(legend.position="bottom") +
-  geom_vline(xintercept=0.5, colour="grey50") +
-  geom_hline(yintercept=0.5, colour="grey50") + 
-  xlab("Predictions from decision tree") +
-  ylab("Predictions from random forest")   +
-  scale_colour_brewer(palette="Paired")
-@
-\caption{\label{fig:tree-forest}Scatterplot of prediction results from the tree (horizontal) and the forest (vertical).
-Each point corresponds to a land-to-land comparison.
-Known matches are indicated in blue.
-Matches predicted by the decision tree are on the right half of the plot, matches predicted by the random forest are in the upper half.
-}
-\end{figure}
+The maximal consecutively matching striae (cms) is found only in seventh position.
 
 \paragraph{Acknowledgment} Thanks to the men and women behind the software R\cite{R}, and the authors of the R packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
+
+This work was funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE).
 \section{Conclusion}

### 2015-12-11 imaging-paper.Rnw
some more info on IBIS

@@
 
-\paragraph{Acknowledgment} Thanks to the men and women behind the software R\cite{R}, and the authors of the R packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
-
-This work was funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE).
 \section{Conclusion}
@@
 \item reducing subjectivity: `the concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner' \cite{miller:1998}.
+
+\item The analysis by Miller \cite{miller:1998} involved comparisons of 50 pairs of bullets against IBIS results (at the time consisting of 1,600 bullets of the same characteristics).
+
+The feasibility of a database of ballistic images was evaluated in a 2008 report \cite{nap:2008} by the National Research Council.
+Part of the evaluation was based  on Integrated Ballistics Identification System (IBIS) -- to be precise, on the IBIS Heritage Line, which is based on 2d microscopic images.
+The findings of the report are
+
+While microscopy allows some 3d effects (using different lighting and the innate ability of our eyes to move focus), in images this is reduced to a color shade of a pixel at best.
+This means that out of the top six features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match, which vastly reduces the quality of automated matches.
+
 \item involve multiple crosscuts for each bullet in a match?
@@
 \end{enumerate}
+
+\paragraph{Acknowledgment} Thanks to the men and women behind the software R\cite{R}, and the authors of the R packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
+
+This work was funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE).
+
 \bibliographystyle{abbrv}

### 2015-12-11 imaging-paper.Rnw
conclusion

--- 2015-12-10 imaging-paper.Rnw
@@
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches  \cite{nichols:2003}.
+For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches  \cite{nichols:1997, nichols:2003}.
 \end{enumerate}
@@
 
-\paragraph{Acknowledgment} Thanks to the men and women behind the software R\cite{R}, and the authors of the R packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
-
-This work was funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE).
 \section{Conclusion}
 
-\hh{Things we discussed}
-\begin{enumerate}
-\item the Hamby study only includes 35 bullets - this is not a particularly realistic scenario for automatic matching. Adding more non-matches into the mix is going to make things more realistic. 
-\item The matching algorithm is sensitive to the parameter choices made.
+We have presented ...
+\hh{XXX Quick summary of the main points of the paper XXX }
+
+The matching algorithm is sensitive to the parameter choices made.
 The distance between crosscuts (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
@@
 While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.
-\item reducing subjectivity: `the concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner' \cite{miller:1998}.
-\item involve multiple crosscuts for each bullet in a match?
-\item we get probabilities for matches. 
-\item We don't have a solution for subclass characteristics
-\end{enumerate}
+
+ The Hamby study serves as our evaluation `database'.
+It consists of only  35 bullets -- this is obviously not a particularly realistic scenario for automatic matching.
+But this is what at the moment  is available to us.
+As the NIST database expands, this might get into more challenging territory.
+
+The feasibility of a database of ballistic images was evaluated in a 2008 report \cite{nap:2008} by the National Research Council.
+The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
+The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found.
+The remaining problem is simply that statistically the quality of the matching algorithm could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
+The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks. 
+The report also found, that results from 2d images are improved upon by matches based on 3d images.
+This is consistent with the importance of features found here: out of the top six features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.
+
+By suggesting an automated algorithm that first identifies lands, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
+[T]`he concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner' \cite{miller:1998}.
+For a fair assessment of the adequacy of an algorithm we need transparency.
+Our matching  algorithm is open: the code is readily available in form of the R package x3rplus \cite{x3prplus}, and the code to produce this paper is available at \hh{XXX, Eric ...
+we need a website}.
+Obviously, this is only a first step, but in times of the Netflix challenge and the many Data Science challenges hosted by Kaggle, we should not let the quality of the matching algorithm be the problem that stands in the way.
+
+Only a much larger database and transparent communication of results will give us  answers to whether automatic matching of ballistic evidence is really feasible.
+Whether there really is individualization of this evidence, as is often silently assumed.
+And in the case, that individualization is a reasonable assumption, whether  there is a way to be able to distinguish between individual and subclass characteristics.
 

### 2015-12-11 imaging-paper.Rnw
little things

@@
 Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar.
-Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence \eh{XXX we probably need a reference here}.
-\hh{XXX have you tried the Exoneration Registry at michigan? http://www.law.umich.edu/special/exoneration/Pages/about.aspx}
+Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence \cite{giannelli:2011}.
 
@@
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
-\hh{In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.
-Both of these estimates are consistent with a 9 mm bullet fired in a Ruger P-85.}
+In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.
+Both of these estimates are consistent with a 9 mm bullet fired in a Ruger P-85.
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
 For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.
-
-%\hh{XXX add parameter estimates? $\hat{r} \approx 4500$ - indicating a 9 mm bullet, amount of degrees covered by land?  }
 
@@
 We call the location of this peak $p_{left}$ (see Figure~\ref{fig:loess_step47}).
-    %\hh{XXX By peak, do you mean the location of the peak or its height or both?} XXX never mind, found the answer below
     \item Similarly, determine the location of the valley of the left groove by finding the first double-smoothed $y_j$ that is the minimum within its smoothing window.
@@
 As mentioned, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a Loess regression, we selected a smoothing factor of 35.
-We can investigate the performance of groove detection across different smoothing factors.
-Figure \ref{fig:varysmooth}...
+In particular, for detecting and removing the grooves prior to fitting a Loess regression, and for detecting the peaks/valleys of the loess residuals, we selected a smoothing factor of 35.
+We can investigate the performance of peak/valley detection across different smoothing factors.
+Figure \ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 35, 5, and 65 respectively.
+It is immediately clear that a small smoothing factor like 5 is a poor choice.
+This results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
+Given that the striation patterns are typically much larger, we're in effect muddying the waters by performing such minimal smoothing.
+
+A larger smoothing factor on the other hand, like 65, seems to be a more plausible option.
+Most of the peaks/valleys present which are detected by a smoothing factor of 35 are also detected at 65.
+However, some notable issues arise.
+Notice that the peak on the left hand side of the image is smoothed out, and thus not detected.
+On the right hand side, the peak is just barely still present and hence detected.
+But in many cases, the large windows are smoothing out some of the structure that we wish to see.
+Furthermore, it can be seen that the peaks/valleys are often shifted relative to their position in the original loess residuals, or in the smoothed data with smaller smoothing factors.
 
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{Groove detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
+\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
 <<smoothfac1, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
-get_grooves(br111_bullet, smoothfactor = 35, smoothplot = TRUE)$plot
-@
-    }
-\end{subfigure}    
-\begin{subfigure}[t]{\linewidth}
-    \caption{Groove detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
+br111.groove <- get_grooves(br111_bullet)
+br111.loess <- fit_loess(br111_bullet, br111.groove)
+get_peaks(br111.loess$data, smoothfactor = 35)$plot
+@
+    }
+\end{subfigure}    
+\begin{subfigure}[t]{\linewidth}
+    \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
 <<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
-get_grooves(br111_bullet, smoothfactor = 5, smoothplot = TRUE)$plot
-@
-    }
-\end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-    \caption{Groove detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
+get_peaks(br111.loess$data, smoothfactor = 5)$plot
+@
+    }
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+    \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
 <<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
-get_grooves(br111_bullet, smoothfactor = 65, smoothplot = TRUE)$plot
-@
-    }
-\end{subfigure}
-\caption{\label{fig:varysmooth} Groove detection at smoothing factors of 35, 5, and 65 respectively.
-Note that a smoothing factor of 5 yields enough variation prior to the groove that the right groove is not detected properly, while a smoothing factor of 65 over-smoothes and causes the peak on the right to disappear, so that the peak and groove are both not detected properly.}
+get_peaks(br111.loess$data, smoothfactor = 65)$plot
+@
+    }
+\end{subfigure}
+\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 35, 5, and 65 respectively.
+Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 65 over-smoothes and causes the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
 \end{figure}
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the lowest crosscut level in this region \hh{(assumption here: lower crosscut levels show better expression of individual marks XXX need to still check)}
+We first identify a stable region for each bullet and pick the lowest crosscut level in this region, because typically, individual characteristics are best express at the lower end of the bullet. 
+%\hh{(assumption here: lower crosscut levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
 

### 2015-12-11 imaging-paper.Rnw
citations connected to previous word

@@
 
-Rough outline:
-\begin{enumerate}
-\item DONE Description of the problem: Standard operating practice, 2009 NAS report
-\item DONE Description of the Results
-\item DONE Walkthrough of the process
-\item \hh{XXX Alternative approaches: IBIS database and the NRC report\cite{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
-NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
-\end{enumerate}
+% Rough outline:
+% \begin{enumerate}
+% \item DONE Description of the problem: Standard operating practice, 2009 NAS report
+% \item DONE Description of the Results
+% \item DONE Walkthrough of the process
+% \item \hh{XXX Alternative approaches: IBIS database and the NRC report\cite{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
+% NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
+% \end{enumerate}
 
@@
 Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar.
-Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence \cite{giannelli:2011}.
-
-The new scrutiny of the procedure culminated with the 2009 National Academy of Sciences report \cite{NAS:2009}.
+Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\cite{giannelli:2011}.
+
+The new scrutiny of the procedure culminated with the 2009 National Academy of Sciences report~\cite{NAS:2009}.
 The report concluded that minimal research had been done exploring the reliability of these ballistic identification methods.
@@
 Rifling in the barrel creates striation marks on the bullet during the firing process (and grooves).
-These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article \cite{afte:1992}.
+These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article~\cite{afte:1992}.
 ``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows." %\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}
@@
 \section{Methodology}
-Throughout this paper, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study \cite{hamby:2009}.
+Throughout this paper, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\cite{hamby:2009}.
 Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
@@
 That is, the smoothed data is once again smoothed by computing a new rolling average with the same smoothing factor.
-This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass and then added back to the result \cite{tukey:1977}.
+This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass and then added back to the result~\cite{tukey:1977}.
 This has the effect of introducing a little bit more variance into the smoothed data.
@@
 We have implemented a web interface to automate many of the steps outlined in the previous section.
-The web interface, developed using Shiny \cite{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
-A 3D viewer, using Plotly \cite{plotly}, allows the user to rotate and zoom to examine surface features of the images.
+The web interface, developed using Shiny~\cite{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
+A 3D viewer, using Plotly~\cite{plotly}, allows the user to rotate and zoom to examine surface features of the images.
 A screenshot of the application is available in Figure \ref{fig:app}.
@@
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti \cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches  \cite{nichols:1997, nichols:2003}.
+For lead bullets, such as this, Biasotti~\cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches ~\cite{nichols:1997, nichols:2003}.
 \end{enumerate}
@@
 Known matches share on average 8.2 maximal CMS, with a standard deviation of 5.8.
-Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller \cite{miller:1998}.
+Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller~\cite{miller:1998}.
 Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share fifteen consecutively matched striae.
@@
 
-Using recursive partitioning, we fit a decision tree \cite{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
+Using recursive partitioning, we fit a decision tree~\cite{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
 The resulting tree is shown in Figure~\ref{fig:tree}.
@@
 
-Another benefit is that we can not only apply a single decision tree, but several hundred of them to combine in a random forest \cite{breiman:2001, randomForest}.
+Another benefit is that we can not only apply a single decision tree, but several hundred of them to combine in a random forest~\cite{breiman:2001, randomForest}.
 In a random forest multiple trees are fitted.
@@
 
-The feasibility of a database of ballistic images was evaluated in a 2008 report \cite{nap:2008} by the National Research Council.
+The feasibility of a database of ballistic images was evaluated in a 2008 report~\cite{nap:2008} by the National Research Council.
 The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
@@
 By suggesting an automated algorithm that first identifies lands, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
-[T]`he concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner' \cite{miller:1998}.
+[T]`he concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner'~\cite{miller:1998}.
 For a fair assessment of the adequacy of an algorithm we need transparency.
-Our matching  algorithm is open: the code is readily available in form of the R package x3rplus \cite{x3prplus}, and the code to produce this paper is available at \hh{XXX, Eric ...
+Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \hh{XXX, Eric ...
 we need a website}.
@@
 
-\paragraph{Acknowledgment} Thanks to the men and women behind the software R\cite{R}, and the authors of the R packages  knitr \cite{knitr} and ggplot2 \cite{ggplot2}.
+\paragraph{Acknowledgment} Thanks to the men and women behind the software R~\cite{R}, and the authors of the R packages  knitr~\cite{knitr} and ggplot2~\cite{ggplot2}.
 

### 2015-12-11 imaging-paper.Rnw
Add question
iadded references

@@
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti~\cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches ~\cite{nichols:1997, nichols:2003}.
+For lead bullets, such as this, Biasotti~\cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches~\cite{nichols:1997, nichols:2003, nichols:2003b}.
 \end{enumerate}
@@
 Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \hh{XXX, Eric ...
-we need a website}.
-Obviously, this is only a first step, but in times of the Netflix challenge and the many Data Science challenges hosted by Kaggle, we should not let the quality of the matching algorithm be the problem that stands in the way.
+we need a website} \eh{XXX can we just put the GitHub link here?} \hh{github is fine!}.
+Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle, we should not let the quality of the matching algorithm be the problem that stands in the way.
 

### 2015-12-11 imaging-paper.Rnw
some conclusion work

@@
 
-We have presented ...
-\hh{XXX Quick summary of the main points of the paper XXX }
+We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a ballistics identification perspective, removes these features, and produces residuals which allows for the easy identification of markings.
+We've generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches.
+We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features.
+Finally, we've provided an easy-to-use open-source backend, and web frontend to many of these features.
 
@@
 For a fair assessment of the adequacy of an algorithm we need transparency.
-Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \hh{XXX, Eric ...
-we need a website} \eh{XXX can we just put the GitHub link here?} \hh{github is fine!}.
+Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
 Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle, we should not let the quality of the matching algorithm be the problem that stands in the way.

### 2015-12-11 imaging-paper.Rnw
color scheme

@@
 \begin{figure}[hbtp]
-\centering
-\begin{subfigure}[t]{\linewidth}
-\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
+%\centering
+%\begin{subfigure}[t]{\linewidth}
+%\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=6, out.width='\\linewidth', warning=FALSE>>=
 br111.groove <- get_grooves(br111_bullet)
 br111.loess <- fit_loess(br111_bullet, br111.groove)
-get_peaks(br111.loess$data, smoothfactor = 35)$plot
-@
-    }
-\end{subfigure}    
-\begin{subfigure}[t]{\linewidth}
-    \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
-<<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
-get_peaks(br111.loess$data, smoothfactor = 5)$plot
-@
-    }
-\end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-    \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
-<<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
-get_peaks(br111.loess$data, smoothfactor = 65)$plot
-@
-    }
-\end{subfigure}
+
+peaks1 <- get_peaks(br111.loess$data, smoothfactor = 15)
+peaks1$lines$smoothfactor = 15
+peaks1$plot$data$smoothfactor = 15
+peaks2 <- get_peaks(br111.loess$data, smoothfactor = 5)
+peaks2$lines$smoothfactor = 5
+peaks2$plot$data$smoothfactor = 5
+peaks3 <- get_peaks(br111.loess$data, smoothfactor = 35)
+peaks3$lines$smoothfactor = 35
+peaks3$plot$data$smoothfactor = 35
+
+peaks <- rbind(peaks1$lines, peaks2$lines, peaks3$lines)
+smooths <- rbind(peaks1$plot$data, peaks2$plot$data, 
+                 peaks3$plot$data)
+
+  ggplot() + theme_bw() +
+    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
+              data=peaks,  alpha=0.2) +
+    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks) +
+    scale_colour_brewer(palette="Set2") + 
+    scale_fill_brewer(palette="Set2") +
+    theme(legend.position="none") + 
+    facet_grid(smoothfactor~., labeller="label_both") +
+    geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
+              colour="grey50") +
+    geom_line(aes(x=y, y=smoothed), data=smooths) +
+  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
+@
+%'     }
+%' \end{subfigure}    
+%' \begin{subfigure}[t]{\linewidth}
+%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
+%' <<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
+%' get_peaks(br111.loess$data, smoothfactor = 5)$plot
+%' @
+%'     }
+%' \end{subfigure}
+%' \begin{subfigure}[t]{\linewidth}
+%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
+%' <<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
+%' get_peaks(br111.loess$data, smoothfactor = 65)$plot
+%' @
+%'     }
+%' \end{subfigure}
 \caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 35, 5, and 65 respectively.
@@
 [T]`he concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner'~\cite{miller:1998}.
+
 For a fair assessment of the adequacy of an algorithm we need transparency.
 Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
-Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle, we should not let the quality of the matching algorithm be the problem that stands in the way.
+Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home:\url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
 

### 2015-12-12 imaging-paper.Rnw
I've updated the results file, I'll update the rest once Lizzie is sleeping

@@
 
-CCFs <- read.csv("data-35-25/bullet-stats.csv")
+CCFs <- read.csv("data-15-25/bullet-stats.csv")
 includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
@@
 %
-After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0056.
-This is a weighted average between false positives of 0.00028 and false negatives of 0.32.
+After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0049.
+This is a weighted average between false positives of 0.00028 and false negatives of 0.2778.
 Note that these error rates are based on land-to-land comparisons and will be much lower for bullet-to-bullet matches.
@@
 Both results for the tree as well as the random forest are shown.
-For each bullet at least two lands are correctly matched.
+For each bullet at least four lands are correctly matched.
 Unknown bullets B and M are the hardest for the algorithm to match.}

### 2015-12-12 code/evaluation-hamby.R

@@ -1,4 +1,4 @@
-span <- 35
+span <- 15
 dataStr <- sprintf("data-%d-25", span)
 datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
 datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
@@ -105,7 +105,7 @@ CCFs <- read.csv(file.path(dataStr, "bullet-stats.csv"))
 
 qplot(factor(cms), data=CCFs)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(cms), fill=match), position="fill")
-sum(CCFs$cms >= 16)
+sum(CCFs$cms >= 13)
 ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match))
 ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=ccf, colour=match)) + facet_wrap(~match)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(non_cms), fill=match), position="fill")
@@ -123,8 +123,8 @@ prp(rp1, extra = 101)
 
 qplot(ccf, num.matches, geom="jitter", data=CCFs, colour=match, alpha=0.1) + facet_wrap(~match)
 
-CCFs$pred35 <- predict(rp1)
-xtabs(~(pred35>.5)+match, data=CCFs)
+CCFs$pred <- predict(rp1)
+xtabs(~(pred>.5)+match, data=CCFs)
 
 # the ccf/num.matches distribution of non-matches looks like a normal distribution - 
 # let's find a mean and compute (a standardized) distance from that

### 2015-12-12 imaging-paper.Rnw
changes made to smooth factor of 15

@@
 options(replace.assign=TRUE,scipen=3, digits=2)
+bstats <- read.csv("data-15-25/bullet-stats.csv")
 @
@@
 %\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below}
-An opportunity that arises from comparing multiple crosscuts of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
-In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
-This process is shown in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
-`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9. 
+
 \begin{figure}[hbtp]
@@
 \end{figure}
-\afterpage{\clearpage}
+An opportunity that arises from comparing multiple crosscuts of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
+In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
+This process is shown in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
+`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9.
 
@@
 We can investigate the performance of peak/valley detection across different smoothing factors.
-Figure \ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 35, 5, and 65 respectively.
+Figure \ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 15, and 35 respectively.
 It is immediately clear that a small smoothing factor like 5 is a poor choice.
 This results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
-Given that the striation patterns are typically much larger, we're in effect muddying the waters by performing such minimal smoothing.
-
-A larger smoothing factor on the other hand, like 65, seems to be a more plausible option.
-Most of the peaks/valleys present which are detected by a smoothing factor of 35 are also detected at 65.
+Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
+
+A larger smoothing factor on the other hand, like 35, seems to be a more plausible option.
+Most of the peaks/valleys present which are detected by a smoothing factor of 15 are also detected at 35.
 However, some notable issues arise.
-Notice that the peak on the left hand side of the image is smoothed out, and thus not detected.
-On the right hand side, the peak is just barely still present and hence detected.
-But in many cases, the large windows are smoothing out some of the structure that we wish to see.
+Notice that the valley on the right hand side of the image is smoothed out, and thus not detected.
+On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$ or the triple-peak pattern around $y=1,100$.
+That is, in many cases, large windows are smoothing out some of the structure that we wish to see.
 Furthermore, it can be seen that the peaks/valleys are often shifted relative to their position in the original loess residuals, or in the smoothed data with smaller smoothing factors.
@@
 %\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=6, out.width='\\linewidth', warning=FALSE>>=
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='\\linewidth', warning=FALSE>>=
 br111.groove <- get_grooves(br111_bullet)
@@
 we are running $15 \times 6 = 90$ lands from unkown bullets against $2 \times 10 \times 6 = 120$ lands from known bullets, yielding a total of $10,800$ comparisons with 180 known land-to-land matches.
- When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land F-3 and all 120 known lands.
+ When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land D-5 and all 120 known lands.
 Two lands show a high CMS.
@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Maximal number of CMS between unknown Bullet F land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown Bullet D land 5 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
-load("data-35-25/unkn27.RData")
+load("data-15-25/unkn17.RData")
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Overlaid lines of the land of the unknown bullet F-3 and the land with the top matching CMS.}
+\caption{Overlaid lines of D-5 and the land with the top matching CMS.}
 <<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
@@
 #res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
+
+print(ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
+  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
+  scale_linetype_discrete("") + ylab("") +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+
+@
+
+\caption{Overlaid lines of D-5 and the land with the second top matching CMS.}
+<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
+res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
+
+print(ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
+  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
+  scale_linetype_discrete("") + ylab("") +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+
+@
+\end{subfigure}
+\caption{\label{fig:hamby-perfect}Showcase scenario  when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
+\end{figure}
+
+Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
+All 36 pairs of lands with at least sixteen CMS in common are matches.
+There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
+On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}.
+Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}.
+Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller~\cite{miller:1998}.
+Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae.
+Visually we can easily tell these two lands do not match well.
+%
+\begin{figure}[hbtp]
+\begin{subfigure}[t]{\linewidth}
+\caption{Distribution of maximal CMS.}
+<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
+ggplot(data=bstats) + 
+  geom_bar(aes(x=factor(cms))) + theme_bw() +
+  theme(legend.position="bottom") + 
+  xlab("maximum CMS")
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{All land-to-land comparisons a maximum number of CMS of thirteen or more are matches.}
+<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
+ggplot(data=bstats) + 
+  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
+  theme(legend.position="bottom") + 
+  scale_fill_brewer(palette="Paired") +
+  xlab("maximum CMS")
+@
+\end{subfigure}
+\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
+\end{figure}
+
+\begin{figure}[hbtp]
+<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\linewidth', warning=FALSE>>=
+load("data-15-25/unkn37.RData")
+res <- reslist[[60]]
 
@@
   geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-
-@
-
-\caption{Overlaid lines of the land of the unknown bullet F-3 and the land with the second top matching CMS.}
-<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
-#res <- reslist[[which.max(cmsdist)]]  
-res <- reslist[[11]]  # number 2
-
-print(ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
-#  geom_hline(yintercept = res$threshold) +
-#  geom_hline(yintercept = - res$threshold) +
-#  scale_colour_brewer("",palette="Set1") +
-  scale_linetype_discrete("") + ylab("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-
-@
-\end{subfigure}
-\caption{\label{fig:hamby-perfect}Showcase scenario  when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
-\end{figure}
-
-Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
-All 29 pairs of lands with at least sixteen CMS in common are matches.
-There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
-On average, there are 2.2 maximal CMS between known non-matches (with a standard deviation of 1.4).
-Known matches share on average 8.2 maximal CMS, with a standard deviation of 5.8.
-Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller~\cite{miller:1998}.
-Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share fifteen consecutively matched striae.
-The number of peaks and valleys, their heights and depths as well as the overall pattern are in relative good agreement.
-\hh{XXX something re-assuring that visually we can tell them apart?
-I don't know - this could just as well be a case where the crosscuts are very different.
-}
-%
-\begin{figure}[hbtp]
-\begin{subfigure}[t]{\linewidth}
-\caption{Distribution of maximal CMS.}
-<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
-bstats <- read.csv("data-35-25/bullet-stats.csv")
-ggplot(data=bstats) + 
-  geom_bar(aes(x=factor(cms))) + theme_bw() +
-  theme(legend.position="bottom") + 
-  xlab("maximum CMS")
-@
-\end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-\caption{All land-to-land comparisons a maximum number of CMS of sixteen or more are matches.}
-<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
-ggplot(data=bstats) + 
-  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
-  theme(legend.position="bottom") + 
-  scale_fill_brewer(palette="Paired") +
-  xlab("maximum CMS")
-@
-\end{subfigure}
-\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
-\end{figure}
-
-\begin{figure}[hbtp]
-<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\linewidth', warning=FALSE>>=
-load("data-35-25/unkn82.RData")
-res <- reslist[[97]]
-
-print(ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
-#  geom_hline(yintercept = res$threshold) +
-#  geom_hline(yintercept = - res$threshold) +
-#  scale_colour_brewer("",palette="Set1") +
-  scale_linetype_discrete("") + ylab("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-@
-\caption{\label{fig:mismatch}Known mismatch with a surprisingly large number of maximal consecutive matching striae (fifteen) and a small distance between the crosscuts. }
+@
+\caption{\label{fig:mismatch}Known mismatch with a relatively large number of maximal consecutive matching striae (twelve) in the middle. Visual inspection easily reveals this as mismatch. }
 \end{figure}
@@
 Out of the 48 pairs of land-to-land crosscuts with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of crosscuts that both are very flat with peaks and valleys of less than 1.5$\mu m$).
-The average distance between known matches in the Hamby study is $1.9\mu m$ with a standard deviation of $2.3\mu m$; while known mismatches have an average  distance of  3.3$\mu m$ with a standard deviation of $1.9\mu m$.
+The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$distr.dist[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$distr.dist[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$distr.dist[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$distr.dist[!bstats$match])}\mu m$.
 Both of these additional variables show large, if not significant, differences between matches and non-matches.
@@
 A total of 123 lands is being matched correctly.
-Interestingly, the number of consecutive matching striae only features as the last step in this evaluation.
-At this point in the tree, the first false positive matches are introduced into the results: out of the ten land-to-land comparisons three are mis-matched.
+Interestingly, the number of consecutive matching striae does not feature in this evaluation.
 Instead of CMS, cross-correlation (ccf) between the crosscuts is very important in the matching process  by the decision tree.
-Aside from  cross correlation, the number of matches (num.matches) between crosscuts and the distance between crosscuts (distr.dist) is involved in the decision.
+Aside from  cross correlation, the number of matches (num.matches) and mismatches (mis.matches) is involved in the decision.
 Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
@@
 
-CCFs <- read.csv("data-15-25/bullet-stats.csv")
-includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
-rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
+includes <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
+rp1 <- rpart(match~., bstats[,includes])  # doesn't include cms at all !!!!
 
@@
 
-CCFs$pred <- predict(rp1)
+bstats$pred <- predict(rp1)
 @
@@
 library(reshape2)
-rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
+rtrees <- randomForest(factor(match)~., data=bstats[,includes], ntree=300)
 
@@
 errors$id <- 1:nrow(errors)
-CCFs$forest <- predict(rtrees, type="prob")[,2]
+bstats$forest <- predict(rtrees, type="prob")[,2]
 @
@@
 <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
-CCFs$bullet <- gsub("-[0-9]$", "", CCFs$b2)
-bullets <- CCFs %>% group_by(bullet) %>% summarize(
+bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
+bullets <- bstats %>% group_by(bullet) %>% summarize(
   n = n(),
@@
 \caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree.
-Both results for the tree as well as the random forest are shown.
+Both results for the tree and the random forest are shown.
 For each bullet at least four lands are correctly matched.
@@
 Matching bullets can only be tested realistically in a much bigger test case.
-Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.
-whenever the predicted probability of a match exceeds 0.5, a match is declared.
+Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared.
 Basing this decision on 0.5 is not necessary.
@@
 Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis).
-Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only 13 cases result in a different decision.
+Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only 12 cases result in a different decision.
 
@@
 <<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
-qplot(pred, forest, geom="jitter", data=CCFs, colour=match) + theme_bw() +
+qplot(pred, forest, geom="jitter", data=bstats, colour=match) + theme_bw() +
   theme(legend.position="bottom") +
@@
 imp$meanMatch <- sapply(imp$Variable, function(var) {
-  mean(subset(CCFs, match)[,var])
+  mean(subset(bstats, match)[,var])
 })
 imp$sdMatch <- sapply(imp$Variable, function(var) {
-  sd(subset(CCFs, match)[,var])
+  sd(subset(bstats, match)[,var])
 })
 imp$meanNonMatch <- sapply(imp$Variable, function(var) {
-  mean(subset(CCFs, !match)[,var])
+  mean(subset(bstats, !match)[,var])
 })
 imp$sdNonMatch <- sapply(imp$Variable, function(var) {
-  sd(subset(CCFs, !match)[,var])
+  sd(subset(bstats, !match)[,var])
 })
@@
 The rows in the table are ordered according to descending importance.
-The variable with the most predictive power by far is cross-correlation (ccf), followed by the number of matches (num.matches) and the Euclidean distance between the crosscuts (distr.dist).
-The maximal consecutively matching striae (cms) is found only in seventh position.
+The variable with the most predictive power by far is cross-correlation (ccf), followed by the number of matches (num.matches) and the sum of the height of joint peaks (sumpeaks).
+The maximal consecutively matching striae (cms) is found only in sixth position.
 
@@
 We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a ballistics identification perspective, removes these features, and produces residuals which allows for the easy identification of markings.
-We've generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches.
+We have generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches.
 We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features.
-Finally, we've provided an easy-to-use open-source backend, and web frontend to many of these features.
+Finally, we have provided an easy-to-use open-source backend, and web frontend to many of these features.
 
@@
 The distance between crosscuts (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
-Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 57.8375$\mu m$ is used, corresponding to a window of 17 values to the left and the right of an observation).
+Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
 While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.

### 2015-12-12 imaging-paper.Rnw
figures moved around a bit, but looks better now

@@
 \usepackage{float}
-\usepackage{afterpage}
+%\usepackage{afterpage}
 
@@
     \renewcommand{\bottomfraction}{0.99}	% max fraction of floats at bottom
+% parameters for text pages
+\setcounter{topnumber}{2}
+    \setcounter{bottomnumber}{2}
+    \setcounter{totalnumber}{4}     % 2 may work better
+    \setcounter{dbltopnumber}{2}    % for 2-column pages
+    \renewcommand{\dbltopfraction}{0.9}	% fit big float above 2-col. text
+    \renewcommand{\textfraction}{0.07}	% allow minimal text w. figs
+    %   Parameters for FLOAT pages (not text pages):
+    \renewcommand{\floatpagefraction}{0.7}	% require fuller float pages
+	% N.B.: floatpagefraction MUST be less than topfraction !!
+    \renewcommand{\dblfloatpagefraction}{0.7}	% require fuller float pages
 
@@
 \end{figure}
-\afterpage{\clearpage}
+%\afterpage{\clearpage}
 
 \subsection{Varying smoothing factor}
-As mentioned, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a Loess regression, and for detecting the peaks/valleys of the loess residuals, we selected a smoothing factor of 35.
+As mentioned earlier, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
+In particular, for detecting and removing the grooves prior to fitting a Loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 15 seems more appropriate.
 We can investigate the performance of peak/valley detection across different smoothing factors.
-Figure \ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 15, and 35 respectively.
+Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 15, and 35, respectively.
+The dark line corresponds to the smoothed values, while the grey line in the back shows the unsmoothed values of the surface measurements.
+The choice of smoothing factor is a classical decision of a bias/variance trade-off.
 It is immediately clear that a small smoothing factor like 5 is a poor choice.
-This results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
+It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
 Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
@@
 %' \end{subfigure}
-\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 35, 5, and 65 respectively.
-Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 65 over-smoothes and causes the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
+\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 15, and 35, respectively.
+Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 35 might over-smoothe and cause the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
 \end{figure}
@@
 
-\hh{XXX next step: how does this matching work in known non-matches? distribution of CMS under a null of no match? }
+In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
 
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{All land-to-land comparisons a maximum number of CMS of thirteen or more are matches.}
+\caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
 <<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=

### 2015-12-12 imaging-paper.Rnw
little edits

@@
               data=peaks,  alpha=0.2) +
-    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks) +
+    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
+               data= peaks, alpha=0.7) +
     scale_colour_brewer(palette="Set2") + 
@@
 \begin{figure}[hbtp]
-%\centering
-%\begin{subfigure}[t]{\linewidth}
-%\caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 35.\label{fig:smoothfac1}}{%
 <<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='\\linewidth', warning=FALSE>>=
@@
               data=peaks,  alpha=0.2) +
-    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks) +
+    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
     scale_colour_brewer(palette="Set2") + 
@@
 @
-%'     }
-%' \end{subfigure}    
-%' \begin{subfigure}[t]{\linewidth}
-%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 5.\label{fig:smoothfac2}}{%
-%' <<smoothfac2, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
-%' get_peaks(br111.loess$data, smoothfactor = 5)$plot
-%' @
-%'     }
-%' \end{subfigure}
-%' \begin{subfigure}[t]{\linewidth}
-%'     \caption{Peak and valley detection for bullet 1-5 at a smoothing factor of 65.\label{fig:smoothfac3}}{%
-%' <<smoothfac3, echo=FALSE, fig.width=7, fig.height=3, out.width='\\textwidth', warning=FALSE>>=
-%' get_peaks(br111.loess$data, smoothfactor = 65)$plot
-%' @
-%'     }
-%' \end{subfigure}
 \caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 15, and 35, respectively.
@@
 @
-\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree.
+\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12).
 Both results for the tree and the random forest are shown.
@@
 <<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
-qplot(pred, forest, geom="jitter", data=bstats, colour=match) + theme_bw() +
+qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
+      shape=match) + theme_bw() +
   theme(legend.position="bottom") +

### 2015-12-12 code/complete-hamby.R
Speed up hamby

@@ -6,10 +6,7 @@ library(gridExtra)
 library(zoo)
 
 knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
-
 unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
-unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
 span <- 35
 dataStr <- sprintf("data-%d-25", span) # using crosscuts-25.csv
@@ -17,20 +14,91 @@ dataStr <- sprintf("data-%d-25", span) # using crosscuts-25.csv
 ###############
 # can we identify the barrels the unknown bullets came from?
 
+new_get_bullet <- function(path, x = 243.75, bullet = NULL) {
+    if (is.null(bullet)) bullet <- read.x3p(path)
+    dbr111 <- fortify_x3p(bullet)
+    
+    pickx <- dbr111$x[which.min(abs(x - unique(dbr111$x)))]
+    
+    dbr111.fixx <- dbr111[dbr111$x == pickx,]
+    
+    return(dbr111.fixx)
+}
+
+new_processBullets <- function(bullet, x = 100) {
+    crosscuts <- unique(fortify_x3p(bullet)$x)
+    crosscuts <- crosscuts[crosscuts >= min(x)]
+    crosscuts <- crosscuts[crosscuts <= max(x)]
+    if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]
+    
+    list_of_fits <- lapply(crosscuts, function(x) {
+        br111 <- new_get_bullet(path = NULL, x = x, bullet = bullet)
+        br111.groove <- get_grooves(br111)
+        fit_loess(br111, br111.groove)$resid$data
+    })
+    lof <- list_of_fits %>% bind_rows
+    
+    cbind(lof, bullet = bullet$path)
+}
+
+new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
+    lof <- rbind(lof1, lof2)
+    bAlign = bulletAlign(lof)
+    lofX <- bAlign$bullet  
+    
+    b12 <- unique(lof$bullet)
+    peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
+    peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)
+    
[…]

### 2015-12-12 imaging-paper.Rnw
finishing off the conclusion

@@
 
-Things to be dicussed in the future
-\begin{enumerate}
-\item reduce the stretch of crosscuts and match smaller pieces. This allows for both a more realistic scenario, as well as a quantification of the individuality of the segment.
+An approach that might help in determining this difference comes from Biometrics, is suggested by Jain et al~\cite{jain:2015:tphil, jain:2015} and has been successfully applied in the Aadhar project (the effort by the Unique Identification Authority of India (UIDAI) to provide a unique 12-digit identification number to approximately 1.2 billion residents of India based on ten fingerprints and two irises).
+This approach uses dictionary learning based on small square patches  of 64x64 or 32x32 pixels.
+For matching bullets, a simpler, one-dimensional window approach might be sufficient to match striation marks because of the physical forces  leading to the existence of these marks.
+
 \end{enumerate}

### 2015-12-12 imaging-paper.Rnw
and now it's latexing again

@@
 \includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
-\caption{\label{fig:manualmatch}\hh{Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch}Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}\hh{Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}}
+\caption{\label{fig:manualmatch-rgl}Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}
 \end{figure}
@@
 
-\end{enumerate}
-
 \paragraph{Acknowledgment} Thanks to the men and women behind the software R~\cite{R}, and the authors of the R packages  knitr~\cite{knitr} and ggplot2~\cite{ggplot2}.

### 2015-12-13 imaging-paper.Rnw
Switch to using 25 as span

@@
 options(replace.assign=TRUE,scipen=3, digits=2)
-bstats <- read.csv("data-15-25/bullet-stats.csv")
+bstats <- read.csv("data-25-25/bullet-stats.csv")
 @
@@
 \begin{figure}[hbtp]
-<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, warning=FALSE, out.width='\\linewidth'>>=
 bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)

### 2015-12-13 code/evaluation-hamby.R

@@ -181,8 +181,8 @@ prp(rp1, extra = 101, fallen.leaves=TRUE)
 # throw in all three evaluations into the mix:
 
 bstats <- NULL
-for (i in c(35, 30, 25)) {
-  dataStr <- sprintf("data-%d", i)
+for (i in c(5, 10, 15, 20, 25, 30, 35, 40)) {
+  dataStr <- sprintf("data-%d-25", i)
   temp <- read.csv(file.path(dataStr, "bullet-stats.csv"))
   includes <- setdiff(names(temp), c("b1", "b2", "data", "resID", "id.x", "id.y"))
   temp$diffx <- with(temp, abs(x1-x2))
@@ -190,9 +190,11 @@ for (i in c(35, 30, 25)) {
   rp <- rpart(match~., data=temp[,includes])
   
   prp(rp, extra = 101)
-  ch <- scan()
+  #ch <- scan()
   temp$pred <- predict(rp)
   temp$span <- i
+  temp$bullet <- NULL
+  temp$crosscutdist <- NULL
   bstats <- rbind(bstats, temp)
 }
 

### 2015-12-13 imaging-paper.Rnw
error rate adjusted and bullet summary plot

--- 2015-12-12 imaging-paper.Rnw
@@
 %
-After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0049.
-This is a weighted average between false positives of 0.00028 and false negatives of 0.2778.
+After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0045.
+This is a weighted average between false positives of 0.000094 and false negatives of 0.27.
 Note that these error rates are based on land-to-land comparisons and will be much lower for bullet-to-bullet matches.
@@
 bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
-bullets <- bstats %>% group_by(bullet) %>% summarize(
+bullets <- bstats %>% filter(match) %>% group_by(bullet) %>% summarize(
   n = n(),

### 2015-12-13 imaging-paper.Rnw
odd mis-match updated

@@
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
-load("data-15-25/unkn17.RData")
+load("data-new-25-25/unkn17.RData")
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
@@
 <<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\linewidth', warning=FALSE>>=
-load("data-15-25/unkn37.RData")
-res <- reslist[[60]]
+load("data-25-25/unkn47.RData")
+res <- reslist[[106]]
+splits <- strsplit(as.character(res$bullets$bullet), split="/")
+last <- sapply(splits, function(x) x[length(x)])
+res$bullets$bullet <- gsub(".x3p","", last)
 
@@
   geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
-#  geom_hline(yintercept = res$threshold) +
-#  geom_hline(yintercept = - res$threshold) +
-#  scale_colour_brewer("",palette="Set1") +
   scale_linetype_discrete("") + ylab("") +
@@
 @
-\caption{\label{fig:mismatch}Known mismatch with a relatively large number of maximal consecutive matching striae (twelve) in the middle. Visual inspection easily reveals this as mismatch. }
+\caption{\label{fig:mismatch}Known mismatch with a relatively large number of maximal consecutive matching striae (twelve) in the middle.
+The pattern in the middle does look surprisingly similar, however the outer ends of the crosscuts easily reveals this as mismatch.
+}
 \end{figure}

### 2015-12-13 code/evaluation-hamby.R

@@ -1,4 +1,4 @@
-span <- 15
+span <- 25
 dataStr <- sprintf("data-new-%d-25", span)
 datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
 datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
@@ -74,7 +74,7 @@ CCFs$span <- span
 
 library(rpart)
 library(rpart.plot)
-includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y"))
+includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "span", "forest"))
 rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
 prp(rp1, extra = 101)
 CCFs$pred <- predict(rp1)

### 2015-12-13 imaging-paper.Rnw
update on bst matches

@@
 bstats <- read.csv("data-15-25/bullet-stats.csv")
+
+scrubPath <- function(x) {
+  splits <- strsplit(as.character(x), split="/")
+  last <- sapply(splits, function(x) x[length(x)])
+  gsub(".x3p","", last)
+}
+
 @
@@
 we are running $15 \times 6 = 90$ lands from unkown bullets against $2 \times 10 \times 6 = 120$ lands from known bullets, yielding a total of $10,800$ comparisons with 180 known land-to-land matches.
- When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land D-5 and all 120 known lands.
+ When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land C-3 and all 120 known lands.
 Two lands show a high CMS.
@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Maximal number of CMS between unknown Bullet D land 5 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown Bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
-load("data-new-25-25/unkn17.RData")
+load("data-25-25/unkn9.RData")
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Overlaid lines of D-5 and the land with the top matching CMS.}
+\caption{Overlaid lines of C-3 and the land with the top matching CMS.}
 <<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
@@
 #res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
+res$bullets$bullet <- scrubPath(res$bullets$bullet)
 
@@
 
-\caption{Overlaid lines of D-5 and the land with the second top matching CMS.}
+\caption{Overlaid lines of C-3 and the land with the second top matching CMS.}
 <<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
-res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
+res <- reslist[[which(cmsdist==14)]]  # number 2
+res$bullets$bullet <- scrubPath(res$bullets$bullet)
 
@@
 res <- reslist[[106]]
-splits <- strsplit(as.character(res$bullets$bullet), split="/")
-last <- sapply(splits, function(x) x[length(x)])
-res$bullets$bullet <- gsub(".x3p","", last)
+res$bullets$bullet <- scrubPath(res$bullets$bullet)
 

### 2015-12-13 imaging-paper.Rnw
set seed for random forest

@@
 Instead of CMS, cross-correlation (ccf) between the crosscuts is very important in the matching process  by the decision tree.
-Aside from  cross correlation, the number of matches (num.matches) and mismatches (mis.matches) is involved in the decision.
+Aside from  cross correlation, the number of matches (num.matches) is involved in the decision.
 Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
@@
 library(reshape2)
+set.seed(20151227)
 rtrees <- randomForest(factor(match)~., data=bstats[,includes], ntree=300)
@@
 %
-After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0045.
+After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0046.
 This is a weighted average between false positives of 0.000094 and false negatives of 0.27.
@@
 Both results for the tree and the random forest are shown.
-For each bullet at least four lands are correctly matched.
-Unknown bullets B and M are the hardest for the algorithm to match.}
+For each bullet at least three lands are correctly matched.
+Unknown bullet B is the hardest for the algorithm to match.}
 \end{figure}
@@
 Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis).
-Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only 12 cases result in a different decision.
+Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only ten cases result in a different decision.
 

### 2015-12-13 code/evaluation-hamby.R

@@ -95,7 +95,7 @@ write.csv(CCFs, file=file.path(dataStr, "bullet-stats.csv"), row.names=FALSE)
 
 idx <- which(CCFs$match & CCFs$cms <= 6)
 idx <- which(!CCFs$match & CCFs$ccf > .7)
-idx <- which(!CCFs$match & CCFs$cms == 15)
+idx <- which(!CCFs$match & CCFs$cms == 12)
 idx <- which(CCFs$cms == 11)
 
 for ( i in idx) {

### 2015-12-13 imaging-paper.Rnw
added missing reference

@@
 We can investigate the performance of peak/valley detection across different smoothing factors.
-Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 15, and 35, respectively.
+Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 25, and 35, respectively.
 The dark line corresponds to the smoothed values, while the grey line in the back shows the unsmoothed values of the surface measurements.
@@
 A larger smoothing factor on the other hand, like 35, seems to be a more plausible option.
-Most of the peaks/valleys present which are detected by a smoothing factor of 15 are also detected at 35.
+Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 35.
 However, some notable issues arise.
 Notice that the valley on the right hand side of the image is smoothed out, and thus not detected.
-On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$ or the triple-peak pattern around $y=1,100$.
+On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$.
 That is, in many cases, large windows are smoothing out some of the structure that we wish to see.
@@
 
-peaks1 <- get_peaks(br111.loess$data, smoothfactor = 15)
-peaks1$lines$smoothfactor = 15
-peaks1$plot$data$smoothfactor = 15
+peaks1 <- get_peaks(br111.loess$data, smoothfactor = 25)
+peaks1$lines$smoothfactor = 25
+peaks1$plot$data$smoothfactor = 25
 peaks2 <- get_peaks(br111.loess$data, smoothfactor = 5)
@@
 @
-\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 15, and 35, respectively.
-Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 35 might over-smoothe and cause the peaks/valleys to either disappear or be horizontally shifted from their position in the original loess residuals.}
+\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 35, respectively.
+Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 35 might over-smooth and cause the peaks/valleys to either disappear or shift  horizontally  from their position in the original loess residuals.}
 \end{figure}
@@
     geom_line(aes(x = y, y = l30, linetype = bullet),  data = lofX, alpha=0.6) +
-#    geom_hline(yintercept = threshold) +
-#    geom_hline(yintercept = - threshold) +
-#    scale_colour_brewer("",palette="Set1") +
     scale_linetype_discrete("") +

### 2015-12-13 imaging-paper.Rnw
Update to use 25-25
Include reference to proper smoothing factor

@@
 options(replace.assign=TRUE,scipen=3, digits=2)
-bstats <- read.csv("data-15-25/bullet-stats.csv")
+bstats <- read.csv("data-25-25/bullet-stats.csv")
 
@@
     b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
 
@@
   b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 35)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 35)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
 
@@
 As mentioned earlier, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a Loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 15 seems more appropriate.
+In particular, for detecting and removing the grooves prior to fitting a Loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
 We can investigate the performance of peak/valley detection across different smoothing factors.
-Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 25, and 35, respectively.
+Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 25, and 45, respectively.
 The dark line corresponds to the smoothed values, while the grey line in the back shows the unsmoothed values of the surface measurements.
@@
 
-A larger smoothing factor on the other hand, like 35, seems to be a more plausible option.
-Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 35.
+A larger smoothing factor on the other hand, like 45, seems to be a more plausible option.
+Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 45.
 However, some notable issues arise.
@@
 peaks2$plot$data$smoothfactor = 5
-peaks3 <- get_peaks(br111.loess$data, smoothfactor = 35)
-peaks3$lines$smoothfactor = 35
-peaks3$plot$data$smoothfactor = 35
+peaks3 <- get_peaks(br111.loess$data, smoothfactor = 45)
+peaks3$lines$smoothfactor = 45
+peaks3$plot$data$smoothfactor = 45
 
@@
 @
-\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 35, respectively.
-Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 35 might over-smooth and cause the peaks/valleys to either disappear or shift  horizontally  from their position in the original loess residuals.}
+\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively.
+Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either disappear or shift  horizontally  from their position in the original loess residuals.}
 \end{figure}

### 2015-12-13 imaging-paper.Rnw
adjusted some of the numbers throughout to match the figures.

@@
 The resulting tree is shown in Figure~\ref{fig:tree}.
-A total of 123 lands is being matched correctly.
+A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} lands is being matched correctly.
 Interestingly, the number of consecutive matching striae does not feature in this evaluation.
@@
 library(reshape2)
-set.seed(20151227)
+set.seed(20151215)
 rtrees <- randomForest(factor(match)~., data=bstats[,includes], ntree=300)
@@
 %
-After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0046.
-This is a weighted average between false positives of 0.000094 and false negatives of 0.27.
-Note that these error rates are based on land-to-land comparisons and will be much lower for bullet-to-bullet matches.
-In the case of the Hamby study, the error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched (see Figure~\ref{fig:treeresult}).
+After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0044.
+This is a weighted average between zero false positives and an error rate of false negatives of 0.26.
+Note that this error rate is based on land-to-land comparisons and will be much lower for bullet-to-bullet matches.
+In the case of the Hamby study, the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched (see Figure~\ref{fig:treeresult}).
 \begin{figure}[hbtp]
@@
 Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis).
-Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only ten cases result in a different decision.
+Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.
 
@@
 The rows in the table are ordered according to descending importance.
-The variable with the most predictive power by far is cross-correlation (ccf), followed by the number of matches (num.matches) and the sum of the height of joint peaks (sumpeaks).
-The maximal consecutively matching striae (cms) is found only in sixth position.
+The variable with the most predictive power by far is cross-correlation (ccf), followed by the number of matches (num.matches) and mismatches(num.mismatches), and the sum of the height of joint peaks (sumpeaks).
+The maximal consecutively matching striae (cms) is found only in sixth place.
 
@@
 The report also found, that results from 2d images are improved upon by matches based on 3d images.
-This is consistent with the importance of features found here: out of the top six features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.
-
-By suggesting an automated algorithm that first identifies lands, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
-[T]`he concept of counting striations is subjective and based on experience, but it does tend to be consistent within a single examiner'~\cite{miller:1998}.
+This is consistent with the importance of features found here: out of the top five features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.
+
+By suggesting an automated algorithm that first removes class characteristics, such as the grooves and the curvature of the bullet to reveal the region of the  land, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
+Counting CMS in particular, has been found to be subjective to the examiner: `the concept of counting striations is subjective and based on experience'~\cite{miller:1998}.
 

### 2015-12-14 imaging-paper.Rnw
Some minor introduction edits

@@
 In 2005, in \emph{United States vs.
-Green}, the court ruled that the expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely describe other casings which are similar.
+Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar.
 Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\cite{giannelli:2011}.
@@
 
-Rifling in the barrel creates striation marks on the bullet during the firing process (and grooves).
+Rifling in the barrel creates striation marks on the bullet during the firing process.
 These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article~\cite{afte:1992}.
-``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows." %\footnote{\url{https://www.ncjrs.gov/app/publications/Abstract.aspx?id=254151}}
-
-The primary goal in this type of forensic analysis is to identify markings in an attempt to match two bullets.
+``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."
+
+The primary goal in this type of forensic analysis is to identify these striation markings in an attempt to match two bullets.
 Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help us determine whether two bullets are a match.
 However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
-We've produced a framework which, in software, allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier for the lay person.
+We've produced a framework which, in software, allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
 

### 2015-12-14 imaging-paper.Rnw
Couple more typos

@@
 Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt.
-The direction of the rifling is a class characteristic.
-
-An initial naive approach leads us to fixing a particular value of a coordinate, and then producing a plot of the heights across values of the other coordinate.
+The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.
+
+An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the heights across values of the other coordinate.
 Figure \ref{fig:fixedX} gives a plot resulting from such a procedure.
-It can be seem that the global structure of the land dominates the appearance of the plot.
+It can be seen that the global structure of the land dominates the appearance of the plot.
 The grooves can be clearly identified on the left and right side, and the curvature of the surface is the most visible feature in the middle.

### 2015-12-14 imaging-paper.Rnw
Rebuild the PDF

@@
 \subsection{Loess Fit}
-In an approach to improve upon the previous algorithm, we first have to eliminate 
-%To improve upon the previous algorithm, the first step must be to eliminate
-the grooves from the images under the assumption that they do not contain relevant information for determining a match.
+In an approach to improve upon the previous algorithm, we first have to eliminate the grooves from the images under the assumption that they do not contain relevant information for determining a match.
 Fortunately, the location and appearance of the grooves in these 2d plots is quite consistent.
-Surface measurements %beginning at $y = 0$ 
-reach local maxima around the peak of the groove at either end of the range of $y$, we can then follow the descent of the surface measurements inwards to the valley of the groove. 
-This marks the point at which we should trim the image. %We repeat a similar procedure for the right-hand groove by starting at the maximum y value and working backwards. 
-The procedure can be described as follows:
+Surface measurements reach local maxima around the peak of the groove at either end of the range of $y$, we can then follow the descent of the surface measurements inwards to the valley of the groove.
+This marks the point at which we should trim the image. The procedure can be described as follows:
 

### 2015-12-15 imaging-paper.Rnw
Remove last few comments, rebuild
typo
references included in the conclusions

@@
 
-\hh{XXX what would be good is to get the line identifications and project them back onto the surface. }
-
 \section{Evaluation}
@@
 Both results for the tree and the random forest are shown.
-For each bullet at least three lands are correctly matched.
+For each bullet at least four lands are correctly matched.
 Unknown bullet B is the hardest for the algorithm to match.}
@@
 The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks.
-The report also found, that results from 2d images are improved upon by matches based on 3d images.
+This finding coincides with the assessment by de Kinder et al \cite{dekinder:2004} based on the IBIS Heritage system.
+A further re-assessment in 2015 by de Ceuster et al \cite{deceuster:2014} came to the same conclusions based on the EvoFinder system.
+The NRC report also found, that results from 2d images are improved upon by matches based on 3d images.
 This is consistent with the importance of features found here: out of the top five features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.

### 2015-12-15 imaging-paper.Rnw
acknowledgement

@@
 This finding coincides with the assessment by de Kinder et al \cite{dekinder:2004} based on the IBIS Heritage system.
-A further re-assessment in 2015 by de Ceuster et al \cite{deceuster:2014} came to the same conclusions based on the EvoFinder system.
+A further re-assessment in 2015 by de Ceuster et al \cite{deceuster:2015} came to the same conclusions based on the EvoFinder system.
 The NRC report also found, that results from 2d images are improved upon by matches based on 3d images.
@@
 
-This work was funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE).
+This research was partially funded through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
 

### 2015-12-15 imaging-paper.Rnw
Introduce terminology for bullet X-Y

@@
 Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
+For notational simplicity, we will refer to a particular land of some bullet as Bullet-X-Y, where X is the bullet number, and Y is the land number.
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.
@@
 The orange rectangle shows an example of the smoothing window.
-Groove and peak will be detected, if they are not within the same window.}
+Grooves and peaks will be detected, if they are not within the same window.}
 <<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
-\caption{\label{fig:manualmatch}Manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}
+\caption{\label{fig:manualmatch}Manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
 \end{figure}
@@
 Both of these situations are assessed and discussed in more detail below:
-Figure~\ref{fig:crosscuts} shows a sequence of crosscuts for bullet land 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$.
+Figure~\ref{fig:crosscuts} shows a sequence of crosscuts for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$.
 These are compared to a crosscut at 100$\mu m$.
@@
 In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
-This process is shown in Figure~\ref{fig:crosscuts2} at the example of land 1-1 from barrel 3.
+This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3.
 `Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9.

### 2015-12-15 imaging-paper.Rnw
Edits from Dr. Carriquiry

@@
 % The first character should be within \initial{}
-\initial{F}\textbf{orensic examiners try to determine whether two guns were fired from the same barrel by analyzing them under a microscope in controlled conditions. Can we make the task simpler?}
+\initial{F}\textbf{orensic examiners try to determine whether two bullets were fired from the same barrel by analyzing them under a microscope in controlled conditions. Can we make the task simpler?}
 
@@
 
-Bullet matching is a process in forensic science done to determine whether two bullets were fired from the same gun barrel.
+Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
 This process has broad applicability in terms of the conviction of criminals in the United States criminal justice system.
@@
 
-The new scrutiny of the procedure culminated with the 2009 National Academy of Sciences report~\cite{NAS:2009}.
-The report concluded that minimal research had been done exploring the reliability of these ballistic identification methods.
-In fact, the report says the forensic evidence ``including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
-
-Rifling in the barrel creates striation marks on the bullet during the firing process.
-These marks are claimed to be unique to the barrel, as described in a 1992 AFTE article~\cite{afte:1992}.
+In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
+~\cite{NAS:2009}.
+The report concluded that minimal research had been done to explore the reliability of these ballistic identification methods.
+In fact, the report states that forensic evidence ``including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
+
+Rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet during the firing process.
+These marks are assumed to be unique to the barrel, as described in a 1992 AFTE article~\cite{afte:1992}.
 ``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."
@@
 However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
-We've produced a framework which, in software, allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
+We propose a framework which allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
 
 \section{Methodology}
-Throughout this paper, we will work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\cite{hamby:2009}.
+Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\cite{hamby:2009}.
 Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
 
-3D topographical images of each bullet were taken using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
+3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
 x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
-x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government Firearm Forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
 
@@
 Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
-For notational simplicity, we will refer to a particular land of some bullet as Bullet-X-Y, where X is the bullet number, and Y is the land number.
-An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}, which show side and top profiles of the land respectively.
+For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
+An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
+These figures show side and top profiles of the land respectively.
 The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process.
@@
 An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the heights across values of the other coordinate.
-Figure \ref{fig:fixedX} gives a plot resulting from such a procedure.
+Figure \ref{fig:fixedX} shows a plot resulting from such a procedure.
 It can be seen that the global structure of the land dominates the appearance of the plot.
@@
 
+The smooth curve on the plot represents a segment of a perfect circle with the same radius as the bullet.
 It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
@@
 
-For this, assume that $n$ data points are given in form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle.
+For this, assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle.
 We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.
@@
 \begin{equation}\label{eq:circle}
-D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2
+D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2,
 \end{equation}
-We do this by differentiating $D$ with respect to $r, a,$ and $b$:
-let us assume that $x_i$ and $y_i$ are centralized (i.e.
+by differentiating $D$ with respect to $r, a,$ and $b$:
+let us assume that $x_i$ and $y_i$ are centered (i.e.
 $\sum x_i = \sum_i y_i = 0$).
@@
 \frac{d}{dr} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 r = \\
-&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right) 
+&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right).
 \end{eqnarray*}
-For a minimum then holds:
+At the minimum:
 \begin{equation}\label{eq:rmin}
-\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2 
+\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2.
 \end{equation}
@@
 \frac{d}{da} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 (x_i - a) = \\
-&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]
+&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right].
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right] 
+\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right].
 \end{eqnarray*}
-For the minimum we therefore get a system of two linear equations in $a$ and $b$:
+To find the minimum we therefore get a system of two linear equations in $a$ and $b$:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-This resolves to 
+The solution to the system is:
 \begin{eqnarray*}
@@
 In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.
-Both of these estimates are consistent with a 9 mm bullet fired in a Ruger P-85.
+Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
@@
 \end{subfigure}
-\caption{\label{fig:residual} Residual structure of circular fits at two different cross-section.
+\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections.
 Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
@@
 A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction.
-Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, we can not address all of these issues.
+Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues.
 While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.
@@
 \subsection{Loess Fit}
-In an approach to improve upon the previous algorithm, we first have to eliminate the grooves from the images under the assumption that they do not contain relevant information for determining a match.
+In an attempt to improve upon the previous algorithm, we first have to eliminate the grooves from the images under the assumption that they do not contain relevant information for determining a match.
 Fortunately, the location and appearance of the grooves in these 2d plots is quite consistent.
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step1}Step 1 the Loess fit algorithm.
+\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm.
 For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
@@
 %\begin{subfigure}[t]{\linewidth}
-% \caption{\label{fig:loess_step2}Step 2 of the Loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
+% \caption{\label{fig:loess_step2}Step 2 of the loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
 <<loess_step2, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step3}Step 2 of the Loess fit algorithm.
+\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm.
 The  height values are smoothed twice with a smoothing factor of 35.
 The orange rectangle shows an example of the smoothing window.
-Grooves and peaks will be detected, if they are not within the same window.}
+Grooves and peaks are detected, if they are not within the same window.}
 <<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_step47}Steps 3 -- 6 of the Loess fitting algorithm.
+\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm.
 After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
@@
 This has the effect of introducing a little bit more variance into the smoothed data.
-We instead performed a second smoothing pass on the smoothed data, which has the effect of weighting observations near the center highest, while the weights linearly drop off as you reach the end of the smoothing window or smoothing factor.
-
-Next, we fit a Loess regression to the data.
-Figure \ref{fig:loess_fit} provides a look at the Loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
+We instead performed a second smoothing pass on the smoothed data, which has the effect of weighting observations near the center highest, while the weights linearly drop off as we reach the end of the smoothing window or smoothing factor.
+
+Next, we fit a loess regression to the data.
+Figure \ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
 The fit seems to do a reasonable job of capturing the structure of the image.
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:loess_resid} Residuals of Loess fit for bullet 1-5.}
+\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
-my.loess$resid + ylab("Residuals of Loess Fit")
+my.loess$resid + ylab("Residuals of loess Fit")
 @
@@
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Bullet 1-5 in black") + ylab("Residuals of Loess Fit")
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Bullet 1-5 in black") + ylab("Residuals of loess Fit")
 @    
@@
 
-The application includes a button which allows the automatic creation of a residual plot from the Loess fits of two bullets.
-An x coordinate value can be selected as a cross cut, and a cross-section of the two bullet at the x value will be taken.
-The previously described Loess regression procedure will be applied to the two cross sections.
-A residual plot will be produced beneath the surface plots allowing for an assessment of the similarity of the markings.
-It is intended that users will perform this procedure at different x values in order to see whether the markings indicate a match at different points along the bullet.
+The application includes a button which allows the automatic creation of a residual plot from the loess fits of two bullets.
+An x coordinate value can be selected as a cross cut, and cross sections of the two bullets at the x value are taken.
+The previously described loess regression procedure is applied to the two cross sections.
+A residual plot is produced beneath the surface plots allowing for an assessment of the similarity of the markings.
+Users need to perform this procedure at different x values in order to see whether the markings indicate a match at different points along the bullet.
 
@@
 Cross cuts of bullet 1 are shown on the left (cross cuts with values below 100) and cross cuts of bullets 2 are shown on the right (cross cuts with values above 100).
-A lot of the striation marks align, allowing for an easy visual assessment of the shared similarities between the two bullets.
+Many of the striation marks align, allowing for an easy visual assessment of the shared similarities between the two bullets.
 
@@
 \includegraphics[width=\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}Different view of the manually adjusted side-by-side comparison of land five of bullet 1 and land 1 of bullet 2.}
+\caption{\label{fig:manualmatch-rgl}Different view of the manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
 \end{figure}
@@
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab("Residuals of Loess fit")
+  ylab("Residuals of loess fit")
 @
@@
 \begin{subfigure}[t]{\linewidth}
-    \caption{Adjusted (Bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
+    \caption{Adjusted (bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
 <<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
@@
   scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position="none") + ylab("Residuals of Loess fit")
+  theme(legend.position="none") + ylab("Residuals of loess fit")
 @
@@
 \end{subfigure}
-\caption{\label{fig:cross100}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of Bullet 1-5 to the right shows the similarity of the striation marks.}
+\caption{\label{fig:cross100}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
 \end{figure}
@@
 \]
-with suitable defined limits of the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).
+with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).
 
@@
 These are compared to a crosscut at 100$\mu m$.
-Initially this comparison constitutes an almost perfect match (consolidated by a CMS of fourteen) between the two crosscuts.
+Initially this comparison constitutes an almost perfect match between the two crosscuts.
 However, the match quickly deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.
@@
 As mentioned earlier, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a Loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
+In particular, for detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
 We can investigate the performance of peak/valley detection across different smoothing factors.
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the lowest crosscut level in this region, because typically, individual characteristics are best express at the lower end of the bullet. 
+We first identify a stable region for each bullet and pick the lowest crosscut level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet. 
 %\hh{(assumption here: lower crosscut levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
@@
 
-In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
+Determining a CMS threshold that needs to be exceeded for us to declare a match is beyond the scope of this work, even though it is critically important in practice.
+We provide some ideas in the next section.
+%In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
 
@@
 \begin{subfigure}[t]{.49\textwidth}
-\caption{Maximal number of CMS between unknown Bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
@@
 There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
-On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}.
+On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}).
 Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}.
@@
 
-Another benefit is that we can not only apply a single decision tree, but several hundred of them to combine in a random forest~\cite{breiman:2001, randomForest}.
+Another benefit is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\cite{breiman:2001, randomForest}.
 In a random forest multiple trees are fitted.
@@
 This is a weighted average between zero false positives and an error rate of false negatives of 0.26.
-Note that this error rate is based on land-to-land comparisons and will be much lower for bullet-to-bullet matches.
+Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
 In the case of the Hamby study, the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched (see Figure~\ref{fig:treeresult}).
@@
 
-This research was partially funded through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
+This research was partially funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
 

### 2015-12-16 imaging-paper.Rnw
update to elsevier format

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Large Colored Title Article
-% LaTeX Template
-% Version 1.1 (25/11/12)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Frits Wenneker (http://www.howtotex.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[DIV=calc, paper=letter, fontsize=10pt, twocolumn]{scrartcl}	 % A4 paper and 11pt font size
+\documentclass[review]{elsarticle}
+
+\usepackage{lineno,hyperref}
+\modulolinenumbers[5]
+
+\journal{Journal of Forensic Science International}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+%\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+%\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \usepackage{float}
-%\usepackage{afterpage}
+\usepackage{subcaption}
 
@@
 
-% Headers - all currently empty
-\lhead{}
-\chead{}
-\rhead{}
-
-% Footers
-\lfoot{}
-\cfoot{}
-\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % "Page 1 of 2"
-
-\renewcommand{\headrulewidth}{0.0pt} % No header rule
-\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule
-
 \usepackage{color}
@@
 \newcommand{\eh}[1]{{\color{blue} #1}}
-
-\usepackage{lettrine} % Package to accentuate the first letter of the text
-\newcommand{\initial}[1]{ % Defines the command and style for the first letter
-\lettrine[lines=3,lhang=0.3,nindent=0em]{
-\color{DarkGoldenrod}
-{\textsf{#1}}}{}}
-
-\usepackage{subcaption}
 
@@
 %----------------------------------------------------------------------------------------
-%	TITLE SECTION
-%----------------------------------------------------------------------------------------
-
-\usepackage{titling} % Allows custom title configuration
-
-\newcommand{\HorRule}{\color{DarkGoldenrod} \rule{\linewidth}{1pt}} % Defines the gold horizontal rule around the title
-
-\pretitle{\vspace{-30pt} \begin{flushleft} \HorRule \fontsize{50}{50} \usefont{OT1}{phv}{b}{n} \color{DarkRed} \selectfont} % Horizontal rule before the title
-
-\title{Matching Bullets} % Your article title
-
-\posttitle{\par\end{flushleft}\vskip 0.5em} % Whitespace under the title
-
-\preauthor{\begin{flushleft}\large \lineskip 0.5em \usefont{OT1}{phv}{b}{sl} \color{DarkRed}} % Author font configuration
-
-\author{Eric Hare, Heike Hofmann} % Your name
-
-\postauthor{\footnotesize \usefont{OT1}{phv}{m}{sl} \color{Black} % Configuration for the institution name
-
-Department of Statistics and Statistical Laboratory,
-Iowa State University% Your institution
-
-\par\end{flushleft}\HorRule} % Horizontal rule after the title
-
-\date{} % Add a date here if you would like one to appear underneath the title block
-
-%----------------------------------------------------------------------------------------
 
@@
 
-\maketitle % Print the title
-
-\thispagestyle{fancy} % Enabling the custom headers/footers for the first page
-
-%----------------------------------------------------------------------------------------
-%	ABSTRACT
-%----------------------------------------------------------------------------------------
-
-% The first character should be within \initial{}
-\initial{F}\textbf{orensic examiners try to determine whether two bullets were fired from the same barrel by analyzing them under a microscope in controlled conditions. Can we make the task simpler?}
-
-%----------------------------------------------------------------------------------------
-%	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\paragraph{Keywords:} Data visualization, Statistical graphics, Statistical computing.
+\begin{frontmatter}
+
+\title{Matching Bullets}
+\tnotetext[mytitlenote]{Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}
+
+%% Group authors per affiliation:
+\author{Eric Hare, Heike Hofmann, Alicia Carriquiry}
+\address{Department of Statistics and Statistical Laboratory, Iowa State University}
+\fntext[myfootnote]{Since 1880.}
+
+%% or include affiliations in footnotes:
+\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
+\ead[url]{www.elsevier.com}
+
+\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Eric Hare}
+\ead{erichare@iastate.edu}
+
+\address[mymainaddress]{Department of Statistics and Statistical Laboratory, Iowa State University}
+\address[mysecondaryaddress]{secondary address -- CS?}
+
+\begin{abstract}
+In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
+Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
+During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet.
+Identifying these striation markings in an attempt to match two bullets is the primary goal.
+We propose a framework for the analysis of the 3D surface measurements which transcribes the markings into a 2D plotting framework.
+This makes identification of matches easier and more rooted in science for both the expert and the lay person.
+The automatic matching routine we propose manages to correctly predict 132 out of 180 known bullet land to bullet land matches.
+\end{abstract}
+
+\begin{keyword}
+\hh{XXX not the right keywords yet} Data visualization, Statistical graphics, Statistical computing.
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
 
@@
 We propose a framework which allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
+
+The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
+We then describe a web application for visualizing and working with bullet surfaces.
+We proceed to describing an automatic matching routine which we evaluate on a small database of reference bullets.
 
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:residuala}Residual structure of a surface crosscut at $x = 1.5625$ (bottom of the bullet).}
+\caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
 <<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:residualb} Residual structure of a surface crosscut at $x = 100.00$}
+\caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
 <<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
@@
 \caption{\label{fig:loess_step1}Step 1 the loess fit algorithm.
-For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
+For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
+\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
 <<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
@@
 \begin{subfigure}[t]{\linewidth}
-    \caption{Adjusted (bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
+    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
 <<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
@@
 \end{subfigure}
-\caption{\label{fig:cross100}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
+\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
 \end{figure}
@@
 \subsection{Horizontal Alignment}
-Using (for now) a single crosscut across the surface, we extract the values and remove the class characteristics using a loess fit as before.
+Using (for now) a single cross section across the surface, we extract the values and remove the class characteristics using a loess fit as before.
 This enables us to focus on the individual characterists.
 Striation marks show up in this representations as peaks and valleys.
-Figure~\ref{fig:cross100} shows lines of the surface crosscut at $x = 100$.
+Figure~\ref{fig:cross100} shows lines of the surface cross section at $x = 100$.
 A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.
-This horizontal shift is based on the cross-correlation between the two surface crosscuts.
-Let $f(t)$ and $g(t)$ define the values of the crosscut at $t$, then the cross-correlation between $f$ and $g$ at lag $k$ is defined as
+This horizontal shift is based on the cross-correlation between the two surface cross sections.
+Let $f(t)$ and $g(t)$ define the values of the cross section at $t$, then the cross-correlation between $f$ and $g$ at lag $k$ is defined as
 \[
@@
 @
-\caption{\label{fig:ccf}Cross-correlation between the two crosscuts shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
-At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the crosscuts.
-Figure~\ref{fig:crosscutX} shows the lag-shifted crosscuts.}
-\end{figure}
-
-\subsection{Varying crosscuts}
-The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up.
-This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which crosscut level was used in a matching, but on the other hand we can use multiple crosscuts of the same bullet for an initial assessment of its quality.
+\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
+At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections.
+Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
+\end{figure}
+
+\subsection{Varying cross sections}
+The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up.
+This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality.
 Both of these situations are assessed and discussed in more detail below:
-Figure~\ref{fig:crosscuts} shows a sequence of crosscuts for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$.
-These are compared to a crosscut at 100$\mu m$.
-Initially this comparison constitutes an almost perfect match between the two crosscuts.
-However, the match quickly deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.
-Only if the crosscut is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
-Given that we have to expect some variation in nominally the same crosscut levels due to manual alignments in microscopes, we should take crosscut levels into account in the automatic matching routine by evaluating matches at several cross sections.
+Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$.
+These are compared to a cross section at 100$\mu m$.
+Initially this comparison constitutes an almost perfect match between the two cross sections.
+However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.
+Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
+Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections.
 \begin{figure}[hbtp]
@@
 @
-\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts.
-Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$.
-With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
+\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections.
+Cross section at $x = 100$ is compared to cross sections every 50$\mu m$.
+With every step away from the original cross section site, the number of differences between the cross sections increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
 }
@@
 @
-\caption{\label{fig:crosscuts2}Varying crosscuts for barrel 3, bullet 1.
-Initially, the match between crosscuts 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet.
-At a crosscut level of $175\mu m$ the bullet pattern stabilizes.
-For this land, matches should not be attempted at lower crosscuts.}
-\end{figure}
-An opportunity that arises from comparing multiple crosscuts of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
-In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern.
+\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet 1.
+Initially, the match between cross sections 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet.
+At a cross section level of $175\mu m$ the bullet pattern stabilizes.
+For this land, matches should not be attempted at lower cross sections.}
+\end{figure}
+An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the cross section is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
+In the approach here, we keep increasing the cross section level until we find a section with a stable pattern.
 This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3.
-`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9.
+`Stability' is here defined as two aligned cross sections chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9.
 
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{Using a rolling median peaks and grooves are identified for each crosscut. \label{fig:smoothcutb}}{%
+\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
 <<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
@@
     geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
-  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
+  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
 @
@@
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
+  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) +
   xlab("y")
@@
 \end{subfigure}
-\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between crosscuts (c).}
+\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between cross sections (c).}
 \end{figure}
@@
     geom_line(aes(x=y, y=smoothed), data=smooths) +
-  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
+  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
 @
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the lowest crosscut level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet. 
-%\hh{(assumption here: lower crosscut levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
+We first identify a stable region for each bullet and pick the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet.
+%\hh{(assumption here: lower cross section levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
 
@@
 \begin{enumerate}
-\item Smooth the two crosscuts by a loess with a very small span (see Figure~\ref{fig:smooth}). 
-\item Use cross-correlation to find the best alignment along the crosscuts: Shift one of the crosscuts by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
-\item Using a rolling average, identify peaks and grooves for each of the crosscuts.
+\item Smooth the two cross sections by a loess with a very small span (see Figure~\ref{fig:smooth}). 
+\item Use cross-correlation to find the best alignment along the cross sections: Shift one of the cross sections by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
+\item Using a rolling average, identify peaks and grooves for each of the cross sections.
 We define a region around the extrema on each side as one third of the distance to the next extrema (see Figure~\ref{fig:smoothcutb}).
-\item Match grooves and peaks between crosscuts: based on the regions around the extrema as defined above, we identify joint regions between two crosscuts as those areas, in which at least two of the individual regions touch.
+\item Match grooves and peaks between cross sections: based on the regions around the extrema as defined above, we identify joint regions between two cross sections as those areas, in which at least two of the individual regions touch.
 A joint region is defined as the smallest interval that covers all of the touching regions.
-A joint region is called a match between the crosscuts, if all of the crosscut-specific regions are of the same type of extrema, i.e.\ either all peaks or all valleys.
+A joint region is called a match between the cross sections, if all of the cross section-specific regions are of the same type of extrema, i.e.\ either all peaks or all valleys.
 In Figure~\ref{fig:match} all matches are colored corresponding to their type of extrema.
 Non-matching regions are left grey.
-\item Identify the maximal number of CMS (consecutive matching striae) between the two crosscuts by counting consecutive matching lines. 
+\item Identify the maximal number of CMS (consecutive matching striae) between the two cross sections by counting consecutive matching lines. 
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
@@
 \caption{\label{fig:mismatch}Known mismatch with a relatively large number of maximal consecutive matching striae (twelve) in the middle.
-The pattern in the middle does look surprisingly similar, however the outer ends of the crosscuts easily reveals this as mismatch.
+The pattern in the middle does look surprisingly similar, however the outer ends of the cross sections easily reveals this as mismatch.
 }
@@
 For smaller numbers of CMS, the percentage of false positives quickly increases.
-However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two crosscuts (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
+However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two cross sections (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
 There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:distr.dist}Relationship between maximal CMS, distances between crosscuts and matches.}
+\caption{\label{fig:distr.dist}Relationship between maximal CMS, distances between cross sections and matches.}
 <<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
@@
    facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
-  ylab("Distance between crosscuts") +
+  ylab("Distance between cross sections") +
   scale_colour_brewer(palette="Paired") +
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:ccfs-match}Relationship between maximal CMS, cross-correlation between crosscuts and matches.}
+\caption{\label{fig:ccfs-match}Relationship between maximal CMS, cross-correlation between cross sections and matches.}
 <<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
@@
 \end{subfigure}
-\caption{A small distance between crosscuts or a large cross-correlation between crosscuts are indicative of matches between lands.}
-\end{figure}
-
-Similarly, the Euclidean distance between two crosscuts across the range of loess residuals is strongly correlated to known matches.
-Out of the 48 pairs of land-to-land crosscuts with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of crosscuts that both are very flat with peaks and valleys of less than 1.5$\mu m$).
+\caption{A small distance between cross sections or a large cross-correlation between cross sections are indicative of matches between lands.}
+\end{figure}
+
+Similarly, the Euclidean distance between two cross sections across the range of loess residuals is strongly correlated to known matches.
+Out of the 48 pairs of land-to-land cross sections with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of cross sections that both are very flat with peaks and valleys of less than 1.5$\mu m$).
 The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$distr.dist[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$distr.dist[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$distr.dist[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$distr.dist[!bstats$match])}\mu m$.
@@
 Interestingly, the number of consecutive matching striae does not feature in this evaluation.
-Instead of CMS, cross-correlation (ccf) between the crosscuts is very important in the matching process  by the decision tree.
+Instead of CMS, cross-correlation (ccf) between the cross sections is very important in the matching process  by the decision tree.
 Aside from  cross correlation, the number of matches (num.matches) is involved in the decision.
@@
 The matching algorithm is sensitive to the parameter choices made.
-The distance between crosscuts (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
+The distance between cross sections (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
 Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
@@
 The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found.
-The remaining problem is simply that statistically the quality of the matching algorithm could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
+The remaining problem is simply that statistically the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
 The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks.
@@
 For a fair assessment of the adequacy of an algorithm we need transparency.
-Our matching  algorithm is open: the code is readily available in form of the R package x3rplus~\cite{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
+Our matching  algorithm is open: the code is readily available in form of the R package x3prplus~\cite{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
 Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home:\url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
@@
 
-\bibliographystyle{abbrv}
-%%use following if all content of bibtex file should be shown
-%\nocite{*}
+\section*{References}
+
 \bibliography{references}

### 2015-12-17 imaging-paper.Rnw
updates to references because of url treatment

@@
 \address{Department of Statistics and Statistical Laboratory, Iowa State University}
-\fntext[myfootnote]{Since 1880.}
+%\fntext[myfootnote]{Since 1880.}
 
 %% or include affiliations in footnotes:
-\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
-\ead[url]{www.elsevier.com}
-
-\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Eric Hare}
+%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
+%\ead[url]{www.elsevier.com}
+
+%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
 \ead{erichare@iastate.edu}
 
-\address[mymainaddress]{Department of Statistics and Statistical Laboratory, Iowa State University}
-\address[mysecondaryaddress]{secondary address -- CS?}
+% \address[mymainaddress]{Department of Statistics and Statistical Laboratory, Iowa State University}
+% \address[mysecondaryaddress]{secondary address -- CS?}
 
@@
 % \item DONE Walkthrough of the process
-% \item \hh{XXX Alternative approaches: IBIS database and the NRC report\cite{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
+% \item \hh{XXX Alternative approaches: IBIS database and the NRC report\citep{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
 % NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
@@
 Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar.
-Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\cite{giannelli:2011}.
+Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\citep{giannelli:2011}.
 
 In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
-~\cite{NAS:2009}.
+~\citep{NAS:2009}.
 The report concluded that minimal research had been done to explore the reliability of these ballistic identification methods.
@@
 Rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet during the firing process.
-These marks are assumed to be unique to the barrel, as described in a 1992 AFTE article~\cite{afte:1992}.
+These marks are assumed to be unique to the barrel, as described in a 1992 AFTE article~\citep{afte:1992}.
 ``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."
@@
 \section{Methodology}
-Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\cite{hamby:2009}.
+Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
 Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
@@
 \caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
-<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
 qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
@@
 \caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
-<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
 #qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
@@
 For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
-<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
-br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
-br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
-br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
-br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
-br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
-br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))
+<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
+br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
+br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
+br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
+br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
+br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
+br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
+br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))
 
@@
 \end{subfigure}
-%\begin{subfigure}[t]{\linewidth}
-% \caption{\label{fig:loess_step2}Step 2 of the loess fit algorithm. The height values are smoothed with a particular smoothing factor to remove noise in the deviations.}
-<<loess_step2, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
-subdata <- br111_bullet
-smoothfactor <- 35
-
-value_filled <- na.fill(subdata$value, "extend")
-smoothed <- rollapply(value_filled, smoothfactor, function(x) mean(x))
-smoothed_truefalse <- rollapply(smoothed, smoothfactor, function(x) mean(x))
-
-p2 <- qplot(1.5625*(1:length(smoothed)), smoothed) + theme_bw() + xlab("y") +
-  ylab("Surface measurements")
-#p2
-@
-%\end{subfigure}
 \begin{subfigure}[t]{\linewidth}
@@
 Grooves and peaks are detected, if they are not within the same window.}
-<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
+subdata <- br111_bullet
+smoothfactor <- 35
+
+value_filled <- na.fill(subdata$value, "extend")
+smoothed <- rollapply(value_filled, smoothfactor, function(x) mean(x))
+smoothed_truefalse <- rollapply(smoothed, smoothfactor, function(x) mean(x))
+
 xloc=250
@@
 Values outside these boundaries are removed (shown in grey)}
-<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
     lengthdiff <- length(subdata$value) - length(smoothed_truefalse)
@@
 That is, the smoothed data is once again smoothed by computing a new rolling average with the same smoothing factor.
-This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass and then added back to the result~\cite{tukey:1977}.
+This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass and then added back to the result~\citep{tukey:1977}.
 This has the effect of introducing a little bit more variance into the smoothed data.
@@
 \caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
-<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
 
@@
 \caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
-<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE>>=
 my.loess$resid + ylab("Residuals of loess Fit")
@@
 We have implemented a web interface to automate many of the steps outlined in the previous section.
-The web interface, developed using Shiny~\cite{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
-A 3D viewer, using Plotly~\cite{plotly}, allows the user to rotate and zoom to examine surface features of the images.
+The web interface, developed using Shiny~\citep{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
+A 3D viewer, using Plotly~\citep{plotly}, allows the user to rotate and zoom to examine surface features of the images.
 A screenshot of the application is available in Figure \ref{fig:app}.
@@
 \begin{figure}[hbtp]
+\centering
 <<side-by-side, echo=FALSE, eval=FALSE>>=
@@
 list_of_fits <- lapply(crosscuts, function(x) {
-  br111 <- get_bullet(paths[5], x = x)
+  br111 <- get_crosscut(paths[5], x = x)
   br111.groove <- get_grooves(br111)
@@
 list_of_fits2 <- lapply(crosscuts, function(x) {
-  br111 <- get_bullet(paths[7], x)
+  br111 <- get_crosscut(paths[7], x)
   br111.groove <- get_grooves(br111)
@@
 
-\includegraphics[width=\columnwidth]{images/side-by-side-manual-match.png}
+\includegraphics[width=.7\columnwidth]{images/side-by-side-manual-match.png}
 \caption{\label{fig:manualmatch}Manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
@@
 \begin{figure}[hbtp]
-\includegraphics[width=\columnwidth]{images/matchup-rgl.png}
+\centering
+\includegraphics[width=.7\columnwidth]{images/matchup-rgl.png}
 \caption{\label{fig:manualmatch-rgl}Different view of the manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
@@
 \caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
-<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
 qplot(y, resid, group=bullet, colour=bullet, data=lof, 
@@
     \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
-<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
 # qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
@@
 \begin{figure}[hbtp]
-<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '\\linewidth'>>=
+<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.7\\linewidth'>>=
 lag <- ccf$lag[which.max(ccf$acf)]
@@
 \begin{figure}[hbtp]
-<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
 \begin{figure}[hbtp]
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
 \caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
-<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
+<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
 qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
@@
 \caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
-<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
+<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
 #images <- file.path(datadir, dir(datadir))
@@
 \label{fig:smoothcutd}}{%
-<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth', warning=FALSE>>=
+<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.7\\textwidth', warning=FALSE>>=
 peaks1$lines$bullet <- b12[1]
@@
 \begin{figure}[hbtp]
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='\\linewidth', warning=FALSE>>=
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.7\\linewidth', warning=FALSE>>=
 br111.groove <- get_grooves(br111_bullet)
@@
 \begin{enumerate}
-\item Smooth the two cross sections by a loess with a very small span (see Figure~\ref{fig:smooth}). 
-\item Use cross-correlation to find the best alignment along the cross sections: Shift one of the cross sections by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
-\item Using a rolling average, identify peaks and grooves for each of the cross sections.
+\item {\bf Smooth the two cross sections} by a loess with a very small span (see Figure~\ref{fig:smooth}). 
+\item Use cross-correlation to {\bf find the best alignment} along the cross sections: Shift one of the cross sections by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
+\item Using a rolling average, {\bf identify peaks and grooves} for each of the cross sections.
 We define a region around the extrema on each side as one third of the distance to the next extrema (see Figure~\ref{fig:smoothcutb}).
-\item Match grooves and peaks between cross sections: based on the regions around the extrema as defined above, we identify joint regions between two cross sections as those areas, in which at least two of the individual regions touch.
+\item {\bf Match grooves and peaks between cross sections:} based on the regions around the extrema as defined above, we identify joint regions between two cross sections as those areas, in which at least two of the individual regions touch.
 A joint region is defined as the smallest interval that covers all of the touching regions.
@@
 Non-matching regions are left grey.
-\item Identify the maximal number of CMS (consecutive matching striae) between the two cross sections by counting consecutive matching lines. 
+\item {\bf Identify the maximal number of CMS} (consecutive matching striae) between the two cross sections by counting consecutive matching lines. 
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti~\cite{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches~\cite{nichols:1997, nichols:2003, nichols:2003b}.
+For lead bullets, such as this, Biasotti~\citep{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}.
 \end{enumerate}
@@
 \begin{figure}[hbtp]
-<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='\\linewidth', warning=FALSE>>=
+<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.7\\linewidth', warning=FALSE>>=
 
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{.49\textwidth}
+\begin{subfigure}[t]{\textwidth}
 \caption{Maximal number of CMS between unknown bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
-<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
+<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\textwidth'>>=
 load("data-25-25/unkn9.RData")
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}
+\begin{subfigure}[t]{\textwidth}
 \caption{Overlaid lines of C-3 and the land with the top matching CMS.}
-<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
+<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.7\\textwidth', warning = FALSE>>=
 res <- reslist[[which.max(cmsdist)]]  
@@
 @
-
+\end{subfigure}
+\begin{subfigure}[t]{\textwidth}
 \caption{Overlaid lines of C-3 and the land with the second top matching CMS.}
-<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
+<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.7\\textwidth', warning = FALSE>>=
 res <- reslist[[which(cmsdist==14)]]  # number 2
@@
 \end{subfigure}
-\caption{\label{fig:hamby-perfect}Showcase scenario  when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
+\caption{\label{fig:hamby-perfect}Showcase scenario  when matching with CMS works very well. Unfortunately the matches are not always that convincing.}
 \end{figure}
@@
 Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}.
-Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by Miller~\cite{miller:1998}.
+Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}.
 Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae.
@@
 \caption{Distribution of maximal CMS.}
-<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
+<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
 ggplot(data=bstats) + 
@@
 \caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
-<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
+<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
 ggplot(data=bstats) + 
@@
 \begin{figure}[hbtp]
-<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\linewidth', warning=FALSE>>=
+<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.7\\linewidth', warning=FALSE>>=
 load("data-25-25/unkn47.RData")
@@
 \caption{\label{fig:distr.dist}Relationship between maximal CMS, distances between cross sections and matches.}
-<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
+<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
 ggplot(data=bstats) + 
@@
 \caption{\label{fig:ccfs-match}Relationship between maximal CMS, cross-correlation between cross sections and matches.}
-<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='\\linewidth'>>=
+<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
 ggplot(data=bstats) + theme_bw() + 
@@
 
-Using recursive partitioning, we fit a decision tree~\cite{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
+Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
 The resulting tree is shown in Figure~\ref{fig:tree}.
@@
 \begin{figure}[hbtp]
-<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\linewidth'>>=
 library(rpart)
@@
 
-Another benefit is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\cite{breiman:2001, randomForest}.
+Another benefit is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\citep{breiman:2001, randomForest}.
 In a random forest multiple trees are fitted.
@@
 \begin{figure}[hbtp]
-<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='\\linewidth'>>=
+<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='.7\\linewidth'>>=
 qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
@@
 \begin{figure}[hbtp]
-<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
+<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\linewidth'>>=
 bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
@@
 \begin{figure}[hbtp]
-<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='\\linewidth'>>=
+<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.7\\linewidth'>>=
 qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
@@
 
-The feasibility of a database of ballistic images was evaluated in a 2008 report~\cite{nap:2008} by the National Research Council.
+The feasibility of a database of ballistic images was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
 The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
@@
 The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks.
-This finding coincides with the assessment by de Kinder et al \cite{dekinder:2004} based on the IBIS Heritage system.
-A further re-assessment in 2015 by de Ceuster et al \cite{deceuster:2015} came to the same conclusions based on the EvoFinder system.
+This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system.
+A further re-assessment by \citet{deceuster:2015} came to the same conclusions based on the EvoFinder system.
 The NRC report also found, that results from 2d images are improved upon by matches based on 3d images.
@@
 By suggesting an automated algorithm that first removes class characteristics, such as the grooves and the curvature of the bullet to reveal the region of the  land, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
-Counting CMS in particular, has been found to be subjective to the examiner: `the concept of counting striations is subjective and based on experience'~\cite{miller:1998}.
+Counting CMS in particular, has been found to be subjective to the examiner: `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.
 
 For a fair assessment of the adequacy of an algorithm we need transparency.
-Our matching  algorithm is open: the code is readily available in form of the R package x3prplus~\cite{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
-Obviously, this is only a first step, but in times of the Netflix prize~\cite{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home:\url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
+Our matching  algorithm is open: the code is readily available in form of the R package x3prplus~\citep{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
+Obviously, this is only a first step, but in times of the Netflix prize~\citep{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home:\url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
 
@@
 
-An approach that might help in determining this difference comes from Biometrics, is suggested by Jain et al~\cite{jain:2015:tphil, jain:2015} and has been successfully applied in the Aadhar project (the effort by the Unique Identification Authority of India (UIDAI) to provide a unique 12-digit identification number to approximately 1.2 billion residents of India based on ten fingerprints and two irises).
+An approach that might help in determining this difference comes from Biometrics, is suggested by \citet{jain:2015:tphil, jain:2015} and has been successfully applied in the Aadhar project (the effort by the Unique Identification Authority of India (UIDAI) to provide a unique 12-digit identification number to approximately 1.2 billion residents of India based on ten fingerprints and two irises).
 This approach uses dictionary learning based on small square patches  of 64x64 or 32x32 pixels.
@@
 
-\paragraph{Acknowledgment} Thanks to the men and women behind the software R~\cite{R}, and the authors of the R packages  knitr~\cite{knitr} and ggplot2~\cite{ggplot2}.
+\paragraph{Acknowledgment} Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.
 

### 2015-12-17 imaging-paper.Rnw
probability ratios

@@
   xlab("maximum CMS")
+
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
+<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+ggplot(data=bstats) + 
+    geom_density(aes(x=cms, fill=match), alpha=0.5) +
+  xlab("maximum CMS") +
+  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  theme_bw() + 
+  theme(legend.position=c(1,1), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) 
 @
@@
 \begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:distr.dist}Relationship between maximal CMS, distances between cross sections and matches.}
+\caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
 <<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
@@
     xlab("maximum CMS")
-@
-\end{subfigure}
-\begin{subfigure}[t]{\linewidth}
-\caption{\label{fig:ccfs-match}Relationship between maximal CMS, cross-correlation between cross sections and matches.}
+
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
+<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+ggplot(data=bstats) + 
+    geom_density(aes(x=distr.dist, fill=match), alpha=0.5) +
+  xlab("Distances between cross sections") +
+  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  theme_bw() + 
+  theme(legend.position=c(1,1), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) 
+@
+\end{subfigure}
+\caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
+\end{figure}
+
+\begin{figure}[hbtp]
+\begin{subfigure}[t]{\linewidth}
+\caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
 <<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
@@
 \end{subfigure}
-\caption{A small distance between cross sections or a large cross-correlation between cross sections are indicative of matches between lands.}
+\begin{subfigure}[t]{\linewidth}
+\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
+<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+ggplot(data=bstats) + 
+    geom_density(aes(x=ccf, fill=match), alpha=0.5) +
+  xlab("Cross-correlation between cross sections") +
+  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  theme_bw() + 
+  theme(legend.position=c(1,1), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) 
+@
+\end{subfigure}
+\begin{subfigure}[t]{\linewidth}
+\caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
+<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+x <- seq(0,1, by=0.001)
+library(sm)
+km <- sm.density(bstats$ccf[bstats$match], display="none", eval.points=x)
+knm <- sm.density(bstats$ccf[!bstats$match], display="none", eval.points=x)
+dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
+dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
+dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
+
+densities <- data.frame(
+  x = rep(c(0, x, 1), times = 2),
+  estimate = c(0, km$estimate, 0, 0, knm$estimate,0),
+  match = rep(c("KM", "KNM"), each=1003))
+
+ggplot() +
+  geom_polygon(aes(x=x, y=estimate, fill=match, colour=match), 
+               data=densities, alpha=0.5) +
+  xlab("Cross-correlation between cross sections") +
+  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  theme_bw() + 
+  ylab("Density estimate") +
+  theme(legend.position=c(.975,1), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) +
+  geom_line(aes(x, pmin(5, kmtail/knmtail)), data=dframe) 
+@
+\end{subfigure}
+\caption{A large cross-correlation between cross sections is indicative of matches between lands.}
 \end{figure}

### 2015-12-17 imaging-paper.Rnw
ratio for cms

@@
 <<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+x <- seq(0,29, by=0.03)
+library(sm)
+km <- sm.density(bstats$cms[bstats$match], display="none", eval.points=x)
+knm <- sm.density(bstats$cms[!bstats$match], display="none", eval.points=x)
+dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
+dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
+dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
+
+densities <- data.frame(
+  match = rep(c("KM", "KNM"), each=(length(x)+2)),
+  x = rep(c(0, x, 29), times = 2),
+  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
+
+ggplot() +
+  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
+               data=densities, alpha=0.5) +
+xlab("maximum CMS") +
+  scale_fill_brewer("", palette="Paired") +
+  ylab("Density estimates") +
+  theme_bw() + 
+  theme(legend.position=c(1,1), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4)))
+
 ggplot(data=bstats) + 
@@
 ggplot() +
-  geom_polygon(aes(x=x, y=estimate, fill=match, colour=match), 
+  geom_polygon(aes(x=x, y=estimate, fill=match), colour="black",
                data=densities, alpha=0.5) +
@@
   scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
-  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+#  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
   theme_bw() + 

### 2015-12-18 imaging-paper.Rnw
ratios included on densities

--- 2015-12-17 imaging-paper.Rnw
@@
 library(xtable)
+library(sm)
 
@@
 <<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
-ggplot(data=bstats) + 
-    geom_density(aes(x=cms, fill=match), alpha=0.5) +
+x <- seq(0,29, by=0.03)
+km <- sm.density(bstats$cms[bstats$match], display="none", eval.points=x)
+knm <- sm.density(bstats$cms[!bstats$match], display="none", eval.points=x)
+dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
+dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
+dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
+
+densities <- data.frame(
+  match = rep(c("KM", "KNM"), each=(length(x)+2)),
+  x = rep(c(0, x, 29), times = 2),
+  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
+
+ggplot() +
+  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
+               data=densities, alpha=0.5) +
 xlab("maximum CMS") +
-  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  scale_fill_brewer("", palette="Paired") +
+  ylab("Density estimates") +
   theme_bw() + 
   theme(legend.position=c(1,1), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) 
+        legend.background = element_rect(fill=alpha('white', 0.4))) +
+  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)
+
+# ggplot(data=bstats) + 
+#     geom_density(aes(x=cms, fill=match), alpha=0.5) +
+#   xlab("maximum CMS") +
+#   scale_fill_brewer("", palette="Paired") +
+#   theme_bw() + 
+#   theme(legend.position=c(1,1), legend.justification=c(1,1),
+#         legend.background = element_rect(fill=alpha('white', 0.4)))
+
 @
@@
     xlab("maximum CMS")
-
 @
@@
   xlab("Distances between cross sections") +
-  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  scale_fill_brewer("", palette="Paired") +
   theme_bw() + 
@@
         legend.background = element_rect(fill=alpha('white', 0.4)))
+
+x <- seq(0,20, by=0.02)
+km <- sm.density(bstats$distr.dist[bstats$match], display="none", eval.points=x)
+knm <- sm.density(bstats$distr.dist[!bstats$match], display="none", eval.points=x)
+dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
+dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
+dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
+
+densities <- data.frame(
+  match = rep(c("KM", "KNM"), each=(length(x)+2)),
+  x = rep(c(0, x, 29), times = 2),
+  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
+
+ggplot() +
+  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
+               data=densities, alpha=0.5) +
+  xlab("Distances between cross sections") +
+  scale_fill_brewer("", palette="Paired") +
+  ylab("Density estimates") +
+  theme_bw() + 
+  theme(legend.position=c(1,1), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) +
+  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)
+
 @
@@
   xlab("Cross-correlation between cross sections") +
-  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  scale_fill_brewer("", palette="Paired") +
   theme_bw() + 
@@
 ggplot() +
-  geom_polygon(aes(x=x, y=estimate, fill=match, colour=match), 
+  geom_polygon(aes(x=x, y=estimate, fill=match), colour="black",
                data=densities, alpha=0.5) +
   xlab("Cross-correlation between cross sections") +
-  scale_fill_brewer("", palette="Paired", labels=c("KNM", "KM")) +
-  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+  scale_fill_brewer("", palette="Paired") +
+#  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
   theme_bw() + 

### 2015-12-18 imaging-paper.Rnw
excluding flagged lands from automatic evaluation

@@
 \begin{keyword}
-\hh{XXX not the right keywords yet} Data visualization, Statistical graphics, Statistical computing.
+\hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Statistical graphics, Statistical computing.
 \end{keyword}
@@
 
-\section{Approach to automatic matching}
+\section{Automatic matching}
 Applying the loess fit to a range of different surface cross cuts (see Figures~\ref{fig:manualmatch} and~\ref{fig:manualmatch-rgl}, with cross cuts from 50 to 150) shows the striation marks across two bullets.
@@
 The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up.
-This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality.
+This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality.
 Both of these situations are assessed and discussed in more detail below:
@@
 \end{figure}
-An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the cross section is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
+An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the cross section is in a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
 In the approach here, we keep increasing the cross section level until we find a section with a stable pattern.
 This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3.
-`Stability' is here defined as two aligned cross sections chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9.
+`Stability' is here defined as two aligned cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
 
@@
   b12 <- unique(lof$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 30)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 30)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
   peaks1$lines$bullet <- b12[1]
@@
 
-\subsection{Automated matching algorithm}
+\subsection{Algorithm}
 
@@
 \section{Evaluation}
+\hh{In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
+}
+\hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
+For this, we ues cross sections 25$\mu m$ apart and require a minimum of 0.9 for the cross correlation.
+Four lands from different bullets are flagged as problematic in this respect.
+A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface.}
+%
+\begin{figure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{Br6 Bullet 2-1}
+%' <<echo=FALSE>>=
+%' b1 <- read.x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
+%' image(b1$surface.matrix)
+%' @
+\includegraphics[width=\linewidth]{images/br6-2-1.png}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{Br9 Bullet 2-4}
+\includegraphics[width=\linewidth]{images/br9-2-4.png}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{Unkn Bullet B-2}
+\includegraphics[width=\linewidth]{images/b-2.png}
+\end{subfigure}
+\begin{subfigure}[t]{.49\textwidth}
+\caption{Unkn Bullet Q-4}
+\includegraphics[width=\linewidth]{images/q-4.png}
+\end{subfigure}
+\caption{\label{fig:fourflags}}
+\end{figure}
+%
 Using the James Hamby study, we run all unknown bullets against all known bullets for matches. 

### 2015-12-18 code/checkBullets.R

@@ -13,7 +13,7 @@ unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 if (!file.exists("csvs/crosscuts.csv")) {
 crosscuts <- sapply(c(knowns, unknowns), function(x) {
   cat(x)
-  crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=25))
+  crosscut <- bulletCheckCrossCut(x, x = seq(100, 750, by=25))
   cat(crosscut)
   cat("\n")
   crosscut

### 2015-12-18 imaging-paper.Rnw
the new random forest is insanely good!

@@
 \centering
-<<side-by-side, echo=FALSE, eval=FALSE>>=
-paths <- file.path(datadir, dir(datadir))
-
-br111 <- read.x3p(paths[5])
-crosscuts <- unique(fortify_x3p(br111)$x)
-crosscuts <- crosscuts[crosscuts > 50]
-crosscuts <- crosscuts[crosscuts < 150]
-
-list_of_fits <- lapply(crosscuts, function(x) {
-  br111 <- get_crosscut(paths[5], x = x)
-  br111.groove <- get_grooves(br111)
-  br111.groove$plot
-  fit_loess(br111, br111.groove)
-})
-
-lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
-
-br111 <- read.x3p(paths[7])
-crosscuts <- unique(fortify_x3p(br111)$x)
-crosscuts <- crosscuts[crosscuts > 50]
-crosscuts <- crosscuts[crosscuts < 150]
-
-list_of_fits2 <- lapply(crosscuts, function(x) {
-  br111 <- get_crosscut(paths[7], x)
-  br111.groove <- get_grooves(br111)
-  br111.groove$plot
-  fit_loess(br111, br111.groove)
-})
-
-lof2 <- lapply(list_of_fits2, function(x) x$resid$data) %>% bind_rows
-lof$bullet <- 1
-lof2$bullet <- 2
-
-LOF <- rbind(lof, lof2)
-
-subLOF1 <- LOF %>% filter(bullet == 1, x <= 100)
-subLOF2 <- LOF %>% filter(bullet == 2, x > 100)
-
-#subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
-subLOF1$y <- subLOF1$y + 19*1.5625 # working now!!!
-subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
-
-qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
-  scale_fill_gradientn(limits=c(-5,5), 
-                       colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
-                       values = c(0,0.5, 0.75, 0.9, 1)) +
-  theme_bw()
-@
+%' <<side-by-side, echo=FALSE, eval=FALSE>>=
+%' paths <- file.path(datadir, dir(datadir))
+%' 
+%' br111 <- read.x3p(paths[5])
+%' crosscuts <- unique(fortify_x3p(br111)$x)
+%' crosscuts <- crosscuts[crosscuts > 50]
+%' crosscuts <- crosscuts[crosscuts < 150]
+%' 
+%' list_of_fits <- lapply(crosscuts, function(x) {
+%'   br111 <- get_crosscut(paths[5], x = x)
+%'   br111.groove <- get_grooves(br111)
+%'   br111.groove$plot
+%'   fit_loess(br111, br111.groove)
+%' })
+%' 
+%' lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
+%' 
+%' br111 <- read.x3p(paths[7])
+%' crosscuts <- unique(fortify_x3p(br111)$x)
+%' crosscuts <- crosscuts[crosscuts > 50]
+%' crosscuts <- crosscuts[crosscuts < 150]
+%' 
+%' list_of_fits2 <- lapply(crosscuts, function(x) {
+%'   br111 <- get_crosscut(paths[7], x)
+%'   br111.groove <- get_grooves(br111)
+%'   br111.groove$plot
+%'   fit_loess(br111, br111.groove)
+%' })
+%' 
+%' lof2 <- lapply(list_of_fits2, function(x) x$resid$data) %>% bind_rows
+%' lof$bullet <- 1
+%' lof2$bullet <- 2
+%' 
+%' LOF <- rbind(lof, lof2)
+%' 
+%' subLOF1 <- LOF %>% filter(bullet == 1, x <= 100)
+%' subLOF2 <- LOF %>% filter(bullet == 2, x > 100)
+%' 
+%' #subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
+%' subLOF1$y <- subLOF1$y + 19*1.5625 # working now!!!
+%' subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
+%' 
+%' qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
+%'   scale_fill_gradientn(limits=c(-5,5), 
+%'                        colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
+%'                        values = c(0,0.5, 0.75, 0.9, 1)) +
+%'   theme_bw()
+%' @
 
@@
 \end{subfigure}
-\caption{\label{fig:fourflags}}
+\caption{\label{fig:fourflags}\hh{Images of the four lands that got flagged during the quality assessment.
+All of them show scratch marks across the striation marks from the barrel.
+They will be excluded from the remainder of the analysis.}}
 \end{figure}
 %
-Using the James Hamby study, we run all unknown bullets against all known bullets for matches. 
+\hh{We exclude these four lands from the further matching considerations and 
+ run all unknown bullets against all known bullets for matches. 
 The algorithm is using land-to-land matches, i.e.
-we are running $15 \times 6 = 90$ lands from unkown bullets against $2 \times 10 \times 6 = 120$ lands from known bullets, yielding a total of $10,800$ comparisons with 180 known land-to-land matches.
+we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unkown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons.
+}
  When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land C-3 and all 120 known lands.
@@
 
-includes <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
-rp1 <- rpart(match~., bstats[,includes])  # doesn't include cms at all !!!!
+includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
+cc <- read.csv("csvs/crosscuts-25-new.csv")
+cc$bullet <- gsub(".*//","",cc$path)
+cc$bullet <- gsub(".x3p", "", cc$bullet)
+bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
+                  bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]
+
+excludesObs <- which(bstats$flagged)
+
+rp1 <- rpart(match~., bstats[-excludesObs,includesVar])  # doesn't include cms at all !!!!
 
@@
 
-bstats$pred <- predict(rp1)
+bstats$pred <- predict(rp1, newdata=bstats)
 @
@@
 set.seed(20151215)
-rtrees <- randomForest(factor(match)~., data=bstats[,includes], ntree=300)
+rtrees <- randomForest(factor(match)~., data=bstats[-excludesObs,includesVar], ntree=300)
 
@@
 errors$id <- 1:nrow(errors)
-bstats$forest <- predict(rtrees, type="prob")[,2]
+bstats$forest <- predict(rtrees, type="prob", newdata=bstats)[,2]
 @
@@
 @
-\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12).
+\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12, except for Q and B, where the maximum is 11).
 Both results for the tree and the random forest are shown.
@@
 
+\hh{XXX Needs still an answer: under what circumstances do we decide on match/no-match successfully?}
+
 \section{Conclusion}

### 2015-12-19 imaging-paper.Rnw
update on random forest

@@
 \hh{We exclude these four lands from the further matching considerations and 
- run all unknown bullets against all known bullets for matches. 
-The algorithm is using land-to-land matches, i.e.
+ run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e.
 we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unkown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons.
+Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
 }
@@
 bstats$pred <- predict(rp1, newdata=bstats)
+bstats$pred[bstats$flagged] <- NA
 @
@@
 bstats$forest <- predict(rtrees, type="prob", newdata=bstats)[,2]
+bstats$forest[bstats$flagged] <- NA
 @
@@
 %
-After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0044.
-This is a weighted average between zero false positives and an error rate of false negatives of 0.26.
+After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0040.
+This is a weighted average between a false positive error rate of 0.000098 and an error rate of false negatives of 0.22.
+\hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known matches and non-matches (see Figures~\ref{fig:treeresults} and ~\ref{fig:tree-forest}).}
 Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
@@
 bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
-bullets <- bstats %>% filter(match) %>% group_by(bullet) %>% summarize(
+bullets <- bstats %>% filter(match, !flagged) %>% group_by(bullet) %>% summarize(
   n = n(),
@@
 bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
-qplot(data=bm, value, reorder(bullet, value/n), size=I(2.5), 
+qplot(data=bm, 100*value/n, reorder(bullet, value/n), size=I(2.5), 
       shape=prediction, colour=prediction, alpha=I(0.6)) + 
@@
 #  scale_shape_manual(values=c(1,2)) +
-  scale_x_continuous("Number of correctly predicted land-to-land matches", 
-                     breaks = 3*0:4, limits=c(0,12),
-                     minor_breaks = 1:11) +
+  scale_x_continuous("Percent of correctly predicted land-to-land matches", 
+                     breaks = c(0,25,50,75,100), limits=c(0,100)) +
   theme(legend.position = "bottom")
@@
 Both results for the tree and the random forest are shown.
-For each bullet at least four lands are correctly matched.
-Unknown bullet B is the hardest for the algorithm to match.}
+In the tree, at least \hh{40\% of all possible} lands of a bullet are  matched correctly, while the random forest achieves perfect results.
+Unknown bullet B is the hardest for the tree to match.}
 \end{figure}
@@
   ylab("Predictions from random forest")   +
-  scale_colour_brewer(palette="Paired")
+  scale_colour_brewer(palette="Paired", labels=c("KNM", "KM")) +
+  scale_shape_discrete(labels=c("KNM", "KM"))
 @

### 2015-12-19 imaging-paper.Rnw
switch to grey scale
comment on bulletTrax 3d study

@@
 
-\title{Matching Bullets}
+\title{Random Forests for Automatic Matching Bullet-Lands}
 \tnotetext[mytitlenote]{Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}
@@
 This makes identification of matches easier and more rooted in science for both the expert and the lay person.
-The automatic matching routine we propose manages to correctly predict 132 out of 180 known bullet land to bullet land matches.
+The automatic matching routine we propose \hh{manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches.}
 \end{abstract}
@@
 \begin{keyword}
-\hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Statistical graphics, Statistical computing.
+\hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Machine Learning, Feature importance.
 \end{keyword}
@@
 % \end{enumerate}
+
+\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'.
+No algorithm implemented/available, no error rates of automated matching are given XXX}
 
@@
 %' @
-\includegraphics[width=\linewidth]{images/br6-2-1.png}
+\includegraphics[width=\linewidth]{images/br6-2-1-grey.png}
 \end{subfigure}
@@
 \caption{Br9 Bullet 2-4}
-\includegraphics[width=\linewidth]{images/br9-2-4.png}
+\includegraphics[width=\linewidth]{images/br9-2-4-grey.png}
 \end{subfigure}
@@
 \caption{Unkn Bullet B-2}
-\includegraphics[width=\linewidth]{images/b-2.png}
+\includegraphics[width=\linewidth]{images/b-2-grey.png}
 \end{subfigure}
@@
 \caption{Unkn Bullet Q-4}
-\includegraphics[width=\linewidth]{images/q-4.png}
+\includegraphics[width=\linewidth]{images/q-4-grey.png}
 \end{subfigure}
@@
 <<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
-ggplot(data=bstats) + 
+ggplot(data=subset(bstats, !flagged)) + 
   geom_bar(aes(x=factor(cms))) + theme_bw() +
@@
 <<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
-ggplot(data=bstats) + 
+ggplot(data=subset(bstats, !flagged)) + 
   geom_bar(aes(x=factor(cms), fill=match), position="fill") +
@@
 Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis).
-Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.
-
-\begin{figure}[hbtp]
-<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.7\\linewidth'>>=
+\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match.
+The random forest predicts all 172 known matches correctly as well as all 10,212 known non-matches.}
+%Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.
+
+\begin{figure}[hbtp]
+<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.7\\linewidth', warning = FALSE>>=
 qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
@@
 
-\hh{XXX Needs still an answer: under what circumstances do we decide on match/no-match successfully?}
-
 \section{Conclusion}

### 2015-12-19 imaging-paper.Rnw
Found and corrected typo
Fix figure reference

@@
 \hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
-For this, we ues cross sections 25$\mu m$ apart and require a minimum of 0.9 for the cross correlation.
+For this, we use cross sections 25$\mu m$ apart and require a minimum of 0.9 for the cross correlation.
 Four lands from different bullets are flagged as problematic in this respect.
@@
 This is a weighted average between a false positive error rate of 0.000098 and an error rate of false negatives of 0.22.
-\hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known matches and non-matches (see Figures~\ref{fig:treeresults} and ~\ref{fig:tree-forest}).}
+\hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known matches and non-matches (see Figures~\ref{fig:treeresult} and ~\ref{fig:tree-forest}).}
 Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.

### 2015-12-20 imaging-paper.Rnw
cash issues

@@
 library(knitr)
-opts_chunk$set(fig.path='Figure/fig-', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, fig.show='hold', par=TRUE, cache=F, tidy=T, concordance=TRUE, autodep=TRUE, root.dir="../", comment="|", dev="CairoPDF")
 imgdir <- "Figure/"
@@
 cc <- read.csv("csvs/crosscuts-25-new.csv")
-cc$bullet <- gsub(".*//","",cc$path)
-cc$bullet <- gsub(".x3p", "", cc$bullet)
-bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
-                  bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]
-
-excludesObs <- which(bstats$flagged)
-
-rp1 <- rpart(match~., bstats[-excludesObs,includesVar])  # doesn't include cms at all !!!!
+# cc$bullet <- gsub(".*//","",cc$path)
+# cc$bullet <- gsub(".x3p", "", cc$bullet)
+# bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
+#                   bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]
+
+#excludesObs <- which(bstats$flagged)
+
+rp1 <- rpart(match~., subset(bstats, !flagged)[,includesVar])  # doesn't include cms at all !!!!
 
@@
 set.seed(20151215)
-rtrees <- randomForest(factor(match)~., data=bstats[-excludesObs,includesVar], ntree=300)
+rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=300)
 

### 2015-12-20 imaging-paper.Rnw
centering all pics

@@
   \centering
-\begin{subfigure}[b]{\linewidth}
+\begin{subfigure}[b]{\linewidth}\centering
 \caption{View from the side.\label{fig:sidex3p}}{%
@@
 \end{subfigure}    
-\begin{subfigure}[b]{\linewidth}
+\begin{subfigure}[b]{\linewidth}\centering
     \caption{View from the top.\label{fig:topx3p}}{%
@@
 \begin{figure}[hbtp]
+  \centering
 <<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{\linewidth}
+  \centering
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
@@
 \end{subfigure}    
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{\linewidth}
+  \centering
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:loess_step1}Step 1 the loess fit algorithm.
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm.
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm.
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{\linewidth}
+  \centering
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
@@
 \begin{figure}[hbtp]
+  \centering
 <<loessfits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
@@
 \centering
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
@@
 \end{subfigure}    
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
     \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
@@
 \begin{figure}[hbtp]
+  \centering
 <<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.7\\linewidth'>>=
@@
 \begin{figure}[hbtp]
+  \centering
 <<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
@@
 \begin{figure}[hbtp]
+  \centering
 <<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
@@
 \centering
-    \begin{subfigure}[t]{\linewidth}
+    \begin{subfigure}[t]{\linewidth}\centering
 \caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
@@
 \end{subfigure}    
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
     \caption{Rectangles in the back identify a line on one of the bullets.
@@
 \begin{figure}[hbtp]
+  \centering
 <<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.7\\linewidth', warning=FALSE>>=
@@
 \begin{figure}[hbtp]
+  \centering
 <<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.7\\linewidth', warning=FALSE>>=
@@
 \begin{figure}
-\begin{subfigure}[t]{.49\textwidth}
+  \centering
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{Br6 Bullet 2-1}
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{Br9 Bullet 2-4}
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{Unkn Bullet B-2}
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{Unkn Bullet Q-4}
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{\textwidth}
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Maximal number of CMS between unknown bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
@@
 \end{subfigure}
-\begin{subfigure}[t]{\textwidth}
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Overlaid lines of C-3 and the land with the top matching CMS.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{\textwidth}
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Overlaid lines of C-3 and the land with the second top matching CMS.}
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{\linewidth}
+  \centering
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{Distribution of maximal CMS.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
@@
 \begin{figure}[hbtp]
+  \centering
 <<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.7\\linewidth', warning=FALSE>>=
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{\linewidth}
+  \centering
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
@@
 \begin{figure}[hbtp]
-\begin{subfigure}[t]{\linewidth}
+  \centering
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}
+\begin{subfigure}[t]{\linewidth}\centering
 \caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
@@
 \begin{figure}[hbtp]
+  \centering
 <<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\linewidth'>>=
@@
 \begin{figure}[hbtp]
+  \centering
 <<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='.7\\linewidth'>>=
@@
 \begin{figure}[hbtp]
+  \centering
 <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\linewidth'>>=
@@
 \begin{figure}[hbtp]
+  \centering
 <<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.7\\linewidth', warning = FALSE>>=

### 2015-12-20 imaging-paper.Rnw
comments for publication

@@
 \tableofcontents
+
+\hh{
+Todo list for journal shape-up: there is a lot of good material, we mainly need to focus on what story to tell and make that a solid argument.
+\begin{enumerate}
+\item take out a few things: no circular fit, no figures 9 and 10 
+\item do we need the web app?
+\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
+\item add: formula for loess fit and its s.e.
+\item add: comparison of error rates to those of FEs in Hamby study.
+\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
+\item add: Why do we take the lowest stable cross section?
+\end{enumerate}
+}
 
@@
 Both of these situations are assessed and discussed in more detail below:
-Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$.
+Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$.
 These are compared to a cross section at 100$\mu m$.
@@
   \centering
-<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 12, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100, 450, by = 50)
+crosscuts <- seq(100, 400, by = 50)
 lof <- processBullets(paths[5], x = crosscuts)
@@
 grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
-             lop[[7]],
              ncol = 1)

### 2015-12-21 imaging-paper.Rnw
removing all specific formatting

@@
-\documentclass[review]{elsarticle}
+\documentclass[11pt]{article}
+\usepackage[margin = 1in]{geometry}
 
@@
 \modulolinenumbers[5]
-
-\journal{Journal of Forensic Science International}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-%\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{setspace}
+\doublespacing
 
@@
 \usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
+\usepackage{natbib}
 \usepackage{float}
@@
 
+\title{Random Forests for Automatic Matching Bullet-Lands}
+\author{Eric Hare{\footnote{Corresponding author: erichare@iastate.edu}}, Heike Hofmann, Alicia Carriquiry\\
+Department of Statistics and Statistical Laboratory, Iowa State University}
+
 \begin{document}
-
-\begin{frontmatter}
-
-\title{Random Forests for Automatic Matching Bullet-Lands}
-\tnotetext[mytitlenote]{Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}
-
-%% Group authors per affiliation:
-\author{Eric Hare, Heike Hofmann, Alicia Carriquiry}
-\address{Department of Statistics and Statistical Laboratory, Iowa State University}
-%\fntext[myfootnote]{Since 1880.}
-
-%% or include affiliations in footnotes:
-%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
-%\ead[url]{www.elsevier.com}
-
-%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{erichare@iastate.edu}
-
-% \address[mymainaddress]{Department of Statistics and Statistical Laboratory, Iowa State University}
-% \address[mysecondaryaddress]{secondary address -- CS?}
+\maketitle
 
@@
 
-\begin{keyword}
-\hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Machine Learning, Feature importance.
-\end{keyword}
-
-\end{frontmatter}
-
+\paragraph{Keywords:} \hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Machine Learning, Feature importance.
+
+\newpage
 \linenumbers
@@
 
+<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
+#rm(list=ls())
+#wd <- getwd()
+library(extrafont)
+library(knitr)
+imgdir <- "Figure/"
+codedir <- "code"
+datadir <- "app/images/Hamby252_3DX3P1of2/"
+
+options(replace.assign=TRUE,scipen=3, digits=2)
+bstats <- read.csv("data-25-25/bullet-stats.csv")
+
+scrubPath <- function(x) {
+  splits <- strsplit(as.character(x), split="/")
+  last <- sapply(splits, function(x) x[length(x)])
+  gsub(".x3p","", last)
+}
+
+@
+
+\section{Introduction}
 \hh{
@@
 }
-
-<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
-#rm(list=ls())
-#wd <- getwd()
-library(extrafont)
-library(knitr)
-imgdir <- "Figure/"
-codedir <- "code"
-datadir <- "app/images/Hamby252_3DX3P1of2/"
-
-options(replace.assign=TRUE,scipen=3, digits=2)
-bstats <- read.csv("data-25-25/bullet-stats.csv")
-
-scrubPath <- function(x) {
-  splits <- strsplit(as.character(x), split="/")
-  last <- sapply(splits, function(x) x[length(x)])
-  gsub(".x3p","", last)
-}
-
-@
-
-\section{Introduction}
-
-% Rough outline:
-% \begin{enumerate}
-% \item DONE Description of the problem: Standard operating practice, 2009 NAS report
-% \item DONE Description of the Results
-% \item DONE Walkthrough of the process
-% \item \hh{XXX Alternative approaches: IBIS database and the NRC report\citep{nap:2008} against it\footnote{Fox report on the NRC report \url{http://goo.gl/XurSVl}.
-% NRC report: \url{http://www.nap.edu/catalog/12162/ballistic-imaging}} }
-% \end{enumerate}
-
-\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'.
-No algorithm implemented/available, no error rates of automated matching are given XXX}
-
 Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
-This process has broad applicability in terms of the conviction of criminals in the United States criminal justice system.
+This process has broad applicability in terms of the conviction  in the United States criminal justice system.
 Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny.
@@
 ~\citep{NAS:2009}.
-The report concluded that minimal research had been done to explore the reliability of these ballistic identification methods.
-In fact, the report states that forensic evidence ``including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
+The report states that ``[m]uch forensic evidence -- including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
 
@@
 We propose a framework which allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
+
+\hh{XXX More detailed discussion of automatic approaches and database discussion XXX}
+\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'.
+No algorithm implemented/available, no error rates of automated matching are given XXX}
 
@@
 \caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
-<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
+<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.5\\linewidth'>>=
 qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
@@
 \caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
-<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.7\\linewidth'>>=
+<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.5\\linewidth'>>=
 #qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
@@
   \centering
-<<circlefits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
       geom="line", size=I(.75), colour=I("grey70")) + 
-  facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
+  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
@@
 For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
-<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
 br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
@@
 Grooves and peaks are detected, if they are not within the same window.}
-<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
 subdata <- br111_bullet
@@
 Values outside these boundaries are removed (shown in grey)}
-<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
     lengthdiff <- length(subdata$value) - length(smoothed_truefalse)
@@
 \caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
-<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
 
@@
 \caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
-<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.7\\linewidth', warning=FALSE>>=
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE>>=
 my.loess$resid + ylab("Residuals of loess Fit")
@@
   \centering
-<<loessfits, echo=FALSE, fig.width=7, fig.height=7, out.width='\\linewidth', warning=FALSE>>=
+<<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
 br121.groove <- get_grooves(br121_bullet)
@@
       geom="line", size=I(.75), colour=I("grey70")) + 
-  facet_wrap(~part, ncol=2) + scale_colour_brewer(palette="Paired") +
+  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
@@
 \caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
-<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
+<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
 qplot(y, resid, group=bullet, colour=bullet, data=lof, 
@@
     \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
-<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.7\\textwidth'>>=
+<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
 # qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
@@
   \centering
-<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.7\\linewidth'>>=
+<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.5\\linewidth'>>=
 lag <- ccf$lag[which.max(ccf$acf)]
@@
   \centering
-<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 12, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 12, out.width = '.5\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
   \centering
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.7\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.5\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
 \caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
-<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
+<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
 qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
@@
 \caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
-<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.7\\textwidth', warning=FALSE>>=
+<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
 #images <- file.path(datadir, dir(datadir))
@@
 \label{fig:smoothcutd}}{%
-<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.7\\textwidth', warning=FALSE>>=
+<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.5\\textwidth', warning=FALSE>>=
 peaks1$lines$bullet <- b12[1]
@@
   \centering
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.7\\linewidth', warning=FALSE>>=
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\linewidth', warning=FALSE>>=
 br111.groove <- get_grooves(br111_bullet)
@@
   \centering
-<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.7\\linewidth', warning=FALSE>>=
+<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.5\\linewidth', warning=FALSE>>=
 
@@
 \caption{Maximal number of CMS between unknown bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
-<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\textwidth'>>=
+<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
 load("data-25-25/unkn9.RData")
@@
 \caption{Overlaid lines of C-3 and the land with the top matching CMS.}
-<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.7\\textwidth', warning = FALSE>>=
+<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\textwidth', warning = FALSE>>=
 res <- reslist[[which.max(cmsdist)]]  
@@
 \caption{Overlaid lines of C-3 and the land with the second top matching CMS.}
-<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.7\\textwidth', warning = FALSE>>=
+<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\textwidth', warning = FALSE>>=
 res <- reslist[[which(cmsdist==14)]]  # number 2
@@
 \caption{Distribution of maximal CMS.}
-<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
 ggplot(data=subset(bstats, !flagged)) + 
@@
 \caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
-<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
 ggplot(data=subset(bstats, !flagged)) + 
@@
 \caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
-<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
 x <- seq(0,29, by=0.03)
@@
   \centering
-<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.7\\linewidth', warning=FALSE>>=
+<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\linewidth', warning=FALSE>>=
 load("data-25-25/unkn47.RData")
@@
 \caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
-<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
+<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\linewidth'>>=
 ggplot(data=bstats) + 
@@
 \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
-<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
 ggplot(data=bstats) + 
@@
 \caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
-<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.7\\linewidth'>>=
+<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\linewidth'>>=
 ggplot(data=bstats) + theme_bw() + 
@@
 \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
-<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
 ggplot(data=bstats) + 
@@
 \caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
-<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.7\\linewidth'>>=
+<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
 x <- seq(0,1, by=0.001)
@@
   \centering
-<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\linewidth'>>=
+<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\linewidth'>>=
 library(rpart)
@@
   \centering
-<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='.7\\linewidth'>>=
+<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='.5\\linewidth'>>=
 qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
@@
   \centering
-<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\linewidth'>>=
+<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\linewidth'>>=
 bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
@@
   \centering
-<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.7\\linewidth', warning = FALSE>>=
+<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.5\\linewidth', warning = FALSE>>=
 qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
@@
 
-\section*{References}
-
+Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.
+
+%% Harvard
+\bibliographystyle{model2-names.bst}
+%\biboptions{authoryear}
 \bibliography{references}

### 2015-12-22 imaging-paper.Rnw
cleaning up

@@
 
-@
-
-\section{Introduction}
-\hh{
-Todo list for journal shape-up: there is a lot of good material, we mainly need to focus on what story to tell and make that a solid argument.
-\begin{enumerate}
-\item take out a few things: no circular fit, no figures 9 and 10 
-\item do we need the web app?
-\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
-\item add: formula for loess fit and its s.e.
-\item add: comparison of error rates to those of FEs in Hamby study.
-\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
-\item add: Why do we take the lowest stable cross section?
-\end{enumerate}
-}
-Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
-This process has broad applicability in terms of the conviction  in the United States criminal justice system.
-Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny.
-In 2005, in \emph{United States vs.
-Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar.
-Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\citep{giannelli:2011}.
-
-In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
-~\citep{NAS:2009}.
-The report states that ``[m]uch forensic evidence -- including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
-
-Rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet during the firing process.
-These marks are assumed to be unique to the barrel, as described in a 1992 AFTE article~\citep{afte:1992}.
-``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."
-
-The primary goal in this type of forensic analysis is to identify these striation markings in an attempt to match two bullets.
-Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help us determine whether two bullets are a match.
-However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
-We propose a framework which allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
-
-\hh{XXX More detailed discussion of automatic approaches and database discussion XXX}
-\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'.
-No algorithm implemented/available, no error rates of automated matching are given XXX}
-
-The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
-We then describe a web application for visualizing and working with bullet surfaces.
-We proceed to describing an automatic matching routine which we evaluate on a small database of reference bullets.
-
-\section{Methodology}
-Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
-Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
-
-3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
-x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
-x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
-
-\begin{figure}[hbtp]
-  \centering
-\begin{subfigure}[b]{\linewidth}\centering
-\caption{View from the side.\label{fig:sidex3p}}{%
-      \includegraphics[width=\linewidth]{images/sidex3p.png}
-    }
-\end{subfigure}    
-\begin{subfigure}[b]{\linewidth}\centering
-    \caption{View from the top.\label{fig:topx3p}}{%
-    \includegraphics[width=\linewidth]{images/topx3p.png}
-    }
-\end{subfigure}
-\caption{An example of a scan of a bullet land from one groove to the next. }
-\end{figure}
-Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
-For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
-An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
-These figures show side and top profiles of the land respectively.
-The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process.
-Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt.
-The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.
-
-An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the heights across values of the other coordinate.
-Figure \ref{fig:fixedX} shows a plot resulting from such a procedure.
-It can be seen that the global structure of the land dominates the appearance of the plot.
-The grooves can be clearly identified on the left and right side, and the curvature of the surface is the most visible feature in the middle.
-
-\begin{figure}[hbtp]
-  \centering
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='\\linewidth'>>=
 library(RColorBrewer)
@@
 library(sm)
+
+@
+
+\section{Introduction}
+\hh{
+Todo list for journal shape-up: there is a lot of good material, we mainly need to focus on what story to tell and make that a solid argument.
+\begin{enumerate}
+\item take out a few things: no circular fit, no figures 9 and 10 
+\item do we need the web app?
+\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
+\item add: formula for loess fit and its s.e.
+\item add: comparison of error rates to those of FEs in Hamby study.
+\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
+\item add: Why do we take the lowest stable cross section?
+\end{enumerate}
+}
+Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
+This process has broad applicability in terms of the conviction  in the United States criminal justice system.
+Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny.
+In 2005, in \emph{United States vs.
+Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar.
+Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\citep{giannelli:2011}.
+
+In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
+~\citep{NAS:2009}.
+The report states that ``[m]uch forensic evidence -- including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
+
+Rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet during the firing process.
+These marks are assumed to be unique to the barrel, as described in a 1992 AFTE article~\citep{afte:1992}.
+``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."
+
+The primary goal in this type of forensic analysis is to identify these striation markings in an attempt to match two bullets.
+Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help us determine whether two bullets are a match.
+However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
+We propose a framework which allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
+
+\hh{XXX More detailed discussion of automatic approaches and database discussion XXX}
+\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'.
+No algorithm implemented/available, no error rates of automated matching are given XXX}
+
+The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
+We then describe a web application for visualizing and working with bullet surfaces.
+We proceed to describing an automatic matching routine which we evaluate on a small database of reference bullets.
+
+\section{Methodology}
+Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
+Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
+
+3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
+x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
+x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+
+\begin{figure}[hbtp]
+  \centering
+\begin{subfigure}[t]{.49\linewidth}\centering
+\caption{View from the side.\label{fig:sidex3p}}{%
+      \includegraphics[width=\linewidth]{images/sidex3p.png}
+    }
+\end{subfigure}    
+\begin{subfigure}[t]{.49\linewidth}\centering
+    \caption{View from the top.\label{fig:topx3p}}{%
+    \includegraphics[width=\linewidth]{images/topx3p.png}
+    }
+\end{subfigure}
+\caption{An example of a scan of a bullet land from one groove to the next. }
+\end{figure}
+Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
+For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
+An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
+These figures show side and top profiles of the land respectively.
+The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process.
+Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt.
+The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.
+
+An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the heights across values of the other coordinate.
+Figure \ref{fig:fixedX} shows a plot resulting from such a procedure.
+It can be seen that the global structure of the land dominates the appearance of the plot.
+The grooves can be clearly identified on the left and right side, and the curvature of the surface is the most visible feature in the middle.
+
+\begin{figure}[hbtp]
+  \centering
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\linewidth'>>=
 
@@
   \centering
-\begin{subfigure}[t]{\linewidth}\centering
-\caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
-<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.5\\linewidth'>>=
+\begin{subfigure}[b]{.49\linewidth}\centering
+\caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
+<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
 qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
@@
 \end{subfigure}    
-\begin{subfigure}[t]{\linewidth}\centering
-\caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
-<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='.5\\linewidth'>>=
+\begin{subfigure}[b]{.49\linewidth}\centering
+\caption{\label{fig:residualb} Residual structure at $x = 100.00$}
+<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
 #qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
@@
 <<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
+db2$land <- db2$part
+db1$land <- db1$part
 qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
       geom="line", size=I(.75), colour=I("grey70")) + 
-  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
+  facet_wrap(~land, ncol=3, labeller="label_both") + 
+  scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 
   geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
-             data = filter(db1, part==5, x == 100)[,c("y", "resid", "x")]) +
+             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
   ylab("Residuals from circular fit") +
@@
 @
-\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with the bullet 2-1.}
+\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
 \end{figure}
@@
   \centering
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{.49\linewidth}\centering
 \caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
-<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
 
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{.49\linewidth}\centering
 \caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
-<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE>>=
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
 my.loess$resid + ylab("Residuals of loess Fit")
@@
 actual.df <- my.loessactual$data[1:minrows,]
-
+melt.df$land <- melt.df$part
 qplot(y, resid, data=melt.df, #colour=factor(x), 
       geom="line", size=I(.75), colour=I("grey70")) + 
-  facet_wrap(~part, ncol=3) + scale_colour_brewer(palette="Paired") +
+  facet_wrap(~land, ncol=3, labeller="label_both") + 
+  scale_colour_brewer(palette="Paired") +
   theme_bw() + theme(legend.position="bottom") + 

### 2015-12-22 imaging-paper.Rnw
re-formatting

@@
 \centering
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[b]{.49\linewidth}\centering
 \caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
-<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
+<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
 qplot(y, resid, group=bullet, colour=bullet, data=lof, 
@@
 \end{subfigure}    
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[b]{.49\linewidth}\centering
     \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
-<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
+<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
 # qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
@@
 #   theme(legend.position="none")
-qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
-      geom="line") + 
+qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
+      geom="line") + xlab("y") +
   scale_colour_manual("", values=cols) + theme_bw() +
@@
   \centering
-<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 12, out.width = '.5\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
 grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
-             ncol = 1)
+             ncol = 2)
 @
@@
   \centering
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 7, fig.height = 10, out.width = '.5\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100,225, by=25)
+crosscuts <- seq(100,250, by=25)
 lof <- processBullets(paths[37], x = crosscuts)
@@
 grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
-             lop[[4]], lop[[5]], 
-             ncol=1)
+             lop[[4]], lop[[5]], lop[[6]],
+             ncol=2)
 @
@@
   \centering
-<<barrel2, echo=FALSE, fig.width = 7, fig.height=12, out.width='.5\\linewidth', warning=FALSE>>=
+<<barrel2, echo=FALSE, fig.width = 10, fig.height=6, out.width='\\linewidth', warning=FALSE>>=
 
@@
              list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
-             ncol = 1)
+             ncol = 2)
 @
@@
 \end{subfigure}
-\begin{subfigure}[t]{\textwidth}\centering
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{Overlaid lines of C-3 and the land with the top matching CMS.}
-<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\textwidth', warning = FALSE>>=
+<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
 res <- reslist[[which.max(cmsdist)]]  
@@
 \end{subfigure}
-\begin{subfigure}[t]{\textwidth}\centering
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{Overlaid lines of C-3 and the land with the second top matching CMS.}
-<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\textwidth', warning = FALSE>>=
+<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
 res <- reslist[[which(cmsdist==14)]]  # number 2
@@
 After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0040.
-This is a weighted average between a false positive error rate of 0.000098 and an error rate of false negatives of 0.22.
+This is a weighted average between a false positive error rate of $9.8 \times 10^{-5}$ and an error rate of false negatives of 0.22.
 \hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known matches and non-matches (see Figures~\ref{fig:treeresult} and ~\ref{fig:tree-forest}).}
@@
 Our matching  algorithm is open: the code is readily available in form of the R package x3prplus~\citep{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
-Obviously, this is only a first step, but in times of the Netflix prize~\citep{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home:\url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
+Obviously, this is only a first step, but in times of the Netflix prize~\citep{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home: \url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
 

### 2015-12-22 imaging-paper.Rnw
reference to NIST standard measurement paper
couple of comments for changes/alternative solutions

@@
 No algorithm implemented/available, no error rates of automated matching are given XXX}
+\hh{XXX Interesting paper by \citet{ma:2004} with  emphasis on ccf and acf.
+They are using a stylus instrument for measureing the big curvature XXX -- the advantage of our approach is that nobody needs to look at anything until after the matching is done.
+}
+\hh{XXX Hamby study: 507 participants returned results.
+If a match was reported, these matches were correct.
+But eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.}
+% \hh{XXX for Hamby study: automated systems were successfully used for matching:
+% \begin{verbatim}
+% In addition to individuals examining the test sets using optical comparison microscopy, five test sets were examined using ‘ballistics’ imaging equipment.
+The test sets were examined using the following systems with correct answers reported by the participants.
+This information indicates that these systems – when properly used – can provide appropriate data:
+% \begin{itemize}
+% \item Intelligent Automation’s SciClopsTM - Dr. Ben Bachrach (Maryland, United States);
+% \item Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
+% \item Integrated Ballistics Identification System (IBIS) TM – Mr. Robert Thompson (California, United States);
+% \item BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)
+% \end{itemize}
+% \end{verbatim}. }
+%
 

### 2015-12-22 imaging-paper.Rnw
just changing

--- 2015-12-22 imaging-paper.Rnw
@@
 \begin{abstract}
+some change, just for fun
 In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
@@
 No algorithm implemented/available, no error rates of automated matching are given XXX}
+\hh{XXX Interesting paper by \citet{ma:2004} with  emphasis on ccf and acf.
+They are using a stylus instrument for measureing the big curvature XXX -- the advantage of our approach is that nobody needs to look at anything until after the matching is done.
+}
+\hh{XXX Hamby study: 507 participants returned results.
+If a match was reported, these matches were correct.
+But eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.}
+\hh{XXX for Hamby study: automated systems were successfully used for matching:
+ Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
+Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
+Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
+BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)
+ }
 

### 2015-12-22 imaging-paper.Rnw
Alicia's color
sweave options out
hopefully last style change

@@
-\documentclass[11pt]{article}
-\usepackage[margin = 1in]{geometry}
-
+% Template for the submission to:
+%   The Annals of Probability           [aop]
+%   The Annals of Applied Probability   [aap]
+%   The Annals of Statistics            [aos] 
+%   The Annals of Applied Statistics    [aoas]
+%   Stochastic Systems                  [ssy]
+%
+%Author: In this template, the places where you need to add information
+%        (or delete line) are indicated by {???}.  Mostly the information
+%        required is obvious, but some explanations are given in lines starting
+%Author:
+%All other lines should be ignored.  After editing, there should be
+%no instances of ??? after this line.
+
+% use option [preprint] to remove info line at bottom
+% journal options: aop,aap,aos,aoas,ssy
+% natbib option: authoryear
+\documentclass[aoas]{imsart}
+
+\usepackage{amsthm,amsmath,natbib}
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+
+% provide arXiv number if available:
+\arxiv{arXiv:0000.0000}
+
+% put your definitions there:
+\startlocaldefs
+\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
+\usepackage{subcaption}
 \usepackage{lineno,hyperref}
 \modulolinenumbers[5]
-\usepackage{setspace}
-\doublespacing
-
-\usepackage[USenglish]{babel}
-\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
-\usepackage{amsmath,amsfonts,amsthm} % Math packages
-\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
-\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
-\usepackage{booktabs} % Horizontal rules in tables
-\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
-\usepackage{natbib}
-\usepackage{float}
-\usepackage{subcaption}
-
-% declare the path(s) where your graphic files are
-\graphicspath{{Figure/}{Images/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
-\DeclareGraphicsExtensions{.pdf,.jpg,.png}
-
-\usepackage{sectsty} % Enables custom section titles
-\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands
-
-\usepackage{fancyhdr} % Needed to define custom headers/footers
-\pagestyle{fancy} % Enables the custom headers/footers
-\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")
-
-\usepackage{url}
-
-\usepackage{color}
+
 \newcommand{\hh}[1]{{\color{magenta} #1}}
 \newcommand{\eh}[1]{{\color{blue} #1}}
-
-    \renewcommand{\topfraction}{0.99}	% max fraction of floats at top
-    \renewcommand{\bottomfraction}{0.99}	% max fraction of floats at bottom
-% parameters for text pages
-\setcounter{topnumber}{2}
-    \setcounter{bottomnumber}{2}
-    \setcounter{totalnumber}{4}     % 2 may work better
-    \setcounter{dbltopnumber}{2}    % for 2-column pages
-    \renewcommand{\dbltopfraction}{0.9}	% fit big float above 2-col. text
-    \renewcommand{\textfraction}{0.07}	% allow minimal text w. figs
-    %   Parameters for FLOAT pages (not text pages):
-    \renewcommand{\floatpagefraction}{0.7}	% require fuller float pages
-	% N.B.: floatpagefraction MUST be less than topfraction !!
-    \renewcommand{\dblfloatpagefraction}{0.7}	% require fuller float pages
-
-%----------------------------------------------------------------------------------------
-
-\title{Random Forests for Automatic Matching Bullet-Lands}
-\author{Eric Hare{\footnote{Corresponding author: erichare@iastate.edu}}, Heike Hofmann, Alicia Carriquiry\\
-Department of Statistics and Statistical Laboratory, Iowa State University}
+\newcommand{\ac}[1]{{\color{darkgreen} #1}}
+
+\endlocaldefs
 
 \begin{document}
-\maketitle
-
+
+\begin{frontmatter}
+
+% "Title of the paper"
+\title{Automatic Matching of Bullet Lands}
+\runtitle{Automatic Matching of Bullet Lands}
+
+% indicate corresponding author with \corref{}
+\begin{aug}
+\author{\fnms{Eric} \snm{Hare}\corref{}\ead[label=e1]{erichare@iastate.edu}\thanksref{t1, m1}},
+\author{\fnms{Heike} \snm{Hofmann}\ead[label=e2]{hofmann@iastate.edu}\thanksref{m1}}
+\and
+\author{\fnms{Alicia} \snm{Carriquiry}\ead[label=e3]{alicia@iastate.edu}\thanksref{m1}}
+
+\thankstext{t1}{Thanks to somebody \hh{XXX acknowledge the grant here? XXX}} 
+\affiliation{Iowa State University\thanksmark{m1}}
+
+\address{Address of the Authors\\
+Department of Statistics and Statistical Laboratory\\ Snedecor Hall\\
+Ames Iowa 50011-1210\\
+\printead{e1}\\
+\phantom{E-mail:\ }\printead*{e2}\\
+\phantom{E-mail:\ }\printead*{e3}}
+
+\runauthor{E. Hare et al.}
+\end{aug}
 \begin{abstract}
-some change, just for fun
 In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
@@
 
-\paragraph{Keywords:} \hh{XXX not the right keywords yet} 3d Topological Surface Measurement, Data visualization, Machine Learning, Feature importance.
-
-\newpage
+%\begin{keyword}[class=MSC]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
+
+\begin{keyword}
+\kwd{3d Topological Surface Measurement}
+\kwd{Data visualization}
+\kwd{Machine Learning}
+\kwd{Feature importance}
+\kwd{Cross-correlation function}
+\end{keyword}
+
+\end{frontmatter}
+
 \linenumbers
-
-\tableofcontents
 
@@
   \centering
-\begin{subfigure}[t]{.49\linewidth}\centering
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{View from the side.\label{fig:sidex3p}}{%
-      \includegraphics[width=\linewidth]{images/sidex3p.png}
-    }
-\end{subfigure}    
-\begin{subfigure}[t]{.49\linewidth}\centering
+      \includegraphics[width=\textwidth]{images/sidex3p.png}
+    }
+\end{subfigure}    
+\begin{subfigure}[t]{.49\textwidth}\centering
     \caption{View from the top.\label{fig:topx3p}}{%
-    \includegraphics[width=\linewidth]{images/topx3p.png}
+    \includegraphics[width=\textwidth]{images/topx3p.png}
     }
@@
   \centering
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\linewidth'>>=
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=
 
@@
   \centering
-\begin{subfigure}[b]{.49\linewidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
 \caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
-<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
 qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
@@
 \end{subfigure}    
-\begin{subfigure}[b]{.49\linewidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
 \caption{\label{fig:residualb} Residual structure at $x = 100.00$}
-<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
+<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
 #qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
@@
   \centering
-<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
+<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
 db2$land <- db2$part
@@
   \centering
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{\label{fig:loess_step1}Step 1 the loess fit algorithm.
 For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
-<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
 br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm.
@@
 Grooves and peaks are detected, if they are not within the same window.}
-<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
 subdata <- br111_bullet
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm.
@@
 Values outside these boundaries are removed (shown in grey)}
-<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
     lengthdiff <- length(subdata$value) - length(smoothed_truefalse)
@@
   \centering
-\begin{subfigure}[t]{.49\linewidth}\centering
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
-<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
+<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE,  message=FALSE>>=
 
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\linewidth}\centering
+\begin{subfigure}[t]{.49\textwidth}\centering
 \caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
-<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE>>=
+<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
 my.loess$resid + ylab("Residuals of loess Fit")
@@
   \centering
-<<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\linewidth', warning=FALSE>>=
+<<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
 br121.groove <- get_grooves(br121_bullet)
@@
 \centering
-\begin{subfigure}[b]{.49\linewidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
 \caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
@@
 \end{subfigure}    
-\begin{subfigure}[b]{.49\linewidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
     \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
@@
   \centering
-<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.5\\linewidth'>>=
+<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.5\\textwidth'>>=
 lag <- ccf$lag[which.max(ccf$acf)]
@@
   \centering
-<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
   \centering
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))
@@
 \centering
-    \begin{subfigure}[t]{\linewidth}\centering
+    \begin{subfigure}[t]{\textwidth}\centering
 \caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
@@
 \end{subfigure}    
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
     \caption{Rectangles in the back identify a line on one of the bullets.
@@
   \centering
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\linewidth', warning=FALSE>>=
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
 br111.groove <- get_grooves(br111_bullet)
@@
   \centering
-<<barrel2, echo=FALSE, fig.width = 10, fig.height=6, out.width='\\linewidth', warning=FALSE>>=
+<<barrel2, echo=FALSE, fig.width = 10, fig.height=6, out.width='\\textwidth', warning=FALSE>>=
 
@@
 %' @
-\includegraphics[width=\linewidth]{images/br6-2-1-grey.png}
+\includegraphics[width=\textwidth]{images/br6-2-1-grey.png}
 \end{subfigure}
@@
 \caption{Br9 Bullet 2-4}
-\includegraphics[width=\linewidth]{images/br9-2-4-grey.png}
+\includegraphics[width=\textwidth]{images/br9-2-4-grey.png}
 \end{subfigure}
@@
 \caption{Unkn Bullet B-2}
-\includegraphics[width=\linewidth]{images/b-2-grey.png}
+\includegraphics[width=\textwidth]{images/b-2-grey.png}
 \end{subfigure}
@@
 \caption{Unkn Bullet Q-4}
-\includegraphics[width=\linewidth]{images/q-4-grey.png}
+\includegraphics[width=\textwidth]{images/q-4-grey.png}
 \end{subfigure}
@@
   \centering
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Distribution of maximal CMS.}
-<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
+<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
 ggplot(data=subset(bstats, !flagged)) + 
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
-<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
+<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
 ggplot(data=subset(bstats, !flagged)) + 
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
-<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
+<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
 x <- seq(0,29, by=0.03)
@@
   \centering
-<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\linewidth', warning=FALSE>>=
+<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\textwidth', warning=FALSE>>=
 load("data-25-25/unkn47.RData")
@@
   \centering
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
-<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\linewidth'>>=
+<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
 ggplot(data=bstats) + 
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
-<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
+<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
 ggplot(data=bstats) + 
@@
   \centering
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
-<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\linewidth'>>=
+<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
 ggplot(data=bstats) + theme_bw() + 
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
-<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
+<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
 ggplot(data=bstats) + 
@@
 \end{subfigure}
-\begin{subfigure}[t]{\linewidth}\centering
+\begin{subfigure}[t]{\textwidth}\centering
 \caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
-<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\linewidth'>>=
+<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
 x <- seq(0,1, by=0.001)
@@
   \centering
-<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\linewidth'>>=
+<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
 library(rpart)
@@
   \centering
-<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='.5\\linewidth'>>=
+<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
 qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
@@
   \centering
-<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\linewidth'>>=
+<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
 bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
@@
   \centering
-<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.5\\linewidth', warning = FALSE>>=
+<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.5\\textwidth', warning = FALSE>>=
 qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
@@
 
+% AOS,AOAS: If there are supplements please fill:
+%\begin{supplement}[id=suppA]
+%  \sname{Supplement A}
+%  \stitle{Title}
+%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
+%  \sdatatype{.pdf}" 
+%  \sdescription{Some text}
+%\end{supplement}
+
 %% Harvard

### 2015-12-23 imaging-paper.Rnw
starting the edit ... cylindrical fit into supplement, web app commented out

@@
 \linenumbers
-
+\tableofcontents
+\newpage
 <<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
@@
 The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
-We then describe a web application for visualizing and working with bullet surfaces.
-We proceed to describing an automatic matching routine which we evaluate on a small database of reference bullets.
+%We then describe a web application for visualizing and working with bullet surfaces.
+We then proceed to describing an automatic matching routine which we evaluate on a small database of reference bullets.
 
@@
 Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt.
-The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual bullet level.
-
-An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the heights across values of the other coordinate.
+The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual barrel/bullet level.
+
+An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across.
 Figure \ref{fig:fixedX} shows a plot resulting from such a procedure.
@@
 It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
-
-\subsection{Cylindrical Fit}
-
-One reasonable approach is to fit a circle to the curve and compute the residuals from this fit.
-
-For this, assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle.
-We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.
-
-We minimize the following expression:
-\begin{equation}\label{eq:circle}
-D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2,
-\end{equation}
-by differentiating $D$ with respect to $r, a,$ and $b$:
-let us assume that $x_i$ and $y_i$ are centered (i.e.
-$\sum x_i = \sum_i y_i = 0$).
-Note, if they are not, make a note of the current means, subtract them now and add them to $(\hat{a}, \hat{b})$ at the end.
-
-\noindent
-The  derivate of $D$ with respect to $r$ is:
-\begin{eqnarray*}
-\frac{d}{dr} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 r = \\
-&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right).
-\end{eqnarray*}
-At the minimum:
-\begin{equation}\label{eq:rmin}
-\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2.
-\end{equation}
-%
-%
-The  derivative of $D$ with respect to $a$ is:
-\begin{eqnarray*}
-\frac{d}{da} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 (x_i - a) = \\
-&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right].
-\end{eqnarray*}
-Using (\ref{eq:rmin}) for $nr^2$  in the equation above we get:
-\begin{eqnarray*}
-\frac{d}{da} D &=& -4 \left[  \sum_i a(x_i-a)^2 +  \sum_i a(y_i-b)^2  + \right. \\
-&& \phantom{-4 \ \ } \left . \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]  = \\
-&=& -4 \left[ \sum_i (x_i-a)^2 (a + x_i - a)  + \right.\\
-&& \phantom{-4 \ \ } \left .\sum_i (x_i - a + a) (y_i - b)^2 \right] = \\
-&=& -4 \left[ \sum_i (x_i-a)^2 x_i   + \sum_i x_i  (y_i - b)^2 \right] 
-\stackrel{\begin{array}{c}\sum_i x_i = 0\\
-\sum_i y_i = 0\end{array}}{=} \\
-&=& -4 \left[ \sum_i x_i^3   + \sum_i x_i y_i^2  - 2a s_{xx} - 2b s_{xy} \right],
-\end{eqnarray*}
-where $s_{xx} = \sum_i x_i^2, s_{xy} = \sum_i x_i y_i$ and $s_{yy} = \sum_i y_i^2$.
-
-\noindent
-Likewise, we get for the derivative of $D$ with respect to $b$:
-\begin{eqnarray*}
-\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right].
-\end{eqnarray*}
-To find the minimum we therefore get a system of two linear equations in $a$ and $b$:
-\begin{eqnarray*}
-2 s_{xx} a + 2 s_{xy} b = c_1 && \text{ with } c_1 = \sum_i x_i^3 + x_i y_i^2 \\
-2 s_{xy} a + 2 s_{yy} b = c_2 &&\text{ with } c_2 = \sum_i x_i^2 y_i + y_i^3.
-\end{eqnarray*}
-The solution to the system is:
-\begin{eqnarray*}
-\hat{a} &=& \frac{c_1 s_{yy} - c_2 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2},\\
-\hat{b} &=& \frac{c_2 s_{xx} - c_1 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2}, \text{ and}\\
-\hat{r^2} &=& \frac{1}{n}s_{xx} + \frac{1}{n}s_{yy} + \hat{a}^2 + \hat{b}^2.
-\end{eqnarray*}
-
-The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
-In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.
-Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
-The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
-For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.
-
-\begin{figure}[hbtp]
-  \centering
-\begin{subfigure}[b]{.49\textwidth}\centering
-\caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
-<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
-qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
-      geom="line", size=I(1)) +
-#  scale_colour_brewer("x", palette="Paired") + 
-  theme_bw() + 
-  geom_hline(yintercept = 0, colour="grey50") +
-  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
-  theme(legend.position="bottom")
-@
-\end{subfigure}    
-\begin{subfigure}[b]{.49\textwidth}\centering
-\caption{\label{fig:residualb} Residual structure at $x = 100.00$}
-<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
-#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
-qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
-      geom="line", size=I(1)) +
-#  scale_colour_brewer("x", palette="Paired") + 
-  geom_hline(yintercept = 0, colour="grey50") +
-  theme_bw() + 
-  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
-  theme(legend.position="bottom")
-@
-\end{subfigure}
-\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections.
-Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
-\end{figure}
-
-A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction.
-Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues.
-While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.
-
-<<bullet1, echo=FALSE, cache=TRUE>>=
-db1 <- NULL
-for (i in 1:6) {
-  bname <- sprintf(file.path(datadir, "Br1 Bullet 1-%d.x3p"), i)
-  dbi <- fortify_x3p(read.x3p(bname))
-  dbi$part <- i
-  db1 <- rbind(db1, dbi)
-}
-
-db1 <- db1 %>% group_by(part, x) %>% do (
-    data.frame(., predCircle(.$y, .$value))
-  )
-@
-
-<<bullet2, echo=FALSE, cache=TRUE>>=
-db2 <- NULL
-for (i in 1:6) {
-  bname <- sprintf(file.path(datadir, "Br1 Bullet 2-%d.x3p"), i)
-  dbi <- fortify_x3p(read.x3p(bname))
-  dbi$part <- i
-  db2 <- rbind(db2, dbi)
-}
-
-db2 <- db2 %>% group_by(part, x) %>% do (
-    data.frame(., predCircle(.$y, .$value))
-  )
-
-@
-
-\begin{figure}[hbtp]
-  \centering
-<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
-db2$land <- db2$part
-db1$land <- db1$part
-qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
-      geom="line", size=I(.75), colour=I("grey70")) + 
-  facet_wrap(~land, ncol=3, labeller="label_both") + 
-  scale_colour_brewer(palette="Paired") +
-  theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
-             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
-  ylab("Residuals from circular fit") +
-  ggtitle("Bullet 1-5 in black")
-@
-\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
-\end{figure}
 
@@
 
-\section{Web Application}
-We have implemented a web interface to automate many of the steps outlined in the previous section.
+% \section{Web Application}
+% We have implemented a web interface to automate many of the steps outlined in the previous section.
 The web interface, developed using Shiny~\citep{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
@@
 The application is available at the url \url{http://erichare.shinyapps.io/x3prproto}.
-
-\begin{figure*}
-\includegraphics[width=\textwidth]{images/app.png}
-\caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
-\end{figure*}
-
-The mouse is the primary control input for interacting with the surface plots.
+% 
+% \begin{figure*}
+% \includegraphics[width=\textwidth]{images/app.png}
+% \caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
+% \end{figure*}
+% 
+% The mouse is the primary control input for interacting with the surface plots.
 One can mouse over the plots to view the coordinate values for each point.
@@
 There are also a set of controls available at the top right of the page, which will allow the selection of some other rotation methods, taking of screenshots, and reseting the camera to the default view.
-
-Several lighting options are available to tweak in order to highlight markings in on the bullets more strongly:
-
-\begin{enumerate}
-    \item \textbf{Ambient Lighting} - Lighting which is used to simulate the light naturally present in the environment.
-    \item \textbf{Diffuse Lighting} - Setting this value higher will cause the surface to re-emit a higher proportion of light.
-    \item \textbf{Specular Lighting} - The lighting which appears on the surface when illuminated.
-    \item \textbf{Roughness Lighting} - The amount of light reflected by the roughness of the surface.
-    \item \textbf{Fresnel Lighting} - Higher values create a wider, softer lighting across the surface.
-\end{enumerate}
-
-The application includes a button which allows the automatic creation of a residual plot from the loess fits of two bullets.
+% 
+% Several lighting options are available to tweak in order to highlight markings in on the bullets more strongly:
+% 
+% \begin{enumerate}
+%     \item \textbf{Ambient Lighting} - Lighting which is used to simulate the light naturally present in the environment.
+%     \item \textbf{Diffuse Lighting} - Setting this value higher will cause the surface to re-emit a higher proportion of light.
+%     \item \textbf{Specular Lighting} - The lighting which appears on the surface when illuminated.
+%     \item \textbf{Roughness Lighting} - The amount of light reflected by the roughness of the surface.
+%     \item \textbf{Fresnel Lighting} - Higher values create a wider, softer lighting across the surface.
+% \end{enumerate}
+% 
+% The application includes a button which allows the automatic creation of a residual plot from the loess fits of two bullets.
 An x coordinate value can be selected as a cross cut, and cross sections of the two bullets at the x value are taken.
@@
 Users need to perform this procedure at different x values in order to see whether the markings indicate a match at different points along the bullet.
-
-Enhancements to the web application are still in progress. In particular, much of the automatic matching functionality described in the next section will be added in the future.
+% 
+% Enhancements to the web application are still in progress. In particular, much of the automatic matching functionality described in the next section will be added in the future.
 
 \section{Automatic matching}
-Applying the loess fit to a range of different surface cross cuts (see Figures~\ref{fig:manualmatch} and~\ref{fig:manualmatch-rgl}, with cross cuts from 50 to 150) shows the striation marks across two bullets.
+Applying the loess fit to a range of different surface cross cuts (see Figure~\ref{fig:manualmatch-rgl}, with cross cuts from 50 to 150) shows the striation marks from two bullets.
 Cross cuts of bullet 1 are shown on the left (cross cuts with values below 100) and cross cuts of bullets 2 are shown on the right (cross cuts with values above 100).
@@
 
-\begin{figure}[hbtp]
-\centering
-%' <<side-by-side, echo=FALSE, eval=FALSE>>=
-%' paths <- file.path(datadir, dir(datadir))
-%' 
-%' br111 <- read.x3p(paths[5])
-%' crosscuts <- unique(fortify_x3p(br111)$x)
-%' crosscuts <- crosscuts[crosscuts > 50]
-%' crosscuts <- crosscuts[crosscuts < 150]
-%' 
-%' list_of_fits <- lapply(crosscuts, function(x) {
-%'   br111 <- get_crosscut(paths[5], x = x)
-%'   br111.groove <- get_grooves(br111)
-%'   br111.groove$plot
-%'   fit_loess(br111, br111.groove)
-%' })
-%' 
-%' lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
-%' 
-%' br111 <- read.x3p(paths[7])
-%' crosscuts <- unique(fortify_x3p(br111)$x)
-%' crosscuts <- crosscuts[crosscuts > 50]
-%' crosscuts <- crosscuts[crosscuts < 150]
-%' 
-%' list_of_fits2 <- lapply(crosscuts, function(x) {
-%'   br111 <- get_crosscut(paths[7], x)
-%'   br111.groove <- get_grooves(br111)
-%'   br111.groove$plot
-%'   fit_loess(br111, br111.groove)
-%' })
-%' 
-%' lof2 <- lapply(list_of_fits2, function(x) x$resid$data) %>% bind_rows
-%' lof$bullet <- 1
-%' lof2$bullet <- 2
-%' 
-%' LOF <- rbind(lof, lof2)
-%' 
-%' subLOF1 <- LOF %>% filter(bullet == 1, x <= 100)
-%' subLOF2 <- LOF %>% filter(bullet == 2, x > 100)
-%' 
-%' #subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
-%' subLOF1$y <- subLOF1$y + 19*1.5625 # working now!!!
-%' subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
-%' 
-%' qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
-%'   scale_fill_gradientn(limits=c(-5,5), 
-%'                        colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
-%'                        values = c(0,0.5, 0.75, 0.9, 1)) +
-%'   theme_bw()
-%' @
-
-\includegraphics[width=.7\columnwidth]{images/side-by-side-manual-match.png}
-\caption{\label{fig:manualmatch}Manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
-\end{figure}
+% \begin{figure}[hbtp]
+% \centering
+% %' <<side-by-side, echo=FALSE, eval=FALSE>>=
+% %' paths <- file.path(datadir, dir(datadir))
+% %' 
+% %' br111 <- read.x3p(paths[5])
+% %' crosscuts <- unique(fortify_x3p(br111)$x)
+% %' crosscuts <- crosscuts[crosscuts > 50]
+% %' crosscuts <- crosscuts[crosscuts < 150]
+% %' 
+% %' list_of_fits <- lapply(crosscuts, function(x) {
+% %'   br111 <- get_crosscut(paths[5], x = x)
+% %'   br111.groove <- get_grooves(br111)
+% %'   br111.groove$plot
+% %'   fit_loess(br111, br111.groove)
+% %' })
+% %' 
+% %' lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
+% %' 
+% %' br111 <- read.x3p(paths[7])
+% %' crosscuts <- unique(fortify_x3p(br111)$x)
+% %' crosscuts <- crosscuts[crosscuts > 50]
+% %' crosscuts <- crosscuts[crosscuts < 150]
+% %' 
+% %' list_of_fits2 <- lapply(crosscuts, function(x) {
+% %'   br111 <- get_crosscut(paths[7], x)
+% %'   br111.groove <- get_grooves(br111)
+% %'   br111.groove$plot
+% %'   fit_loess(br111, br111.groove)
+% %' })
+% %' 
+% %' lof2 <- lapply(list_of_fits2, function(x) x$resid$data) %>% bind_rows
+% %' lof$bullet <- 1
+% %' lof2$bullet <- 2
+% %' 
+% %' LOF <- rbind(lof, lof2)
+% %' 
+% %' subLOF1 <- LOF %>% filter(bullet == 1, x <= 100)
+% %' subLOF2 <- LOF %>% filter(bullet == 2, x > 100)
+% %' 
+% %' #subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
+% %' subLOF1$y <- subLOF1$y + 19*1.5625 # working now!!!
+% %' subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
+% %' 
+% %' qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
+% %'   scale_fill_gradientn(limits=c(-5,5), 
+% %'                        colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
+% %'                        values = c(0,0.5, 0.75, 0.9, 1)) +
+% %'   theme_bw()
+% %' @
+% 
+% \includegraphics[width=.7\columnwidth]{images/side-by-side-manual-match.png}
+% \caption{\label{fig:manualmatch}Manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
+% \end{figure}
+
 \begin{figure}[hbtp]
@@
 \includegraphics[width=.7\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}Different view of the manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
+\caption{\label{fig:manualmatch-rgl}3d view of the manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1 after removing the curvature.}
 \end{figure}

### 2015-12-23 imaging-supplement.Rnw
starting the edit ... cylindrical fit into supplement, web app commented out

% Template for the submission to:
%   The Annals of Probability           [aop]
%   The Annals of Applied Probability   [aap]
%   The Annals of Statistics            [aos] 
%   The Annals of Applied Statistics    [aoas]
%   Stochastic Systems                  [ssy]
%
%Author: In this template, the places where you need to add information
%        (or delete line) are indicated by {???}.  Mostly the information
%        required is obvious, but some explanations are given in lines starting
%Author:
%All other lines should be ignored.  After editing, there should be
%no instances of ??? after this line.

% use option [preprint] to remove info line at bottom
% journal options: aop,aap,aos,aoas,ssy
% natbib option: authoryear
\documentclass[aoas]{imsart}

\usepackage{amsthm,amsmath,natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% provide arXiv number if available:
\arxiv{arXiv:0000.0000}

% put your definitions there:
\startlocaldefs
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{subcaption}
\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\eh}[1]{{\color{blue} #1}}
\newcommand{\ac}[1]{{\color{darkgreen} #1}}

\endlocaldefs

\begin{document}

\begin{frontmatter}

% "Title of the paper"
\title{Supplement to Automatic Matching of Bullet Lands}
\runtitle{Supplement: Automatic Matching of Bullet Lands}

% indicate corresponding author with \corref{}
\begin{aug}
\author{\fnms{Eric} \snm{Hare}\corref{}\ead[label=e1]{erichare@iastate.edu}\thanksref{m1}},
\author{\fnms{Heike} \snm{Hofmann}\ead[label=e2]{hofmann@iastate.edu}\thanksref{m1}}
\and
\author{\fnms{Alicia} \snm{Carriquiry}\ead[label=e3]{alicia@iastate.edu}\thanksref{m1}}

\affiliation{Iowa State University\thanksmark{m1}}

\runauthor{E. Hare et al.}
\end{aug}

\end{frontmatter}

\linenumbers
\tableofcontents
\newpage
<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
#rm(list=ls())
#wd <- getwd()
library(extrafont)
library(knitr)
imgdir <- "Figure/"
codedir <- "code"
datadir <- "app/images/Hamby252_3DX3P1of2/"

options(replace.assign=TRUE,scipen=3, digits=2)
bstats <- read.csv("data-25-25/bullet-stats.csv")

scrubPath <- function(x) {
  splits <- strsplit(as.character(x), split="/")
  last <- sapply(splits, function(x) x[length(x)])
  gsub(".x3p","", last)
}

library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)

@

\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=

cols = c(alpha("grey60", alpha=0.6), alpha("black", 0.5))

br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
dbr111 <- fortify_x3p(br111)

pars <- data.frame(getCircle(dbr111$y, dbr111$value))
dbr111$theta <- acos((dbr111$y-pars$x0)/pars$radius)/pi*180
dbr111 <- dbr111 %>% mutate(
  xpred = cos(theta/180*pi)*pars$radius + pars$x0,
  ypred = sin(theta/180*pi)*pars$radius + pars$y0
)

qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 99*1.5625^2), y, value, geom="line", size=I(1)) +
  geom_line(aes(x=xpred, y=ypred, group=x), 
            colour="grey30", size=0.25) +
#  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
  ylab("") +
  theme_bw() + 
  theme(legend.position="bottom") #+ coord_equal()
@
\caption{\label{fig:fixedX}Side profile of the surface measurements (in $\mu m$) of a bullet land across a fixed $x$ value. Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
\end{figure}

\section{Cylindrical Fit}

One reasonable approach is to fit a circle to the curve and compute the residuals from this fit.

For this, assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle. We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.

We minimize the following expression:
\begin{equation}\label{eq:circle}
D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2,
\end{equation}
by differentiating $D$ with respect to $r, a,$ and $b$:
let us assume that $x_i$ and $y_i$ are centered (i.e. $\sum x_i = \sum_i y_i = 0$). Note, if they are not, make a note of the current means, subtract them now and add them to $(\hat{a}, \hat{b})$ at the end.

\noindent
The  derivate of $D$ with respect to $r$ is:
\begin{eqnarray*}
\frac{d}{dr} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 r = \\
&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right).
\end{eqnarray*}
At the minimum:
\begin{equation}\label{eq:rmin}
\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2.
\end{equation}
%
%
The  derivative of $D$ with respect to $a$ is:
\begin{eqnarray*}
\frac{d}{da} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 (x_i - a) = \\
&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right].
\end{eqnarray*}
Using (\ref{eq:rmin}) for $nr^2$  in the equation above we get:
\begin{eqnarray*}
\frac{d}{da} D &=& -4 \left[  \sum_i a(x_i-a)^2 +  \sum_i a(y_i-b)^2  + \right. \\
&& \phantom{-4 \ \ } \left . \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]  = \\
&=& -4 \left[ \sum_i (x_i-a)^2 (a + x_i - a)  + \right.\\
&& \phantom{-4 \ \ } \left .\sum_i (x_i - a + a) (y_i - b)^2 \right] = \\
&=& -4 \left[ \sum_i (x_i-a)^2 x_i   + \sum_i x_i  (y_i - b)^2 \right] 
\stackrel{\begin{array}{c}\sum_i x_i = 0\\
\sum_i y_i = 0\end{array}}{=} \\
&=& -4 \left[ \sum_i x_i^3   + \sum_i x_i y_i^2  - 2a s_{xx} - 2b s_{xy} \right],
\end{eqnarray*}
where $s_{xx} = \sum_i x_i^2, s_{xy} = \sum_i x_i y_i$ and $s_{yy} = \sum_i y_i^2$.

\noindent
Likewise, we get for the derivative of $D$ with respect to $b$:
\begin{eqnarray*}
\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right].
\end{eqnarray*}
To find the minimum we therefore get a system of two linear equations in $a$ and $b$:
\begin{eqnarray*}
2 s_{xx} a + 2 s_{xy} b = c_1 && \text{ with } c_1 = \sum_i x_i^3 + x_i y_i^2 \\
2 s_{xy} a + 2 s_{yy} b = c_2 &&\text{ with } c_2 = \sum_i x_i^2 y_i + y_i^3.
\end{eqnarray*}
The solution to the system is:
\begin{eqnarray*}
\hat{a} &=& \frac{c_1 s_{yy} - c_2 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2},\\
\hat{b} &=& \frac{c_2 s_{xx} - c_1 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2}, \text{ and}\\
\hat{r^2} &=& \frac{1}{n}s_{xx} + \frac{1}{n}s_{yy} + \hat{a}^2 + \hat{b}^2.
\end{eqnarray*}

The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.  Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
The residuals are dominated, as expected, by the grooves, which show up as large positive residuals. For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:residualb} Residual structure at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction. Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues. While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.

<<bullet1, echo=FALSE, cache=TRUE>>=
db1 <- NULL
for (i in 1:6) {
  bname <- sprintf(file.path(datadir, "Br1 Bullet 1-%d.x3p"), i)
  dbi <- fortify_x3p(read.x3p(bname))
  dbi$part <- i
  db1 <- rbind(db1, dbi)
}

db1 <- db1 %>% group_by(part, x) %>% do (
    data.frame(., predCircle(.$y, .$value))
  )
@

<<bullet2, echo=FALSE, cache=TRUE>>=
db2 <- NULL
for (i in 1:6) {
  bname <- sprintf(file.path(datadir, "Br1 Bullet 2-%d.x3p"), i)
  dbi <- fortify_x3p(read.x3p(bname))
  dbi$part <- i
  db2 <- rbind(db2, dbi)
}

db2 <- db2 %>% group_by(part, x) %>% do (
    data.frame(., predCircle(.$y, .$value))
  )

@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
db2$land <- db2$part
db1$land <- db1$part
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~land, ncol=3, labeller="label_both") + 
  scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
\end{figure}

\end{document}

### 2015-12-23 imaging-paper.Rnw
cross sections are called signatures/profiles in NIST

@@
-% Template for the submission to:
-%   The Annals of Probability           [aop]
-%   The Annals of Applied Probability   [aap]
-%   The Annals of Statistics            [aos] 
-%   The Annals of Applied Statistics    [aoas]
-%   Stochastic Systems                  [ssy]
-%
-%Author: In this template, the places where you need to add information
-%        (or delete line) are indicated by {???}.  Mostly the information
-%        required is obvious, but some explanations are given in lines starting
-%Author:
-%All other lines should be ignored.  After editing, there should be
-%no instances of ??? after this line.
-
-% use option [preprint] to remove info line at bottom
-% journal options: aop,aap,aos,aoas,ssy
-% natbib option: authoryear
 \documentclass[aoas]{imsart}
@@
   \centering
-\begin{subfigure}[t]{.49\textwidth}\centering
+\begin{subfigure}[t]{.54\textwidth}\centering
 \caption{View from the side.\label{fig:sidex3p}}{%
@@
 \end{subfigure}    
-\begin{subfigure}[t]{.49\textwidth}\centering
+\begin{subfigure}[t]{.45\textwidth}\centering
     \caption{View from the top.\label{fig:topx3p}}{%
@@
 Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt.
-The direction of the rifling is a class characteristic, or a feature that pertains to a particular class of firearms and is not unique at the individual barrel/bullet level.
-
-An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across.
-Figure \ref{fig:fixedX} shows a plot resulting from such a procedure.
+The direction of the rifling is a class characteristic, i.e.\ a feature that pertains to a particular class of firearms, and is not unique at the individual barrel/bullet level.
+
+\hh{In an approach to analyze the striation pattern we extract a {\emph bullet profile} or {\emph signature} \cite{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.}
+%An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
+Figure \ref{fig:fixedX} shows a plot of a bullet signature.
 It can be seen that the global structure of the land dominates the appearance of the plot.
@@
 
-%% Harvard
-\bibliographystyle{model2-names.bst}
+\bibliographystyle{imsart-nameyear.bst}
 %\biboptions{authoryear}

### 2015-12-23 imaging-paper.Rnw
grooves identification

@@
 The smooth curve on the plot represents a segment of a perfect circle with the same radius as the bullet.
-It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
-
-\subsection{Loess Fit}
-In an attempt to improve upon the previous algorithm, we first have to eliminate the grooves from the images under the assumption that they do not contain relevant information for determining a match.
-Fortunately, the location and appearance of the grooves in these 2d plots is quite consistent.
+\hh{While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
+A discussion of the fit and the remaining residual structure can be found in the supplement.} \hh{XXX Reference to the supplement?
+XXX}
+%It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
+
+%\subsection{Loess Fit}
+\hh{Instead, we use multiple loess fits to model the overall structure in order to identify the bullet markings. }
+
+\subsection{Identify Grooves}
+%In an attempt to improve upon the previous algorithm, w
+We first have to \hh{identify} the \hh{locations of the left and right}  groove in the image.
+\hh{The grooves are not assumed to contain information relevant for determining matches, they also dominate the structure, and therefore need  to be removed.}
+Fortunately, the location and appearance of the grooves in the surface profiles is quite consistent.
 Surface measurements reach local maxima around the peak of the groove at either end of the range of $y$, we can then follow the descent of the surface measurements inwards to the valley of the groove.
@@
 \begin{enumerate}
-    \item Fix an $x$ value (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
-    \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \textbf{smoothing factor}.
-(Figure~\ref{fig:loess_step3}, smoothing factor of 35, corresponding to 55$\mu m$).
+    \item Fix an $x$ value \hh{to extract a bullet signature} (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
+    \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \emph{smoothing factor} $s$.
+(Figure~\ref{fig:loess_step3}, smoothing factor $s = 35$ corresponding to 55$\mu m$).
 %    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor as above. (Figure~\ref{fig:loess_step3}).
-    \item Determine the location of the peak of the left groove by finding the first double-smoothed value $y_i$ that is the maximum within its smoothing window (e.g.\ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$, $i = 1, ..., 17$).
-We call the location of this peak $p_{left}$ (see Figure~\ref{fig:loess_step47}).
+    \item Determine the location of the peak of the left groove by finding the first double-smoothed value $y_i$ that is the maximum within its smoothing window (e.g.\ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$, where $i$ is between 1  and $\lfloor s/2 \rfloor$).
+We call the location of this peak $p_{\ell}$ (see Figure~\ref{fig:loess_step47}).
     \item Similarly, determine the location of the valley of the left groove by finding the first double-smoothed $y_j$ that is the minimum within its smoothing window.
-Call the location of this valley $v_{left}$.
-    \item Reverse the order of the $y$ values and repeat the previous two steps to find the peak and valley of the right groove, $(p_{right}, v_{right})$.
-    \item Trim the surface measurements to values within the two grooves (i.e.\ remove all records with $y_i < v_{left}$ and $y_i > v_{right}$). (Figure \ref{fig:loess_step47}).
+Call the location of this valley $v_{\ell}$.
+    \item Reverse the order of the $y$ values and repeat the previous two steps to find the peak and valley of the right groove, $(p_{r}, v_{r})$.
+    \item Trim the surface measurements to values within the two grooves (i.e.\ remove all records with $y_i < v_{\ell}$ and $y_i > v_{r}$) (see Figure~\ref{fig:loess_step47}).
 \end{enumerate}
@@
 
-The ``smoothing factor" described in the algorithm represents the window size to use for a rolling average.
-Higher values of the smoothing factor therefore lead to more smoothed results.
+The smoothing factor $s$ introduced in the algorithm represents the window size to use for a rolling average.
+Higher values of $s$ therefore lead to more smoothed results.
 In the groove detection portion of this algorithm, we wish to remove most of the deviation but still maintain enough signal in the groove to detect it.
-Empirically, a value of 35 for the smoothing factor seems to work quite well (later in the manuscript we investigate varying the smoothing factor).
+Empirically, a value of $s = 35$ for the smoothing factor seems to work quite well (the smoothing factor is further discussed in section~\ref{sec:smoothing}).
 It is important to note that the smoothing pass is done twice.
@@
 This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass and then added back to the result~\citep{tukey:1977}.
-This has the effect of introducing a little bit more variance into the smoothed data.
-We instead performed a second smoothing pass on the smoothed data, which has the effect of weighting observations near the center highest, while the weights linearly drop off as we reach the end of the smoothing window or smoothing factor.
+This has the effect of introducing a bit more variance back into the smoothed data.
+We instead performed a second smoothing pass on the smoothed data, which has the effect of weighting observations near the center of the window the highest, while the weights linearly drop off as we reach either end of the smoothing window.
 
 Next, we fit a loess regression to the data.
-Figure \ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
+Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
 The fit seems to do a reasonable job of capturing the structure of the image.
-Figure \ref{fig:loess_resid} shows the residuals from this fit.
+Figure~\ref{fig:loess_resid} shows the residuals from this fit.
 We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet 1-5.
@@
 
-\subsection{Varying smoothing factor}
+\subsection{Varying smoothing factor}\label{sec:smoothing}
 As mentioned earlier, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.

### 2015-12-23 imaging-paper.Rnw
more into supplement

@@
-\documentclass[aoas]{imsart}
+\documentclass[aoas, preprint]{imsart}
 
@@
 
-In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
-~\citep{NAS:2009}.
+In 2009, the National Academy of Sciences published a report~\citep{NAS:2009} questioning the scientific validity of many forensic methods including firearm examination.
 The report states that ``[m]uch forensic evidence -- including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
@@
 
-\section{Methodology}
+\section{Bullet signatures}
 Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
@@
 
-\hh{In an approach to analyze the striation pattern we extract a {\emph bullet profile} or {\emph signature} \cite{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.}
+\hh{In an approach to analyze the striation pattern we extract a {\emph bullet profile} \cite{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.}
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
-Figure \ref{fig:fixedX} shows a plot of a bullet signature.
+Figure \ref{fig:fixedX} shows a plot of the side profile of a bullet land.
 It can be seen that the global structure of the land dominates the appearance of the plot.
@@
 \hh{While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of the fit and the remaining residual structure can be found in the supplement.} \hh{XXX Reference to the supplement?
+A discussion of a circular fit and the remaining residual structure can be found in the supplement.} \hh{XXX Reference to the supplement?
 XXX}
@@
 %\subsection{Loess Fit}
-\hh{Instead, we use multiple loess fits to model the overall structure in order to identify the bullet markings. }
-
-\subsection{Identify Grooves}
+\hh{Instead of a circular fit, we use multiple loess fits to model the overall structure in order to identify the bullet markings. }
+
+\subsection{Identifying grooves}
 %In an attempt to improve upon the previous algorithm, w
@@
 Surface measurements reach local maxima around the peak of the groove at either end of the range of $y$, we can then follow the descent of the surface measurements inwards to the valley of the groove.
-This marks the point at which we should trim the image. The procedure can be described as follows:
+The location of the valleys  mark the points at which we should trim the image. The procedure can be described as follows:
 
 \begin{enumerate}
-    \item Fix an $x$ value \hh{to extract a bullet signature} (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
+    \item Fix an $x$ value \hh{to extract a bullet profile} (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
     \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \emph{smoothing factor} $s$.
@@
 
+\subsection{Removing curvature}
 Next, we fit a loess regression to the data.
-Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of part 1 of bullet 1.
+
+\hh{XXX Introduce Loess formula with all of its details in here. we are using the default of {\tt loess}, which has a span of 0.75 and uses a polynomial of degree 2. XXX}
+Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of bullet 1-5.
 The fit seems to do a reasonable job of capturing the structure of the image.
 Figure~\ref{fig:loess_resid} shows the residuals from this fit.
-We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet 1-5.
-This is shown in Figure~\ref{fig:loessfits}.
-Bullet 2-1 is still the best match for bullet 1-5, but now the details of the match are more pronounced.
-
+These residuals are called the \emph{signature of bullet 1-5.}
+%
 \begin{figure}[hbtp]
@@
 \end{figure}
+%
+Figure~\ref{fig:loessfits} shows the signatures of all lands of bullet 2 (in grey) fired from the same barrel as bullet 1.
+The darker line overlaid on each of the signatures, shows signature from bullet 1-5.
+%We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet 1-5.
+This is shown in Figure~\ref{fig:loessfits}.
+Bullet 2-1 is the best match for bullet 1-5, but the alignment is not quite  correct.
 
@@
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Bullet 1-5 in black") + ylab("Residuals of loess Fit")
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Signatures of Bullet 1 in grey, Bullet 1-5 in black") + ylab("Residuals of loess fit")
 @    
@@
 % The application includes a button which allows the automatic creation of a residual plot from the loess fits of two bullets.
-An x coordinate value can be selected as a cross cut, and cross sections of the two bullets at the x value are taken.
+An x coordinate value can be selected as a cross section, and cross sections of the two bullets at the x value are taken.
 The previously described loess regression procedure is applied to the two cross sections.
@@
 \section{Automatic matching}
-Applying the loess fit to a range of different surface cross cuts (see Figure~\ref{fig:manualmatch-rgl}, with cross cuts from 50 to 150) shows the striation marks from two bullets.
-Cross cuts of bullet 1 are shown on the left (cross cuts with values below 100) and cross cuts of bullets 2 are shown on the right (cross cuts with values above 100).
-Many of the striation marks align, allowing for an easy visual assessment of the shared similarities between the two bullets.
+Applying the loess fit to a range of different surface cross sections (see Figure~\ref{fig:manualmatch-rgl} for cross sections from 50$\mu m$ to 150 $\mu m$) shows the 3d striation marks from two bullets.
+Signatures of bullet 1 are shown on the left (all cross sections with values below 100$\mu m$) and cross sections of bullets 2 are shown on the right (cross sections with values above 100$\mu m$).
+\hh{Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one bullet to the other.
+Visually, this allows for an easy assessment of these two bullet lands as a match.
+However, this match relies on visual inspection and can therefore not be completely objective.
+The goal of this section is to eliminate the need for a visual inspection during the matching process and replace it by an automatic algorithm.
+This also allows for a quantification of the quality of the match.}
 
@@
 @
+
+\subsection{Horizontal Alignment}
+A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.
+This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$.
+Striation marks show up in this representations as peaks and valleys.
+The individual characteristics are quite prominent and, again, suggest a match between the lands.
+A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities.
+%
 \begin{figure}[hbtp]
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
+\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
 <<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
+    \caption{Aligned (by horizontal shift) signatures.\label{fig:crosscutX}}{%
 <<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
-# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
-#       geom="line", alpha=I(0.6)) + 
-#   scale_colour_brewer(palette="Set1") + theme_bw() +
-#   theme(legend.position="none")
 qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
@@
   scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position="none") + ylab("Residuals of loess fit")
-@
-    }
-\end{subfigure}
-\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
-\end{figure}
-
-\subsection{Horizontal Alignment}
-Using (for now) a single cross section across the surface, we extract the values and remove the class characteristics using a loess fit as before.
-This enables us to focus on the individual characterists.
-Striation marks show up in this representations as peaks and valleys.
-Figure~\ref{fig:cross100} shows lines of the surface cross section at $x = 100$.
-A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.
-This horizontal shift is based on the cross-correlation between the two surface cross sections.
-Let $f(t)$ and $g(t)$ define the values of the cross section at $t$, then the cross-correlation between $f$ and $g$ at lag $k$ is defined as
+  theme(legend.position = c(.9, .2),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylab("Residuals of loess fit")
+@
+    }
+\end{subfigure}
+\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$.
+A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
+\end{figure}
+%
+This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0 $\mu m$ and about 2500 $\mu m$, 1.5625 $\mu m$ apart.
+The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
 \[
@@
 @
-\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
-At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections.
-Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
+\caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
+At a lag of $\Sexpr{lag}$ the correlation peaks, indicating the largest amount of agreement between the signatures.
+Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}
 \end{figure}
@@
 \subsection{Varying cross sections}
-The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up.
-This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality.
+The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
+This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can extract signatures from multiple cross sections of the same bullet for an initial assessment of its quality.
+\hh{XXX shift cross section discussion into the supplement XXX}
 Both of these situations are assessed and discussed in more detail below:
-Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$.
-These are compared to a cross section at 100$\mu m$.
-Initially this comparison constitutes an almost perfect match between the two cross sections.
-However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.
-Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
-Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections.
-\begin{figure}[hbtp]
-  \centering
-<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
-paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100, 400, by = 50)
-lof <- processBullets(paths[5], x = crosscuts)
-lof$bullet <- paste(lof$bullet, lof$x)
-
-reslist <- lapply(crosscuts[-1], function(cc) {
-#  browser()
-  b2 <- subset(lof, x %in% c(cc, 100))
-  lofX <- bulletSmooth(b2)
-  bAlign = bulletAlign(lofX)
-  lofX <- bAlign$bullet
-    b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
-
-#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-#  lines <- striation_identify(lofX, threshold = threshold)
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
-})
-
-ccfs <- sapply(reslist, function(res) res$ccf)
-
-lop <- lapply(reslist, function(res) {
-ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
-  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
-  scale_linetype_discrete("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
-    ylab("") + xlab("")
-})
-
-grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
-             ncol = 2)
-@
-\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections.
-Cross section at $x = 100$ is compared to cross sections every 50$\mu m$.
-With every step away from the original cross section site, the number of differences between the cross sections increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
-}
-\end{figure}
+
 %\afterpage{\clearpage}
@@
   \centering
-<<tree:forest, echo=FALSE, fig.width=6, fig.height=6.25, out.width='.5\\textwidth', warning = FALSE>>=
+<<tree:forest, echo=FALSE, fig.width=5, fig.height=5.25, out.width='.5\\textwidth', warning = FALSE>>=
 qplot(pred, forest, geom="jitter", data=bstats, colour=match, 

### 2015-12-23 imaging-supplement.Rnw
more into supplement

@@
 
+\subsection{Assessing cross-correlation between signatures at multiple cross sections}
+
+Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$.
+These are compared to a cross section at 100$\mu m$.
+Initially this comparison constitutes an almost perfect match between the two cross sections.
+However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.
+Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
+Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections.
+\begin{figure}[hbtp]
+  \centering
+<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
+paths <- file.path(datadir, dir(datadir))
+crosscuts <- seq(100, 400, by = 50)
+lof <- processBullets(paths[5], x = crosscuts)
+lof$bullet <- paste(lof$bullet, lof$x)
+
+reslist <- lapply(crosscuts[-1], function(cc) {
+#  browser()
+  b2 <- subset(lof, x %in% c(cc, 100))
+  lofX <- bulletSmooth(b2)
+  bAlign = bulletAlign(lofX)
+  lofX <- bAlign$bullet
+    b12 <- unique(b2$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
+
+#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+#  lines <- striation_identify(lofX, threshold = threshold)
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
+  maxCMS <- maxCMS(lines$match==TRUE)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
+})
+
+ccfs <- sapply(reslist, function(res) res$ccf)
+
+lop <- lapply(reslist, function(res) {
+ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
+  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
+  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
+  scale_linetype_discrete("") +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylim(c(-6,6)) +
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
+    ylab("") + xlab("")
+})
+
+grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
+             ncol = 2)
+@
+\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections.
+Cross section at $x = 100$ is compared to cross sections every 50$\mu m$.
+With every step away from the original cross section site, the number of differences between the cross sections increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
+}
+\end{figure}
 \end{document}

### 2015-12-23 imaging-paper.Rnw
up to stability.

@@
 The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
-This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can extract signatures from multiple cross sections of the same bullet for an initial assessment of its quality.
-\hh{XXX shift cross section discussion into the supplement XXX}
-Both of these situations are assessed and discussed in more detail below:
-
-%\afterpage{\clearpage}
-
-%\hh{XXX Does this give us a way to automatically check for the bottom of the bullet? Attempt below}
+This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: we have to be aware of which cross section level was used in a matching.
+An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement \hh{XXX section XXX} for a more thorough discussion).
+However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. \hh{XXX shift cross section discussion into the supplement XXX}
+This situation is assessed and discussed in more detail below:
 
@@
 \caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet 1.
-Initially, the match between cross sections 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet.
+Initially, the match between cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
 At a cross section level of $175\mu m$ the bullet pattern stabilizes.
@@
 \end{figure}
-An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the cross section is in a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or in a stable section, where we can expect to find matches to other surfaces reasonably well.
-In the approach here, we keep increasing the cross section level until we find a section with a stable pattern.
-This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3.
-`Stability' is here defined as two aligned cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
+%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
+by comparing signatures from cross sections that are not too far apart --  25 $\mu m$ to 50 $\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or from a stable section, where we can have a reasonable expectation of finding matches to other signatures.
+In the approach here, we keep increasing the cross section until we find a section with a stable pattern.
+This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3, where
+`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
 

### 2015-12-23 imaging-paper.Rnw
done for now

@@
 
-\subsection{Horizontal Alignment}
+\subsection{Horizontal alignment}
 A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.
@@
 
-\subsection{Varying cross sections}
+\subsection{Impact of cross sections}
 The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
@@
 
+\hh{XXX Why do we use the lowest possible cross section? XXX}
+
+\subsection{Varying smoothing factor}\label{sec:smoothing}
+As mentioned earlier, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
+In particular, for detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
+We can investigate the performance of peak/valley detection across different smoothing factors.
+Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 25, and 45, respectively.
+The dark line corresponds to the smoothed values, while the grey line in the back shows the unsmoothed values of the surface measurements.
+The choice of smoothing factor is a classical decision of a bias/variance trade-off.
+It is immediately clear that a small smoothing factor like 5 is a poor choice.
+It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
+Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
+
+A larger smoothing factor on the other hand, like 45, seems to be a more plausible option.
+Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 45.
+However, some notable issues arise.
+Notice that the valley on the right hand side of the image is smoothed out, and thus not detected.
+On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$.
+That is, in many cases, large windows are smoothing out some of the structure that we wish to see.
+Furthermore, it can be seen that the peaks/valleys are often shifted relative to their position in the original loess residuals, or in the smoothed data with smaller smoothing factors.
+
+\begin{figure}[hbtp]
+  \centering
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
+br111.groove <- get_grooves(br111_bullet)
+br111.loess <- fit_loess(br111_bullet, br111.groove)
+
+peaks1 <- get_peaks(br111.loess$data, smoothfactor = 25)
+peaks1$lines$smoothfactor = 25
+peaks1$plot$data$smoothfactor = 25
+peaks2 <- get_peaks(br111.loess$data, smoothfactor = 5)
+peaks2$lines$smoothfactor = 5
+peaks2$plot$data$smoothfactor = 5
+peaks3 <- get_peaks(br111.loess$data, smoothfactor = 45)
+peaks3$lines$smoothfactor = 45
+peaks3$plot$data$smoothfactor = 45
+
+peaks <- rbind(peaks1$lines, peaks2$lines, peaks3$lines)
+smooths <- rbind(peaks1$plot$data, peaks2$plot$data, 
+                 peaks3$plot$data)
+
+  ggplot() + theme_bw() +
+    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
+              data=peaks,  alpha=0.2) +
+    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
+    scale_colour_brewer(palette="Set2") + 
+    scale_fill_brewer(palette="Set2") +
+    theme(legend.position="none") + 
+    facet_grid(smoothfactor~., labeller="label_both") +
+    geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
+              colour="grey50") +
+    geom_line(aes(x=y, y=smoothed), data=smooths) +
+  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
+@
+\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively.
+Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either disappear or shift  horizontally  from their position in the original loess residuals.}
+\end{figure}
+
+\subsection{Algorithm}
+
 \begin{figure}[hbtp]
@@
 %\afterpage{\clearpage}
-
-\subsection{Varying smoothing factor}\label{sec:smoothing}
-As mentioned earlier, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
-We can investigate the performance of peak/valley detection across different smoothing factors.
-Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 25, and 45, respectively.
-The dark line corresponds to the smoothed values, while the grey line in the back shows the unsmoothed values of the surface measurements.
-The choice of smoothing factor is a classical decision of a bias/variance trade-off.
-It is immediately clear that a small smoothing factor like 5 is a poor choice.
-It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
-Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
-
-A larger smoothing factor on the other hand, like 45, seems to be a more plausible option.
-Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 45.
-However, some notable issues arise.
-Notice that the valley on the right hand side of the image is smoothed out, and thus not detected.
-On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$.
-That is, in many cases, large windows are smoothing out some of the structure that we wish to see.
-Furthermore, it can be seen that the peaks/valleys are often shifted relative to their position in the original loess residuals, or in the smoothed data with smaller smoothing factors.
-
-\begin{figure}[hbtp]
-  \centering
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
-br111.groove <- get_grooves(br111_bullet)
-br111.loess <- fit_loess(br111_bullet, br111.groove)
-
-peaks1 <- get_peaks(br111.loess$data, smoothfactor = 25)
-peaks1$lines$smoothfactor = 25
-peaks1$plot$data$smoothfactor = 25
-peaks2 <- get_peaks(br111.loess$data, smoothfactor = 5)
-peaks2$lines$smoothfactor = 5
-peaks2$plot$data$smoothfactor = 5
-peaks3 <- get_peaks(br111.loess$data, smoothfactor = 45)
-peaks3$lines$smoothfactor = 45
-peaks3$plot$data$smoothfactor = 45
-
-peaks <- rbind(peaks1$lines, peaks2$lines, peaks3$lines)
-smooths <- rbind(peaks1$plot$data, peaks2$plot$data, 
-                 peaks3$plot$data)
-
-  ggplot() + theme_bw() +
-    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
-              data=peaks,  alpha=0.2) +
-    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
-    scale_colour_brewer(palette="Set2") + 
-    scale_fill_brewer(palette="Set2") +
-    theme(legend.position="none") + 
-    facet_grid(smoothfactor~., labeller="label_both") +
-    geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
-              colour="grey50") +
-    geom_line(aes(x=y, y=smoothed), data=smooths) +
-  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
-@
-\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively.
-Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either disappear or shift  horizontally  from their position in the original loess residuals.}
-\end{figure}
-
-\subsection{Algorithm}
-
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 

### 2015-12-23 imaging-paper.Rnw
major shuffle

@@
 
-% \section{Web Application}
-% We have implemented a web interface to automate many of the steps outlined in the previous section.
-The web interface, developed using Shiny~\citep{shiny}, allows a user to upload two x3p files representing scans of a bullet between two grooves.
-A 3D viewer, using Plotly~\citep{plotly}, allows the user to rotate and zoom to examine surface features of the images.
-A screenshot of the application is available in Figure \ref{fig:app}.
-The application is available at the url \url{http://erichare.shinyapps.io/x3prproto}.
-% 
-% \begin{figure*}
-% \includegraphics[width=\textwidth]{images/app.png}
-% \caption{\label{fig:app}{Screenshot of the web interface for analyzing bullet images.}}
-% \end{figure*}
-% 
-% The mouse is the primary control input for interacting with the surface plots.
-One can mouse over the plots to view the coordinate values for each point.
-Clicking and dragging anywhere on the plot will allow rotation.
-Scrolling up and down allows the user to zoom out and and respectively.
-There are also a set of controls available at the top right of the page, which will allow the selection of some other rotation methods, taking of screenshots, and reseting the camera to the default view.
-% 
-% Several lighting options are available to tweak in order to highlight markings in on the bullets more strongly:
-% 
-% \begin{enumerate}
-%     \item \textbf{Ambient Lighting} - Lighting which is used to simulate the light naturally present in the environment.
-%     \item \textbf{Diffuse Lighting} - Setting this value higher will cause the surface to re-emit a higher proportion of light.
-%     \item \textbf{Specular Lighting} - The lighting which appears on the surface when illuminated.
-%     \item \textbf{Roughness Lighting} - The amount of light reflected by the roughness of the surface.
-%     \item \textbf{Fresnel Lighting} - Higher values create a wider, softer lighting across the surface.
-% \end{enumerate}
-% 
-% The application includes a button which allows the automatic creation of a residual plot from the loess fits of two bullets.
-An x coordinate value can be selected as a cross section, and cross sections of the two bullets at the x value are taken.
-The previously described loess regression procedure is applied to the two cross sections.
-A residual plot is produced beneath the surface plots allowing for an assessment of the similarity of the markings.
-Users need to perform this procedure at different x values in order to see whether the markings indicate a match at different points along the bullet.
-% 
-% Enhancements to the web application are still in progress. In particular, much of the automatic matching functionality described in the next section will be added in the future.
-
 \section{Automatic matching}
@@
 
-% \begin{figure}[hbtp]
-% \centering
-% %' <<side-by-side, echo=FALSE, eval=FALSE>>=
-% %' paths <- file.path(datadir, dir(datadir))
-% %' 
-% %' br111 <- read.x3p(paths[5])
-% %' crosscuts <- unique(fortify_x3p(br111)$x)
-% %' crosscuts <- crosscuts[crosscuts > 50]
-% %' crosscuts <- crosscuts[crosscuts < 150]
-% %' 
-% %' list_of_fits <- lapply(crosscuts, function(x) {
-% %'   br111 <- get_crosscut(paths[5], x = x)
-% %'   br111.groove <- get_grooves(br111)
-% %'   br111.groove$plot
-% %'   fit_loess(br111, br111.groove)
-% %' })
-% %' 
-% %' lof <- lapply(list_of_fits, function(x) x$resid$data) %>% bind_rows
-% %' 
-% %' br111 <- read.x3p(paths[7])
-% %' crosscuts <- unique(fortify_x3p(br111)$x)
-% %' crosscuts <- crosscuts[crosscuts > 50]
-% %' crosscuts <- crosscuts[crosscuts < 150]
-% %' 
-% %' list_of_fits2 <- lapply(crosscuts, function(x) {
-% %'   br111 <- get_crosscut(paths[7], x)
-% %'   br111.groove <- get_grooves(br111)
-% %'   br111.groove$plot
-% %'   fit_loess(br111, br111.groove)
-% %' })
-% %' 
-% %' lof2 <- lapply(list_of_fits2, function(x) x$resid$data) %>% bind_rows
-% %' lof$bullet <- 1
-% %' lof2$bullet <- 2
-% %' 
-% %' LOF <- rbind(lof, lof2)
-% %' 
-% %' subLOF1 <- LOF %>% filter(bullet == 1, x <= 100)
-% %' subLOF2 <- LOF %>% filter(bullet == 2, x > 100)
-% %' 
-% %' #subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
-% %' subLOF1$y <- subLOF1$y + 19*1.5625 # working now!!!
-% %' subLOF <- rbind(data.frame(subLOF1), data.frame(subLOF2))
-% %' 
-% %' qplot(x, y, fill = resid, colour=I(NA), data=subLOF, geom="tile") + 
-% %'   scale_fill_gradientn(limits=c(-5,5), 
-% %'                        colors=c("grey5", "gold4","darkgoldenrod1","lightgoldenrod1", "lemonchiffon"),
-% %'                        values = c(0,0.5, 0.75, 0.9, 1)) +
-% %'   theme_bw()
-% %' @
-% 
-% \includegraphics[width=.7\columnwidth]{images/side-by-side-manual-match.png}
-% \caption{\label{fig:manualmatch}Manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1.}
-% \end{figure}
-
-\begin{figure}[hbtp]
-\centering
-\includegraphics[width=.7\columnwidth]{images/matchup-rgl.png}
+\begin{figure}[hbtp]
+\centering
+\includegraphics[width=.65\columnwidth]{images/matchup-rgl.png}
 \caption{\label{fig:manualmatch-rgl}3d view of the manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1 after removing the curvature.}
@@
 
-<<twolands100, cache=TRUE,  echo=FALSE>>=
-images <- file.path(datadir, dir(datadir))
-
-lof <- processBullets(paths = images[c(5,7)], x = 100)
-
-subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
-lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
-
-# smooth
-lofX <- lofX %>% group_by(bullet) %>% mutate(
-  l30 = smoothloess(y, resid, span = 0.03)
-)
-
-# cut at .75
-threshold <- .75
-lofX$r05 <- threshold* sign(lofX$l30) * as.numeric(abs(lofX$l30) > threshold)
-lofX$type <- factor(lofX$r05)
-levels(lofX$type) <- c("groove", NA, "peak")
-@
-
-<<twolands100adjust, echo=FALSE>>=
-images <- file.path(datadir, dir(datadir))
-
-lof <- processBullets(paths = images[c(5,7)], x = 100)
-
-subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
-subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-ccf <- ccf(subLOFx1$resid, subLOFx2$resid, lag.max=100, plot=FALSE)
-lag <- ccf$lag[which.max(ccf$acf)]
-subLOFx1$y <- subLOFx1$y - lag*1.5625 
-lofY <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
-
-@
-
-<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
-matches <- lofX %>% group_by(y) %>% summarise(
-  potential = (length(unique(type)) == 1),
-  allnas = sum(is.na(type))/n(),
-  type1 = na.omit(type)[1],
-  type = paste(type, sep="|", collapse="|"),
-  n = n()
-)
-
-matches$id <- cumsum(matches$allnas == 1) + 1
-matches$lineid <- as.numeric(matches$allnas != 1) * matches$id
-
-isMatch <- function(id, type) {
-  if (id[1] == 0) return(FALSE)
-#  browser()
-  types <- strsplit(type, split = "|", fixed=TRUE) 
-  t1 <- sapply(types, function(x) x[1])
-  t2 <- sapply(types, function(x) x[2])
-  if (all(t1 == "NA")) return(FALSE)
-  if (all(is.na(t2))) return(FALSE)
-  if (all(t2 == "NA")) return(FALSE)
-
-  peak <- length(grep("peak", c(t1, t2))) > 0
-  groove <- length(grep("groove", c(t1, t2))) > 0
-  if (peak & groove) return(FALSE)
-
-  return(TRUE)
-}
-
-lines <- matches %>% group_by(lineid) %>% summarise(
-  meany = mean(y, na.rm=T),
-  miny = min(y, na.rm=T),
-  maxy = max(y, na.rm=T) + 1.5625,
-  match = isMatch(lineid, type),
-  type = type1[1]
-)
-lines <- subset(lines, lineid != 0)
-@
-
-\subsection{Horizontal alignment}
-A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.
-This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$.
-Striation marks show up in this representations as peaks and valleys.
-The individual characteristics are quite prominent and, again, suggest a match between the lands.
-A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities.
-%
-\begin{figure}[hbtp]
-\centering
-\begin{subfigure}[b]{.49\textwidth}\centering
-\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
-<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
-qplot(y, resid, group=bullet, colour=bullet, data=lof, 
-      geom="line") + 
-  scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position = c(.9, .2),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab("Residuals of loess fit")
-@
-    }
-\end{subfigure}    
-\begin{subfigure}[b]{.49\textwidth}\centering
-    \caption{Aligned (by horizontal shift) signatures.\label{fig:crosscutX}}{%
-<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
-qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
-      geom="line") + xlab("y") +
-  scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position = c(.9, .2),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab("Residuals of loess fit")
-@
-    }
-\end{subfigure}
-\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$.
-A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
-\end{figure}
-%
-This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0 $\mu m$ and about 2500 $\mu m$, 1.5625 $\mu m$ apart.
-The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
-\[
-(f * g) (k) = \sum_t f(t+k) g(t),
-\]
-with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).
-
-\begin{figure}[hbtp]
-  \centering
-<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '.5\\textwidth'>>=
-lag <- ccf$lag[which.max(ccf$acf)]
-qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
-  theme_bw() + ylab("Correlation") + xlab("Lag k") +
-  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
-  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
-                     minor_breaks = c(-75,-25,25, 75))
-@
-\caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
-At a lag of $\Sexpr{lag}$ the correlation peaks, indicating the largest amount of agreement between the signatures.
-Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}
-\end{figure}
-
-\subsection{Impact of cross sections}
-The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
-This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: we have to be aware of which cross section level was used in a matching.
-An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement \hh{XXX section XXX} for a more thorough discussion).
-However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. \hh{XXX shift cross section discussion into the supplement XXX}
-This situation is assessed and discussed in more detail below:
-
-\begin{figure}[hbtp]
-  \centering
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
-paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100,250, by=25)
-lof <- processBullets(paths[37], x = crosscuts)
-lof$bullet <- paste(lof$bullet, lof$x)
-
-reslist <- lapply(1:length(crosscuts[-1]), function(i) {
-  b2 <- subset(lof, x %in% crosscuts[i:(i+1)])
-  lofX <- bulletSmooth(b2)
-  bAlign <- bulletAlign(lofX)
-  lofX <- bAlign$bullet
-
-  b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
-
-#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-#  lines <- striation_identify(lofX, threshold = threshold)
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
-})
-
-ccfs <- sapply(reslist, function(res) res$ccf)
-
-lop <- lapply(reslist, function(res) {
-ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
-  scale_linetype_discrete("") +
-#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6)) + ylab("") + xlab("") + 
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
-})
-
-grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
-             lop[[4]], lop[[5]], lop[[6]],
-             ncol=2)
-@
-\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet 1.
-Initially, the match between cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
-At a cross section level of $175\mu m$ the bullet pattern stabilizes.
-For this land, matches should not be attempted at lower cross sections.}
-\end{figure}
-%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
-by comparing signatures from cross sections that are not too far apart --  25 $\mu m$ to 50 $\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or from a stable section, where we can have a reasonable expectation of finding matches to other signatures.
-In the approach here, we keep increasing the cross section until we find a section with a stable pattern.
-This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3, where
-`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
-
-\hh{XXX Why do we use the lowest possible cross section? XXX}
-
-\subsection{Varying smoothing factor}\label{sec:smoothing}
-As mentioned earlier, the algorithm for peak/valley detection depends on the selection of a smoothing window, called the smoothing factor.
-In particular, for detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
-We can investigate the performance of peak/valley detection across different smoothing factors.
-Figure~\ref{fig:varysmooth} displays the resulting peaks and valleys across smoothing factors of 5, 25, and 45, respectively.
-The dark line corresponds to the smoothed values, while the grey line in the back shows the unsmoothed values of the surface measurements.
-The choice of smoothing factor is a classical decision of a bias/variance trade-off.
-It is immediately clear that a small smoothing factor like 5 is a poor choice.
-It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
-Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
-
-A larger smoothing factor on the other hand, like 45, seems to be a more plausible option.
-Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 45.
-However, some notable issues arise.
-Notice that the valley on the right hand side of the image is smoothed out, and thus not detected.
-On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$.
-That is, in many cases, large windows are smoothing out some of the structure that we wish to see.
-Furthermore, it can be seen that the peaks/valleys are often shifted relative to their position in the original loess residuals, or in the smoothed data with smaller smoothing factors.
-
-\begin{figure}[hbtp]
-  \centering
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
-br111.groove <- get_grooves(br111_bullet)
-br111.loess <- fit_loess(br111_bullet, br111.groove)
-
-peaks1 <- get_peaks(br111.loess$data, smoothfactor = 25)
-peaks1$lines$smoothfactor = 25
-peaks1$plot$data$smoothfactor = 25
-peaks2 <- get_peaks(br111.loess$data, smoothfactor = 5)
-peaks2$lines$smoothfactor = 5
-peaks2$plot$data$smoothfactor = 5
-peaks3 <- get_peaks(br111.loess$data, smoothfactor = 45)
-peaks3$lines$smoothfactor = 45
-peaks3$plot$data$smoothfactor = 45
-
-peaks <- rbind(peaks1$lines, peaks2$lines, peaks3$lines)
-smooths <- rbind(peaks1$plot$data, peaks2$plot$data, 
-                 peaks3$plot$data)
-
-  ggplot() + theme_bw() +
-    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
-              data=peaks,  alpha=0.2) +
-    geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
-    scale_colour_brewer(palette="Set2") + 
-    scale_fill_brewer(palette="Set2") +
-    theme(legend.position="none") + 
-    facet_grid(smoothfactor~., labeller="label_both") +
-    geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
-              colour="grey50") +
-    geom_line(aes(x=y, y=smoothed), data=smooths) +
-  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
-@
-\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively.
-Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either disappear or shift  horizontally  from their position in the original loess residuals.}
-\end{figure}
-
+\hh{In this section, we are going to describe the algorithm for matching signatures first, and discuss the impact of paramater choices in the subsections thereafter.} 
 \subsection{Algorithm}
@@
 }
+\end{figure}
+
+<<twolands100, cache=TRUE,  echo=FALSE>>=
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
+lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+# smooth
+lofX <- lofX %>% group_by(bullet) %>% mutate(
+  l30 = smoothloess(y, resid, span = 0.03)
+)
+
+# cut at .75
+threshold <- .75
+lofX$r05 <- threshold* sign(lofX$l30) * as.numeric(abs(lofX$l30) > threshold)
+lofX$type <- factor(lofX$r05)
+levels(lofX$type) <- c("groove", NA, "peak")
+@
+
+<<twolands100adjust, echo=FALSE>>=
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
+subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
+ccf <- ccf(subLOFx1$resid, subLOFx2$resid, lag.max=100, plot=FALSE)
+lag <- ccf$lag[which.max(ccf$acf)]
+subLOFx1$y <- subLOFx1$y - lag*1.5625 
+lofY <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+@
+
+<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
+matches <- lofX %>% group_by(y) %>% summarise(
+  potential = (length(unique(type)) == 1),
+  allnas = sum(is.na(type))/n(),
+  type1 = na.omit(type)[1],
+  type = paste(type, sep="|", collapse="|"),
+  n = n()
+)
+
+matches$id <- cumsum(matches$allnas == 1) + 1
+matches$lineid <- as.numeric(matches$allnas != 1) * matches$id
+
+isMatch <- function(id, type) {
+  if (id[1] == 0) return(FALSE)
+#  browser()
+  types <- strsplit(type, split = "|", fixed=TRUE) 
+  t1 <- sapply(types, function(x) x[1])
+  t2 <- sapply(types, function(x) x[2])
+  if (all(t1 == "NA")) return(FALSE)
+  if (all(is.na(t2))) return(FALSE)
+  if (all(t2 == "NA")) return(FALSE)
+
+  peak <- length(grep("peak", c(t1, t2))) > 0
+  groove <- length(grep("groove", c(t1, t2))) > 0
+  if (peak & groove) return(FALSE)
+
+  return(TRUE)
+}
+
+lines <- matches %>% group_by(lineid) %>% summarise(
+  meany = mean(y, na.rm=T),
+  miny = min(y, na.rm=T),
+  maxy = max(y, na.rm=T) + 1.5625,
+  match = isMatch(lineid, type),
+  type = type1[1]
+)
+lines <- subset(lines, lineid != 0)
+@
+
+\subsection{Horizontal alignment}
+A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.
+This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$.
+Striation marks show up in this representations as peaks and valleys.
+The individual characteristics are quite prominent and, again, suggest a match between the lands.
+A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities.
+%
+\begin{figure}[hbtp]
+\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
+\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
+<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+qplot(y, resid, group=bullet, colour=bullet, data=lof, 
+      geom="line") + 
+  scale_colour_manual("", values=cols) + theme_bw() +
+  theme(legend.position = c(.9, .2),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylab("Residuals of loess fit")
+@
+    }
+\end{subfigure}    
+\begin{subfigure}[b]{.49\textwidth}\centering
+    \caption{Aligned signatures.\label{fig:crosscutX}}{%
+<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
+qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
+      geom="line") + xlab("y") +
+  scale_colour_manual("", values=cols) + theme_bw() +
+  theme(legend.position = c(.9, .2),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylab("Residuals of loess fit")
+@
+    }
+\end{subfigure}
+\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$.
+A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
+\end{figure}
+%
+This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0 $\mu m$ and about 2500 $\mu m$, 1.5625 $\mu m$ apart.
+The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
+\[
+(f * g) (k) = \sum_t f(t+k) g(t),
+\]
+with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).
+
+\begin{figure}[hbtp]
+  \centering
+<<ccf, echo=FALSE, fig.width = 8, fig.height = 3, out.width = '.65\\textwidth'>>=
+lag <- ccf$lag[which.max(ccf$acf)]
+qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
+  theme_bw() + ylab("Correlation") + xlab("Lag k") +
+  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
+  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
+                     minor_breaks = c(-75,-25,25, 75))
+@
+\caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
+At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures.
+Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}
+\end{figure}
+
+\subsection{Impact of cross sections}
+The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
+This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: we have to be aware of which cross section level was used in a matching.
+An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement \hh{XXX section XXX} for a more thorough discussion).
+However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. 
+This situation is assessed and discussed in more detail below:
+
+\begin{figure}[hbtp]
+  \centering
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
+paths <- file.path(datadir, dir(datadir))
+crosscuts <- seq(100,250, by=25)
+lof <- processBullets(paths[37], x = crosscuts)
+lof$bullet <- paste(lof$bullet, lof$x)
+
+reslist <- lapply(1:length(crosscuts[-1]), function(i) {
+  b2 <- subset(lof, x %in% crosscuts[i:(i+1)])
+  lofX <- bulletSmooth(b2)
+  bAlign <- bulletAlign(lofX)
+  lofX <- bAlign$bullet
+
+  b12 <- unique(b2$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
+
+#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
+#  lines <- striation_identify(lofX, threshold = threshold)
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
+  maxCMS <- maxCMS(lines$match==TRUE)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
+})
+
+ccfs <- sapply(reslist, function(res) res$ccf)
+
+lop <- lapply(reslist, function(res) {
+ggplot() +
+  theme_bw() + 
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
+  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
+  scale_linetype_discrete("") +
+#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
+  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+        legend.background = element_rect(fill=alpha('white', 0.4))) + 
+  ylim(c(-6,6)) + ylab("") + xlab("") + 
+  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
+})
+
+grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
+             lop[[4]], lop[[5]], lop[[6]],
+             ncol=2)
+@
+\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet 1-1.
+Initially, the match between cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
+At a cross section level of $175\mu m$ the bullet pattern stabilizes.
+For this land, matches should not be attempted at lower cross sections.
+}
+\end{figure}
+%An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
+by comparing signatures from cross sections that are not too far apart --  25 $\mu m$ to 50 $\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or from a stable section, where we can have a reasonable expectation of finding matches to other signatures.
+In the approach here, we keep increasing the cross section until we find a section with a stable pattern.
+This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3, where
+`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
+
+\hh{XXX Why do we use the lowest possible cross section? XXX}
+
+\subsection{Varying smoothing factor}\label{sec:smoothing}
+As mentioned earlier, the algorithm for detecting peaks and valleys depend on the selection of a smoothing window, called the smoothing factor.
+For detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
+%We can investigate the performance of peak/valley detection across different smoothing factors.
+Figure~\ref{fig:varysmooth} displays the  peaks and valleys detected in the same signature at smoothing factors of 5, 25, and 45, respectively.
+The dark line corresponds to the smoothed values, while the grey line in the back shows the raw signature.
+The choice of smoothing factor is a classical decision of a bias/variance trade-off.
+It is immediately clear that a small smoothing factor like 5 is a poor choice.
+It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
+Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
+\hh{XXX shouldn't we also make this dependent on the resolution of the equipment?
+XXX}
+
+A larger smoothing factor on the other hand, like 45, seems to be a more plausible option.
+Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 45.
+However, some notable issues arise.
+Notice that the valley on the right hand side of the image is smoothed out, and thus not detected.
+On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$.
+That is, in many cases, large windows are smoothing out some of the structure that we wish to see.
+Furthermore, it can be seen that the peaks/valleys are often shifted relative to their position in the original loess residuals, or in the smoothed data with smaller smoothing factors.
+
+\begin{figure}[hbtp]
+  \centering
+<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
+br111.groove <- get_grooves(br111_bullet)
+br111.loess <- fit_loess(br111_bullet, br111.groove)
+
+peakslist <- lapply(c(5, 25, 45), function(s) {
+  peaks <- get_peaks(br111.loess$data, smoothfactor = s)
+  peaks$lines$smoothfactor = s
+  peaks$plot$data$smoothfactor = s
+
+  peaks
+})
+
+peaks <- plyr::ldply(peakslist, function(x) x$lines)
+smooths <- plyr::ldply(peakslist, function(x) x$plot$data)
+
+ggplot() + theme_bw() +
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
+            data=peaks,  alpha=0.2) +
+  geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
+  scale_colour_brewer(palette="Set2") + 
+  scale_fill_brewer(palette="Set2") +
+  theme(legend.position="none") + 
+  facet_grid(smoothfactor~., labeller="label_both") +
+  geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
+            colour="grey50") +
+  geom_line(aes(x=y, y=smoothed), data=smooths) +
+  ylab(expression(paste("Signature values (in ",mu,"m)", sep=""))) 
+@
+\caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively.
+Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either end disappear or shift  horizontally  from their position in the signature.}
 \end{figure}

### 2015-12-23 imaging-paper.Rnw
major re-work: extend algorithm to include feature extraction & describe each

@@
 \startlocaldefs
+\usepackage{enumitem}
 \usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
@@
 bstats <- read.csv("data-25-25/bullet-stats.csv")
+names(bstats)[9:10] <- c("#matches", "#non-matches")
 
@@
 @
-\caption{\label{fig:fixedX}Side profile of the surface measurements (in $\mu m$) of a bullet land across a fixed $x$ value.
+\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land across a fixed $x$ value.
 Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
@@
 
-\subsection{Identifying grooves}
+\subsection{Identifying groove locations}
 %In an attempt to improve upon the previous algorithm, w
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm.
-For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
+\caption{\label{fig:loess_step1}Step 1 of the loess fit algorithm.
+For a fixed cross section ($x$ held constant at 243.75~$\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
@@
 \caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm.
-The  height values are smoothed twice with a smoothing factor of 35.
+The  surface measurements are smoothed twice with a smoothing factor of $s = 35$.
 The orange rectangle shows an example of the smoothing window.
-Grooves and peaks are detected, if they are not within the same window.}
+Valleys and peaks are detected, if they are not within the same window.}
 <<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
@@
 \caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm.
-After smoothing the $y$ values left and right minima are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
-Values outside these boundaries are removed (shown in grey)}
+After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
+Values outside the blue boundaries are removed (shown in grey)}
 <<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
@@
 \hh{XXX Introduce Loess formula with all of its details in here. we are using the default of {\tt loess}, which has a span of 0.75 and uses a polynomial of degree 2. XXX}
-Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of bullet 1-5.
+Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of bullet~1-5.
 The fit seems to do a reasonable job of capturing the structure of the image.
 Figure~\ref{fig:loess_resid} shows the residuals from this fit.
-These residuals are called the \emph{signature of bullet 1-5.}
+These residuals are called the \emph{signature} of bullet~1-5.
 %
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{\label{fig:loess_fit} Loess fit for bullet 1-5.}
+\caption{\label{fig:loess_fit} Loess fit for bullet~1-5.}
 <<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{\label{fig:loess_resid} Residuals of loess fit for bullet 1-5.}
+\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5.}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
@@
 \end{subfigure}
-\caption{Fit and residuals of a loess fit to bullet 1-5 (Barrel 1).}
+\caption{Fit and residuals of a loess fit to bullet~1-5 (Barrel~1).}
 \end{figure}
 %
-Figure~\ref{fig:loessfits} shows the signatures of all lands of bullet 2 (in grey) fired from the same barrel as bullet 1.
-The darker line overlaid on each of the signatures, shows signature from bullet 1-5.
-%We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet 1-5.
+Figure~\ref{fig:loessfits} shows the signatures of all lands of bullet 2 (in grey) fired from the same barrel as bullet~1.
+The darker line overlaid on each of the signatures, shows signature from bullet~1-5.
+%We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet~1-5.
 This is shown in Figure~\ref{fig:loessfits}.
-Bullet 2-1 is the best match for bullet 1-5, but the alignment is not quite  correct.
+Bullet 2-1 is the best match for bullet~1-5, but the alignment is not quite  correct.
 
@@
   theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Signatures of Bullet 1 in grey, Bullet 1-5 in black") + ylab("Residuals of loess fit")
-@    
-\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet 1-5 is shown in black. }
+  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Signatures of bullet 1 in grey, bullet 1-5 in black") + ylab("Residuals of loess fit")
+@    
+\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet~1-5 is shown in black. }
 \end{figure}
@@
 \section{Automatic matching}
-Applying the loess fit to a range of different surface cross sections (see Figure~\ref{fig:manualmatch-rgl} for cross sections from 50$\mu m$ to 150 $\mu m$) shows the 3d striation marks from two bullets.
-Signatures of bullet 1 are shown on the left (all cross sections with values below 100$\mu m$) and cross sections of bullets 2 are shown on the right (cross sections with values above 100$\mu m$).
-\hh{Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one bullet to the other.
+Applying the loess fit to a range of different surface cross sections (see Figure~\ref{fig:manualmatch-rgl} for cross sections from 50$\mu m$ to 150~$\mu m$) shows the 3d striation marks from two bullets.
+Signatures of bullet~1 are shown on the left (all cross sections with values below 100$\mu m$) and cross sections of bullets 2 are shown on the right (cross sections with values above 100$\mu m$).
+\hh{Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one side to the other.
 Visually, this allows for an easy assessment of these two bullet lands as a match.
@@
 \includegraphics[width=.65\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}3d view of the manually adjusted side-by-side comparison of bullet 1-5 and bullet 2-1 after removing the curvature.}
+\caption{\label{fig:manualmatch-rgl}3d view of the manually adjusted side-by-side comparison of bullet~1-5 and bullet 2-1 after removing the curvature.}
 \end{figure}
@@
 \hh{In this section, we are going to describe the algorithm for matching signatures first, and discuss the impact of paramater choices in the subsections thereafter.}
+
 \subsection{Algorithm}
+<<twolands100, cache=TRUE,  echo=FALSE>>=
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
+lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+# smooth
+lofX <- lofX %>% group_by(bullet) %>% mutate(
+  l30 = smoothloess(y, resid, span = 0.03)
+)
+
+# cut at .75
+threshold <- .75
+lofX$r05 <- threshold* sign(lofX$l30) * as.numeric(abs(lofX$l30) > threshold)
+lofX$type <- factor(lofX$r05)
+levels(lofX$type) <- c("groove", NA, "peak")
+@
+
+<<twolands100adjust, echo=FALSE>>=
+images <- file.path(datadir, dir(datadir))
+
+lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
+subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
+ccf <- ccf(subLOFx1$resid, subLOFx2$resid, lag.max=100, plot=FALSE)
+lag <- ccf$lag[which.max(ccf$acf)]
+subLOFx1$y <- subLOFx1$y - lag*1.5625 
+lofY <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
+
+@
+
+<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
+matches <- lofX %>% group_by(y) %>% summarise(
+  potential = (length(unique(type)) == 1),
+  allnas = sum(is.na(type))/n(),
+  type1 = na.omit(type)[1],
+  type = paste(type, sep="|", collapse="|"),
+  n = n()
+)
+
+matches$id <- cumsum(matches$allnas == 1) + 1
+matches$lineid <- as.numeric(matches$allnas != 1) * matches$id
+
+isMatch <- function(id, type) {
+  if (id[1] == 0) return(FALSE)
+#  browser()
+  types <- strsplit(type, split = "|", fixed=TRUE) 
+  t1 <- sapply(types, function(x) x[1])
+  t2 <- sapply(types, function(x) x[2])
+  if (all(t1 == "NA")) return(FALSE)
+  if (all(is.na(t2))) return(FALSE)
+  if (all(t2 == "NA")) return(FALSE)
+
+  peak <- length(grep("peak", c(t1, t2))) > 0
+  groove <- length(grep("groove", c(t1, t2))) > 0
+  if (peak & groove) return(FALSE)
+
+  return(TRUE)
+}
+
+lines <- matches %>% group_by(lineid) %>% summarise(
+  meany = mean(y, na.rm=T),
+  miny = min(y, na.rm=T),
+  maxy = max(y, na.rm=T) + 1.5625,
+  match = isMatch(lineid, type),
+  type = type1[1]
+)
+lines <- subset(lines, lineid != 0)
+@
 
@@
     \begin{subfigure}[t]{\textwidth}\centering
-\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
-<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
+\caption{Loess smooth of signatures  at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
+<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
 qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
-<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE>>=
+\caption{Using a rolling median peaks and valleys are identified for each signature. \label{fig:smoothcutb}}{%
+<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
 #images <- file.path(datadir, dir(datadir))
@@
     \caption{Rectangles in the back identify a line on one of the bullets.
-Matching striation marks are indicated by color filled rectangles and arked by an `o'.
+Matching striation marks are indicated by color filled rectangles and marked by an `o'.
 Mismatches are filled in grey and  marked by an `x'.
 \label{fig:smoothcutd}}{%
-<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.5\\textwidth', warning=FALSE>>=
+<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.65\\textwidth', warning=FALSE>>=
 peaks1$lines$bullet <- b12[1]
@@
 \end{subfigure}
-\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between cross sections (c).}
+\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between signatures (c).}
 \end{figure}
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet.
-%\hh{(assumption here: lower cross section levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
+We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet.
+\hh{(assumption here: lower cross section levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
 
@@
 \begin{enumerate}
-\item {\bf Smooth the two cross sections} by a loess with a very small span (see Figure~\ref{fig:smooth}). 
-\item Use cross-correlation to {\bf find the best alignment} along the cross sections: Shift one of the cross sections by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
-\item Using a rolling average, {\bf identify peaks and grooves} for each of the cross sections.
-We define a region around the extrema on each side as one third of the distance to the next extrema (see Figure~\ref{fig:smoothcutb}).
-\item {\bf Match grooves and peaks between cross sections:} based on the regions around the extrema as defined above, we identify joint regions between two cross sections as those areas, in which at least two of the individual regions touch.
-A joint region is defined as the smallest interval that covers all of the touching regions.
-A joint region is called a match between the cross sections, if all of the cross section-specific regions are of the same type of extrema, i.e.\ either all peaks or all valleys.
-In Figure~\ref{fig:match} all matches are colored corresponding to their type of extrema.
-Non-matching regions are left grey.
-\item {\bf Identify the maximal number of CMS} (consecutive matching striae) between the two cross sections by counting consecutive matching lines. 
+\item {\bf Smooth the two signatures} by a loess with a very small span (see Figure~\ref{fig:smooth}). 
+\item Use cross-correlation to {\bf find the best alignment} of the two signatures: shift one of the signatures by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
+\item Using a rolling average, {\bf identify peaks and valleys} for each of the signatures.
+We then define an interval around the location of the extrema on each side as one third of the distance to the location of the next extrema (see Figure~\ref{fig:smoothcutb}).
+\item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify joint intervals between two signatures as those areas, in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals.
+A joint interval is then called a match(ing stria) between the signatures, if all of the intervals are of the same type of extrema, i.e.\ they are either all peaks or all valleys.
+In Figure~\ref{fig:match} all matches are shown as color-filled rectangles corresponding to their type of extrema (peaks are shown in orange, and valleys in green).
+Non-matching intervals are left grey.
+\item {\bf Extract features from the aligned signatures and the matches between them:} there is a slew of different features one could extract.
+Here, we describe a few of the ones that can be found in the literature and some that we found to be of practical relevance:
+\begin{enumerate}[label=\roman*]
+\item maximal number of CMS (consecutive matching striae), and, similarly, the number of consecutively non-matching striae (CNMS), 
+\item number of matches and non-matches,
+\item average difference between signatures $D$, defined as the Euclidean vertical distance of surface measurements between aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
+\[
+D^2 = \frac{1}{\text{\#}t}\sum_t (f(t) - g(t))^2,
+\]
+\item the sum $S$ of matched extrema: for each of the matched stria, determine the average of the absolute height of the peak or valley. $S$ is then defined as the sum of all these averages. 
+\end{enumerate}
+\end{enumerate}
+%
+The difference $D$ between signatures is here defined as the Euclidean distance (in $\mu m$).
+In the paper by \citet{ma:2004}, distance is defined as a measure relative to the first signature, which serves as a comparison reference and is therefore a unitless quantity.
+
+Assessing the maximal number of CMS is part of current practice in evaluating bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}.
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
-For lead bullets, such as this, Biasotti~\citep{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match, and which are part of current practice in assessing bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}.
-\end{enumerate}
+For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
+We will assess this claim in the next section.
 
 Determining a CMS threshold that needs to be exceeded for us to declare a match is beyond the scope of this work, even though it is critically important in practice.
-We provide some ideas in the next section.
+We provide some ideas in the next section, but first we assess the impact of the parameters chosen in the matching algorithm.
 %In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
@@
 @
-\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to the first land of the second bullet shot from barrel 2.
-The largest number of CMS is eleven, observed with land six of the first bullet.
-}
-\end{figure}
-
-<<twolands100, cache=TRUE,  echo=FALSE>>=
-images <- file.path(datadir, dir(datadir))
-
-lof <- processBullets(paths = images[c(5,7)], x = 100)
-
-subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
-lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
-
-# smooth
-lofX <- lofX %>% group_by(bullet) %>% mutate(
-  l30 = smoothloess(y, resid, span = 0.03)
-)
-
-# cut at .75
-threshold <- .75
-lofX$r05 <- threshold* sign(lofX$l30) * as.numeric(abs(lofX$l30) > threshold)
-lofX$type <- factor(lofX$r05)
-levels(lofX$type) <- c("groove", NA, "peak")
-@
-
-<<twolands100adjust, echo=FALSE>>=
-images <- file.path(datadir, dir(datadir))
-
-lof <- processBullets(paths = images[c(5,7)], x = 100)
-
-subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
-subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-ccf <- ccf(subLOFx1$resid, subLOFx2$resid, lag.max=100, plot=FALSE)
-lag <- ccf$lag[which.max(ccf$acf)]
-subLOFx1$y <- subLOFx1$y - lag*1.5625 
-lofY <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
-
-@
-
-<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
-matches <- lofX %>% group_by(y) %>% summarise(
-  potential = (length(unique(type)) == 1),
-  allnas = sum(is.na(type))/n(),
-  type1 = na.omit(type)[1],
-  type = paste(type, sep="|", collapse="|"),
-  n = n()
-)
-
-matches$id <- cumsum(matches$allnas == 1) + 1
-matches$lineid <- as.numeric(matches$allnas != 1) * matches$id
-
-isMatch <- function(id, type) {
-  if (id[1] == 0) return(FALSE)
-#  browser()
-  types <- strsplit(type, split = "|", fixed=TRUE) 
-  t1 <- sapply(types, function(x) x[1])
-  t2 <- sapply(types, function(x) x[2])
-  if (all(t1 == "NA")) return(FALSE)
-  if (all(is.na(t2))) return(FALSE)
-  if (all(t2 == "NA")) return(FALSE)
-
-  peak <- length(grep("peak", c(t1, t2))) > 0
-  groove <- length(grep("groove", c(t1, t2))) > 0
-  if (peak & groove) return(FALSE)
-
-  return(TRUE)
-}
-
-lines <- matches %>% group_by(lineid) %>% summarise(
-  meany = mean(y, na.rm=T),
-  miny = min(y, na.rm=T),
-  maxy = max(y, na.rm=T) + 1.5625,
-  match = isMatch(lineid, type),
-  type = type1[1]
-)
-lines <- subset(lines, lineid != 0)
-@
+\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to Bullet 2-1 (barrel 2). The largest number of CMS is eleven, observed with Bullet 1-6. }
+\end{figure}
 
@@
 \caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$.
-A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
+A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
 \end{figure}
 %
-This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0 $\mu m$ and about 2500 $\mu m$, 1.5625 $\mu m$ apart.
+This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
 The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
@@
 @
-\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet 1-1.
+\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet~1-1.
 Initially, the match between cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
@@
 %An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
-by comparing signatures from cross sections that are not too far apart --  25 $\mu m$ to 50 $\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or from a stable section, where we can have a reasonable expectation of finding matches to other signatures.
+by comparing signatures from cross sections that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
 In the approach here, we keep increasing the cross section until we find a section with a stable pattern.
-This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3, where
+This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
 `stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
@@
 ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(cms))) + theme_bw() +
+  geom_bar(aes(x=factor(CMS))) + theme_bw() +
   theme(legend.position="bottom") + 
@@
 ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
+  geom_bar(aes(x=factor(CMS), fill=match), position="fill") +
   theme(legend.position="bottom") + 
@@
 x <- seq(0,29, by=0.03)
-km <- sm.density(bstats$cms[bstats$match], display="none", eval.points=x)
-knm <- sm.density(bstats$cms[!bstats$match], display="none", eval.points=x)
+km <- sm.density(bstats$CMS[bstats$match], display="none", eval.points=x)
+knm <- sm.density(bstats$CMS[!bstats$match], display="none", eval.points=x)
 dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:distr.dist}Relationship between maximal CMS and distances between cross sections.}
-<<distr.dist, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
+\caption{\label{fig:D}Relationship between maximal CMS and distances between cross sections.}
+<<D, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
 ggplot(data=bstats) + 
-    geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match), alpha=0.5) +
+    geom_jitter(aes(x=factor(CMS), y=D, colour=match), alpha=0.5) +
    facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
@@
 ggplot(data=bstats) + 
-    geom_density(aes(x=distr.dist, fill=match), alpha=0.5) +
+    geom_density(aes(x=D, fill=match), alpha=0.5) +
   xlab("Distances between cross sections") +
@@
 x <- seq(0,20, by=0.02)
-km <- sm.density(bstats$distr.dist[bstats$match], display="none", eval.points=x)
-knm <- sm.density(bstats$distr.dist[!bstats$match], display="none", eval.points=x)
+km <- sm.density(bstats$D[bstats$match], display="none", eval.points=x)
+knm <- sm.density(bstats$D[!bstats$match], display="none", eval.points=x)
 dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
@@
 ggplot(data=bstats) + theme_bw() + 
-  geom_jitter(aes(x=factor(cms), y=ccf, colour=match), alpha=.5) + 
+  geom_jitter(aes(x=factor(CMS), y=ccf, colour=match), alpha=.5) + 
   facet_wrap(~match, labeller="label_both") + 
@@
 Out of the 48 pairs of land-to-land cross sections with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of cross sections that both are very flat with peaks and valleys of less than 1.5$\mu m$).
-The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$distr.dist[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$distr.dist[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$distr.dist[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$distr.dist[!bstats$match])}\mu m$.
+The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.
 Both of these additional variables show large, if not significant, differences between matches and non-matches.
@@
 
-includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
+includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged"))
 cc <- read.csv("csvs/crosscuts-25-new.csv")
@@
 set.seed(20151215)
+names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
+includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged"))
+
 rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=300)
@@
 row.names(imp) <- 1:nrow(imp)
+imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
 names(imp)[-(1:2)] <- c("Match", "(sd)", "Non-Match", "(sd)")

### 2015-12-23 code/evaluation-hamby.R

@@ -33,8 +33,8 @@ CCFs <- plyr::ldply(datas, function(x) {
 # browser()    
     # feature extraction
     data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), 
-               distr.dist=distr.dist, 
-               distr.sd = distr.sd,
+               D=distr.dist, 
+               sd.D = distr.sd,
                b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
                num.matches = sum(res$lines$match), 
                num.mismatches = sum(!res$lines$match), 
@@ -133,7 +133,7 @@ CCFs <- read.csv(file.path(dataStr, "bullet-stats.csv"))
 qplot(factor(cms), data=CCFs)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(cms), fill=match), position="fill")
 sum(CCFs$cms >= 13)
-ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=distr.dist, colour=match))
+ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=D, colour=match))
 ggplot(data=CCFs) + geom_jitter(aes(x=factor(cms), y=ccf, colour=match)) + facet_wrap(~match)
 ggplot(data=CCFs) + geom_bar(aes(x=factor(non_cms), fill=match), position="fill")
 qplot(ccf, num.matches, geom="jitter", data=CCFs, colour=match, alpha=I(0.5)) + facet_wrap(~match)

### 2015-12-24 imaging-paper.Rnw
spineplot reference added

@@
 %In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
+Aside of the features listed above, we also keep track of the value of the cross-correlation function found during the alignment, the  lag, at which signatures show the best alignment, and the cross section levels ($x_1$ and $x_2$).
 
@@
 \hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
-For this, we use cross sections 25$\mu m$ apart and require a minimum of 0.9 for the cross correlation.
+For this, we assess cross-correlation of signatures at cross sections 25$\mu m$ apart.
+For a stable region, we require a minimum of 0.9 for the cross correlation.
 Four lands from different bullets are flagged as problematic in this respect.
-A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface.}
+A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface, also known as `tank rash'.}
 %
@@
 }
- When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the number of maximum consecutive matching striae between land C-3 and all 120 known lands.
+ When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 known lands.
 Two lands show a high CMS.
-These correspond to the known matches, shown in Figures~\ref{fig:hamby-perfect}b and c.
+These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and c.
  Unfortunately, not all results are as clear cut. 
-It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches to get an overview of what we can reasonably expect from an automated match.
+It might not be reasonable to assume that we can match all lands, but the idea is to try to maximize the number of matches to get an overview of what we can reasonably expect from an automated match.
 
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{Maximal number of CMS between unknown bullet C land 3 and all of the 120 other (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown bullet C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{Overlaid lines of C-3 and the land with the top matching CMS.}
+\caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
 <<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{Overlaid lines of C-3 and the land with the second top matching CMS.}
+\caption{Overlaid signatures of C-3 and the land with the second top matching CMS.}
 <<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
@@
 Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
-All 36 pairs of lands with at least sixteen CMS in common are matches.
+All 42 pairs of lands with at least thirteen CMS in common are matches.
 There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
@@
   \centering
-\begin{subfigure}[t]{\textwidth}\centering
+\begin{subfigure}[b]{\textwidth}\centering
 \caption{Distribution of maximal CMS.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{\textwidth}\centering
-\caption{All land-to-land comparisons with a maximum number of CMS of thirteen or higher are matches.}
-<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
+\begin{subfigure}[b]{\textwidth}\centering
+\caption{Conditional barchart \citep{hummel}: heights show probability of match/non-match given a specific CMS. All land-to-land comparisons with at least 13 CMS are matches.}
+<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
 ggplot(data=subset(bstats, !flagged)) + 

### 2015-12-24 imaging-paper.Rnw
proof-read up to 1111

@@
 \modulolinenumbers[5]
+
+\renewcommand{\topfraction}{0.99}	% max fraction of floats at top
+\renewcommand{\bottomfraction}{0.99}	% max fraction of floats at bottom
+% parameters for text pages
+\setcounter{topnumber}{2}
+\setcounter{bottomnumber}{2}
+\setcounter{totalnumber}{4}     % 2 may work better
+\setcounter{dbltopnumber}{2}    % for 2-column pages
+\renewcommand{\dbltopfraction}{0.9}	% fit big float above 2-col. text
+\renewcommand{\textfraction}{0.07}	% allow minimal text w. figs
+%   Parameters for FLOAT pages (not text pages):
+\renewcommand{\floatpagefraction}{0.7}	% require fuller float pages
+% N.B.: floatpagefraction MUST be less than topfraction !!
+\renewcommand{\dblfloatpagefraction}{0.7}	% require fuller float pages
 
@@
 bstats <- read.csv("data-25-25/bullet-stats.csv")
-names(bstats)[9:10] <- c("#matches", "#non-matches")
 
@@
 
-\subsection{Algorithm}
+\subsection{Algorithm}\label{sec:algorithm}
 <<twolands100, cache=TRUE,  echo=FALSE>>=
@@
       geom="line") + 
-  scale_colour_manual("", values=cols) + theme_bw() +
+  scale_colour_manual("Matches", values=cols) + theme_bw() +
   theme(legend.position = c(.9, .2),
@@
  Unfortunately, not all results are as clear cut. 
-It might not be reasonable to assume that we can match all lands, but the idea is to try to maximize the number of matches to get an overview of what we can reasonably expect from an automated match.
+It might not be reasonable to assume that we can match all lands, but the idea is to try to maximize the number of matches to get an overview of what we might be able to expect from an automated match.
 
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
 \caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}\centering
-\caption{Overlaid signatures of C-3 and the land with the second top matching CMS.}
+\begin{subfigure}[b]{.49\textwidth}\centering
+\caption{Top 2 match with C-3 based on CMS.}
 <<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
@@
 All 42 pairs of lands with at least thirteen CMS in common are matches.
+%
+\begin{figure}[hbtp]
+  \centering
+%' \begin{subfigure}[b]{\textwidth}\centering
+%' \caption{}
+\begin{minipage}[t]{.47\textwidth}
+<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
+ggplot(data=subset(bstats, !flagged)) + 
+  geom_bar(aes(x=factor(CMS))) + theme_bw() +
+  theme(legend.position="bottom") + 
+  xlab("maximum CMS")
+@
+%\end{subfigure}
+%\begin{subfigure}[b]{\textwidth}\centering
+\end{minipage}
+\begin{minipage}[t]{.52\textwidth}
+<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='\\textwidth'>>=
+bstats$km <- c( "KNM", "KM")[as.numeric(bstats$match)+1]
+ggplot(data=subset(bstats, !flagged)) + 
+  geom_bar(aes(x=factor(CMS), fill=km), position="fill") +
+  theme(legend.position="bottom") + 
+  scale_fill_brewer("", palette="Paired") +
+  xlab("maximum CMS")
+
+@
+\end{minipage}
+%' \end{subfigure}
+%' \begin{subfigure}[t]{\textwidth}\centering
+%' \caption{\label{fig:cms-dens}Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
+%' <<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.6\\textwidth'>>=
+%' x <- seq(0,29, by=0.03)
+%' km <- sm.density(bstats$CMS[bstats$match],h=2, model="Normal", display="none", eval.points=x)
+%' knm <- sm.density(bstats$CMS[!bstats$match],h=1, display="none", eval.points=x)
+%' dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
+%' dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
+%' dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
+%' 
+%' densities <- data.frame(
+%'   match = rep(c("KM", "KNM"), each=(length(x)+2)),
+%'   x = rep(c(0, x, 29), times = 2),
+%'   estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
+%' 
+%' ggplot() +
+%'   geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
+%'                data=densities, alpha=0.5) +
+%' xlab("maximum CMS") +
+%'   scale_fill_brewer("", palette="Paired") +
+%'   ylab("Density estimates") +
+%'   theme_bw() + 
+%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
+%'         legend.background = element_rect(fill=alpha('white', 0.4))) 
+%' #  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe) 
+%' 
+%' # ggplot(data=bstats) + 
+%' #     geom_density(aes(x=cms, fill=match), alpha=0.5) +
+%' #   xlab("maximum CMS") +
+%' #   scale_fill_brewer("", palette="Paired") +
+%' #   theme_bw() + 
+%' #   theme(legend.position=c(1,1), legend.justification=c(1,1),
+%' #         legend.background = element_rect(fill=alpha('white', 0.4))) 
+%' 
+%' @
+%' \end{subfigure}
+\caption{\label{fig:cms}Distribution of maximal CMS (left).
+Conditional barchart \citep{hummel} on the right: heights show probability of match/non-match given a specific CMS.
+All land-to-land comparisons with at least 13 CMS are matches.}
+\end{figure}
+%
 There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
@@
 Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae.
-Visually we can easily tell these two lands do not match well.
-%
-\begin{figure}[hbtp]
-  \centering
-\begin{subfigure}[b]{\textwidth}\centering
-\caption{Distribution of maximal CMS.}
-<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
-ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(CMS))) + theme_bw() +
-  theme(legend.position="bottom") + 
-  xlab("maximum CMS")
-@
-\end{subfigure}
-\begin{subfigure}[b]{\textwidth}\centering
-\caption{Conditional barchart \citep{hummel}: heights show probability of match/non-match given a specific CMS. All land-to-land comparisons with at least 13 CMS are matches.}
-<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
-ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(CMS), fill=match), position="fill") +
-  theme(legend.position="bottom") + 
-  scale_fill_brewer(palette="Paired") +
-  xlab("maximum CMS")
-
-@
-\end{subfigure}
-\begin{subfigure}[t]{\textwidth}\centering
-\caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
-<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
-x <- seq(0,29, by=0.03)
-km <- sm.density(bstats$CMS[bstats$match], display="none", eval.points=x)
-knm <- sm.density(bstats$CMS[!bstats$match], display="none", eval.points=x)
-dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
-dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
-dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
-
-densities <- data.frame(
-  match = rep(c("KM", "KNM"), each=(length(x)+2)),
-  x = rep(c(0, x, 29), times = 2),
-  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
-
-ggplot() +
-  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
-               data=densities, alpha=0.5) +
-xlab("maximum CMS") +
-  scale_fill_brewer("", palette="Paired") +
-  ylab("Density estimates") +
-  theme_bw() + 
-  theme(legend.position=c(1,1), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) +
-  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)
-
-# ggplot(data=bstats) + 
-#     geom_density(aes(x=cms, fill=match), alpha=0.5) +
-#   xlab("maximum CMS") +
-#   scale_fill_brewer("", palette="Paired") +
-#   theme_bw() + 
-#   theme(legend.position=c(1,1), legend.justification=c(1,1),
-#         legend.background = element_rect(fill=alpha('white', 0.4)))
-
-@
-\end{subfigure}
-\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
-\end{figure}
-
-\begin{figure}[hbtp]
-  \centering
-<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\textwidth', warning=FALSE>>=
+Visually we can easily tell that these two lands do not match well.
+
+\begin{figure}[hbtp]
+  \centering
+<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.65\\textwidth', warning=FALSE>>=
 load("data-25-25/unkn47.RData")
@@
 For smaller numbers of CMS, the percentage of false positives quickly increases.
-However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two cross sections (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
+However, if we take other features of the image into account, we can increase the number of correct matches considerably: Figure~\ref{fig:densities} gives an overview of the densities of all of the features derived earlier, distinguishing between known matches (KM) and known non-matches (KNM).
+The densities of almost all of the features show strong differences between matches and non matches, e.g\ a high amount of cross-correlation between two signatures is indicative for a match --  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
 There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
-\begin{figure}[hbtp]
-  \centering
-\begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:D}Relationship between maximal CMS and distances between cross sections.}
-<<D, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
-ggplot(data=bstats) + 
-    geom_jitter(aes(x=factor(CMS), y=D, colour=match), alpha=0.5) +
-   facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
-  ylab("Distance between cross sections") +
-  scale_colour_brewer(palette="Paired") +
-  theme_bw() + 
-  theme(legend.position="none") +
-    xlab("maximum CMS")
-@
-\end{subfigure}
-\begin{subfigure}[t]{\textwidth}\centering
-\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
-<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
-ggplot(data=bstats) + 
-    geom_density(aes(x=D, fill=match), alpha=0.5) +
-  xlab("Distances between cross sections") +
-  scale_fill_brewer("", palette="Paired") +
-  theme_bw() + 
-  theme(legend.position=c(1,1), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4)))
-
-x <- seq(0,20, by=0.02)
-km <- sm.density(bstats$D[bstats$match], display="none", eval.points=x)
-knm <- sm.density(bstats$D[!bstats$match], display="none", eval.points=x)
-dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
-dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
-dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
-
-densities <- data.frame(
-  match = rep(c("KM", "KNM"), each=(length(x)+2)),
-  x = rep(c(0, x, 29), times = 2),
-  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
-
-ggplot() +
-  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
-               data=densities, alpha=0.5) +
-  xlab("Distances between cross sections") +
-  scale_fill_brewer("", palette="Paired") +
-  ylab("Density estimates") +
-  theme_bw() + 
-  theme(legend.position=c(1,1), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) +
-  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)
-
-@
-\end{subfigure}
-\caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
-\end{figure}
-
-\begin{figure}[hbtp]
-  \centering
-\begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
-<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
-ggplot(data=bstats) + theme_bw() + 
-  geom_jitter(aes(x=factor(CMS), y=ccf, colour=match), alpha=.5) + 
-  facet_wrap(~match, labeller="label_both") + 
-    scale_colour_brewer(palette="Paired") +
-  xlab("maximum CMS") + ylab("Cross-correlation") +
-  theme(legend.position="none")
-@
-\end{subfigure}
-\begin{subfigure}[t]{\textwidth}\centering
-\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
-<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
-ggplot(data=bstats) + 
-    geom_density(aes(x=ccf, fill=match), alpha=0.5) +
-  xlab("Cross-correlation between cross sections") +
-  scale_fill_brewer("", palette="Paired") +
-  theme_bw() + 
-  theme(legend.position=c(1,1), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) 
-@
-\end{subfigure}
-\begin{subfigure}[t]{\textwidth}\centering
-\caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
-<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
-x <- seq(0,1, by=0.001)
-library(sm)
-km <- sm.density(bstats$ccf[bstats$match], display="none", eval.points=x)
-knm <- sm.density(bstats$ccf[!bstats$match], display="none", eval.points=x)
-dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
-dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
-dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
-
-densities <- data.frame(
-  x = rep(c(0, x, 1), times = 2),
-  estimate = c(0, km$estimate, 0, 0, knm$estimate,0),
-  match = rep(c("KM", "KNM"), each=1003))
-
-ggplot() +
-  geom_polygon(aes(x=x, y=estimate, fill=match), colour="black",
-               data=densities, alpha=0.5) +
-  xlab("Cross-correlation between cross sections") +
-  scale_fill_brewer("", palette="Paired") +
-#  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
-  theme_bw() + 
-  ylab("Density estimate") +
-  theme(legend.position=c(.975,1), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) +
-  geom_line(aes(x, pmin(5, kmtail/knmtail)), data=dframe) 
-@
-\end{subfigure}
-\caption{A large cross-correlation between cross sections is indicative of matches between lands.}
-\end{figure}
-
-Similarly, the Euclidean distance between two cross sections across the range of loess residuals is strongly correlated to known matches.
-Out of the 48 pairs of land-to-land cross sections with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of cross sections that both are very flat with peaks and valleys of less than 1.5$\mu m$).
+
+\begin{figure}[hbtp]
+  \centering
+<<density-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=7, out.width='\\textwidth'>>=
+features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
+densities <- plyr::ldply(features, function(x) {
+  xr <- range(bstats[,x])
+  xx <- seq(xr[1], xr[2], length.out=500)
+
+  densKM <- sm.density(bstats[,x][bstats$match], display="none", 
+                       eval.points=xx, weights=NA, method="normal")
+  densKNM <- sm.density(bstats[,x][!bstats$match], display="none", 
+                       eval.points=xx, weights=NA, method="normal")
+  dframe <- data.frame(var=x, x = xx, KM=densKM$estimate, KNM = densKNM$estimate)
+  dframe <- rbind(data.frame(var=x, x = xr[1], KM=0, KNM=0),
+        dframe, data.frame(var=x, x = xr[2], KM=0, KNM=0))
+  dframe$order <- 1:nrow(dframe)
+  dframe
+}) 
+library(reshape2)
+dm <- melt(densities, measure.var=c("KM", "KNM"))
+dm$var <- factor(dm$var, levels=features)
+levels(dm$var)[3:4] <- c("#matches", "#non-matches")
+
+qplot(x, value, group=variable, geom="polygon", data=dm, fill=variable, 
+      alpha=I(0.6), colour=I("grey20")) + 
+  facet_wrap(~var, ncol=4, scales="free") +
+  theme_bw() + ylab("density") + xlab("") +
+  scale_fill_brewer("Matches", palette="Paired") +
+  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
+        legend.background = element_rect(colour="grey60"),
+        plot.margin = unit(c(0,0,0,0), unit="cm"))
+@
+\caption{\label{fig:densities}Overview of all the marginal densities for features extracted in section~\ref{sec:algorithm}.}
+\end{figure}
+
+%' \begin{figure}[hbtp]
+%'   \centering
+%' \begin{subfigure}[t]{\textwidth}\centering
+%' \caption{\label{fig:D}Relationship between maximal CMS and distances between cross sections.}
+%' <<D, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
+%' ggplot(data=bstats) + 
+%'     geom_jitter(aes(x=factor(CMS), y=D, colour=match), alpha=0.5) +
+%'    facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
+%'   ylab("Distance between cross sections") +
+%'   scale_colour_brewer(palette="Paired") +
+%'   theme_bw() + 
+%'   theme(legend.position="none") +
+%'     xlab("maximum CMS")
+%' @
+%' \end{subfigure}
+%' \begin{subfigure}[t]{\textwidth}\centering
+%' \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
+%' <<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
+%' ggplot(data=bstats) + 
+%'     geom_density(aes(x=D, fill=match), alpha=0.5) +
+%'   xlab("Distances between cross sections") +
+%'   scale_fill_brewer("", palette="Paired") +
+%'   theme_bw() + 
+%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
+%'         legend.background = element_rect(fill=alpha('white', 0.4))) 
+%' 
+%' 
+%' 
+%' x <- seq(0,20, by=0.02)
+%' km <- sm.density(bstats$D[bstats$match], display="none", eval.points=x)
+%' knm <- sm.density(bstats$D[!bstats$match], display="none", eval.points=x)
+%' dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
+%' dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
+%' dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
+%' 
+%' densities <- data.frame(
+%'   match = rep(c("KM", "KNM"), each=(length(x)+2)),
+%'   x = rep(c(0, x, 29), times = 2),
+%'   estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
+%' 
+%' ggplot() +
+%'   geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
+%'                data=densities, alpha=0.5) +
+%'   xlab("Distances between cross sections") +
+%'   scale_fill_brewer("", palette="Paired") +
+%'   ylab("Density estimates") +
+%'   theme_bw() + 
+%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
+%'         legend.background = element_rect(fill=alpha('white', 0.4))) +
+%'   geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe) 
+%' 
+%' @
+%' \end{subfigure}
+%' \caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
+%' \end{figure}
+
+%' \begin{figure}[hbtp]
+%'   \centering
+%' \begin{subfigure}[t]{\textwidth}\centering
+%' \caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
+%' <<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
+%' ggplot(data=bstats) + theme_bw() + 
+%'   geom_jitter(aes(x=factor(CMS), y=ccf, colour=match), alpha=.5) + 
+%'   facet_wrap(~match, labeller="label_both") + 
+%'     scale_colour_brewer(palette="Paired") +
+%'   xlab("maximum CMS") + ylab("Cross-correlation") +
+%'   theme(legend.position="none")
+%' @
+%' \end{subfigure}
+%' \begin{subfigure}[t]{\textwidth}\centering
+%' \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
+%' <<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
+%' ggplot(data=bstats) + 
+%'     geom_density(aes(x=ccf, fill=match), alpha=0.5) +
+%'   xlab("Cross-correlation between cross sections") +
+%'   scale_fill_brewer("", palette="Paired") +
+%'   theme_bw() + 
+%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
+%'         legend.background = element_rect(fill=alpha('white', 0.4))) 
+%' @
+%' \end{subfigure}
+%' \begin{subfigure}[t]{\textwidth}\centering
+%' \caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
+%' <<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
+%' x <- seq(0,1, by=0.001)
+%' library(sm)
+%' km <- sm.density(bstats$ccf[bstats$match], display="none", eval.points=x)
+%' knm <- sm.density(bstats$ccf[!bstats$match], display="none", eval.points=x)
+%' dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
+%' dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
+%' dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
+%' 
+%' densities <- data.frame(
+%'   x = rep(c(0, x, 1), times = 2),
+%'   estimate = c(0, km$estimate, 0, 0, knm$estimate,0),
+%'   match = rep(c("KM", "KNM"), each=1003))
+%' 
+%' ggplot() +
+%'   geom_polygon(aes(x=x, y=estimate, fill=match), colour="black",
+%'                data=densities, alpha=0.5) +
+%'   xlab("Cross-correlation between cross sections") +
+%'   scale_fill_brewer("", palette="Paired") +
+%' #  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
+%'   theme_bw() + 
+%'   ylab("Density estimate") +
+%'   theme(legend.position=c(.975,1), legend.justification=c(1,1),
+%'         legend.background = element_rect(fill=alpha('white', 0.4))) +
+%'   geom_line(aes(x, pmin(5, kmtail/knmtail)), data=dframe) 
+%' @
+%' \end{subfigure}
+%' \caption{A large cross-correlation between cross sections is indicative of matches between lands.}
+%' \end{figure}
+
+Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches.
+Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of cross sections that both are very flat with peaks and valleys of less than 1.5$\mu m$).
 The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.
-Both of these additional variables show large, if not significant, differences between matches and non-matches.
+All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches.
 Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of  a set of feature variables derived from the 3d topological surface measurements.
@@
 vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
-
-includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged"))
+names(bstats)[9:10] <- c("#matches", "#non-matches")
+includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
 cc <- read.csv("csvs/crosscuts-25-new.csv")
@@
 names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
-includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged"))
+includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
 

### 2015-12-24 imaging-paper.Rnw
first read-through done. next step take care of the issues that came up

@@
 % "Title of the paper"
-\title{Automatic Matching of Bullet Lands}
+\title{Automatic Matching of Bullet Lands}\thanksref{T1}
+\thankstext{T1}{Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}
 \runtitle{Automatic Matching of Bullet Lands}
@@
 
-\thankstext{t1}{Thanks to somebody \hh{XXX acknowledge the grant here? XXX}} 
+\thankstext{t1}{This research was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.}
 \affiliation{Iowa State University\thanksmark{m1}}
@@
 Here, we describe a few of the ones that can be found in the literature and some that we found to be of practical relevance:
-\begin{enumerate}[label=\roman*]
+\begin{enumerate}[label=(\roman*)]
 \item maximal number of CMS (consecutive matching striae), and, similarly, the number of consecutively non-matching striae (CNMS), 
@@
 Instead of CMS, cross-correlation (ccf) between the cross sections is very important in the matching process  by the decision tree.
-Aside from  cross correlation, the number of matches (num.matches) is involved in the decision.
-Within matches and non-matches, the values of CMS and cross-correlation seem to be almost independent (see Figure~\ref{fig:ccfs-match}), but both are highly predictive in distinguishing matches from non-matches.
-Out of the two, cross-correlation has higher  predictive power.
+Aside from  cross correlation, the number of matches is involved in the decision.
+Between cross-correlation and CMS, cross-correlation has higher  predictive power.
 This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2d microscopic images of bullets.
-Neither one of these methods allows for an assessment of cross-correlations. This is one of the benefits of switching to a digitalized version of the images that preserves the 3d surface structure.
+Neither one of these methods allows for an assessment of cross-correlations.
+This is one of the benefits of switching to a digitalized version of the images that preserves the 3d surface structure.
+\hh{The emphasis on cross-correlation is similar to the results of the study by \citet{ma:2004}.
+However, in this study authors did not consider the number of matches and non-matches.}
 
@@
 library(reshape2)
-set.seed(20151215)
+
+set.seed(20151202)
 names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
@@
 
-Another benefit is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\citep{breiman:2001, randomForest}.
-In a random forest multiple trees are fitted.
-For each of these trees, only two thirds of the observations are used for fitting, while predictions for the remaining third  are used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate.
+Another benefit of the digitized version of the images is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\citep{breiman:2001, randomForest}.
+For each of the trees in a random forest, only two thirds of the observations are used for fitting, while  the remaining third  are used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate.
 Because errors are determined from the one third of held-back observations, this error rate is called the out-of bag (OOB) error.
@@
 %
-After about 25 trees, the error rate of land-to-land comparisons stabilizes at 0.0040.
-This is a weighted average between a false positive error rate of $9.8 \times 10^{-5}$ and an error rate of false negatives of 0.22.
+After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0038.
+This is a weighted average between a zero false positive error rate and an error rate of false negatives of 0.23.
 \hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known matches and non-matches (see Figures~\ref{fig:treeresult} and ~\ref{fig:tree-forest}).}
-Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
-In the case of the Hamby study, the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched (see Figure~\ref{fig:treeresult}).
-\begin{figure}[hbtp]
-  \centering
-<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
-bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
-bullets <- bstats %>% filter(match, !flagged) %>% group_by(bullet) %>% summarize(
-  n = n(),
-  pred = sum(pred > 0.5),
-  forest = sum(forest > 0.5)
-)
-library(reshape2)
-bm <- melt(data=bullets, id.var=c("bullet", "n"))
-
-bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
-qplot(data=bm, 100*value/n, reorder(bullet, value/n), size=I(2.5), 
-      shape=prediction, colour=prediction, alpha=I(0.6)) + 
-  theme_bw()  + ylab("") + 
-#  scale_shape_manual(values=c(1,2)) +
-  scale_x_continuous("Percent of correctly predicted land-to-land matches", 
-                     breaks = c(0,25,50,75,100), limits=c(0,100)) +
-  theme(legend.position = "bottom")
-@
-\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12, except for Q and B, where the maximum is 11).
+%Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
+In the case of the Hamby study, even for the single tree the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched (see Figure~\ref{fig:treeresult}).
+%' \begin{figure}[hbtp]
+%'   \centering
+%' <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
+%' bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
+%' bullets <- bstats %>% filter(match, !flagged) %>% group_by(bullet) %>% summarize(
+%'   n = n(),
+%'   pred = sum(pred > 0.5),
+%'   forest = sum(forest > 0.5)
+%' )
+%' library(reshape2)
+%' bm <- melt(data=bullets, id.var=c("bullet", "n"))
+%' 
+%' bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
+%' qplot(data=bm, 100*value/n, reorder(bullet, value/n), size=I(2.5), 
+%'       shape=prediction, colour=prediction, alpha=I(0.6)) + 
+%'   theme_bw()  + ylab("") + 
+%' #  scale_shape_manual(values=c(1,2)) +
+%'   scale_x_continuous("Percent of correctly predicted land-to-land matches", 
+%'                      breaks = c(0,25,50,75,100), limits=c(0,100)) +
+%'   theme(legend.position = "bottom")
+%' @
+%' \caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12, except for Q and B, where the maximum is 11).
 Both results for the tree and the random forest are shown.
@@
 Unknown bullet B is the hardest for the tree to match.}
-\end{figure}
-%
+%' \end{figure}
+%' %
 However, this does not carry a lot of weight because of the few bullet comparisons in question: fifteen unknown bullets are successfully matched to two pairs of ten bullets.
@@
 In practice, examiners are allowed a third option of `inconclusive'.
-On a probability spectrum of outcomes we can therefore introduce an interval of `inconclusive' results in the middle of the spectrum.
+On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be not necessary in the Hamby study, because, here, the results from the random forest are very clear cut.
 Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis).
-\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match.
+%\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match.
 The random forest predicts all 172 known matches correctly as well as all 10,212 known non-matches.}
@@
 The rows in the table are ordered according to descending importance.
-The variable with the most predictive power by far is cross-correlation (ccf), followed by the number of matches (num.matches) and mismatches(num.mismatches), and the sum of the height of joint peaks (sumpeaks).
-The maximal consecutively matching striae (cms) is found only in sixth place.
+The variable with the most predictive power by far is cross-correlation, followed by the number of matches and non-matches.
+CMS is found only in sixth place.
 
@@
 
-An approach that might help in determining this difference comes from Biometrics, is suggested by \citet{jain:2015:tphil, jain:2015} and has been successfully applied in the Aadhar project (the effort by the Unique Identification Authority of India (UIDAI) to provide a unique 12-digit identification number to approximately 1.2 billion residents of India based on ten fingerprints and two irises).
+An approach that might help in determining this difference comes from Biometrics \citep{jain:2015:tphil, jain:2015}.
+It has been successfully applied in the Aadhar project (the effort by the Unique Identification Authority of India (UIDAI) to provide a unique 12-digit identification number to approximately 1.2 billion residents of India based on ten fingerprints and two irises).
 This approach uses dictionary learning based on small square patches  of 64x64 or 32x32 pixels.
@@
 \paragraph{Acknowledgment} Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.
-
-This research was partially funded in part by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement \#70NANB15H176 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, University of California Irvine, and University of Virginia.
-
-Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.
 

### 2015-12-25 imaging-paper.Rnw
cleveland added, standard errors still need discussion

@@
 \begin{enumerate}
-\item take out a few things: no circular fit, no figures 9 and 10 
-\item do we need the web app?
-\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
+%\item take out a few things: no circular fit, no figures 9 and 10 
+%\item do we need the web app?
+%\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
 \item add: formula for loess fit and its s.e.
-\item add: comparison of error rates to those of FEs in Hamby study.
-\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
+%\item add: comparison of error rates to those of FEs in Hamby study.
+%\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
 \item add: Why do we take the lowest stable cross section?
+\item add: more discussion of the density plots: probability ratio for a match
+\item add: more discussion on additional bullet images.
+NIST database claims to have 70 bullets -- that would be two sets of the study, but only 35 bullets are currently available.
+Should we contact Alan Zheng to see whether we can get access to more bullet files?
+\item add: should we add a paragraph on the speed of the evaluation?
+- separate out into pre-processing (needs to be done once) and matching (pairwise comparison of lands).
+If we only do bullet matching, we could speed this process up by eliminating some land comparisons.
 \end{enumerate}
@@
 Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
-This process has broad applicability in terms of the conviction  in the United States criminal justice system.
+This process has broad applicability in terms of  convictions  in the United States criminal justice system.
 Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny.
@@
 
-Rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet during the firing process.
+Rifling, manufacturing defects, and impurities in a barrel create striation marks on the bullet during the firing process.
 These marks are assumed to be unique to the barrel, as described in a 1992 AFTE article~\citep{afte:1992}.
@@
 \hh{XXX More detailed discussion of automatic approaches and database discussion XXX}
-\hh{XXX Alternative approach by \citet{roberge:2006}: using correlation to identify matches between pairs of bullets using a correlation measure, that `could be automated'.
-No algorithm implemented/available, no error rates of automated matching are given XXX}
-\hh{XXX Interesting paper by \citet{ma:2004} with  emphasis on ccf and acf.
-They are using a stylus instrument for measureing the big curvature XXX -- the advantage of our approach is that nobody needs to look at anything until after the matching is done.
-}
-\hh{XXX Hamby study: 507 participants returned results.
-If a match was reported, these matches were correct.
-But eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.}
-\hh{XXX for Hamby study: automated systems were successfully used for matching:
- Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
-Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
-Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
-BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)
- }
-
-The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
-%We then describe a web application for visualizing and working with bullet surfaces.
-We then proceed to describing an automatic matching routine which we evaluate on a small database of reference bullets.
-
-\section{Bullet signatures}
+
 Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
 Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
-
 3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
@@
 The direction of the rifling is a class characteristic, i.e.\ a feature that pertains to a particular class of firearms, and is not unique at the individual barrel/bullet level.
+
+\hh{The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. }
+\hh{XXX The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; however, no false-negative error rates for bullets nor details on land-to-land matches were given.
+% Hamby study: automated systems were successfully used for matching:
+%  Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
+% Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
+% Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
+% BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)
+ }
+\hh{XXX Matching of bullets using the BulletTrax3D system is described based on another set of ten consecutively rifled barrels in \citet{roberge:2006}.
+They describe a process using a bullet-to-bullet correlation score to identify matches.
+The authors state that this process `could be automated', but no implementation of the algorithm is available.
+XXX}
+
+The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
+%We then describe a web application for visualizing and working with bullet surfaces.
+We then proceed to describing an automatic matching routine which we evaluate on the bullets made available through the Hamby study.
+
+\section{Bullet signatures}
 
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step1}Step 1 of the loess fit algorithm.
-For a fixed cross section ($x$ held constant at 243.75~$\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
+\caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed cross section ($x$ held constant at 243.75~$\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step3}Step 2 of the loess fit algorithm.
-The  surface measurements are smoothed twice with a smoothing factor of $s = 35$.
+\caption{\label{fig:loess_step3}Step 2 of identifying groove locations: The  surface measurements are smoothed twice with a smoothing factor of $s = 35$.
 The orange rectangle shows an example of the smoothing window.
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step47}Steps 3 -- 6 of the loess fitting algorithm.
-After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
+\caption{\label{fig:loess_step47}Steps 3 -- 6 of identifying groove locations: After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
 Values outside the blue boundaries are removed (shown in grey)}
@@
 Next, we fit a loess regression to the data.
-
-\hh{XXX Introduce Loess formula with all of its details in here. we are using the default of {\tt loess}, which has a span of 0.75 and uses a polynomial of degree 2. XXX}
+\hh{Loess regression \citep{cleveland:1979} is based on the assumption that the relationship between two random variables $X$ and $Y$ can be described in form of a smooth, continuous function $f$ with $y_i = f(x_i) + \varepsilon_i$ for all values $i = 1, ..., n$.
+The function $f$ is estimated in form of  locally weighted polynomial regressions.
+Parameters of the estimation are $\alpha$, the proportion of all points included in the fit (here, $\alpha = 0.75$), the weighting function (here, tricubic weighting $(1 - (\text{dist}/\text{maxdist})^3)^3$) and the degree of the polynomial fitted (here, degree is two).
+}
+
+\hh{XXX standard errors XXX}
 Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of bullet~1-5.
@@
   \centering
-\begin{subfigure}[t]{.49\textwidth}\centering
+\begin{subfigure}[b]{.49\textwidth}\centering
 \caption{\label{fig:loess_fit} Loess fit for bullet~1-5.}
@@
 \end{subfigure}
-\begin{subfigure}[t]{.49\textwidth}\centering
-\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5.}
+\begin{subfigure}[b]{.49\textwidth}\centering
+\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5. Blue areas show pointwise 95\% confidence intervals.}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
@@
       geom="line") + 
-  scale_colour_manual("Matches", values=cols) + theme_bw() +
-  theme(legend.position = c(.9, .2),
+  scale_colour_manual("", values=cols) + theme_bw() +
+  theme(legend.position = c(1, 1), legend.justification=c("right", "top"),
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
@@
   scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position = c(.9, .2),
+  theme(legend.position = c(1, 1), legend.justification=c("right", "top"),
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
@@
 \hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figures~\ref{fig:treeresult} and ~\ref{fig:tree-forest}).}
-%Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
-In the case of the Hamby study, even for the single tree the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched (see Figure~\ref{fig:treeresult}).
+Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
+In the case of the Hamby study, even for the single tree the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched.
+% (see Figure~\ref{fig:treeresult}).
 %' \begin{figure}[hbtp]
@@
 %' %
-However, this does not carry a lot of weight because of the few bullet comparisons in question: fifteen unknown bullets are successfully matched to two pairs of ten bullets.
+\hh{This makes the errors in the automated approach smaller than the human error in the Hamby study.
+Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.}
+
+However, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets.
 Matching bullets can only be tested realistically in a much bigger test case.
@@
 Table~\ref{tab:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree.
-\begin{table*}[tbhp]
+\begin{table}[tbhp]
+\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches.
+Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
+Match and Non-Match show averages of each of the variables for matches and non-matches only.
+In parentheses behind the means, standard deviations are given}
 \centering
@@
 @
-\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches.
-Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
-Match and Non-Match show averages of each of the variables for matches and non-matches only.
-In parentheses behind the means, standard deviations are given}
-\end{table*}
+\end{table}
 
@@
 
-\paragraph{Acknowledgment} Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.
+\section*{Acknowledgment} 
+Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.
 

### 2015-12-26 imaging-paper.Rnw
standard error comment

@@
 The function $f$ is estimated in form of  locally weighted polynomial regressions.
-Parameters of the estimation are $\alpha$, the proportion of all points included in the fit (here, $\alpha = 0.75$), the weighting function (here, tricubic weighting $(1 - (\text{dist}/\text{maxdist})^3)^3$) and the degree of the polynomial fitted (here, degree is two).
-}
-
-\hh{XXX standard errors XXX}
+Parameters of the estimation are $\alpha$, the proportion of all points included in the fit (here, $\alpha = 0.75$), the weighting function and the degree of the polynomial (here, we fit a quadratic regression).
+}
+
+\hh{The main idea of locally weighted regression is to use a weighting routine that  emphasizes the effect of points close by and de-emphasizes the effect of points as they are further away.
+The weighting function used here, is the tricubic function $w(d) = (1 - d^3)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
+$d$ is defined as the the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
+}
+
+\hh{XXX standard errors: are based on independence of errors (which we don't have because of the high autocorrelation) and the assumption that the amount of bias is neglible.
+Which we don't know.
+We can produce standard errors, but (a) the nominal confidence level will be violated, and (b) the errors only allow for pointwise errors.
+We need a different approach.
+XXX}
+
 Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of bullet~1-5.
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5. Blue areas show pointwise 95\% confidence intervals.}
+\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5. %Blue areas show pointwise 95\% confidence intervals.
+}
 <<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=

### 2015-12-26 imaging-paper.Rnw
bootstrap for CIs

@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5. %Blue areas show pointwise 95\% confidence intervals.
-}
-<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
-my.loess$resid + ylab("Residuals of loess Fit") 
+\caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5.
+This defines the {\it signature} of bullet~1-5. Blue areas show  95\% bootstrap confidence intervals (based on $B=500$ bootstrap samples).
+}
+%<<loess_resid, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
+%my.loess$resid + ylab("Residuals of loess Fit") 
+%@
+<<boot_loess, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
+boot.loess <- boot_fit_loess(br111_bullet, br111.groove)
+poly <- with(boot.loess, data.frame(x = c(y,rev(y)), y=c(high, rev(low))))
+qplot(x=x, y=y, geom="polygon", fill=I("steelblue"), data=poly) +
+  geom_line(aes(x=y, y=resid), data=my.loess$data) +
+  ylab("Residuals of loess fit") + xlab("y") + theme_bw()
 @

### 2015-12-26 imaging-paper.Rnw
alignment in table

@@
 \modulolinenumbers[5]
+\usepackage{dcolumn}
+\newcolumntype{d}[1]{D{.}{.}{#1}}
 
@@
 \caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5.
-This defines the {\it signature} of bullet~1-5. Blue areas show  95\% bootstrap confidence intervals (based on $B=500$ bootstrap samples).
+This defines the {\it signature} of bullet~1-5. Blue areas show  95\% bootstrap confidence intervals (based on $B=1,000$ bootstrap samples).
 }
@@
 Interestingly, the number of consecutive matching striae does not feature in this evaluation.
-Instead of CMS, cross-correlation (ccf) between the cross sections is very important in the matching process  by the decision tree.
+Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree.
 Aside from  cross correlation, the number of matches is involved in the decision.
@@
 This is a weighted average between a zero false positive error rate and an error rate of false negatives of 0.23.
-\hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figures~\ref{fig:treeresult} and ~\ref{fig:tree-forest}).}
+\hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).}
 Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
@@
 Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
-Match and Non-Match show averages of each of the variables for matches and non-matches only.
-In parentheses behind the means, standard deviations are given}
-\centering
-<<importance, echo=FALSE, results='asis'>>=
+Averages (and standard deviations) for known matches (KM) and known non-matches (KNM) are shown in the last four columns.}
+\centering
+<<importance, echo=FALSE, results='asis', warning=FALSE>>=
 imp <- data.frame(importance(rtrees))
@@
 imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
-names(imp)[-(1:2)] <- c("Match", "(sd)", "Non-Match", "(sd)")
-print(xtable(imp, align="clrrcrc"), floating = FALSE)
+names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
+print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
+      floating = FALSE)
 @

### 2015-12-26 imaging-paper.Rnw
some explicit definitions

@@
 If we only do bullet matching, we could speed this process up by eliminating some land comparisons.
+\item definitions: 
+A bullet's {\it signature} are the residual surface measurements after removing the big structure using the loess fit.
+The {\it cross section} defines the location along the $x$ axis  at which the signature is taken.
 \end{enumerate}
@@
     geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
-  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
+  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
 @
@@
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) +
+  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
   xlab("y")
@@
 \caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet~1-1.
-Initially, the match between cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
-At a cross section level of $175\mu m$ the bullet pattern stabilizes.
+Initially, the match between signatures from cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
+At a  level of $175\mu m$ the bullet's signature stabilizes.
 For this land, matches should not be attempted at lower cross sections.
@@
 \caption{\label{fig:mismatch}Known mismatch with a relatively large number of maximal consecutive matching striae (twelve) in the middle.
-The pattern in the middle does look surprisingly similar, however the outer ends of the cross sections easily reveals this as mismatch.
+The pattern in the middle does look surprisingly similar, however the outer ends of the signatures easily reveals this as mismatch.
 }
@@
   \centering
-<<density-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=7, out.width='\\textwidth'>>=
+<<density-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
 features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
@@
 levels(dm$var)[3:4] <- c("#matches", "#non-matches")
+levels(dm$variable) <- c("Known matches (KM)", "Known non-matches (KNM)")
 
@@
   theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
-        legend.background = element_rect(colour="grey60"),
+        legend.background = element_rect(colour="grey75"),
         plot.margin = unit(c(0,0,0,0), unit="cm"))
 @
-\caption{\label{fig:densities}Overview of all the marginal densities for features extracted in section~\ref{sec:algorithm}.}
+\caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}.}
 \end{figure}
@@
 Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches.
-Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of cross sections that both are very flat with peaks and valleys of less than 1.5$\mu m$).
+Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$).
 The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.

### 2015-12-27 imaging-paper.Rnw
discussion of bullets

@@
 \hh{The main idea of locally weighted regression is to use a weighting routine that  emphasizes the effect of points close by and de-emphasizes the effect of points as they are further away.
-The weighting function used here, is the tricubic function $w(d) = (1 - d^3)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
-$d$ is defined as the the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
-}
-
-\hh{XXX standard errors: are based on independence of errors (which we don't have because of the high autocorrelation) and the assumption that the amount of bias is neglible.
+The weighting function used here, is the tricubic function $w(d) = \left(1 - d^3\right)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
+$d$ is defined as the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
+}
+
+%\hh{XXX standard errors: are based on independence of errors (which we don't have because of the high autocorrelation) and the assumption that the amount of bias is neglible.
 Which we don't know.
@@
 The fit seems to do a reasonable job of capturing the structure of the image.
-Figure~\ref{fig:loess_resid} shows the residuals from this fit.
-These residuals are called the \emph{signature} of bullet~1-5.
+\hh{The blue area around the signature are  95\% bootstrapped confidence intervals based on 1,000 bootstrap samples.
+}
+Figure~\ref{fig:loess_resid} shows the residuals from this fit. These residuals are called the \emph{signature} of bullet~1-5.
 %
@@
 %
-Figure~\ref{fig:loessfits} shows the signatures of all lands of bullet 2 (in grey) fired from the same barrel as bullet~1.
+
+%' \hh{XXX Do we need the next example? We could save some space here XXX}
+%' Figure~\ref{fig:loessfits} shows the signatures of all lands of bullet 2 (in grey) fired from the same barrel as bullet~1.
 The darker line overlaid on each of the signatures, shows signature from bullet~1-5.
-%We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet~1-5.
+%' %We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet~1-5.
 This is shown in Figure~\ref{fig:loessfits}.
-Bullet 2-1 is the best match for bullet~1-5, but the alignment is not quite  correct.
-
-\begin{figure}[hbtp]
-  \centering
-<<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
-br121.groove <- get_grooves(br121_bullet)
-br122.groove <- get_grooves(br122_bullet)
-br123.groove <- get_grooves(br123_bullet)
-br124.groove <- get_grooves(br124_bullet)
-br125.groove <- get_grooves(br125_bullet)
-br126.groove <- get_grooves(br126_bullet)
-
-my.loess1 <- fit_loess(br121_bullet, br121.groove)
-my.loess2 <- fit_loess(br122_bullet, br122.groove)
-my.loess3 <- fit_loess(br123_bullet, br123.groove)
-my.loess4 <- fit_loess(br124_bullet, br124.groove)
-my.loess5 <- fit_loess(br125_bullet, br125.groove)
-my.loess6 <- fit_loess(br126_bullet, br126.groove)
-
-my.loessactual <- fit_loess(br111_bullet, br111.groove)
-
-minrows <- min(length(my.loess1$data$resid),
-               length(my.loess2$data$resid),
-               length(my.loess3$data$resid),
-               length(my.loess4$data$resid),
-               length(my.loess5$data$resid),
-               length(my.loess6$data$resid),
-               length(my.loessactual$data$resid))
-
-my.df <- data.frame(y = my.loess1$data$y[1:minrows], resid1 = my.loess1$data$resid[1:minrows],
-                    resid2 = my.loess2$data$resid[1:minrows], resid3 = my.loess3$data$resid[1:minrows],
-                    resid4 = my.loess4$data$resid[1:minrows], resid5 = my.loess5$data$resid[1:minrows],
-                    resid6 = my.loess6$data$resid[1:minrows])
-
-melt.df <- my.df %>%
-    gather(key = part, value = resid, resid1:resid6)
-melt.df$part <- gsub("resid", "", melt.df$part)
-
-actual.df <- my.loessactual$data[1:minrows,]
-melt.df$land <- melt.df$part
-qplot(y, resid, data=melt.df, #colour=factor(x), 
-      geom="line", size=I(.75), colour=I("grey70")) + 
-  facet_wrap(~land, ncol=3, labeller="label_both") + 
-  scale_colour_brewer(palette="Paired") +
-  theme_bw() + theme(legend.position="bottom") + 
-  geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Signatures of bullet 1 in grey, bullet 1-5 in black") + ylab("Residuals of loess fit")
-@    
-\caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet~1-5 is shown in black. }
-\end{figure}
+%' Bullet 2-1 is the best match for bullet~1-5, but the alignment is not quite  correct. \hh{This is one of the issues we are addressing next.}
+%' 
+%' \begin{figure}[hbtp]
+%'   \centering
+%' <<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
+%' br121.groove <- get_grooves(br121_bullet)
+%' br122.groove <- get_grooves(br122_bullet)
+%' br123.groove <- get_grooves(br123_bullet)
+%' br124.groove <- get_grooves(br124_bullet)
+%' br125.groove <- get_grooves(br125_bullet)
+%' br126.groove <- get_grooves(br126_bullet)
+%' 
+%' my.loess1 <- fit_loess(br121_bullet, br121.groove)
+%' my.loess2 <- fit_loess(br122_bullet, br122.groove)
+%' my.loess3 <- fit_loess(br123_bullet, br123.groove)
+%' my.loess4 <- fit_loess(br124_bullet, br124.groove)
+%' my.loess5 <- fit_loess(br125_bullet, br125.groove)
+%' my.loess6 <- fit_loess(br126_bullet, br126.groove)
+%' 
+%' my.loessactual <- fit_loess(br111_bullet, br111.groove)
+%' 
+%' minrows <- min(length(my.loess1$data$resid),
+%'                length(my.loess2$data$resid),
+%'                length(my.loess3$data$resid),
+%'                length(my.loess4$data$resid),
+%'                length(my.loess5$data$resid),
+%'                length(my.loess6$data$resid),
+%'                length(my.loessactual$data$resid))
+%' 
+%' my.df <- data.frame(y = my.loess1$data$y[1:minrows], resid1 = my.loess1$data$resid[1:minrows],
+%'                     resid2 = my.loess2$data$resid[1:minrows], resid3 = my.loess3$data$resid[1:minrows],
+%'                     resid4 = my.loess4$data$resid[1:minrows], resid5 = my.loess5$data$resid[1:minrows],
+%'                     resid6 = my.loess6$data$resid[1:minrows])
+%' 
+%' melt.df <- my.df %>%
+%'     gather(key = part, value = resid, resid1:resid6)
+%' melt.df$part <- gsub("resid", "", melt.df$part)
+%' 
+%' actual.df <- my.loessactual$data[1:minrows,]
+%' melt.df$land <- melt.df$part
+%' qplot(y, resid, data=melt.df, #colour=factor(x), 
+%'       geom="line", size=I(.75), colour=I("grey70")) + 
+%'   facet_wrap(~land, ncol=3, labeller="label_both") + 
+%'   scale_colour_brewer(palette="Paired") +
+%'   theme_bw() + theme(legend.position="bottom") + 
+%'   geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Signatures of bullet 1 in grey, bullet 1-5 in black") + 
+%'   ylab("Residuals of loess fit") +
+%'   theme(plot.margin = unit(c(0,0,0,0), unit="cm"))
+%' @    
+%' \caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet~1-5 is shown in black. }
+%' \end{figure}
+%'
 
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{Using a rolling median peaks and valleys are identified for each signature. \label{fig:smoothcutb}}{%
+\caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
 <<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
@@
 \begin{subfigure}[t]{\textwidth}\centering
-    \caption{Rectangles in the back identify a line on one of the bullets.
+    \caption{Rectangles in the back identify a striation mark on one of the bullets.
 Matching striation marks are indicated by color filled rectangles and marked by an `o'.
@@
 \[
-D^2 = \frac{1}{\text{\#}t}\sum_t (f(t) - g(t))^2,
+D^2 = \frac{1}{\text{\#}t}\sum_t \left(f(t) - g(t)\right)^2,
 \]
@@
 
-\begin{figure}[hbtp]
-  \centering
-<<barrel2, echo=FALSE, fig.width = 10, fig.height=6, out.width='\\textwidth', warning=FALSE>>=
-
-list_of_matches <- lapply(25:30, function(i) {
-  lof <- processBullets(paths = images[c(31,i)], x = 100)
-  lof <- bulletSmooth(lof)
-
-  bAlign = bulletAlign(lof)
-  lofX <- bAlign$bullet
-
-  b12 <- unique(lof$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
-  peaks1$lines$bullet = b12[1]
-  peaks2$lines$bullet = b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  ggplot() +
-    theme_bw() + 
-    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=6, data=lines, alpha=0.2) +
-    geom_line(aes(x = y, y = l30, linetype = bullet),  data = lofX, alpha=0.6) +
-    scale_linetype_discrete("") +
-    scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-    ylim(c(-6,6)) + ylab("") + xlab("") +
-    geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
-    geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
-    theme(plot.margin=unit(c(1,0,0,0), unit="line"))
-})
-
-grid.arrange(list_of_matches[[1]], list_of_matches[[2]], list_of_matches[[3]], 
-             list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
-             ncol = 2)
-@
-\caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to Bullet 2-1 (barrel 2). The largest number of CMS is eleven, observed with Bullet 1-6. }
-\end{figure}
+%' \begin{figure}[hbtp]
+%'   \centering
+%' <<barrel2, echo=FALSE, fig.width = 10, fig.height=6, out.width='\\textwidth', warning=FALSE>>=
+%' 
+%' list_of_matches <- lapply(25:30, function(i) {
+%'   lof <- processBullets(paths = images[c(31,i)], x = 100)
+%'   lof <- bulletSmooth(lof)
+%' 
+%'   bAlign = bulletAlign(lof)
+%'   lofX <- bAlign$bullet  
+%' 
+%'   b12 <- unique(lof$bullet)
+%'   peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
+%'   peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
+%'   peaks1$lines$bullet = b12[1]
+%'   peaks2$lines$bullet = b12[2]
+%'   lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+%' 
+%'   ggplot() +
+%'     theme_bw() + 
+%'     geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=6, data=lines, alpha=0.2) +
+%'     geom_line(aes(x = y, y = l30, linetype = bullet),  data = lofX, alpha=0.6) +
+%'     scale_linetype_discrete("") +
+%'     scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
+%'   theme(legend.position = c(1,1.2), legend.justification=c(1,1),
+%'         legend.background = element_rect(fill=alpha('white', 0.4))) + 
+%'     ylim(c(-6,6)) + ylab("") + xlab("") +
+%'     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
+%'     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
+%' #    theme(plot.margin=unit(c(1,0,0,0), unit="line")) +
+%'     theme(plot.margin=unit(c(0,0,0,0), unit="cm"))
+%' })
+%' 
+%' grid.arrange(list_of_matches[[1]], list_of_matches[[2]], list_of_matches[[3]], 
+%'              list_of_matches[[4]], list_of_matches[[5]], list_of_matches[[6]],
+%'              ncol = 2)
+%' @
+%' \caption{\label{fig:barrel2}All lands from the first bullet shot from barrel 2 are compared to Bullet 2-1 (barrel 2).
+The largest number of CMS is eleven, observed with Bullet 1-6.
+\hh{XXX Do we need this example?
+This might be a place to cut XXX}}
+%' \end{figure}
 

### 2015-12-27 imaging-paper.Rnw
ratios, raw figure

@@
 \runtitle{Automatic Matching of Bullet Lands}
+\runauthor{E. Hare et al.}
 
@@
 \phantom{E-mail:\ }\printead*{e3}}
-
-\runauthor{E. Hare et al.}
 \end{aug}
+%
 \begin{abstract}
@@
 %\item out: lots of the figures can go into a supplement, together with a discussion. In the paper we then have a summary of what can be seen, so that we can focus on the results. 
-\item add: formula for loess fit and its s.e.
+%\item add: formula for loess fit and its s.e.
 %\item add: comparison of error rates to those of FEs in Hamby study.
@@
 \hh{The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. }
-\hh{XXX The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; however, no false-negative error rates for bullets nor details on land-to-land matches were given.
+\hh{The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
 % Hamby study: automated systems were successfully used for matching:
@@
  }
-\hh{XXX Matching of bullets using the BulletTrax3D system is described based on another set of ten consecutively rifled barrels in \citet{roberge:2006}.
-They describe a process using a bullet-to-bullet correlation score to identify matches.
+\hh{A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}.
+This study is based on another set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score.
 The authors state that this process `could be automated', but no implementation of the algorithm is available.
-XXX}
+}
 
@@
 Two lands show a high CMS.
-These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and c.
+These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and~\ref{fig:hamby-perfect}c.
  Unfortunately, not all results are as clear cut. 
@@
 There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
-On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}).
-Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}.
-Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}.
+On average, there are \Sexpr{mean(bstats$CMS[!bstats$match], na.rm=T)} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$CMS[!bstats$match])}).
+Known matches share on average \Sexpr{mean(bstats$CMS[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$CMS[bstats$match])}.
+While the probability for a match increases with the number of maximal CMS, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}.
 Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae.
@@
 \end{figure}
+
+<<prob.ratios, echo=FALSE>>=
+library(reshape2)
+dens.stats <- dm %>% group_by(var, variable) %>% mutate(
+  sum = sum(value),
+  up = cumsum(value)/sum,
+  down = 1 - cumsum(value)/sum
+)
+ratios <- dens.stats %>% group_by(var, x) %>% summarise(
+  upper = sum(up[variable=="KM"])/sum(up[variable=="KNM"]),
+  lower = sum(down[variable=="KM"])/sum(down[variable=="KNM"])
+)
+rm <- melt(ratios, measure.var=c("upper", "lower"))
+levels(rm$variable) <- c("P(X < c | KM) / P(X < c | KNM)", "P(X > c | KM) / P(X > c | KNM)")
+qplot(x, value, colour=variable,  data=rm) + facet_wrap(~var, scales="free") + scale_y_log10() + theme_bw() + geom_hline(yintercept=1) + scale_colour_brewer(palette="Set2") + ylab("Probability Ratios")
+@
 

### 2015-12-28 imaging-paper.Rnw
fixed problem with levels

@@
 dens.stats <- dm %>% group_by(var, variable) %>% mutate(
-  sum = sum(value),
+  sum = sum(value, na.rm=TRUE),
   up = cumsum(value)/sum,
@@
 )
+levels(dens.stats$variable) <- c("KM", "KNM")
 ratios <- dens.stats %>% group_by(var, x) %>% summarise(
-  upper = sum(up[variable=="KM"])/sum(up[variable=="KNM"]),
-  lower = sum(down[variable=="KM"])/sum(down[variable=="KNM"])
+  lower = sum(up[variable=="KM"], na.rm=TRUE)/sum(up[variable=="KNM"], na.rm=TRUE),
+  upper = sum(down[variable=="KM"], na.rm=TRUE)/sum(down[variable=="KNM"], na.rm=TRUE)
 )
-rm <- melt(ratios, measure.var=c("upper", "lower"))
+rm <- melt(ratios, measure.var=c("lower", "upper"))
 levels(rm$variable) <- c("P(X < c | KM) / P(X < c | KNM)", "P(X > c | KM) / P(X > c | KNM)")

### 2015-12-29 imaging-paper.Rnw
cleaned up probability ratios, included ROC curves

--- 2015-12-27 imaging-paper.Rnw
@@
 @
-\caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}.}
-\end{figure}
-
-<<prob.ratios, echo=FALSE>>=
+\caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}.
+Shifts in the mode of the density functions between known matches and known non-matches indicate the variable's predictive power in distinguishing matches and non-matches.
+}
+\end{figure}
+
+\begin{figure}[hbtp]
+  \centering
+<<ratios-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
 library(reshape2)
 dens.stats <- dm %>% group_by(var, variable) %>% mutate(
-  sum = sum(value),
+  sum = sum(value, na.rm=TRUE),
   up = cumsum(value)/sum,
@@
 )
+levels(dens.stats$variable) <- c("KM", "KNM")
 ratios <- dens.stats %>% group_by(var, x) %>% summarise(
-  upper = sum(up[variable=="KM"])/sum(up[variable=="KNM"]),
-  lower = sum(down[variable=="KM"])/sum(down[variable=="KNM"])
+  lower = sum(up[variable=="KM"], na.rm=TRUE)/sum(up[variable=="KNM"], na.rm=TRUE),
+  upper = sum(down[variable=="KM"], na.rm=TRUE)/sum(down[variable=="KNM"], na.rm=TRUE)
 )
-rm <- melt(ratios, measure.var=c("upper", "lower"))
+rm <- melt(ratios, measure.var=c("lower", "upper"))
+rm$value <- pmin(pmax(rm$value, 0.1), 10)
 levels(rm$variable) <- c("P(X < c | KM) / P(X < c | KNM)", "P(X > c | KM) / P(X > c | KNM)")
-qplot(x, value, colour=variable,  data=rm) + facet_wrap(~var, scales="free") + scale_y_log10() + theme_bw() + geom_hline(yintercept=1) + scale_colour_brewer(palette="Set2") + ylab("Probability Ratios")
-@
+qplot(x, value, colour=variable,  data=rm, geom="line", group=variable, 
+      size=I(1)) + xlab("X") +
+  facet_wrap(~var, scales="free", ncol=4) + 
+  scale_y_log10(limits=c(1/10,10)) + theme_bw() + 
+  geom_hline(yintercept=1, size=0.5) + 
+  scale_colour_brewer("Estimated Ratios", palette="Set2") + 
+  ylab("(Logged) Probability Ratios") +
+  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
+        legend.background = element_rect(colour="grey75"),
+        plot.margin = unit(c(0,0,0,0), unit="cm")) 
+@
+\caption{\label{fig:ratios}Estimated probability ratios for all of the features described in section~\ref{sec:algorithm}.
+Ratios are logged and and artificially constrained to an interval between (1/10 and 10).
+Large deviations from one, indicate a strong distinction between matches and known matches based on the variable.
+Both left tail and right tail probabilities are shown.
+}
+\end{figure}
+
+\begin{figure}[hbtp]
+  \centering
+<<rocs-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
+# plot false positive against false negative
+# prob of false positive: P(match | KNM)
+# prob of false negative: P(non-match | KM)
+
+# for  X < c: we get probability for P(X < c | KNM) and P(X < c | KM)
+
+features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
+errors <- plyr::ldply(features, function(x) {
+  xx <- unique(bstats[,x])
+  if (length(xx) > 500) {
+    xr <- range(bstats[,x])
+    xx <- seq(xr[1], xr[2], length.out=500)
+  }
+  # upper and lower rule:
+  # upper: match is defined as X > xx
+  # lower: match is defined as X < xx
+  subbstats <- subset(bstats, flagged == FALSE)
+  subbstats$match <- factor(subbstats$match)
+
+  errors <- plyr::ldply(xx, function(cc) {
+    idx <- which(subbstats[,x] >= cc)
+    upper <- data.frame(xtabs(~match, data= subbstats[idx,,drop=FALSE])/
+                          xtabs(~match, data= subbstats))
+    idx <- which(subbstats[,x] <= cc)
+    lower <- data.frame(xtabs(~match, data= subbstats[idx,,drop=FALSE])/
+                          xtabs(~match, data= subbstats))
+
+    data.frame(value = cc, match=lower[,1], lower=lower[,2], upper=upper[,2])
+  })
+
+  errors$variable <- x
+  errors
+})
+
+rocs.lower <- dcast(errors, variable+value~match, value.var="lower")
+rocs.lower$type <- "lower"
+rocs.upper <- dcast(errors, variable+value~match, value.var="upper")
+rocs.upper$type <- "upper"
+rocs <- rbind(rocs.lower, rocs.upper)
+
+rocs$variable <- factor(rocs$variable, levels=features)
+qplot(`FALSE`, `TRUE`, data=rocs,  geom="line", group=type) + theme_bw() + geom_abline(linetype=2) + xlab("False negative rate") + 
+  ylab("False positive rate") + facet_wrap(facets=~variable, ncol=4)
+@
+\caption{\label{fig:rocs}ROC curves for all of the features described in section~\ref{sec:algorithm}.
+\hh{XXX compare ROC curve of CMS against the other individual features -- ccf should be a lot more predictive than CMS.
+XXX}
+\hh{XXX next steps: get rid of lower curve. find area under the curve, put value on bottom right of the chart, sort by ROC.}}
+\end{figure}
 

### 2015-12-29 imaging-paper.Rnw
ROC and AUC

@@
 
-rocs$variable <- factor(rocs$variable, levels=features)
-qplot(`FALSE`, `TRUE`, data=rocs,  geom="line", group=type) + theme_bw() + geom_abline(linetype=2) + xlab("False negative rate") + 
-  ylab("False positive rate") + facet_wrap(facets=~variable, ncol=4)
+aucs <- plyr::ldply(features, function(x) {
+  subbstats <- subset(bstats, flagged == FALSE)
+  pos.scores <- sample(subbstats[which(subbstats$match),x], 5000, replace=TRUE)
+  neg.scores <- sample(subbstats[which(!subbstats$match),x], 5000, replace=TRUE)
+  data.frame(x, upper=mean(pos.scores > neg.scores), 
+             lower = mean(pos.scores < neg.scores))
+})
+aucs <- melt(aucs, measure.var=c("lower", "upper"))
+names(aucs) <- c("variable", "type", "auc")
+rocs <- merge(rocs, aucs, by=c("variable", "type"))
+labels <- unique(subset(rocs, auc > 0.5)[,c("variable", "auc")])
+labels$labels <- paste("AUC:", labels$auc)
+labels$type <- NA
+
+rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
+qplot(`FALSE`, `TRUE`, data=subset(rocs, auc > 0.5),  geom="line", group=type) + theme_bw() + geom_abline(linetype=2) + xlab("False negative rate") + 
+  ylab("False positive rate") + facet_wrap(facets=~variable, ncol=4) + 
+  geom_label(x=1, y=0, aes(label=labels), data=labels, hjust=1, vjust=0)
 @
 \caption{\label{fig:rocs}ROC curves for all of the features described in section~\ref{sec:algorithm}.
-\hh{XXX compare ROC curve of CMS against the other individual features -- ccf should be a lot more predictive than CMS.
-XXX}
-\hh{XXX next steps: get rid of lower curve. find area under the curve, put value on bottom right of the chart, sort by ROC.}}
+Variables are sorted according to their area under the curve (AUC).
+Except for the distance $D$ between signatures, all features are more predictive than the maximal CMS.
+}
 \end{figure}

### 2015-12-29 imaging-paper.Rnw
shading of ROC curves

@@
 
+ggplot(roc.data, aes(x=fpr, ymin=0, ymax=tpr)) +
+    geom_ribbon(alpha=0.2) +
+    geom_line(aes(y=tpr)) +
+    ggtitle(paste0("ROC Curve w/ AUC=", auc))
+
 rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
-qplot(`FALSE`, `TRUE`, data=subset(rocs, auc > 0.5),  geom="line", group=type) + theme_bw() + geom_abline(linetype=2) + xlab("False negative rate") + 
+ggplot(data=subset(rocs, auc > 0.5)) +
+  geom_ribbon(aes(x=`FALSE`, ymax=`TRUE`, ymin=0), alpha=0.2) +
+  geom_line(aes(x=`FALSE`,y=`TRUE`))+ theme_bw() + xlab("False negative rate") + 
   ylab("False positive rate") + facet_wrap(facets=~variable, ncol=4) + 
-  geom_label(x=1, y=0, aes(label=labels), data=labels, hjust=1, vjust=0)
+  geom_label(x=1, y=0, aes(label=labels), data=labels, hjust=1, vjust=0) + 
+  geom_abline(linetype=2)
 @

### 2015-12-29 imaging-paper.Rnw
probability ratios commented out

@@
 Shifts in the mode of the density functions between known matches and known non-matches indicate the variable's predictive power in distinguishing matches and non-matches.
-}
-\end{figure}
-
-\begin{figure}[hbtp]
-  \centering
-<<ratios-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
-library(reshape2)
-dens.stats <- dm %>% group_by(var, variable) %>% mutate(
-  sum = sum(value, na.rm=TRUE),
-  up = cumsum(value)/sum,
-  down = 1 - cumsum(value)/sum
-)
-levels(dens.stats$variable) <- c("KM", "KNM")
-ratios <- dens.stats %>% group_by(var, x) %>% summarise(
-  lower = sum(up[variable=="KM"], na.rm=TRUE)/sum(up[variable=="KNM"], na.rm=TRUE),
-  upper = sum(down[variable=="KM"], na.rm=TRUE)/sum(down[variable=="KNM"], na.rm=TRUE)
-)
-rm <- melt(ratios, measure.var=c("lower", "upper"))
-rm$value <- pmin(pmax(rm$value, 0.1), 10)
-levels(rm$variable) <- c("P(X < c | KM) / P(X < c | KNM)", "P(X > c | KM) / P(X > c | KNM)")
-qplot(x, value, colour=variable,  data=rm, geom="line", group=variable, 
-      size=I(1)) + xlab("X") +
-  facet_wrap(~var, scales="free", ncol=4) + 
-  scale_y_log10(limits=c(1/10,10)) + theme_bw() + 
-  geom_hline(yintercept=1, size=0.5) + 
-  scale_colour_brewer("Estimated Ratios", palette="Set2") + 
-  ylab("(Logged) Probability Ratios") +
-  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
-        legend.background = element_rect(colour="grey75"),
-        plot.margin = unit(c(0,0,0,0), unit="cm")) 
-@
-\caption{\label{fig:ratios}Estimated probability ratios for all of the features described in section~\ref{sec:algorithm}.
+Predictive power is shown in more detail in Figure~\ref{fig:rocs}.}
+\end{figure}
+
+%' \begin{figure}[hbtp]
+%'   \centering
+%' <<ratios-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
+%' library(reshape2)
+%' dens.stats <- dm %>% group_by(var, variable) %>% mutate(
+%'   sum = sum(value, na.rm=TRUE),
+%'   up = cumsum(value)/sum,
+%'   down = 1 - cumsum(value)/sum
+%' )
+%' levels(dens.stats$variable) <- c("KM", "KNM")
+%' ratios <- dens.stats %>% group_by(var, x) %>% summarise(
+%'   lower = sum(up[variable=="KM"], na.rm=TRUE)/sum(up[variable=="KNM"], na.rm=TRUE),
+%'   upper = sum(down[variable=="KM"], na.rm=TRUE)/sum(down[variable=="KNM"], na.rm=TRUE)
+%' )
+%' rm <- melt(ratios, measure.var=c("lower", "upper"))
+%' rm$value <- pmin(pmax(rm$value, 0.1), 10)
+%' levels(rm$variable) <- c("P(X < c | KM) / P(X < c | KNM)", "P(X > c | KM) / P(X > c | KNM)")
+%' qplot(x, value, colour=variable,  data=rm, geom="line", group=variable, 
+%'       size=I(1)) + xlab("X") +
+%'   facet_wrap(~var, scales="free", ncol=4) + 
+%'   scale_y_log10(limits=c(1/10,10)) + theme_bw() + 
+%'   geom_hline(yintercept=1, size=0.5) + 
+%'   scale_colour_brewer("Estimated Ratios", palette="Set2") + 
+%'   ylab("(Logged) Probability Ratios") +
+%'   theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
+%'         legend.background = element_rect(colour="grey75"),
+%'         plot.margin = unit(c(0,0,0,0), unit="cm")) 
+%' @
+%' \caption{\label{fig:ratios}Estimated probability ratios for all of the features described in section~\ref{sec:algorithm}.
 Ratios are logged and and artificially constrained to an interval between (1/10 and 10).
@@
 }
-\end{figure}
+%' \end{figure}
 
@@
 
-ggplot(roc.data, aes(x=fpr, ymin=0, ymax=tpr)) +
-    geom_ribbon(alpha=0.2) +
-    geom_line(aes(y=tpr)) +
-    ggtitle(paste0("ROC Curve w/ AUC=", auc))
-
 rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
@@
 Variables are sorted according to their area under the curve (AUC).
-Except for the distance $D$ between signatures, all features are more predictive than the maximal CMS.
+Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS.
 }

### 2015-12-29 imaging-paper.Rnw
eaning up ROC discussion

@@
 However, if we take other features of the image into account, we can increase the number of correct matches considerably: Figure~\ref{fig:densities} gives an overview of the densities of all of the features derived earlier, distinguishing between known matches (KM) and known non-matches (KNM).
-The densities of almost all of the features show strong differences between matches and non matches, e.g\ a high amount of cross-correlation between two signatures is indicative for a match --  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
-There are 97 land-to-land comparisons with a cross-correlation of above 0.75.
+The densities of almost all of the features show strong differences between matches and non matches.
+For example, a high amount of cross-correlation between two signatures is indicative for a match --  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
+There are 97 land-to-land comparisons with a cross-correlation that high.
 
@@
 Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$).
-The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.
+%The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.
 All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches.
-Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of  a set of feature variables derived from the 3d topological surface measurements.
-There are quite large differences in these averages, indicating that we can successfully employ machine learning methods to distinguish matches from non-matches.
+The predictive power of each one of these features is shown in form of ROC curves in Figure~\ref{fig:rocs}.
+The features are ordered according to the area under the curve (AUC).
+The feature with the highest individual predictive power is $S$, the sum of two signatures' heights of joint peaks and valleys.
+The maximal number of CMS is only in the seventh position here.
+The overall high AUC values indicate, that we can successfully employ machine learning methods to distinguish matches from non-matches.
+%Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3d topological surface measurements.
+There are quite large differences in these averages, indicating
 

### 2015-12-29 imaging-paper.Rnw
todo items updated

@@
 \item add: Why do we take the lowest stable cross section?
-\item add: more discussion of the density plots: probability ratio for a match
+%\item add: more discussion of the density plots: probability ratio for a match - exchanged probability ratios for ROC curves
 \item add: more discussion on additional bullet images.
@@
 If we only do bullet matching, we could speed this process up by eliminating some land comparisons.
+\item add: we might need a proper definition of the joint striation marks (the result of striation-identify)
+\item change: we could make the table into a labelled dotplot of importance to save space. We don't need the mean and standard deviations anymore -- the density plots and the ROCs take care of that.
+\item change: the focus of the paper is initially quite strongly on the maximal consecutively matching striae (CMS).
+We need to emphasis the role of CMS in the introduction: max CMS is Biasotti's main criterium for matching.
+max CMS is part of current practice (AFTE criteria citation).
+AFIS \cite{AFIS} is suggesting it for an automated system (was there ever a phase II of that project?).
+Problem is, that determining CMS is subjective; counting CMS in particular, has been found to be subjective to the examiner: `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.
+For an automatic routine we are faced with the problem of having to determine what a stria is.
+XXX Should we include one of those microscopic images to bring that point home?
+We would need somebody qualified to tell us how many CMS there are, though.
+\item to consider: AOAS states that accepted papers rarely have more than 20 pages. We are currently at about 23--24 pages.
 \item definitions: 
-A bullet's {\it signature} are the residual surface measurements after removing the big structure using the loess fit.
+A bullet's {\it signature} is defined as the residual surface measurements after removing the big structure using the loess fit.
 The {\it cross section} defines the location along the $x$ axis  at which the signature is taken.
@@
 rocs <- merge(rocs, aucs, by=c("variable", "type"))
+rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
+
 labels <- unique(subset(rocs, auc > 0.5)[,c("variable", "auc")])
@@
 
-rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
 ggplot(data=subset(rocs, auc > 0.5)) +

### 2015-12-30 imaging-supplement.Rnw
up to stability.
crosscuts-sd.csv file and some discussion (In the supplement) added

@@
 
-\section{Cylindrical Fit}
+\section{Cylindrical Fit}\label{supp:cylindrical}
 
@@
 
-\subsection{Assessing cross-correlation between signatures at multiple cross sections}
+\section{Assessing cross-correlation between signatures at multiple cross sections}
 
@@
 \end{figure}
+
+\section{Signature intensities}
+
+\hh{XXX Overview of the signatures across one bullet land. Maybe with color?}
+
+Standard deviation of signatures is one measure to quantify the extent to which they are expressed.
+Figure~\ref{fig:crosscuts} shows the amount of the standard deviation across crosscuts for all bullet lands shown as lines.
+Standard deviation decreases as the crosscut increases, i.e.\ bullets' signatures are most expressed close to the bottom of the bullet.
+\hh{XXX can we find an AFTE citation to back this up?}
+
+\begin{figure}[hbtp]
+\centering
+<<crosscuts-sd, echo=FALSE, fig.height=5, fig.width=8, out.width='.7\\textwidth'>>=
+if (!file.exists("csvs/crosscut-sd.csv")) {
+  bullets_processed <- lapply(all_bullets, function(bul) {
+    cat("Computing processed bullet", basename(bul$path), "\n")
+    crosscuts <- 25*(1:20)
+    dframe <- new_processBullets(bullet = bul, x = crosscuts)
+    dframe$bullet <- with(dframe, paste(bullet, x))
+    dframe
+  })
+  names(bullets_processed) <- as.character(ccs$path)
+
+  bullets_smoothed <- bullets_processed %>% bind_rows %>% bulletSmooth
+
+  stats <- bullets_smoothed %>% group_by(bullet, x) %>% summarize(
+    sd = sd(l30, na.rm=T)
+  )
+  splits <- strsplit(stats$bullet, split=".", fixed=TRUE)
+  bnames <- sapply(splits, function(x) paste(x[1],".x3p", sep=""))
+  stats$bullet <- bnames
+
+  write.csv(stats, "csvs/crosscut-sd.csv", row.names=FALSE)
+} else {
+  stats <- read.csv("csvs/crosscut-sd.csv")
+}
+
+qplot(x, sd, data=stats, geom="line", group=bullet, alpha=I(0.5)) + theme_bw() +
+  geom_smooth(group=1) + 
+  xlab("Crosscut location") +
+  ylab("Standard deviation")
+@
+\caption{\label{fig:crosscuts}Standard deviation reduces as crosscut levels increase.}
+\end{figure}
+
 \end{document}

### 2015-12-30 imaging-supplement.Rnw
discussion of signature intensities

--- 2015-12-23 imaging-supplement.Rnw
@@
 
+\section{Cylindrical Fit}\label{supp:cylindrical}
+Figure~\ref{fig:fixedX} shows the side profile of a crosscut through the surface measurements of bullet 1-5 at a fixed level.
+The smooth line on top is a circle, with estimated radius and center.
+The details of this fit are given below.
+We then remove the fit to investigate residuals.
+%
 \begin{figure}[hbtp]
@@
 
-\section{Cylindrical Fit}
+%One reasonable approach is to fit a circle to the curve and compute the residuals from this fit.
 
-One reasonable approach is to fit a circle to the curve and compute the residuals from this fit.
-
-For this, assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle.
+Assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle.
 We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.
@@
 @
-\caption{Circle fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
+\caption{Circular fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
 \end{figure}
 
-\subsection{Assessing cross-correlation between signatures at multiple cross sections}
+\newpage
+\section{Assessing cross-correlation between signatures at multiple cross sections}\label{supp:ccf}
 
@@
 \end{figure}
+
+\newpage
+\section{Signature intensities}\label{supp:bulletbottom}
+<<setup-signatures, echo=FALSE>>=
+library(x3pr)
+library(x3prplus)
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(zoo)
+
+knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
+unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+
+###############
+# can we identify the barrels the unknown bullets came from?
+
+new_get_bullet <- function(path, x = 243.75, bullet = NULL) {
+  if (is.null(bullet)) bullet <- read.x3p(path)
+  dbr111 <- fortify_x3p(bullet)
+
+  pickx <- dbr111$x[which.min(abs(x - unique(dbr111$x)))]
+
+  dbr111.fixx <- dbr111[dbr111$x == pickx,]
+
+  return(dbr111.fixx)
+}
+
+new_processBullets <- function(bullet, x = 100) {
+  crosscuts <- unique(fortify_x3p(bullet)$x)
+  crosscuts <- crosscuts[crosscuts >= min(x)]
+  crosscuts <- crosscuts[crosscuts <= max(x)]
+  if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]
+
+  list_of_fits <- lapply(crosscuts, function(x) {
+    br111 <- new_get_bullet(path = NULL, x = x, bullet = bullet)
+    br111.groove <- get_grooves(br111)
+    fit_loess(br111, br111.groove)$resid$data
+  })
+  lof <- list_of_fits %>% bind_rows
+
+  cbind(lof, bullet = bullet$path)
+}
+
+new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
+  lof <- rbind(lof1, lof2)
+  bAlign = bulletAlign(lof)
+  lofX <- bAlign$bullet
+
+  b12 <- unique(lof$bullet)
+  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
+  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)
+
+  #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
+  #    theme_bw() +
+  #    geom_rect(data=peaks1$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25)) +
+  #    geom_rect(data=peaks2$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25))
+
+  peaks1$lines$bullet <- b12[1]
+  peaks2$lines$bullet <- b12[2]
+
+  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+
+  #   p <- qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
+  #     theme_bw() +
+  #     geom_rect(data=lines, aes(xmin=xmin, xmax=xmax, fill = factor(type)),  ymin=-6, ymax=6, inherit.aes = FALSE, alpha=I(0.25)) +
+  #     ylim(c(-6,6)) +
+  #     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match), inherit.aes = FALSE) +
+  #     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match), inherit.aes = FALSE)
+
+  maxCMS <- maxCMS(lines$match==TRUE)
+  list(maxCMS = maxCMS, ccf = bAlign$ccf, lag=bAlign$lag, lines=lines, bullets=lofX)
+}
+
+# match unknown land using crosscuts
+ccs <- read.csv("csvs/crosscuts-25.csv")
+all_bullets <- lapply(as.character(ccs$path), function(x) {
+  result <- read.x3p(x)
+  result[[3]] <- x
+  names(result)[3] <- "path"
+
+  return(result)
+})
+
+knowns <- all_bullets[1:120]
+unknowns <- all_bullets[121:210]
+
+if (!file.exists("csvs/crosscut-sd.csv")) {
+  bullets_processed <- lapply(all_bullets, function(bul) {
+    cat("Computing processed bullet", basename(bul$path), "\n")
+    crosscuts <- 25*(1:20)
+    dframe <- new_processBullets(bullet = bul, x = crosscuts)
+    dframe$bullet <- with(dframe, paste(bullet, x))
+    dframe
+  })
+  names(bullets_processed) <- as.character(ccs$path)
+
+  bullets_smoothed <- bullets_processed %>% bind_rows %>% bulletSmooth
+
+  stats <- bullets_smoothed %>% group_by(bullet, x) %>% summarize(
+    sd = sd(l30, na.rm=T)
+  )
+  splits <- strsplit(stats$bullet, split=".", fixed=TRUE)
+  bnames <- sapply(splits, function(x) paste(x[1],".x3p", sep=""))
+  stats$bullet <- bnames
+
+  write.csv(stats, "csvs/crosscut-sd.csv", row.names=FALSE)
+} else {
+  stats <- read.csv("csvs/crosscut-sd.csv")
+}  
+@
+%
+Figure~\ref{fig:overview} shows an overview of the signatures across different crossucts on a single bullet. 
+\begin{figure}[hbtp]
+<<one-bulletland-sd, echo=FALSE, dependson='setup-signatures', fig.height=5, fig.width=8, out.width='0.7\\textwidth', warning=FALSE>>=
+k <- 1
+x <- as.character(ccs$path)[k]
+land <- read.x3p(x)
+land[[3]] <- x
+names(land)[3] <- "path"
+
+crosscuts <- 25*(1:20)
+dframe <- new_processBullets(bullet = land, x = crosscuts)
+dframe$bullet <- with(dframe, paste(bullet, x))
+
+bullet <- dframe %>% bind_rows %>% bulletSmooth 
+subbullet <- subset(bullet, x >= ccs$cc[k])
+qplot(x=y, geom="line", y=l30, colour=x, group=x, data=subbullet) + theme_bw() +
+  scale_colour_gradient("Crosscuts") +
+  theme(legend.position="bottom") +
+  ylab("Residuals from loess fit")
+@
+\caption{\label{fig:overview}Signatures of the same bullet at different crosscuts. With increasing crosscut level, peaks and valleys are less pronounced, resulting in a smaller standard deviation.}
+\end{figure}
+As crosscut levels increase individual characteristics become less distinctive, making true matches to other bullets harder.
+The pattern of decreasing peaks and valleys is generally true for  bullet lands, as can be seen in Figure~\ref{fig:sds}.
+%
+\begin{figure}[hbtp]
+\centering
+<<crosscuts-sd, echo=FALSE, fig.height=5, fig.width=8, out.width='.7\\textwidth', dependson='setup-signatures'>>=
+qplot(x, sd, data=stats, geom="line", group=bullet, alpha=I(0.5)) + theme_bw() +
+  geom_smooth(group=1) + 
+  xlab("Crosscut location") +
+  ylab("Standard deviation")
+@
+\caption{\label{fig:sds}Standard deviation reduces as crosscut levels increase.}
+\end{figure}
+Figure~\ref{fig:sds} shows that the amount of standard deviation of a signature decreases on average for all bullet lands as crosscut levels increase.
+This makes standard deviation of a signature one measure to quantify the extent to which a signature is expressed.
+This implies that for identifying matches we should use the lowest crosscut level to extract a bullet's signature once a stable surface region is detected.
+This is in accordance with current standard practice \citep{afte:1992}.
+
+\bibliographystyle{imsart-nameyear.bst}
+%\biboptions{authoryear}
+\bibliography{references}
+
 \end{document}

### 2015-12-30 imaging-paper.Rnw
reference to supplement

@@
 %\item change: emphasis in varying cross section. once a stable region is found, how little change is happening? 
-\item add: Why do we take the lowest stable cross section?
+%\item add: Why do we take the lowest stable cross section?
 %\item add: more discussion of the density plots: probability ratio for a match - exchanged probability ratios for ROC curves
@@
 \hh{While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of a circular fit and the remaining residual structure can be found in the supplement.} \hh{XXX Reference to the supplement?
-XXX}
+A discussion of a circular fit and the remaining residual structure can be found in supplement~\ref{supp:cylindrical}.}
 %It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet.
-\hh{(assumption here: lower cross section levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
+We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet \hh{(see supplement~\ref{supp:bulletbottom} for a more detailed discussion)}.
 
@@
 This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: we have to be aware of which cross section level was used in a matching.
-An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement \hh{XXX section XXX} for a more thorough discussion).
+An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement~\ref{supp:ccf} for a more thorough discussion).
 However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. 
@@
 `stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
-
-\hh{XXX Why do we use the lowest possible cross section? XXX}
 
@@
 rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
+levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")
 
@@
 All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches.
-The predictive power of each one of these features is shown in form of ROC curves in Figure~\ref{fig:rocs}.
+The predictive power of each one of these features is shown in form of the Receiving Operating Characteristic (ROC) curves in Figure~\ref{fig:rocs}.
 The features are ordered according to the area under the curve (AUC).

### 2015-12-30 imaging-paper.Rnw
took out table to save some space

@@
 Should we contact Alan Zheng to see whether we can get access to more bullet files?
+\item add: we could extend the results of the James Hamby study by trying to match every single land against all other lands.
+This would almost double the number of comparisons ($21,115 = (118+88)\cdot 205/2$).
+Something for the appendix?
+What would be fun, would be to show all of the aligned signatures for all ten barrels.
 \item add: should we add a paragraph on the speed of the evaluation?
@@
 \item change: we could make the table into a labelled dotplot of importance to save space. We don't need the mean and standard deviations anymore -- the density plots and the ROCs take care of that.
-\item change: the focus of the paper is initially quite strongly on the maximal consecutively matching striae (CMS).
+%\item change: the focus of the paper is initially quite strongly on the maximal consecutively matching striae (CMS).
 We need to emphasis the role of CMS in the introduction: max CMS is Biasotti's main criterium for matching.
@@
 We would need somebody qualified to tell us how many CMS there are, though.
-\item to consider: AOAS states that accepted papers rarely have more than 20 pages. We are currently at about 23--24 pages.
+\item to consider: AOAS states that accepted papers rarely have more than 20 pages. We are currently over by a couple of pages.
 \item definitions: 
@@
 The primary goal in this type of forensic analysis is to identify these striation markings in an attempt to match two bullets.
-Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help us determine whether two bullets are a match.
-However, this process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
-We propose a framework which allows for the analysis of the surface topologies, and the transcription of the markings into a 2D plotting framework which we believe makes identification of matches easier and more rooted in science for both the expert and the lay person.
-
-\hh{XXX More detailed discussion of automatic approaches and database discussion XXX}
+Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help in determining whether two bullets are a match.
+This process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
+
+\hh{
+Current standard practice~\cite{afte:1992} relies in part on the assessment of the so-called maximal number of consecutively matching striae (CMS), first defined by \citet{biasotti:1959}.
+One of the problems is that a human inspection to determine CMS is subjective \citep{miller:1998}. 
+}
+
+In this paper, we propose a framework which allows for the automatic analysis of the surface topologies of bullets, and the transcription of the individual characteristics into a 2D plotting framework.
+This allows for an objective and quantitative assessment of striae-based bullet matches.
+
+%\hh{XXX More detailed discussion of automatic approaches and database discussion XXX}
 
@@
   \centering
-<<tree:forest, echo=FALSE, fig.width=5, fig.height=5.25, out.width='.5\\textwidth', warning = FALSE>>=
-qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
-      shape=match) + theme_bw() +
-  theme(legend.position="bottom") +
-  geom_vline(xintercept=0.5, colour="grey50") +
-  geom_hline(yintercept=0.5, colour="grey50") + 
-  xlab("Predictions from decision tree") +
-  ylab("Predictions from random forest")   +
+<<tree:forest, echo=FALSE, fig.width=5, fig.height=2.5, out.width='.6\\textwidth', warning = FALSE>>=
+predm <- melt(bstats, measure.var=c("pred", "forest"))
+levels(predm$variable) <- c("Tree", "Forest")
+predm$match <- c("KNM", "KM")[as.numeric(predm$match)+1]
+qplot(data=predm, value, variable, geom="jitter", colour=match, shape=match) + theme_bw() + facet_grid(match~., labeller="label_both") +
     scale_colour_brewer(palette="Paired", labels=c("KNM", "KM")) +
-  scale_shape_discrete(labels=c("KNM", "KM"))
-@
-\caption{\label{fig:tree-forest}Scatterplot of prediction results from the tree (horizontal) and the forest (vertical).
-Each point corresponds to a land-to-land comparison.
-Known matches are indicated in blue.
-Matches predicted by the decision tree are on the right half of the plot, matches predicted by the random forest are in the upper half.
+  scale_shape_discrete(labels=c("KNM", "KM")) + ylab("") + 
+  xlab("Predicted probability of a match") +
+  theme(legend.position="none")
+
+# qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
+#       shape=match) + theme_bw() +
+#   theme(legend.position="bottom") +
+#   geom_vline(xintercept=0.5, colour="grey50") +
+#   geom_hline(yintercept=0.5, colour="grey50") + 
+#   xlab("Predictions from decision tree") +
+#   ylab("Predictions from random forest")   +
+#   scale_colour_brewer(palette="Paired", labels=c("KNM", "KM")) +
+#   scale_shape_discrete(labels=c("KNM", "KM"))
+# 
+@
+\caption{\label{fig:tree-forest}Prediction results from the tree and the forest.
+Using a cut-off probability of 0.5 the forest correctly predicts every single comparison.
+Compared to the tree, the forest's prediction probabilities are  shrunk towards either end of the prediction range.
 }
@@
 Besides a probabilistic quantification of matches, we get from a random forest an assessment of the importance of each of the features derived from the bullets' 3d topological surface measurements.
-Table~\ref{tab:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree.
+Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree.
+\begin{figure}
+<<echo=FALSE, fig.width=7, fig.height=1.5, out.width='.7\\textwidth'>>=
+imp <- data.frame(importance(rtrees))
+imp$Variable <- row.names(imp)
+imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
+imp1 <- subset(imp, Variable != c("x1", "sd.D"))
+imp2 <- subset(imp, Variable %in% c("x1", "sd.D"))
+qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
+  geom_text(aes(y=1.1, label=Variable), angle=45, hjust=0, data=imp1) + 
+  geom_text(aes(y=0.9, label=Variable), angle=-45, hjust=0, data=imp2) + 
+  ylim(c(0.5, 2.5)) + theme_bw() + ylab("") +
+  theme(axis.ticks.y=element_blank(), axis.text.y=element_blank(),
+        panel.grid.major.y=element_blank(), panel.grid.minor.y=element_blank()) +
+  xlab("Importance (mean decrease in Gini index)")
+@
+\caption{\label{fig:importance}Importance of features in the random forest. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.}
+\end{figure}
 \begin{table}[tbhp]
@@
 
-The rows in the table are ordered according to descending importance.
-The variable with the most predictive power by far is cross-correlation, followed by the number of matches and non-matches.
+%The rows in the table are ordered according to descending importance. 
+The variables with the most predictive power by far are cross-correlation and the overall number of matches, followed by the total depth of joint striations $S$ and non-matches.
 CMS is found only in sixth place.
@@
 By suggesting an automated algorithm that first removes class characteristics, such as the grooves and the curvature of the bullet to reveal the region of the  land, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
-Counting CMS in particular, has been found to be subjective to the examiner: `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.
+In particular, `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.
 
@@
 
-Only a much larger database and transparent communication of results will give us  answers to whether automatic matching of ballistic evidence is really feasible.
-Whether there really is individualization of this evidence, as is often silently assumed.
+Only a much larger database and transparent communication of results will give us  answers to whether automatic matching of ballistic evidence is really feasible,  whether there really is individualization of this evidence, as is often silently assumed.
 And in the case, that individualization is a reasonable assumption, whether  there is a way to be able to distinguish between individual and subclass characteristics.

### 2015-12-30 imaging-paper.Rnw
dotplot instead of table

@@
 imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
-imp1 <- subset(imp, Variable != c("x1", "sd.D"))
+imp1 <- subset(imp, !(Variable %in% c("x1", "sd.D")))
 imp2 <- subset(imp, Variable %in% c("x1", "sd.D"))
@@
   geom_text(aes(y=0.9, label=Variable), angle=-45, hjust=0, data=imp2) + 
-  ylim(c(0.5, 2.5)) + theme_bw() + ylab("") +
+  ylim(c(0, 2.75)) + theme_bw() + ylab("") +
   theme(axis.ticks.y=element_blank(), axis.text.y=element_blank(),
@@
 \end{figure}
-\begin{table}[tbhp]
-\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches.
+%' \begin{table}[tbhp]
+%' \caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches.
 Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
 Averages (and standard deviations) for known matches (KM) and known non-matches (KNM) are shown in the last four columns.}
-\centering
-<<importance, echo=FALSE, results='asis', warning=FALSE>>=
-imp <- data.frame(importance(rtrees))
-imp$Variable <- row.names(imp)
-imp <- imp[order(-imp$MeanDecreaseGini),]
-names(imp)[1] <- "Importance"
-imp <- imp[,2:1]
-imp$meanMatch <- sapply(imp$Variable, function(var) {
-  mean(subset(bstats, match)[,var])
-})
-imp$sdMatch <- sapply(imp$Variable, function(var) {
-  sd(subset(bstats, match)[,var])
-})
-imp$meanNonMatch <- sapply(imp$Variable, function(var) {
-  mean(subset(bstats, !match)[,var])
-})
-imp$sdNonMatch <- sapply(imp$Variable, function(var) {
-  sd(subset(bstats, !match)[,var])
-})
-imp$meanMatch <- round(imp$meanMatch,1)
-imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
-imp$meanNonMatch <- round(imp$meanNonMatch,1)
-imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
-row.names(imp) <- 1:nrow(imp)
-imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
-names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
-print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
-      floating = FALSE)
-@
-\end{table}
+%' \centering
+%' <<importance, echo=FALSE, results='asis', warning=FALSE>>=
+%' imp <- data.frame(importance(rtrees))
+%' imp$Variable <- row.names(imp)
+%' imp <- imp[order(-imp$MeanDecreaseGini),]
+%' names(imp)[1] <- "Importance"
+%' imp <- imp[,2:1]
+%' imp$meanMatch <- sapply(imp$Variable, function(var) {
+%'   mean(subset(bstats, match)[,var])
+%' })
+%' imp$sdMatch <- sapply(imp$Variable, function(var) {
+%'   sd(subset(bstats, match)[,var])
+%' })
+%' imp$meanNonMatch <- sapply(imp$Variable, function(var) {
+%'   mean(subset(bstats, !match)[,var])
+%' })
+%' imp$sdNonMatch <- sapply(imp$Variable, function(var) {
+%'   sd(subset(bstats, !match)[,var])
+%' })
+%' imp$meanMatch <- round(imp$meanMatch,1)
+%' imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
+%' imp$meanNonMatch <- round(imp$meanNonMatch,1)
+%' imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
+%' row.names(imp) <- 1:nrow(imp)
+%' imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
+%' names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
+%' print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
+%'       floating = FALSE)
+%' @
+%' \end{table}
 

### 2015-12-31 imaging-paper.Rnw
cleaning up the last pieces of content

@@
 \usepackage{subcaption}
-\usepackage{lineno,hyperref}
+\usepackage{lineno}
 \modulolinenumbers[5]
+
+\usepackage{xr}
+\externaldocument{imaging-supplement}
+
 \usepackage{dcolumn}
@@
 %   Parameters for FLOAT pages (not text pages):
-\renewcommand{\floatpagefraction}{0.7}	% require fuller float pages
+\renewcommand{\floatpagefraction}{0.98}	% require fuller float pages
 % N.B.: floatpagefraction MUST be less than topfraction !!
-\renewcommand{\dblfloatpagefraction}{0.7}	% require fuller float pages
+\renewcommand{\dblfloatpagefraction}{0.98}	% require fuller float pages
 
@@
 \title{Automatic Matching of Bullet Lands}\thanksref{T1}
-\thankstext{T1}{Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}
+\thankstext{T1}{Resources for this manuscript, including R code to reproduce all of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}
 \runtitle{Automatic Matching of Bullet Lands}
@@
 \linenumbers
-\tableofcontents
-\newpage
+%\tableofcontents
+%\newpage
 <<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
@@
 %\item add: more discussion of the density plots: probability ratio for a match - exchanged probability ratios for ROC curves
-\item add: more discussion on additional bullet images.
+%\item add: more discussion on additional bullet images.
 NIST database claims to have 70 bullets -- that would be two sets of the study, but only 35 bullets are currently available.
 Should we contact Alan Zheng to see whether we can get access to more bullet files?
-\item add: we could extend the results of the James Hamby study by trying to match every single land against all other lands.
+Not enough space for this one.
+%\item add: we could extend the results of the James Hamby study by trying to match every single land against all other lands.
 This would almost double the number of comparisons ($21,115 = (118+88)\cdot 205/2$).
@@
 What would be fun, would be to show all of the aligned signatures for all ten barrels.
+(moved to supplement)
 \item add: should we add a paragraph on the speed of the evaluation?
@@
 If we only do bullet matching, we could speed this process up by eliminating some land comparisons.
-\item add: we might need a proper definition of the joint striation marks (the result of striation-identify)
-\item change: we could make the table into a labelled dotplot of importance to save space. We don't need the mean and standard deviations anymore -- the density plots and the ROCs take care of that.
+%\item add: we might need a proper definition of the joint striation marks (the result of striation-identify)
+%\item change: we could make the table into a labelled dotplot of importance to save space. We don't need the mean and standard deviations anymore -- the density plots and the ROCs take care of that.
 %\item change: the focus of the paper is initially quite strongly on the maximal consecutively matching striae (CMS).
@@
 \item to consider: AOAS states that accepted papers rarely have more than 20 pages. We are currently over by a couple of pages.
-\item definitions: 
-A bullet's {\it signature} is defined as the residual surface measurements after removing the big structure using the loess fit.
+%\item definitions: 
+%A bullet's {\it signature} is defined as the residual surface measurements after removing the big structure using the loess fit.
 The {\it cross section} defines the location along the $x$ axis  at which the signature is taken.
@@
 
-\hh{In an approach to analyze the striation pattern we extract a {\emph bullet profile} \cite{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.}
+\hh{In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.}
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 \hh{While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of a circular fit and the remaining residual structure can be found in supplement~\ref{supp:cylindrical}.}
+A discussion of a circular fit and the remaining residual structure can be found in supplement~\ref{supp:circular}.}
 %It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
@@
 \caption{\label{fig:loess_resid} Residuals of loess fit for bullet~1-5.
-This defines the {\it signature} of bullet~1-5. Blue areas show  95\% bootstrap confidence intervals (based on $B=1,000$ bootstrap samples).
 }
@@
 \end{subfigure}
-\caption{Fit and residuals of a loess fit to bullet~1-5 (Barrel~1).}
+\caption{Fit and residuals of a loess fit to bullet~1-5 (Barrel~1).
+The residuals define the {\it signature} of bullet~1-5.
+Blue areas around the residuals show  95\% bootstrap confidence intervals (based on $B=1,000$ bootstrap samples).
+}
 \end{figure}
@@
 We then define an interval around the location of the extrema on each side as one third of the distance to the location of the next extrema (see Figure~\ref{fig:smoothcutb}).
+Peaks and valleys constitute the \emph{striation marks} of the bullet.
 \item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify joint intervals between two signatures as those areas, in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals.
@@
 On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be not necessary in the Hamby study, because, here, the results from the random forest are very clear cut.
-Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis).
+Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree and the random forest.
 %\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match.
@@
 \begin{figure}
-<<echo=FALSE, fig.width=7, fig.height=1.5, out.width='.7\\textwidth'>>=
+<<echo=FALSE, fig.width=7, fig.height=2, out.width='.7\\textwidth'>>=
 imp <- data.frame(importance(rtrees))
@@
 qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
-  geom_text(aes(y=1.1, label=Variable), angle=45, hjust=0, data=imp1) + 
-  geom_text(aes(y=0.9, label=Variable), angle=-45, hjust=0, data=imp2) + 
+  geom_text(aes(y=1.2, label=Variable), angle=45, hjust=0, data=imp1) + 
+  geom_text(aes(y=0.8, label=Variable), angle=-45, hjust=0, data=imp2) + 
   ylim(c(0, 2.75)) + theme_bw() + ylab("") +
@@
 We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features.
-Finally, we have provided an easy-to-use open-source backend, and web frontend to many of these features.
+%Finally, we have provided an easy-to-use open-source backend, and web frontend to many of these features.
 

### 2015-12-31 imaging-supplement.Rnw
cleaning up the last pieces of content

@@
 \usepackage{subcaption}
-\usepackage{lineno,hyperref}
+\usepackage{lineno}
 \modulolinenumbers[5]
@@
 
+\newpage
+\section{One-on-one comparison of Hamby study}\label{supp:extended}
+\hh{
+add: we could extend the results of the James Hamby study by trying to match every single land against all other lands.
+This would almost double the number of comparisons ($21,115 = (118+88)\cdot 205/2$).
+What would be fun, would be to show all of the aligned signatures for all ten barrels.
+(moved to supplement)
+}
 \bibliographystyle{imsart-nameyear.bst}

### 2015-12-31 imaging-supplement.Rnw
including the modified functions into the old ones.

@@
 
-new_get_bullet <- function(path, x = 243.75, bullet = NULL) {
-  if (is.null(bullet)) bullet <- read.x3p(path)
-  dbr111 <- fortify_x3p(bullet)
-
-  pickx <- dbr111$x[which.min(abs(x - unique(dbr111$x)))]
-
-  dbr111.fixx <- dbr111[dbr111$x == pickx,]
-
-  return(dbr111.fixx)
-}
-
 new_processBullets <- function(bullet, x = 100) {
@@
   list_of_fits <- lapply(crosscuts, function(x) {
-    br111 <- new_get_bullet(path = NULL, x = x, bullet = bullet)
+    br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
     br111.groove <- get_grooves(br111)
@@
 This would almost double the number of comparisons ($21,115 = (118+88)\cdot 205/2$).
-What would be fun, would be to show all of the aligned signatures for all ten barrels.
-(moved to supplement)
 }
+\hh{Show aligned signatures for all ten barrels. }
+<<aligned, echo=FALSE>>=
+matches <- read.csv("csvs/matches.csv", header = FALSE)
+matches$barrel <- rep(1:10, each=6)
+matches$id <- 1:nrow(matches)
+library(reshape2)
+mm <- melt(matches, id.var=c("barrel", "id"), na.rm=TRUE)
+mm <- subset(mm, value != "")
+mm$value <- gsub("Br. Bullet ", "", mm$value)
+
+@
 \bibliographystyle{imsart-nameyear.bst}

### 2015-12-31 code/complete-hamby.R

@@ -12,16 +12,16 @@ unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 ###############
 # can we identify the barrels the unknown bullets came from?
 
-new_get_bullet <- function(path, x = 243.75, bullet = NULL) {
-    if (is.null(bullet)) bullet <- read.x3p(path)
-    dbr111 <- fortify_x3p(bullet)
-    
-    pickx <- dbr111$x[which.min(abs(x - unique(dbr111$x)))]
-    
-    dbr111.fixx <- dbr111[dbr111$x == pickx,]
-    
-    return(dbr111.fixx)
-}
+# new_get_bullet <- function(path, x = 243.75, bullet = NULL) {
+#     if (is.null(bullet)) bullet <- read.x3p(path)
+#     dbr111 <- fortify_x3p(bullet)
+#     
+#     pickx <- dbr111$x[which.min(abs(x - unique(dbr111$x)))]
+#     
+#     dbr111.fixx <- dbr111[dbr111$x == pickx,]
+#     
+#     return(dbr111.fixx)
+# }
 
 new_processBullets <- function(bullet, x = 100) {
     crosscuts <- unique(fortify_x3p(bullet)$x)
@@ -30,7 +30,7 @@ new_processBullets <- function(bullet, x = 100) {
     if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]
     
     list_of_fits <- lapply(crosscuts, function(x) {
-        br111 <- new_get_bullet(path = NULL, x = x, bullet = bullet)
+        br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
         br111.groove <- get_grooves(br111)
         fit_loess(br111, br111.groove)$resid$data
     })

### 2015-12-31 imaging-paper.Rnw
Small typo

@@
 library(sm)
-
 @

### 2015-12-31 imaging-supplement.Rnw
first overview of unaligned signatures

@@
 \hh{Show aligned signatures for all ten barrels. }
-<<aligned, echo=FALSE>>=
+<<aligned, echo=FALSE, warning=FALSE, fig.width=8, fig.height=10, out.width='\\textwidth', cache=TRUE>>=
 matches <- read.csv("csvs/matches.csv", header = FALSE)
@@
 mm <- subset(mm, value != "")
-mm$value <- gsub("Br. Bullet ", "", mm$value)
+mm$value <- gsub("Br.* Bullet ", "", mm$value)
+mm$prefix <- sprintf("Br%d Bullet ", mm$barrel)
+unkns <- grep("[A-Z]",mm$value)
+mm$prefix[unkns] <- "Ukn Bullet "
+mm$path <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P1of2//"
+mm$path[unkns] <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P2of2//"
+mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))
 
+ccs <- read.csv("csvs/crosscuts-25-new.csv")
+mm <- merge(mm, ccs, by="path", all=TRUE)
+mm <- subset(mm, !is.na(mm$cc))
+
+crosscuts <- plyr::ldply(
+  1:nrow(mm), 
+  function(i) {
+    dframe <- get_crosscut(mm$path[i], mm$cc[i])
+    groove <- get_grooves(dframe)
+    data.frame(mm[i,], fit_loess(dframe, groove)$resid$data)
+  })
+qplot(x=y, y=resid, data=crosscuts, geom="line", group=path) + ylim(c(-7,7)) + facet_wrap(facets=~id, ncol=6) + theme_bw() 
 @

### 2016-01-01 imaging-supplement.Rnw
aligned signatures of all 60 lands

--- 2015-12-31 imaging-supplement.Rnw
@@
 add: we could extend the results of the James Hamby study by trying to match every single land against all other lands.
-This would almost double the number of comparisons ($21,115 = (118+88)\cdot 205/2$).
+This  almost doubles the number of comparisons ($21,115 = (118+88)\cdot 205/2$).
 }
 \hh{Show aligned signatures for all ten barrels. }
-<<aligned, echo=FALSE>>=
+
+Figure~\ref{fig:aligned} gives an overview of all the signatures from bullet lands in the Hamby study aligned by barrel.
+Three to five bullets were fired from each barrel.
+The figure gives us both some insight into how well signatures match and how consistent individual characteristics are impregnated on bullets fired from each of the barrels.
+Signatures for some lands match remarkably well - such as land 5 from barrel 1, whereas all lands from barrel 5 show some variability both in the location and depths of peaks and valleys.
+\begin{figure}[hbtp]
+<<aligned, echo=FALSE, warning=FALSE, fig.width=12, fig.height=15, out.width='\\textwidth', cache=TRUE>>=
 matches <- read.csv("csvs/matches.csv", header = FALSE)
@@
 mm <- subset(mm, value != "")
-mm$value <- gsub("Br. Bullet ", "", mm$value)
+mm$value <- gsub("Br.* Bullet ", "", mm$value)
+mm$prefix <- sprintf("Br%d Bullet ", mm$barrel)
+unkns <- grep("[A-Z]",mm$value)
+mm$prefix[unkns] <- "Ukn Bullet "
+mm$path <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P1of2//"
+mm$path[unkns] <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P2of2//"
+mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))
 
+ccs <- read.csv("csvs/crosscuts-25-new.csv")
+mm <- merge(mm, ccs, by="path", all=TRUE)
+mm <- subset(mm, !is.na(mm$cc))
+
+crosscuts <- plyr::ldply(
+  1:nrow(mm), 
+  function(i) {
+    dframe <- get_crosscut(mm$path[i], mm$cc[i])
+    groove <- get_grooves(dframe)
+    data.frame(mm[i,], fit_loess(dframe, groove)$resid$data)
+  })
+
+crossSmooth <- bulletSmooth(crosscuts)
+
+alignme <- function(data) {
+#  browser()
+  data$bullet <- as.character(data$bullet)
+  bullets <- unique(data$bullet)
+  b1 <- bullets[1]
+  lofs <- NULL
+  for (i in bullets[-1]) {
+    lof <- subset(data, bullet %in% c(b1, i))
+    lofs <- rbind(lofs, bulletAlign(lof)$bullet)
+  }
+  lofs <- unique(lofs)
+  lofs
+}
+
+aligned <- crossSmooth %>% group_by(id) %>% do(
+  data.frame(alignme(.))
+)
+aligned$land <- (aligned$id %% 6)
+aligned$land[aligned$land==0] <- 6
+qplot(x=y, y=l30, data=aligned, geom="line", group=bullet) + ylim(c(-7,7)) + facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
+  xlab("") + ylab("Signature")
 @
+\caption{\label{fig:aligned}Overview of aligned signatures for all bullet lands for the ten barrels of the Hamby study. }
+\end{figure}
 \bibliographystyle{imsart-nameyear.bst}

### 2016-01-01 imaging-paper.Rnw
colour taken out of main document

@@
 Identifying these striation markings in an attempt to match two bullets is the primary goal.
-We propose a framework for the analysis of the 3D surface measurements which transcribes the markings into a 2D plotting framework.
-This makes identification of matches easier and more rooted in science for both the expert and the lay person.
-The automatic matching routine we propose \hh{manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches.}
+We propose an automated framework for the analysis of the 3D surface measurements, that first transcribes the markings into a 2D plotting framework.
+This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels.
+The automatic matching routine we propose manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
 \end{abstract}
@@
 
-\hh{
 Current standard practice~\cite{afte:1992} relies in part on the assessment of the so-called maximal number of consecutively matching striae (CMS), first defined by \citet{biasotti:1959}.
 One of the problems is that a human inspection to determine CMS is subjective \citep{miller:1998}.
-}
 
@@
 
-\hh{The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. }
-\hh{The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
+The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. 
+The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
 % Hamby study: automated systems were successfully used for matching:
@@
 % BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)
- }
-\hh{A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}.
+
+A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}.
 This study is based on another set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score.
 The authors state that this process `could be automated', but no implementation of the algorithm is available.
-}
 
@@
 
-\hh{In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.}
+In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 The smooth curve on the plot represents a segment of a perfect circle with the same radius as the bullet.
-\hh{While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of a circular fit and the remaining residual structure can be found in supplement~\ref{supp:circular}.}
+While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
+A discussion of a circular fit and the remaining residual structure can be found in supplement~\ref{supp:circular}.
 %It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
@@
 %\subsection{Loess Fit}
-\hh{Instead of a circular fit, we use multiple loess fits to model the overall structure in order to identify the bullet markings. }
+Instead of a circular fit, we use multiple loess fits to model the overall structure in order to identify the bullet markings.
 
@@
 %In an attempt to improve upon the previous algorithm, w
-We first have to \hh{identify} the \hh{locations of the left and right}  groove in the image.
-\hh{The grooves are not assumed to contain information relevant for determining matches, they also dominate the structure, and therefore need  to be removed.}
+We first have to identify the locations of the left and right  groove in the image.
+The grooves are not assumed to contain information relevant for determining matches, they also dominate the structure, and therefore need  to be removed.
 Fortunately, the location and appearance of the grooves in the surface profiles is quite consistent.
@@
 \begin{enumerate}
-    \item Fix an $x$ value \hh{to extract a bullet profile} (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
+    \item Fix an $x$ value to extract a bullet profile (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
     \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \emph{smoothing factor} $s$.
@@
 Next, we fit a loess regression to the data.
-\hh{Loess regression \citep{cleveland:1979} is based on the assumption that the relationship between two random variables $X$ and $Y$ can be described in form of a smooth, continuous function $f$ with $y_i = f(x_i) + \varepsilon_i$ for all values $i = 1, ..., n$.
+Loess regression \citep{cleveland:1979} is based on the assumption that the relationship between two random variables $X$ and $Y$ can be described in form of a smooth, continuous function $f$ with $y_i = f(x_i) + \varepsilon_i$ for all values $i = 1, ..., n$.
 The function $f$ is estimated in form of  locally weighted polynomial regressions.
 Parameters of the estimation are $\alpha$, the proportion of all points included in the fit (here, $\alpha = 0.75$), the weighting function and the degree of the polynomial (here, we fit a quadratic regression).
-}
-
-\hh{The main idea of locally weighted regression is to use a weighting routine that  emphasizes the effect of points close by and de-emphasizes the effect of points as they are further away.
+
+The main idea of locally weighted regression is to use a weighting routine that  emphasizes the effect of points close by and de-emphasizes the effect of points as they are further away.
 The weighting function used here, is the tricubic function $w(d) = \left(1 - d^3\right)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
 $d$ is defined as the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
-}
 
@@
 The fit seems to do a reasonable job of capturing the structure of the image.
-\hh{The blue area around the signature are  95\% bootstrapped confidence intervals based on 1,000 bootstrap samples.
-}
+The blue area around the signature are  95\% bootstrapped confidence intervals based on 1,000 bootstrap samples.
 Figure~\ref{fig:loess_resid} shows the residuals from this fit. These residuals are called the \emph{signature} of bullet~1-5.
@@
 Signatures of bullet~1 are shown on the left (all cross sections with values below 100$\mu m$) and cross sections of bullets 2 are shown on the right (cross sections with values above 100$\mu m$).
-\hh{Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one side to the other.
+Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one side to the other.
 Visually, this allows for an easy assessment of these two bullet lands as a match.
@@
 The goal of this section is to eliminate the need for a visual inspection during the matching process and replace it by an automatic algorithm.
-This also allows for a quantification of the quality of the match.}
+This also allows for a quantification of the quality of the match.
 
@@
 
-\hh{In this section, we are going to describe the algorithm for matching signatures first, and discuss the impact of paramater choices in the subsections thereafter.}
+In this section, we are going to describe the algorithm for matching signatures first, and discuss the impact of paramater choices in the subsections thereafter.
 
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet \hh{(see supplement~\ref{supp:bulletbottom} for a more detailed discussion)}.
+We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see supplement~\ref{supp:bulletbottom} for a more detailed discussion).
 
@@
 The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
-This poses both a caveat to matching attempts as well as an opportunity \hh{for quality control}: we have to be aware of which cross section level was used in a matching.
+This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of which cross section level was used in a matching.
 An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement~\ref{supp:ccf} for a more thorough discussion).
@@
 %An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
-by comparing signatures from cross sections that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off \hh{or some other damage}, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
+by comparing signatures from cross sections that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
 In the approach here, we keep increasing the cross section until we find a section with a stable pattern.
@@
 %We can investigate the performance of peak/valley detection across different smoothing factors.
+\hh{XXX the smoothing factor needs a few more details: a smoothing factor of 5 means that we consider the two measurements on the left and the right of the observations under consideration XXX}
 Figure~\ref{fig:varysmooth} displays the  peaks and valleys detected in the same signature at smoothing factors of 5, 25, and 45, respectively.
@@
 Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
-\hh{XXX shouldn't we also make this dependent on the resolution of the equipment?
-XXX}
+Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artefacts in the analysis.
 
@@
 \section{Evaluation}
-\hh{In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
-}
-\hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
+In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
+As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
 For this, we assess cross-correlation of signatures at cross sections 25$\mu m$ apart.
@@
 Four lands from different bullets are flagged as problematic in this respect.
-A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface, also known as `tank rash'.}
+A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface, also known as `tank rash'.
 %
@@
 \end{subfigure}
-\caption{\label{fig:fourflags}\hh{Images of the four lands that got flagged during the quality assessment.
+\caption{\label{fig:fourflags}Images of the four lands that got flagged during the quality assessment.
 All of them show scratch marks across the striation marks from the barrel.
-They will be excluded from the remainder of the analysis.}}
+They will be excluded from the remainder of the analysis.}
 \end{figure}
 %
-\hh{We exclude these four lands from the further matching considerations and 
+We exclude these four lands from the further matching considerations and 
  run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e.
@@
 Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
-}
  When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 known lands.
@@
 This is one of the benefits of switching to a digitalized version of the images that preserves the 3d surface structure.
-\hh{The emphasis on cross-correlation is similar to the results of the study by \citet{ma:2004}.
-However, in this study authors did not consider the number of matches and non-matches.}
+The emphasis on cross-correlation is similar to the results of the study by \citet{ma:2004}.
+However, in this study authors did not consider the number of matches and non-matches.
 
@@
 This is a weighted average between a zero false positive error rate and an error rate of false negatives of 0.23.
-\hh{This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).}
+This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).
 Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
@@
 %' %
-\hh{This makes the errors in the automated approach smaller than the human error in the Hamby study.
-Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.}
+This makes the errors in the automated approach smaller than the human error in the Hamby study.
+Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.
 

### 2016-01-01 imaging-supplement.Rnw
colour taken out of main document

@@
 qplot(x=y, y=l30, data=aligned, geom="line", group=bullet) + ylim(c(-7,7)) + facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
-  xlab("") + ylab("Signature")
+  xlab("") + ylab("Signature") + 
+  theme(plot.margin=unit(c(0,0,0,0), unit="cm"))
 @

### 2016-01-01 imaging-paper.Rnw
missing link

@@
 The NRC report also found, that results from 2d images are improved upon by matches based on 3d images.
-This is consistent with the importance of features found here: out of the top five features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.
+This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the number of matches and mismatches are available for a match based on 2d features.
 

### 2016-01-01 imaging-paper.Rnw
speed consideration

@@
 If we only do bullet matching, we could speed this process up by eliminating some land comparisons.
+- Thinking about this more, I'm against it.
+The speed massively depends on the kind of computer used - the important thing is, that on a server all of these calculations are very fast -- even my machine can do 11 comparisons per second, if the files are pre-processed.
+Pre-processing might take a couple of seconds - at most 20-30 seconds for each bullet land.
+That is an initial cost one has to pay.
 %\item add: we might need a proper definition of the joint striation marks (the result of striation-identify)

### 2016-01-03 code/complete-hamby.R
moved last function into the package

@@ -39,35 +39,35 @@ unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 #     data.frame(lof, bullet = bullet$path, stringsAsFactors = FALSE)
 # }
 
-new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
-    lof <- rbind(lof1, lof2)
-    bAlign = bulletAlign(lof)
-    lofX <- bAlign$bullet  
-    
-    b12 <- unique(lof$bullet)
-    peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
-    peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)
-    
-    #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
-    #    theme_bw() +
-    #    geom_rect(data=peaks1$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25)) +
-    #    geom_rect(data=peaks2$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25))
-    
-    peaks1$lines$bullet <- b12[1]
-    peaks2$lines$bullet <- b12[2]
-    
-    lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-    
-    #   p <- qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
-    #     theme_bw() +
-    #     geom_rect(data=lines, aes(xmin=xmin, xmax=xmax, fill = factor(type)),  ymin=-6, ymax=6, inherit.aes = FALSE, alpha=I(0.25)) +
-    #     ylim(c(-6,6)) +
-    #     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match), inherit.aes = FALSE) +
-    #     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match), inherit.aes = FALSE) 
-    
-    maxCMS <- maxCMS(lines$match==TRUE)
-    list(maxCMS = maxCMS, ccf = bAlign$ccf, lag=bAlign$lag, lines=lines, bullets=lofX)
-}  
+# new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
[…]

### 2016-01-03 imaging-paper.Rnw
Update to be compatible with new processBullets

@@
 
-lof <- processBullets(paths = images[c(5,7)], x = 100)
-
-subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+#lof <- processBullets(paths = images[c(5,7)], x = 100)
+subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
+subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
+#subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+#subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
 subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
@@
 
-lof <- processBullets(paths = images[c(5,7)], x = 100)
-
-subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
+#lof <- processBullets(paths = images[c(5,7)], x = 100)
+subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
+subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
+#subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
+#subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
 subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
@@
 #images <- file.path(datadir, dir(datadir))
-  lof <- processBullets(paths = images[c(5,7)], x = 100)
+  #lof <- processBullets(paths = images[c(5,7)], x = 100)
+
+  subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
+  subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
+  lof <- rbind(subLOFx1, subLOFx2)
+
   lof <- bulletSmooth(lof)
@@
 crosscuts <- seq(100,250, by=25)
-lof <- processBullets(paths[37], x = crosscuts)
+lof <- processBullets(read.x3p(paths[37]), name = paths[37], x = crosscuts)
 lof$bullet <- paste(lof$bullet, lof$x)

### 2016-01-04 imaging-paper.Rnw
little blurb on smoothing factor

@@
 \subsection{Varying smoothing factor}\label{sec:smoothing}
-As mentioned earlier, the algorithm for detecting peaks and valleys depend on the selection of a smoothing window, called the smoothing factor.
+As mentioned earlier, the algorithm for detecting peaks and valleys depend on the selection of a smoothing window, called the smoothing factor or span.
+A smoothing factor of $k$ means that the  $k$ closest observations to $x_o$ are considered for a fit for $x_o$.
+Because surface measurements are recorded at an equidistand resolution (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$..
 For detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor  of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
 %We can investigate the performance of peak/valley detection across different smoothing factors.
-\hh{XXX the smoothing factor needs a few more details: a smoothing factor of 5 means that we consider the two measurements on the left and the right of the observations under consideration XXX}
+
 Figure~\ref{fig:varysmooth} displays the  peaks and valleys detected in the same signature at smoothing factors of 5, 25, and 45, respectively.

### 2016-01-04 imaging-paper.Rnw
random thing from git

@@
 Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
-Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artefacts in the analysis.
+Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artifacts in the analysis.
 

### 2016-01-05 code/complete-hamby-all.R
Commit what i have so far for complete study

@@ -1,0 +1,107 @@
+library(x3pr)
+library(x3prplus)
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(zoo)
+
+knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
+unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+
+
+###############
+# can we identify the barrels the unknown bullets came from?
+
+# new_get_bullet <- function(path, x = 243.75, bullet = NULL) {
+#     if (is.null(bullet)) bullet <- read.x3p(path)
+#     dbr111 <- fortify_x3p(bullet)
+#     
+#     pickx <- dbr111$x[which.min(abs(x - unique(dbr111$x)))]
+#     
+#     dbr111.fixx <- dbr111[dbr111$x == pickx,]
+#     
+#     return(dbr111.fixx)
+# }
+
+# new_processBullets <- function(bullet, x = 100) {
+#     crosscuts <- unique(fortify_x3p(bullet)$x)
+#     crosscuts <- crosscuts[crosscuts >= min(x)]
+#     crosscuts <- crosscuts[crosscuts <= max(x)]
+#     if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]
+#     
+#     list_of_fits <- lapply(crosscuts, function(x) {
+#         br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
+#         br111.groove <- get_grooves(br111)
+#         fit_loess(br111, br111.groove)$resid$data
+#     })
+#     lof <- list_of_fits %>% bind_rows
+#     
+#     data.frame(lof, bullet = bullet$path, stringsAsFactors = FALSE)
+# }
+
+# new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
+#     lof <- rbind(lof1, lof2)
+#     bAlign = bulletAlign(lof)
+#     lofX <- bAlign$bullet  
+#     
+#     b12 <- unique(lof$bullet)
+#     peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
+#     peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)
+#     
+#     #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
+#     #    theme_bw() +
[…]

### 2016-01-05 code/complete-hamby-all.R
Switch to using only span 25

@@ -83,13 +83,13 @@ knowns <- all_bullets[1:120]
 unknowns <- all_bullets[121:210]
 bullets_processed <- lapply(all_bullets, function(bul) {
     cat("Computing processed bullet", basename(bul$path), "\n")
-    processBullets(bullet = bul, x = ccs$cc[which(ccs$path == bul$path)])
+    processBullets(bullet = bul, name = bul$path, x = ccs$cc[which(ccs$path == bul$path)])
 })
 names(bullets_processed) <- as.character(ccs$path)
 
 bullets_smoothed <- bullets_processed %>% bind_rows %>% bulletSmooth
 
-for (span in c(10, 20, 25, 30, 40)) {
+for (span in c(25)) {
     dataStr <- sprintf("data-new-all-%d-25", span) # using crosscuts-25.csv
     
     if (!file.exists(dataStr)) dir.create(dataStr)
@@ -99,6 +99,8 @@ for (span in c(10, 20, 25, 30, 40)) {
             
             br1 <- filter(bullets_smoothed, bullet == x$path)
             br2 <- filter(bullets_smoothed, bullet == all_bullets[[j]]$path)
+            
+            if (all_bullets[[j]]$path == x$path) br2$bullet <- paste0(br2$bullet, "_2")
             
             bulletGetMaxCMSXXX(br1, br2, span=span)
         })

### 2016-01-05 code/complete-hamby.R

@@ -83,7 +83,7 @@ knowns <- all_bullets[1:120]
 unknowns <- all_bullets[121:210]
 bullets_processed <- lapply(all_bullets, function(bul) {
     cat("Computing processed bullet", basename(bul$path), "\n")
-    processBullets(bullet = bul, x = ccs$cc[which(ccs$path == bul$path)])
+    processBullets(bullet = bul, name = bul$path, x = ccs$cc[which(ccs$path == bul$path)])
 })
 names(bullets_processed) <- as.character(ccs$path)
 

### 2016-01-05 imaging-paper.Rnw
forest-tree image

@@
 predm$match <- c("KNM", "KM")[as.numeric(predm$match)+1]
-qplot(data=predm, value, variable, geom="jitter", colour=match, shape=match) + theme_bw() + facet_grid(match~., labeller="label_both") +
+qplot(data=predm, value, match, geom="jitter", colour=match, shape=match) + theme_bw() + facet_grid(variable~.) +
     scale_colour_brewer(palette="Paired", labels=c("KNM", "KM")) +

### 2016-01-05 code/evaluation-hamby-all.R
Evaluation code for the full data

@@ -1,0 +1,225 @@
+span <- 25
+dataStr <- sprintf("data-new-all-%d-25", span)
+datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
+
+CCFs <- plyr::ldply(datas, function(x) {
+  load(x)
+  cmsdist <- sapply(reslist, function(x) x$maxCMS)
+  ccf <- plyr::ldply(reslist, function(res) {
+    lofX <- res$bullets
+    b12 <- unique(lofX$bullet)
+    
+    subLOFx1 <- subset(lofX, bullet==b12[1])
+    subLOFx2 <- subset(lofX, bullet==b12[2]) 
+    
+    subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
+    subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
+
+    ccf <- ccf(subLOFx1$val, subLOFx2$val, plot = FALSE, lag.max=200, na.action = na.omit)
+    lag <- ccf$lag[which.max(ccf$acf)]
+    incr <- min(diff(sort(unique(subLOFx1$y))))
+    
+    subLOFx1$y <- subLOFx1$y -  lag * incr # amount of shifting should just be lag * y.inc
+    ys <- intersect(subLOFx1$y, subLOFx2$y)
+    idx1 <- which(subLOFx1$y %in% ys)
+    idx2 <- which(subLOFx2$y %in% ys)
+    distr.dist <- mean((subLOFx1$val[idx1] - subLOFx2$val[idx2])^2, na.rm=TRUE)
+    distr.sd <- sd(subLOFx1$val, na.rm=TRUE) + sd(subLOFx2$val, na.rm=TRUE)
+    km <- which(res$lines$match)
+    knm <- which(!res$lines$match)
+    if (length(km) == 0) km <- c(length(knm)+1,0)
+    if (length(knm) == 0) knm <- c(length(km)+1,0)
+# browser()    
+    # feature extraction
+    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), 
+               D=distr.dist, 
+               sd.D = distr.sd,
+               b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
+               num.matches = sum(res$lines$match), 
+               num.mismatches = sum(!res$lines$match), 
+               non_cms = x3prplus::maxCMS(!res$lines$match),
+               left_cms = max(knm[1] - km[1], 0),
[…]

### 2016-01-05 imaging-paper.Rnw
bullet names

@@
     \begin{subfigure}[t]{\textwidth}\centering
-\caption{Loess smooth of signatures  at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
+\caption{Loess smooth of signatures  at cross section $x = 100$ (span is 0.03).\label{fig:smooth}}{%
 <<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
-qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
+lofX$bulletname <- scrubPath(lofX$bullet)
+qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bulletname) +
   geom_line(aes(y=l30), colour="grey70", size=0.5) +
-  facet_grid(bullet~.) +
+  facet_grid(bulletname~.) +
 #  scale_colour_manual("", values=cols) +
@@
 
+  peaks$bulletname <- scrubPath(peaks$bullet)
+  lofX$bulletname <- scrubPath(lofX$bullet)
   ggplot() + theme_bw() +
@@
     theme(legend.position="none") + 
-    facet_grid(bullet~.) +
-    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
+    facet_grid(bulletname~.) +
+    geom_line(aes(x=y, y=l30, group=bulletname), data=lofX) +
   ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
@@
   ylim(c(-6,6.5)) + theme_bw() +
-  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
+  geom_line(data=lofX, aes(x=y, y=l30, group=bulletname, linetype=bulletname)) +
   scale_linetype_discrete("") +
@@
 <<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
-qplot(y, resid, group=bullet, colour=bullet, data=lof, 
+lof$bulletname <- scrubPath(lof$bullet)
+qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
       geom="line") + 
@@
 <<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
-qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
+lofY$bulletname <- scrubPath(lofY$bullet)
+qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
       geom="line") + xlab("y") +
@@
 lof <- processBullets(read.x3p(paths[37]), name = paths[37], x = crosscuts)
-lof$bullet <- paste(lof$bullet, lof$x)
+lof$bullet <- paste(scrubPath(lof$bullet), lof$x)
 

### 2016-01-05 imaging-supplement.Rnw
moved table to supplement

@@
 \modulolinenumbers[5]
+
+\usepackage{dcolumn}
+\newcolumntype{d}[1]{D{.}{.}{#1}}
 
@@
 \end{figure}
+
+\newpage
+\section{Results from Random Forest}\label{supp:randomforest}
+
+<<setup-rtrees, echo=FALSE, message=FALSE>>=
+bstats <- read.csv("data-25-25/bullet-stats.csv", stringsAsFactors = FALSE)
+
+library(randomForest)
+set.seed(20151202)
+includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
+
+rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=300)
+@
+
+\begin{table}[tbhp]
+\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches.
+Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
+Averages (and standard deviations) for known matches (KM) and known non-matches (KNM) are shown in the last four columns.}
+\centering
+<<importance, echo=FALSE, results='asis', warning=FALSE>>=
+imp <- data.frame(importance(rtrees))
+imp$Variable <- row.names(imp)
+imp <- imp[order(-imp$MeanDecreaseGini),]
+names(imp)[1] <- "Importance"
+imp <- imp[,2:1]
+imp$meanMatch <- sapply(imp$Variable, function(var) {
+  mean(subset(bstats, match)[,var])
+})
+imp$sdMatch <- sapply(imp$Variable, function(var) {
+  sd(subset(bstats, match)[,var])
+})
+imp$meanNonMatch <- sapply(imp$Variable, function(var) {
+  mean(subset(bstats, !match)[,var])
+})
+imp$sdNonMatch <- sapply(imp$Variable, function(var) {
+  sd(subset(bstats, !match)[,var])
+})
+imp$meanMatch <- round(imp$meanMatch,1)
+imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
+imp$meanNonMatch <- round(imp$meanNonMatch,1)
+imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
+row.names(imp) <- 1:nrow(imp)
+imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
+names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
+print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
+      floating = FALSE)
+@
+\end{table}
+
 \bibliographystyle{imsart-nameyear.bst}

### 2016-01-05 code/evaluation-hamby-all.R

@@ -107,7 +107,7 @@ CCFs <- CCFs[!dpls,]
 library(rpart)
 library(rpart.plot)
 
-includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "span", "forest"))
+includes <- setdiff(names(CCFs), c("id","b1", "b2", "data", "resID", "id.x", "id.y", "pred", "span", "forest"))
 rp1 <- rpart(match~., data=subset(CCFs, !flagged)[,includes])  # doesn't include cms at all !!!!
 prp(rp1, extra = 101)
 CCFs$pred <- predict(rp1, newdata=CCFs)
@@ -118,6 +118,7 @@ library(randomForest)
 set.seed(20151202)
 rtrees <- randomForest(factor(match)~., data=subset(CCFs, !flagged)[,includes], ntree=300)
 CCFs$forest <- predict(rtrees, newdata=CCFs, type="prob")[,2]
+xtabs(~(forest>0.5)+match, data=CCFs[!CCFs$flagged,])
 
 write.csv(CCFs, file=file.path(dataStr, "bullet-stats-single.csv"), row.names=FALSE)
 

### 2016-01-05 imaging-paper.Rnw
moved table to supplement
error rates for all of hamby

@@
 \externaldocument{imaging-supplement}
-
-\usepackage{dcolumn}
-\newcolumntype{d}[1]{D{.}{.}{#1}}
 
@@
 %
-After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0038.
-This is a weighted average between a zero false positive error rate and an error rate of false negatives of 0.23.
+After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0039.
+This is a weighted average between false positive error rate of 0.0002 and an error rate of false negatives of 0.2267.
 This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).
@@
 \end{figure}
-%' \begin{table}[tbhp]
-%' \caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches.
-Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.
-Averages (and standard deviations) for known matches (KM) and known non-matches (KNM) are shown in the last four columns.}
-%' \centering
-%' <<importance, echo=FALSE, results='asis', warning=FALSE>>=
-%' imp <- data.frame(importance(rtrees))
-%' imp$Variable <- row.names(imp)
-%' imp <- imp[order(-imp$MeanDecreaseGini),]
-%' names(imp)[1] <- "Importance"
-%' imp <- imp[,2:1]
-%' imp$meanMatch <- sapply(imp$Variable, function(var) {
-%'   mean(subset(bstats, match)[,var])
-%' })
-%' imp$sdMatch <- sapply(imp$Variable, function(var) {
-%'   sd(subset(bstats, match)[,var])
-%' })
-%' imp$meanNonMatch <- sapply(imp$Variable, function(var) {
-%'   mean(subset(bstats, !match)[,var])
-%' })
-%' imp$sdNonMatch <- sapply(imp$Variable, function(var) {
-%'   sd(subset(bstats, !match)[,var])
-%' })
-%' imp$meanMatch <- round(imp$meanMatch,1)
-%' imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
-%' imp$meanNonMatch <- round(imp$meanNonMatch,1)
-%' imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
-%' row.names(imp) <- 1:nrow(imp)
-%' imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
-%' names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
-%' print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
-%'       floating = FALSE)
-%' @
-%' \end{table}
 

### 2016-01-05 imaging-supplement.Rnw
error rates for all of hamby

@@
 \newpage
-\section{One-on-one comparison of Hamby study}\label{supp:extended}
-\hh{
-add: we could extend the results of the James Hamby study by trying to match every single land against all other lands.
-This  almost doubles the number of comparisons ($21,115 = (118+88)\cdot 205/2$).
+\section{Complete evaluation of the Hamby study}\label{supp:extended}
+\hh{One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land.
+This effectively doubles the number of comparisons from 10,384 pairwise comparisons of useable bullet lands to 21,115 ($= (118+88)\cdot 205/2$) comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons.
 }
+
+\hh{When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the estimated error rates of 0.2267 and 0.0002.
+}
+
+\hh{However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0.
+The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.}
+<<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=
+knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
+knowns <- dir(path=knowndatadir)
+knowns <- gsub(".x3p", "", knowns)
+
+unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+unknowns <- dir(path=unknowndatadir)
+unknowns <- gsub(".x3p", "", unknowns)
+
+#bstats <- read.csv("data-25-25/bullet-stats.csv")
+#flagged <- c("Br6 Bullet 2-1", "Br9 Bullet 2-4", "Ukn Bullet B-2", "Ukn Bullet Q-4")
+includesVar <- setdiff(names(ballstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "id", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km", "forestsmall",
+                                        "left_cms", "right_cms", "left_noncms", "right_noncms"))
+
+library(randomForest)
+#set.seed(20151202)
+#rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=100)
+
+ballstats <- read.csv("data-new-all-25-25/bullet-stats-single.csv")
+# names(ballstats)[10] <- "num.nonmatches"
+# names(ballstats)[11] <- "CNMS"
+# names(ballstats)[16] <- "S"
+# names(ballstats)[17] <- "CMS"
+#ballstats$forestsmall <- predict(rtrees, newdata=ballstats, type="prob")[,2]
+###########
+
+ballstats$insmall <- with(ballstats, ((b1 %in% knowns) & (b2 %in% unknowns)) | ((b2 %in% knowns) & (b1 %in% unknowns)))
+
+set.seed(20160512)
+rtrees2 <- randomForest(factor(match)~., data=subset(ballstats, !flagged & insmall)[,includesVar], ntree=300)
+
+ballstats$forestsmall <- predict(rtrees2, newdata=ballstats, type="prob")[,2]
+
+# xtabs(~(forestsmall>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
+
+set.seed(20151201+i); i+1
+rtrees3 <- randomForest(factor(match)~., data=subset(ballstats, !flagged)[,includesVar], ntree=300); rtrees3
+ballstats$forest <- predict(rtrees3, newdata=ballstats, type="prob")[,2]
+xtabs(~(forest>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
+qplot(forest, data=subset(ballstats, !flagged), geom="jitter", y=match)
+@
+% 9/(10628+9)
+% 18/(76+18)
+
 \hh{Show aligned signatures for all ten barrels. }
@@
 \newpage
-\section{Results from Random Forest}\label{supp:randomforest}
+\section{Table of feature importance}\label{supp:randomforest}
 

### 2016-01-05 imaging-supplement.Rnw
now without the plot

@@
 ballstats$forest <- predict(rtrees3, newdata=ballstats, type="prob")[,2]
-xtabs(~(forest>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
-qplot(forest, data=subset(ballstats, !flagged), geom="jitter", y=match)
+#xtabs(~(forest>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
+#qplot(forest, data=subset(ballstats, !flagged), geom="jitter", y=match)
 @

### 2016-01-05 imaging-supplement.Rnw
errors fixed in last code chunk

--- 2016-01-05 imaging-supplement.Rnw
@@
 
-set.seed(20151201+i); i+1
-rtrees3 <- randomForest(factor(match)~., data=subset(ballstats, !flagged)[,includesVar], ntree=300); rtrees3
+set.seed(20151201)
+rtrees3 <- randomForest(factor(match)~., data=subset(ballstats, !flagged)[,includesVar], ntree=300)
 ballstats$forest <- predict(rtrees3, newdata=ballstats, type="prob")[,2]
-xtabs(~(forest>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
-qplot(forest, data=subset(ballstats, !flagged), geom="jitter", y=match)
+#xtabs(~(forest>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
+#qplot(forest, data=subset(ballstats, !flagged), geom="jitter", y=match)
 @

### 2016-01-06 imaging-supplement.Rnw
reference to file corrected

@@
 #flagged <- c("Br6 Bullet 2-1", "Br9 Bullet 2-4", "Ukn Bullet B-2", "Ukn Bullet Q-4")
+ballstats <- read.csv("data-new-all-25-25/bullet-stats-single.csv")
+
 includesVar <- setdiff(names(ballstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "id", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km", "forestsmall",
@@
 
-ballstats <- read.csv("data-new-all-25-25/bullet-stats-single.csv")
 # names(ballstats)[10] <- "num.nonmatches"

### 2016-01-06 imaging-supplement.Rnw
fixed call in cross section code chunk

@@
 <<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
+
+scrubPath <- function(x) {
+  splits <- strsplit(as.character(x), split="/")
+  last <- sapply(splits, function(x) x[length(x)])
+  gsub(".x3p","", last)
+}
+
 paths <- file.path(datadir, dir(datadir))
 crosscuts <- seq(100, 400, by = 50)
-lof <- processBullets(paths[5], x = crosscuts)
+lof <- processBullets(read.x3p(paths[5]), name = scrubPath(paths[5]), x = crosscuts)
 lof$bullet <- paste(lof$bullet, lof$x)
@@
 \caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections.
-Cross section at $x = 100$ is compared to cross sections every 50$\mu m$.
-With every step away from the original cross section site, the number of differences between the cross sections increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
+The signature extracted at $x = 100$ is compared to signatures at every 50$\mu m$.
+With every step away from the original cross section site, the number of differences between the signatures increases; the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
 }
@@
 
-\hh{Show aligned signatures for all ten barrels. }
+%\hh{Show aligned signatures for all ten barrels. }
 

### 2016-01-06 imaging-supplement.Rnw
tons of cross sections to signatures

@@
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
-For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.
+For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all signatures.
 
@@
 A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction.
-Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues.
+Even when we fit a circle at each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues.
 While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.
@@
 @
-\caption{Circular fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
+\caption{Circular fit to the signature of each land of bullet 2, with signature from bullet 1-5 overlaid.\label{fig:circlefits} The signature of bullet 1-5 matches best with bullet 2-1.}
 \end{figure}
@@
 
-Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$.
-These are compared to a cross section at 100$\mu m$.
-Initially this comparison constitutes an almost perfect match between the two cross sections.
-However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.
-Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
+Figure~\ref{fig:crosscuts} shows a sequence of signatures for bullet 1-5 (barrel 1) that are taken at cross sections 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$.
+These are compared to the signature at cross section 100$\mu m$.
+Initially, this comparison constitutes an almost perfect match between the two signatures.
+However, the match quickly deteriorates with increasing distance between the cross sections.
+Only if signatures are from cross sections  within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
 Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections.
@@
 The signature extracted at $x = 100$ is compared to signatures at every 50$\mu m$.
-With every step away from the original cross section site, the number of differences between the signatures increases; the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher.
+With every step away from the original cross section site, the number of differences between the signatures increases; the number of maximum CMS decreases from initially 22 to  four or fewer at $x = 300\mu m$ and higher.
 }

### 2016-01-06 imaging-supplement.Rnw
first attempt at labelling the aligned signatures.

@@
 
-\newpage
 \section{Signature intensities}\label{supp:bulletbottom}
@@
 aligned$land[aligned$land==0] <- 6
-qplot(x=y, y=l30, data=aligned, geom="line", group=bullet) + ylim(c(-7,7)) + facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
-  xlab("") + ylab("Signature") + 
-  theme(plot.margin=unit(c(0,0,0,0), unit="cm"))
+aligned$bulletland <- aligned$bullet
+aligned$bullet <- gsub("-[0-9]","",aligned$bullet)
+library(RColorBrewer)
+cols <- rep(brewer.pal(n=10, name="Paired"), length=35)[sample(35, 35, replace=TRUE)]
+labels <- aligned %>% group_by(bulletland) %>% summarize(
+  x = max(y),
+  y = min(l30, na.rm=T),
+  label = bullet[1],
+  barrel = barrel[1],
+  land = land[1]
+)
+
+qplot(x=y, y=l30, data=aligned, geom="line", group=bullet, colour=bullet) + 
+  ylim(c(-7,7)) + xlab("") + ylab("Signature") + 
+  facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
+  theme(plot.margin=unit(c(0,0,0,0), unit="cm"), legend.position="none") + 
+  scale_colour_manual(values=cols) +
+  geom_label(aes(x=x, y=y, label=label), data=labels, 
+             inherit.aes = FALSE)
 @
@@
 <<importance, echo=FALSE, results='asis', warning=FALSE>>=
-imp <- data.frame(importance(rtrees))
+imp <- data.frame(importance(rtrees2))
 imp$Variable <- row.names(imp)
+
+imp2 <- data.frame(importance(rtrees3))
+imp2$Variable <- row.names(imp2)
+names(imp2)[1] <- "Bigforest"
+imp <- merge(imp, imp2, by="Variable")
+
 imp <- imp[order(-imp$MeanDecreaseGini),]
-names(imp)[1] <- "Importance"
-imp <- imp[,2:1]
+names(imp)[2:3] <- c("Importance 1", "Importance 2")
+
+bstats <- subset(ballstats, insmall == TRUE)
 imp$meanMatch <- sapply(imp$Variable, function(var) {
@@
 })
+
 imp$meanMatch <- round(imp$meanMatch,1)
@@
 row.names(imp) <- 1:nrow(imp)
-imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
-names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
-print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
+imp$Variable[c(2,3,4,6,8)] <- c("#matches", "S", "#non-matches", "CMS", "CNMS")
+names(imp)[-(1:3)] <- c("KM", "(sd)", "KNM", "(sd)")
+print(xtable(imp, align="clrrrd{3.2}rd{3.2}", digits=c(0,0,1,1,1,2,1,2)), 
       floating = FALSE)

### 2016-01-06 imaging-paper.Rnw
tiny edits

@@
   scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position = c(1, 1), legend.justification=c("right", "top"),
+  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
@@
   theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5.5, data=res$lines, alpha=0.2) +
   geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
@@
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6)) + ylab("") + xlab("") + 
+  ylim(c(-6,6.5)) + ylab("") + xlab("") + 
   geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{Br6 Bullet 2-1}
+\caption{Barrel 6 Bullet 2-1}
 %' <<echo=FALSE>>=
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{Br9 Bullet 2-4}
+\caption{Barrel 9 Bullet 2-4}
 \includegraphics[width=\textwidth]{images/br9-2-4-grey.png}
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{Unkn Bullet B-2}
+\caption{Unknown Bullet B-2}
 \includegraphics[width=\textwidth]{images/b-2-grey.png}
@@
 \begin{subfigure}[t]{.49\textwidth}\centering
-\caption{Unkn Bullet Q-4}
+\caption{Unknown Bullet Q-4}
 \includegraphics[width=\textwidth]{images/q-4-grey.png}
@@
 \caption{\label{fig:fourflags}Images of the four lands that got flagged during the quality assessment.
-All of them show scratch marks across the striation marks from the barrel.
-They will be excluded from the remainder of the analysis.}
+All of them show scratch marks (tank rash) across the striation marks from the barrel.
+They are excluded from the remainder of the analysis.}
 \end{figure}
@@
   \centering
-<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
+<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.65\\textwidth'>>=
 library(rpart)
@@
   \centering
-<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='.5\\textwidth'>>=
+<<oob, echo=FALSE, fig.width=6, fig.height=3, out.width='.6\\textwidth'>>=
 qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")

### 2016-01-06 imaging-supplement.Rnw
labelling of aligned signatures

@@
 \newcolumntype{d}[1]{D{.}{.}{#1}}
+
+\usepackage{pdflscape}
 
@@
 Signatures for some lands match remarkably well - such as land 5 from barrel 1, whereas all lands from barrel 5 show some variability both in the location and depths of peaks and valleys.
-\begin{figure}[hbtp]
-<<aligned, echo=FALSE, warning=FALSE, fig.width=12, fig.height=15, out.width='\\textwidth', cache=TRUE>>=
+
+<<aligned, echo=FALSE, warning=FALSE,  cache=TRUE>>=
 matches <- read.csv("csvs/matches.csv", header = FALSE)
@@
 library(RColorBrewer)
+set.seed(20160106)
 cols <- rep(brewer.pal(n=10, name="Paired"), length=35)[sample(35, 35, replace=TRUE)]
@@
   y = min(l30, na.rm=T),
-  label = bullet[1],
+  label = gsub(" Bullet","",bulletland[1]),
   barrel = barrel[1],
-  land = land[1]
+  land = land[1],
+  bullet = bullet[1]
 )
+labels <- labels %>% group_by(barrel, land) %>% mutate(id=order(bulletland))
+@
 
-qplot(x=y, y=l30, data=aligned, geom="line", group=bullet, colour=bullet) + 
-  ylim(c(-7,7)) + xlab("") + ylab("Signature") + 
+\begin{landscape}
+\begin{figure}[hbtp]
+<<aligned-first, dependson='aligned', echo=FALSE, warning=FALSE, fig.width=15, fig.height=10, out.width='7.5in'>>=
+qplot(x=y, y=l30, data=subset(aligned, barrel <=5), 
+      geom="line", group=bullet, colour=bullet) + 
+  ylim(c(-7,7)) + xlab("") + ylab("") + xlim(c(0,3000)) +
   facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
-  theme(plot.margin=unit(c(0,0,0,0), unit="cm"), legend.position="none") + 
-  scale_colour_manual(values=cols) +
-  geom_label(aes(x=x, y=y, label=label), data=labels, 
-             inherit.aes = FALSE)
+  theme(plot.margin=unit(c(0,0,-1,-1), unit="line"), legend.position="none") + 
+  geom_label(aes(y=2.5*(id-3), label=label, colour=bullet), 
+             x = 3000, data=subset(labels, barrel <= 5), 
+             inherit.aes = FALSE, hjust="right", size=3) +
+  scale_colour_manual(values=cols)
 @
-\caption{\label{fig:aligned}Overview of aligned signatures for all bullet lands for the ten barrels of the Hamby study. }
+\caption{\label{fig:aligned}Overview of aligned signatures for all bullet lands for  barrels 1 to 5 of the Hamby study. }
 \end{figure}
 
+\begin{figure}[hbtp]
+<<aligned-second, dependson='aligned', echo=FALSE, warning=FALSE, fig.width=15, fig.height=10, out.width='7.5in'>>=
+qplot(x=y, y=l30, data=subset(aligned, barrel > 5), 
+      geom="line", group=bullet, colour=bullet) + 
+  ylim(c(-7,7)) + xlab("") + ylab("") + xlim(c(0,3000)) +
+  facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
+  theme(plot.margin=unit(c(0,0,-1,-1), unit="line"), legend.position="none") + 
+  geom_label(aes(y=2.5*(id-3), label=label, colour=bullet), 
+             x = 3000, data=subset(labels, barrel > 5), 
+             inherit.aes = FALSE, hjust="right", size=3) +
+  scale_colour_manual(values=cols)
+@
+\caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet lands for  barrels 5 to 10 of the Hamby study. }
+\end{figure}
+\end{landscape}
 \newpage

### 2016-01-06 imaging-supplement.Rnw
feature importance including both random forests

@@
 \hh{One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land.
-This effectively doubles the number of comparisons from 10,384 pairwise comparisons of useable bullet lands to 21,115 ($= (118+88)\cdot 205/2$) comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons.
+This effectively doubles the number of comparisons from 10,384 pairwise comparisons of usable bullet lands to 21,115 $\left[= (118+88)\cdot 205/2\right]$ comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons.
 }
@@
 
-Figure~\ref{fig:aligned} gives an overview of all the signatures from bullet lands in the Hamby study aligned by barrel.
+Figures~\ref{fig:aligned} and~\ref{fig:aligned-second} give an overview of all the signatures from bullet lands in the Hamby study aligned by barrel.
 Three to five bullets were fired from each barrel.
-The figure gives us both some insight into how well signatures match and how consistent individual characteristics are impregnated on bullets fired from each of the barrels.
-Signatures for some lands match remarkably well - such as land 5 from barrel 1, whereas all lands from barrel 5 show some variability both in the location and depths of peaks and valleys.
+The figures give us both some insight into how well signatures match and how consistent individual characteristics are impregnated on bullets fired from each of the barrels.
+Signatures for some lands match remarkably well -- such as land 5 from barrel 1, whereas all lands from barrel 5 show some variability both in the location and depths of peaks and valleys.
 
@@
 \section{Table of feature importance}\label{supp:randomforest}
-
+Two random forests were calculated for the Hamby study.
+For the first random forest only comparisons of bullet lands from known bullets and unknown bullets were used.
+The second random forest is based on a full comparison of every land with every other land, increasing the number of comparisons from originally 10,384 (10,212 known non-matches and 172 known matches) by another 10,931 comparisons (10,637 known non-matches and 94 known matches).
+Random forests allow an assessment of variable importance (also called feature importance) as the mean decrease in Gini index when including each variable.
+Table~\ref{tab:importance} shows the results for feature importance for both of these random forests.
+Importance~1 refers to the smaller subset, Importance~2 is the feature importance derived from the random forest based on all pairwise comparisons.
 <<setup-rtrees, echo=FALSE, message=FALSE>>=

### 2016-01-06 imaging-paper.Rnw
fixing small tidbits

@@
 While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of a circular fit and the remaining residual structure can be found in supplement~\ref{supp:circular}.
+A discussion of a circular fit and the remaining residual structure can be found in supplement section~\ref{supp:circular}.
 %It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see supplement~\ref{supp:bulletbottom} for a more detailed discussion).
+We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see supplement section~\ref{supp:bulletbottom} for a more detailed discussion).
 
@@
 This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of which cross section level was used in a matching.
-An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement~\ref{supp:ccf} for a more thorough discussion).
+An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement section~\ref{supp:ccf} for a more thorough discussion).
 However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. 
@@
 Besides a probabilistic quantification of matches, we get from a random forest an assessment of the importance of each of the features derived from the bullets' 3d topological surface measurements.
-Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree.
+Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree (for the exact values please refer to supplement section~\ref{supp:randomforest}).
 \begin{figure}

### 2016-01-06 imaging-supplement.Rnw
fixing small tidbits

@@
 <<importance, echo=FALSE, results='asis', warning=FALSE>>=
-imp <- data.frame(importance(rtrees2))
-imp$Variable <- row.names(imp)
 
@@
 names(imp2)[1] <- "Bigforest"
+
+imp <- data.frame(importance(rtrees))
+imp$Variable <- row.names(imp2)
+
 imp <- merge(imp, imp2, by="Variable")

### 2016-01-06 imaging-paper.Rnw
Include the number of bullets in the intro

@@
 Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
-Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce known test bullets and unknown bullets for comparison.
+Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce 20 known test bullets and 15 unknown bullets for comparison.
 3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).

### 2016-01-06 imaging-paper.Rnw
Fix a couple typos i found

@@
 
-In this section, we are going to describe the algorithm for matching signatures first, and discuss the impact of paramater choices in the subsections thereafter.
+In this section, we are going to describe the algorithm for matching signatures first, and discuss the impact of parameter choices in the subsections thereafter.
 
@@
 A smoothing factor of $k$ means that the  $k$ closest observations to $x_o$ are considered for a fit for $x_o$.
-Because surface measurements are recorded at an equidistand resolution (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$..
+Because surface measurements are recorded at an equidistant resolution (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$..
 For detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor  of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.

### 2016-01-06 imaging-paper.Rnw
One more typo corrected

@@
  run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e.
-we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unkown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons.
+we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unknown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons.
 Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
@@
 Neither one of these methods allows for an assessment of cross-correlations.
-This is one of the benefits of switching to a digitalized version of the images that preserves the 3d surface structure.
+This is one of the benefits of switching to a digitized version of the images that preserves the 3d surface structure.
 The emphasis on cross-correlation is similar to the results of the study by \citet{ma:2004}.

### 2016-01-06 imaging-paper.Rnw
refer to extended hamby results

@@
 
+\hh{Besides including results from known matches againts known non-matches, we can increase the number of comparisons in the Hamby study to include all possible land-to-land comparisons.
+This effectively doubles the number of data points available.
+Comparisons not previously included in fitting the random forest can also be used as an additional source for assessing error rates.
+Results for this and a more detailed discussion can be found in supplement section~\ref{supp:extended}.}
+
 \section{Conclusion}

### 2016-01-06 imaging-paper.Rnw
to do list out

@@
 \section{Introduction}
-\hh{
-Todo list for journal shape-up: there is a lot of good material, we mainly need to focus on what story to tell and make that a solid argument.
-\begin{enumerate}
+%\hh{
+%Todo list for journal shape-up: there is a lot of good material, we mainly need to focus on what story to tell and make that a solid argument.
+%\begin{enumerate}
 %\item take out a few things: no circular fit, no figures 9 and 10 
@@
 (moved to supplement)
-\item add: should we add a paragraph on the speed of the evaluation?
+%\item add: should we add a paragraph on the speed of the evaluation?
 - separate out into pre-processing (needs to be done once) and matching (pairwise comparison of lands).
@@
 We would need somebody qualified to tell us how many CMS there are, though.
-\item to consider: AOAS states that accepted papers rarely have more than 20 pages. We are currently over by a couple of pages.
+%\item to consider: AOAS states that accepted papers rarely have more than 20 pages. We are currently over by a couple of pages.
 %\item definitions: 
@@
 The {\it cross section} defines the location along the $x$ axis  at which the signature is taken.
-\end{enumerate}
-}
+%\end{enumerate}
+%}
 Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
@@
 
+\hh{XXX speed? XXX}
+
 An approach that might help in determining this difference comes from Biometrics \citep{jain:2015:tphil, jain:2015}.

### 2016-01-07 imaging-paper.Rnw
couple more words on the extended study

@@
 
-set.seed(20151202)
+set.seed(20160105)
 names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
@@
 After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0039.
-This is a weighted average between false positive error rate of 0.0002 and an error rate of false negatives of 0.2267.
+This is a weighted average between false positive error rate of 0.0001 and an error rate of false negatives of 0.2267.
 This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).

### 2016-01-07 imaging-supplement.Rnw
couple more words on the extended study

@@
 The smooth line on top is a circle, with estimated radius and center.
-The details of this fit are given below.
-We then remove the fit to investigate residuals.
+The details of this fit are given below:
 %
@@
 
-Assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle.
+Assume that $n$ data points are given in the form of data tuples $(x_1, y_1)$, $(x_2, y_2)$, $...$, $(x_n, y_n)$ that are (approximately) located on a circle.
 We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.
@@
 
+\newpage
 \section{Signature intensities}\label{supp:bulletbottom}
@@
 
-\newpage
+%\newpage
 \section{Complete evaluation of the Hamby study}\label{supp:extended}
@@
 
-\hh{When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the estimated error rates of 0.2267 and 0.0002.
+\hh{When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the random forest's estimated OOB error rates of 0.226744 and 0.000098.
 }
@@
 \hh{However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0.
-The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.}
+The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.
+The false positive rate is therefore virtually unchanged, while we see a slight improvement in the false negative rate for an overall OOB error rate of 0.3\%.}
 <<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=

### 2016-01-07 imaging-supplement.Rnw
comment on study size

@@
 The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.
-The false positive rate is therefore virtually unchanged, while we see a slight improvement in the false negative rate for an overall OOB error rate of 0.3\%.}
+The false positive rate is therefore virtually unchanged, while we see a slight improvement in the false negative rate for an overall OOB error rate of 0.3\%, i.e.\ an increase to twice the number of comparisons leads to a decrease of 25\% of the estimated error rate.
+This is yet another argument in favor of a larger database for training algorithms.}
 <<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=

### 2016-01-07 imaging-supplement.Rnw
see-through labels

@@
              x = 3000, data=subset(labels, barrel <= 5), 
-             inherit.aes = FALSE, hjust="right", size=3) +
+             inherit.aes = FALSE, hjust="right", size=3, 
+             fill=alpha("white", .5)) +
   scale_colour_manual(values=cols)

### 2016-01-08 imaging-paper.Rnw
added acknowledgement

@@
 \section*{Acknowledgment} 
+%Thanks to David Baldwin for pointing us to the NIST database and introducing us to the basic terms.
 Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.

### 2016-01-08 imaging-paper.Rnw
Couple small tweaks here and there

@@
 We first have to identify the locations of the left and right  groove in the image.
-The grooves are not assumed to contain information relevant for determining matches, they also dominate the structure, and therefore need  to be removed.
+The grooves are not assumed to contain information relevant for determining matches.
+They also dominate the structure, and therefore need  to be removed.
 Fortunately, the location and appearance of the grooves in the surface profiles is quite consistent.

### 2016-01-08 imaging-paper.Rnw
Typo

@@
 A smoothing factor of $k$ means that the  $k$ closest observations to $x_o$ are considered for a fit for $x_o$.
-Because surface measurements are recorded at an equidistant resolution (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$..
+Because surface measurements are recorded at an equidistant resolution (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$.
 For detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor  of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.

### 2016-01-12 imaging-paper.Rnw
fixed y label

@@
   scale_fill_brewer("", palette="Paired") +
-  xlab("maximum CMS")
-
+  xlab("maximum CMS") + ylab("Proportion")
 @

### 2016-01-14 imaging-paper.Rnw
Reword mission statement

@@
 
-The primary goal in this type of forensic analysis is to identify these striation markings in an attempt to match two bullets.
-Since gun barrels should produce similar markings on bullets fired consecutively, the careful analysis of these surface images can help in determining whether two bullets are a match.
-This process has traditionally required forensic experts to perform the analysis, and required physical examination of the bullets.
+From a statistical standpoint, identification of the gun that fired the bullet(s) requires that we compare the probabilities of observing matching striae under the competing hypotheses that the gun fired, or did not fire, the crime scene bullet.
+If indeed the uniqueness assumption is plausible, the latter probability approaches zero.
 
@@
 
-In this paper, we propose a framework which allows for the automatic analysis of the surface topologies of bullets, and the transcription of the individual characteristics into a 2D plotting framework.
+Here, we focus on the question of defining what constitutes a match, since there is no agreement in the forensics community.
+We propose a framework which allows for the automatic analysis of the surface topologies of bullets, and the transcription of the individual characteristics into a 2D plotting framework.
 This allows for an objective and quantitative assessment of striae-based bullet matches.

### 2016-01-14 imaging-paper.Rnw
Fix broken supplement reference

@@
 \end{figure}
-Each fired bullet is provided in form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
+Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
 For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
@@
 
-In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements for a fixed value of $x$.
+In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements at a fixed value of the bullet's height, denoted by $x$.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of a circular fit and the remaining residual structure can be found in supplement section~\ref{supp:circular}.
+A discussion of a circular fit and the remaining residual structure can be found in Supplement section~\ref{supp:cylindrical}.
 %It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.

### 2016-01-14 imaging-paper.Rnw
Fix some more typos

@@
 %    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor as above. (Figure~\ref{fig:loess_step3}).
-    \item Determine the location of the peak of the left groove by finding the first double-smoothed value $y_i$ that is the maximum within its smoothing window (e.g.\ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$, where $i$ is between 1  and $\lfloor s/2 \rfloor$).
+    \item Determine the location of the peak of the left groove by finding the first doubly-smoothed value $y_i$ that is the maximum within its smoothing window (e.g.\ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$, where $i$ is between 1  and $\lfloor s/2 \rfloor$).
 We call the location of this peak $p_{\ell}$ (see Figure~\ref{fig:loess_step47}).
@@
 In the groove detection portion of this algorithm, we wish to remove most of the deviation but still maintain enough signal in the groove to detect it.
-Empirically, a value of $s = 35$ for the smoothing factor seems to work quite well (the smoothing factor is further discussed in section~\ref{sec:smoothing}).
+Empirically, a value of $s = 35$ for the smoothing factor seems to work quite well (the smoothing factor is further discussed in Section~\ref{sec:smoothing}).
 It is important to note that the smoothing pass is done twice.
@@
 This has the effect of introducing a bit more variance back into the smoothed data.
-We instead performed a second smoothing pass on the smoothed data, which has the effect of weighting observations near the center of the window the highest, while the weights linearly drop off as we reach either end of the smoothing window.
+We instead performed a second smoothing pass on the smoothed data, which has the effect of weighting observations near the center of the window the highest, with the weights linearly dropping off as we reach either end of the smoothing window.
 
@@
 Loess regression \citep{cleveland:1979} is based on the assumption that the relationship between two random variables $X$ and $Y$ can be described in form of a smooth, continuous function $f$ with $y_i = f(x_i) + \varepsilon_i$ for all values $i = 1, ..., n$.
-The function $f$ is estimated in form of  locally weighted polynomial regressions.
+The function $f$ is approximated via locally weighted polynomial regressions.
 Parameters of the estimation are $\alpha$, the proportion of all points included in the fit (here, $\alpha = 0.75$), the weighting function and the degree of the polynomial (here, we fit a quadratic regression).
@@
 The weighting function used here, is the tricubic function $w(d) = \left(1 - d^3\right)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
-$d$ is defined as the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
+Here, $d$ is defined as the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
 
@@
 
-Figure~\ref{fig:loess_fit} provides a look at the loess fit, in blue, overlaid against the processed image of bullet~1-5.
+Figure~\ref{fig:loess_fit} shows the loess fit, in blue, overlaid on the processed image of bullet~1-5.
 The fit seems to do a reasonable job of capturing the structure of the image.
@@
 
-In this section, we are going to describe the algorithm for matching signatures first, and discuss the impact of parameter choices in the subsections thereafter.
+In this section, we describe the algorithm for matching signatures first, and the impact of parameter choices in the subsections thereafter.
 
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see supplement section~\ref{supp:bulletbottom} for a more detailed discussion).
+We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement section~\ref{supp:bulletbottom} for a more detailed discussion).
 
@@
 Non-matching intervals are left grey.
-\item {\bf Extract features from the aligned signatures and the matches between them:} there is a slew of different features one could extract.
+\item {\bf Extract features from the aligned signatures and the matches between them:} many different features could be extracted.
 Here, we describe a few of the ones that can be found in the literature and some that we found to be of practical relevance:
 \begin{enumerate}[label=(\roman*)]
-\item maximal number of CMS (consecutive matching striae), and, similarly, the number of consecutively non-matching striae (CNMS), 
-\item number of matches and non-matches,
-\item average difference between signatures $D$, defined as the Euclidean vertical distance of surface measurements between aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
+\item Maximal number of CMS (consecutive matching striae), and, similarly, the number of consecutively non-matching striae (CNMS), 
+\item Number of matches and non-matches,
+\item Average difference $D$ between signatures, defined as the Euclidean vertical distance between surface measurements of aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
 \[
@@
 \]
-\item the sum $S$ of matched extrema: for each of the matched stria, determine the average of the absolute height of the peak or valley. $S$ is then defined as the sum of all these averages. 
+\item The sum $S$ of average absolute heights of matched extrema: for each of the two matched stria, compute the average of the absolute heights of the peaks or valleys.
+$S$ is then defined as the sum of all these averages.
 \end{enumerate}
@@
 
-Assessing the maximal number of CMS is part of current practice in evaluating bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}.
+Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}.
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets.
@@
 
-Determining a CMS threshold that needs to be exceeded for us to declare a match is beyond the scope of this work, even though it is critically important in practice.
-We provide some ideas in the next section, but first we assess the impact of the parameters chosen in the matching algorithm.
+Determining a threshold such that CMS values above the threshold indicate a match with high reliability is beyond the scope of this work, even though it is critically important in practice.
+We provide some ideas in the next section, but first we assess the robustness of the matching algorithm to different choices of the parameter values.
 %In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
-Aside of the features listed above, we also keep track of the value of the cross-correlation function found during the alignment, the  lag, at which signatures show the best alignment, and the cross section levels ($x_1$ and $x_2$).
+In addition to the features listed above, we also keep track of the value of the cross-correlation function found during the alignment, the lag at which signatures show the best alignment, and the cross section levels ($x_1$ and $x_2$).
 
@@
 This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of which cross section level was used in a matching.
-An assessment of this situation shows that, visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see supplement section~\ref{supp:ccf} for a more thorough discussion).
+Visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion).
 However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. 
@@
 by comparing signatures from cross sections that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
-In the approach here, we keep increasing the cross section until we find a section with a stable pattern.
+In the approach here, we keep increasing the height $x$ of the cross section until we find a section with a stable pattern.
 This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
@@
 \subsection{Varying smoothing factor}\label{sec:smoothing}
-As mentioned earlier, the algorithm for detecting peaks and valleys depend on the selection of a smoothing window, called the smoothing factor or span.
+As mentioned earlier, the algorithm for detecting peaks and valleys depends on the selection of a smoothing window, called the smoothing factor or span.
 A smoothing factor of $k$ means that the  $k$ closest observations to $x_o$ are considered for a fit for $x_o$.
@@
 
-A larger smoothing factor on the other hand, like 45, seems to be a more plausible option.
+A larger smoothing factor on the other hand (like 45), seems to be a more plausible option.
 Most of the peaks/valleys present which are detected by a smoothing factor of 25 are also detected at 45.
@@
 Notice that the valley on the right hand side of the image is smoothed out, and thus not detected.
-On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are being smoothed out, for example the peak at around $y = 750$.
+On the left hand side, a double peak is detected - that might be a questionable decision - but there are several peaks in the middle, that are smoothed out, for example the peak at around $y = 750$.
 That is, in many cases, large windows are smoothing out some of the structure that we wish to see.
@@
 As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
-For this, we assess cross-correlation of signatures at cross sections 25$\mu m$ apart.
+For this, we compute the cross-correlation of signatures at cross sections 25$\mu m$ apart.
 For a stable region, we require a minimum of 0.9 for the cross correlation.
 Four lands from different bullets are flagged as problematic in this respect.
-A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface, also known as `tank rash'.
+A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands has scratch marks across the surface, also known as `tank rash'.
 %
@@
 %
-We exclude these four lands from the further matching considerations and 
+We exclude these four lands from further matching considerations and 
  run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e.
@@
 The densities of almost all of the features show strong differences between matches and non matches.
-For example, a high amount of cross-correlation between two signatures is indicative for a match --  in the Hamby study only known matches have a cross-correlation of 0.75 or higher.
+For example, a high amount of cross-correlation between two signatures is indicative of a match --  in the Hamby study, only known matches have a cross-correlation of 0.75 or higher.
 There are 97 land-to-land comparisons with a cross-correlation that high.

### 2016-01-14 imaging-paper.Rnw
Final set of edits

@@
 All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches.
-The predictive power of each one of these features is shown in form of the Receiving Operating Characteristic (ROC) curves in Figure~\ref{fig:rocs}.
-The features are ordered according to the area under the curve (AUC).
-The feature with the highest individual predictive power is $S$, the sum of two signatures' heights of joint peaks and valleys.
+The predictive power of each one of these features is shown in the form of the Receiving Operating Characteristic (ROC) curves in Figure~\ref{fig:rocs}.
+The features are arranged in descending order according to the area under the curve (AUC).
+The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys.
 The maximal number of CMS is only in the seventh position here.
@@
 Instead of CMS, cross-correlation (ccf) between the signatures is very important in the matching process  by the decision tree.
-Aside from  cross correlation, the number of matches is involved in the decision.
+Aside from  cross correlation, the total number of matches is also included in the decision rule.
 Between cross-correlation and CMS, cross-correlation has higher  predictive power.
-This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2d microscopic images of bullets.
+This  does not  contradict earlier findings emphasizing the value of CMS on visual assessments of bullet matches: in those papers, assessments were based on purely visual inspection of either actual bullets or 2D microscopic images of bullets.
 Neither one of these methods allows for an assessment of cross-correlations.
-This is one of the benefits of switching to a digitized version of the images that preserves the 3d surface structure.
-The emphasis on cross-correlation is similar to the results of the study by \citet{ma:2004}.
+This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure.
+The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}.
 However, in this study authors did not consider the number of matches and non-matches.
@@
 
-Another benefit of the digitized version of the images is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\citep{breiman:2001, randomForest}.
-For each of the trees in a random forest, only two thirds of the observations are used for fitting, while  the remaining third  are used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate.
+Another benefit of the digitized version of the images is that we can apply several hundred decision trees to combine in a random forest~\citep{breiman:2001, randomForest}.
+For each of the trees in a random forest, only two thirds of the observations are used for fitting, while  the remaining third is used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate.
 Because errors are determined from the one third of held-back observations, this error rate is called the out-of bag (OOB) error.
@@
 However, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets.
-Matching bullets can only be tested realistically in a much bigger test case.
+Matching bullets can only be tested realistically in a much bigger experiment.
 Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared.
@@
 
-Besides a probabilistic quantification of matches, we get from a random forest an assessment of the importance of each of the features derived from the bullets' 3d topological surface measurements.
-Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the gini index when the variable in question is included in a tree (for the exact values please refer to supplement section~\ref{supp:randomforest}).
+Besides resulting in a probabilistic quantification of matches, random forests also provide an assessment of the importance of each of the features derived from the bullets' 3D topological surface measurements.
+Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to supplement section~\ref{supp:randomforest}).
 \begin{figure}
@@
 %The rows in the table are ordered according to descending importance. 
-The variables with the most predictive power by far are cross-correlation and the overall number of matches, followed by the total depth of joint striations $S$ and non-matches.
+The variables with the most predictive power are cross-correlation and the overall number of matches, followed by the total depth of joint striations $S$ and total number of non-matches.
 CMS is found only in sixth place.
@@
 
-We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a ballistics identification perspective, removes these features, and produces residuals which allows for the easy identification of markings.
+We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a ballistics identification perspective, removes these features, and produces residuals which allow for the easy identification of markings.
 We have generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches.
@@
 The matching algorithm is sensitive to the parameter choices made.
-The distance between cross sections (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum are affecting the final outcome.
-Another parameter is the amount of smoothing in identifying peaks and grooves (currently a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
-While we tried to lay out in the paper the impact each of the parameter choices has on the matching, we are still far from an optimized scenario.
+The distance between cross sections (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum threshold do affect the final outcome.
+Another parameter that must be selected is the amount of smoothing when identifying peaks and grooves (currently, a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
+While we tried to lay out in the paper the impact that each of the parameter choices has on the matching performance, we are still far from an optimized scenario.
 
 The Hamby study serves as our evaluation `database'.
-It consists of only  35 bullets -- this is obviously not a particularly realistic scenario for automatic matching.
-But this is what at the moment  is available to us.
-As the NIST database expands, this might get into more challenging territory.
-
-The feasibility of a database of ballistic images was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
+It consists of only  35 bullets -- this is obviously not a particularly realistic scenario for an automatic matching procedure, but for now we are unaware of other databases containing bullets in the x3p format.
+
+The feasibility of creating a database of ballistic images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
 The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
 The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found.
-The remaining problem is simply that statistically the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
-The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks.
+The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
+The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks.
 This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system.
 A further re-assessment by \citet{deceuster:2015} came to the same conclusions based on the EvoFinder system.
-The NRC report also found, that results from 2d images are improved upon by matches based on 3d images.
-This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the number of matches and mismatches are available for a match based on 2d features.
+The NRC report also found, that results from 2D images can be improved when matches are based on 3D images.
+This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the total number of matches and mismatches are available for a match based on 2D features.
 
@@
 
-For a fair assessment of the adequacy of an algorithm we need transparency.
+For a fair assessment of the performance of an algorithm, we need transparency.
 Our matching  algorithm is open: the code is readily available in form of the R package x3prplus~\citep{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
-Obviously, this is only a first step, but in times of the Netflix prize~\citep{koren:2009, toescher:2009, piotte:2009} and the many Data Science challenges hosted by Kaggle\footnote{kaggle home: \url{https://www.kaggle.com/solutions/competitions}}, we should not let the quality of the matching algorithm be the problem that stands in the way.
-
-Only a much larger database and transparent communication of results will give us  answers to whether automatic matching of ballistic evidence is really feasible,  whether there really is individualization of this evidence, as is often silently assumed.
-And in the case, that individualization is a reasonable assumption, whether  there is a way to be able to distinguish between individual and subclass characteristics.
+To understand whether an automated approach along the lines of the one we propose can accurately identify sets of bullets with undistinguishable markings, it will be necessary to assemble a much larger database that includes a wide range of ammunition types, degrees of damage, gun makes, etc.
+We are unaware of the existence of any such database.
+In addition to serving as a realistic testbed for the performance of the automated matching algorithm, such a database would also permit testing the underlying, as of yet untested, assumptions of uniqueness and reproducibility of the markings left by a gun on bullets.
 
@@
 
-An approach that might help in determining this difference comes from Biometrics \citep{jain:2015:tphil, jain:2015}.
+An approach that might be helpful comes from Biometrics \citep{jain:2015:tphil, jain:2015}.
 It has been successfully applied in the Aadhar project (the effort by the Unique Identification Authority of India (UIDAI) to provide a unique 12-digit identification number to approximately 1.2 billion residents of India based on ten fingerprints and two irises).

### 2016-01-15 imaging-supplement.Rnw
Fix some more typos
typos

@@
 This is yet another argument in favor of a larger database for training algorithms.}
+
+\hh{XXX maybe use log of D to remove the left boundary XXX}
 <<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=

### 2016-01-15 imaging-paper.Rnw
typos
calculating CMS differently (named cms2)

@@
 Non-matching intervals are left grey.
+\hh{XXX this definition does not coincide with the one given in \citet{thompson:2013}.
+Here, CMS is defined only in terms of matching peaks.
+Additionally, \citet{thompson:2013} peaks are only used, if they can be identified and matched `within a tolerable range' between lands.
+The definition given here, is computationally far less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land.
+XXX}
 \item {\bf Extract features from the aligned signatures and the matches between them:} many different features could be extracted.
@@
 \[
-D^2 = \frac{1}{\text{\#}t}\sum_t \left(f(t) - g(t)\right)^2,
+D^2 = \frac{1}{\text{\#}t}\sum_t \left[f(t) - g(t)\right]^2,
 \]

### 2016-01-15 code/evaluation-hamby.R

@@ -5,7 +5,7 @@ datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
 
 CCFs <- plyr::ldply(datas, function(x) {
   load(x)
-  cmsdist <- sapply(reslist, function(x) x$maxCMS)
+#  cmsdist <- sapply(reslist, function(x) x$maxCMS)
   ccf <- plyr::ldply(reslist, function(res) {
     lofX <- res$bullets
     b12 <- unique(lofX$bullet)
@@ -30,7 +30,7 @@ CCFs <- plyr::ldply(datas, function(x) {
     knm <- which(!res$lines$match)
     if (length(km) == 0) km <- c(length(knm)+1,0)
     if (length(knm) == 0) knm <- c(length(km)+1,0)
-# browser()    
+ #browser()    
     # feature extraction
     data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), 
                D=distr.dist, 
@@ -38,6 +38,8 @@ CCFs <- plyr::ldply(datas, function(x) {
                b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
                num.matches = sum(res$lines$match), 
                num.mismatches = sum(!res$lines$match), 
+               cms = res$maxCMS,
+               cms2 = x3prplus::maxCMS(subset(res$lines, type==1 | is.na(type))$match),
                non_cms = x3prplus::maxCMS(!res$lines$match),
                left_cms = max(knm[1] - km[1], 0),
                right_cms = max(km[length(km)] - knm[length(knm)],0),
@@ -46,7 +48,7 @@ CCFs <- plyr::ldply(datas, function(x) {
                sumpeaks = sum(abs(res$lines$heights[res$lines$match]))
                )
   })
-  ccf$cms <- cmsdist
+#  ccf$cms <- cmsdist
   ccf$data <- x
   ccf  
 })
@@ -87,9 +89,19 @@ rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!
 prp(rp1, extra = 101)
 CCFs$pred <- predict(rp1)
 
+includes2 <- setdiff(includes, c("left_cms", "right_cms", "left_noncms", "right_noncms", "cms2"))
 library(randomForest)
-rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
+set.seed(20160105)
[…]

### 2016-01-15 code/evaluation-hamby.R
cms and cms2 values for comparison

@@ -103,7 +103,7 @@ imp1b <- data.frame(importance(rtrees1b))
 
 
 
-write.csv(CCFs, file=file.path(dataStr, "bullet-stats.csv"), row.names=FALSE)
+write.csv(CCFs, file=file.path(dataStr, "bullet-stats-cms.csv"), row.names=FALSE)
 
 
 ##################################################

### 2016-01-15 imaging-paper.Rnw
bit more on the difference/similarity of the CMS definition

@@
 \hh{XXX this definition does not coincide with the one given in \citet{thompson:2013}.
-Here, CMS is defined only in terms of matching peaks.
+There, CMS is defined only in terms of matching peaks.
 Additionally, \citet{thompson:2013} peaks are only used, if they can be identified and matched `within a tolerable range' between lands.
-The definition given here, is computationally far less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land.
+The definition given in this paper, is computationally far less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land.
+In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92).
 XXX}

### 2016-01-15 imaging-paper.Rnw
moved the CMS discussion to the proper place

@@
 Non-matching intervals are left grey.
-\hh{XXX this definition does not coincide with the one given in \citet{thompson:2013}.
-There, CMS is defined only in terms of matching peaks.
-Additionally, \citet{thompson:2013} peaks are only used, if they can be identified and matched `within a tolerable range' between lands.
-The definition given in this paper, is computationally far less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land.
-In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92).
-XXX}
 \item {\bf Extract features from the aligned signatures and the matches between them:} many different features could be extracted.
@@
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
+\hh{XXX Note that the definition given in this paper does not coincide with the one given in \citet{thompson:2013}.
+There, CMS is defined only in terms of matching peaks.
+Additionally, \citet{thompson:2013} peaks are only used, if they can be identified and matched `within a tolerable range' between lands.
+The definition given in here, is computationally less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land.
+In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92).
+XXX}
 For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
-We will assess this claim in the next section.
+\hh{XXX Biasotti's four or more consecutive lines is closer to eight or more with our definition.
+Should we just multiply our CMS by 0.5, to account for the fact that between every two peaks has to be a valley?
+XXX} We will assess this claim in the next section.
 

### 2016-01-16 imaging-paper.Rnw
A few minor revisions to new CMS paragraph

@@
 There, CMS is defined only in terms of matching peaks.
-Additionally, \citet{thompson:2013} peaks are only used, if they can be identified and matched `within a tolerable range' between lands.
-The definition given in here, is computationally less complex, but should come close in values, because of the requirement to only consider signatures from a stable region in the land.
+Additionally, \citet{thompson:2013} peaks are only used if they can be identified and matched `within a tolerable range' between lands.
+The definition given here is computationally less complex, but should yield similar values, because of the requirement to only consider signatures from a stable region in the land.
 In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92).

### 2016-01-16 imaging-paper.Rnw
Lands from known bullets

@@
 Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
- When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 known lands.
+ When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 lands from known bullets.
 Two lands show a high CMS.

### 2016-01-16 imaging-paper.Rnw
AUC calculation upped to 10,000 samples - still a tie between ccf and S

@@
 
+set.seed(20140501)
 aucs <- plyr::ldply(features, function(x) {
   subbstats <- subset(bstats, flagged == FALSE)
-  pos.scores <- sample(subbstats[which(subbstats$match),x], 5000, replace=TRUE)
-  neg.scores <- sample(subbstats[which(!subbstats$match),x], 5000, replace=TRUE)
+  pos.scores <- sample(subbstats[which(subbstats$match),x], 10000, replace=TRUE)
+  neg.scores <- sample(subbstats[which(!subbstats$match),x], 10000, replace=TRUE)
   data.frame(x, upper=mean(pos.scores > neg.scores), 

### 2016-01-17 imaging-paper.Rnw
reference added to stable region

--- 2016-01-16 imaging-paper.Rnw
@@
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
-\hh{XXX Note that the definition given in this paper does not coincide with the one given in \citet{thompson:2013}.
+\hh{XXX Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}.
 There, CMS is defined only in terms of matching peaks.
-Additionally, \citet{thompson:2013} peaks are only used if they can be identified and matched `within a tolerable range' between lands.
-The definition given here is computationally less complex, but should yield similar values, because of the requirement to only consider signatures from a stable region in the land.
+Additionally, peaks in  \citet{thompson:2013} are only used if they can be identified and matched `within a tolerable range' between lands.
+The definition given here is computationally less complex, but should yield similar values, because of the requirement to only consider signatures from a stable region in the land (see Section~\ref{sec:heights} for further details on stability of regions).
 In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92).
@@
 
-\subsection{Impact of cross sections}
+\subsection{Impact of cross sections}\label{sec:heights}
 The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
@@
 This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
-`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.9.
+`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.
 
@@
 
+set.seed(20140105)
 aucs <- plyr::ldply(features, function(x) {
   subbstats <- subset(bstats, flagged == FALSE)
-  pos.scores <- sample(subbstats[which(subbstats$match),x], 5000, replace=TRUE)
-  neg.scores <- sample(subbstats[which(!subbstats$match),x], 5000, replace=TRUE)
+  pos.scores <- sample(subbstats[which(subbstats$match),x], 10000, replace=TRUE)
+  neg.scores <- sample(subbstats[which(!subbstats$match),x], 10000, replace=TRUE)
   data.frame(x, upper=mean(pos.scores > neg.scores), 

### 2016-01-17 imaging-paper.Rnw
typo

@@
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
-\hh{XXX Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}.
-There, CMS is defined only in terms of matching peaks.
-Additionally, peaks in  \citet{thompson:2013} are only used if they can be identified and matched `within a tolerable range' between lands.
-The definition given here is computationally less complex, but should yield similar values, because of the requirement to only consider signatures from a stable region in the land (see Section~\ref{sec:heights} for further details on stability of regions).
+\hh{ Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}.
+There, CMS is defined only in terms of matching peaks without regarding valleys.
+Additionally, peaks in  \citet{thompson:2013} are  used only if they can be identified and matched `within a tolerable range' between lands.
+The definition given here is computationally less complex, but should yield highly correlated values, because of the requirement to only consider signatures from a stable region in the land (see Section~\ref{sec:heights} for further details on stability of regions).
 In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92).
-XXX}
-For lead bullets, such as this, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
-\hh{XXX Biasotti's four or more consecutive lines is closer to eight or more with our definition.
-Should we just multiply our CMS by 0.5, to account for the fact that between every two peaks has to be a valley?
-XXX} We will assess this claim in the next section.
+}
+For lead bullets, such as used in the Hamby study, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
+\hh{Biasotti's four or more consecutive lines translates to eight or more with our definition.
+} %We will assess the claim that eight or more consecutive lines are sufficient for a match in the next section.
 
@@
   subbstats <- subset(bstats, flagged == FALSE)
-  pos.scores <- sample(subbstats[which(subbstats$match),x], 10000, replace=TRUE)
-  neg.scores <- sample(subbstats[which(!subbstats$match),x], 10000, replace=TRUE)
+  pos.scores <- sample(subbstats[which(subbstats$match),x], 50000, replace=TRUE)
+  neg.scores <- sample(subbstats[which(!subbstats$match),x], 50000, replace=TRUE)
   data.frame(x, upper=mean(pos.scores > neg.scores), 
@@
 imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
-imp1 <- subset(imp, !(Variable %in% c("x1", "sd.D")))
-imp2 <- subset(imp, Variable %in% c("x1", "sd.D"))
+imp1 <- subset(imp, !(Variable %in% c("x1", "sd.D", "S")))
+imp2 <- subset(imp, Variable %in% c("x1", "sd.D", "S"))
 qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
@@
 
-\hh{Besides including results from known matches againts known non-matches, we can increase the number of comparisons in the Hamby study to include all possible land-to-land comparisons.
+\hh{Besides including results from known matches against known non-matches, we can increase the number of comparisons in the Hamby study to include all possible land-to-land comparisons.
 This effectively doubles the number of data points available.

### 2016-01-17 imaging-paper.Rnw
A couple typos

@@
 Besides resulting in a probabilistic quantification of matches, random forests also provide an assessment of the importance of each of the features derived from the bullets' 3D topological surface measurements.
-Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to supplement section~\ref{supp:randomforest}).
+Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to Supplement section~\ref{supp:randomforest}).
 \begin{figure}
@@
 Comparisons not previously included in fitting the random forest can also be used as an additional source for assessing error rates.
-Results for this and a more detailed discussion can be found in supplement section~\ref{supp:extended}.}
+Results for this and a more detailed discussion can be found in Supplement section~\ref{supp:extended}.}
 

### 2016-01-17 imaging-supplement.Rnw
Disable messages in the supplement

@@
 \section{Signature intensities}\label{supp:bulletbottom}
-<<setup-signatures, echo=FALSE>>=
+<<setup-signatures, echo=FALSE, message=FALSE>>=
 library(x3pr)

### 2016-01-19 imaging-paper.Rnw
size of figures
captions of first figure

@@
   \centering
-\begin{subfigure}[t]{.54\textwidth}\centering
-\caption{View from the side.\label{fig:sidex3p}}{%
+\begin{subfigure}[t]{.65\textwidth}\centering
+\caption{View from the side.
+\hh{XXX it's ambigious to talk about side and top.
+because it's not the side of the bullet(land), but the `side' of the data.
+I'm not sure, how to phrase it, though.
+How about: View of the data along the circumference of the bullet (circular segment of about 30 degrees).
+}\label{fig:sidex3p}}{%
       \includegraphics[width=\textwidth]{images/sidex3p.png}
@@
 \end{subfigure}    
-\begin{subfigure}[t]{.45\textwidth}\centering
-    \caption{View from the top.\label{fig:topx3p}}{%
+\begin{subfigure}[t]{.65\textwidth}\centering
+    \caption{View from the top. \hh{XXX same problem as above. XXX How about: View of a bullet land (lower end of the view is the bottom of the bullet). }\label{fig:topx3p}}{%
     \includegraphics[width=\textwidth]{images/topx3p.png}
@@
   \centering
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=
+<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.65\\textwidth'>>=
 

### 2016-01-19 imaging-paper.Rnw
bullet land sketches
language: cross section to height (some), 3d to 3D (all)

--- 2016-01-17 imaging-paper.Rnw
@@
 \begin{keyword}
-\kwd{3d Topological Surface Measurement}
+\kwd{3D Topological Surface Measurement}
 \kwd{Data visualization}
@@
   \centering
-\begin{subfigure}[t]{.54\textwidth}\centering
-\caption{View from the side.\label{fig:sidex3p}}{%
-      \includegraphics[width=\textwidth]{images/sidex3p.png}
-    }
-\end{subfigure}    
-\begin{subfigure}[t]{.45\textwidth}\centering
-    \caption{View from the top.\label{fig:topx3p}}{%
-    \includegraphics[width=\textwidth]{images/topx3p.png}
-    }
-\end{subfigure}
-\caption{An example of a scan of a bullet land from one groove to the next. }
+\begin{subfigure}[t]{\textwidth}\centering
+\caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
+      \includegraphics[width=.65\textwidth]{images/sidex3p.png}
+      \hspace{1cm}
+      \includegraphics[width=.25\textwidth]{images/side-sketch.png}
+    }
+\end{subfigure}    
+\begin{subfigure}[t]{\textwidth}\centering
+    \caption{Frontal view of a bullet land (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
+    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
+    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
+    }
+\end{subfigure}
+\caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
 \end{figure}
@@
 
-In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements at a fixed value of the bullet's height, denoted by $x$.
+In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements at a fixed height $x$ of the bullet.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 @
-\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land across a fixed $x$ value.
+\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at height $x$.
 Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
@@
 %\subsection{Loess Fit}
-Instead of a circular fit, we use multiple loess fits to model the overall structure in order to identify the bullet markings.
+Instead of a circular fit, we use multiple loess fits to model the overall structure and extract the bullet markings.
 
@@
 \begin{enumerate}
-    \item Fix an $x$ value to extract a bullet profile (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
+    \item At a fixed height $x$ extract a bullet's profile (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
     \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \emph{smoothing factor} $s$.
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed cross section ($x$ held constant at 243.75~$\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
+\caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
@@
 \section{Automatic matching}
-Applying the loess fit to a range of different surface cross sections (see Figure~\ref{fig:manualmatch-rgl} for cross sections from 50$\mu m$ to 150~$\mu m$) shows the 3d striation marks from two bullets.
-Signatures of bullet~1 are shown on the left (all cross sections with values below 100$\mu m$) and cross sections of bullets 2 are shown on the right (cross sections with values above 100$\mu m$).
+Applying the loess fit to a range of different signatures (see Figure~\ref{fig:manualmatch-rgl} for signatures extracted at heights from 50$\mu m$ to 150~$\mu m$) shows the 3D striation marks from two bullets.
+Signatures of bullet~1 are shown on the left (all extracted from heights below 100$\mu m$) and signatures of bullets 2 are shown on the right (extracted at heights above 100$\mu m$).
 Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one side to the other.
@@
 \includegraphics[width=.65\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}3d view of the manually adjusted side-by-side comparison of bullet~1-5 and bullet 2-1 after removing the curvature.}
+\caption{\label{fig:manualmatch-rgl}3D view of the manually adjusted side-by-side comparison of bullet~1-5 and bullet 2-1 after removing the curvature.}
 \end{figure}
@@
 The overall high AUC values indicate, that we can successfully employ machine learning methods to distinguish matches from non-matches.
-%Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3d topological surface measurements.
+%Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3D topological surface measurements.
 There are quite large differences in these averages, indicating

### 2016-01-20 imaging-paper.Rnw
'height' instead of cross section up to and including section 3.3

@@
 
-In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section across the surface measurements at a fixed height $x$ of the bullet.
+In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 @
-\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at height $x$.
+\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at a fixed height $x$.
 Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
@@
 \begin{enumerate}
-    \item At a fixed height $x$ extract a bullet's profile (Figure~\ref{fig:loess_step1}, with $x$ set to 243.75$\mu m$).
+    \item At a fixed height $x$ extract a bullet's profile (Figure~\ref{fig:loess_step1}, with $x = 243.75$\mu m$).
     \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \emph{smoothing factor} $s$.
@@
 \section{Automatic matching}
-Applying the loess fit to a range of different signatures (see Figure~\ref{fig:manualmatch-rgl} for signatures extracted at heights from 50$\mu m$ to 150~$\mu m$) shows the 3D striation marks from two bullets.
+Applying the loess fit to a range of different signatures (see Figure~\ref{fig:manualmatch-rgl} for signatures extracted at heights between 50$\mu m$ and 150~$\mu m$) shows the 3D striation marks from two bullets.
 Signatures of bullet~1 are shown on the left (all extracted from heights below 100$\mu m$) and signatures of bullets 2 are shown on the right (extracted at heights above 100$\mu m$).
@@
     \begin{subfigure}[t]{\textwidth}\centering
-\caption{Loess smooth of signatures  at cross section $x = 100$ (span is 0.03).\label{fig:smooth}}{%
+\caption{Loess smooth of signatures  at a height of $x = 100\mu m$ (span is 0.03).\label{fig:smooth}}{%
 <<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement section~\ref{supp:bulletbottom} for a more detailed discussion).
+We first identify a stable region for each bullet land and extract the signature at the lowest height in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement section~\ref{supp:bulletbottom} for a more detailed discussion).
 
@@
 %In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
-In addition to the features listed above, we also keep track of the value of the cross-correlation function found during the alignment, the lag at which signatures show the best alignment, and the cross section levels ($x_1$ and $x_2$).
+In addition to the features listed above, we also keep track of the value of the cross-correlation function found during the alignment, the lag at which signatures show the best alignment, and the height at which signatures are extracted ($x_1$ and $x_2$).
 
@@
 \subsection{Horizontal alignment}
-A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.
-This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$.
+Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$.
 Striation marks show up in this representations as peaks and valleys.
 The individual characteristics are quite prominent and, again, suggest a match between the lands.
-A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities.
+A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities.
 %
@@
 \end{subfigure}
-\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$.
+\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 taken at  heights of $x = 100\mu m$.
 A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
@@
 
-\subsection{Impact of cross sections}\label{sec:heights}
-The cross section on which a comparison between signatures is based matters -- signatures from cross sections that are further apart, show more pronounced differences between the signatures. 
-This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of which cross section level was used in a matching.
-Visually, matches degrade if the signatures upon which the match is based are from cross sections further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion).
-However, we can extract signatures from multiple cross sections of the same bullet land for an initial assessment of its quality. 
+\subsection{Impact of bullet height}\label{sec:heights}
+The height at which signatures are extracted for a comparison between bullet lands matters -- signatures taken from heights that are further apart, show more pronounced differences between the signatures.
+This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of the height that was used in a matching.
+Visually, matches degrade if the signatures upon which the match is based are from heights further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion).
+However, we can extract signatures from multiple heights of the same bullet land for an initial assessment of its quality. 
 This situation is assessed and discussed in more detail below:
@@
 @
-\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet~1-1.
-Initially, the match between signatures from cross sections 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
+\caption{\label{fig:crosscuts2}Signatures  for barrel 3, bullet~1-1 extracted from varying heights.
+Initially, the match between signatures taken at heights 25$\mu m$ apart is affected strongly by some break off at the bottom of the bullet.
 At a  level of $175\mu m$ the bullet's signature stabilizes.
-For this land, matches should not be attempted at lower cross sections.
+For this land, matches should not be attempted at lower heights.
 }
@@
 %An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
-by comparing signatures from cross sections that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
-In the approach here, we keep increasing the height $x$ of the cross section until we find a section with a stable pattern.
+by comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
+In the approach here, we keep increasing the height $x$ at which the signature is taken until we find a section with a stable pattern.
 This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
-`stability' is  defined as two aligned signatures from cross sections chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.
+`stability' is  defined as two aligned signatures from heights chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.
 

### 2016-01-20 imaging-paper.Rnw
proof-reading main text done

@@
 \begin{enumerate}
-    \item At a fixed height $x$ extract a bullet's profile (Figure~\ref{fig:loess_step1}, with $x = 243.75$\mu m$).
+    \item At a fixed height $x$ extract a bullet's profile (Figure~\ref{fig:loess_step1}, with $x = 243.75\mu m$).
     \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \emph{smoothing factor} $s$.
@@
 As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
-For this, we compute the cross-correlation of signatures at cross sections 25$\mu m$ apart.
-For a stable region, we require a minimum of 0.9 for the cross correlation.
+For this, we compute the cross-correlation of signatures extracted from heights 25$\mu m$ apart.
+For a stable region, we require a minimum of 0.95 for the cross correlation.
 Four lands from different bullets are flagged as problematic in this respect.
@@
  run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e.
-we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unknown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons.
+we are comparing $15 \times 6 -2 = 90 - 2 = 88$ lands from unknown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons.
 Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
- When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 lands from known bullets.
+ When things go well, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 lands from known bullets.
 Two lands show a high CMS.
@@
 While the probability for a match increases with the number of maximal CMS, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}.
-Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae.
+Figure~\ref{fig:mismatch} shows a known mismatch between two lands that share twelve consecutively matched striae.
 Visually we can easily tell that these two lands do not match well.
@@
 \caption{\label{fig:mismatch}Known mismatch with a relatively large number of maximal consecutive matching striae (twelve) in the middle.
-The pattern in the middle does look surprisingly similar, however the outer ends of the signatures easily reveals this as mismatch.
+The pattern in the middle does look surprisingly similar, however the outer ends of the signatures easily reveals this comparison as mismatch.
 }
@@
   \centering
-<<oob, echo=FALSE, fig.width=6, fig.height=3, out.width='.6\\textwidth'>>=
+<<oob, echo=FALSE, fig.width=6, fig.height=3, out.width='.7\\textwidth'>>=
 qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
@@
 This makes the errors in the automated approach smaller than the human error in the Hamby study.
-Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.
+Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, corresponding to a rate of 0.0011.
+%\hh{XXX calling this a false negative rate is problematic.
+FEs will be up in arms, because inconclusive doesn't necessarily mean something was missed, it could also mean that there was something wrong with the bullets.
+XXX }
 
@@
 Matching bullets can only be tested realistically in a much bigger experiment.
-Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared.
+Another thing to note about  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared.
 Basing this decision on 0.5 is not necessary.
@@
 
-\hh{Besides including results from known matches against known non-matches, we can increase the number of comparisons in the Hamby study to include all possible land-to-land comparisons.
+Besides including results from known matches against known non-matches, we can increase the number of comparisons in the Hamby study to include all possible land-to-land comparisons.
 This effectively doubles the number of data points available.
 Comparisons not previously included in fitting the random forest can also be used as an additional source for assessing error rates.
-Results for this and a more detailed discussion can be found in Supplement section~\ref{supp:extended}.}
+Results for this and a more detailed discussion can be found in Supplement section~\ref{supp:extended}.
 
@@
 The matching algorithm is sensitive to the parameter choices made.
-The distance between cross sections (currently 25$\mu m$) to evaluate stability, as well as the cross-correlation factor (currently 0.9) we set as a minimum threshold do affect the final outcome.
+The heights at which signatures are extracted (currently 25$\mu m$ apart) to evaluate stability, as well as the cross-correlation factor (currently 0.95) we set as a minimum threshold do affect the final outcome.
 Another parameter that must be selected is the amount of smoothing when identifying peaks and grooves (currently, a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
@@
 
-\hh{XXX speed? XXX}
+%\hh{XXX speed? XXX}
 

### 2016-01-20 imaging-supplement.Rnw
proof-read done for supplement

@@
 \section{Cylindrical Fit}\label{supp:cylindrical}
-Figure~\ref{fig:fixedX} shows the side profile of a crosscut through the surface measurements of bullet 1-5 at a fixed level.
+Figure~\ref{fig:fixedX} shows the  profile of surface measurements of bullet 1-5 at a fixed height.
 The smooth line on top is a circle, with estimated radius and center.
@@
 @
-\caption{\label{fig:fixedX}Side profile of the surface measurements (in $\mu m$) of a bullet land across a fixed $x$ value.
+\caption{\label{fig:fixedX}Side profile of the surface measurements (in $\mu m$) of a bullet land at a fixed height of $x$.
 Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
@@
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
-For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all signatures.
+For a profile at height $x = 100\mu m$ there is a residual circular structure that does not show up for all signatures.
 
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
+\caption{\label{fig:residuala}Residual structure at height $x = 1.5625\mu m$ (bottom of the bullet).}
 <<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{\label{fig:residualb} Residual structure at $x = 100.00$}
+\caption{\label{fig:residualb} Residual structure at height $x = 100.00\mu m$}
 <<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
@@
 A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction.
-Even when we fit a circle at each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues.
+Even when we fit a circle at each distinct height of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues.
 While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.
@@
 \newpage
-\section{Assessing cross-correlation between signatures at multiple cross sections}\label{supp:ccf}
+\section{Assessing cross-correlation between signatures at multiple levels of height}\label{supp:ccf}
 
-Figure~\ref{fig:crosscuts} shows a sequence of signatures for bullet 1-5 (barrel 1) that are taken at cross sections 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$.
-These are compared to the signature at cross section 100$\mu m$.
+Figure~\ref{fig:crosscuts} shows a sequence of signatures for bullet 1-5 (barrel 1) that are taken at heights 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$.
+These are compared to the signature at a height of 100$\mu m$.
 Initially, this comparison constitutes an almost perfect match between the two signatures.
-However, the match quickly deteriorates with increasing distance between the cross sections.
-Only if signatures are from cross sections  within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
-Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections.
+However, the match quickly deteriorates with increasing distance between the heights at which signatures are extracted.
+Only if signatures are from heights  within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used.
+Given that we have to expect some variation in nominally the same height values due to (manual) alignments in microscopes, we should take height values into account in the automatic matching routine by evaluating matches at several heights.
 \begin{figure}[hbtp]
@@
 @
-\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections.
-The signature extracted at $x = 100$ is compared to signatures at every 50$\mu m$.
-With every step away from the original cross section site, the number of differences between the signatures increases; the number of maximum CMS decreases from initially 22 to  four or fewer at $x = 300\mu m$ and higher.
+\caption{\label{fig:crosscuts}Overview of the variations in the signatures at different heights.
+The signature extracted at $x = 100\mu m$ is compared to signatures at every 50$\mu m$.
+With every step away from the original height, the number of differences between the signatures increases; the number of maximum CMS decreases from initially 22 to  four or fewer at a height of $x = 300\mu m$ and above.
 }
@@
 %
-Figure~\ref{fig:overview} shows an overview of the signatures across different crossucts on a single bullet. 
+Figure~\ref{fig:overview} shows an overview of the signatures at different heights on a single bullet. 
 \begin{figure}[hbtp]
@@
 @
-\caption{\label{fig:overview}Signatures of the same bullet at different crosscuts. With increasing crosscut level, peaks and valleys are less pronounced, resulting in a smaller standard deviation.}
+\caption{\label{fig:overview}Signatures of the same bullet at different heights.  With increasing height, peaks and valleys are less pronounced, resulting in a smaller standard deviation.}
 \end{figure}
-As crosscut levels increase individual characteristics become less distinctive, making true matches to other bullets harder.
+At larger heights  individual characteristics become less distinctive, making true matches to other bullets harder.
 The pattern of decreasing peaks and valleys is generally true for  bullet lands, as can be seen in Figure~\ref{fig:sds}.
@@
   geom_smooth(group=1) + 
-  xlab("Crosscut location") +
+  xlab("Height (from the bottom of the bullet)") +
   ylab("Standard deviation")
 @
-\caption{\label{fig:sds}Standard deviation reduces as crosscut levels increase.}
+\caption{\label{fig:sds}Standard deviation reduces as height increase.}
 \end{figure}
-Figure~\ref{fig:sds} shows that the amount of standard deviation of a signature decreases on average for all bullet lands as crosscut levels increase.
+Figure~\ref{fig:sds} shows that the amount of standard deviation of a signature decreases on average for all bullet lands at larger heights.
 This makes standard deviation of a signature one measure to quantify the extent to which a signature is expressed.
-This implies that for identifying matches we should use the lowest crosscut level to extract a bullet's signature once a stable surface region is detected.
+For identifying matches we should therefore use the lowest height to extract a bullet's signature once a stable surface region is detected.
 This is in accordance with current standard practice \citep{afte:1992}.
@@
 \section{Complete evaluation of the Hamby study}\label{supp:extended}
-\hh{One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land.
+One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land.
 This effectively doubles the number of comparisons from 10,384 pairwise comparisons of usable bullet lands to 21,115 $\left[= (118+88)\cdot 205/2\right]$ comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons.
-}
 
-\hh{When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the random forest's estimated OOB error rates of 0.226744 and 0.000098.
-}
+When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the random forest's estimated OOB error rates of 0.226744 and 0.000098.
 
-\hh{However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0.
+However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0.
 The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.
 The false positive rate is therefore virtually unchanged, while we see a slight improvement in the false negative rate for an overall OOB error rate of 0.3\%, i.e.\ an increase to twice the number of comparisons leads to a decrease of 25\% of the estimated error rate.
-This is yet another argument in favor of a larger database for training algorithms.}
+This is yet another argument in favor of a larger database for training algorithms.
 
-\hh{XXX maybe use log of D to remove the left boundary XXX}
+%\hh{XXX maybe use log of D to remove the left boundary XXX}
 <<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=

### 2016-01-20 imaging-paper.Rnw
last massaging

@@
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 
-\hh{ Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}.
+ Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}.
 There, CMS is defined only in terms of matching peaks without regarding valleys.
@@
 In the Hamby study, the definition of CMS by \citet{thompson:2013} leads to approximately half of the values of CMS defined in this paper (with a correlation coefficient between the values of the two definitions of about 0.92).
-}
-For lead bullets, such as used in the Hamby study, \citet{biasotti:1959} considered four or more consecutive lines to be sufficient evidence of a match.
-\hh{Biasotti's four or more consecutive lines translates to eight or more with our definition.
-} %We will assess the claim that eight or more consecutive lines are sufficient for a match in the next section.
+For lead bullets, such as used in the Hamby study, \citet{biasotti:1959} considered four or more consecutive peaks (corresponding to eight or more consecutive lines in our definition) to be sufficient evidence of a match.
+%We will assess the claim that eight or more consecutive lines are sufficient for a match in the next section.
 
@@
   \centering
-<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.65\\textwidth'>>=
+<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\textwidth'>>=
 library(rpart)
@@
 \section*{Acknowledgment} 
-%Thanks to David Baldwin for pointing us to the NIST database and introducing us to the basic terms.
+Thanks to David Baldwin for pointing us to the NIST database and doing a Ballistics 101 for us.
 Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.

### 2016-01-21 imaging-paper.tex
Include the submission folder

--- 2016-01-20 imaging-paper.Rnw
@@
-\documentclass[aoas, preprint]{imsart}
+\documentclass[aoas, preprint]{imsart}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \endlocaldefs
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 %\newpage
-<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
-#rm(list=ls())
-#wd <- getwd()
-library(extrafont)
-library(knitr)
-imgdir <- "Figure/"
-codedir <- "code"
-datadir <- "app/images/Hamby252_3DX3P1of2/"
-
-options(replace.assign=TRUE,scipen=3, digits=2)
-bstats <- read.csv("data-25-25/bullet-stats.csv")
-
-scrubPath <- function(x) {
-  splits <- strsplit(as.character(x), split="/")
-  last <- sapply(splits, function(x) x[length(x)])
-  gsub(".x3p","", last)
-}
-
-library(RColorBrewer)
-library(ggplot2)
-library(scales)
-library(dplyr)
-library(x3pr)
-library(x3prplus)
-library(grid)
-library(gridExtra)
-library(zoo)
-library(tidyr)
-library(rpart)
-library(rpart.plot)
-library(xtable)
-library(sm)
-@
 
@@
   \centering
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.65\\textwidth'>>=
-
-cols = c(alpha("grey60", alpha=0.6), alpha("black", 0.5))
-
-br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
-dbr111 <- fortify_x3p(br111)
-
-pars <- data.frame(getCircle(dbr111$y, dbr111$value))
-dbr111$theta <- acos((dbr111$y-pars$x0)/pars$radius)/pi*180
-dbr111 <- dbr111 %>% mutate(
-  xpred = cos(theta/180*pi)*pars$radius + pars$x0,
-  ypred = sin(theta/180*pi)*pars$radius + pars$y0
-)
-
-qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 99*1.5625^2), y, value, geom="line", size=I(1)) +
-  geom_line(aes(x=xpred, y=ypred, group=x), 
-            colour="grey30", size=0.25) +
-#  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
-  ylab("") +
-  theme_bw() + 
-  theme(legend.position="bottom") #+ coord_equal()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=0.65\textwidth]{figure/fixedX-1}
+
+\end{knitrout}
 \caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at a fixed height $x$.
@@
 \caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
-<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
-br111_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
-br121_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
-br122_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
-br123_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
-br124_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
-br125_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
-br126_bullet <- get_crosscut(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))
-
-p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
-p1
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.5\textwidth]{figure/loess_step1-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 Valleys and peaks are detected, if they are not within the same window.}
-<<loess_step3, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
-subdata <- br111_bullet
-smoothfactor <- 35
-
-value_filled <- na.fill(subdata$value, "extend")
-smoothed <- rollapply(value_filled, smoothfactor, function(x) mean(x))
-smoothed_truefalse <- rollapply(smoothed, smoothfactor, function(x) mean(x))
-
-xloc=250
-p3 <- ggplot() +
-  annotate("rect",xmin=xloc-17*1.5625, xmax=xloc+18*1.5625, ymin=80, ymax=200, fill=alpha("orange", alpha=0.5)) + 
-  geom_point(aes(x=(1:length(smoothed_truefalse))*1.5625, y=smoothed_truefalse)) + theme_bw() +
-    geom_vline(xintercept=xloc, colour="grey50") + xlab("y") + ylab("Surface measurements")
-
-p3
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.5\textwidth]{figure/loess_step3-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 Values outside the blue boundaries are removed (shown in grey)}
-<<loess_step47, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
-    lengthdiff <- length(subdata$value) - length(smoothed_truefalse)
-
-    peak_ind_smoothed <- head(which(rollapply(smoothed_truefalse, 3, function(x) which.max(x) == 2)), n = 1)
-    peak_ind <- peak_ind_smoothed + floor(lengthdiff / 2)
-    groove_ind <- head(which(rollapply(tail(smoothed_truefalse, n = -peak_ind_smoothed), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind
-
-    peak_ind2_smoothed_temp <- head(which(rollapply(rev(smoothed_truefalse), 3, function(x) which.max(x) == 2)), n = 1)
-    peak_ind2_temp <- peak_ind2_smoothed_temp + floor(lengthdiff / 2)
-    groove_ind2_temp <- head(which(rollapply(tail(rev(smoothed_truefalse), n = -peak_ind2_smoothed_temp), 3, function(x) which.min(x) == 2)), n = 1) + peak_ind2_temp
-
-    peak_ind2 <- length(subdata$value) - peak_ind2_temp + 1
-    groove_ind2 <- length(subdata$value) - groove_ind2_temp + 1
-
-    p4 <- qplot(subdata$y[subdata$y < subdata$y[groove_ind]], subdata$value[subdata$y < subdata$y[groove_ind]], colour=I("grey60"), alpha=I(.25)) +
-        ylim(c(min(subdata$value[!is.nan(subdata$value)]) - 25, max(subdata$value[!is.nan(subdata$value)]) + 25)) +
-        xlim(c(min(subdata$y) - 25, max(subdata$y) + 25)) +
-        geom_point(aes(x, y), data = data.frame(x = subdata$y[subdata$y > subdata$y[groove_ind2]], y = subdata$value[subdata$y > subdata$y[groove_ind2]]), colour=I("grey60"), alpha=I(.25)) +
-        geom_point(inherit.aes = FALSE, aes(x, y), data = data.frame(x = subdata$y[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]], y = subdata$value[subdata$y < subdata$y[groove_ind2] & subdata$y > subdata$y[groove_ind]])) +
-        theme_bw() +
-        geom_vline(xintercept = subdata$y[peak_ind], colour = "red") +
-        geom_vline(xintercept = subdata$y[groove_ind], colour = "blue") +
-        geom_vline(xintercept = subdata$y[peak_ind2], colour = "red") +
-        geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue") +
-      xlab("y") + ylab("Surface measurements")
-
-    p4
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.5\textwidth]{figure/loess_step47-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 \caption{\label{fig:loess_fit} Loess fit for bullet~1-5.}
-<<loess_fit, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE,  message=FALSE>>=
-
-br111.groove <- get_grooves(br111_bullet)
-
-my.loess <- fit_loess(br111_bullet, br111.groove)
-
-my.loess$fitted + ylab("Surface measurements")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/loess_fit-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 %@
-<<boot_loess, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
-boot.loess <- boot_fit_loess(br111_bullet, br111.groove)
-poly <- with(boot.loess, data.frame(x = c(y,rev(y)), y=c(high, rev(low))))
-qplot(x=x, y=y, geom="polygon", fill=I("steelblue"), data=poly) +
-  geom_line(aes(x=y, y=resid), data=my.loess$data) +
-  ylab("Residuals of loess fit") + xlab("y") + theme_bw()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/boot_loess-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 \subsection{Algorithm}\label{sec:algorithm}
-<<twolands100, cache=TRUE,  echo=FALSE>>=
-images <- file.path(datadir, dir(datadir))
-
-#lof <- processBullets(paths = images[c(5,7)], x = 100)
-subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
-subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
-#subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-#subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
-lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
-
-# smooth
-lofX <- lofX %>% group_by(bullet) %>% mutate(
-  l30 = smoothloess(y, resid, span = 0.03)
-)
-
-# cut at .75
-threshold <- .75
-lofX$r05 <- threshold* sign(lofX$l30) * as.numeric(abs(lofX$l30) > threshold)
-lofX$type <- factor(lofX$r05)
-levels(lofX$type) <- c("groove", NA, "peak")
-@
-
-<<twolands100adjust, echo=FALSE>>=
-images <- file.path(datadir, dir(datadir))
-
-#lof <- processBullets(paths = images[c(5,7)], x = 100)
-subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
-subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
-#subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
-#subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
-subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-ccf <- ccf(subLOFx1$resid, subLOFx2$resid, lag.max=100, plot=FALSE)
-lag <- ccf$lag[which.max(ccf$acf)]
-subLOFx1$y <- subLOFx1$y - lag*1.5625 
-lofY <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
-
-@
-
-<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
-matches <- lofX %>% group_by(y) %>% summarise(
-  potential = (length(unique(type)) == 1),
-  allnas = sum(is.na(type))/n(),
-  type1 = na.omit(type)[1],
-  type = paste(type, sep="|", collapse="|"),
-  n = n()
-)
-
-matches$id <- cumsum(matches$allnas == 1) + 1
-matches$lineid <- as.numeric(matches$allnas != 1) * matches$id
-
-isMatch <- function(id, type) {
-  if (id[1] == 0) return(FALSE)
-#  browser()
-  types <- strsplit(type, split = "|", fixed=TRUE) 
-  t1 <- sapply(types, function(x) x[1])
-  t2 <- sapply(types, function(x) x[2])
-  if (all(t1 == "NA")) return(FALSE)
-  if (all(is.na(t2))) return(FALSE)
-  if (all(t2 == "NA")) return(FALSE)
-
-  peak <- length(grep("peak", c(t1, t2))) > 0
-  groove <- length(grep("groove", c(t1, t2))) > 0
-  if (peak & groove) return(FALSE)
-
-  return(TRUE)
-}
-
-lines <- matches %>% group_by(lineid) %>% summarise(
-  meany = mean(y, na.rm=T),
-  miny = min(y, na.rm=T),
-  maxy = max(y, na.rm=T) + 1.5625,
-  match = isMatch(lineid, type),
-  type = type1[1]
-)
-lines <- subset(lines, lineid != 0)
-@
 
@@
 \caption{Loess smooth of signatures  at a height of $x = 100\mu m$ (span is 0.03).\label{fig:smooth}}{%
-<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
-lofX$bulletname <- scrubPath(lofX$bullet)
-qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bulletname) +
-  geom_line(aes(y=l30), colour="grey70", size=0.5) +
-  facet_grid(bulletname~.) +
-#  scale_colour_manual("", values=cols) +
-  theme_bw() + ylab("") + 
-  theme(legend.position="none")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.65\textwidth]{figure/smooth-1}
+
+\end{knitrout}
     }
@@
 \caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
-<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
-#images <- file.path(datadir, dir(datadir))
-  #lof <- processBullets(paths = images[c(5,7)], x = 100)
-
-  subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
-  subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
-  lof <- rbind(subLOFx1, subLOFx2)
-
-  lof <- bulletSmooth(lof)
-  bAlign = bulletAlign(lof)
-  lofX <- bAlign$bullet
-
-  b12 <- unique(lof$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-  peaks <- rbind(peaks1$lines, peaks2$lines)
-
-  peaks$bulletname <- scrubPath(peaks$bullet)
-  lofX$bulletname <- scrubPath(lofX$bullet)
-  ggplot() + theme_bw() +
-    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
-              data=peaks,  alpha=0.2) +
-    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
-               data= peaks, alpha=0.7) +
-    scale_colour_brewer(palette="Set2") + 
-    scale_fill_brewer(palette="Set2") +
-    theme(legend.position="none") + 
-    facet_grid(bulletname~.) +
-    geom_line(aes(x=y, y=l30, group=bulletname), data=lofX) +
-  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.65\textwidth]{figure/smoothcut-1}
+
+\end{knitrout}
 }
@@
 \label{fig:smoothcutd}}{%
-<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.65\\textwidth', warning=FALSE>>=
-peaks1$lines$bullet <- b12[1]
-peaks2$lines$bullet <- b12[2]
-
-lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-ggplot() + 
-  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
-  theme(legend.position="bottom") +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
-  ylim(c(-6,6.5)) + theme_bw() +
-  geom_line(data=lofX, aes(x=y, y=l30, group=bulletname, linetype=bulletname)) +
-  scale_linetype_discrete("") +
-  scale_colour_manual("", values=cols) +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
-    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
-  xlab("y")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.65\textwidth]{figure/smoothmatch-1}
+
+\end{knitrout}
     }
@@
 \caption{Raw bullet land signatures.\label{fig:crosscut}}{%
-<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
-lof$bulletname <- scrubPath(lof$bullet)
-qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
-      geom="line") + 
-  scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab("Residuals of loess fit")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/crosscuts-1}
+
+\end{knitrout}
     }
@@
     \caption{Aligned signatures.\label{fig:crosscutX}}{%
-<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
-lofY$bulletname <- scrubPath(lofY$bullet)
-qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
-      geom="line") + xlab("y") +
-  scale_colour_manual("", values=cols) + theme_bw() +
-  theme(legend.position = c(1, 1.1), legend.justification=c("right", "top"),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab("Residuals of loess fit")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/crosscutsX-1}
+
+\end{knitrout}
     }
@@
   \centering
-<<ccf, echo=FALSE, fig.width = 8, fig.height = 3, out.width = '.65\\textwidth'>>=
-lag <- ccf$lag[which.max(ccf$acf)]
-qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
-  theme_bw() + ylab("Correlation") + xlab("Lag k") +
-  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
-  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
-                     minor_breaks = c(-75,-25,25, 75))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.65\textwidth]{figure/ccf-1}
+
+\end{knitrout}
 \caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
@@
   \centering
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
-paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100,250, by=25)
-lof <- processBullets(read.x3p(paths[37]), name = paths[37], x = crosscuts)
-lof$bullet <- paste(scrubPath(lof$bullet), lof$x)
-
-reslist <- lapply(1:length(crosscuts[-1]), function(i) {
-  b2 <- subset(lof, x %in% crosscuts[i:(i+1)])
-  lofX <- bulletSmooth(b2)
-  bAlign <- bulletAlign(lofX)
-  lofX <- bAlign$bullet
-
-  b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
-
-#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-#  lines <- striation_identify(lofX, threshold = threshold)
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
-})
-
-ccfs <- sapply(reslist, function(res) res$ccf)
-
-lop <- lapply(reslist, function(res) {
-ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5.5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),   data = res$bullets) +
-  scale_linetype_discrete("") +
-#  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6.5)) + ylab("") + xlab("") + 
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
-})
-
-grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
-             lop[[4]], lop[[5]], lop[[6]],
-             ncol=2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/crosscuts-vary-b31-1}
+
+\end{knitrout}
 \caption{\label{fig:crosscuts2}Signatures  for barrel 3, bullet~1-1 extracted from varying heights.
@@
   \centering
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
-br111.groove <- get_grooves(br111_bullet)
-br111.loess <- fit_loess(br111_bullet, br111.groove)
-
-peakslist <- lapply(c(5, 25, 45), function(s) {
-  peaks <- get_peaks(br111.loess$data, smoothfactor = s)
-  peaks$lines$smoothfactor = s
-  peaks$plot$data$smoothfactor = s
-
-  peaks
-})
-
-peaks <- plyr::ldply(peakslist, function(x) x$lines)
-smooths <- plyr::ldply(peakslist, function(x) x$plot$data)
-
-ggplot() + theme_bw() +
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
-            data=peaks,  alpha=0.2) +
-  geom_vline(aes(xintercept=extrema, colour=factor(type)), data= peaks, alpha=.7) +
-  scale_colour_brewer(palette="Set2") + 
-  scale_fill_brewer(palette="Set2") +
-  theme(legend.position="none") + 
-  facet_grid(smoothfactor~., labeller="label_both") +
-  geom_line(aes(x=y, y=resid, group=x), data=br111.loess$data,  
-            colour="grey50") +
-  geom_line(aes(x=y, y=smoothed), data=smooths) +
-  ylab(expression(paste("Signature values (in ",mu,"m)", sep=""))) 
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.5\textwidth]{figure/smoothfac1-1}
+
+\end{knitrout}
 \caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively.
@@
 \caption{Maximal number of CMS between unknown bullet C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
-<<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
-load("data-25-25/unkn9.RData")
-cmsdist <- sapply(reslist, function(x) x$maxCMS)
-
-qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.5\textwidth]{figure/cms-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 \caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
-<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
-res <- reslist[[which.max(cmsdist)]]  
-#res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
-res$bullets$bullet <- scrubPath(res$bullets$bullet)
-
-print(ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
-  scale_linetype_discrete("") + ylab("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/top-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 \caption{Top 2 match with C-3 based on CMS.}
-<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
-res <- reslist[[which(cmsdist==14)]]  # number 2
-res$bullets$bullet <- scrubPath(res$bullets$bullet)
-
-print(ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
-  scale_linetype_discrete("") + ylab("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/top2-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 \begin{minipage}[t]{.47\textwidth}
-<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
-ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(CMS))) + theme_bw() +
-  theme(legend.position="bottom") + 
-  xlab("maximum CMS")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/cms-bars-1}
+
+\end{knitrout}
 %\end{subfigure}
@@
 \begin{minipage}[t]{.52\textwidth}
-<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='\\textwidth'>>=
-bstats$km <- c( "KNM", "KM")[as.numeric(bstats$match)+1]
-ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(CMS), fill=km), position="fill") +
-  theme(legend.position="bottom") + 
-  scale_fill_brewer("", palette="Paired") +
-  xlab("maximum CMS") + ylab("Proportion")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/cms-spines-1}
+
+\end{knitrout}
 \end{minipage}
@@
 There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches.
-On average, there are \Sexpr{mean(bstats$CMS[!bstats$match], na.rm=T)} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$CMS[!bstats$match])}).
-Known matches share on average \Sexpr{mean(bstats$CMS[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$CMS[bstats$match])}.
+On average, there are 2.31 maximal CMS between known non-matches (with a standard deviation of 1.4).
+Known matches share on average 8.49 maximal CMS, with a standard deviation of 5.65.
 While the probability for a match increases with the number of maximal CMS, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}.
@@
   \centering
-<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.65\\textwidth', warning=FALSE>>=
-load("data-25-25/unkn47.RData")
-res <- reslist[[106]]
-res$bullets$bullet <- scrubPath(res$bullets$bullet)
-
-print(ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
-  geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
-  scale_linetype_discrete("") + ylab("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.65\textwidth]{figure/strange-res-1}
+
+\end{knitrout}
 \caption{\label{fig:mismatch}Known mismatch with a relatively large number of maximal consecutive matching striae (twelve) in the middle.
@@
   \centering
-<<density-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
-features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
-densities <- plyr::ldply(features, function(x) {
-  xr <- range(bstats[,x])
-  xx <- seq(xr[1], xr[2], length.out=500)
-
-  densKM <- sm.density(bstats[,x][bstats$match], display="none", 
-                       eval.points=xx, weights=NA, method="normal")
-  densKNM <- sm.density(bstats[,x][!bstats$match], display="none", 
-                       eval.points=xx, weights=NA, method="normal")
-  dframe <- data.frame(var=x, x = xx, KM=densKM$estimate, KNM = densKNM$estimate)
-  dframe <- rbind(data.frame(var=x, x = xr[1], KM=0, KNM=0),
-        dframe, data.frame(var=x, x = xr[2], KM=0, KNM=0))
-  dframe$order <- 1:nrow(dframe)
-  dframe
-}) 
-library(reshape2)
-dm <- melt(densities, measure.var=c("KM", "KNM"))
-dm$var <- factor(dm$var, levels=features)
-levels(dm$var)[3:4] <- c("#matches", "#non-matches")
-levels(dm$variable) <- c("Known matches (KM)", "Known non-matches (KNM)")
-
-qplot(x, value, group=variable, geom="polygon", data=dm, fill=variable, 
-      alpha=I(0.6), colour=I("grey20")) + 
-  facet_wrap(~var, ncol=4, scales="free") +
-  theme_bw() + ylab("density") + xlab("") +
-  scale_fill_brewer("Matches", palette="Paired") +
-  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
-        legend.background = element_rect(colour="grey75"),
-        plot.margin = unit(c(0,0,0,0), unit="cm"))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/density-overview-1}
+
+\end{knitrout}
 \caption{\label{fig:densities}Overview of all the marginal densities for features described in section~\ref{sec:algorithm}.
@@
   \centering
-<<rocs-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
-# plot false positive against false negative
-# prob of false positive: P(match | KNM)
-# prob of false negative: P(non-match | KM)
-
-# for  X < c: we get probability for P(X < c | KNM) and P(X < c | KM)
-
-features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
-errors <- plyr::ldply(features, function(x) {
-  xx <- unique(bstats[,x])
-  if (length(xx) > 500) {
-    xr <- range(bstats[,x])
-    xx <- seq(xr[1], xr[2], length.out=500)
-  }
-  # upper and lower rule:
-  # upper: match is defined as X > xx
-  # lower: match is defined as X < xx
-  subbstats <- subset(bstats, flagged == FALSE)
-  subbstats$match <- factor(subbstats$match)
-
-  errors <- plyr::ldply(xx, function(cc) {
-    idx <- which(subbstats[,x] >= cc)
-    upper <- data.frame(xtabs(~match, data= subbstats[idx,,drop=FALSE])/
-                          xtabs(~match, data= subbstats))
-    idx <- which(subbstats[,x] <= cc)
-    lower <- data.frame(xtabs(~match, data= subbstats[idx,,drop=FALSE])/
-                          xtabs(~match, data= subbstats))
-
-    data.frame(value = cc, match=lower[,1], lower=lower[,2], upper=upper[,2])
-  })
-
-  errors$variable <- x
-  errors
-})
-
-rocs.lower <- dcast(errors, variable+value~match, value.var="lower")
-rocs.lower$type <- "lower"
-rocs.upper <- dcast(errors, variable+value~match, value.var="upper")
-rocs.upper$type <- "upper"
-rocs <- rbind(rocs.lower, rocs.upper)
-
-set.seed(20140105)
-aucs <- plyr::ldply(features, function(x) {
-  subbstats <- subset(bstats, flagged == FALSE)
-  pos.scores <- sample(subbstats[which(subbstats$match),x], 50000, replace=TRUE)
-  neg.scores <- sample(subbstats[which(!subbstats$match),x], 50000, replace=TRUE)
-  data.frame(x, upper=mean(pos.scores > neg.scores), 
-             lower = mean(pos.scores < neg.scores))
-})
-aucs <- melt(aucs, measure.var=c("lower", "upper"))
-names(aucs) <- c("variable", "type", "auc")
-rocs <- merge(rocs, aucs, by=c("variable", "type"))
-rocs$variable <- reorder(rocs$variable, -rocs$auc, min)
-levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")
-
-labels <- unique(subset(rocs, auc > 0.5)[,c("variable", "auc")])
-labels$labels <- paste("AUC:", labels$auc)
-labels$type <- NA
-
-ggplot(data=subset(rocs, auc > 0.5)) +
-  geom_ribbon(aes(x=`FALSE`, ymax=`TRUE`, ymin=0), alpha=0.2) +
-  geom_line(aes(x=`FALSE`,y=`TRUE`))+ theme_bw() + xlab("False negative rate") + 
-  ylab("False positive rate") + facet_wrap(facets=~variable, ncol=4) + 
-  geom_label(x=1, y=0, aes(label=labels), data=labels, hjust=1, vjust=0) + 
-  geom_abline(linetype=2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/rocs-overview-1}
+
+\end{knitrout}
 \caption{\label{fig:rocs}ROC curves for all of the features described in section~\ref{sec:algorithm}.
@@
 Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$).
-%The average distance between known matches in the Hamby study is $\Sexpr{mean(bstats$D[bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[bstats$match])}\mu m$; while known mismatches have an average  distance of  $\Sexpr{mean(bstats$D[!bstats$match])}\mu m$ with a standard deviation of $\Sexpr{sd(bstats$D[!bstats$match])}\mu m$.
+%The average distance between known matches in the Hamby study is $1.86\mu m$ with a standard deviation of $2.31\mu m$; while known mismatches have an average  distance of  $3.27\mu m$ with a standard deviation of $1.93\mu m$.
 All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches.
@@
 The resulting tree is shown in Figure~\ref{fig:tree}.
-A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} lands is being matched correctly.
+A total of 132 lands is being matched correctly.
 Interestingly, the number of consecutive matching striae does not feature in this evaluation.
@@
   \centering
-<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\textwidth'>>=
-library(rpart)
-library(rpart.plot)
-library(RColorBrewer)
-vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
-names(bstats)[9:10] <- c("#matches", "#non-matches")
-includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
-cc <- read.csv("csvs/crosscuts-25-new.csv")
-# cc$bullet <- gsub(".*//","",cc$path)
-# cc$bullet <- gsub(".x3p", "", cc$bullet)
-# bstats$flagged <- bstats$b2 %in% cc$bullet[which(is.na(cc$cc))] |
-#                   bstats$b1 %in% cc$bullet[which(is.na(cc$cc))]
-
-#excludesObs <- which(bstats$flagged)
-
-rp1 <- rpart(match~., subset(bstats, !flagged)[,includesVar])  # doesn't include cms at all !!!!
-
-per <- rp1$frame$yval # predictive probability for each node in the tree
-
-prp(rp1, extra = 101, box.col=vals[as.numeric(per > 0.5)+1])
-
-bstats$pred <- predict(rp1, newdata=bstats)
-bstats$pred[bstats$flagged] <- NA
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.7\textwidth]{figure/tree-1}
+
+\end{knitrout}
 \caption{\label{fig:tree}Decision tree of matching bullets based on recursive partitioning.
@@
 \end{figure}
-
-<<rforest, echo=FALSE, message=FALSE>>=
-library(randomForest)
-library(reshape2)
-
-set.seed(20160105)
-names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
-includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
-
-rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=300)
-
-errors <- data.frame(rtrees$err.rate)
-errors$id <- 1:nrow(errors)
-bstats$forest <- predict(rtrees, type="prob", newdata=bstats)[,2]
-bstats$forest[bstats$flagged] <- NA
-@
 
@@
   \centering
-<<oob, echo=FALSE, fig.width=6, fig.height=3, out.width='.7\\textwidth'>>=
-qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.7\textwidth]{figure/oob-1}
+
+\end{knitrout}
 \caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
@@
   \centering
-<<tree:forest, echo=FALSE, fig.width=5, fig.height=2.5, out.width='.6\\textwidth', warning = FALSE>>=
-predm <- melt(bstats, measure.var=c("pred", "forest"))
-levels(predm$variable) <- c("Tree", "Forest")
-predm$match <- c("KNM", "KM")[as.numeric(predm$match)+1]
-qplot(data=predm, value, match, geom="jitter", colour=match, shape=match) + theme_bw() + facet_grid(variable~.) +
-    scale_colour_brewer(palette="Paired", labels=c("KNM", "KM")) +
-  scale_shape_discrete(labels=c("KNM", "KM")) + ylab("") + 
-  xlab("Predicted probability of a match") +
-  theme(legend.position="none")
-
-# qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
-#       shape=match) + theme_bw() +
-#   theme(legend.position="bottom") +
-#   geom_vline(xintercept=0.5, colour="grey50") +
-#   geom_hline(yintercept=0.5, colour="grey50") + 
-#   xlab("Predictions from decision tree") +
-#   ylab("Predictions from random forest")   +
-#   scale_colour_brewer(palette="Paired", labels=c("KNM", "KM")) +
-#   scale_shape_discrete(labels=c("KNM", "KM"))
-# 
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.6\textwidth]{figure/tree:forest-1}
+
+\end{knitrout}
 \caption{\label{fig:tree-forest}Prediction results from the tree and the forest.
@@
 \begin{figure}
-<<echo=FALSE, fig.width=7, fig.height=2, out.width='.7\\textwidth'>>=
-imp <- data.frame(importance(rtrees))
-imp$Variable <- row.names(imp)
-imp$Variable[c(7,8)] <- c("#matches", "#non-matches")
-imp1 <- subset(imp, !(Variable %in% c("x1", "sd.D", "S")))
-imp2 <- subset(imp, Variable %in% c("x1", "sd.D", "S"))
-qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
-  geom_text(aes(y=1.2, label=Variable), angle=45, hjust=0, data=imp1) + 
-  geom_text(aes(y=0.8, label=Variable), angle=-45, hjust=0, data=imp2) + 
-  ylim(c(0, 2.75)) + theme_bw() + ylab("") +
-  theme(axis.ticks.y=element_blank(), axis.text.y=element_blank(),
-        panel.grid.major.y=element_blank(), panel.grid.minor.y=element_blank()) +
-  xlab("Importance (mean decrease in Gini index)")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.7\textwidth]{figure/unnamed-chunk-1-1}
+
+\end{knitrout}
 \caption{\label{fig:importance}Importance of features in the random forest. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree.}

### 2016-01-21 imaging-supplement.tex
Include the submission folder

--- 2016-01-20 imaging-supplement.Rnw
@@
 % natbib option: authoryear
-\documentclass[aoas]{imsart}
+\documentclass[aoas]{imsart}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \newpage
-<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
-#rm(list=ls())
-#wd <- getwd()
-library(extrafont)
-library(knitr)
-imgdir <- "Figure/"
-codedir <- "code"
-datadir <- "app/images/Hamby252_3DX3P1of2/"
-
-options(replace.assign=TRUE,scipen=3, digits=2)
-bstats <- read.csv("data-25-25/bullet-stats.csv")
-
-scrubPath <- function(x) {
-  splits <- strsplit(as.character(x), split="/")
-  last <- sapply(splits, function(x) x[length(x)])
-  gsub(".x3p","", last)
-}
-
-library(RColorBrewer)
-library(ggplot2)
-library(scales)
-library(dplyr)
-library(x3pr)
-library(x3prplus)
-library(grid)
-library(gridExtra)
-library(zoo)
-library(tidyr)
-library(rpart)
-library(rpart.plot)
-library(xtable)
-library(sm)
-
-@
 
@@
   \centering
-<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=0.5\textwidth]{figure/fixedX-1}
 
-cols = c(alpha("grey60", alpha=0.6), alpha("black", 0.5))
-
-br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
-dbr111 <- fortify_x3p(br111)
-
-pars <- data.frame(getCircle(dbr111$y, dbr111$value))
-dbr111$theta <- acos((dbr111$y-pars$x0)/pars$radius)/pi*180
-dbr111 <- dbr111 %>% mutate(
-  xpred = cos(theta/180*pi)*pars$radius + pars$x0,
-  ypred = sin(theta/180*pi)*pars$radius + pars$y0
-)
-
-qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 99*1.5625^2), y, value, geom="line", size=I(1)) +
-  geom_line(aes(x=xpred, y=ypred, group=x), 
-            colour="grey30", size=0.25) +
-#  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
-  ylab("") +
-  theme_bw() + 
-  theme(legend.position="bottom") #+ coord_equal()
-@
+\end{knitrout}
 \caption{\label{fig:fixedX}Side profile of the surface measurements (in $\mu m$) of a bullet land at a fixed height of $x$.
@@
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
-In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.
-Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
+In this instance, the radius is estimated as $\hat{r} = 4666.49\mu m = 4.67mm$ and the land covers about 29.5 degrees.  Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
 The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
@@
 \caption{\label{fig:residuala}Residual structure at height $x = 1.5625\mu m$ (bottom of the bullet).}
-<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
-qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
-      geom="line", size=I(1)) +
-#  scale_colour_brewer("x", palette="Paired") + 
-  theme_bw() + 
-  geom_hline(yintercept = 0, colour="grey50") +
-  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
-  theme(legend.position="bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/residual2-1}
+
+\end{knitrout}
 \end{subfigure}    
@@
 \caption{\label{fig:residualb} Residual structure at height $x = 100.00\mu m$}
-<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
-#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
-qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
-      geom="line", size=I(1)) +
-#  scale_colour_brewer("x", palette="Paired") + 
-  geom_hline(yintercept = 0, colour="grey50") +
-  theme_bw() + 
-  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
-  theme(legend.position="bottom")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/residual-1}
+
+\end{knitrout}
 \end{subfigure}
@@
 
-<<bullet1, echo=FALSE, cache=TRUE>>=
-db1 <- NULL
-for (i in 1:6) {
-  bname <- sprintf(file.path(datadir, "Br1 Bullet 1-%d.x3p"), i)
-  dbi <- fortify_x3p(read.x3p(bname))
-  dbi$part <- i
-  db1 <- rbind(db1, dbi)
-}
-
-db1 <- db1 %>% group_by(part, x) %>% do (
-    data.frame(., predCircle(.$y, .$value))
-  )
-@
-
-<<bullet2, echo=FALSE, cache=TRUE>>=
-db2 <- NULL
-for (i in 1:6) {
-  bname <- sprintf(file.path(datadir, "Br1 Bullet 2-%d.x3p"), i)
-  dbi <- fortify_x3p(read.x3p(bname))
-  dbi$part <- i
-  db2 <- rbind(db2, dbi)
-}
-
-db2 <- db2 %>% group_by(part, x) %>% do (
-    data.frame(., predCircle(.$y, .$value))
-  )
-
-@
-
 \begin{figure}[hbtp]
   \centering
-<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
-db2$land <- db2$part
-db1$land <- db1$part
-qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
-      geom="line", size=I(.75), colour=I("grey70")) + 
-  facet_wrap(~land, ncol=3, labeller="label_both") + 
-  scale_colour_brewer(palette="Paired") +
-  theme_bw() + theme(legend.position="bottom") + 
-  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
-             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
-  ylab("Residuals from circular fit") +
-  ggtitle("Bullet 1-5 in black")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/circlefits-1}
+
+\end{knitrout}
 \caption{Circular fit to the signature of each land of bullet 2, with signature from bullet 1-5 overlaid.\label{fig:circlefits} The signature of bullet 1-5 matches best with bullet 2-1.}
@@
   \centering
-<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\textwidth]{figure/crosscuts-vary-1}
 
-scrubPath <- function(x) {
-  splits <- strsplit(as.character(x), split="/")
-  last <- sapply(splits, function(x) x[length(x)])
-  gsub(".x3p","", last)
-}
-
-paths <- file.path(datadir, dir(datadir))
-crosscuts <- seq(100, 400, by = 50)
-lof <- processBullets(read.x3p(paths[5]), name = scrubPath(paths[5]), x = crosscuts)
-lof$bullet <- paste(lof$bullet, lof$x)
-
-reslist <- lapply(crosscuts[-1], function(cc) {
-#  browser()
-  b2 <- subset(lof, x %in% c(cc, 100))
-  lofX <- bulletSmooth(b2)
-  bAlign = bulletAlign(lofX)
-  lofX <- bAlign$bullet
-    b12 <- unique(b2$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)
-
-#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
-#  lines <- striation_identify(lofX, threshold = threshold)
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
-})
-
-ccfs <- sapply(reslist, function(res) res$ccf)
-
-lop <- lapply(reslist, function(res) {
-ggplot() +
-  theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
-  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
-  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
-  scale_linetype_discrete("") +
-  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
-  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
-        legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylim(c(-6,6)) +
-  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
-    ylab("") + xlab("")
-})
-
-grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
-             ncol = 2)
-@
+\end{knitrout}
 \caption{\label{fig:crosscuts}Overview of the variations in the signatures at different heights.
@@
 \section{Signature intensities}\label{supp:bulletbottom}
-<<setup-signatures, echo=FALSE, message=FALSE>>=
-library(x3pr)
-library(x3prplus)
-library(dplyr)
-library(ggplot2)
-library(gridExtra)
-library(zoo)
 
-knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
-
-###############
-# can we identify the barrels the unknown bullets came from?
-
-new_processBullets <- function(bullet, x = 100) {
-  crosscuts <- unique(fortify_x3p(bullet)$x)
-  crosscuts <- crosscuts[crosscuts >= min(x)]
-  crosscuts <- crosscuts[crosscuts <= max(x)]
-  if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]
-
-  list_of_fits <- lapply(crosscuts, function(x) {
-    br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
-    br111.groove <- get_grooves(br111)
-    fit_loess(br111, br111.groove)$resid$data
-  })
-  lof <- list_of_fits %>% bind_rows
-
-  cbind(lof, bullet = bullet$path)
-}
-
-new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
-  lof <- rbind(lof1, lof2)
-  bAlign = bulletAlign(lof)
-  lofX <- bAlign$bullet
-
-  b12 <- unique(lof$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)
-
-  #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
-  #    theme_bw() +
-  #    geom_rect(data=peaks1$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25)) +
-  #    geom_rect(data=peaks2$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25))
-
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  #   p <- qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
-  #     theme_bw() +
-  #     geom_rect(data=lines, aes(xmin=xmin, xmax=xmax, fill = factor(type)),  ymin=-6, ymax=6, inherit.aes = FALSE, alpha=I(0.25)) +
-  #     ylim(c(-6,6)) +
-  #     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match), inherit.aes = FALSE) +
-  #     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match), inherit.aes = FALSE)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lag=bAlign$lag, lines=lines, bullets=lofX)
-}
-
-# match unknown land using crosscuts
-ccs <- read.csv("csvs/crosscuts-25.csv")
-all_bullets <- lapply(as.character(ccs$path), function(x) {
-  result <- read.x3p(x)
-  result[[3]] <- x
-  names(result)[3] <- "path"
-
-  return(result)
-})
-
-knowns <- all_bullets[1:120]
-unknowns <- all_bullets[121:210]
-
-if (!file.exists("csvs/crosscut-sd.csv")) {
-  bullets_processed <- lapply(all_bullets, function(bul) {
-    cat("Computing processed bullet", basename(bul$path), "\n")
-    crosscuts <- 25*(1:20)
-    dframe <- new_processBullets(bullet = bul, x = crosscuts)
-    dframe$bullet <- with(dframe, paste(bullet, x))
-    dframe
-  })
-  names(bullets_processed) <- as.character(ccs$path)
-
-  bullets_smoothed <- bullets_processed %>% bind_rows %>% bulletSmooth
-
-  stats <- bullets_smoothed %>% group_by(bullet, x) %>% summarize(
-    sd = sd(l30, na.rm=T)
-  )
-  splits <- strsplit(stats$bullet, split=".", fixed=TRUE)
-  bnames <- sapply(splits, function(x) paste(x[1],".x3p", sep=""))
-  stats$bullet <- bnames
-
-  write.csv(stats, "csvs/crosscut-sd.csv", row.names=FALSE)
-} else {
-  stats <- read.csv("csvs/crosscut-sd.csv")
-}  
-@
 %
@@
 \begin{figure}[hbtp]
-<<one-bulletland-sd, echo=FALSE, dependson='setup-signatures', fig.height=5, fig.width=8, out.width='0.7\\textwidth', warning=FALSE>>=
-k <- 1
-x <- as.character(ccs$path)[k]
-land <- read.x3p(x)
-land[[3]] <- x
-names(land)[3] <- "path"
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=0.7\textwidth]{figure/one-bulletland-sd-1}
 
-crosscuts <- 25*(1:20)
-dframe <- new_processBullets(bullet = land, x = crosscuts)
-dframe$bullet <- with(dframe, paste(bullet, x))
-
-bullet <- dframe %>% bind_rows %>% bulletSmooth 
-subbullet <- subset(bullet, x >= ccs$cc[k])
-qplot(x=y, geom="line", y=l30, colour=x, group=x, data=subbullet) + theme_bw() +
-  scale_colour_gradient("Crosscuts") +
-  theme(legend.position="bottom") +
-  ylab("Residuals from loess fit")
-@
+\end{knitrout}
 \caption{\label{fig:overview}Signatures of the same bullet at different heights.  With increasing height, peaks and valleys are less pronounced, resulting in a smaller standard deviation.}
@@
 \centering
-<<crosscuts-sd, echo=FALSE, fig.height=5, fig.width=8, out.width='.7\\textwidth', dependson='setup-signatures'>>=
-qplot(x, sd, data=stats, geom="line", group=bullet, alpha=I(0.5)) + theme_bw() +
-  geom_smooth(group=1) + 
-  xlab("Height (from the bottom of the bullet)") +
-  ylab("Standard deviation")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.7\textwidth]{figure/crosscuts-sd-1}
+
+\end{knitrout}
 \caption{\label{fig:sds}Standard deviation reduces as height increase.}
@@
 %\hh{XXX maybe use log of D to remove the left boundary XXX}
-<<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=
-knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-knowns <- dir(path=knowndatadir)
-knowns <- gsub(".x3p", "", knowns)
 
-unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
-unknowns <- dir(path=unknowndatadir)
-unknowns <- gsub(".x3p", "", unknowns)
-
-#bstats <- read.csv("data-25-25/bullet-stats.csv")
-#flagged <- c("Br6 Bullet 2-1", "Br9 Bullet 2-4", "Ukn Bullet B-2", "Ukn Bullet Q-4")
-ballstats <- read.csv("data-new-all-25-25/bullet-stats-single.csv")
-
-includesVar <- setdiff(names(ballstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "id", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km", "forestsmall",
-                                        "left_cms", "right_cms", "left_noncms", "right_noncms"))
-
-library(randomForest)
-#set.seed(20151202)
-#rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=100)
-
-# names(ballstats)[10] <- "num.nonmatches"
-# names(ballstats)[11] <- "CNMS"
-# names(ballstats)[16] <- "S"
-# names(ballstats)[17] <- "CMS"
-#ballstats$forestsmall <- predict(rtrees, newdata=ballstats, type="prob")[,2]
-###########
-
-ballstats$insmall <- with(ballstats, ((b1 %in% knowns) & (b2 %in% unknowns)) | ((b2 %in% knowns) & (b1 %in% unknowns)))
-
-set.seed(20160512)
-rtrees2 <- randomForest(factor(match)~., data=subset(ballstats, !flagged & insmall)[,includesVar], ntree=300)
-
-ballstats$forestsmall <- predict(rtrees2, newdata=ballstats, type="prob")[,2]
-
-# xtabs(~(forestsmall>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
-
-set.seed(20151201)
-rtrees3 <- randomForest(factor(match)~., data=subset(ballstats, !flagged)[,includesVar], ntree=300)
-ballstats$forest <- predict(rtrees3, newdata=ballstats, type="prob")[,2]
-#xtabs(~(forest>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
-#qplot(forest, data=subset(ballstats, !flagged), geom="jitter", y=match)
-@
 % 9/(10628+9)
@@
 
-<<aligned, echo=FALSE, warning=FALSE,  cache=TRUE>>=
-matches <- read.csv("csvs/matches.csv", header = FALSE)
-matches$barrel <- rep(1:10, each=6)
-matches$id <- 1:nrow(matches)
-library(reshape2)
-mm <- melt(matches, id.var=c("barrel", "id"), na.rm=TRUE)
-mm <- subset(mm, value != "")
-mm$value <- gsub("Br.* Bullet ", "", mm$value)
-mm$prefix <- sprintf("Br%d Bullet ", mm$barrel)
-unkns <- grep("[A-Z]",mm$value)
-mm$prefix[unkns] <- "Ukn Bullet "
-mm$path <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P1of2//"
-mm$path[unkns] <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P2of2//"
-mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))
-
-ccs <- read.csv("csvs/crosscuts-25-new.csv")
-mm <- merge(mm, ccs, by="path", all=TRUE)
-mm <- subset(mm, !is.na(mm$cc))
-
-crosscuts <- plyr::ldply(
-  1:nrow(mm), 
-  function(i) {
-    dframe <- get_crosscut(mm$path[i], mm$cc[i])
-    groove <- get_grooves(dframe)
-    data.frame(mm[i,], fit_loess(dframe, groove)$resid$data)
-  })
-
-crossSmooth <- bulletSmooth(crosscuts)
-
-alignme <- function(data) {
-#  browser()
-  data$bullet <- as.character(data$bullet)
-  bullets <- unique(data$bullet)
-  b1 <- bullets[1]
-  lofs <- NULL
-  for (i in bullets[-1]) {
-    lof <- subset(data, bullet %in% c(b1, i))
-    lofs <- rbind(lofs, bulletAlign(lof)$bullet)
-  }
-  lofs <- unique(lofs)
-  lofs
-}
-
-aligned <- crossSmooth %>% group_by(id) %>% do(
-  data.frame(alignme(.))
-)
-aligned$land <- (aligned$id %% 6)
-aligned$land[aligned$land==0] <- 6
-aligned$bulletland <- aligned$bullet
-aligned$bullet <- gsub("-[0-9]","",aligned$bullet)
-library(RColorBrewer)
-set.seed(20160106)
-cols <- rep(brewer.pal(n=10, name="Paired"), length=35)[sample(35, 35, replace=TRUE)]
-labels <- aligned %>% group_by(bulletland) %>% summarize(
-  x = max(y),
-  y = min(l30, na.rm=T),
-  label = gsub(" Bullet","",bulletland[1]),
-  barrel = barrel[1],
-  land = land[1],
-  bullet = bullet[1]
-)
-labels <- labels %>% group_by(barrel, land) %>% mutate(id=order(bulletland))
-@
-
 \begin{landscape}
 \begin{figure}[hbtp]
-<<aligned-first, dependson='aligned', echo=FALSE, warning=FALSE, fig.width=15, fig.height=10, out.width='7.5in'>>=
-qplot(x=y, y=l30, data=subset(aligned, barrel <=5), 
-      geom="line", group=bullet, colour=bullet) + 
-  ylim(c(-7,7)) + xlab("") + ylab("") + xlim(c(0,3000)) +
-  facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
-  theme(plot.margin=unit(c(0,0,-1,-1), unit="line"), legend.position="none") + 
-  geom_label(aes(y=2.5*(id-3), label=label, colour=bullet), 
-             x = 3000, data=subset(labels, barrel <= 5), 
-             inherit.aes = FALSE, hjust="right", size=3, 
-             fill=alpha("white", .5)) +
-  scale_colour_manual(values=cols)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=7.5in]{figure/aligned-first-1}
+
+\end{knitrout}
 \caption{\label{fig:aligned}Overview of aligned signatures for all bullet lands for  barrels 1 to 5 of the Hamby study. }
@@
 \begin{figure}[hbtp]
-<<aligned-second, dependson='aligned', echo=FALSE, warning=FALSE, fig.width=15, fig.height=10, out.width='7.5in'>>=
-qplot(x=y, y=l30, data=subset(aligned, barrel > 5), 
-      geom="line", group=bullet, colour=bullet) + 
-  ylim(c(-7,7)) + xlab("") + ylab("") + xlim(c(0,3000)) +
-  facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
-  theme(plot.margin=unit(c(0,0,-1,-1), unit="line"), legend.position="none") + 
-  geom_label(aes(y=2.5*(id-3), label=label, colour=bullet), 
-             x = 3000, data=subset(labels, barrel > 5), 
-             inherit.aes = FALSE, hjust="right", size=3) +
-  scale_colour_manual(values=cols)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=7.5in]{figure/aligned-second-1}
+
+\end{knitrout}
 \caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet lands for  barrels 5 to 10 of the Hamby study. }
@@
 Importance~1 refers to the smaller subset, Importance~2 is the feature importance derived from the random forest based on all pairwise comparisons.
-<<setup-rtrees, echo=FALSE, message=FALSE>>=
-bstats <- read.csv("data-25-25/bullet-stats.csv", stringsAsFactors = FALSE)
-
-library(randomForest)
-set.seed(20151202)
-includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
-
-rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=300)
-@
 
@@
 \centering
-<<importance, echo=FALSE, results='asis', warning=FALSE>>=
+% latex table generated in R 3.2.3 by xtable 1.8-0 package
+% Wed Jan 20 18:06:55 2016
+\begin{tabular}{clrrrd{3.2}rd{3.2}}
+  \hline
+ & Variable & Importance 1 & Importance 2 & KM & (sd) & KNM & (sd) \\ 
+  \hline
+1 & ccf & 87.0 & 134.6 & 0.7 & ( 0.25) & 0.3 & (   0.10) \\ 
+  2 & \#matches & 81.9 & 128.3 & 15.5 & ( 7.91) & 4.3 & (   2.51) \\ 
+  3 & S & 46.4 & 53.6 & 18.3 & ( 8.95) & 5.5 & (   3.41) \\ 
+  4 & \#non-matches & 35.9 & 62.7 & 9.8 & ( 5.80) & 18.8 & (   3.92) \\ 
+  5 & D & 26.1 & 45.7 & 1.9 & ( 2.32) & 3.3 & (   1.94) \\ 
+  6 & CMS & 15.9 & 25.1 & 8.5 & ( 5.65) & 2.3 & (   1.40) \\ 
+  7 & sd.D & 15.0 & 23.6 & 2.9 & ( 0.80) & 2.9 & (   0.70) \\ 
+  8 & CNMS & 13.3 & 20.9 & 4.8 & ( 4.10) & 10.2 & (   4.35) \\ 
+  9 & lag & 9.9 & 17.0 & 200.8 & (48.29) & 204.0 & ( 113.63) \\ 
+  10 & x2 & 4.2 & 6.2 & 166.5 & (81.07) & 163.5 & (  72.30) \\ 
+  11 & x1 & 3.9 & 6.1 & 175.0 & (71.59) & 175.0 & (  71.40) \\ 
+   \hline
+\end{tabular}
 
-imp2 <- data.frame(importance(rtrees3))
-imp2$Variable <- row.names(imp2)
-names(imp2)[1] <- "Bigforest"
-
-imp <- data.frame(importance(rtrees))
-imp$Variable <- row.names(imp2)
-
-imp <- merge(imp, imp2, by="Variable")
-
-imp <- imp[order(-imp$MeanDecreaseGini),]
-names(imp)[2:3] <- c("Importance 1", "Importance 2")
-
-bstats <- subset(ballstats, insmall == TRUE)
-imp$meanMatch <- sapply(imp$Variable, function(var) {
-  mean(subset(bstats, match)[,var])
-})
-imp$sdMatch <- sapply(imp$Variable, function(var) {
-  sd(subset(bstats, match)[,var])
-})
-imp$meanNonMatch <- sapply(imp$Variable, function(var) {
-  mean(subset(bstats, !match)[,var])
-})
-imp$sdNonMatch <- sapply(imp$Variable, function(var) {
-  sd(subset(bstats, !match)[,var])
-})
-
-imp$meanMatch <- round(imp$meanMatch,1)
-imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
-imp$meanNonMatch <- round(imp$meanNonMatch,1)
-imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
-row.names(imp) <- 1:nrow(imp)
-imp$Variable[c(2,3,4,6,8)] <- c("#matches", "S", "#non-matches", "CMS", "CNMS")
-names(imp)[-(1:3)] <- c("KM", "(sd)", "KNM", "(sd)")
-print(xtable(imp, align="clrrrd{3.2}rd{3.2}", digits=c(0,0,1,1,1,2,1,2)), 
-      floating = FALSE)
-@
 \end{table}

### 2016-04-05 imaging-paper.Rnw
First set of edits

@@
 % provide arXiv number if available:
-\arxiv{arXiv:0000.0000}
+\arxiv{arXiv:1601.05788}
 
@@
 3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
-x3p is a container format conforming to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
-x3p was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+The x3p format conforms to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
+It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
 
@@
 Next, we fit a loess regression to the data.
-Loess regression \citep{cleveland:1979} is based on the assumption that the relationship between two random variables $X$ and $Y$ can be described in form of a smooth, continuous function $f$ with $y_i = f(x_i) + \varepsilon_i$ for all values $i = 1, ..., n$.
+Loess regression \citep{cleveland:1979} is based on the assumption that the relationship between two random variables $X$ and $Y$ can be described in the form of a smooth, continuous function $f$ with $y_i = f(x_i) + \varepsilon_i$ for all values $i = 1, ..., n$.
 The function $f$ is approximated via locally weighted polynomial regressions.
@@
 The main idea of locally weighted regression is to use a weighting routine that  emphasizes the effect of points close by and de-emphasizes the effect of points as they are further away.
-The weighting function used here, is the tricubic function $w(d) = \left(1 - d^3\right)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
+The weighting function used here is the tricubic function $w(d) = \left(1 - d^3\right)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
 Here, $d$ is defined as the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
@@
 Applying the loess fit to a range of different signatures (see Figure~\ref{fig:manualmatch-rgl} for signatures extracted at heights between 50$\mu m$ and 150~$\mu m$) shows the 3D striation marks from two bullets.
-Signatures of bullet~1 are shown on the left (all extracted from heights below 100$\mu m$) and signatures of bullets 2 are shown on the right (extracted at heights above 100$\mu m$).
+Signatures of bullet~1 are shown on the left (all extracted from heights below 100$\mu m$) and signatures of bullet 2 are shown on the right (extracted at heights above 100$\mu m$).
 Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one side to the other.

### 2016-04-05 imaging-paper.Rnw
Fix all the build issues

@@
 In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination.
-Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
+Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel.
 During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet.
@@
 options(replace.assign=TRUE,scipen=3, digits=2)
-bstats <- read.csv("data-25-25/bullet-stats.csv")
+bstats <- read.csv("data-25-25/bullet-stats-old.csv")
 
@@
 
-lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+lines <- striation_identify(peaks1$lines, peaks2$lines)
 
@@
 However, we can extract signatures from multiple heights of the same bullet land for an initial assessment of its quality.
-This situation is assessed and discussed in more detail below:
+By comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
+In the approach here, we keep increasing the height $x$ at which the signature is taken until we find a section with a stable pattern.
+This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where `stability' is  defined as two aligned signatures from heights chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.
 
@@
   peaks2$lines$bullet <- b12[2]
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
+  lines <- striation_identify(peaks1$lines, peaks2$lines)
 
@@
 %An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following:
-by comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
-In the approach here, we keep increasing the height $x$ at which the signature is taken until we find a section with a stable pattern.
-This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet~1-1 from barrel 3, where
-`stability' is  defined as two aligned signatures from heights chosen 25$\mu m$ apart having a cross-correlation of at least 0.95.
 
@@
 It results in a significant amount of noise in the data such that even just a point or two can skew the rolling average enough for a peak or valley to be detected.
-Given that the striation patterns are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
+Given that striation widths are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
 Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artifacts in the analysis.
@@
 
-Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches.
+%Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches.
 Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$).
@@
 The maximal number of CMS is only in the seventh position here.
-The overall high AUC values indicate, that we can successfully employ machine learning methods to distinguish matches from non-matches.
+The overall high AUC values indicate that we can successfully employ machine learning methods to distinguish matches from non-matches.
 %Table~\ref{tab:importance} shows an overview in the averages and corresponding standard deviations of the set of feature variables derived in section~\ref{sec:algorithm} from the 3D topological surface measurements.
@@
 The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}.
-However, in this study authors did not consider the number of matches and non-matches.
+However, in that study, authors did not consider the number of matches and non-matches.
 
@@
 includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
-cc <- read.csv("csvs/crosscuts-25-new.csv")
+cc <- read.csv("csvs/crosscuts-25.csv")
 # cc$bullet <- gsub(".*//","",cc$path)
@@
 The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
-The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found.
+The bottom line of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found.
 The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
@@
 A further re-assessment by \citet{deceuster:2015} came to the same conclusions based on the EvoFinder system.
-The NRC report also found, that results from 2D images can be improved when matches are based on 3D images.
+The NRC report also found that results from 2D images can be improved when matches are based on 3D images.
 This is consistent with the importance of features found here: out of the top five features (see Figure~\ref{fig:importance}), only the total number of matches and mismatches are available for a match based on 2D features.

### 2016-04-05 imaging-paper.tex
Include the arXiv id
Fix all the build issues

@@
 
-Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches.
+%Similarly, the Euclidean distance $D$ between two signatures is strongly correlated to known matches.
 Out of the 48 pairs of land-to-land signatures with a distance of less than .25, 47 correspond to known matches (the single known mismatch consists of a pair of signatures that both are very flat with peaks and valleys of less than 1.5$\mu m$).
-%The average distance between known matches in the Hamby study is $1.86\mu m$ with a standard deviation of $2.31\mu m$; while known mismatches have an average  distance of  $3.27\mu m$ with a standard deviation of $1.93\mu m$.
+%The average distance between known matches in the Hamby study is $mean(bstats$D[bstats$match])\mu m$ with a standard deviation of $sd(bstats$D[bstats$match])\mu m$; while known mismatches have an average  distance of  $mean(bstats$D[!bstats$match])\mu m$ with a standard deviation of $sd(bstats$D[!bstats$match])\mu m$.
 All of the features in Figure~\ref{fig:densities} show large, if not significant, differences between matches and non-matches.

### 2016-04-05 imaging-supplement.Rnw
Fix all the build issues

@@
 
-new_processBullets <- function(bullet, x = 100) {
-  crosscuts <- unique(fortify_x3p(bullet)$x)
-  crosscuts <- crosscuts[crosscuts >= min(x)]
-  crosscuts <- crosscuts[crosscuts <= max(x)]
-  if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]
-
-  list_of_fits <- lapply(crosscuts, function(x) {
-    br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
-    br111.groove <- get_grooves(br111)
-    fit_loess(br111, br111.groove)$resid$data
-  })
-  lof <- list_of_fits %>% bind_rows
-
-  cbind(lof, bullet = bullet$path)
-}
-
-new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
-  lof <- rbind(lof1, lof2)
-  bAlign = bulletAlign(lof)
-  lofX <- bAlign$bullet
-
-  b12 <- unique(lof$bullet)
-  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
-  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)
-
-  #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
-  #    theme_bw() +
-  #    geom_rect(data=peaks1$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25)) +
-  #    geom_rect(data=peaks2$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25))
-
-  peaks1$lines$bullet <- b12[1]
-  peaks2$lines$bullet <- b12[2]
-
-  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)
-
-  #   p <- qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
-  #     theme_bw() +
-  #     geom_rect(data=lines, aes(xmin=xmin, xmax=xmax, fill = factor(type)),  ymin=-6, ymax=6, inherit.aes = FALSE, alpha=I(0.25)) +
-  #     ylim(c(-6,6)) +
-  #     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match), inherit.aes = FALSE) +
-  #     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match), inherit.aes = FALSE)
-
-  maxCMS <- maxCMS(lines$match==TRUE)
-  list(maxCMS = maxCMS, ccf = bAlign$ccf, lag=bAlign$lag, lines=lines, bullets=lofX)
-}
-
 # match unknown land using crosscuts
@@
 
-if (!file.exists("csvs/crosscut-sd.csv")) {
+if (!file.exists("csvs/crosscuts-sd.csv")) {
   bullets_processed <- lapply(all_bullets, function(bul) {
-    cat("Computing processed bullet", basename(bul$path), "\n")
+    #cat("Computing processed bullet", basename(bul$path), "\n")
     crosscuts <- 25*(1:20)
-    dframe <- new_processBullets(bullet = bul, x = crosscuts)
+    dframe <- processBullets(bullet = bul, x = crosscuts)
     dframe$bullet <- with(dframe, paste(bullet, x))
@@
 
-  write.csv(stats, "csvs/crosscut-sd.csv", row.names=FALSE)
+  write.csv(stats, "csvs/crosscuts-sd.csv", row.names=FALSE)
 } else {
-  stats <- read.csv("csvs/crosscut-sd.csv")
+  stats <- read.csv("csvs/crosscuts-sd.csv")
 }
@@
 crosscuts <- 25*(1:20)
-dframe <- new_processBullets(bullet = land, x = crosscuts)
+dframe <- processBullets(bullet = land, x = crosscuts)
 dframe$bullet <- with(dframe, paste(bullet, x))
@@
 
-However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0.
+However, if we use all of the available comparisons to fit another random forest of 300 trees, the defacto error rates for false positives and false negatives are again at 0.
 The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.
@@
 mm$prefix[unkns] <- "Ukn Bullet "
-mm$path <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P1of2//"
-mm$path[unkns] <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P2of2//"
+mm$path <- "app/images/Hamby252_3DX3P1of2//"
+mm$path[unkns] <- "app/images/Hamby252_3DX3P2of2//"
 mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))
 
-ccs <- read.csv("csvs/crosscuts-25-new.csv")
+ccs <- read.csv("csvs/crosscuts-25.csv")
 mm <- merge(mm, ccs, by="path", all=TRUE)
@@
 
+crosscuts$bullet <- crosscuts$path
 crossSmooth <- bulletSmooth(crosscuts)
@@
 <<setup-rtrees, echo=FALSE, message=FALSE>>=
-bstats <- read.csv("data-25-25/bullet-stats.csv", stringsAsFactors = FALSE)
+bstats <- read.csv("data-25-25/bullet-stats-old.csv", stringsAsFactors = FALSE)
 

### 2016-04-05 imaging-supplement.tex
Include the arXiv id
Fix all the build issues

@@
 \centering
-% latex table generated in R 3.2.3 by xtable 1.8-0 package
-% Wed Jan 20 18:06:55 2016
+% latex table generated in R 3.2.4 by xtable 1.8-2 package
+% Tue Apr  5 08:28:49 2016
 \begin{tabular}{clrrrd{3.2}rd{3.2}}

### 2016-04-06 mean-profiles.Rmd
mean profiles

---
title: "extracting mean profiles"
author: "Heike Hofmann"
date: "April 6, 2016"
output: html_document
---

Extracting the mean profile is based on the really nice paper by Chu et al (2010).

```{r, echo=FALSE, warning=FALSE}

library(x3pr)
library(x3prplus)
library(ggplot2)
library(dplyr)

br115 <- "~/CSAFE/Bullets/Hamby252_3DX3P1of2/Br1 Bullet 2-1.x3p"
im115 <- fortify_x3p(read.x3pplus(br115))
qplot(y, x, fill=value, data=im115, geom="tile")
ggsave(file="im115.pdf")

gg115 <- processBullets(read.x3pplus(br115), x=unique(im115$x))
qplot(y, x, fill=resid, data=gg115, geom="tile") + scale_fill_gradient2(limits=c(-7,7))
ggsave(file="im115-resid.pdf")

# get the 'twist'
xs <- unique(gg115$x)
twist <- NULL
ccf <- NULL
aligned <- list()

for (i in seq_along(xs)[-1]) {
  profiles <- subset(gg115, x %in% xs[c(i,i-1)])  
  profiles$bullet <- sprintf("br1-1-5-%s",profiles$x)
  aligned[[i]] <- bulletAlign(profiles, value="resid")
  twist <- c(twist, aligned[[i]]$lag)
  ccf <- c(ccf, aligned[[i]]$ccf)
}

#qplot(xs, c(0, twist)) + ylim(c(-10,10))
#qplot(xs, c(1, ccf)) + geom_hline(yintercept=0.95, colour="red")
qplot(xs, cumsum(c(0, twist))) +ylim(c(700,1100))
dframe <- data.frame(x = xs, twist=cumsum(c(0, twist)), ccf = c(0,ccf))
qplot(x, twist, data=subset(dframe, between(x, 220, 600))) +geom_smooth(method="lm")

summary(m115 <- lm(twist~x, data=subset(dframe, between(x, 220, 600))))

gg115 <- merge(gg115, dframe, by="x")
qplot(y-twist, x, fill=resid, data=gg115, geom="tile") + scale_fill_gradient2(limits=c(-7,7))
ggsave(file="im115-untwist.pdf")

gg115$modeltwist <- predict(m115, newdata=gg115)
# round to precision of y:
gg115$roundtwist <- 1.5625*round(gg115$modeltwist/1.5625, 0)

qplot(y-roundtwist, x, fill=resid, data=gg115, geom="tile") + scale_fill_gradient2(limits=c(-7,7))
ggsave(file="im115-roundtwist.pdf")

qplot(x, ccf, data=dframe) + geom_hline(yintercept=0.975)

mean_profile_all <- gg115 %>% filter(ccf > .975) %>% group_by(y) %>%
  summarize(
    resid = mean(resid)
  )

mean_profile_500 <- gg115 %>% filter(ccf > .975, between(x,50,500)) %>% group_by(y) %>%
  summarize(
    resid = mean(resid)
  )

mean_profile_300 <- gg115 %>% filter(ccf > .975, between(x,50,300)) %>% group_by(y) %>%
  summarize(
    resid = mean(resid)
  )
```

in black: average of all, in blue average based on profiles below 500, in green average based on profiles below 300, in red profile at 100.
```{r, echo=FALSE}
qplot(y, resid, data=mean_profile_all, geom="line") + theme_bw() +
  geom_line(aes(y, resid), data=mean_profile_300, colour="green") +
  geom_line(aes(y, resid), data=mean_profile_500, colour="blue") +
  geom_line(aes(y, resid), data=gg115 %>% filter(x==100), colour="red") + ylim(c(-5,5))

```

### 2016-04-07 mean-profiles.Rmd
bitof cleaning up in the script

@@
   summarize(
-    resid = mean(resid)
+    resid = mean(resid),
+    type="mean-all"
   )
 
-mean_profile_500 <- gg115 %>% filter(ccf > .975, between(x,50,500)) %>% group_by(y) %>%
+mean_high_200 <- gg115 %>% filter(ccf > .975, between(x,50,500)) %>% group_by(y) %>%
   summarize(
-    resid = mean(resid)
+    resid = mean(resid),
+    type="mean-high-200"
   )
 
-mean_profile_300 <- gg115 %>% filter(ccf > .975, between(x,50,300)) %>% group_by(y) %>%
+mean_low_200 <- gg115 %>% filter(ccf > .975, between(x,50,375)) %>% group_by(y) %>%
   summarize(
-    resid = mean(resid)
+    resid = mean(resid),
+    type="mean-low-200"
   )
+
+profiles <- gg115 %>% filter(x==100) %>% select(y, resid) %>%  transform(type="profile-at-100")
+profiles <- rbind(profiles, mean_profile_all, 
+                 mean_high_200, mean_low_200)
 ```
@@
 ```{r, echo=FALSE}
-qplot(y, resid, data=mean_profile_all, geom="line") + theme_bw() +
-  geom_line(aes(y, resid), data=mean_profile_300, colour="green") +
-  geom_line(aes(y, resid), data=mean_profile_500, colour="blue") +
-  geom_line(aes(y, resid), data=gg115 %>% filter(x==100), colour="red") + ylim(c(-5,5))
-
+qplot(y, resid, data=profiles, geom="line", group=type, colour=type) + theme_bw() + ylim(c(-5,5))
 ```

### 2016-04-07 mean-profiles.Rmd
the rollapply seems to do a reasonable job for estimating the twist

@@
 
+Rs <- data.frame(zoo::rollapply(data=dframe$twist, width=200, FUN=function(twist) {
+  x <- 1:length(twist)
+  m <- lm(twist~x)
+  data.frame(r.squared=summary(m)$r.squared, twist=coef(m)[2])
+}, by=1))
+
+twist <- median(subset(Rs, r.squared > .8)$twist)
+
 summary(m115 <- lm(twist~x, data=subset(dframe, between(x, 220, 600))))
@@
 gg115$modeltwist <- predict(m115, newdata=gg115)
+gg115$modeltwist2 <- gg115$x*twist
 # round to precision of y:
 gg115$roundtwist <- 1.5625*round(gg115$modeltwist/1.5625, 0)
+gg115$roundtwist2 <- 1.5625*round(gg115$modeltwist2/1.5625, 0)
 
@@
 ggsave(file="im115-roundtwist.pdf")
+
+qplot(y-roundtwist2, x, fill=resid, data=gg115, geom="tile") + scale_fill_gradient2(limits=c(-7,7))
 

### 2016-04-08 mean-profiles.Rmd
getting the twist rate calculated from median

@@
 
-br115 <- "~/CSAFE/Bullets/Hamby252_3DX3P1of2/Br1 Bullet 2-1.x3p"
+br115 <- "~/CSAFE/Bullets/Hamby252_3DX3P1of2/Br1 Bullet 1-5.x3p"
 im115 <- fortify_x3p(read.x3pplus(br115))

### 2016-04-13 imaging-paper.Rnw
EER location marked on ROC

@@
 labels <- unique(subset(rocs, auc > 0.5)[,c("variable", "auc")])
-labels$labels <- paste("AUC:", labels$auc)
+labels$labels <- sprintf("AUC: %.2f", labels$auc)
 labels$type <- NA
+
+eer <- rocs %>% filter(auc > 0.5) %>% 
+  mutate(differror = abs(1-`TRUE` - `FALSE`))
+eer <- eer %>% group_by(variable) %>% mutate(
+  minerror = differror==min(differror),
+  labels = sprintf("EER: %.2f", 0.5*(`FALSE`+1-`TRUE`))
+) %>% filter(minerror==TRUE)
 
@@
   geom_label(x=1, y=0, aes(label=labels), data=labels, hjust=1, vjust=0) + 
-  geom_abline(linetype=2)
+  geom_abline(linetype=2) + 
+  geom_point(aes(x=`FALSE`, y=`TRUE`), data=eer, size=2.5) +
+  geom_label(x=1, y=0.125, aes(label=labels), data=eer, hjust=1, vjust=0)
 @
@@
 Variables are sorted according to their area under the curve (AUC).
-Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS.
+\hh{The equal error rate (EER) is marked by a point on the ROC curve.} Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS.
 }
@@
 The features are arranged in descending order according to the area under the curve (AUC).
+\hh{XXX The dots mark the equal error rate, i.e.
+the location on the ROC curve, where false positive and false negative error rates are the same.
+The smaller the value, the better.
+We see that in this instance a lowe EER goes hand in hand with high predictive power as measured in AUC.
+XXX HH: do we need to say more about EERs?}
 The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys.

### 2016-04-13 imaging-paper.tex
EER location marked on ROC

@@
 Variables are sorted according to their area under the curve (AUC).
-Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS.
+\hh{The equal error rate (EER) is marked by a point.} Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS.
 }

### 2016-04-13 mean-profile-papers/hamby-twistrates.R
constrained twist rates are done now for the original hamby study

@@ -1,50 +1,0 @@
-library(XML)
-ptm <- proc.time()
-
-# Loop through the vector, adding one
-twist <- getTwist(path="~/CSAFE/Bullets/Hamby252_3DX3P1of2/Br1 Bullet 1-5.x3p")
-# Stop the clock
-proc.time() - ptm
-
-library(dplyr)
-
-knowns <- dir("~/CSAFE/Bullets/Hamby252_3DX3P1of2/", pattern="x3p", full.names=TRUE)
-unknowns <- dir("~/CSAFE/Bullets/Hamby252_3DX3P2of2/", pattern="x3p", full.names=TRUE)
-
-pattern <- "Bullet B"
-pattern <- "Bullet C"
-pattern <- "Bullet D"
-pattern <- "Bullet E"
-pattern <- "Bullet [FHJ]"
-pattern <- "Bullet [LMQ]"
-pattern <- "Bullet [SUX]"
-pattern <- "Bullet [YZ]"
-pattern <- "Br[12] "
-pattern <- "Br[34] "
-pattern <- "Br[56] "
-pattern <- "Br[78] "
-pattern <- "Br9 "
-pattern <- "Br10 "
-twistlist <- c(knowns, unknowns) %>% 
-  grep(pattern=pattern, x=., value=TRUE) %>% 
-  purrr::map_df(getTwist)
-# .id only keeps track of the order of the results at the moment. that's useless.
-twistlist$path <- c(knowns, unknowns) %>% 
-  grep(pattern=pattern, x=., value=TRUE)
-
-twistlist$bulletland <- with(twistlist, gsub(".*//(.*).x3p", "\\1", path))
-
-file <- "mean-profile-papers/twistRobust.csv"
-fileexist = file.exists(file)
-write.table(twistlist[,c(4,3,1,2)], file=file, sep=",", col.names=!fileexist,
-            row.names=FALSE, append=fileexist)
-
-
-twists <- read.csv("mean-profile-papers/twists.csv")
-qplot(twist, min.r.squared, data=twists)
-twists$bullet <- gsub("(.*)-[0-9]*","\\1",twists$bulletland)
-qplot(twist, min.r.squared, data=twists) + facet_wrap(~bullet)
-summary(twists$twist)
-twists$constraint <- "none"
-twists$method <- "q75"
-write.csv(twists, "mean-profile-papers/twists.csv", row.names=FALSE)

### 2016-04-13 code/untwist-hamby.R
untwisting code

@@ -1,0 +1,48 @@
+library(XML)
+ptm <- proc.time()
+
+# Loop through the vector, adding one
+twist <- getTwist(path="~/CSAFE/Bullets/Hamby252_3DX3P1of2/Br1 Bullet 1-5.x3p")
+# Stop the clock
+proc.time() - ptm
+
+library(dplyr)
+
+knowns <- dir("~/CSAFE/Bullets/Hamby252_3DX3P1of2/", pattern="x3p", full.names=TRUE)
+unknowns <- dir("~/CSAFE/Bullets/Hamby252_3DX3P2of2/", pattern="x3p", full.names=TRUE)
+
+pattern <- "Bullet B"
+pattern <- "Bullet C"
+pattern <- "Bullet [DE]"
+pattern <- "Bullet [FHJ]"
+pattern <- "Bullet [LMQ]"
+pattern <- "Bullet [SUX]"
+pattern <- "Bullet [YZ]"
+pattern <- "Br[12] "
+pattern <- "Br[345] "
+pattern <- "Br[678] "
+pattern <- "Br9 "
+pattern <- "Br10 "
+twistlist <- c(knowns, unknowns) %>% 
+  grep(pattern=pattern, x=., value=TRUE) %>% 
+  purrr::map_df(getTwist, twistlimit=c(-2,0)*1.5625)
+# .id only keeps track of the order of the results at the moment. that's useless.
+twistlist$path <- c(knowns, unknowns) %>% 
+  grep(pattern=pattern, x=., value=TRUE)
+
+twistlist$bulletland <- with(twistlist, gsub(".*//(.*).x3p", "\\1", path))
+
+file <- "/Users/heike/papers/2015-imaging-paper/mean-profile-papers/twists-constrained.csv"
+fileexist = file.exists(file)
+write.table(twistlist, file=file, sep=",", col.names=!fileexist,
+            row.names=FALSE, append=fileexist)
+
+
+twists <- read.csv("mean-profile-papers/twists.csv")
+qplot(twist, min.r.squared, data=twists)
+twists$bullet <- gsub("(.*)-[0-9]*","\\1",twists$bulletland)
+qplot(twist, min.r.squared, data=twists) + facet_wrap(~bullet)
+summary(twists$twist)
+twists$constraint <- "none"
+twists$method <- "q75"
+write.csv(twists, "mean-profile-papers/twists.csv", row.names=FALSE)

### 2016-04-13 revision/response.Rmd
Rmd file for response to the reviewers

---
title: "Response to reviewers"
author: "Heike Hofmann, Eric Hare, Alicia Carriquiry"
date: "April 14, 2016"
output: html_document
---

AE:

1. The paper should better explain the role of the new methods for forensic science. The authors should consider referencing some of the previous work the reviewers mention.
2. Methodologically, the crux of the data analysis seems to be a generic application of decision trees. It would be good to clarify that aspect, and to more formally explain how the ensemble of the steps took by the authors to complete this work can be applied to other forensic science questions.

Reviewer A:

4. Random forests, as used and described on pages 18-21 of the paper comes out looking like a winner, but other than that success---worth reporting *somewhere*---, there is no new information on that method. No new insights. And certainly no comparisons with the tens of other things that could have been tried.

Reviewer B:

5. The authors should discuss their work in the context of Nicholas Petraco’s work as well as other ongoing research endeavors in impression and pattern evidence. See https://www.ncjrs.gov/pdffiles1/nij/grants/239048.pdf The alignment method should have been compared with a standard method- See Patraco’s NIJ report for a couple of examples. As presented the novel component of the paper is the ``twicing" for grooves combined with a loess fit to the remove curves from bullet images. I am not sure this is the authors intention. It would be interesting to compare this method with some of the classical methods from functional data analysis used for curve registration.

6. page 7- figure 4- What exactly are you bootstrapping? It is unclear in the paper whether you are bootstrapping residual or bullets or something else entirely. Please clarify.

7. There should have been a comparison of different classification methods to show that the random forest is reasonable.

8. On the random forest classifier- are the prior probabilities being estimated from the cross comparisons described on page 15? What exactly are the prior probabilities used in the random forest?

9. Equal Error Rates on the ROC curves would be a helpful metric as well as the AUC.

***included***

10. Would it be possible to stress the classification system by only using part of the profile? Such as would occur when a partial bullet is recovered? By stressing the random forest you would be able to compare it to other classifiers.

11. What is the standard error of the error rates presented for the matching algorithms? The number of cross comparisons gives false impression the standard errors the when presenting the results.

12. In the conclusions section the authors mention an assumption of ``uniqueness". In general forensic science have moved beyond these questions- see the work of Saks and Kaye for details.

***probably refers to articles such as ``how good is good enough?", books ``modern scientific evidence", "judging science" .... I am not sure that those are really necessary or helpful. It might be better to re-phrase the sentence in the conclusion.***

13. A discussion on how the authors view the use of their matching algorithm would be useful to the forensic community. For example are they intending to use this to build a closed set identification system similar to IAFIS or are them intending this to work in a manner similar to a biometric verification system? etc.

Reviewer C:

14. there is a lack of knowledge of the existing literature on statistical work performed on tool marks and firearms. For example, the work performed by Nicholas Petraco, Fabiano Riva, or Alan Zheng in the past 5 years is not mentioned. Instead, old research projects are mentioned, which mostly focus on manual identification. The authors need to discuss the novelty of their method (and how it potentially addresses the shortcomings of these recent methods)

15.  The authors would need to study (or at least discuss extensively) the ability of their method to capture the variability in bullet profiles over time (at interval of several dozen of bullets fired by the same weapon, when the weapon has been cleaned, or when the bullet was degraded by hitting a target). In addition, a dataset of 10 firearms (even consecutively manufactured) does not provide any information on the rarity of a particular bullet profile, and therefore does not address the second probability statement. Alternatively, the proposed metrics could be tested in a database searching situation, which has not been done.

16. On a related subject, I believe that a random forest provides the probability that two bullets were fired by the same gun given the set of observations (Fig 19), not the probability of interest: the probability of two sets of observations given that the two bullets were fired by the same gun. The authors may want to revise some of the statements in their discussion and conclusion.

17. I am concerned by the assumption of independence of between the different lands on a given bullet that the authors seems to use in their error rates calculations (and ROC curves). The assumption seems unreasonable since as soon as a land has been paired between two bullets, it seems that it becomes more likely to pair the other lands. In other words, I believe that the authors need to redo their error rate measurements, but using bullets as their experimental units and not lands. I am also concerned by the overwhelming majority of non-matches, this must certainly artificially increase the rate of correct exclusion and distort the data.

18. the authors attempt to define a "stable" profile (see Figure 9) and use the stable profile for comparison. What happens if such profile does not exist. Would it not be best to compare surfaces instead of sections? a 2D version of the proposed algorithm should be easily implementable (and has been done at least by Riva).

19. L323: how do the authors know that there are "172 known matches" since the 15 "trace bullets" are supposed to be of unknown sources? which raise the question on how where the ROC and other density graphs generated? only based on the bullets of known origin?

20. If would be helpful if the legend from Fig15 could include a definition of the CMS, CNMS, CCF, ... acronyms such that the reader does not need to flip back and forth between several pages.

Reviewer 1:

21. First, the term "land" is used in the title but not defined until p. 3.  I recommend changing the title, to ensure broad comprehenson; perhaps "Automatic Matching of Bullet Marks," or something of that kind.

22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.

DONE 23. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.

24. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.

25. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.

DONE 26. p. 7, l. 1:  Make this "... can be described in the form ...".

DONE 27. p. 7, l. 9:  Remove the comma after "here" in "... function used here, ...".

DONE 28. p. 7, l. -6:  I'm confused.  The text says "bullets 2 are shown" but the caption refers to bullet 2- which I think should be a specific land on a single bullet.

DONE 29. p. 12, l. -11:  The use of a colon to transition to a new paragraph seems odd.  I recommend just rewording so is a single paragraph and does not use a colon at all.  It is almost as if you are trying to quote rather than state.

DONE 30. p. 13, l. -3:  You write "Given that striation patterns are typically much larger, ..." and I think it would be more clear to write "Given that striation widths are typically much larger, ...".

31. p. 16, l. -1/-2:  Could you elaborate on why a "large number of maximal CMS by itself is not indicative of a match"?  I think the reason is that there could be other marks that are completely incompatible and preclude a match, based on the next sentence, but I am not sure and a few words would be helpful.

DONE (MAYBE) 32. p. 17, l. -13:  I'd prefer another term than "correlation" here.  In fact, the entire sentence could be improved.

DONE 33. p. 17, l. -2:  Remove the comma after "indicates".

DONE 34. p. 18, l. -4:  Change "this study" to "that study", so the reader is certain you mean Ma et al. (2001) and not your own.

35. p. 20,  l. -3:  I really like the idea of including an inconclusive zone.

36. p. 21, Fig. 20:  Most of the features plotted have been defined, but I don't recall x1 and x2.  You may want to remind the reader about all the definitions, or have a table of features appear early in the random forests discussion.

DONE 37. p. 22, l. -9:  Change "bottomline" to "bottom line".

DONE 38. p. 23, l. 4:  Remove the comma after "found".

39. p. 23, l. -13/-5:  I don't think this mention of the Aadhar project adds much, and it distracts from the points you want to highlight about the work in this paper.

DONE 40. Supplement, p. 8, l. 18:  Change "de-facto" to "de facto".

DONE 41. So, my suggestion here is to change the sentence in the abstract to something like ``Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel." Or ``...help the trier of fact determine...".

42. the concept of ``match": The primary goal in forensic science is NOT to attempt to match two bullets. It is to quantify the value of the observations made on the two bullets with regard to the two competing propositions of interest to the court (i.e., assign a likelihood ratio or Bayes factor).

43. I strongly encourage the authors to not follow this trend and use the term ``ballistics" only when you are actually talking about ballistics.

44. 2nd to last line in the abstract: In point (a) ``correctly identify lands with too much damage to be suitable for matching" I suggest replacing the word ``matching" with ``comparison" to use the same terminology as the other pattern evidence disciplines.

45. Line 55: Replace ``in the forensic sciences" with ``in forensic science". (See Margot P, Commentary on The Need for a Research Culture in the Forensic Sciences, UCLA Law Review (2011) 58: 795-801.)

46. Lines 156-158: The sentence ``The blue area around the signature are 95% bootstrapped confidence intervals based on 1,000 bootstrap samples." is confusing. The term ``signature" is defined two sentences further. Is the sentence referring to Figure 4b? If yes, it should probably come at the end of this paragraph.

47. Figure 5: In this figure, where is the limit between bullet 1-5 and bullet 2-1?

48. Line 246: Replace ``this representations" with ``this representation" or ``these representations"

49. Caption of Figure 8: ``At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures." is not a complete sentence.

50. Line 266: Capitalize the ``b" in ``by".

### 2016-04-15 revision/response.Rmd
couple of comments added to review

@@
 
+***why should the probability of two sets of observations given that the two bullets were fired by the same gun be the probability of interest?
+I would agree that the random forest and any of its conclusions are restricted to how good the underlying 'database' is.
+The more representative the set of bullets, the better the predictions should be.
+It is not the purpose of this paper to find the probability of two sets of observations given that the two bullets were fired by the same gun.
+We made sure to remove any misleading statements that might suggest the latter probability.
+***
+
 17.
@@
 I am also concerned by the overwhelming majority of non-matches, this must certainly artificially increase the rate of correct exclusion and distort the data.
+
+***Assuming independence between the lands of each barrel is certainly not realistic in practice, however, by assuming independence we make the task of matching harder for the algorithm.
+XXX still need to do XXX Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel.
+The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.
+***
 
@@
 
+***Figure 11 shows images of all bullet lands we encountered in the Hamby set where no stable profile could be identified.
+All four of these images show massive signs of tank rash - this motivates our suggestion to not consider bullet lands without a stable region for the further analysis.
+Concentrating on one-dimensional profiles is quite common in the literature, and we have expanded the discussion on this choice with appropriate citations.***
+
 19.
@@
 
+***For the Hamby study there is a known ground truth of known matches and known non-matches, which we used for determining the accuracy of the algorithm.
+As state in the paper, the ROC curves are based on comparisons of 88 lands from unknown bullets (90 minus the two lands that showed too much tank rash) and 118 lands from known bullets (120 minus two bullet lands affected by tank rash), resulting in a total of 10,384 land to land comparisons.***
+
 20. If would be helpful if the legend from Fig15 could include a definition of the CMS, CNMS, CCF, ... acronyms such that the reader does not need to flip back and forth between several pages.
+
+***Done for figure 15 and 16***
 
@@
 I recommend changing the title, to ensure broad comprehenson; perhaps "Automatic Matching of Bullet Marks," or something of that kind.
+
+***Let's discuss that on Monday***
 

### 2016-04-15 imaging-paper.Rnw
typo
couple of comments added to review

@@
   \centering
-<<density-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
+<<density-overview,echo=FALSE, warning=FALSE, fig.width=12.5, fig.height=6.25, out.width='\\textwidth'>>=
 features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
@@
 levels(dm$variable) <- c("Known matches (KM)", "Known non-matches (KNM)")
+dm$varLabel <- dm$var
+levels(dm$varLabel) <- c(
+  "Consecutive Matching Striae (CMS)", 
+  "Consecutive Non-Matching Striae (CNMS)",
+  "#matches",
+  "#non-matches",
+  "Average difference (D)",
+  "Sum of peaks (S)",
+  "Cross-correlation function (ccf)")
 
@@
       alpha=I(0.6), colour=I("grey20")) + 
-  facet_wrap(~var, ncol=4, scales="free") +
-  theme_bw() + ylab("density") + xlab("") +
+  facet_wrap(~varLabel, ncol=4, scales="free") +
+  theme_bw() + ylab("") + xlab("") +
   scale_fill_brewer("Matches", palette="Paired") +
-  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
+  theme(legend.position=c(1,0), 
+        legend.justification = c("right", "bottom"),
         legend.background = element_rect(colour="grey75"),
+        axis.title.y=element_blank(),
         plot.margin = unit(c(0,0,0,0), unit="cm"))
@@
   \centering
-<<rocs-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width='\\textwidth'>>=
+<<rocs-overview,echo=FALSE, warning=FALSE, fig.width=12.5, fig.height=6.25, out.width='\\textwidth'>>=
 # plot false positive against false negative
@@
 levels(rocs$variable)[3:4] <- c("#non-matches", "#matches")
+rocs$varLabel <- rocs$variable
+levels(rocs$varLabel) <- c(
+  "Sum of peaks (S)",
+  "Cross-correlation function (ccf)",
+  "#non-matches",
+  "#matches",
+  "Consecutive Non-Matching Striae (CNMS)",
+  "Consecutive Matching Striae (CMS)", 
+  "Average difference (D)"
+  )
 
@@
   geom_ribbon(aes(x=`FALSE`, ymax=`TRUE`, ymin=0), alpha=0.2) +
-  geom_line(aes(x=`FALSE`,y=`TRUE`))+ theme_bw() + xlab("False negative rate") + 
-  ylab("False positive rate") + facet_wrap(facets=~variable, ncol=4) + 
+  geom_line(aes(x=`FALSE`,y=`TRUE`))+ theme_bw() + 
+  xlab("False negative rate") + 
+  ylab("False positive rate") + facet_wrap(facets=~varLabel, ncol=4) + 
   geom_label(x=1, y=0, aes(label=labels), data=labels, hjust=1, vjust=0) + 
@@
 The smaller the value, the better.
-We see that in this instance a lowe EER goes hand in hand with high predictive power as measured in AUC.
+We see that in this instance a low EER goes hand in hand with high predictive power as measured in AUC.
 XXX HH: do we need to say more about EERs?}

### 2016-04-15 revision/response.Rmd
couple more comments

@@
 
+***I am not quite sure what a 'reasonable' classification method is supposed to mean. Shouldn't a classification method be judged by its success in predicting the outcome? 
+Apart from that, random forests are based on decision trees, which are an extremely flexible tool for classification, that can deal with quantitative and categorical types of variables and, in the implementation we used, also deal with some missing values (which was not needed in the data here).***
+
 8. On the random forest classifier- are the prior probabilities being estimated from the cross comparisons described on page 15? What exactly are the prior probabilities used in the random forest?
+
+***No data-based adjustment was made to the prior probabilities, i.e. prior probabilities of 50:50 for matches/non-matches were used.
+HH: what does the reviewer mean by cross-comparison? On page 15 we discuss some results from matching based on CMS.***
 
@@
 
+***Yes, it would be possible to stress the classification system by using only parts of the profile, but a systematic evaluation of this by far exceeds the scope of this paper.***
+
 11. What is the standard error of the error rates presented for the matching algorithms? The number of cross comparisons gives false impression the standard errors the when presenting the results.
+
+***HH: I'm sure I am misunderstanding the comment, but I don't know how to answer this question.
+The standard error of a rate in (0,1) would be sqrt(p*(1-p)).
+So because p is small the square root is going to be still pretty small.
+We are presenting out of bag error rates for the random forest - maybe we need to explain what out-of bag means in more detail?***
 

### 2016-04-15 imaging-paper.Rnw
light grey shading for figure 5

@@
 \centering
-\includegraphics[width=.65\columnwidth]{images/matchup-rgl.png}
-\caption{\label{fig:manualmatch-rgl}3D view of the manually adjusted side-by-side comparison of bullet~1-5 and bullet 2-1 after removing the curvature.}
+\includegraphics[width=.65\columnwidth]{images/matchup-rgl-copy.png}
+\caption{\label{fig:manualmatch-rgl}3D view of the manually adjusted side-by-side comparison of bullet~1-5 and bullet 2-1 after removing the curvature.
+\hh{Bullet 2-1 is shaded light grey in the background.}}
 \end{figure}

### 2016-04-15 imaging-paper.Rnw
A couple small changes
A few other small changes

@@
 This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels.
-The automatic matching routine we propose manages to (a) correctly identify lands with too much damage to be suitable for matching, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
+The automatic matching routine we propose manages to (a) correctly identify lands with too much damage to be suitable for comparison, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
 \end{abstract}
@@
 The x3p format conforms to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
-It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in the forensic sciences.
+It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in forensic science.
 
@@
 Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$.
-Striation marks show up in this representations as peaks and valleys.
+Striation marks show up in these representations as peaks and valleys.
 The individual characteristics are quite prominent and, again, suggest a match between the lands.
@@
 \caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
-At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures.
+The correlation peaks at a lag of -17, which indicates the largest amount of agreement between the signatures.
 Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}

### 2016-04-16 imaging-paper.Rnw
unkn47 didn't match with the description

@@
 <<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='\\textwidth'>>=
-bstats$km <- c( "KNM", "KM")[as.numeric(bstats$match)+1]
+bstats$km <- c( "Known non-match", "Known match")[as.numeric(bstats$match)+1]
 ggplot(data=subset(bstats, !flagged)) + 

### 2016-04-16 revision/response.Rmd
light grey shading for figure 5
A couple small changes
A few other small changes
edited the comments

@@
 6. page 7- figure 4- What exactly are you bootstrapping? It is unclear in the paper whether you are bootstrapping residual or bullets or something else entirely. Please clarify.
+
+***HH: will do***
 
@@
 
+***HH: in my view we ourselves are not building a closed set identification system.
+The problem with existing closed set systems is that it is unclear how the matching is exactly done and what statistical properties it has (number of parameters, distribution of them, sensitivities).
+Repeatability and reproducibility is a huge issue in this area.
+'Matching' is also not just a single step - it consists, as we have laid out in the paper, of a series of steps, which all deserve attention.
+While the reviewer pointed out to us to use 'standard methods' - there are no actual standards in place.
+There might be a set of alternatives, and some evidence as to which methods might work better in some situations than others, but there is a lack of an in-depth discussion of the alternatives at each step as well as an overall discussion.
+In this manuscript we are detailing one route to divvy up the process and provide the code for others to reproduce our findings.
+***
+
 Reviewer C:
@@
 Alternatively, the proposed metrics could be tested in a database searching situation, which has not been done.
+
+***At the time of the submission of this manuscript, the 35 bullets from Hamby set 252 were the only ones accessible to the public.
+They are part of Dr Zheng's efforts at NIST to provide a ballistics reference database.
+In the future more results will be available, but until we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.***
 
@@
 
-DONE 23. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.
+23. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.
+
+***DONE***
 
@@
 
-DONE 26. p. 7, l. 1:  Make this "... can be described in the form ...".
+26. p. 7, l. 1:  Make this "... can be described in the form ...".
 
-DONE 27. p. 7, l. 9:  Remove the comma after "here" in "... function used here, ...".
+***DONE***
 
-DONE 28. p. 7, l. -6:  I'm confused.  The text says "bullets 2 are shown" but the caption refers to bullet 2- which I think should be a specific land on a single bullet.
+27. p. 7, l. 9:  Remove the comma after "here" in "... function used here, ...".
 
-DONE 29.
+***DONE***
+
+28. p. 7, l. -6:  I'm confused.  The text says "bullets 2 are shown" but the caption refers to bullet 2- which I think should be a specific land on a single bullet.
+
+***DONE***
+
+29.
 p.
@@
 
-DONE 30.
-p.
-13, l.
--3:  You write "Given that striation patterns are typically much larger, ..." and I think it would be more clear to write "Given that striation widths are typically much larger, ...".
+***DONE***
+
+30. p. 13, l. -3:  You write "Given that striation patterns are typically much larger, ..." and I think it would be more clear to write "Given that striation widths are typically much larger, ...".
+
+***DONE***
 
@@
 
-DONE (MAYBE) 32. p. 17, l. -13:  I'd prefer another term than "correlation" here.  In fact, the entire sentence could be improved.
+32. p. 17, l. -13:  I'd prefer another term than "correlation" here.  In fact, the entire sentence could be improved.
 
-DONE 33. p. 17, l. -2:  Remove the comma after "indicates".
+***DONE (Maybe)***
 
-DONE 34. p. 18, l. -4:  Change "this study" to "that study", so the reader is certain you mean Ma et al. (2001) and not your own.
+33. p. 17, l. -2:  Remove the comma after "indicates".
+
+***DONE***
+
+34. p. 18, l. -4:  Change "this study" to "that study", so the reader is certain you mean Ma et al. (2001) and not your own.
+
+***DONE***
 
@@
 
-DONE 37. p. 22, l. -9:  Change "bottomline" to "bottom line".
+37. p. 22, l. -9:  Change "bottomline" to "bottom line".
 
-DONE 38. p. 23, l. 4:  Remove the comma after "found".
+***DONE***
+
+38. p. 23, l. 4:  Remove the comma after "found".
+
+***DONE***
 
@@
 
-DONE 40. Supplement, p. 8, l. 18:  Change "de-facto" to "de facto".
+40. Supplement, p. 8, l. 18:  Change "de-facto" to "de facto".
 
-DONE 41.
+***DONE***
+
+41.
 So, my suggestion here is to change the sentence in the abstract to something like ``Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel." Or ``...help the trier of fact determine...".
+
+***DONE***
 
@@
 
+***DONE***
+
 45.
@@
 (See Margot P, Commentary on The Need for a Research Culture in the Forensic Sciences, UCLA Law Review (2011) 58: 795-801.)
+
+***DONE***
 
@@
 
+***Light grey shading in the background was added to help with the distinction between the two flattened bullet lands.***
+
 48. Line 246: Replace ``this representations" with ``this representation" or ``these representations"
+
+***DONE***
 
@@
 
+***DONE***
+***HH: would a removal of the comma make it a complete sentence?***
+
 50. Line 266: Capitalize the ``b" in ``by".
+
+***DONE***

### 2016-04-16 revision/response.Rmd
discussion of references

@@
 It would be interesting to compare this method with some of the classical methods from functional data analysis used for curve registration.
+
+***Additional papers considered (not complete, and not yet included in the paper): 
+chumbley:2010 
+uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening)
+
+chu:2010
+describe extracting mean profile of an area of the image after untwisting the image
+
+bachrach:2002
+general description of an automated matching system based on 3D imaging. 
+Uses cross-correlation function for alignment.
+***
 

### 2016-04-16 revision/response.Rmd
another reference
Riva reference

@@
 Uses cross-correlation function for alignment.
+
+petraco:2012
+I am not sure that this book is the same as the report for the NIJ grant.
+This is a huge effort, and we should certainly cite it, but there are not a lot of actual results included or comparisons done.
+
+riva:2014
+automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.
 ***

### 2016-04-17 revision/response.Rmd
added citation

--- 2016-04-16 revision/response.Rmd
@@
 It would be interesting to compare this method with some of the classical methods from functional data analysis used for curve registration.
+
+***Using the cross correlation function as a method of alignment is actually pretty standard - the approriate references are now included.
+***
+
+***Additional papers considered (not complete, and not yet included in the paper, but I think all of them should be cited): 
+chumbley:2010 
+uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening)
+chubley:2010 is a much more detailed version over David Faden's thesis***
+
+***chu:2010
+describe extracting mean profile of an area of the image after untwisting the image ***
+
+***bachrach:2002
+general description of an automated matching system based on 3D imaging. 
+Uses cross-correlation function for alignment.***
+
+***petraco:2012
+I am not sure that this book is the same as the report for the NIJ grant.
+This is a huge effort, and we should certainly cite it, but there are not a lot of actual results included or comparisons done.***
+
+***riva:2014
+automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.
+***
 

### 2016-04-17 imaging-paper.Rnw
added citation

@@
 %
+\hh{For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, thompson:2013}.}
 This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.

### 2016-04-18 revision/response.Rmd
More comments and responses

@@
 We are presenting out of bag error rates for the random forest - maybe we need to explain what out-of bag means in more detail?***
+
+***EH: Yeah, I was a little unsure of this. Maybe we can ask Alicia, and if she also is confused by it I can type up an explanation on how the OOB error is computed.***
 
@@
 
-***Let's discuss that on Monday***
+***HH: Let's discuss that on Monday***
 
 22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.
+
+***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***
 
@@
 
+***EH: We could slightly rotate the image in RGL, but the top-down view is nice. Alternatively I could just disable the axis labels - We don't really need to see them in this image, right?***
+
 25. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.
+
+***EH: Hm.
+I think we phrased this correctly and the reviewer misunderstood - We literally ARE showing groove to groove impressions.
+Maybe we need to be more clear about groove impressions...
+but in the following paragraph + the image of the bullet I think its fairly clear.
+We could add in a little disclaimer like "Grooves, not to be confused with striae, are..."***
 
@@
 ***DONE***
+
 ***HH: would a removal of the comma make it a complete sentence?***
+
+***EH: I actually thought it already was a complete sentence... but I changed it to "The correlation peaks at a lag of -17, which indicates the largest amount of agreement between the signatures."***
 

### 2016-04-19 revision/response.Rmd
discussion of additional papers

@@
 
+***bachrach:2006
+67page final report for NIJ grant (for Intelligent Automation Inc.)
+wear study: "difference between comparing the nth and (n+1)th bullet fired by a barrel as opposed to the nth and (n+100)th bullet fired by a barrel?
+In other words, does the proximity between firings make a difference?
+Although some wear effect could be detected in both of these scenarios, these effects do not appear to be significant enough to prevent correct identification"***
+
 ***petraco:2012
@@
 ***riva:2014
-automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.
-***
+automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.***
+
+***xie:2009
+first sentence: "Every firearm has individual characteristics that are as unique to it as fingerprints are to human beings." the paper itself doesn't back up any of these claims. very disappointing***
+
+***leon:2006
+first sentence: "Fired bullets bear striation marks that can be thought of as a ‘‘fingerprint’’ left by the firearm." based on multiple 2d images. They extract a 1d 'signature'***
 
@@
 They are part of Dr Zheng's efforts at NIST to provide a ballistics reference database.
-In the future more results will be available, but until we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.***
+In the future more results will be available, but until then we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.***
 
@@
 ***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***
+***HH: I am not sure about the width of a striation - the width depends on where it is measured (in both x and z direction).
+I'm sure there are definitions of that available.
+We might need some more literature.***
 
@@
 ***EH: We could slightly rotate the image in RGL, but the top-down view is nice. Alternatively I could just disable the axis labels - We don't really need to see them in this image, right?***
+***HH: I just deleted the values from the axis***
 

### 2016-04-19 revision/response.Rmd
vorburger:2011 reference for ccf for alignment

@@
 Uses cross-correlation function for alignment.***
+
+***vorburger:2011
+different uses of the cross-correlation function. One of them aligning profiles***
 
@@
 
+***HH: this is clearly a misunderstanding, so we need to be nice about breaking this to the reviewer, and emphasize it in the paper some more.***
 ***EH: Hm.
@@
 ***EH: I actually thought it already was a complete sentence... but I changed it to "The correlation peaks at a lag of -17, which indicates the largest amount of agreement between the signatures."***
+***HH: Ta, as they say around here :)***
 

### 2016-04-19 imaging-paper.Rnw
vorburger:2011 reference for ccf for alignment

--- 2016-04-16 imaging-paper.Rnw
@@
 %
+\hh{For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, vorburger:2011, thompson:2013}.}
 This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.

### 2016-04-19 revision/response.Rmd
Comments from our meeting

--- 2016-04-19 revision/response.Rmd
@@
 
+***vorburger:2011
+different uses of the cross-correlation function. One of them aligning profiles***
+
 ***bachrach:2006
@@
 
-***No data-based adjustment was made to the prior probabilities, i.e. prior probabilities of 50:50 for matches/non-matches were used.
-HH: what does the reviewer mean by cross-comparison? On page 15 we discuss some results from matching based on CMS.***
+***No data-based adjustment was made to the prior probabilities, i.e. prior probabilities of 50:50 for matches/non-matches were used.***
+
+***HH: what does the reviewer mean by cross-comparison? On page 15 we discuss some results from matching based on CMS.***
 
@@
 
+***EH: Quick definition of land in abstract***
+
+25. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.
+
+***HH: this is clearly a misunderstanding, so we need to be nice about breaking this to the reviewer, and emphasize it in the paper some more.***
+
+***EH: Hm.
+I think we phrased this correctly and the reviewer misunderstood - We literally ARE showing groove to groove impressions.
+Maybe we need to be more clear about groove impressions...
+but in the following paragraph + the image of the bullet I think its fairly clear.
+We could add in a little disclaimer like "Grooves, not to be confused with striae, are..."***
+
 22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth. 
+***EH: Add typical width of groove + lands***
 
@@
 We might need some more literature.***
+
+* Make note of the width 35 * 6
 
@@
 ***EH: We could slightly rotate the image in RGL, but the top-down view is nice. Alternatively I could just disable the axis labels - We don't really need to see them in this image, right?***
+
 ***HH: I just deleted the values from the axis***
-
-25. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.
-
-***EH: Hm.
-I think we phrased this correctly and the reviewer misunderstood - We literally ARE showing groove to groove impressions.
-Maybe we need to be more clear about groove impressions...
-but in the following paragraph + the image of the bullet I think its fairly clear.
-We could add in a little disclaimer like "Grooves, not to be confused with striae, are..."***
 
@@
 
+***EH: From our data out of non matches its a 2 / 10212 event...
+but with a lot of pairwise comparisons, it will happen by chance even if its very unlikely.
+Adding a couple more bullets / comparisons we may see higher values.***
+
 32. p. 17, l. -13:  I'd prefer another term than "correlation" here.  In fact, the entire sentence could be improved.
@@
 
+***EH: Add thank you, and explain the ones that could be considered inconclusive. Lots more discussion would be needed about the extent of that zone.***
+
 36.
@@
 You may want to remind the reader about all the definitions, or have a table of features appear early in the random forests discussion.
+
+***EH: Take out all references to x1 and x2***
 
@@
 39. p. 23, l. -13/-5:  I don't think this mention of the Aadhar project adds much, and it distracts from the points you want to highlight about the work in this paper.
+
+***EH: We can take this out***
 
@@
 
+***EH: Don't call it primary goal, call it one of the primary goals***
+
 43. I strongly encourage the authors to not follow this trend and use the term ``ballistics" only when you are actually talking about ballistics.
+
+***EH: Change this, HH agrees.**
 
@@
 
+***HH: Remove blue area / references to it... maybe?***
+
 47. Figure 5: In this figure, where is the limit between bullet 1-5 and bullet 2-1?
@@
 
+***HH: Ta, as they say around here :)***
+
+***EH: Change to the correlation is maximizes.***
+
 50. Line 266: Capitalize the ``b" in ``by".

### 2016-04-20 revision/response.Rmd
Big set of changes

@@
 
-***EH: We can take this out***
+***We have removed references to the Aadhar project.***
 
@@
 
-***EH: Don't call it primary goal, call it one of the primary goals***
+***We have clarified that matching two bullets is one of the goals, but not the sole or primary goal.***
 
@@
 
-***EH: Change this, HH agrees.**
+***We have removed references to the term ``ballistics" except in the context of citing prior work which uses the term.***
 
@@
 
-***DONE***
-
-***HH: would a removal of the comma make it a complete sentence?***
-
-***EH: I actually thought it already was a complete sentence... but I changed it to "The correlation peaks at a lag of -17, which indicates the largest amount of agreement between the signatures."***
-
-***HH: Ta, as they say around here :)***
-
-***EH: Change to the correlation is maximizes.***
+***We have revised this caption to remove the ambiguity from the word ``peaks", which we had intended to mean "maximizes."***
 

### 2016-04-20 imaging-paper.Rnw
added ccf to the list of features - where was that before??
Big set of changes

@@
 During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet.
-Identifying these striation markings in an attempt to match two bullets is the primary goal.
+Identifying these striation markings in an attempt to match two bullets is one of the primary goals.
 We propose an automated framework for the analysis of the 3D surface measurements, that first transcribes the markings into a 2D plotting framework.
@@
 This process has broad applicability in terms of convictions in the United States criminal justice system.
-Ballistics identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny.
+Firearms identification has long been considered an accepted and reliable procedure, but in the past ten years has undergone more significant scrutiny.
 In 2005, in \emph{United States vs.
 Green}, the court ruled that the forensic expert could not confirm that the bullet casings came from a specific weapon with certainty, but could merely ``describe" other casings which are similar.
-Further court cases in the late 2000s expressed caution about the use of ballistics identification evidence~\citep{giannelli:2011}.
+Further court cases in the late 2000s expressed caution about the use of firearms identification evidence~\citep{giannelli:2011}.
 
@@
 \item Number of matches and non-matches,
+\item \hh{The value of the cross-correlation function (ccf) between the aligned signatures \cite{vorburger:2011}, }
 \item Average difference $D$ between signatures, defined as the Euclidean vertical distance between surface measurements of aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
@@
 \caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
-The correlation peaks at a lag of -17, which indicates the largest amount of agreement between the signatures.
+The correlation is maximized at a lag of -17, indicating the largest amount of agreement between the signatures.
 Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}
@@
 
-We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a ballistics identification perspective, removes these features, and produces residuals which allow for the easy identification of markings.
+We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a firearms identification perspective, removes these features, and produces residuals which allow for the easy identification of markings.
 We have generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches.
@@
 
-An approach that might be helpful comes from Biometrics \citep{jain:2015:tphil, jain:2015}.
+%An approach that might be helpful comes from Biometrics \citep{jain:2015:tphil, jain:2015}.
 It has been successfully applied in the Aadhar project (the effort by the Unique Identification Authority of India (UIDAI) to provide a unique 12-digit identification number to approximately 1.2 billion residents of India based on ten fingerprints and two irises).
-This approach uses dictionary learning based on small square patches  of 64x64 or 32x32 pixels.
+%This approach uses dictionary learning based on small square patches  of 64x64 or 32x32 pixels.
 For matching bullets, a simpler, one-dimensional window approach might be sufficient to match striation marks because of the physical forces  leading to the existence of these marks.
@@
 \section*{Acknowledgment} 
-Thanks to David Baldwin for pointing us to the NIST database and doing a Ballistics 101 for us.
+Thanks to David Baldwin for pointing us to the NIST database and doing a Firearms 101 for us.
 Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.

### 2016-04-20 revision/response.Rmd
More comments

@@
 
-***DONE (Maybe)***
+***DONE***
 
@@
 
-***EH: Add thank you, and explain the ones that could be considered inconclusive. Lots more discussion would be needed about the extent of that zone.***
+***Thank you for the comment.
+We have chosen to leave off a more detailed discussion on an inconclusive zone.
+Certainly, formulating such a zone would have important implications in terms of the criminal justice system, but it is one that we felt was outside the scope of this particular paper.
+Certainly, should an automated algorithm like we propose be used as an aid in the conviction or the exoneration of suspects, much attention would be needed on the extent of this zone.***
 

### 2016-04-20 imaging-paper.Rnw
More comments

@@
 imp2 <- subset(imp, Variable %in% c("x1", "sd.D", "S"))
-qplot(MeanDecreaseGini,1, data=imp, geom="point") + 
-  geom_text(aes(y=1.2, label=Variable), angle=45, hjust=0, data=imp1) + 
-  geom_text(aes(y=0.8, label=Variable), angle=-45, hjust=0, data=imp2) + 
+imp3 <- subset(imp, !(Variable %in% c("x1", "x2", "sd.D", "lag")))
+qplot(MeanDecreaseGini,1, data=imp3, geom="point") + 
+  geom_text(aes(y=1.2, label=Variable), angle=45, hjust=0, data=subset(imp3, Variable != "S")) + 
+  geom_text(aes(y=1.2, label=Variable), angle=45, vjust=1.2, hjust=0, data=subset(imp3, Variable == "S")) + 
   ylim(c(0, 2.75)) + theme_bw() + ylab("") +

### 2016-04-20 imaging-supplement.Rnw
More comments

@@
 names(imp)[-(1:3)] <- c("KM", "(sd)", "KNM", "(sd)")
-print(xtable(imp, align="clrrrd{3.2}rd{3.2}", digits=c(0,0,1,1,1,2,1,2)), 
+
+result <- subset(imp, !(Variable %in% c("x1", "x2", "lag", "sd.D")))
+rownames(result) <- 1:nrow(result)
+
+print(xtable(result, align="clrrrd{3.2}rd{3.2}", digits=c(0,0,1,1,1,2,1,2)), 
       floating = FALSE)

### 2016-04-20 imaging-supplement.tex
More comments

@@
 % latex table generated in R 3.2.4 by xtable 1.8-2 package
-% Tue Apr  5 08:28:49 2016
+% Wed Apr 20 10:17:06 2016
 \begin{tabular}{clrrrd{3.2}rd{3.2}}
   \hline
- & Variable & Importance 1 & Importance 2 & KM & (sd) & KNM & (sd) \\ 
+ & Variable & Importance 1 & Importance 2 & KM & (sd) & KNM & (sd).1 \\ 
   \hline
@@
   6 & CMS & 15.9 & 25.1 & 8.5 & ( 5.65) & 2.3 & (   1.40) \\ 
-  7 & sd.D & 15.0 & 23.6 & 2.9 & ( 0.80) & 2.9 & (   0.70) \\ 
-  8 & CNMS & 13.3 & 20.9 & 4.8 & ( 4.10) & 10.2 & (   4.35) \\ 
-  9 & lag & 9.9 & 17.0 & 200.8 & (48.29) & 204.0 & ( 113.63) \\ 
-  10 & x2 & 4.2 & 6.2 & 166.5 & (81.07) & 163.5 & (  72.30) \\ 
-  11 & x1 & 3.9 & 6.1 & 175.0 & (71.59) & 175.0 & (  71.40) \\ 
+  7 & CNMS & 13.3 & 20.9 & 4.8 & ( 4.10) & 10.2 & (   4.35) \\ 
    \hline

### 2016-04-20 revision/response.Rmd
Update response, make more paper changes

@@
 
-***Using the cross correlation function as a method of alignment is actually pretty standard - the approriate references are now included.
-***
+***Using the cross correlation function as a method of alignment is actually pretty standard - the approriate references are now included.***
 
-***Additional papers considered (not complete, and not yet included in the paper, but I think all of them should be cited): 
-chumbley:2010 
-uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening)
-chubley:2010 is a much more detailed version over David Faden's thesis***
+***Additional papers considered (not complete, and not yet included in the paper, but I think all of them should be cited): chumbley:2010 uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening) chubley:2010 is a much more detailed version over David Faden's thesis***
 
-***chu:2010
-describe extracting mean profile of an area of the image after untwisting the image ***
+***chu:2010 describe extracting mean profile of an area of the image after untwisting the image ***
 
-***bachrach:2002
-general description of an automated matching system based on 3D imaging. 
-Uses cross-correlation function for alignment.***
+***bachrach:2002 general description of an automated matching system based on 3D imaging. Uses cross-correlation function for alignment.***
 
-***vorburger:2011
-different uses of the cross-correlation function. One of them aligning profiles***
+***vorburger:2011 different uses of the cross-correlation function. One of them aligning profiles***
 
-***bachrach:2006
-67page final report for NIJ grant (for Intelligent Automation Inc.)
-wear study: "difference between comparing the nth and (n+1)th bullet fired by a barrel as opposed to the nth and (n+100)th bullet fired by a barrel?
+***bachrach:2006 67page final report for NIJ grant (for Intelligent Automation Inc.) wear study: "difference between comparing the nth and (n+1)th bullet fired by a barrel as opposed to the nth and (n+100)th bullet fired by a barrel?
 In other words, does the proximity between firings make a difference?
@@
 
-***petraco:2012
-I am not sure that this book is the same as the report for the NIJ grant.
+***petraco:2012 I am not sure that this book is the same as the report for the NIJ grant.
 This is a huge effort, and we should certainly cite it, but there are not a lot of actual results included or comparisons done.***
 
-***riva:2014
-automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.***
+***riva:2014 automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.***
 
-***xie:2009
-first sentence: "Every firearm has individual characteristics that are as unique to it as fingerprints are to human beings." the paper itself doesn't back up any of these claims. very disappointing***
+***xie:2009 first sentence: "Every firearm has individual characteristics that are as unique to it as fingerprints are to human beings." the paper itself doesn't back up any of these claims.
+very disappointing***
 
-***leon:2006
-first sentence: "Fired bullets bear striation marks that can be thought of as a ‘‘fingerprint’’ left by the firearm." based on multiple 2d images. They extract a 1d 'signature'***
+***leon:2006 first sentence: "Fired bullets bear striation marks that can be thought of as a ‘‘fingerprint’’ left by the firearm." based on multiple 2d images. They extract a 1d 'signature'***
 
@@
 
-***I am not quite sure what a 'reasonable' classification method is supposed to mean. Shouldn't a classification method be judged by its success in predicting the outcome? 
+***I am not quite sure what a 'reasonable' classification method is supposed to mean.
+Shouldn't a classification method be judged by its success in predicting the outcome?
 Apart from that, random forests are based on decision trees, which are an extremely flexible tool for classification, that can deal with quantitative and categorical types of variables and, in the implementation we used, also deal with some missing values (which was not needed in the data here).***
@@
 
-***HH: Let's discuss that on Monday***
+***To clarify the title, we have modified the abstract to provide a brief definition of bullet lands.***
 
-***EH: Quick definition of land in abstract***
+22. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.
 
-25. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.
+***Grooves and striations are actually not synonyms in this case.
+The grooves are impressions induced by the rifling of the gun barrel, and properties associated with them are class characteristics not unique to individual gun barrels.
+We've taken care to ensure the distinction is clear between the striae of interest, which are extracted from the lands of the bullet (between grooves), and the grooves themselves.***
 
-***HH: this is clearly a misunderstanding, so we need to be nice about breaking this to the reviewer, and emphasize it in the paper some more.***
-
-***EH: Hm.
-I think we phrased this correctly and the reviewer misunderstood - We literally ARE showing groove to groove impressions.
-Maybe we need to be more clear about groove impressions...
-but in the following paragraph + the image of the bullet I think its fairly clear.
-We could add in a little disclaimer like "Grooves, not to be confused with striae, are..."***
-
-22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth. 
+23. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth. 
 ***EH: Add typical width of groove + lands***
@@
 
-* Make note of the width 35 * 6
-
-23. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.
+24. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.
 
@@
 
-24. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.
+25. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.
 
-***EH: We could slightly rotate the image in RGL, but the top-down view is nice. Alternatively I could just disable the axis labels - We don't really need to see them in this image, right?***
-
-***HH: I just deleted the values from the axis***
+***We have removed the values from the y-axis, as they are not necessary in interpreting the image.***
 
@@
 
-***EH: From our data out of non matches its a 2 / 10212 event...
-but with a lot of pairwise comparisons, it will happen by chance even if its very unlikely.
-Adding a couple more bullets / comparisons we may see higher values.***
+***In the Hamby study, with 10,212 land-to-land comparisons which are known non-matches, two of these comparisons still yielded a CMS of 12.
+Figure 13 emphasizes this.
+In our particular data, all comparisons with a CMS of 13 or more were in fact matches.
+Still, it is likely due to pure chance alone that, particularly if more bullets were used in comparison, a small portion would have high pairwise CMS despite being non-matches.***
 

### 2016-04-20 imaging-paper.Rnw
Update response, make more paper changes

@@
 This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels.
-The automatic matching routine we propose manages to (a) correctly identify lands with too much damage to be suitable for comparison, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
+The automatic matching routine we propose manages to (a) correctly identify lands (the surface between two bullet grooves) with too much damage to be suitable for comparison, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
 \end{abstract}

### 2016-04-20 revision/response.Rmd
Update response

@@
 
-AE:
+# AE
 
@@
 
-Reviewer A:
+# Reviewer A
 
@@
 
-Reviewer B:
+# Reviewer B
 
@@
 See https://www.ncjrs.gov/pdffiles1/nij/grants/239048.pdf The alignment method should have been compared with a standard method- See Patraco’s NIJ report for a couple of examples.
-As presented the novel component of the paper is the ``twicing" for grooves combined with a loess fit to the remove curves from bullet images.
+As presented the novel component of the paper is the "twicing" for grooves combined with a loess fit to the remove curves from bullet images.
 I am not sure this is the authors intention.
@@
 
-12. In the conclusions section the authors mention an assumption of ``uniqueness". In general forensic science have moved beyond these questions- see the work of Saks and Kaye for details.
+12. In the conclusions section the authors mention an assumption of "uniqueness". In general forensic science have moved beyond these questions- see the work of Saks and Kaye for details.
 
-***probably refers to articles such as ``how good is good enough?", books ``modern scientific evidence", "judging science" ....
+***probably refers to articles such as "how good is good enough?", books "modern scientific evidence", "judging science" ....
 I am not sure that those are really necessary or helpful.
@@
 
-Reviewer C:
+# Reviewer C
 
@@
 
-Reviewer 1:
+# Reviewer 1
 
@@
 23. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.
+
 ***EH: Add typical width of groove + lands***
@@
 ***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***
+
 ***HH: I am not sure about the width of a striation - the width depends on where it is measured (in both x and z direction).
@@
 
+***EH: Talk about adding features decreasing error rate. CMS by itself is not a sufficient condition for a match...***
+
 32. p. 17, l. -13:  I'd prefer another term than "correlation" here.  In fact, the entire sentence could be improved.
@@
 
+# Reviewer 2
+
 41.
-So, my suggestion here is to change the sentence in the abstract to something like ``Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel." Or ``...help the trier of fact determine...".
+So, my suggestion here is to change the sentence in the abstract to something like "Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel." Or "...help the trier of fact determine...".
 
@@
 42.
-the concept of ``match": The primary goal in forensic science is NOT to attempt to match two bullets.
+the concept of "match": The primary goal in forensic science is NOT to attempt to match two bullets.
 It is to quantify the value of the observations made on the two bullets with regard to the two competing propositions of interest to the court (i.e., assign a likelihood ratio or Bayes factor).
@@
 
-43. I strongly encourage the authors to not follow this trend and use the term ``ballistics" only when you are actually talking about ballistics.
+43. I strongly encourage the authors to not follow this trend and use the term "ballistics" only when you are actually talking about ballistics.
 
-***We have removed references to the term ``ballistics" except in the context of citing prior work which uses the term.***
+***We have removed references to the term "ballistics" except in the context of citing prior work which uses the term.***
 
 44.
-2nd to last line in the abstract: In point (a) ``correctly identify lands with too much damage to be suitable for matching" I suggest replacing the word ``matching" with ``comparison" to use the same terminology as the other pattern evidence disciplines.
+2nd to last line in the abstract: In point (a) "correctly identify lands with too much damage to be suitable for matching" I suggest replacing the word "matching" with "comparison" to use the same terminology as the other pattern evidence disciplines.
 
@@
 
-45.
-Line 55: Replace ``in the forensic sciences" with ``in forensic science".
-(See Margot P, Commentary on The Need for a Research Culture in the Forensic Sciences, UCLA Law Review (2011) 58: 795-801.)
+45. Line 55: Replace "in the forensic sciences" with "in forensic science". (See Margot P, Commentary on The Need for a Research Culture in the Forensic Sciences, UCLA Law Review (2011) 58: 795-801.)
 
@@
 46.
-Lines 156-158: The sentence ``The blue area around the signature are 95% bootstrapped confidence intervals based on 1,000 bootstrap samples." is confusing.
-The term ``signature" is defined two sentences further.
+Lines 156-158: The sentence "The blue area around the signature are 95% bootstrapped confidence intervals based on 1,000 bootstrap samples." is confusing.
+The term "signature" is defined two sentences further.
 Is the sentence referring to Figure 4b?
@@
 
-48. Line 246: Replace ``this representations" with ``this representation" or ``these representations"
+48. Line 246: Replace "this representations" with "this representation" or "these representations"
 
@@
 
-49. Caption of Figure 8: ``At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures." is not a complete sentence.
+49. Caption of Figure 8: "At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures." is not a complete sentence.
 
-***We have revised this caption to remove the ambiguity from the word ``peaks", which we had intended to mean "maximizes."***
+***We have revised this caption to remove the ambiguity from the word "peaks", which we had intended to mean "maximizes."***
 
-50. Line 266: Capitalize the ``b" in ``by".
+50. Line 266: Capitalize the "b" in "by".
 
 ***DONE***
+
+***EH: Add reviewers 2 and 3 to the document***

### 2016-04-23 revision/response.Rmd
Tweak a few more of the responses

@@
 XXX still need to do XXX Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel.
-The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.
-***
+The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.***
 
@@
 ***Grooves and striations are actually not synonyms in this case.
-The grooves are impressions induced by the rifling of the gun barrel, and properties associated with them are class characteristics not unique to individual gun barrels.
+The grooves are impressions induced by the rifling of the gun, and properties associated with them are class characteristics not unique to individual gun barrels.
 We've taken care to ensure the distinction is clear between the striae of interest, which are extracted from the lands of the bullet (between grooves), and the grooves themselves.***
@@
 In our particular data, all comparisons with a CMS of 13 or more were in fact matches.
-Still, it is likely due to pure chance alone that, particularly if more bullets were used in comparison, a small portion would have high pairwise CMS despite being non-matches.***
-
-***EH: Talk about adding features decreasing error rate. CMS by itself is not a sufficient condition for a match...***
+Still, it is likely due to pure chance alone that, particularly if more bullets were used in comparison, a small portion would have high pairwise CMS despite being non-matches.
+Adding features beyond just CMS will decrease this error rate - In other words, a high CMS value is not a sufficient condition for a match.***
 
@@
 
-***EH: Take out all references to x1 and x2***
+***We have removed references to x1 and x2.
+Thanks for pointing them out.
+These were features representing the height of the bullet at which the signature was extracted for each of the two bullets, but they were not intended to be part of the random forest.***
 
@@
 ***DONE***
-
-***EH: Add reviewers 2 and 3 to the document***

### 2016-04-23 code/test_groovelocations.R
Update manual groove detection

@@ -3,24 +3,41 @@ library(x3prplus)
 library(ggplot2)
 library(dplyr)
 
-grooves <- read.csv("csvs/grooves-mean.csv")
+ccs <- read.csv("csvs/crosscuts-25.csv")
+grooves <- read.csv("csvs/grooves.csv")
 
-lapply(as.data.frame(t(grooves)), function(x) {
-    mybullet <- read.x3pplus(as.character(x[1]))
-    fortified <- fortify_x3p(mybullet)
+apply(ccs, 1, function(row) {
+    cat(row[1])
     
-    cc <- fortified %>%
-        group_by(y) %>%
-        summarise(value = mean(value, na.rm = TRUE))
+    mybullet <- get_crosscut(row[1], x = as.numeric(row[2]))
+    relevant.groove <- filter(grooves, bullet == row[1], x == mybullet$x[1])
     
-    cc2 <- get_crosscut(bullet = mybullet, x = 136)
-    
-    print(qplot(y, value, data = cc, geom = "line",colour = I("red")) +
+    print(qplot(y, value, data = mybullet, geom = "line") +
         theme_bw() +
-        geom_vline(xintercept = as.numeric(as.character(x[2])), colour = I("blue")) +
-        geom_vline(xintercept = as.numeric(as.character(x[3])), colour = I("blue")) +
-        geom_line(data = cc2, alpha = I(0.8)))
-    
-    value <- readline("Press Enter to Continue, or q to Quit")
+        geom_vline(xintercept = relevant.groove$groove_left, color = "blue") +
+        geom_vline(xintercept = relevant.groove$groove_right, color = "blue"))
+
+    value <- readline("Press Enter to Continue, g to enter groove, or q to Quit")
     if (value == "q") break;
+    
+    if (value == "g") {
+        newvalue <- ""
+        while (newvalue != "g") {
+            left <- as.numeric(readline("Enter the coordinate of the left groove "))
+            right <- as.numeric(readline("Enter the coordinate of the right groove "))
+            
+            print(qplot(y, value, data = mybullet, geom = "line") +
+                theme_bw() +
+                geom_vline(xintercept = left, color = "blue") +
+                geom_vline(xintercept = right, color = "blue"))
+            
+            newvalue <- readline("Press g to confirm, or any other key to try again")
+        }
+        
[…]

### 2016-04-23 code/complete-hamby.R
Bullet stats updated

@@ -39,12 +39,12 @@ for (span in c(25)) {
     dataStr <- sprintf("data-%d-25", span) # using crosscuts-25.csv
     
     if (!file.exists(dataStr)) dir.create(dataStr)
-    for (j in 1:length(all_bullets)) {
-        reslist <- lapply(all_bullets, function(x) {
+    for (j in 1:length(unknowns)) {
+        reslist <- lapply(knowns, function(x) {
             cat("Processing", j, "vs", basename(x$path), "with span", span, "\n")
             
             br1 <- filter(bullets_smoothed, bullet == x$path)
-            br2 <- filter(bullets_smoothed, bullet == all_bullets[[j]]$path)
+            br2 <- filter(bullets_smoothed, bullet == unknowns[[j]]$path)
             
             if (all_bullets[[j]]$path != x$path) {
                 bulletGetMaxCMS(br1, br2, span=span)

### 2016-04-25 imaging-paper.Rnw
Small typos

@@
 We first have to identify the locations of the left and right groove in the image.
-The grooves are not assumed to contain information relevant for determining matches.
+The grooves are assumed to not contain information relevant for determining matches.
 They also dominate the structure, and therefore need  to be removed.
@@
 
-Counting the maximal number of CMS is part of current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
+Counting the maximal number of CMS is part of the current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
 In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a  high number indicative of a match between the bullets. 

### 2016-04-25 code/complete-hamby-set44.R
New set44 grooves
> xtabs(~(pred>0.5)+match+span, data=bstats)
, , span = 25

          match
pred > 0.5 FALSE TRUE
     FALSE  9784   54
     TRUE     10  112

@@ -38,10 +38,10 @@ for (span in c(25)) {
     dataStr <- sprintf("data-set44-%d-25", span) # using crosscuts-25.csv
     
     if (!file.exists(dataStr)) dir.create(dataStr)
-    for (j in 1:210) {
-        reslist <- lapply(c(knowns, unknowns), function(x) {
+    for (j in 1:90) {
+        reslist <- lapply(knowns, function(x) {
             cat("Processing", j, "vs", basename(x$path), "with span", span, "\n")
-            sigh <- c(knowns, unknowns)[[j]]
+            sigh <- unknowns[[j]]
             
             br1 <- filter(bullets_smoothed, bullet == x$path)
             br2 <- filter(bullets_smoothed, bullet == sigh$path)

### 2016-04-25 code/test_groovelocations.R

@@ -3,13 +3,13 @@ library(x3prplus)
 library(ggplot2)
 library(dplyr)
 
-ccs <- read.csv("csvs/crosscuts-25.csv")
-grooves <- read.csv("csvs/grooves.csv")
+ccs <- read.csv("csvs/crosscuts-25-set44.csv")
+grooves <- read.csv("csvs/grooves-set44.csv")
 
-apply(ccs, 1, function(row) {
+apply(ccs[-(1:27),], 1, function(row) {
     cat(row[1])
     
-    mybullet <- get_crosscut(row[1], x = as.numeric(row[2]))
+    mybullet <- get_crosscut(row[1], x = as.numeric(row[2]), transpose = TRUE)
     relevant.groove <- filter(grooves, bullet == row[1], x == mybullet$x[1])
     
     print(qplot(y, value, data = mybullet, geom = "line") +
@@ -40,4 +40,4 @@ apply(ccs, 1, function(row) {
     }
 })
 
-write.csv(grooves, file = "csvs/grooves.csv", row.names = FALSE)
+write.csv(grooves, file = "csvs/grooves-set44.csv", row.names = FALSE)

### 2016-04-27 imaging-paper.Rnw
Merge the set44 with the regular set

@@
 images <- file.path(datadir, dir(datadir))
+images <- images[grep(" ", images)]
 
@@
 paths <- file.path(datadir, dir(datadir))
+paths <- paths[grep(" ", paths)]
+
 crosscuts <- seq(100,250, by=25)
@@
 includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))
-cc <- read.csv("csvs/crosscuts-25.csv")
+#cc <- read.csv("csvs/crosscuts-25.csv")
 # cc$bullet <- gsub(".*//","",cc$path)

### 2016-04-27 imaging-supplement.Rnw
Merge the set44 with the regular set

@@
 paths <- file.path(datadir, dir(datadir))
+paths <- paths[grep(" ", paths)]
+
 crosscuts <- seq(100, 400, by = 50)
@@
 # match unknown land using crosscuts
-ccs <- read.csv("csvs/crosscuts-25.csv")
+ccs <- read.csv("csvs/crosscuts-25-old.csv")
 all_bullets <- lapply(as.character(ccs$path), function(x) {
@@
 knowns <- dir(path=knowndatadir)
+knowns <- knowns[grep(" ", knowns)]
 knowns <- gsub(".x3p", "", knowns)
@@
 unknowns <- dir(path=unknowndatadir)
+unknowns <- unknowns[grep(" ", unknowns)]
 unknowns <- gsub(".x3p", "", unknowns)
@@
 <<aligned, echo=FALSE, warning=FALSE,  cache=TRUE>>=
-matches <- read.csv("csvs/matches.csv", header = FALSE)
+matches <- read.csv("csvs/matches-old.csv", header = FALSE)
 matches$barrel <- rep(1:10, each=6)
@@
 
-ccs <- read.csv("csvs/crosscuts-25.csv")
+ccs <- read.csv("csvs/crosscuts-25-old.csv")
 mm <- merge(mm, ccs, by="path", all=TRUE)

### 2016-04-27 code/checkBullets-degraded.R

@@ -1,29 +1,0 @@
-library(x3pr)
-library(x3prplus)
-library(dplyr)
-library(ggplot2)
-library(zoo)
-
-knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
-
-unknowndatadir <- "app/degraded_images/Hamby252_3DX3P1of2/"
-unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
-
-if (!file.exists("csvs/crosscuts-25-degraded.csv")) {
-crosscuts <- sapply(c(knowns, unknowns), function(x) {
-    cat(x)
-    crosscut <- bulletCheckCrossCut(x, x = seq(100, 500, by=25), span = 0.15, minccf = .99)
-    cat(crosscut)
-    cat("\n")
-    crosscut
-})
-
-dframe <- data.frame(
-  path = c(knowns, unknowns),
-  cc = crosscuts
-)
-
-
-write.csv(dframe, file="csvs/crosscuts-25-degraded.csv", row.names=FALSE)
-}

### 2016-04-27 code/checkBullets-set44.R

@@ -1,30 +1,0 @@
-library(x3pr)
-library(x3prplus)
-library(dplyr)
-library(ggplot2)
-library(zoo)
-
-knowndatadir <- "~/Downloads/Hamby Set 44/known"
-knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
-
-unknowndatadir <- "~/Downloads/Hamby Set 44/unknown"
-unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
-
-if (!file.exists("csvs/crosscuts-25-set44.csv")) {
-crosscuts <- sapply(c(knowns, unknowns), function(x) {
-    cat(x)
-    crosscut <- bulletCheckCrossCut(x, xlimits = seq(25, 500, by = 25), transpose = TRUE)
-    cat(crosscut)
-    cat("\n")
-    crosscut
-})
-crosscuts[is.na(crosscuts)] <- 150
-
-dframe <- data.frame(
-  path = c(knowns, unknowns),
-  cc = crosscuts
-)
-
-
-write.csv(dframe, file="csvs/crosscuts-25-set44.csv", row.names=FALSE)
-}

### 2016-04-27 code/checkBullets.R

@@ -12,8 +12,9 @@ unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
 if (!file.exists("csvs/crosscuts-25.csv")) {
 crosscuts <- sapply(c(knowns, unknowns), function(x) {
-  cat(x)
-  crosscut <- bulletCheckCrossCut(x, xlimits = seq(25, 500, by = 25))
+    transpose <- (length(grep(" ", x)) == 0)
+  cat(x, "- transposing:", transpose)
+  crosscut <- bulletCheckCrossCut(x, xlimits = seq(25, 500, by = 25), transpose = transpose)
   cat(crosscut)
   cat("\n")
   crosscut

### 2016-04-27 code/complete-hamby-degraded.R

@@ -1,108 +1,0 @@
-library(x3pr)
-library(x3prplus)
-library(dplyr)
-library(ggplot2)
-library(gridExtra)
-library(zoo)
-
-knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-unknowndatadir <- "app/degraded_images/Hamby252_3DX3P1of2/"
-
-
-###############
-# can we identify the barrels the unknown bullets came from?
-
-# new_get_bullet <- function(path, x = 243.75, bullet = NULL) {
-#     if (is.null(bullet)) bullet <- read.x3p(path)
-#     dbr111 <- fortify_x3p(bullet)
-#     
-#     pickx <- dbr111$x[which.min(abs(x - unique(dbr111$x)))]
-#     
-#     dbr111.fixx <- dbr111[dbr111$x == pickx,]
-#     
-#     return(dbr111.fixx)
-# }
-
-# new_processBullets <- function(bullet, x = 100) {
-#     crosscuts <- unique(fortify_x3p(bullet)$x)
-#     crosscuts <- crosscuts[crosscuts >= min(x)]
-#     crosscuts <- crosscuts[crosscuts <= max(x)]
-#     if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]
-#     
-#     list_of_fits <- lapply(crosscuts, function(x) {
-#         br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
-#         br111.groove <- get_grooves(br111)
-#         fit_loess(br111, br111.groove)$resid$data
-#     })
-#     lof <- list_of_fits %>% bind_rows
-#     
-#     data.frame(lof, bullet = bullet$path, stringsAsFactors = FALSE)
-# }
-
-# new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
-#     lof <- rbind(lof1, lof2)
-#     bAlign = bulletAlign(lof)
-#     lofX <- bAlign$bullet  
-#     
-#     b12 <- unique(lof$bullet)
-#     peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
-#     peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)
-#     
-#     #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
-#     #    theme_bw() +
[…]

### 2016-04-27 code/complete-hamby-set44.R

@@ -1,54 +1,0 @@
-library(x3pr)
-library(x3prplus)
-library(dplyr)
-library(ggplot2)
-library(gridExtra)
-library(zoo)
-
-knowndatadir <- "~/Downloads/Hamby Set 44/known"
-unknowndatadir <- "~/Downloads/Hamby Set 44/unknown"
-
-# match unknown land using crosscuts
-ccs <- read.csv("csvs/crosscuts-25-set44.csv")
-all_bullets <- lapply(as.character(ccs$path), function(x) {
-    result <- read.x3pplus(x, transpose = TRUE)
-    result[[3]] <- x
-    names(result)[3] <- "path"
-        
-    return(result)
-})
-
-knowns <- all_bullets[1:120]
-unknowns <- all_bullets[121:210]
-grooves <- read.csv("csvs/grooves-set44.csv")
-bullets_processed <- lapply(all_bullets, function(bul) {
-    cat("Computing processed bullet", basename(bul$path), "\n")
-    
-    xval <- ccs$cc[which(ccs$path == bul$path)]
-    left <- grooves$groove_left[grooves$bullet == bul$path & grooves$x == xval]
-    right <- grooves$groove_right[grooves$bullet == bul$path & grooves$x == xval]
-    
-    processBullets(bullet = bul, name = bul$path, x = xval, grooves = c(left, right))
-})
-names(bullets_processed) <- as.character(ccs$path)
-
-bullets_smoothed <- bullets_processed %>% bind_rows %>% bulletSmooth
-
-for (span in c(25)) {
-    dataStr <- sprintf("data-set44-%d-25", span) # using crosscuts-25.csv
-    
-    if (!file.exists(dataStr)) dir.create(dataStr)
-    for (j in 1:90) {
-        reslist <- lapply(knowns, function(x) {
-            cat("Processing", j, "vs", basename(x$path), "with span", span, "\n")
-            sigh <- unknowns[[j]]
-            
-            br1 <- filter(bullets_smoothed, bullet == x$path)
-            br2 <- filter(bullets_smoothed, bullet == sigh$path)
-            
-            if (x$path == sigh$path) return(NULL)
-            bulletGetMaxCMS(br1, br2, span=span)
-        })
-        save(reslist, file=file.path(dataStr, sprintf("unkn%d.RData", j)))
-    }
-}

### 2016-04-27 code/complete-hamby.R

@@ -11,7 +11,9 @@ unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 # match unknown land using crosscuts
 ccs <- read.csv("csvs/crosscuts-25.csv")
 all_bullets <- lapply(as.character(ccs$path), function(x) {
-    result <- read.x3pplus(x)
+    transpose <- (length(grep(" ", x)) == 0)
+    
+    result <- read.x3pplus(x, transpose = transpose)
     result[[3]] <- x
     names(result)[3] <- "path"
         

### 2016-04-27 code/evaluation-hamby-degraded.R

@@ -1,175 +1,0 @@
-span <- 25
-dataStr <- sprintf("data-degraded-%d-25", span)
-datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
-datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
-
-CCFs <- plyr::ldply(datas, function(x) {
-  load(x)
-#  cmsdist <- sapply(reslist, function(x) x$maxCMS)
-  ccf <- plyr::ldply(reslist, function(res) {
-    lofX <- res$bullets
-    b12 <- unique(lofX$bullet)
-    
-    subLOFx1 <- subset(lofX, bullet==b12[1])
-    subLOFx2 <- subset(lofX, bullet==b12[2]) 
-    
-    subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
-    subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-
-    ccf <- ccf(subLOFx1$val, subLOFx2$val, plot = FALSE, lag.max=200, na.action = na.omit)
-    lag <- ccf$lag[which.max(ccf$acf)]
-    incr <- min(diff(sort(unique(subLOFx1$y))))
-    
-    subLOFx1$y <- subLOFx1$y -  lag * incr # amount of shifting should just be lag * y.inc
-    ys <- intersect(subLOFx1$y, subLOFx2$y)
-    idx1 <- which(subLOFx1$y %in% ys)
-    idx2 <- which(subLOFx2$y %in% ys)
-    distr.dist <- mean((subLOFx1$val[idx1] - subLOFx2$val[idx2])^2, na.rm=TRUE)
-    distr.sd <- sd(subLOFx1$val, na.rm=TRUE) + sd(subLOFx2$val, na.rm=TRUE)
-    km <- which(res$lines$match)
-    knm <- which(!res$lines$match)
-    if (length(km) == 0) km <- c(length(knm)+1,0)
-    if (length(knm) == 0) knm <- c(length(km)+1,0)
- #browser()    
-    # feature extraction
-    signature.length <- min(nrow(subLOFx1), nrow(subLOFx2))
-    
-    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), 
-               D=distr.dist, 
-               sd.D = distr.sd,
-               b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
-               num.matches = sum(res$lines$match), 
-               signature.length = signature.length,
[…]

### 2016-04-27 code/evaluation-hamby-set44.R

@@ -1,162 +1,0 @@
-span <- 25
-dataStr <- sprintf("data-set44-%d-25", span)
-datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
-datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
-
-CCFs <- plyr::ldply(datas, function(x) {
-  load(x)
-#  cmsdist <- sapply(reslist, function(x) x$maxCMS)
-  ccf <- plyr::ldply(reslist, function(res) {
-     if (is.null(res)) return(NULL)
-    lofX <- res$bullets
-    b12 <- unique(lofX$bullet)
-    
-    subLOFx1 <- subset(lofX, bullet==b12[1])
-    subLOFx2 <- subset(lofX, bullet==b12[2]) 
-    
-    subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
-    subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-
-    ccf <- ccf(subLOFx1$val, subLOFx2$val, plot = FALSE, lag.max=200, na.action = na.omit)
-    lag <- ccf$lag[which.max(ccf$acf)]
-    incr <- min(diff(sort(unique(subLOFx1$y))))
-    
-    subLOFx1$y <- subLOFx1$y -  lag * incr # amount of shifting should just be lag * y.inc
-    ys <- intersect(subLOFx1$y, subLOFx2$y)
-    idx1 <- which(subLOFx1$y %in% ys)
-    idx2 <- which(subLOFx2$y %in% ys)
-    distr.dist <- mean((subLOFx1$val[idx1] - subLOFx2$val[idx2])^2, na.rm=TRUE)
-    distr.sd <- sd(subLOFx1$val, na.rm=TRUE) + sd(subLOFx2$val, na.rm=TRUE)
-    km <- which(res$lines$match)
-    knm <- which(!res$lines$match)
-    if (length(km) == 0) km <- c(length(knm)+1,0)
-    if (length(knm) == 0) knm <- c(length(km)+1,0)
- #browser()    
-    # feature extraction
-    signature.length <- min(nrow(subLOFx1), nrow(subLOFx2))
-    
-    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), 
-               D=distr.dist, 
-               sd.D = distr.sd,
-               b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
[…]

### 2016-04-27 code/evaluation-hamby.R

@@ -1,7 +1,7 @@
 span <- 25
 dataStr <- sprintf("data-%d-25", span)
 datas <- file.path(dataStr, dir(dataStr, pattern="RData"))
-datas <- datas[grep(paste0(dataStr,"/u.*"), datas)]
+datas <- datas[grep(paste0(dataStr,"/u.*-new.RData"), datas)]
 
 CCFs <- plyr::ldply(datas, function(x) {
   load(x)
@@ -14,14 +14,6 @@ CCFs <- plyr::ldply(datas, function(x) {
     subLOFx1 <- subset(lofX, bullet==b12[1])
     subLOFx2 <- subset(lofX, bullet==b12[2]) 
     
-    subLOFx1$y <- subLOFx1$y - min(subLOFx1$y)
-    subLOFx2$y <- subLOFx2$y - min(subLOFx2$y)
-
-    ccf <- ccf(subLOFx1$val, subLOFx2$val, plot = FALSE, lag.max=200, na.action = na.omit)
-    lag <- ccf$lag[which.max(ccf$acf)]
-    incr <- min(diff(sort(unique(subLOFx1$y))))
-    
-    subLOFx1$y <- subLOFx1$y -  lag * incr # amount of shifting should just be lag * y.inc
     ys <- intersect(subLOFx1$y, subLOFx2$y)
     idx1 <- which(subLOFx1$y %in% ys)
     idx2 <- which(subLOFx2$y %in% ys)
@@ -35,7 +27,7 @@ CCFs <- plyr::ldply(datas, function(x) {
     # feature extraction
     signature.length <- min(nrow(subLOFx1), nrow(subLOFx2))
     
-    data.frame(ccf=max(ccf$acf), lag=which.max(ccf$acf), 
+    data.frame(ccf=res$ccf, lag=res$lag, 
                D=distr.dist, 
                sd.D = distr.sd,
                b1=b12[1], b2=b12[2], x1 = subLOFx1$x[1], x2 = subLOFx2$x[1],
@@ -81,9 +73,6 @@ CCFs$b2 <- gsub(".x3p","", CCFs$b2)
 
 
 matches <- read.csv("csvs/matches.csv", header=FALSE, stringsAsFactors = FALSE)
-matches$V3 <- paste("Ukn Bullet",matches$V3)
-matches$V4 <- paste("Ukn Bullet",matches$V4)
-matches$V5 <- paste("Ukn Bullet",matches$V5)
 matches$id <- 1:nrow(matches)
 
 library(reshape2)

### 2016-04-27 code/groove_locations.R

@@ -11,8 +11,10 @@ unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
 unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
 
 groove_locations_means <- lapply(c(knowns, unknowns), function(bul) {
+    transpose <- (length(grep(" ", bul)) == 0)
+    
     cat(bul, "\n")
-    bullet <- read.x3pplus(bul)
+    bullet <- read.x3pplus(bul, transpose = transpose)
     fortified <- fortify_x3p(bullet)
     
     fort.mean <- fortified %>%
@@ -37,9 +39,11 @@ lapply(as.data.frame(t(groove.means)), function(test) {
     left <- as.numeric(as.character(test[2]))
     right <- as.numeric(as.character(test[3]))
     
+    transpose <- (length(grep(" ", bul)) == 0)
+    
     cat(bul, "\n")
-    bullet <- read.x3pplus(bul)
-    cc <- get_crosscut(bullet = bullet, x = 150)
+    bullet <- read.x3pplus(bul, transpose = transpose)
+    cc <- get_crosscut(bullet = bullet, x = 150, transpose = transpose)
     result <- get_grooves(cc, mean_left = left, mean_right = right, mean_window = mean_window, smoothfactor = 25)
     
     print(result$plot)
@@ -53,13 +57,15 @@ groove_locations <- lapply(as.data.frame(t(groove.means)), function(test) {
     left <- as.numeric(as.character(test[2]))
     right <- as.numeric(as.character(test[3]))
     
+    transpose <- (length(grep(" ", bul)) == 0)
+    
     cat(bul, "\n")
-    bullet <- read.x3pplus(bul)
+    bullet <- read.x3pplus(bul, transpose = transpose)
     fortified <- fortify_x3p(bullet)
     all_grooves <- lapply(unique(fortified$x), function(x) {
         myresult <- try({
             cat("Processing", x, "\n")
-            cc <- get_crosscut(bullet = bullet, x = x)
+            cc <- get_crosscut(bullet = bullet, x = x, transpose = transpose)
             result <- get_grooves(cc, mean_left = left, mean_right = right, mean_window = mean_window, smoothfactor = 25)
             
             result$groove

### 2016-04-27 code/groove_locations_set44.R

@@ -1,80 +1,0 @@
-library(x3pr)
-library(x3prplus)
-library(dplyr)
-library(ggplot2)
-library(zoo)
-
-knowndatadir <- "~/Downloads/Hamby Set 44/known"
-knowns <- file.path(knowndatadir, dir(knowndatadir, pattern="x3p"))
-
-unknowndatadir <- "~/Downloads/Hamby Set 44/unknown"
-unknowns <- file.path(unknowndatadir, dir(unknowndatadir))
-
-groove_locations_means <- lapply(c(knowns, unknowns), function(bul) {
-    cat(bul, "\n")
-    bullet <- read.x3pplus(bul, transpose = TRUE)
-    fortified <- fortify_x3p(bullet)
-    
-    fort.mean <- fortified %>%
-        filter(x >= 100, x <= 300) %>%
-        group_by(y) %>%
-        summarise(value = mean(value, na.rm = TRUE))
-    
-    cc <- cbind(x = NA, fort.mean)
-    result <- get_grooves(cc, smoothfactor = 25, adjust = 0, groove_cutoff = 500)
-
-    return(result$groove)
-})
-groove.means <- cbind(data.frame(bullet = c(knowns, unknowns), do.call(rbind, groove_locations_means)))
-names(groove.means) <- c("bullet", "groove_left", "groove_right")
-write.csv(groove.means, file = "csvs/grooves-mean-set44.csv", row.names = FALSE)
-
-groove.means <- read.csv("csvs/grooves-mean-set44.csv")
-mean_window <- 150
-
-lapply(as.data.frame(t(groove.means)), function(test) {
-    bul <- as.character(test[1])
-    left <- as.numeric(as.character(test[2]))
-    right <- as.numeric(as.character(test[3]))
-    
-    cat(bul, "\n")
-    bullet <- read.x3pplus(bul, transpose = TRUE)
-    cc <- get_crosscut(bullet = bullet, x = 150)
-    result <- get_grooves(cc, mean_left = left, mean_right = right, mean_window = mean_window, smoothfactor = 25)
-    
-    print(result$plot)
-    
-    value <- readline("Press Enter to Continue, or q to Quit")
-    if (value == "q") break;    
-})
-
-groove_locations <- lapply(as.data.frame(t(groove.means)), function(test) {
-    bul <- as.character(test[1])
-    left <- as.numeric(as.character(test[2]))
[…]

### 2016-04-27 code/test_align.R

@@ -2,8 +2,6 @@ library(ggplot2)
 library(x3pr)
 library(x3prplus)
 library(dplyr)
-
-
     
 matches <- read.csv("csvs/matches.csv", header = FALSE)
 ccs <- read.csv("csvs/crosscuts-25.csv")

### 2016-04-27 code/complete-hamby.R
Oops, increase the indices now

@@ -22,8 +22,8 @@ all_bullets <- lapply(as.character(ccs$path), function(x) {
 
 grooves <- read.csv("csvs/grooves.csv")
 
-knowns <- all_bullets[1:120]
-unknowns <- all_bullets[121:210]
+knowns <- all_bullets[1:240]
+unknowns <- all_bullets[241:420]
 bullets_processed <- lapply(all_bullets, function(bul) {
     cat("Computing processed bullet", basename(bul$path), "\n")
     

### 2016-04-28 revision/response.Rmd
Update response about width of striae

@@
 1. The paper should better explain the role of the new methods for forensic science. The authors should consider referencing some of the previous work the reviewers mention.
+
 2.
@@
 
-***why should the probability of two sets of observations given that the two bullets were fired by the same gun be the probability of interest?
+***HH: why should the probability of two sets of observations given that the two bullets were fired by the same gun be the probability of interest?
 I would agree that the random forest and any of its conclusions are restricted to how good the underlying 'database' is.
@@
 ***Grooves and striations are actually not synonyms in this case.
-The grooves are impressions induced by the rifling of the gun, and properties associated with them are class characteristics not unique to individual gun barrels.
+The grooves are impressions induced by the rifling of the barrel, and properties associated with them are class characteristics not unique to individual gun barrels.
 We've taken care to ensure the distinction is clear between the striae of interest, which are extracted from the lands of the bullet (between grooves), and the grooves themselves.***
@@
 We might need some more literature.***
+
+***EH: http://nvlpubs.nist.gov/nistpubs/jres/116/3/v116.n03.a01.pdf - "The number and width of striations in a land impression can vary depending on the size of the bullet and the characteristics of the barrel.
+For example, according to Biasotti’s study, the Smith and Wesson barrel with a land impression width of 2.4 mm has an average of 60 striations with total striation counts between 16 and 97 [13].
+Observing the land impression images used in our study, 50 pixels (about 0.08 mm) is a reasonable choice for the average width of a structuring element."***
 

### 2016-05-03 imaging-paper.Rnw
Refactor some library code

@@
 library(sm)
+library(reshape2)
+library(randomForest)
 @
@@
 })
-library(reshape2)
+
 dm <- melt(densities, measure.var=c("KM", "KNM"))
@@
 <<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.7\\textwidth'>>=
-library(rpart)
-library(rpart.plot)
-library(RColorBrewer)
 vals = alpha(brewer.pal(3, name="Paired"), alpha=0.5)
@@
 <<rforest, echo=FALSE, message=FALSE>>=
-library(randomForest)
-library(reshape2)
-
 set.seed(20160105)

### 2016-05-04 revision/response.Rmd
Work on some more of the responses

@@
 
+***AC: TODO***
+
 2.
@@
 
+***We have detailed a multi-stage analysis process that begins with a source image, processes the image, extracts features, and ultimately uses a random forest to compute predicted match probabilities.
+While the forest is a component of the process, we've taken care to emphasize that the preceeding steps outline a more general framework for analysis of striae in images that could theoretically apply to other forensic science questions as well.
+
+***EH: TODO: Add to paper***
+
 # Reviewer A
 
-4.
+3.
 Random forests, as used and described on pages 18-21 of the paper comes out looking like a winner, but other than that success---worth reporting *somewhere*---, there is no new information on that method.
@@
 
+***HH: Could you look at this response i wrote?***
+
+***We were not intending to provide new insights on random forests.
+The insights were in the feature extraction and image processing leading up to the random forest fit.
+As the forest ultimately predicted every land-to-land comparison properly, we felt that a comparison to other learning methods wass outside the scope of this paper.
+Had a wider variety of data been available, there would likely be more variability in the predictions and a comparison to other learning methods would be prudent.
+For this particular paper, however, the groove detection, profile extraction, and feature definitions were the intended focus.***
+
 # Reviewer B
 
-5.
+4.
 The authors should discuss their work in the context of Nicholas Petraco’s work as well as other ongoing research endeavors in impression and pattern evidence.
@@
 
-6. page 7- figure 4- What exactly are you bootstrapping? It is unclear in the paper whether you are bootstrapping residual or bullets or something else entirely. Please clarify.
+5. page 7- figure 4- What exactly are you bootstrapping? It is unclear in the paper whether you are bootstrapping residual or bullets or something else entirely. Please clarify.
 
@@
 
-7. There should have been a comparison of different classification methods to show that the random forest is reasonable.
+6. There should have been a comparison of different classification methods to show that the random forest is reasonable.
 
-***I am not quite sure what a 'reasonable' classification method is supposed to mean.
+***HH: I am not quite sure what a 'reasonable' classification method is supposed to mean.
 Shouldn't a classification method be judged by its success in predicting the outcome?
@@
 
-8. On the random forest classifier- are the prior probabilities being estimated from the cross comparisons described on page 15? What exactly are the prior probabilities used in the random forest?
+7. On the random forest classifier- are the prior probabilities being estimated from the cross comparisons described on page 15? What exactly are the prior probabilities used in the random forest?
 
@@
 
-9. Equal Error Rates on the ROC curves would be a helpful metric as well as the AUC.
+8. Equal Error Rates on the ROC curves would be a helpful metric as well as the AUC.
 
-***included***
+***Included***
 
-10.
+9.
 Would it be possible to stress the classification system by only using part of the profile?
@@
 
-11. What is the standard error of the error rates presented for the matching algorithms? The number of cross comparisons gives false impression the standard errors the when presenting the results.
+10. What is the standard error of the error rates presented for the matching algorithms? The number of cross comparisons gives false impression the standard errors the when presenting the results.
 
@@
 
-12. In the conclusions section the authors mention an assumption of "uniqueness". In general forensic science have moved beyond these questions- see the work of Saks and Kaye for details.
+11. In the conclusions section the authors mention an assumption of "uniqueness". In general forensic science have moved beyond these questions- see the work of Saks and Kaye for details.
+
+***AC: TODO***
 
@@
 
-13.
+12.
 A discussion on how the authors view the use of their matching algorithm would be useful to the forensic community.
@@
 
-14.
+13.
 there is a lack of knowledge of the existing literature on statistical work performed on tool marks and firearms.
@@
 
-15.
+***TODO***
+
+14.
 The authors would need to study (or at least discuss extensively) the ability of their method to capture the variability in bullet profiles over time (at interval of several dozen of bullets fired by the same weapon, when the weapon has been cleaned, or when the bullet was degraded by hitting a target).
@@
 
-16.
+15.
 On a related subject, I believe that a random forest provides the probability that two bullets were fired by the same gun given the set of observations (Fig 19), not the probability of interest: the probability of two sets of observations given that the two bullets were fired by the same gun.
@@
 
-17.
+16.
 I am concerned by the assumption of independence of between the different lands on a given bullet that the authors seems to use in their error rates calculations (and ROC curves).
@@
 
-18.
+17.
 the authors attempt to define a "stable" profile (see Figure 9) and use the stable profile for comparison.
@@
 
-19.
+18.
 L323: how do the authors know that there are "172 known matches" since the 15 "trace bullets" are supposed to be of unknown sources?
@@
 
-20. If would be helpful if the legend from Fig15 could include a definition of the CMS, CNMS, CCF, ... acronyms such that the reader does not need to flip back and forth between several pages.
+19. If would be helpful if the legend from Fig15 could include a definition of the CMS, CNMS, CCF, ... acronyms such that the reader does not need to flip back and forth between several pages.
 
@@
 
-21.
+20.
 First, the term "land" is used in the title but not defined until p.
@@
 
-22. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.
+21. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.
 
@@
 
-23. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.
+22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.
 
@@
 
-24. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.
+23. p. 2, l. -1:  It would be good to reword so the sentence doesn't start with "x3p".  Also p. 3, l. 2.
 
@@
 
-25. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.
+24. p. 3, Fig. 1:  The numbers on the y-axis are superimposed.
 
@@
 
-26. p. 7, l. 1:  Make this "... can be described in the form ...".
+25. p. 7, l. 1:  Make this "... can be described in the form ...".
 
@@
 
-27. p. 7, l. 9:  Remove the comma after "here" in "... function used here, ...".
+26. p. 7, l. 9:  Remove the comma after "here" in "... function used here, ...".
 
@@
 
-28. p. 7, l. -6:  I'm confused.  The text says "bullets 2 are shown" but the caption refers to bullet 2- which I think should be a specific land on a single bullet.
+27. p. 7, l. -6:  I'm confused.  The text says "bullets 2 are shown" but the caption refers to bullet 2- which I think should be a specific land on a single bullet.
 
@@
 
-29.
+28.
 p.
@@
 
-30. p. 13, l. -3:  You write "Given that striation patterns are typically much larger, ..." and I think it would be more clear to write "Given that striation widths are typically much larger, ...".
+29. p. 13, l. -3:  You write "Given that striation patterns are typically much larger, ..." and I think it would be more clear to write "Given that striation widths are typically much larger, ...".
 
@@
 
-31.
+30.
 p.
@@
 
-32. p. 17, l. -13:  I'd prefer another term than "correlation" here.  In fact, the entire sentence could be improved.
+31. p. 17, l. -13:  I'd prefer another term than "correlation" here.  In fact, the entire sentence could be improved.
 
@@
 
-33. p. 17, l. -2:  Remove the comma after "indicates".
+32. p. 17, l. -2:  Remove the comma after "indicates".
 
@@
 
-34. p. 18, l. -4:  Change "this study" to "that study", so the reader is certain you mean Ma et al. (2001) and not your own.
+33. p. 18, l. -4:  Change "this study" to "that study", so the reader is certain you mean Ma et al. (2001) and not your own.
 
@@
 
-35. p. 20,  l. -3:  I really like the idea of including an inconclusive zone.
+34. p. 20,  l. -3:  I really like the idea of including an inconclusive zone.
 
@@
 
-36.
+35.
 p.
@@
 
-37. p. 22, l. -9:  Change "bottomline" to "bottom line".
+36. p. 22, l. -9:  Change "bottomline" to "bottom line".
 
@@
 
-38. p. 23, l. 4:  Remove the comma after "found".
+37. p. 23, l. 4:  Remove the comma after "found".
 
@@
 
-39. p. 23, l. -13/-5:  I don't think this mention of the Aadhar project adds much, and it distracts from the points you want to highlight about the work in this paper.
+38. p. 23, l. -13/-5:  I don't think this mention of the Aadhar project adds much, and it distracts from the points you want to highlight about the work in this paper.
 
@@
 
-40. Supplement, p. 8, l. 18:  Change "de-facto" to "de facto".
+39. Supplement, p. 8, l. 18:  Change "de-facto" to "de facto".
 
@@
 
-41.
+40.
 So, my suggestion here is to change the sentence in the abstract to something like "Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel." Or "...help the trier of fact determine...".
@@
 
-42.
+41.
 the concept of "match": The primary goal in forensic science is NOT to attempt to match two bullets.
@@
 
-43. I strongly encourage the authors to not follow this trend and use the term "ballistics" only when you are actually talking about ballistics.
+42. I strongly encourage the authors to not follow this trend and use the term "ballistics" only when you are actually talking about ballistics.
 
@@
 
-44.
+43.
 2nd to last line in the abstract: In point (a) "correctly identify lands with too much damage to be suitable for matching" I suggest replacing the word "matching" with "comparison" to use the same terminology as the other pattern evidence disciplines.
@@
 
-45. Line 55: Replace "in the forensic sciences" with "in forensic science". (See Margot P, Commentary on The Need for a Research Culture in the Forensic Sciences, UCLA Law Review (2011) 58: 795-801.)
+44. Line 55: Replace "in the forensic sciences" with "in forensic science". (See Margot P, Commentary on The Need for a Research Culture in the Forensic Sciences, UCLA Law Review (2011) 58: 795-801.)
 
@@
 
-46.
+45.
 Lines 156-158: The sentence "The blue area around the signature are 95% bootstrapped confidence intervals based on 1,000 bootstrap samples." is confusing.
@@
 
-47. Figure 5: In this figure, where is the limit between bullet 1-5 and bullet 2-1?
+46. Figure 5: In this figure, where is the limit between bullet 1-5 and bullet 2-1?
 
@@
 
-48. Line 246: Replace "this representations" with "this representation" or "these representations"
+47. Line 246: Replace "this representations" with "this representation" or "these representations"
 
@@
 
-49. Caption of Figure 8: "At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures." is not a complete sentence.
+48. Caption of Figure 8: "At a lag of -17 the correlation peaks, indicating the largest amount of agreement between the signatures." is not a complete sentence.
 
@@
 
-50. Line 266: Capitalize the "b" in "by".
+49. Line 266: Capitalize the "b" in "by".
 

### 2016-05-06 revision/response.Rmd
Update response, add in new section to conclusion

@@
 ***We have detailed a multi-stage analysis process that begins with a source image, processes the image, extracts features, and ultimately uses a random forest to compute predicted match probabilities.
-While the forest is a component of the process, we've taken care to emphasize that the preceeding steps outline a more general framework for analysis of striae in images that could theoretically apply to other forensic science questions as well.
-
-***EH: TODO: Add to paper***
+While the forest is a component of the process, we've taken care to emphasize that the preceeding steps outline a more general framework for analysis of striae in images that could theoretically apply to other forensic science questions as well.***
 

### 2016-05-07 imaging-paper.Rnw
Update response, add in new section to conclusion
added boxplots - densities do not work, because they are too skew

@@
 
+qplot(data=predm, value, geom="boxplot", fill=match, x=match, alpha=I(0.5)) + 
+  theme_bw() + facet_grid(variable~.) +
+  scale_fill_brewer(palette="Paired", labels=c("KNM", "KM")) +
+   ylab("") + 
+   xlab("Predicted probability of a match") +
+  coord_flip() + 
+  theme(legend.position="none")
+
 # qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
@@
 In particular, `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.
+\eh{The steps outlined in this paper could also help explore other important forensic science problems.
+In particular, more general toolmark examination can benefit from this detection and removal of the overall structure and the extraction of features from deviations of this structure.}
 

### 2016-05-07 imaging-paper.Rnw
Few smaller typos

@@
 The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. 
-The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
+The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
 % Hamby study: automated systems were successfully used for matching:
@@
 
-In an approach to analyze the striation pattern we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land.
+In an approach to analyze the striation pattern, we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 We first have to identify the locations of the left and right groove in the image.
-The grooves are assumed to not contain information relevant for determining matches.
+The grooves are assumed to contain no information relevant for determining matches.
 They also dominate the structure, and therefore need to be removed.

### 2016-05-07 revision/response.Rmd
Add a new response on OOB error

@@
 
+***Under the assumption of land-to-land independence, the number of land-to-land matches is distributed according to a Binomial(N, p) distribution where N is the total number of bullet lands and p is the probability of a match between two bullet lands.
+However, the bullet lands from the Hamby study are all known to be either pairwise matches or non-matches.
+Thus, we present the Out-of-bag (OOB) error for the random forest, representing the error rate as determined from the observations not present in the particular bootstrap samples.
+This is discussed further on Page 19.***
+
 ***HH: I'm sure I am misunderstanding the comment, but I don't know how to answer this question.
@@
 The authors may want to revise some of the statements in their discussion and conclusion.
+
+***AC: TODO***
 

### 2016-05-07 revision/response.Rmd
References for striae width

@@
 ***No data-based adjustment was made to the prior probabilities, i.e. prior probabilities of 50:50 for matches/non-matches were used.***
-
-***HH: what does the reviewer mean by cross-comparison? On page 15 we discuss some results from matching based on CMS.***
 
@@
 
-***EH: Add typical width of groove + lands***
-
-***EH: Presumably the typical width is some kind of class characteristic dependent on the gun itself, right?***
-
-***HH: I am not sure about the width of a striation - the width depends on where it is measured (in both x and z direction).
-I'm sure there are definitions of that available.
-We might need some more literature.***
-
-***EH: http://nvlpubs.nist.gov/nistpubs/jres/116/3/v116.n03.a01.pdf - "The number and width of striations in a land impression can vary depending on the size of the bullet and the characteristics of the barrel.
-For example, according to Biasotti’s study, the Smith and Wesson barrel with a land impression width of 2.4 mm has an average of 60 striations with total striation counts between 16 and 97 [13].
-Observing the land impression images used in our study, 50 pixels (about 0.08 mm) is a reasonable choice for the average width of a structuring element."***
+***We have added the typical width of both the groove-to-groove (land) width, as well as the typical width of a striation.***
 

### 2016-05-07 imaging-paper.Rnw
References for striae width

@@
 \end{figure}
+
+<<width, echo=FALSE>>=
+typical_width <- read.csv("csvs/grooves.csv") %>%
+    summarise(length = mean(groove_right - groove_left)) %>%
+    as.numeric %>%
+    round(digits = 2)
+@
+
 Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
+In the Hamby data, typical length (groove-to-groove) of a land is about \Sexpr{typical_width} micrometers or \Sexpr{typical_width / 1000} millimeters.
 For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
@@
 The direction of the rifling is a class characteristic, i.e.\ a feature that pertains to a particular class of firearms, and is not unique at the individual barrel/bullet level.
+
+The typical number and width of striation markings on bullets varies significantly depending on the gun barrel.
+For instance, a Smith and Wesson barrel with a land-width of 2.4 millimeters contained an average 60 striae, with an average width of about 0.08 millimeters~\citep{chu:2011}.
 

### 2016-05-07 imaging-paper.tex
added boxplots - densities do not work, because they are too skew
References for striae width

@@
 \end{figure}
+
 Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
+In the Hamby data, typical length (groove-to-groove) of a land is about 2005.9 micrometers or 2.01 millimeters.
 For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
@@
 
+The typical number and width of striation markings on bullet varies significantly depending on the gun barrel.
+For instance, a Smith and Wesson barrel with a land-width of 2.4 millimeters contained an average 60 striae, with an average width of about 0.08 millimeters~\citep{chu:2011}.
+
 The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. 
-The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
+The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
 % Hamby study: automated systems were successfully used for matching:
@@
 
+\includegraphics[width=.6\textwidth]{figure/tree:forest-2}
+
 \end{knitrout}

### 2016-05-08 imaging-paper.Rnw
Some further edits, typos, remove a paragraph

@@
 \begin{enumerate}
-\item {\bf Smooth the two signatures} by a loess with a very small span (see Figure~\ref{fig:smooth}). 
+\item {\bf Smooth the two signatures} using a loess with a very small span (see Figure~\ref{fig:smooth}). 
 \item Use cross-correlation to {\bf find the best alignment} of the two signatures: shift one of the signatures by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
@@
 Peaks and valleys constitute the \emph{striation marks} of the bullet.
-\item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify joint intervals between two signatures as those areas, in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals.
+\item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify joint intervals between two signatures as the areas in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals.
 A joint interval is then called a match(ing stria) between the signatures, if all of the intervals are of the same type of extrema, i.e.\ they are either all peaks or all valleys.
@@
 \item Number of matches and non-matches,
-\item \hh{The value of the cross-correlation function (ccf) between the aligned signatures \cite{vorburger:2011}, }
+\item The value of the cross-correlation function (ccf) between the aligned signatures \cite{vorburger:2011},
 \item Average difference $D$ between signatures, defined as the Euclidean vertical distance between surface measurements of aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
@@
 %In order to get a better understanding of  how  this matching works in known matches and non-matches, we next investigate the algorithm's performance in a test scenario.
-In addition to the features listed above, we also keep track of the value of the cross-correlation function found during the alignment, the lag at which signatures show the best alignment, and the height at which signatures are extracted ($x_1$ and $x_2$).
+%In addition to the features listed above, we also keep track of the value of the cross-correlation function found during the alignment, the lag at which signatures show the best alignment, and the height at which signatures are extracted ($x_1$ and $x_2$).
 

### 2016-05-10 revision/response.Rmd
Update response based on meeting

@@
 
+# Overall TODO
+
+1. Read the paper start to finish
+2. Look at boxplots vs densities (maybe use jittered boxplots?)
+
 # AE
@@
 
-***AC: TODO***
+***The AE makes two very good points.
+We have greatly extended our review of the existing literature and have included references to relevant papers.
+Where possible, we have also cocntrasted the proposed approach with existing methods.
+We have also added a discussion of the potential uses of a score such as ours in the forensic and legal applications.***
+
+***TODO: Add to paper***
 
@@
 
-***We have detailed a multi-stage analysis process that begins with a source image, processes the image, extracts features, and ultimately uses a random forest to compute predicted match probabilities.
-While the forest is a component of the process, we've taken care to emphasize that the preceeding steps outline a more general framework for analysis of striae in images that could theoretically apply to other forensic science questions as well.***
+***The basic objective is to find the most discriminating combination of attributes to construct a score that provides a quantitative summary, or signature, for each bullet.
+In that sense, yes, random forests are a standard alternative.
+Random forests are extraordinarily flexible tools, and permit the use of both continuous and discrete variables in the construction of the score.
+This is one reason to think that they are likely to be useful in many other forensic applications.
+We have added a comment about this in the manuscript.
+One aspect of our work that is (sadly) novel, at least relative to the forensics community, is that we are making our algorithm and code completely public, so that the scientific community can evaluate the approach in detail.***
+
+***TODO: Add to paper***
 
@@
 
-***HH: Could you look at this response i wrote?***
-
 ***We were not intending to provide new insights on random forests.
@@
 For this particular paper, however, the groove detection, profile extraction, and feature definitions were the intended focus.***
+
+***We selected random forests because the method has desirable attributes and can be easily ported to other forensic applications.
+The focus of the paper was not to find the best algorithm to develop a score but rather show the potential of combining as large a number of attributes as are available to discriminate among bullets.
+***
+
+***TODO: Combine into one***
 
@@
 
-***HH: I am not quite sure what a 'reasonable' classification method is supposed to mean.
-Shouldn't a classification method be judged by its success in predicting the outcome?
-Apart from that, random forests are based on decision trees, which are an extremely flexible tool for classification, that can deal with quantitative and categorical types of variables and, in the implementation we used, also deal with some missing values (which was not needed in the data here).***
+***A classification method should be judged by its success in predicting the outcome.
+Random forests are based on decision trees, which are an extremely flexible tool for classification, that can deal with quantitative and categorical types of variables and, in the implementation we used, also deal with some missing values (which was not needed in the data here).***
 
@@
 
-***HH: I'm sure I am misunderstanding the comment, but I don't know how to answer this question.
-The standard error of a rate in (0,1) would be sqrt(p*(1-p)).
-So because p is small the square root is going to be still pretty small.
-We are presenting out of bag error rates for the random forest - maybe we need to explain what out-of bag means in more detail?***
-
-***EH: Yeah, I was a little unsure of this. Maybe we can ask Alicia, and if she also is confused by it I can type up an explanation on how the OOB error is computed.***
-
 11. In the conclusions section the authors mention an assumption of "uniqueness". In general forensic science have moved beyond these questions- see the work of Saks and Kaye for details.
 
-***AC: TODO***
-
-***probably refers to articles such as "how good is good enough?", books "modern scientific evidence", "judging science" ....
-I am not sure that those are really necessary or helpful.
-It might be better to re-phrase the sentence in the conclusion.***
+***The reviewer is absolutely correct, but we are talking about a different ``uniqueness'' concept here.
+The discipline of firearm and toolmark examining relies on two foundational assumptions: that every gun leaves a unique set of markings and that the marks are reproducible.
+This is still the credo among tool and firearms examiners.
+The reviewer is referring to the fact that in testimony, forensic scientists are beginning to move away from declaring that, for example, an individual gun, to the exclusion of all others, was the one that fired the bullet.
+The work we present addresses the fundamental assumption of uniqueness, and will hopefully help move the practice away from statements of unique source.***
 
@@
 There might be a set of alternatives, and some evidence as to which methods might work better in some situations than others, but there is a lack of an in-depth discussion of the alternatives at each step as well as an overall discussion.
-In this manuscript we are detailing one route to divvy up the process and provide the code for others to reproduce our findings.
+In this manuscript we are detailing one route to divvy up the process and provide the code for others to reproduce our findings.***
+
+*** AC:  Actually I think that the meaning is different.
+I wonder whether ``closed system'' not in the sense of secret but rather in the sense of building a big database of signatures that can be searched (like AFIS, but open).
+This is what I interpret given the question about how forensic scientists might use this method.
 ***
+
+***TODO: Focus on open algorithms (opening up the closed system) / merge responses***
 
@@
 
-***TODO***
+***The point that the reviewer makes is valid.
+While we were familiar with the computer-assisted approaches that have been proposed, practice still relies on human examination of potential candidates presented by a system such as NIBIN.
+We have added what we hope is an acceptably complete review of the literature in the revised manuscript.***
+
+***TODO: Cite riva paper, explain work has been done on cartridges (not bullets)***
 
@@
 
-***At the time of the submission of this manuscript, the 35 bullets from Hamby set 252 were the only ones accessible to the public.
+***At the time of the submission of this manuscript, the 35 bullets from 10 barrels from Hamby set 252 were the only ones accessible to the public.
 They are part of Dr Zheng's efforts at NIST to provide a ballistics reference database.
@@
 
-***AC: TODO***
-
-***HH: why should the probability of two sets of observations given that the two bullets were fired by the same gun be the probability of interest?
-I would agree that the random forest and any of its conclusions are restricted to how good the underlying 'database' is.
-The more representative the set of bullets, the better the predictions should be.
-It is not the purpose of this paper to find the probability of two sets of observations given that the two bullets were fired by the same gun.
-We made sure to remove any misleading statements that might suggest the latter probability.
-***
+***The reviewer is correct in stating that two probabilities of interest are the probability of observing the data under the ``same source'' (same gun) hypothesis or under the ``different source'' hypothesis.
+An empirical approach to get at those two probabilities is to derive the distribution of scores when two bullets are known to have been fired by the same gun and the distribution of scores when two bullets are known to have been fired by different guns.
+We have actually derived those to distributions and include them in the revised manuscript.***
 
@@
 ***Assuming independence between the lands of each barrel is certainly not realistic in practice, however, by assuming independence we make the task of matching harder for the algorithm.
-XXX still need to do XXX Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel.
+Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel.
 The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.***
+
+***TODO: We will figure out***
 
@@
 
+*** AC:  The reviewer is correct.
+Here, we talk about ``match'' only as a means to test whether our algorithm is actually pairing the samples that are true pairs.
+This is a term used only in this development stage and is not intended for the court.
+****
+
+***TODO: Clarify that we are discussing "match" only in terms of assessing our algorithm. Merge AC's response***
+
 42. I strongly encourage the authors to not follow this trend and use the term "ballistics" only when you are actually talking about ballistics.

### 2016-05-20 revision/response.Rmd
Response update

@@
 We have greatly extended our review of the existing literature and have included references to relevant papers.
-Where possible, we have also cocntrasted the proposed approach with existing methods.
+Where possible, we have also contrasted the proposed approach with existing methods.
 We have also added a discussion of the potential uses of a score such as ours in the forensic and legal applications.***
@@
 
-***We were not intending to provide new insights on random forests.
-The insights were in the feature extraction and image processing leading up to the random forest fit.
-As the forest ultimately predicted every land-to-land comparison properly, we felt that a comparison to other learning methods wass outside the scope of this paper.
-Had a wider variety of data been available, there would likely be more variability in the predictions and a comparison to other learning methods would be prudent.
-For this particular paper, however, the groove detection, profile extraction, and feature definitions were the intended focus.***
-
 ***We selected random forests because the method has desirable attributes and can be easily ported to other forensic applications.
-The focus of the paper was not to find the best algorithm to develop a score but rather show the potential of combining as large a number of attributes as are available to discriminate among bullets.
-***
-
-***TODO: Combine into one***
+We were not intending to provide new insights on random forests.
+The focus of the paper was to show the potential of combining a large number ofimage attributes to discriminate among bullets.
+As the forest ultimately predicted every land-to-land comparison properly, we felt that a comparison to other learning methods was outside the scope of this paper.
+Had a wider variety of data been available, there would likely be more variability in the predictions and a comparison to other learning methods would be prudent.***
 
@@
 
-***We have clarified that matching two bullets is one of the goals, but not the sole or primary goal.***
-
-*** AC:  The reviewer is correct.
-Here, we talk about ``match'' only as a means to test whether our algorithm is actually pairing the samples that are true pairs.
+***The reviewer is correct.
+Here, we talk about "match" only as a means to test whether our algorithm is actually pairing the samples that are true pairs.
 This is a term used only in this development stage and is not intended for the court.
-****
-
-***TODO: Clarify that we are discussing "match" only in terms of assessing our algorithm. Merge AC's response***
+As such, we have clarified that matching two bullets is one of our goals, but not the sole or primary goal of this procedure in forensic science.***
 

### 2016-05-20 imaging-paper.tex
Response update

@@
 
-The typical number and width of striation markings on bullet varies significantly depending on the gun barrel.
+The typical number and width of striation markings on bullets varies significantly depending on the gun barrel.
 For instance, a Smith and Wesson barrel with a land-width of 2.4 millimeters contained an average 60 striae, with an average width of about 0.08 millimeters~\citep{chu:2011}.

### 2016-05-20 imaging-paper.Rnw
removed bootstrap

@@
 The fit seems to do a reasonable job of capturing the structure of the image.
-The blue area around the signature are  95\% bootstrapped confidence intervals based on 1,000 bootstrap samples.
+%The blue area around the signature are  95\% bootstrapped confidence intervals based on 1,000 bootstrap samples.
 Figure~\ref{fig:loess_resid} shows the residuals from this fit. These residuals are called the \emph{signature} of bullet~1-5.
@@
 <<boot_loess, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
-boot.loess <- boot_fit_loess(br111_bullet, br111.groove)
-poly <- with(boot.loess, data.frame(x = c(y,rev(y)), y=c(high, rev(low))))
-qplot(x=x, y=y, geom="polygon", fill=I("steelblue"), data=poly) +
+#boot.loess <- boot_fit_loess(br111_bullet, br111.groove)
+#poly <- with(boot.loess, data.frame(x = c(y,rev(y)), y=c(high, rev(low))))
+#qplot(x=x, y=y, geom="polygon", fill=I("steelblue"), data=poly) +
+ggplot() +
   geom_line(aes(x=y, y=resid), data=my.loess$data) +
@@
 The residuals define the {\it signature} of bullet~1-5.
-Blue areas around the residuals show  95\% bootstrap confidence intervals (based on $B=1,000$ bootstrap samples).
+%Blue areas around the residuals show  95\% bootstrap confidence intervals (based on $B=1,000$ bootstrap samples).
 }

### 2016-05-20 imaging-paper.Rnw
just jittered plot for matches

@@
 predm$match <- c("KNM", "KM")[as.numeric(predm$match)+1]
+predm$barrel1 <- gsub("(.*) .*-[0-9]+", "\\1", predm$b1)
+predm$barrel2 <- gsub("(.*)-[0-9]+", "\\1", predm$b2)
+
 qplot(data=predm, value, match, geom="jitter", colour=match, shape=match) + theme_bw() + facet_grid(variable~.) +
@@
 
-qplot(data=predm, value, geom="boxplot", fill=match, x=match, alpha=I(0.5)) + 
-  theme_bw() + facet_grid(variable~.) +
-  scale_fill_brewer(palette="Paired", labels=c("KNM", "KM")) +
-   ylab("") + 
-   xlab("Predicted probability of a match") +
-  coord_flip() + 
-  theme(legend.position="none")
-
-# qplot(pred, forest, geom="jitter", data=bstats, colour=match, 
-#       shape=match) + theme_bw() +
-#   theme(legend.position="bottom") +
-#   geom_vline(xintercept=0.5, colour="grey50") +
-#   geom_hline(yintercept=0.5, colour="grey50") + 
-#   xlab("Predictions from decision tree") +
-#   ylab("Predictions from random forest")   +
-#   scale_colour_brewer(palette="Paired", labels=c("KNM", "KM")) +
-#   scale_shape_discrete(labels=c("KNM", "KM"))
-# 
 @

### 2016-05-23 revision/response.Rmd
Notes from mtg

@@
 2. Look at boxplots vs densities (maybe use jittered boxplots?)
+3. Talk to Alen, send paper, ask to add to intro? Add as author?
 
@@
 
-***HH: will do***
+***We have decided to remove the bootstrap confidence intervals for the loess fit residuals because they did not add enough information.***
 
@@
 
-***HH: in my view we ourselves are not building a closed set identification system.
-The problem with existing closed set systems is that it is unclear how the matching is exactly done and what statistical properties it has (number of parameters, distribution of them, sensitivities).
+***The problem with existing closed set systems is that it is unclear how the matching is exactly done and what statistical properties it has (number of parameters, distribution of them, sensitivities).
 Repeatability and reproducibility is a huge issue in this area.
 'Matching' is also not just a single step - it consists, as we have laid out in the paper, of a series of steps, which all deserve attention.
-While the reviewer pointed out to us to use 'standard methods' - there are no actual standards in place.
+While it was pointed out to us to use 'standard methods' - there are no actual standards in place.
+The status quo is NIBIN, which is not accessible to anyone outside of law enforcement.
 There might be a set of alternatives, and some evidence as to which methods might work better in some situations than others, but there is a lack of an in-depth discussion of the alternatives at each step as well as an overall discussion.
@@
 
-*** AC:  Actually I think that the meaning is different.
-I wonder whether ``closed system'' not in the sense of secret but rather in the sense of building a big database of signatures that can be searched (like AFIS, but open).
-This is what I interpret given the question about how forensic scientists might use this method.
-***
-
-***TODO: Focus on open algorithms (opening up the closed system) / merge responses***
+***TODO: Add some of this discussion to the intro + conclusion***
 
@@
 
-***The reviewer is correct in stating that two probabilities of interest are the probability of observing the data under the ``same source'' (same gun) hypothesis or under the ``different source'' hypothesis.
+***The reviewer is correct in stating that two probabilities of interest are the probability of observing the data under the "same source" (same gun) hypothesis or under the "different source" hypothesis.
 An empirical approach to get at those two probabilities is to derive the distribution of scores when two bullets are known to have been fired by the same gun and the distribution of scores when two bullets are known to have been fired by different guns.
@@
 ***Assuming independence between the lands of each barrel is certainly not realistic in practice, however, by assuming independence we make the task of matching harder for the algorithm.
+This also has the benefit of potentially being more useful for degraded data, where we may not have a full set of six groove-to-groove scans.
 Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel.
@@
 
-***TODO: We will figure out***
+***TODO: We will figure out whether there is clustering associated with the barrel. Add discussion of why we are comparing land to land rather than bullet to bullet***
 

### 2016-05-26 revision/response.Rmd
Continue working on many of the changes

@@
 
-***TODO: Add some of this discussion to the intro + conclusion***
-
 # Reviewer C
@@
 
-***TODO: We will figure out whether there is clustering associated with the barrel. Add discussion of why we are comparing land to land rather than bullet to bullet***
+***TODO: We will figure out whether there is clustering associated with the barrel.***
 

### 2016-05-26 imaging-paper.Rnw
Continue working on many of the changes

@@
 The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match.
+\eh{We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet.
+Although doing so introduces an implicit assumption of independence between lands, assuming independence only serves to make the task more challenging.
+XXX need more?}
 The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
@@
 We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features.
-%Finally, we have provided an easy-to-use open-source backend, and web frontend to many of these features.
+\eh{Matching bullets is clearly not a single step, but rather a composition of a number of data analysis tasks which each deserve attention.
+With no scientific standards in place, our intent is to explain one approach to addressing these tasks, while documenting all steps and providing all code so other researchers and forensic scientists can reproduce and expand on our findings.}
 

### 2016-05-26 imaging-paper.tex
recompile
Continue working on many of the changes

@@
 
-\includegraphics[width=.6\textwidth]{figure/tree:forest-2}
-
 \end{knitrout}

### 2016-07-11 imaging-paper.Rnw
Some small typos

@@
 \caption{\label{fig:varysmooth} Peak/valley detection at smoothing factors of 5, 25, and 45, respectively.
-Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either end disappear or shift  horizontally  from their position in the signature.}
+Note that a smoothing factor of 5 yields enough noise that many very minimal overlapping peaks and valleys are detected, while a smoothing factor of 45 might over-smooth and cause the peaks/valleys to either end disappear or shift  horizontally from their original position in the signature.}
 \end{figure}
@@
 This is one of the benefits of switching to a digitized version of the images that preserves the 3D surface structure.
-The findings about the discriminating power of cross-correlation is consistent with the results of the study by \citet{ma:2004}.
-However, in that study, authors did not consider the number of matches and non-matches.
+The findings about the discriminating power of cross-correlation are consistent with the results of the study by \citet{ma:2004}.
+However, in that study, the authors did not consider the number of matches and non-matches.
 
@@
 In practice, examiners are allowed a third option of `inconclusive'.
-On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be not necessary in the Hamby study, because, here, the results from the random forest are very clear cut.
+On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be unnecessary in the Hamby study, because, here, the results from the random forest are very clear cut.
 Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree and the random forest.

### 2016-07-11 code/test_getpeaks.R

@@ -3,11 +3,11 @@ library(x3prplus)
 library(ggplot2)
 
 for (file in dir("app/images/Hamby252_3DX3P1of2")) {
-    mybullet <- get_bullet(file.path("app/images/Hamby252_3DX3P1of2", file), x = 150)
+    mybullet <- get_crosscut(file.path("app/images/Hamby252_3DX3P1of2", file), x = 150)
     x <- get_grooves(mybullet)
     cat(file, "\n")
     my.loess <- fit_loess(mybullet, x)
-    result <- get_peaks(my.loess$data)
+    result <- get_peaks(my.loess$data, smoothfactor = 15)
     print(result$plot)
     value <- readline("Press Enter to Continue, or q to Quit")
     if (value == "q") break;

### 2016-07-11 code/filtering.Rmd
Include some new filtering test code

---
title: "Testing Filtering Parameters"
output: html_notebook
---

```{r, message=FALSE}
library(x3pr)
library(x3prplus)
library(ggplot2)
library(smoother)
library(changepoint)

mybullet <- read.x3pplus("~/GitHub/imaging-paper/app/images/Hamby252_3DX3P1of2/Br1 Bullet 1-1.x3p")
```

```{r}
mycc <- get_crosscut(x = 100, bullet = mybullet)
my.groove <- get_grooves(mycc)
my.loess <- fit_loess(mycc, my.groove)
qplot(y, resid, data = my.loess$data, geom = "line") + theme_bw()
```

```{r}
my.loess$data$smoothed <- smth.gaussian(my.loess$data$resid, window = 16)
my.inds <- which(diff(sign(diff(my.loess$data$smoothed))) == -2) + 1
my.inds2 <- which(diff(sign(diff(my.loess$data$smoothed))) == 2) + 1

length(my.inds) + length(my.inds2)

qplot(y, smoothed, data = my.loess$data, geom = "line") + theme_bw() +
    geom_vline(xintercept = my.loess$data$y[my.inds], colour = "red") +
    geom_vline(xintercept = my.loess$data$y[my.inds2], colour = "blue")
```

### 2016-07-12 code/filtering.Rmd
Try the NIST double pass smoothing

@@
 ```{r}
-my.loess$data$smoothed <- smth.gaussian(my.loess$data$resid, window = 16)
+my.loess$data$smoothed_first <- smth.gaussian(my.loess$data$resid, window = 160)
+my.loess$data$smoothed <- smth.gaussian(my.loess$data$smoothed_first, window = 16)
 my.inds <- which(diff(sign(diff(my.loess$data$smoothed))) == -2) + 1

### 2016-07-12 code/filtering.Rmd
Maybe switch it around?

--- 2016-07-11 code/filtering.Rmd
@@
 ```{r}
-my.loess$data$smoothed <- smth.gaussian(my.loess$data$resid, window = 16)
+my.loess$data$smoothed_first <- my.loess$data$resid - smth.gaussian(my.loess$data$resid, window = 16)
+my.loess$data$smoothed <- my.loess$data$smoothed_first - smth.gaussian(my.loess$data$smoothed_first, window = 160)
 my.inds <- which(diff(sign(diff(my.loess$data$smoothed))) == -2) + 1

### 2016-07-13 code/filtering.Rmd
The new hamby code

--- 2016-07-11 code/filtering.Rmd
@@
 ```{r}
-my.loess$data$smoothed <- smth.gaussian(my.loess$data$resid, window = 16)
+my.loess$data$smoothed_first <- smth.gaussian(my.loess$data$resid, window = 16)
+my.loess$data$smoothed <- my.loess$data$smoothed_first - smth.gaussian(my.loess$data$resid, window = 160)
 my.inds <- which(diff(sign(diff(my.loess$data$smoothed))) == -2) + 1

### 2016-07-13 code/complete-hamby-nist.R

@@ -1,0 +1,59 @@
+library(x3pr)
+library(x3prplus)
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(zoo)
+
+knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
+unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+
+# match unknown land using crosscuts
+ccs <- read.csv("csvs/crosscuts-25.csv")
+all_bullets <- lapply(as.character(ccs$path), function(x) {
+    transpose <- (length(grep(" ", x)) == 0)
+    
+    result <- read.x3pplus(x, transpose = transpose)
+    result[[3]] <- x
+    names(result)[3] <- "path"
+        
+    return(result)
+})
+
+grooves <- read.csv("csvs/grooves.csv")
+
+knowns <- all_bullets[1:240][grep(" ", as.character(ccs$path)[1:240])]
+unknowns <- all_bullets[241:420][grep(" ", as.character(ccs$path)[241:420])]
+bullets_processed <- lapply(c(knowns, unknowns), function(bul) {
+    cat("Computing processed bullet", basename(bul$path), "\n")
+    
+    xval <- ccs$cc[which(ccs$path == bul$path)]
+    
+    grooves_sub <- filter(grooves, bullet == bul$path)
+    
+    left <- grooves_sub$groove_left[which.min(abs(xval - grooves_sub$x))]
+    right <- grooves_sub$groove_right[which.min(abs(xval - grooves_sub$x))]
+    
+    processBullets(bullet = bul, name = bul$path, x = grooves_sub$x[which.min(abs(xval - grooves_sub$x))], grooves = c(left, right))
+})
+names(bullets_processed) <- as.character(ccs$path)[grep(" ", as.character(ccs$path))]
+
+#bullets_smoothed <- bullets_processed %>% bind_rows %>% bulletSmooth
+bullets_smoothed <- bullets_processed %>% bind_rows
+
+for (span in c(1)) {
+    dataStr <- sprintf("data-%d-25", span) # using crosscuts-25.csv
+    
+    if (!file.exists(dataStr)) dir.create(dataStr)
+    for (j in 1:length(unknowns)) {
+        reslist <- lapply(knowns, function(x) {
+            cat("Processing", j, "vs", basename(x$path), "with span", span, "\n")
+            
+            br1 <- filter(bullets_smoothed, bullet == x$path)
[…]

### 2016-07-19 revision/response.Rmd
Updates based on mtg

@@
 ---
-
-# Overall TODO
-
-1. Read the paper start to finish
-2. Look at boxplots vs densities (maybe use jittered boxplots?)
-3. Talk to Alen, send paper, ask to add to intro? Add as author?
 
@@
 It would be interesting to compare this method with some of the classical methods from functional data analysis used for curve registration.
+
+***TODO: Check which references are and are not in the paper. Add Chu 2010, Chu 2009, Bachrach 2002, Vorburger 2011, Riva 2014. Cite Bachrach 2006 in response to reviewer about barrel effects***
 
@@
 
-***TODO: Cite riva paper, explain work has been done on cartridges (not bullets)***
+***TODO: Cite riva paper, explain work has been done on cartridges (not bullets). Cite John Song's papers***
 
@@
 The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.***
-
-***TODO: We will figure out whether there is clustering associated with the barrel.***
 

### 2016-07-25 imaging-paper.Rnw
Reproducible research comments

@@
 The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match.
-\eh{We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet.
+Our algorithm is fully open source and available on GitHub \citep{x3prplus}.
+This transparency allows for a greater understanding of the individual steps involved in the bullet matching process, and allows other forensic examiners, as well as outside observers, examine the factors that separate known bullet matches from non-matches.
+We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet.
 Although doing so introduces an implicit assumption of independence between lands, assuming independence only serves to make the task more challenging.
-XXX need more?}
 The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.

### 2016-07-25 revision/response.Rmd
Reproducible research comments
More responses

@@
 author: "Heike Hofmann, Eric Hare, Alicia Carriquiry"
-date: "April 14, 2016"
+date: "July 26, 2016"
 output: html_document
@@
 
-***TODO: Add to paper***
-
 2.
@@
 One aspect of our work that is (sadly) novel, at least relative to the forensics community, is that we are making our algorithm and code completely public, so that the scientific community can evaluate the approach in detail.***
-
-***TODO: Add to paper***
 
@@
 
-***TODO: Check which references are and are not in the paper. Add Chu 2010, Chu 2009, Bachrach 2002, Vorburger 2011, Riva 2014. Cite Bachrach 2006 in response to reviewer about barrel effects***
+***Using the cross correlation function as a method of alignment is actually pretty standard - the appropriate references are now included.***
 
-***Using the cross correlation function as a method of alignment is actually pretty standard - the approriate references are now included.***
-
-***Additional papers considered (not complete, and not yet included in the paper, but I think all of them should be cited): chumbley:2010 uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening) chubley:2010 is a much more detailed version over David Faden's thesis***
+***chumbley:2010 uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening) chubley:2010 is a much more detailed version over David Faden's thesis***
 
@@
 
-***TODO: Cite riva paper, explain work has been done on cartridges (not bullets). Cite John Song's papers***
-
 14.
@@
 This also has the benefit of potentially being more useful for degraded data, where we may not have a full set of six groove-to-groove scans.
-Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel.
+Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel (Bachrach 2006).
 The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.***
@@
 
-***HH: Remove blue area / references to it... maybe?***
+***DONE***
 

### 2016-07-25 imaging-paper.Rnw
More responses

@@
 The authors state that this process `could be automated', but no implementation of the algorithm is available.
+Modern automated techniques using 3D images have also been conducted by \citet{riva:2014}, but this work applies to striae found on cartridge cases and not bullets.
 
@@
 %
-\hh{For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, vorburger:2011, thompson:2013}.}
+\hh{For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, chu:2010, vorburger:2011, thompson:2013}.}
 This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.

### 2016-07-26 revision/response.Rmd
Commit new response

@@
 title: "Response to reviewers"
-author: "Heike Hofmann, Eric Hare, Alicia Carriquiry"
+author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
 date: "July 26, 2016"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 
-***Using the cross correlation function as a method of alignment is actually pretty standard - the appropriate references are now included.***
-
-***chumbley:2010 uses 1d profiles and discusses several different angles of extraction (no alignment necessary, no grooves and no flattening) chubley:2010 is a much more detailed version over David Faden's thesis***
-
-***chu:2010 describe extracting mean profile of an area of the image after untwisting the image ***
-
-***bachrach:2002 general description of an automated matching system based on 3D imaging. Uses cross-correlation function for alignment.***
-
-***vorburger:2011 different uses of the cross-correlation function. One of them aligning profiles***
-
-***bachrach:2006 67page final report for NIJ grant (for Intelligent Automation Inc.) wear study: "difference between comparing the nth and (n+1)th bullet fired by a barrel as opposed to the nth and (n+100)th bullet fired by a barrel?
-In other words, does the proximity between firings make a difference?
-Although some wear effect could be detected in both of these scenarios, these effects do not appear to be significant enough to prevent correct identification"***
-
-***petraco:2012 I am not sure that this book is the same as the report for the NIJ grant.
-This is a huge effort, and we should certainly cite it, but there are not a lot of actual results included or comparisons done.***
-
-***riva:2014 automatic matching system based on 3d images of cartridge cases. Matching involves breech faces, but no striation marks.***
-
-***xie:2009 first sentence: "Every firearm has individual characteristics that are as unique to it as fingerprints are to human beings." the paper itself doesn't back up any of these claims.
-very disappointing***
-
-***leon:2006 first sentence: "Fired bullets bear striation marks that can be thought of as a ‘‘fingerprint’’ left by the firearm." based on multiple 2d images. They extract a 1d 'signature'***
+    ***Thank you for the comments.
+Using the cross correlation function as a method of alignment is actually pretty standard - the appropriate references are now included.
+We have attempted to better clarify the novel component of this work, by explaining openly and transparently describing each individual step of the analysis process, and providing the source code used to reproduce the findings.***
 

### 2016-07-26 revision/response.Rmd
Paper updates

@@
   pdf_document: default
-  html_document: default
+bibliography: references.bib
 ---
@@
 Using the cross correlation function as a method of alignment is actually pretty standard - the appropriate references are now included.
-We have attempted to better clarify the novel component of this work, by explaining openly and transparently describing each individual step of the analysis process, and providing the source code used to reproduce the findings.***
+We have attempted to better clarify the novel component of this work, by describing openly and transparently each individual step of the analysis process, and providing the source code used to reproduce the findings.***
 
@@
 They are part of Dr Zheng's efforts at NIST to provide a ballistics reference database.
-In the future more results will be available, but until then we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.***
+In the future more results will be available, but until then we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.
+With that said, it has been shown that with repeated firings, minor degradation does occur, but that such minor degradation likely wouldn't prevent correct matching [@bachrach:2006].
+It would be very interesting to study this further, but this unfortunately is outside the scope of this paper.***
 
@@
 An empirical approach to get at those two probabilities is to derive the distribution of scores when two bullets are known to have been fired by the same gun and the distribution of scores when two bullets are known to have been fired by different guns.
-We have actually derived those to distributions and include them in the revised manuscript.***
+We have actually derived those two distributions and included them in the revised manuscript.***
 
@@
 This also has the benefit of potentially being more useful for degraded data, where we may not have a full set of six groove-to-groove scans.
-Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel (Bachrach 2006).
+Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel.
 The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.***
@@
     ***DONE***
+
+# References

### 2016-07-26 imaging-paper.Rnw
Paper updates

@@
 \caption{\label{fig:manualmatch-rgl}3D view of the manually adjusted side-by-side comparison of bullet~1-5 and bullet 2-1 after removing the curvature.
-\hh{Bullet 2-1 is shaded light grey in the background.}}
+Bullet 2-1 is shaded light grey in the background.}
 \end{figure}
@@
 %
-\hh{For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, chu:2010, vorburger:2011, thompson:2013}.}
+For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, chu:2010, vorburger:2011, thompson:2013}.
 This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
@@
 Variables are sorted according to their area under the curve (AUC).
-\hh{The equal error rate (EER) is marked by a point on the ROC curve.} Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS.
+The equal error rate (EER) is marked by a point on the ROC curve.
+Except for the distance $D$ between signatures, all individual features derived from the surface measurements and the aligned striation marks are more predictive than the maximal CMS.
 }
@@
 The features are arranged in descending order according to the area under the curve (AUC).
-\hh{XXX The dots mark the equal error rate, i.e.
+The dots mark the equal error rate, i.e.
 the location on the ROC curve, where false positive and false negative error rates are the same.
@@
 We see that in this instance a low EER goes hand in hand with high predictive power as measured in AUC.
-XXX HH: do we need to say more about EERs?}
 The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys.
@@
 We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features.
-\eh{Matching bullets is clearly not a single step, but rather a composition of a number of data analysis tasks which each deserve attention.
-With no scientific standards in place, our intent is to explain one approach to addressing these tasks, while documenting all steps and providing all code so other researchers and forensic scientists can reproduce and expand on our findings.}
+Matching bullets is clearly not a single step, but rather a sequence of a number of data analysis tasks which each deserve attention.
+As there is no scientific standard in place at this point in time, our intent is to explain one approach to addressing these tasks, while documenting all steps and providing all code so other researchers and forensic scientists can reproduce and expand on our findings.
 
@@
 In particular, `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.
-\eh{The steps outlined in this paper could also help explore other important forensic science problems.
-In particular, more general toolmark examination can benefit from this detection and removal of the overall structure and the extraction of features from deviations of this structure.}
+The steps outlined in this paper could also help explore other important forensic science problems.
+In particular, more general toolmark examination can benefit from this detection and removal of the overall structure and the extraction of features from deviations of this structure.
 

### 2016-08-08 revision/response.Rmd
New Arxiv submission

@@
 author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
-date: "July 26, 2016"
+date: "August 8, 2016"
 output:

### 2016-10-07 imaging-paper.Rnw
New data directory and updated paper compilation

@@
 codedir <- "code"
-datadir <- "app/images/Hamby252_3DX3P1of2/"
+datadir <- "images/Hamby (2009) Barrel/bullets"
 
 options(replace.assign=TRUE,scipen=3, digits=2)
-bstats <- read.csv("data-25-25/bullet-stats-old.csv")
+bstats <- read.csv("data/data-25-25/bullet-stats-old.csv")
 
@@
 
-%' \hh{XXX Do we need the next example? We could save some space here XXX}
-%' Figure~\ref{fig:loessfits} shows the signatures of all lands of bullet 2 (in grey) fired from the same barrel as bullet~1.
-The darker line overlaid on each of the signatures, shows signature from bullet~1-5.
-%' %We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet~1-5.
-This is shown in Figure~\ref{fig:loessfits}.
-%' Bullet 2-1 is the best match for bullet~1-5, but the alignment is not quite  correct. \hh{This is one of the issues we are addressing next.}
-%' 
-%' \begin{figure}[hbtp]
-%'   \centering
-%' <<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
-%' br121.groove <- get_grooves(br121_bullet)
-%' br122.groove <- get_grooves(br122_bullet)
-%' br123.groove <- get_grooves(br123_bullet)
-%' br124.groove <- get_grooves(br124_bullet)
-%' br125.groove <- get_grooves(br125_bullet)
-%' br126.groove <- get_grooves(br126_bullet)
-%' 
-%' my.loess1 <- fit_loess(br121_bullet, br121.groove)
-%' my.loess2 <- fit_loess(br122_bullet, br122.groove)
-%' my.loess3 <- fit_loess(br123_bullet, br123.groove)
-%' my.loess4 <- fit_loess(br124_bullet, br124.groove)
-%' my.loess5 <- fit_loess(br125_bullet, br125.groove)
-%' my.loess6 <- fit_loess(br126_bullet, br126.groove)
-%' 
-%' my.loessactual <- fit_loess(br111_bullet, br111.groove)
-%' 
-%' minrows <- min(length(my.loess1$data$resid),
-%'                length(my.loess2$data$resid),
-%'                length(my.loess3$data$resid),
-%'                length(my.loess4$data$resid),
-%'                length(my.loess5$data$resid),
-%'                length(my.loess6$data$resid),
-%'                length(my.loessactual$data$resid))
-%' 
-%' my.df <- data.frame(y = my.loess1$data$y[1:minrows], resid1 = my.loess1$data$resid[1:minrows],
-%'                     resid2 = my.loess2$data$resid[1:minrows], resid3 = my.loess3$data$resid[1:minrows],
-%'                     resid4 = my.loess4$data$resid[1:minrows], resid5 = my.loess5$data$resid[1:minrows],
-%'                     resid6 = my.loess6$data$resid[1:minrows])
-%' 
-%' melt.df <- my.df %>%
-%'     gather(key = part, value = resid, resid1:resid6)
-%' melt.df$part <- gsub("resid", "", melt.df$part)
-%' 
-%' actual.df <- my.loessactual$data[1:minrows,]
-%' melt.df$land <- melt.df$part
-%' qplot(y, resid, data=melt.df, #colour=factor(x), 
-%'       geom="line", size=I(.75), colour=I("grey70")) + 
-%'   facet_wrap(~land, ncol=3, labeller="label_both") + 
-%'   scale_colour_brewer(palette="Paired") +
-%'   theme_bw() + theme(legend.position="bottom") + 
-%'   geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Signatures of bullet 1 in grey, bullet 1-5 in black") + 
-%'   ylab("Residuals of loess fit") +
-%'   theme(plot.margin = unit(c(0,0,0,0), unit="cm"))
-%' @    
-%' \caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet~1-5 is shown in black. }
-%' \end{figure}
-%'
-
 \section{Automatic matching}
@@
 paths <- file.path(datadir, dir(datadir))
-paths <- paths[grep(" ", paths)]
+paths <- paths[grep("Br[0-9]", paths)]
 
@@
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
-load("data-25-25/unkn9.RData")
+load("data/data-25-25/unkn9.RData")
 cmsdist <- sapply(reslist, function(x) x$maxCMS)
@@
 <<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.65\\textwidth', warning=FALSE>>=
-load("data-25-25/unkn47.RData")
+load("data/data-25-25/unkn47.RData")
 res <- reslist[[106]]

### 2016-10-07 imaging-supplement.Rnw
Finally fix up all the paths

@@
 library(sm)
-
+library(reshape2)
 @
@@
 
-knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
-unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+knowndatadir <- "images/Hamby (2009) Barrel/bullets"
+unknowndatadir <- "images/Hamby (2009) Barrel/bullets"
 
@@
 ccs <- read.csv("csvs/crosscuts-25-old.csv")
+ccs$path <- file.path(knowndatadir, basename(as.character(ccs$path)))
 all_bullets <- lapply(as.character(ccs$path), function(x) {
@@
 <<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=
-knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
+knowndatadir <- "images/Hamby (2009) Barrel/bullets"
 knowns <- dir(path=knowndatadir)
-knowns <- knowns[grep(" ", knowns)]
+knowns <- knowns[grep("Br[0-9]", knowns)]
 knowns <- gsub(".x3p", "", knowns)
 
-unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
+unknowndatadir <- "images/Hamby (2009) Barrel/bullets"
 unknowns <- dir(path=unknowndatadir)
-unknowns <- unknowns[grep(" ", unknowns)]
+unknowns <- unknowns[grep("Ukn", unknowns)]
 unknowns <- gsub(".x3p", "", unknowns)
 
-#bstats <- read.csv("data-25-25/bullet-stats.csv")
+#bstats <- read.csv("data/data-25-25/bullet-stats.csv")
 #flagged <- c("Br6 Bullet 2-1", "Br9 Bullet 2-4", "Ukn Bullet B-2", "Ukn Bullet Q-4")
-ballstats <- read.csv("data-new-all-25-25/bullet-stats-single.csv")
+ballstats <- read.csv("data/data-new-all-25-25/bullet-stats-single.csv")
 
@@
 matches$id <- 1:nrow(matches)
-library(reshape2)
 mm <- melt(matches, id.var=c("barrel", "id"), na.rm=TRUE)
@@
 mm$prefix[unkns] <- "Ukn Bullet "
-mm$path <- "app/images/Hamby252_3DX3P1of2//"
-mm$path[unkns] <- "app/images/Hamby252_3DX3P2of2//"
+mm$path <- "images/Hamby (2009) Barrel/bullets/"
 mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))
@@
 ccs <- read.csv("csvs/crosscuts-25-old.csv")
+ccs$path <- file.path(knowndatadir, basename(as.character(ccs$path)))
 mm <- merge(mm, ccs, by="path", all=TRUE)
@@
 <<setup-rtrees, echo=FALSE, message=FALSE>>=
-bstats <- read.csv("data-25-25/bullet-stats-old.csv", stringsAsFactors = FALSE)
+bstats <- read.csv("data/data-25-25/bullet-stats-old.csv", stringsAsFactors = FALSE)
 

### 2016-10-07 imaging-paper.Rnw
Include reference to x3pr

@@
 It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in forensic science.
+An open-source package for working with x3p files has been written in R \citep{R} by Dr.
+Nicholas Petraco which we build upon in the development of our algorithm \citep{x3pr}.
+This package enables the direct reading and manipulation of x3p files.
 
@@
 Modern automated techniques using 3D images have also been conducted by \citet{riva:2014}, but this work applies to striae found on cartridge cases and not bullets.
+Other work on 3D images has been conducted by \citet{petraco:2012}, which also describes work carried out on cartridge cases, as well as screwdriver striation patterns, with promising results.
+We aim to use similar ideas of automated 3D image analysis, and apply this work to bullet surfaces with the modern tools enabled by the R statistical software.
 

### 2016-10-07 imaging-paper.tex
New data directory and updated paper compilation
Include reference to x3pr

@@
 \newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
 
@@
 %
-
-%' \hh{XXX Do we need the next example? We could save some space here XXX}
-%' Figure~\ref{fig:loessfits} shows the signatures of all lands of bullet 2 (in grey) fired from the same barrel as bullet~1.
-The darker line overlaid on each of the signatures, shows signature from bullet~1-5.
-%' %We can reproduce a similar figure to the one from the circular fits by overlaying the residuals from each loess fit for bullet 2 onto an image of the loess residuals from bullet~1-5.
-This is shown in Figure~\ref{fig:loessfits}.
-%' Bullet 2-1 is the best match for bullet~1-5, but the alignment is not quite  correct. \hh{This is one of the issues we are addressing next.}
-%' 
-%' \begin{figure}[hbtp]
-%'   \centering
-%' <<loessfits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
-%' br121.groove <- get_grooves(br121_bullet)
-%' br122.groove <- get_grooves(br122_bullet)
-%' br123.groove <- get_grooves(br123_bullet)
-%' br124.groove <- get_grooves(br124_bullet)
-%' br125.groove <- get_grooves(br125_bullet)
-%' br126.groove <- get_grooves(br126_bullet)
-%' 
-%' my.loess1 <- fit_loess(br121_bullet, br121.groove)
-%' my.loess2 <- fit_loess(br122_bullet, br122.groove)
-%' my.loess3 <- fit_loess(br123_bullet, br123.groove)
-%' my.loess4 <- fit_loess(br124_bullet, br124.groove)
-%' my.loess5 <- fit_loess(br125_bullet, br125.groove)
-%' my.loess6 <- fit_loess(br126_bullet, br126.groove)
-%' 
-%' my.loessactual <- fit_loess(br111_bullet, br111.groove)
-%' 
-%' minrows <- min(length(my.loess1$data$resid),
-%'                length(my.loess2$data$resid),
-%'                length(my.loess3$data$resid),
-%'                length(my.loess4$data$resid),
-%'                length(my.loess5$data$resid),
-%'                length(my.loess6$data$resid),
-%'                length(my.loessactual$data$resid))
-%' 
-%' my.df <- data.frame(y = my.loess1$data$y[1:minrows], resid1 = my.loess1$data$resid[1:minrows],
-%'                     resid2 = my.loess2$data$resid[1:minrows], resid3 = my.loess3$data$resid[1:minrows],
-%'                     resid4 = my.loess4$data$resid[1:minrows], resid5 = my.loess5$data$resid[1:minrows],
-%'                     resid6 = my.loess6$data$resid[1:minrows])
-%' 
-%' melt.df <- my.df %>%
-%'     gather(key = part, value = resid, resid1:resid6)
-%' melt.df$part <- gsub("resid", "", melt.df$part)
-%' 
-%' actual.df <- my.loessactual$data[1:minrows,]
-%' melt.df$land <- melt.df$part
-%' qplot(y, resid, data=melt.df, #colour=factor(x), 
-%'       geom="line", size=I(.75), colour=I("grey70")) + 
-%'   facet_wrap(~land, ncol=3, labeller="label_both") + 
-%'   scale_colour_brewer(palette="Paired") +
-%'   theme_bw() + theme(legend.position="bottom") + 
-%'   geom_line(data = actual.df, aes(y, resid), colour="black", size=.75, alpha=0.5) + ggtitle("Signatures of bullet 1 in grey, bullet 1-5 in black") + 
-%'   ylab("Residuals of loess fit") +
-%'   theme(plot.margin = unit(c(0,0,0,0), unit="cm"))
-%' @    
-%' \caption{\label{fig:loessfits} Residuals of a loess fit to bullets 1 and 2 from barrel 1. Bullet~1-5 is shown in black. }
-%' \end{figure}
-%'
 

### 2016-10-07 arxiv/imaging-paper.tex
New Arxiv submission
Remove line numbers for Arxiv
Quick Arxiv update

--- 2016-07-26 imaging-paper.tex
@@
 \caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
-      \includegraphics[width=.65\textwidth]{images/sidex3p.png}
+      \includegraphics[width=.65\textwidth]{sidex3p.png}
       \hspace{1cm}
-      \includegraphics[width=.25\textwidth]{images/side-sketch.png}
+      \includegraphics[width=.25\textwidth]{side-sketch.png}
     }
@@
     \caption{Frontal view of a bullet land (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
-    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
-    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
+    \includegraphics[width=.65\textwidth]{topx3p.png} \hspace{1cm}
+    \includegraphics[width=.15\textwidth]{top-sketch.png}
     }
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=0.65\textwidth]{figure/fixedX-1}
+\includegraphics[width=0.65\textwidth]{fixedX-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.5\textwidth]{figure/loess_step1-1}
+\includegraphics[width=.5\textwidth]{loess_step1-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.5\textwidth]{figure/loess_step3-1}
+\includegraphics[width=.5\textwidth]{loess_step3-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.5\textwidth]{figure/loess_step47-1}
+\includegraphics[width=.5\textwidth]{loess_step47-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/loess_fit-1}
+\includegraphics[width=\textwidth]{loess_fit-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/boot_loess-1}
+\includegraphics[width=\textwidth]{boot_loess-1}
 
@@
 \centering
-\includegraphics[width=.65\columnwidth]{images/matchup-rgl-copy.png}
+\includegraphics[width=.65\columnwidth]{matchup-rgl-copy.png}
 \caption{\label{fig:manualmatch-rgl}3D view of the manually adjusted side-by-side comparison of bullet~1-5 and bullet 2-1 after removing the curvature.
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/smooth-1}
+\includegraphics[width=.65\textwidth]{smooth-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/smoothcut-1}
+\includegraphics[width=.65\textwidth]{smoothcut-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/smoothmatch-1}
+\includegraphics[width=.65\textwidth]{smoothmatch-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/crosscuts-1}
+\includegraphics[width=\textwidth]{crosscuts-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/crosscutsX-1}
+\includegraphics[width=\textwidth]{crosscutsX-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/ccf-1}
+\includegraphics[width=.65\textwidth]{ccf-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/crosscuts-vary-b31-1}
+\includegraphics[width=\textwidth]{crosscuts-vary-b31-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.5\textwidth]{figure/smoothfac1-1}
+\includegraphics[width=.5\textwidth]{smoothfac1-1}
 
@@
 %' @
-\includegraphics[width=\textwidth]{images/br6-2-1-grey.png}
+\includegraphics[width=\textwidth]{br6-2-1-grey.png}
 \end{subfigure}
@@
 \caption{Barrel 9 Bullet 2-4}
-\includegraphics[width=\textwidth]{images/br9-2-4-grey.png}
+\includegraphics[width=\textwidth]{br9-2-4-grey.png}
 \end{subfigure}
@@
 \caption{Unknown Bullet B-2}
-\includegraphics[width=\textwidth]{images/b-2-grey.png}
+\includegraphics[width=\textwidth]{b-2-grey.png}
 \end{subfigure}
@@
 \caption{Unknown Bullet Q-4}
-\includegraphics[width=\textwidth]{images/q-4-grey.png}
+\includegraphics[width=\textwidth]{q-4-grey.png}
 \end{subfigure}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.5\textwidth]{figure/cms-1}
+\includegraphics[width=.5\textwidth]{cms-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/top-1}
+\includegraphics[width=\textwidth]{top-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/top2-1}
+\includegraphics[width=\textwidth]{top2-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/cms-bars-1}
+\includegraphics[width=\textwidth]{cms-bars-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/cms-spines-1}
+\includegraphics[width=\textwidth]{cms-spines-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/strange-res-1}
+\includegraphics[width=.65\textwidth]{strange-res-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/density-overview-1}
+\includegraphics[width=\textwidth]{density-overview-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/rocs-overview-1}
+\includegraphics[width=\textwidth]{rocs-overview-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.7\textwidth]{figure/tree-1}
+\includegraphics[width=.7\textwidth]{tree-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.7\textwidth]{figure/oob-1}
+\includegraphics[width=.7\textwidth]{oob-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.6\textwidth]{figure/tree:forest-1}
+\includegraphics[width=.6\textwidth]{treeforest-1}
 
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.7\textwidth]{figure/unnamed-chunk-1-1}
+\includegraphics[width=.7\textwidth]{unnamed-chunk-1-1}
 

### 2016-10-11 imaging-paper.tex
Make sure to include RData

@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in readChar(con, 5L, useBytes = TRUE): cannot open compressed file 'data/data-25-25/unkn9.RData', probable reason 'No such file or directory'}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in readChar(con, 5L, useBytes = TRUE): cannot open the connection}}\end{kframe}
 \includegraphics[width=.5\textwidth]{figure/cms-1}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in reslist[[which(cmsdist == 14)]]: attempt to select less than one element in get1index}}\end{kframe}
 \includegraphics[width=\textwidth]{figure/top2-1}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in readChar(con, 5L, useBytes = TRUE): cannot open the connection}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in reslist[[106]]: subscript out of bounds}}\end{kframe}
 \includegraphics[width=.65\textwidth]{figure/strange-res-1}

### 2016-10-17 revision/response2.Rmd
It's go time.

---
title: "Response to reviewers"
author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
date: "October 17, 2016"
output:
  pdf_document: default
bibliography: references.bib
---

# Reviewer 1

1. The authors have not provided a comprehensive literature review, to the point that the paper appears to be taking credit for the work that has been done by Dr. Nick Petraco, his collaborators, and others. I provided the references for Dr. Petraco’s work in my last review.

    **Please see the attached document for an in-depth discussion of our use of Dr. Petraco's work**

2. As I read the paper it is unclear how the matching criteria is different than the matching criteria currently in use by the tool mark’s community, especially those proposed by Monkres et al (2013) in Comparison and Statistical Analysis of Land
Impressions from Consecutively Rifled Barrels and the PhD dissertation work published in Technometrics by Amy B. Hoeksema with her PhD advisor Max Morris in 2013. Dr. Hoeksema’s dissertation also has a wonderful explanation of Chumbley’s U-statistic, which appears to be similar to the author’s proposed statistic, which is a PhD dissertation in the department of the authors of the paper currently under review. My opinion is that a discussion and comparison of these classical methods with their new comparison
statistic is necessary to demonstrate how the author’s work advances the science of land impression
comparisons.

    **On October 11th, 2016, we sat down with Dr. Morris in order to discuss the similarities and differences between the Chumbley U-statistic, and the methods used and features derived in our paper. He was surprised to see this comment, as the differences are rather stark. The Chumbley U-statistic characterizes toolmark signatures by producing two sets of correlation values. One set is produced by doing "coordinated shifts" after finding the ideal matching window, and computing each correlation value. The other set is computed by doing "independent shifts". Given that two toolmarks match, we'd expect that the correlation values from the coordinated shifts would be significantly higher than the independent shifts, and this is assessed using a Mann-Whitney U statistic. While very interesting and worth further discussion, Dr. Morris agreed that this is far different from the methods we have proposed. We use full pairs bullet signatures and extract features based on these. The features including correlation, as well as the number of matches, the number of consecutive matches, and more. At no time is a Mann-Whitney U statistic used, windows extracted, or any coordinated/independent shifts used.**

3. The main methods used in the paper under review are extremely dependent on Dr. Petraco’s work (implemented in an R library x3pr and available at https://github.com/npetraco). The authors built a new library titled x3prplus (available at https://github.com/heike/x3prplus) on top of x3pr. While working on my review, I went through a large amount of the two R libraries on github. The new library mainly appears to be wrappers for Dr. Petraco’s work. I find it very disconcerting and highly inappropriate that the authors are relying on Dr. Petraco’s work to such a high degree and that they are not referencing his work in their paper.

    **Please see the attached document for a rebuttal to these claims**

# Reviewer 2

4. This paper is uninteresting because it is not novel, even when considering the applied nature of the proposed work. It is very common for statisticians recently involved in forensic science to assume that nothing has been done prior to their own involvement, and this seems to be the assumption made by the authors. Unfortunately, much work on the subject of automated bullet comparisons has been done and published (but not necessarily in the statistical literature) and the authors of this paper are simply re-inventing the wheel without providing anything new to the debate.

    **XXX**

5. The paper is pedestrian and incomplete because extremely similar work has been proposed by Jain, Chumbley, Prokoski, Petraco, Zheng and so on (the list is available to this reviewed, who is happy to share it with the editor if needed). The authors even ignored a milestone paper published by their own department head... The authors of the current paper have not made any attempt to review the relevant literature: if anything, they merely took the references that I provided in my first review and added them to their bibliographic review, while dismissing their work. For example, the authors' treatment of Riva et al's work is not appropriate: it is not because they consider cartridge cases that their methodology could not be applied to landmarks on bullets! I believe that if the authors had bothered reviewing the literature they would not have ended up re-inventing the wheel and they may have actually done something useful for the community.

    **XXX**

6. The paper is scientifically questionable because (1) nothing in this paper is novel (as discussed above), (2) the authors are extremely confused as to what type of inference they are proposing. They emphasize that legal and scientific scholars advocate the use of a Bayesian framework for the quantification of the weight of forensic evidence, yet the entire paper is about calculating error rates. I assume that the authors understand that in order to make an error, one has to make a decision, and whether in a Bayesian or frequentist framework, a decision is related to posterior probabilities and utility functions, not to Bayes factors. Bottom line, if the authors are endorsing the use of a Bayesian framework, they spent a great deal of effort addressing the wrong problem. If the authors are not endorsing a Bayesian framework, then they should say so, and scientifically support their work on based on its own merits.

    **XXX**

7. The paper is ethically problematic because it seems that most of the code made available by the author is similar to, if not require the use of, legacy code provided by another author working in the same area, and who is absolutely not referenced in the current paper.

    **We referenced the OpenFMC organization and devoted a paragraph to its discussion. This is the organization that is the recommended citation for the package which we use a function of, `x3pr`. We have added a reference to Dr. Petraco directly in this discussion. See the below citation, given by `x3pr` as of October 17th, 2016:**
    ```{r}
    citation("x3pr")
    ```

8. Overall, this paper does not advance the field of forensic science one bit: (1) the statistical analysis proposed by the authors is not novel at all, thus we are not learning anything from it; (2) the final results reported by the authors are useless since (i) they are related to the wrong type of inference (error rates vs. Bayes factors, and (ii) they are based on such a small and distorted dataset that they can't even be used in court to support the admissibility (or rejection) of firearm evidence.

    **We are very clear about the limitations presented by a small dataset, particularly in the conclusion. Unfortunately, this is the data that we had to work with. With that said, the idea of statistical modeling is that once new data is obtained, its rather trivial to compute the set of features used in our model, and retrain the model with a broader and more varied dataset. Whether performance on holdout samples would remain as strong remains to be seen. But because this code is open-source and in the public domain, the algorithm can be assessed on new data by anyone willing to invest the time to run it.**

# Reviewer 4

9. This paper describes a statistical experiment to to determine whether two bullets were fired from the same gun barrel. The authors formulate a simple model, and perfom a version of least squares.

    **We feel this is a rather inaccurate summary of the work done in our paper. We do use a random forest model, but this is not the focus. Rather, the data analysis steps culminating in feature extraction are the focus. These features, computed on pairs of bullet lands, are then fed into a random forest model, but we do formulate this model. Furthermore, we do not perform a version of least squares - The random forest uses decision trees. We felt the purpose of our work was to highlight the individual data analysis steps that compromise a full analysis of bullet striations, beginning from an open-source data file, and ending with a predicted probability of a match.**

10. No new statistical methodology is introduced, so the impact of this work is very narrow.  There is also no attempt to relate the work to any previous statistical studies or tests, as there is only one reference.

    **The revised copy of our paper contains over thirty references, not just one. Many are from previous statistical studies on bullet matching (e.g., Biasotti 1959, Chu 2010, Riva 2014) while many more are references on statisitcal tests themselves (e.g., Therneau 2015, Liaw 2002, Cleveland 1979)**

11. I believe that this work should be submitted to a more specialized journal, but Annals of Applied Statistics is definitely the wrong venue.

    **XXX**

# Reviewer 5

12. The authors have addressed the majority of my comments. Exceptions are: Comment 1): The authors corrected the second sentence in the abstract as suggested, yet did not correct the first sentence of the Introduction, which was a copy-paste of the second sentence in the abstract, and this first sentence in the Introduction is now still incorrect. This sentence, currently reading “Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.” should be replaced by “Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel.” in order to be correct and consistent with the abstract.

    **Thank you. We apologize for the oversight, and have made this correction.**

13. Comment 2): The weakness of the paper is that it lacks an explanation for why a binary “match” vs “non-match” approach was chosen. It does not explain what the value of the observations is with regard to the two competing propositions, i.e., the question of interest to the forensic scientist. If the authors want this to be a good paper, I encourage them to provide such an explanation.

    **We have attempted to extend this discussion. Although we have assessed our algorithm in terms of a binary "match" vs. "non-match" scale, we didn't attempt to reduce the full scope of the problem down to this decision. As a matter of fact, the output from our model are predicted probabilities of a match, or scores from the random forest. These scores can be seen in Figure 19. For purposes of assessment, we considered a predicted probability of at least 50% as indicative of a match. However, in a real world application of this algorithm, the scores themselves are most certainly relevant. We apologize that we didn't make this distinction as clear in the previous revision, and we hope this alleviates some of the reviewer's concerns.**

14. Comment 3): The term “ballistics” was corrected as suggested. Yet there remain two occurrences of “ballistic” as an adjective in lines 477 (“ballistic images”) and 490 (“ballistic imaging”). I interpret these as referring to images or imaging of the trajectory of bullets. If this is not what was meant, I strongly encourage the authors to replace “ballistic images” and “ballistic imaging” with phrases that represent what they actually meant.

    **We have corrected the use of ballistics in these instances.**

15. Comment 7): An attempt was made to visually distinguish between the two bullets, yet note that the shading that was added is barely visible when the figure is printed on a color laser printer.

    **Thank you for pointing this out, as we did not attempt to check the printed copy of the paper. We have corrected this to improve the printed appearance.**

# Reviewer 6

16. The revised paper "Automatic Matching of Bullet Lands" describes an automated procedure for matching bullets that have been fired from the same gun.  It also compares the accuracy of the procedure on a test set of bullet images from the Hamby, Brundage and Thorpe study.  The key conclusion is that the automated algorithm outperforms human matching, and achieves high accuracy. The methods used in the paper are statistically simple, but also robust.  The problem is important and the paper fills a need.

    **We thank you for the kind words.**

17. Some aspects of the algorithm are judgmental, and I expect that at each step in the algorithm one might be able to find an alternative that is incrementally smarter.  But I think that is permitting the best to be the enemy of the good.

    **We agree wholeheartedly with the reviewer. Several best judgment decisions were made during the course of the algorithm. We attempted to be explicit when certain parameters were chosen (e.g., the smoothing factor parameter) and include a discussion of how the results change when these parameters are varied. As such, thus algorithm could be modified to use best parameters as chosen from cross-validation, for instance.**

18. The paper is clearly written and sufficiently concise.  The description and motivation for the work is good.  And the code is in the public domain, which is a great contribution.

    **Thank you once again for the comments.**

## x3prplus

```{r, child='x3prplus.Rmd'}

```

### 2016-10-17 revision/x3prplus.Rmd
It's go time.

---
title: "x3prplus vs x3pr"
author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
date: "10/11/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

We have compiled this document in response to a serious accusation raised by a reviewer of our paper Automatic Matching of Bullet Lands. The accusation is reproduced below:

*The main methods used in the paper under review are **extremely dependent on Dr. Petraco's work** (implemented in an R library x3pr and available at https://github.com/npetraco). The authors built a new library titled x3prplus (available at https://github.com/heike/x3prplus) on top of x3pr. While working on my review, I went through a large amount of the two R libraries on github. **The new library mainly appears to be wrappers for Dr. Petraco’s work**. I find it very disconcerting and highly inappropriate that the authors are relying on Dr. Petraco’s work to such a high degree and that they are not referencing his work in their paper.*

In this document, we will show quite easily that this accusation is demonstrably false. To begin, we can easily access all exported functions of the `x3pr` package with a call to `getNamespace`. If we do so for `x3pr`, we obtain the following list:

```{r}
ls(getNamespace("x3pr"))
```

Performing the same task for `x3prplus`:

```{r}
ls(getNamespace("x3prplus"))
```

We will use these functions to illustrate the vast separation of functionality between the two packages.

## Function Similarities

Did we attempt to take credit for work that was done by Dr. Petraco? Below is a list of all the functions available in the `x3pr` package (accessed 10/10/2016). Further down, you can see the list of all the functions available in `x3prplus`. As is clear from the function names and descriptions, **there is virtually no overlapping functionality**. In general, Dr. Petraco's functions are about reading and parsing file formats. Our functions assume the reading step has already taken place, and compromise a much broader set of statistics and data science functionality including fitting models, deriving features, detecting grooves in bullet lands, and more.

x3pr Function              Comparable x3prplus Function Description                   
-------------------------- ---------------------------- -----------
launch_3Dview              None                         Shiny app to view a 3D surface
lms2x3p.file               None                         Convert a Zeiss LMS file to an X3P format
make.x3p.header            None                         Put together the XML format x3p header
plot3D.x3p.file            $\text{None}^1$              3D plot of a surface in X3P format
print.x3p.file.info        None                         Format (most) x3p header file info
read.digital.surf.file     None                         Read in a Digital Surf file
read.digital.surf.header   None                         Read the header off a Digital Surf file
read.digital.surf.profiles None                         Read in the surface data of a Digital Surf file
read.x3p                   read.x3pplus                 Read an x3p file
read.zeiss.lms.file        None                         Read a surface file in Zeiss LMS format
read.zeiss.lms.header      None                         Read the header off a Zeiss LMS file
read.zeiss.lms.profiles    None                         Read out surface data in Zeiss LMS format
sur2x3p.file               None                         Convert a read in .sur file to x3p
write.x3p                  $\text{None}^2$              Write a x3p format file

1: `x3prplus` includes an internal function, `plot_3d_land`, which uses the plotly package to produce a surface plot of a bullet land. This function is not accessible to users and uses no functionality from the `x3pr` function `plot3D.x3p.file`.

2: The `unfortify_x3p` function converts a data frame representation of a bullet land into the surface matrix version native to `x3pr`. However, this does not depend on the `write.x3p` code, nor does it have any of the disk writing functionality.

## Function Wrappers

It is rather trivial to assess the claim that `x3prplus` "mainly appears to be wrappers" for `x3pr`. In fact, we can very simply determine the unique number of `x3pr` functions in use by `x3prplus` to determine whether there is any evidence that `x3prplus` is merely a wrapper.

x3prplus Function    x3pr Function Used Description
-------------------- ------------------ -----------
boot_fit_loess       None               Fit a LOESS model with bootstrap samples
bulletAlign          None               Align two surface cross cuts according to maximal correlation
bulletAlign_dtw      None               Align two surface cross cuts according to maximal correlation
bulletAlign_nist     None               Align two surface cross cuts according to maximal correlation
bulletCheckCrossCut  None               Identifying a reliable cross section
bulletGetMaxCMS      None               Identify the number of maximum CMS between two bullet lands
bulletGetMaxCMS_nist None               Identify the number of maximum CMS between two bullet lands
bulletSmooth         None               Smooth the surface of a bullet
CMS                  None               Table of the number of consecutive matches
fit_loess            None               Fit a loess curve to a bullet data frame
fortify_x3p          None               Convert a list of x3p file into a data frame
get_bullet           None               Deprecated function use get_crosscut
get_crosscut         None               Read a crosscut from a 3d surface file
get_grooves          None               Find the grooves of a bullet land
get_peaks            None               Identify the location of peaks and heights at a crosscut
get_peaks_nist       None               Identify the location of peaks and heights at a crosscut
getCircle            None               Estimate center and radius
getTwist             None               Estimate the twist in a bullet land
maxCMS               None               Number of maximum consecutively matching striae
predCircle           None               Estimate predictions and residuals for a circle fit of x and y
predSmooth           None               Estimate predictions and residuals for a smooth of x and y
processBullets       None               Process x3p file
read.x3pplus         read.x3p           Read X3P File with option to transpose matrix
sample.x3p           None               Sample every X element of a data frame
smoothloess          None               Predict smooth from a fit
striation_identify   None               Match striation marks across two cross sections
unfortify_x3p        None               Convert a data frame into an x3p file

In total, **there is exactly one single function call made to an `x3pr` function**. This is a function we wrote, `read.x3pplus`, which extends `read.x3p` with a simple option to transpose the resulting matrix. Not a single other `x3pr` function is ever called in `x3prplus`. Comparatively, we call dozens of unique functions belonging to the `ggplot2` package. It is clear that we aren't attempting to be a wrapper for `x3pr` anymore than we are a wrapper for `ggplot2`. In fact, the set of functionality of both packages is wildly different, as can be seen from a simple read of the function descriptions in the two tables above.

In other words, the claim that `x3prplus` is "extremely dependent on Dr. Petraco's work" is demonstrably false. As a matter of fact, we have an internal version of the code which replaces the single use of Dr. Petraco's code, in our `read.x3pplus` function, with a natively written XML parser. Nonetheless, we would prefer to build on Dr. Petraco's work, which is most certainly in the spirit of open-source software and academic research. This is also likely why Dr. Petraco has made his code publicly available on GitHub.

## Referencing x3pr

The question remains of whether despite only using a single function of `x3pr`, and our package bearing virtually no similarlity in terms of functionality, did we fail to provide proper attribution of the package. Dr. Petraco contacted Eric regarding this issue, in an email which was CCed to his major professor (Dr. Heike Hofmann), the director of CSAFE (Dr. Alicia Carriquiry), the Department Chair of Statistics at Iowa State University (Dr. Max Morris) and the CSAFE contract officer at NIST (Susan Ballou). Dr. Petraco claimed "I am puzzled however as to why no citation appears to this work anywhere in your text as `x3pr` contributed critically to it." The paper referenced was a pre-publication version published to Arxiv, and was still in the revision stage. Eric acknowledged the omission of the direct reference to `x3pr` in the pre-publication paper, and added it that afternoon, within hours of receiving the email:

https://github.com/erichare/imaging-paper/commit/c63cfeee7134222a9b760752fca80084aadad866

Further, numerous other references to Dr. Petraco and `x3pr` have existed within the context of our package `x3prplus`, including but not limited to:

1. The README for the package: https://github.com/heike/x3prplus
2. Presentations given on x3prplus and our algorithm
	- https://github.com/heike/x3prplus/blob/presentations/jsm-2016/slides.Rmd
	- https://github.com/heike/x3prplus/blob/presentations/graphics-group-2015/slides.Rmd
	- https://github.com/heike/x3prplus/blob/presentations/forensics-group-late2016/slides.Rmd
3. Within the package description itself.
4. Directly in the name of the package `x3prplus`

As a matter of fact, we submitted a pull request in order to fix an issue in `x3pr` which was detected over a year ago. Dr. Petraco merged it in to fix the bug, as can be seen by this screenshot taken October 11th. This is also the most recent change that Dr. Petraco has made to his package, and predates serious work on `x3prplus`.

```{r, echo=FALSE}
knitr::include_graphics("github.png")
```

Although using `x3pr` merely to read in the data, we felt that it was important to reference the work that had been put into it. Given this, the reviewer finding our package "very disconcerting" and "highly inappropriate" is quite frankly a very strange assertion, particularly given the lack of any evidence provided in the review.

### 2016-10-25 revision/response2.Rmd
Typo fixes and response rebuilding

@@
 author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
-date: "October 17, 2016"
+date: "October 24, 2016"
 output:
@@
 I provided the references for Dr.
-Petraco’s work in my last review.
+Petraco's work in my last review.
 
-    **Please see the attached document for an in-depth discussion of our use of Dr. Petraco's work**
+    **Most of the references that were suggested by reviewers earlier were incorporated into the revised version of the manuscript.
+Regarding the "overlap" with Petraco's work, please see the attached document entitled x3prplus for an in-depth discussion of our use of exactly one function in Petraco's package.**
 
@@
 My opinion is that a discussion and comparison of these classical methods with their new comparison
-statistic is necessary to demonstrate how the author’s work advances the science of land impression
-comparisons.
+statistic is necessary to demonstrate how the author’s work advances the science of land impression comparisons.
 
-    **On October 11th, 2016, we sat down with Dr.
-Morris in order to discuss the similarities and differences between the Chumbley U-statistic, and the methods used and features derived in our paper.
-He was surprised to see this comment, as the differences are rather stark.
-The Chumbley U-statistic characterizes toolmark signatures by producing two sets of correlation values.
+    **First, it seems more appropriate to refer to Hoeksema's U statistic since she was the one to propose its use in her dissertation.
+As the reviewer notes, this work was carried out in our own department, and we were well aware of it.
+Thus our surprise with the review, since to say that our work is a repeat of Amy Hoeksema's work is so inaccurate as to make us wonder whether the reviewer is familiar with either approaches.
+On October 11th, 2016, we sat down with Max Morris to discuss the similarities and differences between the Hoeksema U-statistic, and the methods used and features derived in our paper.
+Max was at least as surprised as we were to read the reviewer's comment, since the differences between Amy's work and ours are stark.
+The Hoeksema U-statistic characterizes toolmark signatures by producing two sets of correlation values.
 One set is produced by doing "coordinated shifts" after finding the ideal matching window, and computing each correlation value.
 The other set is computed by doing "independent shifts".
-Given that two toolmarks match, we'd expect that the correlation values from the coordinated shifts would be significantly higher than the independent shifts, and this is assessed using a Mann-Whitney U statistic.
-While very interesting and worth further discussion, Dr.
-Morris agreed that this is far different from the methods we have proposed.
-We use full pairs bullet signatures and extract features based on these.
-The features including correlation, as well as the number of matches, the number of consecutive matches, and more.
-At no time is a Mann-Whitney U statistic used, windows extracted, or any coordinated/independent shifts used.**
+When the two toolmarks match, we'd expect that the correlation values from the coordinated shifts would be significantly higher than those computed from the independent shifts, and this is assessed using a Mann-Whitney U statistic.
+While very interesting and worth further discussion, Max Morris agreed that this is far different from the methods we have proposed.
+In our work, we use full bullet signatures from pairs of bullets and extract several features based on these.
+The features include a cross-correlation, as well as the number of matching striae, the number of consecutive matching striae, the difference in the height of the signatures at each peak and at each valley, and several more.
+At no time is a Mann-Whitney U statistic used, windows extracted, or any coordinated/independent shifts used.
+These are completely different approaches where the only "similarity" is that we also compute a correlation, albeit between different variables!**
 
@@
 
-    **Please see the attached document for a rebuttal to these claims**
+    **The reviewer must have looked at the wrong code because no one who is reasonably familiar with R would claim that x3prplus is a "wrapper" for x3pr!
+Please see the attached document for a rebuttal to these claims.
+In x3prplus, all we did is call ONE of Petraco's functions, read.x3p, to read 3D files.
+Everything else is completely different, as can be seen in the attached document.
+Perhaps our choice of name for the package has contributed to the confusion.**
+
+\newpage
 
@@
 
-    **XXX**
+    **We thank the reviewer for his/her lecture, but it is difficult to take his/her comments seriously when they are offered with absolutely no evidence.
+Exactly whose work are we re-inventing?
+And where has work similar to ours been published?
+If the reviewer is going to make statements such as the ones above, please, try to back them up with some facts.**
 
@@
 
-    **XXX**
+    **Really?
+We have referred to many of those papers in the manuscript and have shown how our approach is quite different.
+Indeed, even PCAST (the President's Council of Advisors in Science and Technology) in its report, mentioned that our work was the only promising work and they actually looked at all the literature themselves.
+So we respectfully disagree with Reviewer 2.
+The Riva and Champod paper published in 2014 is indeed referred to in the manuscript and of all the published work, it is the one that most resembles what we have done.
+The reviewer poo-poos our saying that one difference is that Riva and Champod worked on casings and we worked on bullets.
+The reviewer would be well advised to read our comment more carefully;  the reason it is important is that Riva and Champod make use of the circular structure of their data to pre-process their images and develop their score.
+A big component of their algorithm is the alignment of images, which they do by rotating the images on the xY plane.
+This is not something that cannot be extended to the case of bullets, were the XY plane is not flat and where rotation makes no sense.
+The approach they use to extract features from the aligned images is completely different from ours and for the reviewer to say that we are re-inventing the wheel is a bit of a scandal.
+The fact that Riva and Champod have come up with a nice automated algorithm to match breech face and firing pin images does not mean that we cannot come up with a DIFFERENT automated algorithm to match bullet lands.
+This said, we have added additional commentary about Riva and Champod in our manuscript.**
 
@@
 
-    **XXX**
+    **Again, we have to disagree with the reviewer and again we take issue at his/her bandying about such incendiary comments with nothing to back them up.
+We are not confused, in particular not on the issue of inference.
+In the intro we mentioned the type of inference that is often drawn in legal settings, mostly to satisfy an earlier reviewer (an actual constructive reviewer) who requested that we add a bit of general background in the introduction.
+We then proceed to explain that this is not what we pursue in our paper;  all we want to present is an automated algorithm to decide whether two bullets may have been fired from the same gun.
+No more and no less.
+Yes, we are focusing on error rates, and for the reviewer's peace of mind, yes, we realize that there is a decision implicit in our trying to compute an error rate.
+The decision is simply whether two bullets may have been fired from the same gun.
+We have gone back to the manuscript and added some additional language to explain the purposes of the work we present.**
 
@@
     **We are very clear about the limitations presented by a small dataset, particularly in the conclusion.
-Unfortunately, this is the data that we had to work with.
-With that said, the idea of statistical modeling is that once new data is obtained, its rather trivial to compute the set of features used in our model, and retrain the model with a broader and more varied dataset.
+Unfortunately, this are the data that we had to work with.
+With that said, the idea of statistical modeling is that once new data are obtained, it will be rather trivial to compute the set of features used in our model, and retrain the model with a broader and more varied dataset.
 Whether performance on holdout samples would remain as strong remains to be seen.
-But because this code is open-source and in the public domain, the algorithm can be assessed on new data by anyone willing to invest the time to run it.**
+But because this code is open-source and in the public domain, the algorithm can be assessed on new data by anyone willing to invest the time to run it.
+While the dataset is small (only 10 barrels) we have been told by multiple firearm examiners that it is a challenging dataset on which to develop methods because the 10 barrels were consecutively manufactured.
+Thus, one would expect that thre striae made by those barrels will be more similar across them than across two randomly selected barrels from a large population.
+We are collaborating with Alan Zhang in NIST to assemble a larger database that will be publicly available within a year or two.**
+
+\newpage
 
@@
 Furthermore, we do not perform a version of least squares - The random forest uses decision trees.
-We felt the purpose of our work was to highlight the individual data analysis steps that compromise a full analysis of bullet striations, beginning from an open-source data file, and ending with a predicted probability of a match.**
+(Is the reviewer thinking of regression trees?
+If so, that is not what we are implementing.) One purpose of our work is to highlight the individual data analysis steps that compromise a full analysis of bullet striations, beginning from an open-source data file, and ending with a predicted probability of a match.**
 
@@
 
-    **The revised copy of our paper contains over thirty references, not just one.
+    **Did the reviewer read the correct paper?
+The revised copy of our paper contains over thirty references, not just one.
 Many are from previous statistical studies on bullet matching (e.g., Biasotti 1959, Chu 2010, Riva 2014) while many more are references on statisitcal tests themselves (e.g., Therneau 2015, Liaw 2002, Cleveland 1979)**
@@
 
-    **XXX**
+    **We believe that the focus on a timely application and the novel combination of statistical and machine learning methods to address the problem warrant publication in the Annals of Applied Statistics.**
+
+\newpage
 
@@
 
+\newpage
+
 # Reviewer 6
@@
 
-## x3prplus
+\newpage
+
+# x3prplus
 

### 2016-10-26 imaging-paper.Rnw
More updates

@@
 %}
-Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel.
+Firearm examination is a forensic tool used to help the court determine whether two bullets were fired from the same gun barrel.
 This process has broad applicability in terms of convictions in the United States criminal justice system.
@@
 The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match.
+While we assess the performance of the algorithm in terms of a binary decision of match vs.
+non-match using a 50\% probability cut-off, our primary goal is to highlight the features that are statistically associated with matches and non-matches, and to provide a quantitative assessment of this association.
+In a real-world application of our algorithm, the raw scores would need further analysis and scrutiny, and it is likely a 50\% cut-off would be an inappropriate choice on the basis of reasonable doubt.
+
 Our algorithm is fully open source and available on GitHub \citep{x3prplus}.
@@
 
+Work has been done to quantify toolmark similarity by \citet{lock2013significance}.
+This algorithm determines an optimal matching window between two toolmark signatures, and then performs a set of both coordinated and independent shifts.
+Given a match, the coordinated shifts would be expected to yield correlation values higher than those obtained from independent shifts.
+This is assessed using a Mann-Whitney U Statistic.
 A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}.
@@
 The authors state that this process `could be automated', but no implementation of the algorithm is available.
+
 Modern automated techniques using 3D images have also been conducted by \citet{riva:2014}, but this work applies to striae found on cartridge cases and not bullets.
+This algorithm performs alignment of striae by rotation of the XY plane, which is not generalizable to bullets in which the XY plane is not flat.
 Other work on 3D images has been conducted by \citet{petraco:2012}, which also describes work carried out on cartridge cases, as well as screwdriver striation patterns, with promising results.

### 2016-10-26 imaging-paper.Rnw
No more ballistic imaging

@@
 
-The feasibility of creating a database of ballistic images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
+The feasibility of creating a database of images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
 The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
@@
 The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
-The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks.
+The findings of the NRC report on imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks.
 This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system.

### 2016-11-30 imaging-paper.Rnw
Add bulletr references

@@
 library(dplyr)
-library(x3pr)
-library(x3prplus)
+library(bulletr)
 library(grid)
@@
 
-br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
+br111 <- read_x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
 dbr111 <- fortify_x3p(br111)
@@
 #lof <- processBullets(paths = images[c(5,7)], x = 100)
-subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
-subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
+subLOFx1 <- processBullets(read_x3p(images[5]), name = images[5], x = 100)
+subLOFx2 <- processBullets(read_x3p(images[7]), name = images[7], x = 100)
 #subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
@@
 
-  subLOFx1 <- processBullets(read.x3p(images[5]), name = images[5], x = 100)
-  subLOFx2 <- processBullets(read.x3p(images[7]), name = images[7], x = 100)
+  subLOFx1 <- processBullets(read_x3p(images[5]), name = images[5], x = 100)
+  subLOFx2 <- processBullets(read_x3p(images[7]), name = images[7], x = 100)
   lof <- rbind(subLOFx1, subLOFx2)
@@
 crosscuts <- seq(100,250, by=25)
-lof <- processBullets(read.x3p(paths[37]), name = paths[37], x = crosscuts)
+lof <- processBullets(read_x3p(paths[37]), name = paths[37], x = crosscuts)
 lof$bullet <- paste(scrubPath(lof$bullet), lof$x)
@@
 %' <<echo=FALSE>>=
-%' b1 <- read.x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
+%' b1 <- read_x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
 %' image(b1$surface.matrix)

### 2016-11-30 imaging-paper.tex
Add bulletr references

@@
 Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
-In the Hamby data, typical length (groove-to-groove) of a land is about 2005.9 micrometers or 2.01 millimeters.
+In the Hamby data, typical length (groove-to-groove) of a land is about 1998.28 micrometers or 2 millimeters.
 For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
@@
 \subsection{Algorithm}\label{sec:algorithm}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in 1:info\$num.pts.line: argument of length 0}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in 1:info\$num.pts.line: argument of length 0}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'subLOFx1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in data.frame(subLOFx1): object 'subLOFx1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in factor(lofX\$r05): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in levels(lofX\$type) <- c("{}groove"{}, NA, "{}peak"{}): object 'lofX' not found}}\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in matches\$allnas: object of type 'closure' is not subsettable}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in matches\$allnas: object of type 'closure' is not subsettable}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in UseMethod("{}group\_by\_"{}): no applicable method for 'group\_by\_' applied to an object of class "{}function"{}}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in subset.default(lines, lineid != 0): object 'lineid' not found}}\end{kframe}
+\end{knitrout}
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/smooth-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(data, aesthetics, environment = env): object 'lofX' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/crosscuts-vary-b31-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in 1:info\$num.pts.line: argument of length 0}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'lof' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in subset(lof, x \%in\% crosscuts[i:(i + 1)]): object 'lof' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(X = X, FUN = FUN, ...): object 'reslist' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(reslist, function(res) \{: object 'reslist' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in arrangeGrob(...): object 'lop' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in readChar(con, 5L, useBytes = TRUE): cannot open compressed file 'data/data-25-25/unkn9.RData', probable reason 'No such file or directory'}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in readChar(con, 5L, useBytes = TRUE): cannot open the connection}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(X = X, FUN = FUN, ...): object 'reslist' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'cmsdist' not found}}\end{kframe}
 \includegraphics[width=.5\textwidth]{figure/cms-1}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/top-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'reslist' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'res' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in fortify(data): object 'res' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/top2-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'reslist' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'res' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in fortify(data): object 'res' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/strange-res-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in readChar(con, 5L, useBytes = TRUE): cannot open the connection}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'reslist' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'res' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in fortify(data): object 'res' not found}}\end{kframe}
 \end{knitrout}

### 2016-11-30 imaging-supplement.Rnw
Add bulletr references

@@
   bname <- sprintf(file.path(datadir, "Br1 Bullet 1-%d.x3p"), i)
-  dbi <- fortify_x3p(read.x3p(bname))
+  dbi <- fortify_x3p(read_x3p(bname))
   dbi$part <- i
@@
   bname <- sprintf(file.path(datadir, "Br1 Bullet 2-%d.x3p"), i)
-  dbi <- fortify_x3p(read.x3p(bname))
+  dbi <- fortify_x3p(read_x3p(bname))
   dbi$part <- i
@@
 crosscuts <- seq(100, 400, by = 50)
-lof <- processBullets(read.x3p(paths[5]), name = scrubPath(paths[5]), x = crosscuts)
+lof <- processBullets(read_x3p(paths[5]), name = scrubPath(paths[5]), x = crosscuts)
 lof$bullet <- paste(lof$bullet, lof$x)
@@
 <<setup-signatures, echo=FALSE, message=FALSE>>=
-library(x3pr)
-library(x3prplus)
+library(bulletr)
 library(dplyr)
@@
 all_bullets <- lapply(as.character(ccs$path), function(x) {
-  result <- read.x3p(x)
+  result <- read_x3p(x)
   result[[3]] <- x
@@
 x <- as.character(ccs$path)[k]
-land <- read.x3p(x)
+land <- read_x3p(x)
 land[[3]] <- x

### 2016-11-30 imaging-paper.Rnw
Rebuild both papers with no cache

@@
 \subsection{Algorithm}\label{sec:algorithm}
-<<twolands100, cache=TRUE,  echo=FALSE>>=
+<<twolands100,  echo=FALSE>>=
 images <- file.path(datadir, dir(datadir))
@@
 
-<<twolands100match, cache=TRUE, echo=FALSE, dependson='twolands100'>>=
+<<twolands100match, echo=FALSE, dependson='twolands100'>>=
 matches <- lofX %>% group_by(y) %>% summarise(
@@
   \centering
-<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', warning = FALSE>>=
 paths <- file.path(datadir, dir(datadir))

### 2016-11-30 imaging-supplement.Rnw
Rebuild both papers with no cache

@@
 
-<<bullet1, echo=FALSE, cache=TRUE>>=
+<<bullet1, echo=FALSE>>=
 db1 <- NULL
@@
 
-<<bullet2, echo=FALSE, cache=TRUE>>=
+<<bullet2, echo=FALSE>>=
 db2 <- NULL
@@
   \centering
-<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
+<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', warning = FALSE>>=
 
@@
 
-<<aligned, echo=FALSE, warning=FALSE,  cache=TRUE>>=
+<<aligned, echo=FALSE, warning=FALSE>>=
 matches <- read.csv("csvs/matches-old.csv", header = FALSE)

### 2016-11-30 imaging-paper.Rnw
Fix paper building on linux

@@
 
+im1 <- "images/Hamby (2009) Barrel/bullets/Br1 Bullet 1-5.x3p"
+im2 <- "images/Hamby (2009) Barrel/bullets/Br1 Bullet 2-1.x3p"
+
 #lof <- processBullets(paths = images[c(5,7)], x = 100)
-subLOFx1 <- processBullets(read_x3p(images[5]), name = images[5], x = 100)
-subLOFx2 <- processBullets(read_x3p(images[7]), name = images[7], x = 100)
+subLOFx1 <- processBullets(read_x3p(im1), name = im1, x = 100)
+subLOFx2 <- processBullets(read_x3p(im2), name = im2, x = 100)
 #subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
@@
 #lof <- processBullets(paths = images[c(5,7)], x = 100)
-subLOFx1 <- processBullets(read_x3p(images[5]), name = images[5], x = 100)
-subLOFx2 <- processBullets(read_x3p(images[7]), name = images[7], x = 100)
+subLOFx1 <- processBullets(read_x3p(im1), name = im1, x = 100)
+subLOFx2 <- processBullets(read_x3p(im2), name = im2, x = 100)
 #subLOFx1 <- subset(lof, bullet=="Br1 Bullet 1-5")
@@
 
-  subLOFx1 <- processBullets(read_x3p(images[5]), name = images[5], x = 100)
-  subLOFx2 <- processBullets(read_x3p(images[7]), name = images[7], x = 100)
+  subLOFx1 <- processBullets(read_x3p(im1), name = im1, x = 100)
+  subLOFx2 <- processBullets(read_x3p(im2), name = im2, x = 100)
   lof <- rbind(subLOFx1, subLOFx2)

### 2016-11-30 imaging-supplement.Rnw
Fix paper building on linux

@@
 
+im1 <- "images/Hamby (2009) Barrel/bullets/Br1 Bullet 1-5.x3p"
+
 crosscuts <- seq(100, 400, by = 50)
-lof <- processBullets(read_x3p(paths[5]), name = scrubPath(paths[5]), x = crosscuts)
+lof <- processBullets(read_x3p(im1), name = scrubPath(im1), x = crosscuts)
 lof$bullet <- paste(lof$bullet, lof$x)

### 2016-11-30 code/test_readx3p.R

@@ -6,7 +6,7 @@ datamat <- mybullet$surface.matrix
 
 plot_ly(z = ~datamat) %>% add_surface()
 
-mybullet2 <- read_x3p("images/Hamby (2009) Barrel/bullets/br1_1_land1.x3p", transpose = TRUE)
+mybullet2 <- read_x3p("images/Hamby (2009) Barrel/bullets/br1_1_land1.x3p")
 datamat2 <- mybullet2$surface.matrix
 
 plot_ly(z = ~datamat2) %>% add_surface()

### 2016-11-30 code/bullet-app.R
Bullet db app

@@ -1,0 +1,35 @@
+library(shiny)
+library(plotly)
+library(bulletr)
+
+hamby_dir <- "../images/Hamby (2009) Barrel/bullets"
+cary_dir <- "../images/Cary Persistence/bullets/"
+
+all_images <- c(file.path(hamby_dir, dir(hamby_dir)), file.path(cary_dir, dir(cary_dir)))
+
+ui <- fluidPage(
+   
+   titlePanel("Bullet Database"),
+   
+   sidebarLayout(
+      sidebarPanel(
+          selectInput("land", "Bullet Land", choices = all_images)
+      ),
+      
+      mainPanel(
+          plotlyOutput("viewland")
+      )
+   )
+)
+
+server <- function(input, output) {
+    mybullet <- reactive({
+        read_x3p(input$land)
+    })
+    
+    output$viewland <- renderPlotly({
+        plot_3d_land(bullet = mybullet())
+    })
+}
+
+shinyApp(ui = ui, server = server)

### 2017-01-12 imaging-paper.Rnw
Paper edits from Dr. Carriquiry

@@
 During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet.
-Identifying these striation markings in an attempt to match two bullets is one of the primary goals.
-We propose an automated framework for the analysis of the 3D surface measurements, that first transcribes the markings into a 2D plotting framework.
+Identifying these striation markings in an attempt to match two bullets is one of the primary goals of firearm examination.
+We propose an automated framework for the analysis of the 3D surface measurements of bullet lands that first transcribes the markings into a 2D plotting framework.
 This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels.
@@
 These marks are assumed to be unique to the barrel, as described in a 1992 AFTE article~\citep{afte:1992}.
-``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement" The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."
+``The theory of identification as it pertains to the comparison of toolmarks enables opinions of common origin to be made when the \emph{unique surface contours} of two toolmarks are in sufficient agreement".
+The article goes on to state that ``Significance is determined by the comparative examination of two or more sets of surface contour patterns comprised of individual peaks, ridges and furrows."
 
@@
 
-Current standard practice~\cite{afte:1992} relies in part on the assessment of the so-called maximal number of consecutively matching striae (CMS), first defined by \citet{biasotti:1959}.
-One of the problems is that a human inspection to determine CMS is subjective \citep{miller:1998}.
-
-Here, we focus on the question of defining what constitutes a match, since there is no agreement in the forensics community.
+Current firearm examination practice relies mostly on visual assessment and comparison of striation.
+Indeed, the AFTE Theory of Identification (https://afte.org/about-us/what-is-afte/afte-theory-of-identification) explicitly requires that examiners evaluate the strength of similarity between two samples relative to other comparisons they may have carried out in the past.
+An attempt to quantify the degree of similarity consists in counting the number of consecutively matching striae (CMS) between two bullets, first proposed by \citet{biasotti:1959}.
+This approach has two drawbacks, however.
+First, determining matching striae is still a subjective activity.
+Second, as discussed by \citet{miller:1998}, the number of CMS may be high even if the bullets were not fired by the same gun.
+
+Here, we focus on the question of defining a metric that can be used to objectively compare two bullets.
 We propose a framework which allows for the automatic analysis of the surface topologies of bullets, and the transcription of the individual characteristics into a 2D plotting framework.
@@
 
-Throughout this paper, we work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
+We work with images from the James Hamby Consecutively Rifled Ruger Barrel Study~\citep{hamby:2009}.
 Ten consecutively rifled Ruger P-85 pistol barrels were obtained from the manufacturer and fired to produce 20 known test bullets and 15 unknown bullets for comparison.
@@
 It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in forensic science.
-An open-source package for working with x3p files has been written in R \citep{R} by Dr.
-Nicholas Petraco which we build upon in the development of our algorithm \citep{x3pr}.
-This package enables the direct reading and manipulation of x3p files.
+We have developed an open-source package for analyzing bullet lands written in R \citep{R}.
+This package is called bulletr \citep{bulletr} and enables direct reading and manipulation of x3p files.
+It also implements all of the methods we propose in this paper.
+A different package exists for reading x3p files called x3pr \citep{x3pr}, developed by Petraco (2014), but it is not designed to carry out calculations like the ones we propose after the x3p files have been read.
 
@@
 In the Hamby data, typical length (groove-to-groove) of a land is about \Sexpr{typical_width} micrometers or \Sexpr{typical_width / 1000} millimeters.
-For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
+For notational simplicity, we refer to a particular land of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
@@
 
-The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match.
+The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the strength of a match between two bullet lands.
 While we assess the performance of the algorithm in terms of a binary decision of match vs.
@@
 
-Our algorithm is fully open source and available on GitHub \citep{x3prplus}.
-This transparency allows for a greater understanding of the individual steps involved in the bullet matching process, and allows other forensic examiners, as well as outside observers, examine the factors that separate known bullet matches from non-matches.
-We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet.
+Our algorithm is fully open source and available on GitHub \citep{bulletr}.
+This transparency allows for a greater understanding of the individual steps involved in the bullet matching process, and allows other forensic examiners, as well as outside observers, to examine the factors that discriminate between known bullet matches and non-matches.
+We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet level.
 Although doing so introduces an implicit assumption of independence between lands, assuming independence only serves to make the task more challenging.
-The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
-% Hamby study: automated systems were successfully used for matching:
-%  Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
-% Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
-% Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
-% BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)
-
-Work has been done to quantify toolmark similarity by \citet{lock2013significance}.
-This algorithm determines an optimal matching window between two toolmark signatures, and then performs a set of both coordinated and independent shifts.
+
+The remainder of this paper is structured as follows: We first briefly review some earlier work.
+We then discuss two methods of modeling the class structure of the bullet surfaces.
+Finally, we proceed to describing an automatic matching routine which we evaluate on the bullets made available through the Hamby study.
+
+\section{Previous work}
+
+There have been attempts to develop automatic or semi-automatic matching protocols, but most have focused on breech face and firing pin marks \citep[e.g.][]{riva:2014} or discuss a single attribute for comparison \citep[e.g.][]{vorburger:2011, chu:2011}.
+Still others refer to proprietary algorithms \citep{roberge:2006}.
+We briefly review some of this earlier work in what follows.
+
+The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods.
+However, aside from a zero false positive rate, false-negative error rates for bullets are not given nor are error rates for land-to-land matches mentioned.
+
+\citet{lock2013significance} proposed an approach to quantify similarity of toolmarks.
+Their algorithm determines an optimal matching window between two toolmark signatures, and then performs a set of both coordinated and independent shifts.
 Given a match, the coordinated shifts would be expected to yield correlation values higher than those obtained from independent shifts.
 This is assessed using a Mann-Whitney U Statistic.
+
 A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}.
-This study is based on another set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score.
+Their study used a different set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score.
 The authors state that this process `could be automated', but no implementation of the algorithm is available.
 
-Modern automated techniques using 3D images have also been conducted by \citet{riva:2014}, but this work applies to striae found on cartridge cases and not bullets.
-This algorithm performs alignment of striae by rotation of the XY plane, which is not generalizable to bullets in which the XY plane is not flat.
-Other work on 3D images has been conducted by \citet{petraco:2012}, which also describes work carried out on cartridge cases, as well as screwdriver striation patterns, with promising results.
-We aim to use similar ideas of automated 3D image analysis, and apply this work to bullet surfaces with the modern tools enabled by the R statistical software.
-
-The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
-%We then describe a web application for visualizing and working with bullet surfaces.
-We then proceed to describing an automatic matching routine which we evaluate on the bullets made available through the Hamby study.
+Modern automated techniques using 3D images have also been proposed by \citet{riva:2014}.
+However, the authors focused on cartridge cases and not bullets.
+This might seem like a trivial distinction, but it has implications for the development of the algorithm.
+The algorithm performs alignment of striae by rotation of the XY plane, which is not generalizable to bullets in which the XY plane is not flat.
+
+Other work on 3D images has been described by \citet{petraco:2012}, who also focus on cartridge cases, as well as screwdriver striation patterns, and by others \citep[e.g.,][]{chu:2010, chu:2011, vorburger:2011}.
 
@@
 
-In an approach to analyze the striation pattern, we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land.
+To analyze the striation pattern, we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of a circular fit and the remaining residual structure can be found in Supplement section~\ref{supp:cylindrical}.
+A discussion of a circular fit and the remaining residual structure can be found in Supplement Section~\ref{supp:cylindrical}.
 %It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
@@
 %In an attempt to improve upon the previous algorithm, w
-We first have to identify the locations of the left and right groove in the image.
+We first identify the location of the left and right grooves in the image.
 The grooves are assumed to contain no information relevant for determining matches.
@@
 Fortunately, the location and appearance of the grooves in the surface profiles is quite consistent.
-Surface measurements reach local maxima around the peak of the groove at either end of the range of $y$, we can then follow the descent of the surface measurements inwards to the valley of the groove.
-The location of the valleys mark the points at which we should trim the image. The procedure can be described as follows:
+Surface measurements reach local maxima around the peak of the groove at either end of the range of $y$, and we can then follow the descent of the surface measurements inwards to the valley of the groove.
+The location of the valleys mark the points at which we trim the image. The procedure can be described as follows:
 
@@
 The smoothing factor $s$ introduced in the algorithm represents the window size to use for a rolling average.
-Higher values of $s$ therefore lead to more smoothed results.
-In the groove detection portion of this algorithm, we wish to remove most of the deviation but still maintain enough signal in the groove to detect it.
-Empirically, a value of $s = 35$ for the smoothing factor seems to work quite well (the smoothing factor is further discussed in Section~\ref{sec:smoothing}).
+Higher values of $s$ therefore lead to more smoothing.
+Empirically, a value of $s = 35$ for the smoothing factor seems to work well (the smoothing factor is further discussed in Section~\ref{sec:smoothing}).
 It is important to note that the smoothing pass is done twice.
-That is, the smoothed data is once again smoothed by computing a new rolling average with the same smoothing factor.
+That is, the smoothed data are once again smoothed by computing a new rolling average with the same smoothing factor.
 This bears some similarities to the ideas of John Tukey in his book Exploratory Data Analysis, where he describes a smoothing process called ``twicing" in which a second pass is made on the residuals computed from the first pass and then added back to the result~\citep{tukey:1977}.
@@
 
-The main idea of locally weighted regression is to use a weighting routine that  emphasizes the effect of points close by and de-emphasizes the effect of points as they are further away.
+The main idea of locally weighted regression is to use a weighting routine that  emphasizes the effect of points close to the fitting location and de-emphasizes the effect of points as they are further away.
 The weighting function used here is the tricubic function $w(d) = \left(1 - d^3\right)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
@@
 Visually, this allows for an easy assessment of these two bullet lands as a match.
-However, this match relies on visual inspection and can therefore not be completely objective.
+However, this match relies on visual inspection and is therefore subjective.
 The goal of this section is to eliminate the need for a visual inspection during the matching process and replace it by an automatic algorithm.
-This also allows for a quantification of the quality of the match.
+This also allows for a quantification of the strength of the match.
 
@@
 Figure~\ref{fig:match} gives an overview of the automated matching routine: 
-We first identify a stable region for each bullet land and extract the signature at the lowest height in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement section~\ref{supp:bulletbottom} for a more detailed discussion).
+We first identify a stable region for each bullet land and extract the signature at the lowest height in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see Supplement Section~\ref{supp:bulletbottom} for a more detailed discussion).
 
@@
 We then define an interval around the location of the extrema on each side as one third of the distance to the location of the next extrema (see Figure~\ref{fig:smoothcutb}).
-Peaks and valleys constitute the \emph{striation marks} of the bullet.
-\item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify joint intervals between two signatures as the areas in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals.
+Peaks and valleys constitute the \emph{striation marks} on the bullet.
+\item {\bf Match striations across signatures:} based on the intervals around the extrema as defined above, we identify common intervals as the areas in which two or more of the individual intervals overlap: a joint interval is defined as the smallest interval that encompasses all of the overlapping intervals.
 A joint interval is then called a match(ing stria) between the signatures, if all of the intervals are of the same type of extrema, i.e.\ they are either all peaks or all valleys.
@@
 Non-matching intervals are left grey.
-\item {\bf Extract features from the aligned signatures and the matches between them:} many different features could be extracted.
+\item {\bf Extract features from the aligned signatures and the matches between them:} many different features can be extracted from the aligned signatures.
 Here, we describe a few of the ones that can be found in the literature and some that we found to be of practical relevance:
@@
 \item Number of matches and non-matches,
-\item The value of the cross-correlation function (ccf) between the aligned signatures \cite{vorburger:2011},
+\item The value of the cross-correlation function (ccf) between the aligned signatures \citep{vorburger:2011},
 \item Average difference $D$ between signatures, defined as the Euclidean vertical distance between surface measurements of aligned signatures. Let $f(t)$ and $g(t)$ be smoothed, aligned signatures:
@@
 Counting the maximal number of CMS is part of the current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
-In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a high number indicative of a match between the bullets. 
- Note that the definition of CMS given in this paper does not match the one given in \citet{thompson:2013}.
+In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a high number suggestive of a match between the lands.
+Note that the definition of CMS we use does not match the one given in \citet{thompson:2013}.
 There, CMS is defined only in terms of matching peaks without regarding valleys.
@@
 Striation marks show up in these representations as peaks and valleys.
-The individual characteristics are quite prominent and, again, suggest a match between the lands.
-A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities.
+The individual characteristics are prominent and, again, suggest a match between the lands.
+A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities between signatures.
 %
@@
 %
-For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures~\cite{bachrach:2002, chu:2010, vorburger:2011, thompson:2013}.
+For this alignment we use the cross-correlation function to find a maximal amount of agreement between the signatures \citep{bachrach:2002, chu:2010, vorburger:2011, thompson:2013}.
 This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
@@
 This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of the height that was used in a matching.
-Visually, matches degrade if the signatures upon which the match is based are from heights further than 200$\mu m$ apart  (see Supplement section~\ref{supp:ccf} for a more thorough discussion).
+Visually, matches degrade if the signatures upon which the match is based are from heights further than 200$\mu m$ apart  (see Supplement Section~\ref{supp:ccf} for more discussion).
 However, we can extract signatures from multiple heights of the same bullet land for an initial assessment of its quality.
@@
 Given that striation widths are typically much larger, we are in effect muddying the waters by performing such minimal smoothing.
-Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- in order to not introduce artifacts in the analysis.
+Another consideration is that the smoothing should not fall below the  resolution of the equipment at which the surface measurements are taken -- so as to not introduce artifacts in the analysis.
 
@@
 \section{Evaluation}
-In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
+In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance using the James Hamby study data.
 As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
@@
 Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
- When things go well, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 lands from known bullets.
+Ideally, results look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 lands from known bullets.
 Two lands show a high CMS.
@@
 For smaller numbers of CMS, the percentage of false positives quickly increases.
-However, if we take other features of the image into account, we can increase the number of correct matches considerably: Figure~\ref{fig:densities} gives an overview of the densities of all of the features derived earlier, distinguishing between known matches (KM) and known non-matches (KNM).
+However, if we take other features of the image into account, we can increase the number of correct matches considerably: Figure~\ref{fig:densities} gives an overview of the densities of all of the features derived earlier, for known matches (KM) and known non-matches (KNM).
 The densities of almost all of the features show strong differences between matches and non matches.
@@
 The smaller the value, the better.
-We see that in this instance a low EER goes hand in hand with high predictive power as measured in AUC.
+We see that in this instance a low equal error rate (EER) goes hand in hand with high predictive power as measured in AUC.
 The feature with the highest individual predictive power is $S$, the sum of the average heights of two signatures at peaks and valleys.
@@
 This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).
-Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches.
-In the case of the Hamby study, even for the single tree the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched.
+Note that this error rate is based on land-to-land comparisons and is expected to be much lower for bullet-to-bullet comparisons.
+In the case of the Hamby data, even a single tree results in an overall error rate of zero, if we require that a match of two bullets occurs when at least two of the bullet's lands are matched.
 % (see Figure~\ref{fig:treeresult}).
@@
 This makes the errors in the automated approach smaller than the human error in the Hamby study.
-Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, corresponding to a rate of 0.0011.
+Out of the 507 participants who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, corresponding to a rate of 0.0011.
 %\hh{XXX calling this a false negative rate is problematic.
@@
 
-However, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets.
+For the Hamby data, error rates based on bullet-to-bullet matches do not carry a lot of weight because of the small size of the study: fifteen unknown bullets are successfully matched to two pairs of ten bullets.
 Matching bullets can only be tested realistically in a much bigger experiment.
 Another thing to note about  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared.
-Basing this decision on 0.5 is not necessary.
+Basing this decision on a threshold fixed at 0.5 may not be the best approach.
 In practice, examiners are allowed a third option of `inconclusive'.
@@
 Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree and the random forest.
+As expected, the random forest provides a more realistic estimate of the uncertainty in the classification.
 %\hh{The tree predicts 132 known matches correctly, with 40 false negatives and one false positive match.
@@
 Besides resulting in a probabilistic quantification of matches, random forests also provide an assessment of the importance of each of the features derived from the bullets' 3D topological surface measurements.
-Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to Supplement section~\ref{supp:randomforest}).
+Figure~\ref{fig:importance} shows an overview of the importance of each variable measured as the mean decrease in the Gini index when the variable in question is included in a tree (for the exact values please refer to Supplement Section~\ref{supp:randomforest}).
 \begin{figure}
@@
 %The rows in the table are ordered according to descending importance. 
-The variables with the most predictive power are cross-correlation and the overall number of matches, followed by the total depth of joint striations $S$ and total number of non-matches.
+The variables with the most predictive power are cross-correlation and the overall number of matching extrema, followed by the total depth of joint striations $S$ and total number of non-matches.
 CMS is found only in sixth place.
@@
 Comparisons not previously included in fitting the random forest can also be used as an additional source for assessing error rates.
-Results for this and a more detailed discussion can be found in Supplement section~\ref{supp:extended}.
-
-\section{Conclusion}
-
-We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a firearms identification perspective, removes these features, and produces residuals which allow for the easy identification of markings.
-We have generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches.
-We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features.
-Matching bullets is clearly not a single step, but rather a sequence of a number of data analysis tasks which each deserve attention.
-As there is no scientific standard in place at this point in time, our intent is to explain one approach to addressing these tasks, while documenting all steps and providing all code so other researchers and forensic scientists can reproduce and expand on our findings.
+Results for this and a more detailed discussion can be found in Supplement Section~\ref{supp:extended}.
+
+\section{Discussion}
+
+We present an algorithm which detects the most prominent but least relevant structure of a bullet from a firearms identification perspective, removes these features, and produces residuals which allow for the easy identification of markings.
+We have generalized this algorithm to align the residuals from two bullets to automatically determine whether they are indistinguishable.
+A random forest model provides a probabilistic assessment of the strength of a match, along with an ordering of the relevance of features.
+Matching bullets is clearly not a one-step process, but rather a sequence of data analysis tasks each deserving attention.
+As there is no scientific standard in place at this point in time, our intent is to explain an approach to addressing these tasks, while documenting all steps and providing all code so other researchers and forensic scientists can reproduce and expand on our findings.
 
@@
 Another parameter that must be selected is the amount of smoothing when identifying peaks and grooves (currently, a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
-While we tried to lay out in the paper the impact that each of the parameter choices has on the matching performance, we are still far from an optimized scenario.
+We try to lay out in the paper the impact that each of the parameter choices has on the matching performance, but more research and better data are needed to define an optimized scenario.
 
 The Hamby study serves as our evaluation `database'.
-It consists of only  35 bullets -- this is obviously not a particularly realistic scenario for an automatic matching procedure, but for now we are unaware of other databases containing bullets in the x3p format.
+It consists of only  35 bullets -- this is obviously not a particularly realistic scenario for an automatic matching procedure, but for now we are unaware of other databases containing bullets in the x3p format that we could add to our study.
 
 The feasibility of creating a database of images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
-The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
+The committee investigated the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
 The bottom line of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found.
-The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
+The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records without overwhelming forensic examiners, who have to examine possible matches suggested by the system.
 The findings of the NRC report on imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks.
@@
 The steps outlined in this paper could also help explore other important forensic science problems.
-In particular, more general toolmark examination can benefit from this detection and removal of the overall structure and the extraction of features from deviations of this structure.
+In particular, more general toolmark examination can benefit from the approach we discuss.
 
 For a fair assessment of the performance of an algorithm, we need transparency.
-Our matching  algorithm is open: the code is readily available in form of the R package x3prplus~\citep{x3prplus}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
+Our matching  algorithm is open: the code is readily available in form of the R package bulletr~\citep{bulletr}, and the code to produce this paper is available at \url{http://www.github.com/erichare/imaging-paper}.
 To understand whether an automated approach along the lines of the one we propose can accurately identify sets of bullets with undistinguishable markings, it will be necessary to assemble a much larger database that includes a wide range of ammunition types, degrees of damage, gun makes, etc.

### 2017-01-12 imaging-paper.tex
Rebuild both papers with no cache
Paper edits from Dr. Carriquiry

--- 2016-10-26 imaging-paper.tex
@@
 Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
-In the Hamby data, typical length (groove-to-groove) of a land is about 2005.9 micrometers or 2.01 millimeters.
-For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
+In the Hamby data, typical length (groove-to-groove) of a land is about 1998.28 micrometers or 2 millimeters.
+For notational simplicity, we refer to a particular land of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
 An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.

### 2017-01-12 revision/response2.Rmd
More updates
Reviewer response with Alicia's edits

@@
 author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
-date: "October 24, 2016"
+date: "January 12, 2017"
 output:
@@
     **Most of the references that were suggested by reviewers earlier were incorporated into the revised version of the manuscript.
-Regarding the "overlap" with Petraco's work, please see the attached document entitled x3prplus for an in-depth discussion of our use of exactly one function in Petraco's package.**
+Regarding the "overlap" with Petraco's work, please see the attached document entitled x3prplus for an in-depth discussion of our use of exactly one function in Petraco's package.
+To avoid further confusion, we are no longer using any of Petraco's functions and have instead published a 100% non-overlapping package called bulletr.**
 
 2.
-As I read the paper it is unclear how the matching criteria is different than the matching criteria currently in use by the tool mark’s community, especially those proposed by Monkres et al (2013) in Comparison and Statistical Analysis of Land
-Impressions from Consecutively Rifled Barrels and the PhD dissertation work published in Technometrics by Amy B.
+As I read the paper it is unclear how the matching criteria is different than the matching criteria currently in use by the tool mark’s community, especially those proposed by Monkres et al (2013) in Comparison and Statistical Analysis of Land Impressions from Consecutively Rifled Barrels and the PhD dissertation work published in Technometrics by Amy B.
 Hoeksema with her PhD advisor Max Morris in 2013.
@@
 Hoeksema’s dissertation also has a wonderful explanation of Chumbley’s U-statistic, which appears to be similar to the author’s proposed statistic, which is a PhD dissertation in the department of the authors of the paper currently under review.
-My opinion is that a discussion and comparison of these classical methods with their new comparison
-statistic is necessary to demonstrate how the author’s work advances the science of land impression comparisons.
+My opinion is that a discussion and comparison of these classical methods with their new comparison statistic is necessary to demonstrate how the author’s work advances the science of land impression comparisons.
 
@@
 As the reviewer notes, this work was carried out in our own department, and we were well aware of it.
-Thus our surprise with the review, since to say that our work is a repeat of Amy Hoeksema's work is so inaccurate as to make us wonder whether the reviewer is familiar with either approaches.
+Thus our surprise with the review, since to say that our work is a repeat of Amy Hoeksema's work is so inaccurate as to make us wonder whether the reviewer understood what we propose to do.
 On October 11th, 2016, we sat down with Max Morris to discuss the similarities and differences between the Hoeksema U-statistic, and the methods used and features derived in our paper.
@@
 Everything else is completely different, as can be seen in the attached document.
-Perhaps our choice of name for the package has contributed to the confusion.**
+Perhaps our choice of name for the package has contributed to the confusion.
+Therefore, we have written our own function to read x3p files and have renamed our new package bulletr.
+bulletr is completely separate from x3pr and has many more capabilities.**
 
@@
 We have referred to many of those papers in the manuscript and have shown how our approach is quite different.
-Indeed, even PCAST (the President's Council of Advisors in Science and Technology) in its report, mentioned that our work was the only promising work and they actually looked at all the literature themselves.
+Indeed, even PCAST (the President's Council of Advisors in Science and Technology) in its report, mentioned that our work was the only promising approach and they actually looked at all the literature themselves.
 So we respectfully disagree with Reviewer 2.
@@
 The reviewer would be well advised to read our comment more carefully;  the reason it is important is that Riva and Champod make use of the circular structure of their data to pre-process their images and develop their score.
-A big component of their algorithm is the alignment of images, which they do by rotating the images on the xY plane.
-This is not something that cannot be extended to the case of bullets, were the XY plane is not flat and where rotation makes no sense.
+A big component of their algorithm is the alignment of images, which they do by rotating the images on the XY plane.
+This is not something that can be extended to the case of bullets, were the XY plane is not flat and where rotation makes no sense.
 The approach they use to extract features from the aligned images is completely different from ours and for the reviewer to say that we are re-inventing the wheel is a bit of a scandal.
@@
 See the below citation, given by `x3pr` as of October 17th, 2016:**
+
     ```{r}
@@
     ```
+
+    **However, in the end we decided to go a step further and completely remove the dependency on Petraco's code. Right now, we are not using *any* of Petraco's code in any of our functions.**
 
@@
 While the dataset is small (only 10 barrels) we have been told by multiple firearm examiners that it is a challenging dataset on which to develop methods because the 10 barrels were consecutively manufactured.
-Thus, one would expect that thre striae made by those barrels will be more similar across them than across two randomly selected barrels from a large population.
+Thus, one would expect that the striae made by those barrels will be more similar across them than across two randomly selected barrels from a large population.
 We are collaborating with Alan Zhang in NIST to assemble a larger database that will be publicly available within a year or two.**
@@
     **Did the reviewer read the correct paper?
-The revised copy of our paper contains over thirty references, not just one.
-Many are from previous statistical studies on bullet matching (e.g., Biasotti 1959, Chu 2010, Riva 2014) while many more are references on statisitcal tests themselves (e.g., Therneau 2015, Liaw 2002, Cleveland 1979)**
+The copy of our paper that we sent out for revision contains over thirty references, not just one.
+Many are from previous statistical studies on bullet matching (e.g., Biasotti 1959, Chu 2010, Riva 2014) while many more are references on statistical tests themselves (e.g., Therneau 2015, Liaw 2002, Cleveland 1979)**
 
@@
 
-# x3prplus
+# bulletr - January 2017
+
+The three reviewers that were brought in for the second round of reviews based many of their criticisms on our allged "lifting" of Petraco's code.
+Nothing could be further from reality.
+Our original package x3prplus only made use of a single function in Petraco's x3pr package, that was used to read in the 3D data files.
+The rest of our package shares absolutely nothing else with Petraco's as explained below.
+The differences are so clear, that perhaps one reviewer could have been confused; but to have three reviewers make the exact same completely inaccurate accusation seems more like a coordinated effort to undermine our work.
+
+Regardless, we have decided that we do not wish to rely on Petraco's package at all.
+Toward this end, we have written our own function to read in 3D files (which does not have any similarity with Petraco's read.x3p function) and have renamed our new package bulletr.
+In this way, we no longer share a single line of code with Petraco.
+
+# x3prplus - October 2016
 

### 2017-02-02 arxiv/imaging-paper.tex
Remove line numbers
One more linenumber removal
Arxiv update and AoAS resubmission

@@
 
-\linenumbers
 %\tableofcontents
@@
 
-The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match.
-Our algorithm is fully open source and available on GitHub \citep{x3prplus}.
-This transparency allows for a greater understanding of the individual steps involved in the bullet matching process, and allows other forensic examiners, as well as outside observers, examine the factors that separate known bullet matches from non-matches.
-We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet.
+The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the strength of a match between two bullet lands.
+While we assess the performance of the algorithm in terms of a binary decision of match vs.
+non-match using a 50\% probability cut-off, our primary goal is to highlight the features that are statistically associated with matches and non-matches, and to provide a quantitative assessment of this association.
+In a real-world application of our algorithm, the raw scores would need further analysis and scrutiny, and it is likely a 50\% cut-off would be an inappropriate choice on the basis of reasonable doubt.
+
+Our algorithm is fully open source and available on GitHub \citep{bulletr}.
+This transparency allows for a greater understanding of the individual steps involved in the bullet matching process, and allows other forensic examiners, as well as outside observers, to examine the factors that discriminate between known bullet matches and non-matches.
+We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet level.
 Although doing so introduces an implicit assumption of independence between lands, assuming independence only serves to make the task more challenging.
-The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned.
-% Hamby study: automated systems were successfully used for matching:
-%  Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
-% Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
-% Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
-% BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)
+
+The remainder of this paper is structured as follows: We first briefly review some earlier work.
+We then discuss two methods of modeling the class structure of the bullet surfaces.
+Finally, we proceed to describing an automatic matching routine which we evaluate on the bullets made available through the Hamby study.
+
+\section{Previous work}
+
+There have been attempts to develop automatic or semi-automatic matching protocols, but most have focused on breech face and firing pin marks \citep[e.g.][]{riva:2014} or discuss a single attribute for comparison \citep[e.g.][]{vorburger:2011, chu:2011}.
+Still others refer to proprietary algorithms \citep{roberge:2006}.
+We briefly review some of this earlier work in what follows.
+
+The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods.
+However, aside from a zero false positive rate, false-negative error rates for bullets are not given nor are error rates for land-to-land matches mentioned.
+
+\citet{lock2013significance} proposed an approach to quantify similarity of toolmarks.
+Their algorithm determines an optimal matching window between two toolmark signatures, and then performs a set of both coordinated and independent shifts.
+Given a match, the coordinated shifts would be expected to yield correlation values higher than those obtained from independent shifts.
+This is assessed using a Mann-Whitney U Statistic.
 
 A procedure for bullet matching using the BulletTrax3D system is described in \citet{roberge:2006}.
-This study is based on another set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score.
+Their study used a different set of ten consecutively rifled barrels; matches are identified based on a bullet-to-bullet correlation score.
 The authors state that this process `could be automated', but no implementation of the algorithm is available.
-Modern automated techniques using 3D images have also been conducted by \citet{riva:2014}, but this work applies to striae found on cartridge cases and not bullets.
-Other work on 3D images has been conducted by \citet{petraco:2012}, which also describes work carried out on cartridge cases, as well as screwdriver striation patterns, with promising results.
-We aim to use similar ideas of automated 3D image analysis, and apply this work to bullet surfaces with the modern tools enabled by the R statistical software.
-
-The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces.
-%We then describe a web application for visualizing and working with bullet surfaces.
-We then proceed to describing an automatic matching routine which we evaluate on the bullets made available through the Hamby study.
+
+Modern automated techniques using 3D images have also been proposed by \citet{riva:2014}.
+However, the authors focused on cartridge cases and not bullets.
+This might seem like a trivial distinction, but it has implications for the development of the algorithm.
+The algorithm performs alignment of striae by rotation of the XY plane, which is not generalizable to bullets in which the XY plane is not flat.
+
+Other work on 3D images has been described by \citet{petraco:2012}, who also focus on cartridge cases, as well as screwdriver striation patterns, and by others \citep[e.g.,][]{chu:2010, chu:2011, vorburger:2011}.
 
@@
 Another parameter that must be selected is the amount of smoothing when identifying peaks and grooves (currently, a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
-While we tried to lay out in the paper the impact that each of the parameter choices has on the matching performance, we are still far from an optimized scenario.
+We try to lay out in the paper the impact that each of the parameter choices has on the matching performance, but more research and better data are needed to define an optimized scenario.
 
 The Hamby study serves as our evaluation `database'.
-It consists of only  35 bullets -- this is obviously not a particularly realistic scenario for an automatic matching procedure, but for now we are unaware of other databases containing bullets in the x3p format.
-
-The feasibility of creating a database of ballistic images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
-The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
+It consists of only  35 bullets -- this is obviously not a particularly realistic scenario for an automatic matching procedure, but for now we are unaware of other databases containing bullets in the x3p format that we could add to our study.
+
+The feasibility of creating a database of images that could be used to identify guns used in crimes was evaluated in a 2008 report~\citep{nap:2008} by the National Research Council.
+The committee investigated the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS.
 The bottom line of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found.
-The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system.
-The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks.
+The problem that remained is that statistically, the quality of the matching algorithm (in this case, of breech-face marks and firing pin impressions) could not withstand a hugely increased number of records without overwhelming forensic examiners, who have to examine possible matches suggested by the system.
+The findings of the NRC report on imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing between fine marks.
 This finding coincides with the assessment by \citet{dekinder:2004} based on the IBIS Heritage system.

### 2017-02-09 code/full_run.R
Update to use new csafe server

@@ -19,7 +19,7 @@ library(stringr)
 dbname <- "bullets"
 user <- "buser"
 password <- readLines("buser_pass.txt")
-host <- "50.81.214.252"
+host <- "10.25.122.176"
 
 con <- dbConnect(MySQL(), user = user, password = password,
                  dbname = dbname, host = host)
@@ -316,7 +316,7 @@ ccf <- as.data.frame(do.call(rbind, ccf_temp)) %>%
 dbWriteTable(con, "ccf", ccf, row.names = FALSE, append = TRUE)
 
 my_matches <- dbReadTable(con, "matches")
-CCFs_withlands <- ccf %>%
+CCFs_withlands <- dbReadTable(con, "ccf") %>%
     left_join(select(profiles, profile_id, land_id), by = c("profile1_id" = "profile_id")) %>%
     left_join(select(profiles, profile_id, land_id), by = c("profile2_id" = "profile_id")) %>%
     left_join(my_matches, by = c("land_id.x" = "land1_id", "land_id.y" = "land2_id")) %>%

### 2017-03-23 code/bullet-app.R
Fix the bullets backend app

@@ -8,52 +8,55 @@ library(plotly)
 
 dbname <- "bullets"
 user <- "buser"
-password <- "TihrcdTJn#7pWv"
+password <- readLines("buser_pass.txt")
 host <- "127.0.0.1"
 
 con <- dbConnect(MySQL(), user = user, password = password,
                  dbname = dbname, host = host)
 
-bullet_choices <- dbGetQuery(con, "SELECT id,name FROM metadata")
-bullet_choices_vec <- bullet_choices$id
-names(bullet_choices_vec) <- bullet_choices$name
+bullet_choices <- dbGetQuery(con, "SELECT land_id,study,barrel,bullet,land FROM metadata") %>%
+    mutate(name = paste0(study, " Br", barrel, "_B", bullet, "_L", land))
+
+bullids <- bullet_choices$land_id
+names(bullids) <- bullet_choices$name
 
 ui <- fluidPage(
-   
-   titlePanel("Bullet Database"),
-   
-   sidebarLayout(
-      sidebarPanel(
-          selectInput("land", "Bullet Land", choices = bullet_choices_vec),
-          numericInput("crosscut", "Crosscut", min = 0, max = 500, value = 0)
-      ),
-      
-      mainPanel(
-          tableOutput("metadata"),
-          tableOutput("profile_metadata"),
-          plotOutput("profile"),
-          plotOutput("signature")
-          #plotlyOutput("viewland")
-      )
-   )
+    
+    titlePanel("Bullet Database"),
+    
+    sidebarLayout(
+        sidebarPanel(
+            selectInput("land", "Bullet Land", choices = bullids),
+            numericInput("crosscut", "Crosscut", min = 0, max = 500, value = 75),
+            numericInput("run_id", "Run ID", value = 3)
+        ),
+        
+        mainPanel(
+            tableOutput("metadata"),
+            tableOutput("profile_metadata"),
+            plotOutput("profile"),
+            plotOutput("signature")
+            #plotlyOutput("viewland")
+        )
+    )
 )
 
 server <- function(input, output, session) {
     bullet_metadata <- reactive({
-        result <- dbGetQuery(con, paste0("SELECT * FROM metadata WHERE id = ", input$land, ""))
+        result <- dbGetQuery(con, paste0("SELECT * FROM metadata WHERE land_id = ", input$land, ""))
         updateNumericInput(session, "crosscut", max = result$num_profiles[1])
         
         return(result)
     })
     
     observe({
[…]

### 2017-06-18 imaging-paper.Rnw
Paper updates based on review

@@
 This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels.
-The automatic matching routine we propose manages to (a) correctly identify lands (the surface between two bullet grooves) with too much damage to be suitable for comparison, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
+The automatic matching routine we propose manages to (a) correctly identify land impressions (the surface between two bullet groove impressions) with too much damage to be suitable for comparison, and (b) correctly identify all 10,384 land-to-land matches of the James Hamby study \citep{hamby:2009}.
 \end{abstract}
@@
 Here, we focus on the question of defining a metric that can be used to objectively compare two bullets.
-We propose a framework which allows for the automatic analysis of the surface topologies of bullets, and the transcription of the individual characteristics into a 2D plotting framework.
+We propose a framework which allows for the automatic analysis of the surface topologies of bullets, and the transcription of the individual characteristics into a set of features that quantify their similarities.
 This allows for an objective and quantitative assessment of striae-based bullet matches.
@@
 \caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
+\label{fig:allx3p}
 \end{figure}
@@
 
-Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
-In the Hamby data, typical length (groove-to-groove) of a land is about \Sexpr{typical_width} micrometers or \Sexpr{typical_width / 1000} millimeters.
+Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent groove impressions on the bullet, called a ``land" or land impression.
+The shoulders are the raised portion of the surface between the groove impression and the land impression.
+In the Hamby data, typical length (shoulder-to-shoulder) of a land is about \Sexpr{typical_width} micrometers or \Sexpr{typical_width / 1000} millimeters.
 For notational simplicity, we refer to a particular land of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
@@
 
-To analyze the striation pattern, we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land.
+To analyze the striation pattern, we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land, as illustrated in Figure \ref{fig:allx3p}.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
@@
 It can be seen that the global structure of the land dominates the appearance of the plot.
-The grooves can be clearly identified on the left and right side, and the curvature of the surface is the most visible feature in the middle.
+The shoulders can be clearly identified on the left and right side, and the curvature of the surface is the most visible feature in the middle.
 
@@
 
-\subsection{Identifying groove locations}
+\subsection{Identifying shoulder locations}
 %In an attempt to improve upon the previous algorithm, w
-We first identify the location of the left and right grooves in the image.
-The grooves are assumed to contain no information relevant for determining matches.
-They also dominate the structure, and therefore need to be removed.
-Fortunately, the location and appearance of the grooves in the surface profiles is quite consistent.
-Surface measurements reach local maxima around the peak of the groove at either end of the range of $y$, and we can then follow the descent of the surface measurements inwards to the valley of the groove.
+We first identify the location of the left and right shoulders in the image.
+The groove impressions are assumed to contain no information relevant for determining matches, and the shoulders identify the location at which the land impression begins.
+The shoulders also dominate the structure, and therefore need to be removed.
+Fortunately, the location and appearance of the shoulders in the surface profiles is quite consistent.
+Surface measurements reach local maxima around the peak of the shoulder at either end of the range of $y$, and we can then follow the descent of the surface measurements inwards to the valley of the shoulder.
 The location of the valleys mark the points at which we trim the image. The procedure can be described as follows:
@@
     \item For each $y$ value, smooth out any deviations occurring near the minima by twice applying a rolling average with a pre-set \emph{smoothing factor} $s$.
-(Figure~\ref{fig:loess_step3}, smoothing factor $s = 35$ corresponding to 55$\mu m$).
+(Figure~\ref{fig:loess_step3}, smoothing factor $s = 35$(data points) corresponding to 55$\mu m$).
 %    \item For each smoothed $y$ value, compute another rolling average using the same smoothing factor as above. (Figure~\ref{fig:loess_step3}).
-    \item Determine the location of the peak of the left groove by finding the first doubly-smoothed value $y_i$ that is the maximum within its smoothing window (e.g.\ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$, where $i$ is between 1  and $\lfloor s/2 \rfloor$).
+    \item Determine the location of the peak of the left shoulder by finding the first doubly-smoothed value $y_i$ that is the maximum within its smoothing window (e.g.\ such that $y_i > y_{i - 1}$ and $y_i > y_{i + 1}$, where $i$ is between 1  and $\lfloor s/2 \rfloor$).
 We call the location of this peak $p_{\ell}$ (see Figure~\ref{fig:loess_step47}).
-    \item Similarly, determine the location of the valley of the left groove by finding the first double-smoothed $y_j$ that is the minimum within its smoothing window.
+    \item Similarly, determine the location of the valley of the left shoulder by finding the first double-smoothed $y_j$ that is the minimum within its smoothing window.
 Call the location of this valley $v_{\ell}$.
-    \item Reverse the order of the $y$ values and repeat the previous two steps to find the peak and valley of the right groove, $(p_{r}, v_{r})$.
-    \item Trim the surface measurements to values within the two grooves (i.e.\ remove all records with $y_i < v_{\ell}$ and $y_i > v_{r}$) (see Figure~\ref{fig:loess_step47}).
+    \item Reverse the order of the $y$ values and repeat the previous two steps to find the peak and valley of the right shoulder, $(p_{r}, v_{r})$.
+    \item Trim the surface measurements to values within the two shoulder valleys (i.e.\ remove all records with $y_i < v_{\ell}$ and $y_i > v_{r}$) (see Figure~\ref{fig:loess_step47}).
 \end{enumerate}
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
+\caption{\label{fig:loess_step1}Step 1 of identifying shoulder locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
 <<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='.5\\textwidth', warning=FALSE,  message=FALSE>>=
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step3}Step 2 of identifying groove locations: The  surface measurements are smoothed twice with a smoothing factor of $s = 35$.
+\caption{\label{fig:loess_step3}Step 2 of identifying shoulder locations: The  surface measurements are smoothed twice with a smoothing factor of $s = 35$.
 The orange rectangle shows an example of the smoothing window.
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step47}Steps 3 -- 6 of identifying groove locations: After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
+\caption{\label{fig:loess_step47}Steps 3 -- 6 of identifying shoulder locations: After smoothing the surface measurements extrema on the left and right are detected (marked by vertical lines, red indicating peaks and blue indicating valleys).
 Values outside the blue boundaries are removed (shown in grey)}
@@
 \end{subfigure}
-\caption{Overview of all six steps of the smoothing algorithm to identify and remove grooves from the bullet images.}
+\caption{Overview of all six steps of the smoothing algorithm to identify and remove shoulders and groove impressions from the bullet images.}
 \end{figure}
@@
 The weighting function used here is the tricubic function $w(d) = \left(1 - d^3\right)^3$, for $d \in [0,1]$ and $w(d) = 0$ otherwise.
-Here, $d$ is defined as the distance between $x_i$ and the location of the fit $x_o$ and the maximum distance of the range of the $x$-values for span $\alpha$ in $x_o$.
+Here, $d$ is defined as the distance between $x_i$ and the location of the fit $x_o$, divided by the overall range of the included data, so as to map the results to a $[0,1]$ range as the definition requires.
 
@@
 #subLOFx2 <- subset(lof, bullet=="Br1 Bullet 2-1")
-subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
+#subLOFx1$y <- subLOFx1$y + 23*1.5625 # working now!!!
 lofX <- rbind(data.frame(subLOFx1), data.frame(subLOFx2))
@@
 \item Use cross-correlation to {\bf find the best alignment} of the two signatures: shift one of the signatures by the lag indicated by the cross-correlation function (see Figure~\ref{fig:ccf} for the cross-correlation function and Figure~\ref{fig:crosscutX} for the resulting shift).
-\item Using a rolling average, {\bf identify peaks and valleys} for each of the signatures.
+\item Using a rolling average, {\bf identify peaks and valleys} for each of the signatures by identifying points at which the derivative of the signature is equal to zero.
 We then define an interval around the location of the extrema on each side as one third of the distance to the location of the next extrema (see Figure~\ref{fig:smoothcutb}).
@@
 \caption{\label{fig:ccf}Cross-correlation function between the two signatures shown in Figure~\ref{fig:crosscut} at lags between -100 and 100.
-The correlation is maximized at a lag of -17, indicating the largest amount of agreement between the signatures.
-Figure~\ref{fig:crosscutX} shows the lag-shifted signatures.}
+The correlation is maximized at a lag of -2, indicating the largest amount of agreement between the signatures.
+Figure~\ref{fig:crosscutX} shows the slight change resulting from the lag-shifted signatures.}
 \end{figure}
@@
 A smoothing factor of $k$ means that the  $k$ closest observations to $x_o$ are considered for a fit for $x_o$.
-Because surface measurements are recorded at an equidistant resolution (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$.
-For detecting and removing the grooves prior to fitting a loess regression we selected a smoothing factor  of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
+Because surface measurements are recorded at equal sampling intervals (here, of 1.5625$\mu m$), we decided to only consider odd smoothing factors $2k + 1$, which means that the  $k$ observations to the left and right of $x_o$ are considered for a local fit of $x_o$.
+For detecting and removing the shoulders prior to fitting a loess regression we selected a smoothing factor of 35, while for detecting the peaks/valleys of the loess residuals a smoothing factor of 25 seems more appropriate.
 %We can investigate the performance of peak/valley detection across different smoothing factors.
@@
 Four lands from different bullets are flagged as problematic in this respect.
-A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands has scratch marks across the surface, also known as `tank rash'.
+A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands has scratch marks across the surface, also known as `tank rash' \citep{hamby:2009}.
 %
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{Maximal number of CMS between unknown bullet C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown bullet land C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
@@
 labels <- unique(subset(rocs, auc > 0.5)[,c("variable", "auc")])
-labels$labels <- sprintf("AUC: %.2f", labels$auc)
+labels$lbls <- sprintf("AUC: %.2f", labels$auc)
 labels$type <- NA
@@
 ) %>% filter(minerror==TRUE)
+eer <- eer %>% left_join(select(labels, variable, auc_lbl = lbls), by = c("variable" = "variable"))
 
@@
   ylab("False positive rate") + facet_wrap(facets=~varLabel, ncol=4) + 
-  geom_label(x=1, y=0, aes(label=labels), data=labels, hjust=1, vjust=0) + 
+  geom_label(x=1, y=0, aes(label=auc_lbl), data=eer, hjust=1, vjust=0) + 
   geom_abline(linetype=2) + 
@@
 Using a cut-off probability of 0.5 the forest correctly predicts every single comparison.
-Compared to the tree, the forest's prediction probabilities are  shrunk towards either end of the prediction range.
-}
+Compared to the tree, the forest's prediction probabilities are  shrunk towards either end of the prediction range.}
 \end{figure}
@@
 The heights at which signatures are extracted (currently 25$\mu m$ apart) to evaluate stability, as well as the cross-correlation factor (currently 0.95) we set as a minimum threshold do affect the final outcome.
-Another parameter that must be selected is the amount of smoothing when identifying peaks and grooves (currently, a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
+Another parameter that must be selected is the amount of smoothing when identifying peaks and valleys (currently, a window of 23.4375$\mu m$ is used, corresponding to a window of 7 values to the left and the right of an observation).
 We try to lay out in the paper the impact that each of the parameter choices has on the matching performance, but more research and better data are needed to define an optimized scenario.
@@
 
-By suggesting an automated algorithm that first removes class characteristics, such as the grooves and the curvature of the bullet to reveal the region of the  land, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
+By suggesting an automated algorithm that first removes class characteristics, such as the groove impressions, shoulders, and the curvature of the bullet to reveal the region of the land, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
 In particular, `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.

### 2017-06-18 imaging-paper.tex
Paper updates based on review

@@
 \caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
-\end{figure}
-
-Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land".
-In the Hamby data, typical length (groove-to-groove) of a land is about 1998.28 micrometers or 2 millimeters.
+\label{fig:allx3p}
+\end{figure}
+
+Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent groove impressions on the bullet, called a ``land" or land impression.
+The shoulders are the raised portion of the surface between the groove impression and the land impression.
+In the Hamby data, typical length (shoulder-to-shoulder) of a land is about 1998.28 micrometers or 2 millimeters.
 For notational simplicity, we refer to a particular land of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{\label{fig:loess_step1}Step 1 of identifying groove locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
+\caption{\label{fig:loess_step1}Step 1 of identifying shoulder locations: For a fixed height ($x = 243.75\mu m$)  surface measurements for bullet~1-5 are plotted across the range of $y$.}
 \begin{knitrout}
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{Maximal number of CMS between unknown bullet C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown bullet land C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
 \begin{knitrout}

### 2017-06-18 imaging-supplement.Rnw
Paper updates based on review

@@
 Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
-The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
+The residuals are dominated, as expected, by the shoulders, which show up as large positive residuals.
 For a profile at height $x = 100\mu m$ there is a residual circular structure that does not show up for all signatures.
@@
 \centering
-<<crosscuts-sd, echo=FALSE, fig.height=5, fig.width=8, out.width='.7\\textwidth', dependson='setup-signatures'>>=
+<<crosscuts-sd, message=FALSE, echo=FALSE, fig.height=5, fig.width=8, out.width='.7\\textwidth', dependson='setup-signatures'>>=
 qplot(x, sd, data=stats, geom="line", group=bullet, alpha=I(0.5)) + theme_bw() +

### 2017-06-18 revision/response3.Rmd
Paper updates based on review

---
title: "Response to reviewers"
author: "Eric Hare, Heike Hofmann, Alicia Carriquiry"
date: "June 19, 2017"
output:
  pdf_document: default
bibliography: references.bib
---

# Reviewer 1

# Reviewer 2

1. Page 2, line 46-47: I do not understand the phrase, “the transcription of the individual characteristics into a 2D plotting framework.” Please write the idea more clearly.

    **Thank you. We have clarified this statement.**

2. Figures. 1, 3, 4, 5, 6, 7, 8, 9, 10, 12b, 12c, 14: These figures all have labels that are too small to be readable or they lack units or both.  The authors must fix all of them in one way or the other.  Without the units, the reader cannot be sure whether the scales along the x direction represent pixels (data points) or um.

    **XXX TODO: Heike, could you take a stab at this?**

3. Page 8, lines 207-209: I do not understand the sentence, “Here, d is defined as the distance between xi and the location of the fit xo and the maximum distance of the range of the x-values for span alpha in xo.” It appears that the quantity d has two definitions, “distance between xi and the location of the fit xo” and “maximum distance of the range of the x-values”.  This cannot be correct. Please re-write.

    **Apologies for the confusion. This was intended to convey the fact that in the traditional definition, the value d is in an interval between 0 and 1, and hence to map our distances to this scale, we divide by the range of the x values. We have rewritten this statement to hopefully make it more clear.**

4. Page 11, line 240: Something is missing in the description of Step 3.  How can the calculation of a rolling average by itself yield a criterion for identifying peaks and valleys in the profile?  Please rewrite the description of the procedure more completely.

    **We have added some clarification to this point. The rolling average is then used to determine peaks and valleys by identifying points at which the derivative is equal to zero. Given sufficient smoothing, this should correspond to identification of a striation marking.**

5. Page 13, figure 7: Even though the figures are tiny and the scales are too small, I should be able to discern a shift of (1.5625 x 17 um) = 27 um in one of the profiles between graphs 7a and 7b.  Instead the two graphs look identical.  I think the wrong graph was plotted somewhere.

    **You are correct below that the actual optimal lag was -2. We have corrected this error.**

6. Page 13, figure 8: The graph does not agree with its caption.  The caption states that correlation is maximized at a lag of minus 17 whereas the graph clearly shows the maximum at a lag of about minus 2.  There must be an error to rectify here. If the lag is only 2, that would answer the issue I raised about figure 7.

    **As stated above, you are correct. The caption has been corrected.**

7. Page 20, figure 16: Why are the AUC values the same (0.91) for all graphs although the graphs are all different.  This seems to result from clerical errors.

    **This clerical error has been corrected and the labels are now accurate.**

8. Page 22, figure 19: The perfect separation observed between matching and non-matching data points is not consistent with the modest predicted probabilities shown on the abscissa.  For example, at predicted probabilities of 0.6 to 0.75, I would have expected some dark blue non-matching data points to appear on the right side of the graph.  It would be good to explain what “predicted probability” means precisely for this graph.

    **XXX TODO: Heike, can we explain this better in the paper?**

    **The primary difference in the numbers comes from the fact that one is an out-of-bag error rate, i.e., an error rate computed on the land comparisons that were held out of the trees during the computation of the random forest. Meanwhile, figure 19 is the predicted probability on all land-to-land comparisons provided by our final forest.**

9. Page 3, lines 68-70: The authors’ term for the transition region between land impressions and groove impressions on a bullet surface is unconventional. The authors have chosen to use the word “groove” instead of “shoulder” to describe the transition region between the land impression and the groove impression areas. This could be confusing to a firearms examiner.

    **We have carefully gone through the paper and adjusted the language so that it will be more clear to a firearms examiner.**

10. Page 5, line 142: For better clarity, the authors could refer to Fig. 1 when describing the coordinate “fixed height x along the bullet land”.

    **We have included this reference.**

11. Page 6, line 168: I think it would avoid confusion later if the authors make it clear that the smoothing factor s is specified in terms of data points and not in terms of distance.  Adding “(pixels)” or (data points)” to the number “35” on this line would make the difference more explicit at the start.

    **This does make it much more clear, and we have made the change.**

12. Page 11, lines 240-254: There appear to be logical gaps in the explanations for Steps 3 and 4.  It is difficult to grasp the difference between “common” intervals and “joint” intervals.  From Step 3 it seems clear that common intervals must be specified for a pair of profiles, whereas joint intervals seem to be specified for single profiles.  This leads to confusion about why the region labelled “X” in Fig. 6c is not considered to contain matching striations when the intervals on both profiles overlap.

    **We hope we have clarified portions of these steps sufficiently. First, we've made it more clear about how the peak and valley detection operates. Secondly, you are correct that the region labeled X in Figure 6c was an error. This has since been correct, and hence your intuition about common intervals was in fact correct. If there are further steps that can be taken to improve this portion, please let us know.**

13. Page 14, line 334: The phrase “an equidistant resolution” is confusing here.  I think that “equal sampling intervals” would be more conventional.

    **We agree, and have accordingly made this change.**

14. Page 15, line 372: The authors should supply a reference for the term “tank rash.”

    **This reference has now been included (Hamby 2009).**

15. Page 18, figure caption 13a: Change “bullet C-3” to “bullet land C-3”.

    **Done.**

16. Page 19, figure 15: The histograms seem unrealistically smooth, particularly the histograms for matches.  How were they calculated?

    **These are actually density plots rather than histograms. Essentially, they're curve fits to the actual histograms. This is why they appear smoother than the raw results would appear.**

17. Supplement, page 6, figure caption 4: The caption claims a max CMS (consecutive matching striae) of 22, but I was only able to count a max CMS of 11. How is CMS defined here?

    **The location of the Os at the bottom of the figure identify the locations that our algorithm detects a striation marking. The primary reason for the discrepancy is that we treat valleys as a striation marking for the purpose of the CMS computation as well as peaks. This is described further on page 12 in the main paper.**

### 2017-06-22 imaging-paper.Rnw
first couple of figures - had to change path, will change back

@@
 codedir <- "code"
-datadir <- "images/Hamby (2009) Barrel/bullets"
+datadir <- "/Users/heike/papers/dissertations/eric-dissertation/images/Hamby (2009) Barrel/bullets/"
+#datadir <- "images/Hamby (2009) Barrel/bullets"
 
@@
             colour="grey30", size=0.25) +
-#  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
-  ylab("") +
+  ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+#  ylab("") +
   theme_bw() + 
@@
 
-p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + ylab("Surface measurements")
+p1 <- qplot(y, value, data = br111_bullet, size=I(1)) + theme_bw() + 
+    ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20))
+
 p1
@@
   geom_point(aes(x=(1:length(smoothed_truefalse))*1.5625, y=smoothed_truefalse)) + theme_bw() +
-    geom_vline(xintercept=xloc, colour="grey50") + xlab("y") + ylab("Surface measurements")
+    geom_vline(xintercept=xloc, colour="grey50") + 
+    ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20))
 
@@
         geom_vline(xintercept = subdata$y[groove_ind2], colour = "blue") +
-      xlab("y") + ylab("Surface measurements")
+      ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20))
 
@@
 
-my.loess$fitted + ylab("Surface measurements")
+my.loess$fitted + ylab(expression(paste("Relative Height (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20))
 @
@@
   geom_line(aes(x=y, y=resid), data=my.loess$data) +
-  ylab("Residuals of loess fit") + xlab("y") + theme_bw()
+  theme_bw() +
+  ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20))
+
 @

### 2017-06-22 imaging-paper.tex
first couple of figures - had to change path, will change back

@@
 \subsection{Algorithm}\label{sec:algorithm}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in unzip(path, exdir = mydir): error 1 in extracting from zip file}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'NA' does not exist in current working directory ('/Users/heike/papers/2015-imaging-paper').}}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in unzip(path, exdir = mydir): error 1 in extracting from zip file}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'NA' does not exist in current working directory ('/Users/heike/papers/2015-imaging-paper').}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in data.frame(subLOFx1): object 'subLOFx1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in factor(lofX\$r05): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in levels(lofX\$type) <- c("{}groove"{}, NA, "{}peak"{}): object 'lofX' not found}}\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in unzip(path, exdir = mydir): error 1 in extracting from zip file}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'NA' does not exist in current working directory ('/Users/heike/papers/2015-imaging-paper').}}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in unzip(path, exdir = mydir): error 1 in extracting from zip file}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'NA' does not exist in current working directory ('/Users/heike/papers/2015-imaging-paper').}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'subLOFx1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'subLOFx2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ccf(subLOFx1\$resid, subLOFx2\$resid, lag.max = 100, plot = FALSE): object 'subLOFx1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ccf\$lag: object of type 'closure' is not subsettable}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'subLOFx1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in data.frame(subLOFx1): object 'subLOFx1' not found}}\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in matches\$allnas: object of type 'closure' is not subsettable}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in matches\$allnas: object of type 'closure' is not subsettable}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in UseMethod("{}group\_by\_"{}): no applicable method for 'group\_by\_' applied to an object of class "{}function"{}}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in subset.default(lines, lineid != 0): object 'lineid' not found}}\end{kframe}
+\end{knitrout}
 
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/smoothcut-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'NA' does not exist in current working directory ('/Users/heike/papers/2015-imaging-paper').}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: 'NA' does not exist in current working directory ('/Users/heike/papers/2015-imaging-paper').}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in rbind(subLOFx1, subLOFx2): object 'subLOFx1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'lof' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unique(data\$bullet): object 'lof' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'bAlign' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unique(lof\$bullet): object 'lof' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in subset(lofX, bullet == b12[1]): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in subset(lofX, bullet == b12[2]): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'b12' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'b12' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in rbind(peaks1\$lines, peaks2\$lines): object 'peaks1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'peaks' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'lofX' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in fortify(data): object 'peaks' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.65\textwidth]{figure/smoothmatch-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'b12' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(expr, envir, enclos): object 'b12' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in rbind(lines1, lines2): object 'peaks1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in !match: invalid argument type}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/crosscuts-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'lof' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(data, aesthetics, environment = env): object 'lof' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=\textwidth]{figure/crosscutsX-1}
-
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strsplit(as.character(x), split = "{}/"{}): object 'lofY' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ggplot(data, aesthetics, environment = env): object 'lofY' not found}}\end{kframe}
 \end{knitrout}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ccf\$lag: object of type 'closure' is not subsettable}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in ccf\$lag: object of type 'closure' is not subsettable}}\end{kframe}
 \includegraphics[width=.65\textwidth]{figure/ccf-1}

### 2017-06-22 imaging-paper.Rnw
text size increased

@@
 
-im1 <- "images/Hamby (2009) Barrel/bullets/Br1 Bullet 1-5.x3p"
-im2 <- "images/Hamby (2009) Barrel/bullets/Br1 Bullet 2-1.x3p"
+im1 <- file.path(datadir, "Br1 Bullet 1-5.x3p")
+im2 <- file.path(datadir, "Br1 Bullet 2-1.x3p")
 
@@
 #  scale_colour_manual("", values=cols) +
-  theme_bw() + ylab("") + 
-  theme(legend.position="none")
+  theme_bw() + #ylab("") + 
+  theme(legend.position="none") +
+    ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=12))
+
 @
@@
     geom_line(aes(x=y, y=l30, group=bulletname), data=lofX) +
-  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
+    ylab(expression(paste("Residuals (in ",mu,m,")", sep=""))) + 
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=12))
 @
@@
   ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
-  xlab("y")
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=12)) 
 @
@@
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab("Residuals of loess fit")
+ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20)) 
 @
@@
         legend.background = element_rect(fill=alpha('white', 0.4))) + 
-  ylab("Residuals of loess fit")
+  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20)) 
 @
@@
 <<ccf, echo=FALSE, fig.width = 8, fig.height = 3, out.width = '.65\\textwidth'>>=
-lag <- ccf$lag[which.max(ccf$acf)]
-qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
+maxlag <- ccf$lag[which.max(ccf$acf)]
+dframe <- data.frame(lag = ccf$lag, acf = ccf$acf)
+qplot(data = dframe, x=lag, xend=lag, y = 0, yend= acf, geom="segment", colour = I("grey50")) + 
   theme_bw() + ylab("Correlation") + xlab("Lag k") +
-  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
-  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
-                     minor_breaks = c(-75,-25,25, 75))
+  geom_segment(aes(x = maxlag, xend = maxlag, y = 0, yend =  max(ccf$acf)), colour="black") + 
+  scale_x_continuous(breaks = c(-100, -50, maxlag, 0, 50, 100),
+                     minor_breaks = c(-75,-25,25, 75)) +
+  theme(text = element_text(size=20)) 
 @
@@
   geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match))
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
+#      ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
+#  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20)) 
 })
@@
   \centering
-<<smoothfac1, echo=FALSE, fig.width=7, fig.height=5.2, out.width='.5\\textwidth', warning=FALSE>>=
+<<smoothfac1, echo=FALSE, fig.width=10, fig.height=5.2, out.width='.75\\textwidth', warning=FALSE>>=
 br111.groove <- get_grooves(br111_bullet)
@@
   geom_line(aes(x=y, y=smoothed), data=smooths) +
-  ylab(expression(paste("Signature values (in ",mu,"m)", sep=""))) 
+  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20)) 
 @
@@
 
-qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
+qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS") + 
+  theme(text = element_text(size=20)) 
 @
@@
 \caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
-<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
+<<top, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
 res <- reslist[[which.max(cmsdist)]]  
@@
 
-print(ggplot() +
+ggplot() +
   theme_bw() + 
-  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=5, data=res$lines, alpha=0.2, show.legend=FALSE) +
+  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, data=res$lines, alpha=0.2, show.legend=FALSE) +
   geom_line(aes(x = y, y = l30, linetype = bullet),  data = res$bullets) +
@@
   theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
+  ylim(c(-6,7)) +
   geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
-
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
+    ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20))
 @
@@
 \caption{Top 2 match with C-3 based on CMS.}
-<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
+<<top2, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
 res <- reslist[[which(cmsdist==14)]]  # number 2
@@
   theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
+  ylim(c(-6,7)) +
   geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
+    ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=20))
 
@@
   theme(legend.position="bottom") + 
-  xlab("maximum CMS")
+  xlab("maximum CMS") +
+  theme(text = element_text(size=20)) 
 @
@@
   scale_fill_brewer("", palette="Paired") +
-  xlab("maximum CMS") + ylab("Proportion")
+  xlab("maximum CMS") + ylab("Proportion") +
+  theme(text = element_text(size=20)) 
 @
@@
   \centering
-<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.65\\textwidth', warning=FALSE>>=
+<<strange-res, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
 load("data/data-25-25/unkn47.RData")
@@
   theme(legend.position = c(1,1), legend.justification=c(1,1)) + 
-  ylim(c(-6,6)) +
+  ylim(c(-6,7)) +
   geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
-  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)))
+  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
+    ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
+  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
+  theme(text = element_text(size=15))
 @
@@
 levels(dm$var)[3:4] <- c("#matches", "#non-matches")
-levels(dm$variable) <- c("Known matches (KM)", "Known non-matches (KNM)")
+levels(dm$variable) <- c("Known matches (KM)", "Known non-matches\n (KNM)")
 dm$varLabel <- dm$var
 levels(dm$varLabel) <- c(
-  "Consecutive Matching Striae (CMS)", 
-  "Consecutive Non-Matching Striae (CNMS)",
+  "Consecutive Matching\n Striae (CMS)", 
+  "Consecutive Non-Matching\n Striae (CNMS)",
   "#matches",
@@
   "Sum of peaks (S)",
-  "Cross-correlation function (ccf)")
+  "Cross-correlation\n function (ccf)")
 
@@
         axis.title.y=element_blank(),
-        plot.margin = unit(c(0,0,0,0), unit="cm"))
+        plot.margin = unit(c(0,0,0,0), unit="cm")) +
+  theme(text = element_text(size=20)) 
 @
@@
   "Sum of peaks (S)",
-  "Cross-correlation function (ccf)",
+  "Cross-correlation\n function (ccf)",
   "#non-matches",
   "#matches",
-  "Consecutive Non-Matching Striae (CNMS)",
-  "Consecutive Matching Striae (CMS)", 
+  "Consecutive Non-Matching\n Striae (CNMS)",
+  "Consecutive Matching\n Striae (CMS)", 
   "Average difference (D)"
@@
   geom_point(aes(x=`FALSE`, y=`TRUE`), data=eer, size=2.5) +
-  geom_label(x=1, y=0.125, aes(label=labels), data=eer, hjust=1, vjust=0)
+  geom_label(x=1, y=0.125, aes(label=labels), data=eer, hjust=1, vjust=0) +
+  theme(text = element_text(size=20)) 
 @
@@
         panel.grid.major.y=element_blank(), panel.grid.minor.y=element_blank()) +
-  xlab("Importance (mean decrease in Gini index)")
+  xlab("Importance (mean decrease in Gini index)") +
+   theme(text = element_text(size=15)) 
 @

### 2017-06-22 imaging-paper.Rnw
fig 12 and 13

@@
   xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  theme(text = element_text(size=20))
+  theme(text = element_text(size=18), legend.background = element_rect(fill="transparent"))
 @
@@
   xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  theme(text = element_text(size=20))
+  theme(text = element_text(size=18), legend.background = element_rect(fill="transparent"))
 
@@
 <<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
+bstats$cmstop <- bstats$CMS
+bstats$cmstop[bstats$CMS >=20] <- "20+"
+bstats$cmstop <- factor(bstats$cmstop)
+bstats$cmstop <- reorder(bstats$cmstop, bstats$CMS)
 ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(CMS))) + theme_bw() +
+  geom_bar(aes(x=cmstop)) + theme_bw() +
   theme(legend.position="bottom") + 
@@
 ggplot(data=subset(bstats, !flagged)) + 
-  geom_bar(aes(x=factor(CMS), fill=km), position="fill") +
+  geom_bar(aes(x=cmstop, fill=km), position="fill") +
   theme(legend.position="bottom") + 
@@
 levels(dm$var)[3:4] <- c("#matches", "#non-matches")
-levels(dm$variable) <- c("Known matches (KM)", "Known non-matches\n (KNM)")
+levels(dm$variable) <- c("Known matches\n(KM)", "Known non-matches\n(KNM)")
 dm$varLabel <- dm$var

### 2017-06-22 imaging-paper.Rnw
line break in facet

@@
   xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
-  theme(text = element_text(size=15))
+  theme(text = element_text(size=15), legend.background = element_rect(fill="transparent"))
 @
@@
   "Consecutive Matching\n Striae (CMS)", 
-  "Consecutive Non-Matching\n Striae (CNMS)",
+  "Consecutive Non-Match-\ning Striae (CNMS)",
   "#matches",

### 2017-06-23 imaging-paper.Rnw
Update reviewer response

@@
 codedir <- "code"
-datadir <- "/Users/heike/papers/dissertations/eric-dissertation/images/Hamby (2009) Barrel/bullets/"
-#datadir <- "images/Hamby (2009) Barrel/bullets"
+datadir <- "images/Hamby (2009) Barrel/bullets"
 

### 2017-06-28 imaging-paper.Rnw
Resubmission again!

@@
 % "Title of the paper"
-\title{Automatic Matching of Bullet Lands}\thanksref{T1}
+\title{Automatic Matching of Bullet Land Impressions}\thanksref{T1}
 \thankstext{T1}{Resources for this manuscript, including R code to reproduce all of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}
-\runtitle{Automatic Matching of Bullet Lands}
+\runtitle{Automatic Matching of Bullet Land Impressions}
 \runauthor{E. Hare et al.}
@@
 Identifying these striation markings in an attempt to match two bullets is one of the primary goals of firearm examination.
-We propose an automated framework for the analysis of the 3D surface measurements of bullet lands that first transcribes the markings into a 2D plotting framework.
+We propose an automated framework for the analysis of the 3D surface measurements of bullet land impressions which transcribes the individual characteristics into a set of features that quantify their similarities.
 This makes identification of matches easier and allows for a quantification of both matches and matchability of barrels.
@@
 In 2009, the National Academy of Sciences published a report~\citep{NAS:2009} questioning the scientific validity of many forensic methods including firearm examination.
-The report states that ``[m]uch forensic evidence -- including, for example, bite marks and firearm and toolmark identification is introduced in criminal trials without any meaningful scientific validation, determination of error rates, or reliability testing to explain the limits of the discipline."
+The report states that ``[m]uch forensic evidence -- including, for example, bite marks and firearm and toolmark identification -- is introduced in criminal trials without any meaningful scientific validation, determination of error rates , or reliability testing to explain the limits of the discipline."
 
@@
 From a statistical standpoint, identification of the gun that fired the bullet(s) requires that we compare the probabilities of observing matching striae under the competing hypotheses that the gun fired, or did not fire, the crime scene bullet.
-If indeed the uniqueness assumption is plausible, the latter probability approaches zero.
+If indeed the uniqueness assumption is plausible, the latter probability approaches zero and the former approaches (but never reaches) one.
 
@@
 3D topographical images of each bullet were obtained using a NanoFocus lens at 20x magnification and made publicly available on the NIST Ballistics Database Project\footnote{\url{http://www.nist.gov/forensics/ballisticsdb/hamby-consecutively-rifled-barrels.cfm}} in a format called x3p (XML 3-D Surface Profile).
-The x3p format conforms to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}}, implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
+The x3p format conforms to the ISO5436-2 standard\footnote{\url{http://sourceforge.net/p/open-gps/mwiki/X3p/}} and is implemented to provide a simple and standard conforming way to exchange 2D and 3D profile data.
 It was adopted by the OpenFMC (Open Forensic Metrology Consortium\footnote{\url{http://www.openfmc.org/}}), a group of academic, industry, and government firearm forensics researchers whose aim is to establish best practices for researchers using metrology in forensic science.
-We have developed an open-source package for analyzing bullet lands written in R \citep{R}.
+We have developed an open-source package for analyzing bullet land impressions written in R \citep{R}.
 This package is called bulletr \citep{bulletr} and enables direct reading and manipulation of x3p files.
@@
 \begin{subfigure}[t]{\textwidth}\centering
-    \caption{Frontal view of a bullet land (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
+    \caption{Frontal view of a bullet land impression (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
     \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
@@
 \end{subfigure}
-\caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
+\caption{Example of a groove-to-groove scan of a bullet land impression. The red-dotted rectangle on the right shows the location and orientation of the segment. }
 \label{fig:allx3p}
@@
 The shoulders are the raised portion of the surface between the groove impression and the land impression.
-In the Hamby data, typical length (shoulder-to-shoulder) of a land is about \Sexpr{typical_width} micrometers or \Sexpr{typical_width / 1000} millimeters.
-For notational simplicity, we refer to a particular land of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
-An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
+In the Hamby data, typical length (shoulder-to-shoulder) of a land impression is about \Sexpr{typical_width} micrometers or \Sexpr{typical_width / 1000} millimeters.
+For notational simplicity, we refer to a particular land impression of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
+An example of plotting one of these land impressions is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
 These figures show side and top profiles of the land respectively.
@@
 
-The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the strength of a match between two bullet lands.
+The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the strength of a match between two bullet land impressions.
 While we assess the performance of the algorithm in terms of a binary decision of match vs.
 non-match using a 50\% probability cut-off, our primary goal is to highlight the features that are statistically associated with matches and non-matches, and to provide a quantitative assessment of this association.
-In a real-world application of our algorithm, the raw scores would need further analysis and scrutiny, and it is likely a 50\% cut-off would be an inappropriate choice on the basis of reasonable doubt.
+In a real-world application of our algorithm, the raw scores would need further analysis and scrutiny, and it is likely that a 50\% cut-off would be an inappropriate choice on the basis of reasonable doubt.
 
@@
 We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet level.
-Although doing so introduces an implicit assumption of independence between lands, assuming independence only serves to make the task more challenging.
+Although doing so introduces an implicit assumption of independence between land impressions, assuming independence only serves to make the task more challenging.
 
 The remainder of this paper is structured as follows: We first briefly review some earlier work.
-We then discuss two methods of modeling the class structure of the bullet surfaces.
+We then discuss two methods for modeling the class structure of the bullet surfaces.
 Finally, we proceed to describing an automatic matching routine which we evaluate on the bullets made available through the Hamby study.
@@
 This might seem like a trivial distinction, but it has implications for the development of the algorithm.
-The algorithm performs alignment of striae by rotation of the XY plane, which is not generalizable to bullets in which the XY plane is not flat.
+Their algorithm performs alignment of striae by rotation of the XY plane, which is not generalizable to bullets in which the XY plane is not flat.
 
@@
 
-To analyze the striation pattern, we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land, as illustrated in Figure \ref{fig:allx3p}.
+To analyze the striation pattern, we extract a \emph{bullet profile} \citep{ma:2004} by taking a cross section of the surface measurements at a fixed height $x$ along the bullet land impression, as previously illustrated in Figure \ref{fig:allx3p}.
 %An initial naive approach to analyzing the striation patterns leads us to fixing a particular value of a coordinate, and then producing a plot of the surface measurements across. 
-Figure \ref{fig:fixedX} shows a plot of the side profile of a bullet land.
+Figure \ref{fig:fixedX} shows a plot of the side profile of a bullet land impression.
 It can be seen that the global structure of the land dominates the appearance of the plot.
@@
 @
-\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at a fixed height $x$.
+\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land impression at a fixed height $x$.
 Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
@@
 Signatures are manually aligned, resulting in many of the striation marks to continuously pass from one side to the other.
-Visually, this allows for an easy assessment of these two bullet lands as a match.
+Visually, this allows for an easy assessment of these two bullet land impressions as a match.
 However, this match relies on visual inspection and is therefore subjective.
@@
 
-All of the other steps are done on pairs of bullet lands:
+All of the other steps are done on pairs of bullet land impressions:
 \begin{enumerate}
@@
 Counting the maximal number of CMS is part of the current practice to identify bullet matches~\citep{nichols:1997, nichols:2003, nichols:2003b}. 
-In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a high number suggestive of a match between the lands.
+In the example of Figure~\ref{fig:match}, the number of consecutive matching striations (CMS) is fifteen, a high number suggestive of a match between the land impressions.
 Note that the definition of CMS we use does not match the one given in \citet{thompson:2013}.
 There, CMS is defined only in terms of matching peaks without regarding valleys.
-Additionally, peaks in  \citet{thompson:2013} are  used only if they can be identified and matched `within a tolerable range' between lands.
+Additionally, peaks in  \citet{thompson:2013} are  used only if they can be identified and matched `within a tolerable range' between land impressions.
 The definition given here is computationally less complex, but should yield highly correlated values, because of the requirement to only consider signatures from a stable region in the land (see Section~\ref{sec:heights} for further details on stability of regions).
@@
 \subsection{Horizontal alignment}
-Signatures of each of the two lands, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$.
+Signatures of each of the two land impressions, 1-5 and 2-1,  in Figure~\ref{fig:manualmatch-rgl} are shown in Figure~\ref{fig:cross100} extracted at a height of $x = 100\mu m$.
 Striation marks show up in these representations as peaks and valleys.
-The individual characteristics are prominent and, again, suggest a match between the lands.
+The individual characteristics are prominent and, again, suggest a match between the land impressions.
 A horizontal shift of one of the signatures (result shown in Fig~\ref{fig:crosscutX}) emphasizes the strong similarities between signatures.
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
+\caption{Raw bullet land impression signatures.\label{fig:crosscut}}{%
 <<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
@@
   geom_segment(aes(x = maxlag, xend = maxlag, y = 0, yend =  max(ccf$acf)), colour="black") + 
-  scale_x_continuous(breaks = c(-100, -50, maxlag, 0, 50, 100),
+  scale_x_continuous(breaks = c(-100, -50, 0, 50, 100),
                      minor_breaks = c(-75,-25,25, 75)) +
@@
 \subsection{Impact of bullet height}\label{sec:heights}
-The height at which signatures are extracted for a comparison between bullet lands matters -- signatures taken from heights that are further apart, show more pronounced differences between the signatures.
+The height at which signatures are extracted for a comparison between bullet land impressions matters -- signatures taken from heights that are further apart, show more pronounced differences between the signatures.
 This poses both a caveat to matching attempts as well as an opportunity for quality control: we have to be aware of the height that was used in a matching.
 Visually, matches degrade if the signatures upon which the match is based are from heights further than 200$\mu m$ apart  (see Supplement Section~\ref{supp:ccf} for more discussion).
-However, we can extract signatures from multiple heights of the same bullet land for an initial assessment of its quality.
+However, we can extract signatures from multiple heights of the same bullet land impression for an initial assessment of its quality.
 By comparing signatures from heights that are not too far apart --  25~$\mu m$ to 50~$\mu m$ -- we get an indication whether the signatures come from a rapidly changing section of the surface, indicative of a break-off or some other damage, or from a stable section, where we have a reasonable expectation of finding matches to other signatures.
@@
 At a  level of $175\mu m$ the bullet's signature stabilizes.
-For this land, matches should not be attempted at lower heights.
+For this land impression, matches should not be attempted at lower heights.
 }
@@
 In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance using the James Hamby study data.
-As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land.
+As a first step, we automatically assess the quality of each of the land impressions by checking that we can identify a stable region.
 For this, we compute the cross-correlation of signatures extracted from heights 25$\mu m$ apart.
 For a stable region, we require a minimum of 0.95 for the cross correlation.
-Four lands from different bullets are flagged as problematic in this respect.
-A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands has scratch marks across the surface, also known as `tank rash' \citep{hamby:2009}.
+Four land impressions from different bullets are flagged as problematic in this respect.
+A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these land impressions has scratch marks across the surface, also known as `tank rash' \citep{hamby:2009}.
 %
@@
 \end{subfigure}
-\caption{\label{fig:fourflags}Images of the four lands that got flagged during the quality assessment.
+\caption{\label{fig:fourflags}Images of the four land impressions that got flagged during the quality assessment.
 All of them show scratch marks (tank rash) across the striation marks from the barrel.
@@
 %
-We exclude these four lands from further matching considerations and
- run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e.
-we are comparing $15 \times 6 -2 = 90 - 2 = 88$ lands from unknown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons.
+We exclude these four land impressions from further matching considerations and
+ run all remaining land impressions from the unknown bullets against all remaining land impressions from known bullets for matches, i.e.
+we are comparing $15 \times 6 -2 = 90 - 2 = 88$ land impressions from unknown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ land impressions from known bullets, yielding a total of $10,384$ land-to-land comparisons.
 Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM).
-Ideally, results look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 lands from known bullets.
-Two lands show a high CMS.
+Ideally, results look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land impression C-3 and all 118 land impressions from known bullets.
+Two land impressions show a high CMS.
 These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and~\ref{fig:hamby-perfect}c.
  Unfortunately, not all results are as clear cut. 
-It might not be reasonable to assume that we can match all lands, but the idea is to try to maximize the number of matches to get an overview of what we might be able to expect from an automated match.
+It might not be reasonable to assume that we can match all land impressions, but the idea is to try to maximize the number of matches to get an overview of what we might be able to expect from an automated match.
 
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{Maximal number of CMS between unknown bullet land C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown bullet land impression C-3 and all of the other 118 considered (known) land impressions. For two land impressions the number of maximum CMS is high. }
 <<cms, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
+\caption{Overlaid signatures of C-3 and the land impression with the top matching CMS.}
 <<top, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth', warning = FALSE>>=
@@
 Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study.
-All 42 pairs of lands with at least thirteen CMS in common are matches.
+All 42 pairs of land impressions with at least thirteen CMS in common are matches.
 %
@@
 While the probability for a match increases with the number of maximal CMS, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}.
-Figure~\ref{fig:mismatch} shows a known mismatch between two lands that share twelve consecutively matched striae.
-Visually we can easily tell that these two lands do not match well.
+Figure~\ref{fig:mismatch} shows a known mismatch between two land impressions that share twelve consecutively matched striae.
+Visually we can easily tell that these two land impressions do not match well.
 
@@
 
-Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
+Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between land impressions based on features derived from the image files.
 The resulting tree is shown in Figure~\ref{fig:tree}.
-A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} lands is being matched correctly.
+A total of \Sexpr{sum(bstats$match[bstats$pred>0.5])} land impressions is being matched correctly.
 Interestingly, the number of consecutive matching striae does not feature in this evaluation.
@@
 Note that this error rate is based on land-to-land comparisons and is expected to be much lower for bullet-to-bullet comparisons.
-In the case of the Hamby data, even a single tree results in an overall error rate of zero, if we require that a match of two bullets occurs when at least two of the bullet's lands are matched.
+In the case of the Hamby data, even a single tree results in an overall error rate of zero, if we require that a match of two bullets occurs when at least two of the bullet's land impressions are matched.
 % (see Figure~\ref{fig:treeresult}).
@@
 
-By suggesting an automated algorithm that first removes class characteristics, such as the groove impressions, shoulders, and the curvature of the bullet to reveal the region of the land, then identifies peaks and valleys on this land, we reduce subjectivity and with it possible sources of bias.
+By suggesting an automated algorithm that first removes class characteristics, such as the groove impressions, shoulders, and the curvature of the bullet to reveal the region of the land impression, then identifies peaks and valleys on this land impression, we reduce subjectivity and with it possible sources of bias.
 In particular, `the concept of counting striations is subjective and based on experience'~\citep{miller:1998}.
@@
 Thanks to the men and women behind the software R~\citep{R}, and the authors of the R packages  knitr~\citep{knitr} and ggplot2~\citep{ggplot2}.
+We also wish to thank an anonymous reviewer, whose constructive and insightful comments contributed greatly to the improvement of the manuscript.
 

### 2017-06-28 imaging-paper.tex
text size increased
Resubmission again!

--- 2017-06-18 imaging-paper.tex
@@
 The shoulders are the raised portion of the surface between the groove impression and the land impression.
-In the Hamby data, typical length (shoulder-to-shoulder) of a land is about 1998.28 micrometers or 2 millimeters.
-For notational simplicity, we refer to a particular land of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
-An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
+In the Hamby data, typical length (shoulder-to-shoulder) of a land impression is about 1998.28 micrometers or 2 millimeters.
+For notational simplicity, we refer to a particular land impression of a bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number.
+An example of plotting one of these land impressions is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}.
 These figures show side and top profiles of the land respectively.
@@
 \end{knitrout}
-\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land at a fixed height $x$.
+\caption{\label{fig:fixedX}Side profile of the surface measurements (in~$\mu m$) of a bullet land impression at a fixed height $x$.
 Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
+\caption{Raw bullet land impression signatures.\label{fig:crosscut}}{%
 \begin{knitrout}
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.5\textwidth]{figure/smoothfac1-1}
+\includegraphics[width=.75\textwidth]{figure/smoothfac1-1}
 
@@
 \begin{subfigure}[t]{\textwidth}\centering
-\caption{Maximal number of CMS between unknown bullet land C-3 and all of the other 118 considered (known) lands. For two lands the number of maximum CMS is high. }
+\caption{Maximal number of CMS between unknown bullet land impression C-3 and all of the other 118 considered (known) land impressions. For two land impressions the number of maximum CMS is high. }
 \begin{knitrout}
@@
 \begin{subfigure}[b]{.49\textwidth}\centering
-\caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
+\caption{Overlaid signatures of C-3 and the land impression with the top matching CMS.}
 \begin{knitrout}
@@
 
-Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between lands based on features derived from the image files.
+Using recursive partitioning, we fit a decision tree~\citep{breiman:1984, rpart, rpart.plot} to predict matches between land impressions based on features derived from the image files.
 The resulting tree is shown in Figure~\ref{fig:tree}.
-A total of 132 lands is being matched correctly.
+A total of 132 land impressions is being matched correctly.
 Interestingly, the number of consecutive matching striae does not feature in this evaluation.

### 2017-06-28 imaging-supplement.Rnw
Resubmission again!

@@
 % "Title of the paper"
-\title{Supplement to Automatic Matching of Bullet Lands}
-\runtitle{Supplement: Automatic Matching of Bullet Lands}
+\title{Supplement to Automatic Matching of Bullet Land Impressions}
+\runtitle{Supplement: Automatic Matching of Bullet Land Impressions}
 
@@
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
-In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.
+In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land impression covers about \Sexpr{diff(range(dbr111$theta))} degrees.
 Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
@@
 @
-\caption{Circular fit to the signature of each land of bullet 2, with signature from bullet 1-5 overlaid.\label{fig:circlefits} The signature of bullet 1-5 matches best with bullet 2-1.}
+\caption{Circular fit to the signature of each land impression of bullet 2, with signature from bullet 1-5 overlaid.\label{fig:circlefits} The signature of bullet 1-5 matches best with bullet 2-1.}
 \end{figure}
@@
 At larger heights  individual characteristics become less distinctive, making true matches to other bullets harder.
-The pattern of decreasing peaks and valleys is generally true for  bullet lands, as can be seen in Figure~\ref{fig:sds}.
+The pattern of decreasing peaks and valleys is generally true for bullet land impressions, as can be seen in Figure~\ref{fig:sds}.
 %
@@
 \end{figure}
-Figure~\ref{fig:sds} shows that the amount of standard deviation of a signature decreases on average for all bullet lands at larger heights.
+Figure~\ref{fig:sds} shows that the amount of standard deviation of a signature decreases on average for all bullet land impressions at larger heights.
 This makes standard deviation of a signature one measure to quantify the extent to which a signature is expressed.
@@
 \section{Complete evaluation of the Hamby study}\label{supp:extended}
-One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land.
-This effectively doubles the number of comparisons from 10,384 pairwise comparisons of usable bullet lands to 21,115 $\left[= (118+88)\cdot 205/2\right]$ comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons.
+One way to expand the use of the James Hamby study is to not only match all of the unknown bullet land impressions against the known bullet land impressions, but to compare every land impression against every other land impression.
+This effectively doubles the number of comparisons from 10,384 pairwise comparisons of usable bullet land impressions to 21,115 $\left[= (118+88)\cdot 205/2\right]$ comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons.
 
@@
 
-Figures~\ref{fig:aligned} and~\ref{fig:aligned-second} give an overview of all the signatures from bullet lands in the Hamby study aligned by barrel.
+Figures~\ref{fig:aligned} and~\ref{fig:aligned-second} give an overview of all the signatures from bullet land impressions in the Hamby study aligned by barrel.
 Three to five bullets were fired from each barrel.
 The figures give us both some insight into how well signatures match and how consistent individual characteristics are impregnated on bullets fired from each of the barrels.
-Signatures for some lands match remarkably well -- such as land 5 from barrel 1, whereas all lands from barrel 5 show some variability both in the location and depths of peaks and valleys.
+Signatures for some land impressions match remarkably well -- such as land 5 from barrel 1, whereas all land impressions from barrel 5 show some variability both in the location and depths of peaks and valleys.
 
@@
 @
-\caption{\label{fig:aligned}Overview of aligned signatures for all bullet lands for  barrels 1 to 5 of the Hamby study. }
+\caption{\label{fig:aligned}Overview of aligned signatures for all bullet land impressions for  barrels 1 to 5 of the Hamby study. }
 \end{figure}
@@
 @
-\caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet lands for  barrels 5 to 10 of the Hamby study. }
+\caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet land impressions for  barrels 5 to 10 of the Hamby study. }
 \end{figure}
@@
 Two random forests were calculated for the Hamby study.
-For the first random forest only comparisons of bullet lands from known bullets and unknown bullets were used.
-The second random forest is based on a full comparison of every land with every other land, increasing the number of comparisons from originally 10,384 (10,212 known non-matches and 172 known matches) by another 10,931 comparisons (10,637 known non-matches and 94 known matches).
+For the first random forest only comparisons of bullet land impressions from known bullets and unknown bullets were used.
+The second random forest is based on a full comparison of every land impression with every other land impression, increasing the number of comparisons from originally 10,384 (10,212 known non-matches and 172 known matches) by another 10,931 comparisons (10,637 known non-matches and 94 known matches).
 Random forests allow an assessment of variable importance (also called feature importance) as the mean decrease in Gini index when including each variable.

### 2017-06-28 imaging-supplement.tex
Update response, make more paper changes
Merge the set44 with the regular set
Refactor some library code
Response update
Notes from mtg
Continue working on many of the changes
Finally fix up all the paths
Rebuild the papers
Rebuild both papers with no cache
Fix paper building on linux
Update to use new csafe server
Fix the bullets backend app
Paper updates based on review
Resubmission again!

@@
 The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
-In this instance, the radius is estimated as $\hat{r} = 4666.49\mu m = 4.67mm$ and the land covers about 29.5 degrees.  Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
-The residuals are dominated, as expected, by the grooves, which show up as large positive residuals.
+In this instance, the radius is estimated as $\hat{r} = 4666.49\mu m = 4.67mm$ and the land impression covers about 29.5 degrees.
+Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
+The residuals are dominated, as expected, by the shoulders, which show up as large positive residuals.
 For a profile at height $x = 100\mu m$ there is a residual circular structure that does not show up for all signatures.
@@
 \end{knitrout}
-\caption{Circular fit to the signature of each land of bullet 2, with signature from bullet 1-5 overlaid.\label{fig:circlefits} The signature of bullet 1-5 matches best with bullet 2-1.}
+\caption{Circular fit to the signature of each land impression of bullet 2, with signature from bullet 1-5 overlaid.\label{fig:circlefits} The signature of bullet 1-5 matches best with bullet 2-1.}
 \end{figure}
@@
 \end{knitrout}
-\caption{\label{fig:aligned}Overview of aligned signatures for all bullet lands for  barrels 1 to 5 of the Hamby study. }
+\caption{\label{fig:aligned}Overview of aligned signatures for all bullet land impressions for  barrels 1 to 5 of the Hamby study. }
 \end{figure}
@@
 \end{knitrout}
-\caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet lands for  barrels 5 to 10 of the Hamby study. }
+\caption{\label{fig:aligned-second}Overview of aligned signatures for all bullet land impressions for  barrels 5 to 10 of the Hamby study. }
 \end{figure}
@@
 \centering
-% latex table generated in R 3.2.4 by xtable 1.8-2 package
-% Wed Apr 20 10:17:06 2016
+% latex table generated in R 3.4.0 by xtable 1.8-2 package
+% Wed Jun 28 07:16:37 2017
 \begin{tabular}{clrrrd{3.2}rd{3.2}}

### 2017-06-28 revision/cover_letter.Rmd
Resubmission again!

---
author: 
 - Eric Hare
 - Department of Statistics 
 - Iowa State University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: Dear Dr. Airoldi,
closing: Best regards,
address: 
 - Edoardo M. Airoldi
 - Editor
 - The Annals of Applied Statistics
return-address:
 - Eric Hare
 - Department of Statistics
 - Iowa State University
 - Ames, IA 50011
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

Re: Manuscript reference No. AOAS1702-003

We have uploaded the revised version on the manuscript as well as our response to the constructive comments of Reviewer 1.  We did not respond to Reviewer 2’s comments, but can do so if needed.  Hopefully, this version is now acceptable for publication in AoAS.  Both the data and the code used to develop the methodology have already been put in the public domain (Github), but if AoAS has a different repository, we will be happy to upload elsewhere as well (assuming that the paper is accepted).

We are grateful for having had an opportunity to respond to the reviews we received in round 2.

Let us know if there are any further concerns or requests.

With best wishes,

### 2017-06-28 submission/imaging-paper.tex
Resubmission again!

--- 2017-06-28 imaging-paper.tex
@@
 
-\linenumbers
+%\linenumbers
 %\tableofcontents
@@
     \caption{Frontal view of a bullet land impression (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
-    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
-    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
+    \includegraphics[width=.65\textwidth]{topx3p.png} \hspace{1cm}
+    \includegraphics[width=.15\textwidth]{top-sketch.png}
     }
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.75\textwidth]{figure/smoothfac1-1}
+\includegraphics[width=.75\textwidth]{smoothfac1-1}
 
@@
 This is a weighted average between false positive error rate of 0.0001 and an error rate of false negatives of 0.2267.
-This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).
+This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:treeforest}).
 Note that this error rate is based on land-to-land comparisons and is expected to be much lower for bullet-to-bullet comparisons.
@@
 On a probability spectrum of outcomes we could therefore introduce an interval of `inconclusive' results in the middle of the spectrum -- which turns out to be unnecessary in the Hamby study, because, here, the results from the random forest are very clear cut.
-Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree and the random forest.
+Figure~\ref{fig:treeforest} shows a comparison of the predicted probabilities of a match by the tree and the random forest.
 As expected, the random forest provides a more realistic estimate of the uncertainty in the classification.
@@
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
-\includegraphics[width=.6\textwidth]{figure/tree:forest-1}
-
-\end{knitrout}
-\caption{\label{fig:tree-forest}Prediction results from the tree and the forest.
+\includegraphics[width=.6\textwidth]{treeforest-1}
+
+\end{knitrout}
+\caption{\label{fig:treeforest}Prediction results from the tree and the forest.
 Using a cut-off probability of 0.5 the forest correctly predicts every single comparison.

### 2017-06-29 arxiv-submission/imaging-paper.tex
submission final

--- 2017-06-28 submission/imaging-paper.tex
@@
 While the circle is an obvious first choice for fitting the structure, it does not completely capture the bullet surface after it was fired.
-A discussion of a circular fit and the remaining residual structure can be found in Supplement Section~\ref{supp:cylindrical}.
+A discussion of a circular fit and the remaining residual structure can be found in Supplement Section~\ref{supp:cylindrical} \citep{supplement}.
 %It is clear that our next approach must be to model the overall structure in order to focus on the deviations, given that the deviations should correspond to bullet markings.
@@
 % AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
-%  \sdatatype{.pdf}"
-%  \sdescription{Some text}
-%\end{supplement}
+\begin{supplement}[id=suppA]
+  \sname{Supplement A}
+  \stitle{Supplement to Automatic Matching of Bullet Land Impressions}
+  \slink[doi]{COMPLETED BY THE TYPESETTER}
+  \sdatatype{.pdf}" 
+  \sdescription{Supplementary derivations for Automatic matching of bullet land impressions}
+\end{supplement}
 

### 2017-06-29 imaging-paper.Rnw
submission final

@@
 % AOS,AOAS: If there are supplements please fill:
-%\begin{supplement}[id=suppA]
-%  \sname{Supplement A}
-%  \stitle{Title}
-%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
-%  \sdatatype{.pdf}" 
-%  \sdescription{Some text}
-%\end{supplement}
+\begin{supplement}[id=suppA]
+  \sname{Supplement A}
+  \stitle{Supplement to Automatic Matching of Bullet Land Impressions}
+  \slink[doi]{COMPLETED BY THE TYPESETTER}
+  \sdatatype{.pdf}" 
+  \sdescription{Supplementary derivations for Automatic matching of bullet land impressions}
+\end{supplement}
 
